### 2024-06-05 conteudo/01-exemplo-normas-ime.tex
Add work done on 04/06

%!TeX root=../tese.tex
%("dica" para o editor de texto: este arquivo é parte de um documento maior)
% para saber mais: https://tex.stackexchange.com/q/78101

\en

\chapter{Background}

\section{Code Duplication Detection}

Code Duplication Detection is a field in computer science studies that gets the attention of researchers back to 1988 \citep{firstman}.
The occurrence of Code Duplication is a harmful artifact to have in a software, affecting software's related tasks such as code readability,
introdution of bugs, addition of co-changed requirement, etc.   \citep{harmone}. 
On most of the cases, the occurrence of Code Duplication tends to create a more unstable software then the nonduplicate code counterpart   \citep{harmtwo}.

Through the years, the subject of studies about Code Duplication Detection branched out mainly in two fields of study,
Code Clone Detection and Code Plagiarism Detection,
with the former one focused on the technical aspect while the later one also introduces an social aspect on the field of research
\citep{litreview}. For our research we are only interested on the Code Clone Detection branch given the nature of our work.

\subsection{Types of Code Duplication}

The categorization of two code artifacts as a duplication of each other is a subjective task (REFERENCE HERE). 
To mitigate this human aspect on the given problem, the literature classified Code Duplication on the scale of changes in between 
the code artifacts in four main types of Code Duplication \citep{litreview}. Given two code artifacts F and G, 
the types of Code Duplication are described as follow \citep{litreview}:

\begin{itemize}
	\item \textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names, white spaces and any other kind of irrelevant elements. (ADD FIGURE) shows a typical example of Type-1 Code Duplication \citep{litreview}.

	\item \textbf{Type-2:} The differences between F and G is the same as the Type-1, with the addition of considering addition and deletion of redudant codes. (ADD FIGURE) shows a typical example of Type-2 Code Duplication \citep{litreview}.

	\item \textbf{Type-3:} The difference between F and G is the same as the Type-2, with the addition of considering the reorder of code blocks as well as statements within code blocks. (ADD FIGURE) shows a typical example of Type-3 Code Duplication \citep{litreview}.

	\item \textbf{Type-4:} The difference between F and G is the same as the Type-3, with the addtionf of considering changes on data sctrucure, the order of operands/operators in expresssions, or replacing part of codes with equivalent composition. (ADD FIGURE) shows a typical example of Type-4 Code Duplication \citep{litreview}. 
\end{itemize}

\index{Tese/Dissertação!versões}

### 2024-06-06 conteudo/01-exemplo-normas-ime.tex
Add work done in 05/06

@@
 Code Duplication Detection is a field in computer science studies that gets the attention of researchers back to 1988 \citep{firstman}.
-The occurrence of Code Duplication is a harmful artifact to have in a software, affecting software's related tasks such as code readability,
-introdution of bugs, addition of co-changed requirement, etc.   \citep{harmone}. 
-On most of the cases, the occurrence of Code Duplication tends to create a more unstable software then the nonduplicate code counterpart   \citep{harmtwo}.
+The occurrence of Code Duplication is a harmful artifact to have in a software, affecting software related tasks such as code readability,
+introdution of bugs, etc.   \citep{harmone}. 
+On most of the cases, the occurrence of code duplication tends to create a more unstable software then the nonduplicate code counterpart.   \citep{harmtwo}
 
@@
 
-\subsection{Types of Code Duplication}
+\subsection{Types of code duplication}
 
 The categorization of two code artifacts as a duplication of each other is a subjective task (REFERENCE HERE). 
-To mitigate this human aspect on the given problem, the literature classified Code Duplication on the scale of changes in between 
-the code artifacts in four main types of Code Duplication \citep{litreview}. Given two code artifacts F and G, 
+To mitigate this human aspect on the given problem, the literature classified code duplication on the scale of changes between 
+the code artifacts in four main types of code dlone duplication \citep{litreview}. Given two code artifacts F and G, 
 the types of Code Duplication are described as follow \citep{litreview}:
@@
 
-\index{Tese/Dissertação!versões}
+\subsection{Literature approches for code clone detection}
+
+Given the passage of time, the literature developed many techniques and methods to approach the code clone detection problem. 
+For better understanding and analyze of these approachs, the literature divided the researches done in
+five main methodologies given the main nature of their work \citep{litreview} , which are:
+
+\begin{itemize}
+	\item \textbf{Textual-based approaches:} The literature review made by Chen \citep{litreview}  states that this methodology
+	is the first to be investigated by the literature. This line of methodology is based on see code as a text artifact and apply
+	textual approachs for text similarity. One example of this methodology applied is the work of Roy and Cordy \citep{textexample}, 
+	which uses parsing and grammar techniques to detect code clones.
+
+	\item \textbf{Token-based approaches:} This methodoly tries to approach the problem by seeing the code in a similar way to
+	the lexical analysis of the compiling process \citep{litreview} . That means doing work in the line of recognizing constants, keywords and other 
+	specific programming languages tokens, mapping variables to <pair,value> representations to become naming-independent, etc. One example of 
+	this methodology applied  is the work of Toomey et al. \citep{tokenexample} to detect code clones through hashed token sequences.
+
+	\item \textbf{Tree-based approaches:} While the token-based approach uses the lexical analysis of the compiling process, the 
+	tree-based approach uses the syntax tree, the structure that  stores the semantic information of a code artifact in respect of the 
+	programming language \citep{compiler}. One example of research that uses this information to detect code clones is the one done by
+	Chilowicz and Duris \citep{treeexample}.
+
+	\item \textbf{Metric-based approaches:} This methodology approches the problem by extracting metrics from the code artifact and use 
+	those metrics as a embedding representation of the artifact to measure similarity \citep{litreview}, similarly to the famous  
+	Word2Vec work done by Mikolov et al \citep{wordtovec}. Examples of metrics can be program size, number of variables, number of 
+	access to memory, etc. One example of this methodology applied is the work done by Kaur and Sharma. \citep{metricexample}
+
+	\item \textbf{Graph-based approaches:} Similarly to the Tree-based approaches, this methotodology depends heavily on an intermediate
+	program representation, called program dependence graph (PDG) \citep{prodg}. The representation store information 
+	about program's control, statement and data dependencies. One example of usage of the program dependence graph is 
+	the recently work done by Liu et al. \citep{tailor}.
+\end{itemize}

### 2024-06-06 conteudo/01-exemplo-normas-ime.tex
More work of 05/06/2024

@@
 
-\subsection{Types of code duplication}
+\subsection{Types of code clone duplication}
 
@@
 To mitigate this human aspect on the given problem, the literature classified code duplication on the scale of changes between 
-the code artifacts in four main types of code dlone duplication \citep{litreview}. Given two code artifacts F and G, 
+the code artifacts in four main types of code clone duplication \citep{litreview}. Given two code artifacts F and G, 
 the types of Code Duplication are described as follow \citep{litreview}:
@@
 \end{itemize}
+
+Given the complexity of Types 3 and 4 of code duplications, the state-of-art approches that aims at this type of
+duplication constantly have a high cost in computational time and memory space \citep{letreview}, which is a concern in our research 
+that aim to work in large codebases scopes. For this reason we will look after mainly at Type 1 and 2 of code duplications.
 

### 2024-06-06 conteudo/01-exemplo-normas-ime.tex
Fix error

@@
 
-Given the complexity of Types 3 and 4 of code duplications, the state-of-art approches that aims at this type of
-duplication constantly have a high cost in computational time and memory space \citep{letreview}, which is a concern in our research 
-that aim to work in large codebases scopes. For this reason we will look after mainly at Type 1 and 2 of code duplications.
+Given the complexity of  types 3 and 4 of code duplications, the state-of-art approches that aims at this type of
+duplication constantly have a high cost in computational time and memory space \citep{litreview}, which is a concern in our research 
+that aim to work in large codebases scopes. For this reason we will look after mainly at type 1 and 2 of code duplications.
 

### 2024-06-06 conteudo/01-exemplo-normas-ime.tex
move to a new file the codeduplication part

--- 2024-06-06 conteudo/01-exemplo-normas-ime.tex
@@
 
-\section{Code Duplication Detection}
-
-Code Duplication Detection is a field in computer science studies that gets the attention of researchers back to 1988 \citep{firstman}.
-The occurrence of Code Duplication is a harmful artifact to have in a software, affecting software related tasks such as code readability,
-introdution of bugs, etc.   \citep{harmone}. 
-On most of the cases, the occurrence of code duplication tends to create a more unstable software then the nonduplicate code counterpart.   \citep{harmtwo}
-
-Through the years, the subject of studies about Code Duplication Detection branched out mainly in two fields of study,
-Code Clone Detection and Code Plagiarism Detection,
-with the former one focused on the technical aspect while the later one also introduces an social aspect on the field of research
-\citep{litreview}. For our research we are only interested on the Code Clone Detection branch given the nature of our work.
-
-\subsection{Types of code duplication}
-
-The categorization of two code artifacts as a duplication of each other is a subjective task (REFERENCE HERE). 
-To mitigate this human aspect on the given problem, the literature classified code duplication on the scale of changes between 
-the code artifacts in four main types of code dlone duplication \citep{litreview}. Given two code artifacts F and G, 
-the types of Code Duplication are described as follow \citep{litreview}:
-
-\begin{itemize}
-	\item \textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names, white spaces and any other kind of irrelevant elements.
-(ADD FIGURE) shows a typical example of Type-1 Code Duplication \citep{litreview}.
-
-	\item \textbf{Type-2:} The differences between F and G is the same as the Type-1, with the addition of considering addition and deletion of redudant codes.
-(ADD FIGURE) shows a typical example of Type-2 Code Duplication \citep{litreview}.
-
-	\item \textbf{Type-3:} The difference between F and G is the same as the Type-2, with the addition of considering the reorder of code blocks as well as statements within code blocks.
-(ADD FIGURE) shows a typical example of Type-3 Code Duplication \citep{litreview}.
-
-	\item \textbf{Type-4:} The difference between F and G is the same as the Type-3, with the addtionf of considering changes on data sctrucure, the order of operands/operators in expresssions, or replacing part of codes with equivalent composition.
-(ADD FIGURE) shows a typical example of Type-4 Code Duplication \citep{litreview}.
-\end{itemize}
-
-\subsection{Literature approches for code clone detection}
-
-Given the passage of time, the literature developed many techniques and methods to approach the code clone detection problem. 
-For better understanding and analyze of these approachs, the literature divided the researches done in
-five main methodologies given the main nature of their work \citep{litreview} , which are:
-
-\begin{itemize}
-	\item \textbf{Textual-based approaches:} The literature review made by Chen \citep{litreview}  states that this methodology
-	is the first to be investigated by the literature. This line of methodology is based on see code as a text artifact and apply
-	textual approachs for text similarity. One example of this methodology applied is the work of Roy and Cordy \citep{textexample}, 
-	which uses parsing and grammar techniques to detect code clones.
-
-	\item \textbf{Token-based approaches:} This methodoly tries to approach the problem by seeing the code in a similar way to
-	the lexical analysis of the compiling process \citep{litreview} . That means doing work in the line of recognizing constants, keywords and other 
-	specific programming languages tokens, mapping variables to <pair,value> representations to become naming-independent, etc. One example of 
-	this methodology applied  is the work of Toomey et al. \citep{tokenexample} to detect code clones through hashed token sequences.
-
-	\item \textbf{Tree-based approaches:} While the token-based approach uses the lexical analysis of the compiling process, the 
-	tree-based approach uses the syntax tree, the structure that  stores the semantic information of a code artifact in respect of the 
-	programming language \citep{compiler}. One example of research that uses this information to detect code clones is the one done by
-	Chilowicz and Duris \citep{treeexample}.
-
-	\item \textbf{Metric-based approaches:} This methodology approches the problem by extracting metrics from the code artifact and use 
-	those metrics as a embedding representation of the artifact to measure similarity \citep{litreview}, similarly to the famous  
-	Word2Vec work done by Mikolov et al \citep{wordtovec}. Examples of metrics can be program size, number of variables, number of 
-	access to memory, etc. One example of this methodology applied is the work done by Kaur and Sharma. \citep{metricexample}
-
-	\item \textbf{Graph-based approaches:} Similarly to the Tree-based approaches, this methotodology depends heavily on an intermediate
-	program representation, called program dependence graph (PDG) \citep{prodg}. The representation store information 
-	about program's control, statement and data dependencies. One example of usage of the program dependence graph is 
-	the recently work done by Liu et al. \citep{tailor}.
-\end{itemize}
+\input{conteudo/background/codeduplication}
+\par

### 2024-06-06 conteudo/background/codeduplication.tex
move to a new file the codeduplication part

\en

\section{Code Duplication Detection}

Code Duplication Detection is a field in computer science studies that gets the attention of researchers back to 1988 \citep{firstman}.
The occurrence of Code Duplication is a harmful artifact to have in a software, affecting software related tasks such as code readability,
introdution of bugs, etc.   \citep{harmone}. 
On most of the cases, the occurrence of code duplication tends to create a more unstable software then the nonduplicate code counterpart.   \citep{harmtwo}

Through the years, the subject of studies about Code Duplication Detection branched out mainly in two fields of study,
Code Clone Detection and Code Plagiarism Detection,
with the former one focused on the technical aspect while the later one also introduces an social aspect on the field of research
\citep{litreview}. For our research we are only interested on the Code Clone Detection branch given the nature of our work.

\subsection{Types of code clone duplication}

The categorization of two code artifacts as a duplication of each other is a subjective task (REFERENCE HERE). 
To mitigate this human aspect on the given problem, the literature classified code duplication on the scale of changes between 
the code artifacts in four main types of code clone duplication \citep{litreview}. Given two code artifacts F and G, 
the types of Code Duplication are described as follow \citep{litreview}:

\begin{itemize}
	\item \textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names, white spaces and any other kind of irrelevant elements. (ADD FIGURE) shows a typical example of Type-1 Code Duplication \citep{litreview}.

	\item \textbf{Type-2:} The differences between F and G is the same as the Type-1, with the addition of considering addition and deletion of redudant codes. (ADD FIGURE) shows a typical example of Type-2 Code Duplication \citep{litreview}.

	\item \textbf{Type-3:} The difference between F and G is the same as the Type-2, with the addition of considering the reorder of code blocks as well as statements within code blocks. (ADD FIGURE) shows a typical example of Type-3 Code Duplication \citep{litreview}.

	\item \textbf{Type-4:} The difference between F and G is the same as the Type-3, with the addtionf of considering changes on data sctrucure, the order of operands/operators in expresssions, or replacing part of codes with equivalent composition. (ADD FIGURE) shows a typical example of Type-4 Code Duplication \citep{litreview}. 
\end{itemize}

Given the complexity of  types 3 and 4 of code duplications, the state-of-art approches that aims at this type of
duplication constantly have a high cost in computational time and memory space \citep{litreview}, which is a concern in our research 
that aim to work in large codebases scopes. For this reason we will look after mainly at type 1 and 2 of code duplications.

\subsection{Literature approches for code clone detection}

Given the passage of time, the literature developed many techniques and methods to approach the code clone detection problem. 
For better understanding and analyze of these approachs, the literature divided the researches done in
five main methodologies given the main nature of their work \citep{litreview} , which are:

\begin{itemize}
	\item \textbf{Textual-based approaches:} The literature review made by Chen \citep{litreview}  states that this methodology
	is the first to be investigated by the literature. This line of methodology is based on see code as a text artifact and apply
	textual approachs for text similarity. One example of this methodology applied is the work of Roy and Cordy \citep{textexample}, 
	which uses parsing and grammar techniques to detect code clones.

	\item \textbf{Token-based approaches:} This methodoly tries to approach the problem by seeing the code in a similar way to
	the lexical analysis of the compiling process \citep{litreview} . That means doing work in the line of recognizing constants, keywords and other 
	specific programming languages tokens, mapping variables to <pair,value> representations to become naming-independent, etc. One example of 
	this methodology applied  is the work of Toomey et al. \citep{tokenexample} to detect code clones through hashed token sequences.

	\item \textbf{Tree-based approaches:} While the token-based approach uses the lexical analysis of the compiling process, the 
	tree-based approach uses the syntax tree, the structure that  stores the semantic information of a code artifact in respect of the 
	programming language \citep{compiler}. One example of research that uses this information to detect code clones is the one done by
	Chilowicz and Duris \citep{treeexample}.

	\item \textbf{Metric-based approaches:} This methodology approches the problem by extracting metrics from the code artifact and use 
	those metrics as a embedding representation of the artifact to measure similarity \citep{litreview}, similarly to the famous  
	Word2Vec work done by Mikolov et al \citep{wordtovec}. Examples of metrics can be program size, number of variables, number of 
	access to memory, etc. One example of this methodology applied is the work done by Kaur and Sharma. \citep{metricexample}

	\item \textbf{Graph-based approaches:} Similarly to the Tree-based approaches, this methotodology depends heavily on an intermediate
	program representation, called program dependence graph (PDG) \citep{prodg}. The representation store information 
	about program's control, statement and data dependencies. One example of usage of the program dependence graph is 
	the recently work done by Liu et al. \citep{tailor}.
\end{itemize}

### 2024-06-06 conteudo/background/codeduplication.tex
Add temporary end text for code duplication detection

@@
 
-	\item \textbf{Type-2:} The differences between F and G is the same as the Type-1, with the addition of considering addition and deletion of redudant codes.
+	\item \textbf{Type-2:} The differences between F and G is the same as the type-1, with the addition of considering addition and deletion of redudant codes.
 (ADD FIGURE) shows a typical example of Type-2 Code Duplication \citep{litreview}.
 
-	\item \textbf{Type-3:} The difference between F and G is the same as the Type-2, with the addition of considering the reorder of code blocks as well as statements within code blocks.
+	\item \textbf{Type-3:} The difference between F and G is the same as the type-2, with the addition of considering the reorder of code blocks as well as statements within code blocks.
 (ADD FIGURE) shows a typical example of Type-3 Code Duplication \citep{litreview}.
 
-	\item \textbf{Type-4:} The difference between F and G is the same as the Type-3, with the addtionf of considering changes on data sctrucure, the order of operands/operators in expresssions, or replacing part of codes with equivalent composition.
+	\item \textbf{Type-4:} The difference between F and G is the same as the type-3, with the addtionf of considering changes on data sctrucure, the order of operands/operators in expresssions, or replacing part of codes with equivalent composition.
 (ADD FIGURE) shows a typical example of Type-4 Code Duplication \citep{litreview}.
 \end{itemize}
-
-Given the complexity of  types 3 and 4 of code duplications, the state-of-art approches that aims at this type of
-duplication constantly have a high cost in computational time and memory space \citep{litreview}, which is a concern in our research 
-that aim to work in large codebases scopes. For this reason we will look after mainly at type 1 and 2 of code duplications.
 
@@
 \end{itemize}
+
+Given the necessity to evaluate and compare approches for code clone detection, it was developed and widely-adopted two datasets: 
+OJ-Clone, a dataset collected from a pedagogical online judge system \citep{ojclone}, and BigCloneBench, a datatset collected from over 
+25,000 Java software systems \citep{bigclonebench}. The existence of only two spreaded datasets is a open concern in literature as 
+the limited scope of OJ-Clone and the problems of BigCloneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail}
+
+The state-of-art methods for code clone detection usually have high computational time and memory space usage, as well being
+programming language specific approches, which cannot be easily implemented in programming languagues that were not used in 
+their research \citep{litreview}. These are main concern for our research and for this reason we will use a generalist textual-based 
+approach based solely on Natural Languague Processing (NLP) text similarity detection methods, which we will introduce and compare
+with the literature approches later on this document.

### 2024-06-07 conteudo/background/nlpsimilarity.tex
Add initial NLP, work of 06/06

\en

\section{Text similarity detection methods}

The standard methodology for text similarity detection are based on building vector embeddings of
the texts and apply a distance function in these embeddings to measure the similarity between them. 
There are multiple methods for building vector embeddings, such as bag-of-words, TF-IDF, LSI, Word2Vec \citep{gensimlivro}. 
The recent works related to build vector embeddings uses Largue Language Models (LLM), for example the usage of BERT, ELMo 
and GPT \citep{llmsimilar}. The most popular distance function used by the literature are the cosine similarity 
defined by the formula \citep{cosineref}:

$$\text{cosine similarity} = SC(A,B) = \frac{ A \cdot B}{ \lVert A \rVert \lVert B \rVert }$$

Where $A$ and $B$ are the vector embeddings of the two texts. One of the advantages of the cosine similarity is that $SC(A,B)$ 
is a number between $0$ and $1$, where $SC(A,B)$ values more close to the number $1$ means that two vectors are more similar.

It is a famous fact that LLM have high computational cost, one example of such is a experiment done by Reimers and Gurevych for 
finding the most similar pair in a collection of 10,000 sentences, that required about 65 hours with BERT \citep{bertsimilar}. This 
is a similar magnitude of comparation pairs we want to do through our work and for this reason we chose to not use this kind of 
vector embedding method, but it is a valid solution for smaller codebases.

### 2024-06-07 conteudo/background/nlpsimilarity.tex
Add gensim information, 07/06

@@
 vector embedding method, but it is a valid solution for smaller codebases.
+
+\subsection{Gensim}
+
+Gensim is a open source python library \citep{gensim} which the owners declare themselves as "The fastest library 
+for training of vector embeddings - Python or otherwise. The core algorithms in Gensim use battle-hardened, 
+highly optimized and parallelized C routines." \citep{gensimsite}. For the velocity of gensim we chose to use this library
+for doing our vector embeddings.
+
+In this work we will not explore more alternatives for vector embeddings. We do not want to optimize the computational 
+cost or the code clone detection to the state-of-art as they are not the main objectives of our research.

### 2024-06-11 conteudo/background/linux.tex
Add start linux background, 11/06

\en

\section{The Linux Kernel}

The Linux kernel is an Unix-like operating system published by Linus Torvalds in 1992 as a free software and
it started becoming popular in late 1990s \citep{linuxbook}. Against popular belief, linux is not a full 
operating system as it do not includes applications such as filesystem utilities, windowing systems. The "Linux" 
distributions know to the public are in fact GNU/Linux operating systems \citep{gnuref}.

TEXT ABOUT LINUX NUMBER SCOPE ETC.

The Linux project is organized as a set of subsystem such as process scheduler, memory management, 
device driver infrastruture, networking and filesystems \citep{melissa}.
Each subsystem usually have a maintainer or a group of maintainers, which are people responsible 
for managing and accepting contributions into the subsystem they care of it \citep{melissa}.

The contributions to the Linux are coordinate with the usage of the Git source management tool. The contributions 
made to the project are formated as patches, which are text documents describing differences between two different 
versions of a source tree \citep{melissa}.

TALK ABOUT HOW THE TOOL ENABLE A NEW WAY TO START CONTRITUTING TO THE KERNEL

AMD Display Driver is a linux's subsystem responsible for the management of AMD GPU's. This subsystem have a problem 
with code clone duplication acknowledged by the subsystem's maintainters. For this reason, we selected this
subsystem as the primary focus for our empirical analysis in our research.

### 2024-06-12 conteudo/background/linux.tex
Add another part of the process

@@
 
-TEXT ABOUT LINUX NUMBER SCOPE ETC.
-
-The Linux project is organized as a set of subsystem such as process scheduler, memory management, 
+The Linux kernel is monolithic organized as a set of subsystem such as process scheduler, memory management, 
 device driver infrastruture, networking and filesystems \citep{melissa}.
 Each subsystem usually have a maintainer or a group of maintainers, which are people responsible 
-for managing and accepting contributions into the subsystem they care of it \citep{melissa}.
+for managing and accepting contributions into the subsystem they care of \citep{melissa}.
 
@@
 
-TALK ABOUT HOW THE TOOL ENABLE A NEW WAY TO START CONTRITUTING TO THE KERNEL
-
 AMD Display Driver is a linux's subsystem responsible for the management of AMD GPU's. This subsystem have a problem 
-with code clone duplication acknowledged by the subsystem's maintainters. For this reason, we selected this
-subsystem as the primary focus for our empirical analysis in our research.
+with code clone duplication, which are acknowledged by the subsystem's maintainters. For this reason, we selected this
+subsystem as the primary focus for refactoring the detected code clone duplication by our tool.
+(MAYBE CHANGE TO: For this reason, we selected this subsystem as the primary focus for our empirical analysis in our research).

### 2024-06-12 conteudo/background/refactor.tex
Add another part of the process

\en

\section{Refactoring}

Fowler defines refactoring in his book as "a controlled technique for imporving the design of an existing code base. Its
essence is applying a series of small behavior-preserving transformations, each of which too small to be worth doing. However
the cumulative effect of each of these transformations is quite significant." \citep{refactorbook}. The elimination of detected 
code clone duplication is a refactoring task that necessitates the application of refactoring methods. 
Fowler presents multiples refactoring methods such as parameterize method, replace parameter with explicit methods and 
replace parameter with method \citep{refactorbook}. There is a briefly explanation of the methods used in our research below.

\subsection{Parameterize method}

TODO. I EXPECT TO FILL THIS SECTION AS THE WORK EVOLVES.

### 2024-06-12 conteudo/tool/tool.tex
Add starting information about the tool, 12/06}

\en

Our proposed tool is a Command Line Interface (CLI) software, the tool purpose is to enable developers to query 
information about code clone duplication in a function level. The tool is separated in two steps. the 
first one, denoted as Setup, is responsible for processing the codebase doing tasks such as extracting functions of the code files 
and searching for code clone duplications. The second step, denoted as Query Responder, is
responsible to execute the tool functionalities using the information obtained in the Setup step. More information about the 
Setup and Query Responder behavior can be found in \ref{subsec:architecture} and in \ref{subsec:func}.

\subsection{Tool architecture}
\label{subsec:architecture}

\subsection{Tool functionalites}
\label{subsec:func}

### 2024-06-13 conteudo/tool/tool.tex
add initial tool, 13/06

@@
 Our proposed tool is a Command Line Interface (CLI) software, the tool purpose is to enable developers to query 
-information about code clone duplication in a function level. The tool is separated in two steps. the 
-first one, denoted as Setup, is responsible for processing the codebase doing tasks such as extracting functions of the code files 
+information about code clone duplication in a function level, by function level we mean to find functions in a codebase that 
+are code duplications. The tool is separated in two steps. The 
+first one, denoted as Setup, is executed only once for each codebase and it is responsible for processing the
+codebase doing tasks such as extracting functions of the code files 
 and searching for code clone duplications. The second step, denoted as Query Responder, is
@@
 
+The Setup step is composed of two components, the Function Breaker and the Code Duplication Finder. the pipeline of this step work as
+follow: the Function Parser receives the codebase we are interested, extract the functions of the codebase along with metadata and
+creates a new temporary codebase where it function becomes a new code file. The Code Duplication Finder iterate over every pair of 
+files in the temporary codebase computing the code duplications and save them in the Code Duplication Database, which is a text file
+that stores every code duplication as a triple <function1,function2,similarity>, where function1 and function2 are the functions that 
+are a duplication of each other, and similarity is the metric given by the code duplication method utilized, which in our case, it is
+the cossine similiratiy explained in \ref{sec:similarity}. Finally, the Query Responder consumes the temporary codebase and the 
+Code Duplication Database to extract code duplication related information as per user request. A diagram to ilustrate the pipeline
+can be found in the Figure (ADD FIGURE HERE). There is a explanation of how each component work below.
+
+\subsubsection{Input codebase}
+
+The input codebase is expected to be a folder in the machine the tool is running into. All files in the codebase that is not a source
+code file from a programming language supported by the tool will be ignored.
+
+\subsubsection{Function Parser}
+
+\subsubsection{Code Duplication Finder}
+
+\subsubsection{Code Duplication Database}
+
+\subsubsection{Setup step}
+
+\subsubsection{Query Responder step}
+
 \subsection{Tool functionalites}

### 2024-06-14 conteudo/tool/architecture.tex
Add additional work of 14/06

--- 2024-06-13 conteudo/tool/tool.tex
@@
 \en
-
-Our proposed tool is a Command Line Interface (CLI) software, the tool purpose is to enable developers to query 
-information about code clone duplication in a function level, by function level we mean to find functions in a codebase that 
-are code duplications. The tool is separated in two steps. The 
-first one, denoted as Setup, is executed only once for each codebase and it is responsible for processing the
-codebase doing tasks such as extracting functions of the code files 
-and searching for code clone duplications. The second step, denoted as Query Responder, is
-responsible to execute the tool functionalities using the information obtained in the Setup step. More information about the 
-Setup and Query Responder behavior can be found in \ref{subsec:architecture} and in \ref{subsec:func}.
 
@@
 
+The Function Parser receives the input codebase and transforms it into the temporary codebase. The Function Parser iterates through
+every source code file from the programming language it support and use a specific programming languague's function 
+extractor to extract
+every function in the file, for each function extracted we create a new source code file and a metadata file in the temporary codebase
+which is represented by the pair \textbf{<source code file, function extracted>}, which represents the source code file the 
+function is extracted from and the 
+proper function. The new source code file only contains the source code from the function it represents, while the metadata file 
+contains extra relevant information such as function name, the line the function it began to be declared in the original souce code 
+file and the line the function's body ends in the original source code file. The programming language we support at the moment 
+is C and Java. Can be found a example of a transform of a source code in the (CRIAR UM APENDICE E COLOCAR AQUI).
+
+About the programming languague's function extractor, we approached this problem in a way that any person comfortable with a language
+can adapt our approach to their specific programming language. Usually, the code blocks can be easily extracted from 
+source code files, independently if the code blocks are defined by curly brackets or by indentation, and it is possible to
+infer if it represents a function's body given how deeply it is in the code context and a few lines before the start 
+of the code block. We followed exactly this approach to implement the function extractor for the languagues we implemented. 
+As a alternative approach, we can build the syntax tree \citep{compiler} of the programming language and use it to 
+extract the functions, which is a more complex task then our main approach. For the programming languages we support, we implemented
+it as follow:
+
+\begin{itemize}
+  \item \textbf{C}:
+  \item \textbf{Java}:
+\end{itemize}
+
 \subsubsection{Code Duplication Finder}
@@
 \subsubsection{Query Responder step}
-
-\subsection{Tool functionalites}
-\label{subsec:func}

### 2024-06-14 conteudo/tool/functionalities.tex
Add additional work of 14/06

--- 2024-06-12 conteudo/tool/tool.tex
@@
 \en
-
-Our proposed tool is a Command Line Interface (CLI) software, the tool purpose is to enable developers to query 
-information about code clone duplication in a function level. The tool is separated in two steps. the 
-first one, denoted as Setup, is responsible for processing the codebase doing tasks such as extracting functions of the code files 
-and searching for code clone duplications. The second step, denoted as Query Responder, is
-responsible to execute the tool functionalities using the information obtained in the Setup step. More information about the 
-Setup and Query Responder behavior can be found in \ref{subsec:architecture} and in \ref{subsec:func}.
-
-\subsection{Tool architecture}
-\label{subsec:architecture}
 

### 2024-06-14 conteudo/tool/tool.tex
Add additional work of 14/06

--- 2024-06-12 conteudo/tool/tool.tex
@@
 Setup and Query Responder behavior can be found in \ref{subsec:architecture} and in \ref{subsec:func}.
-
-\subsection{Tool architecture}
-\label{subsec:architecture}
-
-\subsection{Tool functionalites}
-\label{subsec:func}

### 2024-06-14 conteudo/tool/architecture.tex
Add more work, 14/06

@@
 
-\subsection{Tool architecture}
+\section{Tool architecture}
 \label{subsec:architecture}
@@
 
-\subsubsection{Input codebase}
+\subsection{Input codebase}
 
@@
 
-\subsubsection{Function Parser}
+\subsection{Function Parser}
 
@@
 file and the line the function's body ends in the original source code file. The programming language we support at the moment 
-is C and Java. Can be found a example of a transform of a source code in the (CRIAR UM APENDICE E COLOCAR AQUI).
+is C and Java. Can be found a example of a transform of a source code in the (CREATE APENDIX AND ADD HERE).
 
@@
 
-\subsubsection{Code Duplication Finder}
+\subsection{Code Duplication Finder}
 
-\subsubsection{Code Duplication Database}
+The Code Duplication Finder iterates through every pair of source code files in the temporary codebase, which in this context represents
+functions of the input codebase, and for every pair of files, we execute a code duplication detection method that computes a metric
+that measures of how similar the pair of files are, which we call similarity. Finally, if the similarity is greater of equal then the
+minimum similarity threshold (a parameter passed by the user when it executes the Setup step), we store this pair of files along with
+it similarity in the Code Duplication Database.
 
-\subsubsection{Setup step}
+For the code duplication detection method, we visualize the source code files as texts and execute the TF-IDF vector embedding method
+implemented by the Gensim library \citep{gensim} and compute the cosine similarity as the similarity metric. We chose this method 
+given the alegated Gensim statements about itself performance, being a programming language independent method and the nonnecessity of 
+compilable code, which is expected to happen in the temporary codebase as it not stores complete code artifacts. Platis implemented
+this code cuplication detection method and distributed it to the public with a MIT license \citep{platistool} \citep{mitlicense}, 
+we utilize his implementation as it has some nice to have components such as removing code comments, we do the appropriated changes
+in the input and output of his implementation to respect the expected format of our tool. 
+It is feasible to change the duplication detection method in a manner similar to what we did with Platis's implementation.
 
-\subsubsection{Query Responder step}
+\subsection{Code Duplication Database}
+
+\subsection{Setup step}
+
+\subsection{Query Responder step}

### 2024-06-15 conteudo/tool/architecture.tex
Add more work, 14/06

@@
 extractor to extract
-every function in the file, for each function extracted we create a new source code file and a metadata file in the temporary codebase
+every function in the file, for each function extracted we create two new source code files and a metadata file in the temporary codebase
 which is represented by the pair \textbf{<source code file, function extracted>}, which represents the source code file the 
 function is extracted from and the 
-proper function. The new source code file only contains the source code from the function it represents, while the metadata file 
-contains extra relevant information such as function name, the line the function it began to be declared in the original souce code 
+proper function. The first new source code file contains the function's body from the function it represents, while the other
+new source code file contains the function's signature, the metadata file 
+contains extra relevant information such as function name, the line the function signature starts in the original souce code 
 file and the line the function's body ends in the original source code file. The programming language we support at the moment 
-is C and Java. Can be found a example of a transform of a source code in the (CREATE APENDIX AND ADD HERE).
+is C and Java. Can be found a example of a transform of a source code in the Figure (CREATE APENDIX AND ADD HERE).
 
@@
 \begin{itemize}
-  \item \textbf{C}:
-  \item \textbf{Java}:
+	\begin{item}
+		\textbf{C}:
+	\end{item}
+	\begin{item}
+		\textbf{Java}:
+	\end{item}
 \end{itemize}
+
+\subsection{Temporary codebase}
+
+The temporary codebase is a transformation of the input codebase done by the Function Parser, every function in the input codebase
+is represented in the temporary codebase as three files. There are description of each file below and an example of each can found
+in the figure (CREATE FIGURE AND ADD HERE).
+
+\begin{itemize}
+	\begin{item}
+		\textbf{Header file}: This file contains the signature of the function it represents;
+	\end{item}
+	\begin{item}
+		\textbf{Source file:} This file contains the body of the function it represents;
+	\end{item}
+	\begin{item}
+		\textbf{Info file:} This file contains metadata of the function. The information this file contains is:
+		the function name, the file name that has the function, the relative path of the file related to the input codebase, 
+		the line that the function's signatures starts, the line that the functions body starts, the line that the function's body 
+		ends and if there is a line break between the end of the function's signature and the start of the function's body.
+
+	\end{item}
+\end{itemize}
+
+There is a example of each file in the Appendix
 
@@
 
+The Code Duplication Database is a text file that contains the duplicated pairs of functions found by the Code Duplication Finder, 
+the first line of the file contains a number that represents the number of the duplicated pairs, after that, there is 
+a line of each duplicated pairs that contains the path of the functions in the temporary codebase and the similarity metric of the 
+pair. A example of the Code Duplication Database can be found in the figure (ADD FIGURE HERE).
+
 \subsection{Setup step}

### 2024-06-15 conteudo/background/codeduplication.tex
format a little bit the weird things done before

@@
 
-Code Duplication Detection is a field in computer science studies that gets the attention of researchers back to 1988 \citep{firstman}.
-The occurrence of Code Duplication is a harmful artifact to have in a software, affecting software related tasks such as code readability,
+Code Duplication Detection is a field in computer science studies that gets the attention of researchers 
+back to 1988 \citep{firstman}.
+The occurrence of Code Duplication is a harmful artifact to have in a software, affecting software related 
+tasks such as code readability,
 introdution of bugs, etc.   \citep{harmone}. 
-On most of the cases, the occurrence of code duplication tends to create a more unstable software then the nonduplicate code counterpart.   \citep{harmtwo}
+On most of the cases, the occurrence of code duplication tends to create a more unstable software then the 
+nonduplicate code counterpart.   \citep{harmtwo}
 
@@
 \begin{itemize}
-	\item \textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names, white spaces and any other kind of irrelevant elements.
-(ADD FIGURE) shows a typical example of Type-1 Code Duplication \citep{litreview}.
+	\begin{item}
+		\textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names, white spaces 
+		and any other kind of irrelevant elements. (ADD FIGURE) shows a typical example of Type-1 Code 
+		Duplication \citep{litreview}. 
+	\end{item}
+	\begin{item}
+		\textbf{Type-2:} The differences between F and G is the same as the type-1, with the addition of considering 
+		addition and deletion of redudant codes. (ADD FIGURE) shows a typical example of Type-2 Code Duplication \citep{litreview}. 
+	\end{item}
+	\begin{item}
+		\textbf{Type-3:} The difference between F and G is the same as the type-2, with the addition of considering 
+		the reorder of code blocks as well as statements within code blocks. (ADD FIGURE) shows a typical example of 
+		Type-3 Code Duplication \citep{litreview}. 
+	\end{item}
+	\begin{item}
+		\textbf{Type-4:} The difference between F and G is the same as the type-3, with the addition of considering 
+		changes on data sctrucure, the order of operands/operators in expresssions, or replacing part of codes with 
+		equivalent composition. (ADD FIGURE) shows a typical example of Type-4 Code Duplication \citep{litreview}. 
+	\end{item}
 
-	\item \textbf{Type-2:} The differences between F and G is the same as the type-1, with the addition of considering addition and deletion of redudant codes.
-(ADD FIGURE) shows a typical example of Type-2 Code Duplication \citep{litreview}.
-
-	\item \textbf{Type-3:} The difference between F and G is the same as the type-2, with the addition of considering the reorder of code blocks as well as statements within code blocks.
-(ADD FIGURE) shows a typical example of Type-3 Code Duplication \citep{litreview}.
-
-	\item \textbf{Type-4:} The difference between F and G is the same as the type-3, with the addtionf of considering changes on data sctrucure, the order of operands/operators in expresssions, or replacing part of codes with equivalent composition.
-(ADD FIGURE) shows a typical example of Type-4 Code Duplication \citep{litreview}.
 \end{itemize}

### 2024-06-15 conteudo/tool/architecture.tex
Add some work, 15/06

@@
 
-There is a example of each file in the Appendix
-
 \subsection{Code Duplication Finder}
@@
 
+The Setup step is a procedure executed by the user once per codebase. The Setup step receives the input codebase and the minimum
+similarity metric value to a pair of functions to be considered a duplication, then it executes the Function Parser and the Code 
+Duplication Finder, to generate the temporary codebase and the Code Duplication Database, which will be later used by the Query
+Responder step. The value of the minimum similarity metric is to reduce the Code Duplication Database, which optimize the
+memory usage and the computational cost of the Query Responder step. For example, if this parameter does not exist and we work in 
+a codebase of $10000$ functions and supposing each relative path plus the function name have $50$ charactes, the Code Duplication 
+Database will be a file of size approximately $5000^2 \times 50 ~= 5$ Gigabytes, as it is not expected to a real large codebase 
+to the majority of pairs of functions be a duplication of each other, this number can be reduced dramastically.
+
 \subsection{Query Responder step}
+
+The Query Responder step is the component that the user executes multiples times to consult information about code duplication in
+a codebase that was previously preprocessed by the Setup step. The Query Responder step consumes the temporary codebase and the 
+Code Duplication Database created by the Setup step. This component implements all the tool's functionalities exposed 
+for the user, which are described in the \ref{subsec:func}.

### 2024-06-17 conteudo/tool/architecture.tex
complete initial version of architecture, 17/06

@@
 	\begin{item}
-		\textbf{C}:
+		\textbf{C}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in
+		the code context and som lines of code before the definition of the code block. By the C programming language's grammar, 
+		the functions are defined in a global context, with depth of 0. There are other elements with the same depth as functions,
+		we differentiate them by analysing if the first non-empty character before the curly brackets of the code block is a close 
+		parenthesis, since functions are the only elements that respect this pattern in the C programming language.
 	\end{item}
 	\begin{item}
-		\textbf{Java}:
+		\textbf{Java}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in
+		the code context and som lines of code before the definition of the code block. By the Java's grammar, 
+		the functions are defined inside classes. There are different depths which we can have functions, as Java accepts inner
+		classes. We differentiate functions of other elements by analysing if the first non-empty character before the curly brackets of the code block is a close 
+		parenthesis, since there are others elements which also respect this pattern such as for loop, we would need to apply other 
+		filters. We chose to only analyse functions of depth 1, which ignores every others elements that respect the previously
+		filters, but as counterpart, functions declared inside Inner Classes are not supported by our extractor. We made this choice
+		as there are not inner classes on the codebases explored in this work.
+
 	\end{item}

### 2024-06-18 conteudo/tool/functionalities.tex
Add more work, 14/06
add initial functionality

@@
 
-\subsection{Tool functionalites}
+\section{Tool functionalites}
 \label{subsec:func}
+
+We proposes three main functionalities for the user in the tool, which are named duplication explorer, function information
+and duplication relatory. Each of the functionalities recevies their adequated parameters to execute, with a exception
+of a optional parameter shared by all of them, which is minimum similarity per query. The minimum similarity per query is different
+from the siminum similarity presented in \ref{subsec:setup}, as it enables a volatile change in the minimum similarity interested in a query specific
+context. There is a explanation of each of the main functionalities throughout this section.
+
+\subsection{Duplication explorer}
+
+\subsection{Function information}
+
+The Function information purpose it to give a more in depth information about a specific function. The functionality inform the user
+about the relative path, the function name and the lines the function exist in the code file that the function is contained for the
+function that the user is interested and every function that is a duplication of it. As there ca nbe multiple functions with the 
+same name in a codebase, we distinguish then by the user providing a pattern string and we use the first function find that match
+the pattern. We define that the a function matches a pattern if the string formed by the concatenation of the relative path of the 
+code file plus the function name. 
+A example of the functionality usage can be found in (ADD FIGURE HERE).
+
+\subsection{Duplication relatory}
+
+The duplication relatory purpose is to give a overview of the input codebase. The functionality extract the information
+of how much lines of duplicated code is found per folder in the codebase and present it to the user in a reasonable format. To extract
+the information, we use the Code Duplication Database to list the code duplications, and the temporary code to know how much lines of
+duplicated code contains in each of the functions. (DO I TALK ABOUT THE TRIE USAGE TO MAINTAIN THE HIERARCHICAL TREE STRUCTURE? I 
+THINK NOT). A example of the functionality usage can be found in (ADD FIGURE HERE).

### 2024-06-20 conteudo/tool/architecture.tex
add initial functionality
add local work because got problems with latex, 19/06

@@
 are a duplication of each other, and similarity is the metric given by the code duplication method utilized, which in our case, it is
-the cossine similiratiy explained in \ref{sec:similarity}. Finally, the Query Responder consumes the temporary codebase and the 
+the cossine similarity explained in \ref{sec:similarity}. Finally, the Query Responder consumes the temporary codebase and the 
 Code Duplication Database to extract code duplication related information as per user request. A diagram to ilustrate the pipeline
-can be found in the Figure (ADD FIGURE HERE). There is a explanation of how each component work below.
+can be found in the Figure \ref{fig:diagram}. There is a explanation of how each component work below.
+
+\begin{figure}
+\includegraphics{diagrama_mestrado}
+\caption{The architecture diagram demonstrating the relationship between the components.}
+\label{fig:diagram}
+\end{figure}
 
@@
 
+\label{subsec:setup}
+
 The Setup step is a procedure executed by the user once per codebase. The Setup step receives the input codebase and the minimum

### 2024-06-20 conteudo/tool/functionalities.tex
add local work because got problems with latex, 19/06

@@
 of a optional parameter shared by all of them, which is minimum similarity per query. The minimum similarity per query is different
-from the siminum similarity presented in \ref{subsec:setup}, as it enables a volatile change in the minimum similarity interested in a query specific
-context. There is a explanation of each of the main functionalities throughout this section.
+from the siminum similarity presented in \ref{subsec:setup}, as it enables a volatile change in the minimum similarity
+interested in a query specific context. There is a explanation of each of the main functionalities throughout this section.
 
 \subsection{Duplication explorer}
+
+The duplication explorer is the principal functionality of our tool, the functionality purpose is to expose for the user the code duplication
+pairs found by our tool. We implement optional filters methods to enable more complex queries, which are described below. A example
+of the functionality usage can be found in (ADD FIGURE HERE).
+
+\begin{itemize}
+	\begin{item}
+		\textbf{Sorting filter:} This filter enables the user to sort the resuls by similarity of the functions pairs or 
+		by the number of duplicated lines. By default the tool sorts by similarity and the user can pass a sort parameter to change
+		the sorting method.
+	\end{item}
+
+	\begin{item}
+		\textbf{Limiter filter:} This filter receives a number parameter named limiter by the user, which limits the number of results
+		showed to the user by this given number.
+	\end{item}
+
+	\begin{item}
+		\textbf{Pattern Filter:} This filter make the explorer ignore every pair of duplicated functions that does not match a pattern
+		given by the user as a parameter. We say that a function matches a pattern if the string formed by the concatenation of
+		the relative path of the code file plus the function name. The user can define if both functions or at least one of the 
+		functions in the pair need to match the pattern by given a specific parameter.
+	\end{item}
+\end{itemize}
 
@@
 same name in a codebase, we distinguish then by the user providing a pattern string and we use the first function find that match
-the pattern. We define that the a function matches a pattern if the string formed by the concatenation of the relative path of the 
+the pattern. We say that a function matches a pattern if the string formed by the concatenation of the relative path of the 
 code file plus the function name. 

### 2024-06-20 conteudo/tool/functionalities.tex
do not know what is, done in 19/06

@@
 	\begin{item}
-		\textbf{Pattern Filter:} This filter make the explorer ignore every pair of duplicated functions that does not match a pattern
+		\textbf{Pattern filter:} This filter make the explorer ignore every pair of duplicated functions that does not match a pattern
 		given by the user as a parameter. We say that a function matches a pattern if the string formed by the concatenation of

### 2024-06-20 conteudo/preliminar/results.tex
Removing template artifacts and add final structure

\en

Base Content of the Preliminar Results of the research. TODO.

### 2024-06-20 conteudo/tool/functionalities.tex
add example figures

@@
 pairs found by our tool. We implement optional filters methods to enable more complex queries, which are described below. A example
-of the functionality usage can be found in (ADD FIGURE HERE).
+of the functionality usage can be found in the figure \ref{fig:explorer_ex}.
 
@@
 
+\begin{figure}
+\includegraphics{explorer_example}
+\caption{Explorer Example}
+\label{fig:explorer_ex}
+\end{figure}
+
 \subsection{Function information}
@@
 code file plus the function name. 
-A example of the functionality usage can be found in (ADD FIGURE HERE).
+A example of the functionality usage can be found in the figure \ref{fig:function_ex}.
+
+\begin{figure}
+\includegraphics{function_example}
+\caption{Function Example}
+\label{fig:function_ex}
+\end{figure}
 
@@
 duplicated code contains in each of the functions. (DO I TALK ABOUT THE TRIE USAGE TO MAINTAIN THE HIERARCHICAL TREE STRUCTURE? I 
-THINK NOT). A example of the functionality usage can be found in (ADD FIGURE HERE).
+THINK NOT). A example of the functionality usage can be found in the figure \ref{fig:relatory_ex}.
+
+\begin{figure}
+\includegraphics{relatory_example}
+\caption{Function Example}
+\label{fig:relatory_ex}
+\end{figure}

### 2024-06-20 conteudo/introduction/intro.tex
Add non numeric sections

\en

\section**{Problem outline}

\section**{Research Questions}

\section**{Research Design}

\section**{Manuscript Structure}

### 2024-06-21 conteudo/introduction/intro.tex
Add initial intro

@@
 \section**{Problem outline}
+
+The Linux kernel is a major project in the open source community, containing more then 19 million lines of code and more
+then thirteen thousand contributors along the lifespan of the project \citep{linuxquantity}. Given the magnitude of the project,
+it is unrealistic to expect that all contributors followed the best programming practices and principles throughout every 
+contributions they made to the project. It is expected to find a good amount of bad quality code artifacts in the source code, 
+which could result in hindering or compromising the maintenance of this project, as well as the indication of new features for it.
+
+Device driver, as stated by Madieu, are a "piece of software whose aim is to control and manage a particular hardware device, hence the
+name device driver" \citep{driverdef}. Device drivers are major part of the Linux kernel, representing 66 \% of the source code and
+responsible for the management of crucial element such as keyboards, mouses and GPU \citep{marcelo}. 
+The AMD Display driver of the Linux kernel is a core driver to enable AMD GPU's to operate correctly in desktop linux environments,
+which is a important duty, given that AMD GPU represented 19 \% of the personal GPU market in 2023 \citep{gpumarket}.
+
+When we iteracted with the maintainers of the AMD Display driver, they shared with us their pain with one of the characteristics
+bad quality code artifacts, that is duplicated code artifacts, which hinders the maintainment of the driver. While searching for a
+solution to them in the literature, we only found tools that given two code artifacts, infer if they are duplication of each other,
+and the research in this topic mainly focused in optimizing the accuracy of the tools in this given task. As we desired to the tools
+to find the duplicated code artifacts for us, instead of given two code artifact that we know it is a duplication, confirm to us that
+is in fact a duplication, the literature cannot be applied purely in our problem.
+
+As a alternative, we tried to search solutions in the gray literature, but without success. In the gray literature, the best solution
+we found were primitive solutions that finds every pair of code files that are duplications. As the AMD Display driver is a driver 
+created in 20XX and maintained until today, the code files that contains duplicated code normally also have specific code that is 
+only implemented in this file. Thus, a solution that finds duplication in a more deeply aspect is necessary to correct find 
+duplicatons.
 

### 2024-06-21 conteudo/introduction/intro.tex
finish initial version of problem outlined

@@
 solution to them in the literature, we only found tools that given two code artifacts, infer if they are duplication of each other,
-and the research in this topic mainly focused in optimizing the accuracy of the tools in this given task. As we desired to the tools
+and the research in this topic mainly focused in optimizing the accuracy of these tools in this given task. As we desired to the tools
 to find the duplicated code artifacts for us, instead of given two code artifact that we know it is a duplication, confirm to us that
@@
 we found were primitive solutions that finds every pair of code files that are duplications. As the AMD Display driver is a driver 
-created in 20XX and maintained until today, the code files that contains duplicated code normally also have specific code that is 
-only implemented in this file. Thus, a solution that finds duplication in a more deeply aspect is necessary to correct find 
-duplicatons.
+created in 20XX (TODO HERE) and maintained until today, the code files that contains duplicated code normally also have specific code that is 
+only implemented in this file. Thus, a solution that finds duplication in a more deeply context is necessary to correct find 
+duplicatons in the problem outlined.
+
+Since our searches for solution to the AMD Display driver problem with duplicated code became fruitless. We propose to create a tool
+that approaches the given problem and can be adapted with a moderated amount of effort to support other open source projects facing
+similar issues. Given our proposed tool, we desire to validate if it is capable of detecting duplicated code and if the tool 
+is in fact useful to mitigate the AMD Display driver problem with duplicated code.
 

### 2024-06-25 conteudo/introduction/intro.tex
Add update of intro, 24/06

@@
 
-The Linux kernel is a major project in the open source community, containing more then 19 million lines of code and more
+Representing 3.77\% of the personal desktops and 70,69\% of the smartphones worldwide, the Linux kernel 
+is a open source project providing value to billions of people in their daily lives. 
+The magnitude that the project has reached over the years makes the addition of features, as well as handling bugs and security 
+vunerabilities, potentially affect a significant number of people, demonstrating the importance of maintaining the 
+kernel.
+
+The maintainment of the Linux kernel is a complex task, as the project contains more then 19 million lines of code and have more
 then thirteen thousand contributors along the lifespan of the project \citep{linuxquantity}. Given the magnitude of the project,
@@
 contributions they made to the project. It is expected to find a good amount of bad quality code artifacts in the source code, 
-which could result in hindering or compromising the maintenance of this project, as well as the indication of new features for it.
+which could result in hindering or compromising the maintenance of this project, as well as the addition of new features for it.
 
-Device driver, as stated by Madieu, are a "piece of software whose aim is to control and manage a particular hardware device, hence the
-name device driver" \citep{driverdef}. Device drivers are major part of the Linux kernel, representing 66 \% of the source code and
+Device driver, as stated by Madieu, are a "piece of software whose aim is to control and manage a particular hardware device, hence 
+the name device driver" \citep{driverdef}. Device drivers are major part of the Linux kernel, representing 66 \% of the source code and
 responsible for the management of crucial element such as keyboards, mouses and GPU \citep{marcelo}. 
@@
 we found were primitive solutions that finds every pair of code files that are duplications. As the AMD Display driver is a driver 
-created in 20XX (TODO HERE) and maintained until today, the code files that contains duplicated code normally also have specific code that is 
-only implemented in this file. Thus, a solution that finds duplication in a more deeply context is necessary to correct find 
+created in 20XX (TODO HERE) and maintained until today, the code files that contains duplicated code normally also have specific code
+that is only implemented in this file. Thus, a solution that finds duplication in a more deeply context is necessary to correct find 
 duplicatons in the problem outlined.
 
-Since our searches for solution to the AMD Display driver problem with duplicated code became fruitless. We propose to create a tool
-that approaches the given problem and can be adapted with a moderated amount of effort to support other open source projects facing
-similar issues. Given our proposed tool, we desire to validate if it is capable of detecting duplicated code and if the tool 
-is in fact useful to mitigate the AMD Display driver problem with duplicated code.
+Since our searches for solutions to the duplicated code problem become fruitless. We
+intend to propose an approach to indentify and mitigate the code duplications in the Linux Kernel. For that, 
+we propose a tool capable to indentify code duplications in the kernel, and by participant observation analyse and mitigate the 
+code duplications found while cataloging the software engineer methods used in process such as refactoring methods.
+
+An additional usefulness of our proposed approach is to give a new method to people become contributors to the Linux kernel. The 
+process of starting to contribute to the Linux kernel can be unintuitive, which can cause people to distance themselves.
+Our approach provides to the community a new form of newcomers to start contributing by removing simple code duplications, 
+while they learn the whole process of being a Linux contribuitor.
 

### 2024-06-26 conteudo/introduction/intro.tex
Trying do something more right for intro, 25/06

@@
 \section**{Problem outline}
+\label{sec:problem}
 
@@
 the name device driver" \citep{driverdef}. Device drivers are major part of the Linux kernel, representing 66 \% of the source code and
-responsible for the management of crucial element such as keyboards, mouses and GPU \citep{marcelo}. 
-The AMD Display driver of the Linux kernel is a core driver to enable AMD GPU's to operate correctly in desktop linux environments,
-which is a important duty, given that AMD GPU represented 19 \% of the personal GPU market in 2023 \citep{gpumarket}.
+responsible for the management of crucial element such as keyboards, mouses and GPU \citep{marcelo}. The moment the kernel addopts
+to support new hardwares, it requires to make alteration of existent device drivers or the creation of new ones, 
+this makes the maintainment of the devices drivers a important duty of the Linux community.
 
+The AMD Display driver of the Linux kernel is driver responsible to enable AMD GPU's to operate correctly in desktop linux
+environments, which is a important duty, given that AMD GPU represented 19 \% of the personal GPU market in 2023 \citep{gpumarket}.
 When we iteracted with the maintainers of the AMD Display driver, they shared with us their pain with one of the characteristics
-bad quality code artifacts, that is duplicated code artifacts, which hinders the maintainment of the driver. While searching for a
-solution to them in the literature, we only found tools that given two code artifacts, infer if they are duplication of each other,
-and the research in this topic mainly focused in optimizing the accuracy of these tools in this given task. As we desired to the tools
-to find the duplicated code artifacts for us, instead of given two code artifact that we know it is a duplication, confirm to us that
-is in fact a duplication, the literature cannot be applied purely in our problem.
+bad quality code artifacts, that is duplicated code artifacts, which hinders the maintainment of the driver.
 
-As a alternative, we tried to search solutions in the gray literature, but without success. In the gray literature, the best solution
-we found were primitive solutions that finds every pair of code files that are duplications. As the AMD Display driver is a driver 
-created in 20XX (TODO HERE) and maintained until today, the code files that contains duplicated code normally also have specific code
-that is only implemented in this file. Thus, a solution that finds duplication in a more deeply context is necessary to correct find 
-duplicatons in the problem outlined.
+Searching for solutions to the AMD Display driver problem with duplicated code artifacts, we were able to find tools that given 
+two code artifacts, infer if they are duplication of each other, and the formal literature in the code duplication mainly focused 
+in optimizing the accuracy of these tools in this given task. These solutions do not approach our problem, as we are not able to give them a codebase
+and they returns to us the code duplication in the codebase, and they do not guide us to how mitigate the code duplications after
+we detected them, requiring us to search for others alternatives.
 
-Since our searches for solutions to the duplicated code problem become fruitless. We
-intend to propose an approach to indentify and mitigate the code duplications in the Linux Kernel. For that, 
-we propose a tool capable to indentify code duplications in the kernel, and by participant observation analyse and mitigate the 
-code duplications found while cataloging the software engineer methods used in process such as refactoring methods.
+Given the unsuccessful results on our serach in the formal literature, we started to search for solutions in the gray literature, 
+but we ended up without success. In the gray literature, the best solutions we found were primitive solutions that finds every pair
+of code files that are duplications. As the AMD Display driver is a driver created in 20XX (TODO HERE) and maintained until today,
+the code 
+files that contains duplicated code normally also have specific code that is only implemented in this file. Thus, 
+a solution that finds duplication in a more deeply context is necessary to correct find duplicatons in the Linux kernel context.
 
-An additional usefulness of our proposed approach is to give a new method to people become contributors to the Linux kernel. The 
-process of starting to contribute to the Linux kernel can be unintuitive, which can cause people to distance themselves.
-Our approach provides to the community a new form of newcomers to start contributing by removing simple code duplications, 
-while they learn the whole process of being a Linux contribuitor.
+\section**{Research Design}
 
-\section**{Research Questions}
+The code duplication problem outlined in Section \ref{sec:problem} cannot be resolved by the solutions presented in the formal 
+literature nor the gray literature, we intend to propose an approach to indentify and mitigate the code duplications in the 
+Linux Kernel. For that, we propose a tool capable to indentify code duplications in the kernel, and we mitigate code duplications in the kernel. 
+On this research, we expect to extract patterns of software engineers techniques, such as refactoring methods, which we commonly 
+used to mitigate the code duplications, thus, serving as guidance to future contributors to approach the code duplications problem 
+in the given context. To properly investigate our proposed approach, we developed the research questions below to guie our research.
+
+\textbf{RQ1:} How effective is our approach in mitigating duplicate code in the Linux kernel?
+
+\textbf{RQ1:} How effective is our approach in identifying duplicate code in the Linux kernel?
+
+This research is divided into two phases. On the first phase, we will propose, build and validate a tool capable to indentify code 
+duplications in function level for the C programming language, we expects this proposed tool to indentify code duplications
+in the Linux kernel, as in our preliminary investigation the code files in the kernel is commonly a union of duplicated and 
+nonduplicated functions. To validate the tool, we will do a triangulation of results by evaluate our tool in the code duplication
+databases used in the formal literature and (DO NOT KNOW IF I REALLY WILL DO THIS) doing a empirical analyses in a set
+of randomly selected codes of the AMD Display driver which our proposed tool claims that is a duplications. 
+The detailed explication of how we will validate our tool is demonstrated in the Chapter \ref{cha:method}.
+
+Given the first phase completion, we have a validated tool to indentify code duplications in the Linux kernel 
+context and we can move on to how mitigate this duplications, which compreend the second phase of our research.
+
+(I DO NOT HAVE MORE A PLACE TO FIT THIS PARAGRAPH) An additional usefulness of our proposed approach, which we do not verify in our research, is to give a new method to people become 
+contributors to the Linux kernel. The process of starting to contribute to the Linux kernel can be unintuitive, which can cause 
+people to distance themselves. Our approach provides to the community a new form of newcomers to start contributing by removing 
+simple code duplications, while they learn the whole process of being a Linux contribuitor.
 

### 2024-06-26 conteudo/introduction/intro.tex
Add initial reformuled version of the introduction

@@
 
-This research is divided into two phases. On the first phase, we will propose, build and validate a tool capable to indentify code 
+This research is divided into two phases. On the \textbf{Phase I}, we will propose, build and validate a tool capable to indentify code 
 duplications in function level for the C programming language, we expects this proposed tool to indentify code duplications
@@
 
-Given the first phase completion, we have a validated tool to indentify code duplications in the Linux kernel 
-context and we can move on to how mitigate this duplications, which compreend the second phase of our research.
+Given the first phase completion, we would have a validated tool to indentify code duplications in the Linux kernel 
+context and we can move on to how mitigate these duplications, which comprehend the second phase of our research. On the 
+\textbf{Phase 2}, we intend to discover ways to mitigate the code duplication detected in the linux kernel. 
+To achieve our objective, we designed a etchnographic approach, which enables us to interact with the linux community
+to collect artifacts validating or refuting our mitigation. The detailed explication of the etchnographic approach is demonstrated
+in the Chatper \ref{cha:method}.
 
+(POINTS TO CONSIDER: I do not like much this introduction, but is the best I could do)
 (I DO NOT HAVE MORE A PLACE TO FIT THIS PARAGRAPH) An additional usefulness of our proposed approach, which we do not verify in our research, is to give a new method to people become 

### 2024-06-26 conteudo/introduction/intro.tex
Add final work of the 26/06, only needs the initial text of intro now

@@
 To achieve our objective, we designed a etchnographic approach, which enables us to interact with the linux community
-to collect artifacts validating or refuting our mitigation. The detailed explication of the etchnographic approach is demonstrated
-in the Chatper \ref{cha:method}.
+to collect artifacts validating or refuting our mitigation. The figure (CREATE A FIGURE HERE) ilustrate the research
+process we addopted in this research.
 
 (POINTS TO CONSIDER: I do not like much this introduction, but is the best I could do)
-(I DO NOT HAVE MORE A PLACE TO FIT THIS PARAGRAPH) An additional usefulness of our proposed approach, which we do not verify in our research, is to give a new method to people become 
+(I DO NOT HAVE MORE A PLACE TO FIT THIS PARAGRAPH) An additional usefulness of our proposed approach, which we do not verify
+in our research, is to give a new method to people become 
 contributors to the Linux kernel. The process of starting to contribute to the Linux kernel can be unintuitive, which can cause 
 people to distance themselves. Our approach provides to the community a new form of newcomers to start contributing by removing 
-simple code duplications, while they learn the whole process of being a Linux contribuitor.
-
-\section**{Research Design}
+simple code duplications, while they learn the whole process of being a Linux contributor.
 
 \section**{Manuscript Structure}
+
+This manuscript consists of four more chapters. 
+Chapter \ref{cha:back} presents the literature overview for the code duplication detection (main definitions, current approches 
+methods in the literature), a briefly description of the Linux kernel and a review of refactoring methods used throughtout 
+this research.
+Chapter \ref{cha:tool} presents our proposed tool to detect code duplication in the Linux kernel, describing all of the main 
+components.
+Chapter \ref{cha:method} describes the research methods selected to guide our work.
+Chapter \ref{cha:results} presents the current research stage and the work plan, with addition of preliminary results obtained.

### 2024-06-27 conteudo/background/codeduplication.tex
Add initial research methods, 27/06

@@
 \subsection{Types of code clone duplication}
+
+\label{subsec:types}
 
@@
 
-Given the necessity to evaluate and compare approches for code clone detection, it was developed and widely-adopted two datasets: 
-OJ-Clone, a dataset collected from a pedagogical online judge system \citep{ojclone}, and BigCloneBench, a datatset collected from over 
-25,000 Java software systems \citep{bigclonebench}. The existence of only two spreaded datasets is a open concern in literature as 
-the limited scope of OJ-Clone and the problems of BigCloneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail}
-
 The state-of-art methods for code clone detection usually have high computational time and memory space usage, as well being
@@
 with the literature approches later on this document.
+
+\subsection{Methodology to evaluate code duplication methods}
+
+\label{subsec:codumethods}
+
+Given the necessity to evaluate and compare approches for code clone detection, it was developed and widely-adopted two datasets: 
+OJ-Clone, a dataset collected from a pedagogical online judge system \citep{ojclone}, and BigCloneBench, a dataset collected from over 
+25,000 Java software systems \citep{bigclonebench}. The existence of only two spreaded datasets is a open concern in literature as 
+the limited scope of OJ-Clone and the problems of BigCloneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail}
+
+(ADD INFORMATION ABOUT THE MORE SEPARATION OF CLONE TYPES THE BIGCLONEBENCH makes)

### 2024-06-27 conteudo/background/nlpsimilarity.tex
add initial tool, 13/06
Add initial research methods, 27/06

@@
 \section{Text similarity detection methods}
+\label{sec:similarity}.
 
@@
 Where $A$ and $B$ are the vector embeddings of the two texts. One of the advantages of the cosine similarity is that $SC(A,B)$ 
-is a number between $0$ and $1$, where $SC(A,B)$ values more close to the number $1$ means that two vectors are more similar.
+is a number between $0$ and $1$, where $SC(A,B)$ values more close to the number $1$ means that two vectors are more similar,
+it is common to compare the cosine similarity to aproximation to a probability value, transforming the values between $0$ and $1$ to a
+porcentage between $0 \% $ and $100\%$, through our research we will use the porcentage form.
 

### 2024-06-27 conteudo/introduction/intro.tex
Add initial research methods, 27/06

@@
 literature nor the gray literature, we intend to propose an approach to indentify and mitigate the code duplications in the 
-Linux Kernel. For that, we propose a tool capable to indentify code duplications in the kernel, and we mitigate code duplications in the kernel. 
-On this research, we expect to extract patterns of software engineers techniques, such as refactoring methods, which we commonly 
-used to mitigate the code duplications, thus, serving as guidance to future contributors to approach the code duplications problem 
-in the given context. To properly investigate our proposed approach, we developed the research questions below to guie our research.
+Linux Kernel. For that, we propose a tool capable to indentify duplicated functions in the kernel, and we intend to mitigate 
+code duplications in the kernel. On this research, we expect to extract patterns of software engineers techniques, 
+such as refactoring methods, which we commonly used to mitigate the code duplications, thus, serving as guidance to future 
+contributors to approach the code duplications problem in the given context. To properly investigate our proposed approach, 
+we developed our primarily research question below to guide our research.
 
-\textbf{RQ1:} How effective is our approach in mitigating duplicate code in the Linux kernel?
+\textbf{RQ1:} How effective is our approach in itendifying and mitigating duplicate code in the Linux kernel?
 
-\textbf{RQ1:} How effective is our approach in identifying duplicate code in the Linux kernel?
+Given the complexity of the proposed question, we created a supplementary research question, with the objective to verify midway 
+results, which is related to the performance of our proposed tool in the Linux kernel, the supplementary research question
+is defined as:
+
+\textbf{RQ1.1:} How accurate is the proposed tool for detecting duplicate functions?
 

### 2024-06-27 conteudo/methods/methods.tex
Add initial research methods, 27/06

\en

To answer our proposed research questions, we opted to divide our research. Firstly, we will approach the supplementary research 
RQ1.1, which will serve as a background to our research question RQ1. To answer the RQ1.1, we propose a multiple-method investigation,
by validating our proposed tool in the BigCloneBench databaset \citep{bigclonebench}, which is the standard method used by the
literature, and a empirical analyses of a selected subset of functions in which the tool infers.
With the question RQ1.1 answered, we will return our focus to the RQ1. To answer the RQ1, we opted to create a etchnographic study, 
using participant-observation methodology, we expect to mitigate a selected subset of duplicated functions by refactoring them using
a designed sistematic approach, while iteracting with the Linux community to collect feedback of the mitigation made by us.

\section{Multiple-method Investigation}

The multiple-method investigation to verify the efficiency of our tool in a methodical way is a method in which we consider 
trustworthy.
The first method used is the evaluation of the proposed tool in the BigCloneBench database \citep{bigclonebench}, one of the 
standard code duplication databases used by validation of tool presented in the formal literature. 
As presented by Krinke and Ragkhitwetsagul, the BigCloneBench is not a perfect dataset \citep{bigfail}, and for that reason we opted to do
a additional evalution, in which we will do a empirical analyses on a select subset of functions of the AMD Display driver in which
our tool infers it is a duplication. To a concrete evaluatiion, in both of the methods we will apply a  variation of the similarity 
metric (The similarity metric is a metric measuring the similarity of two code duplications, a proper explanation is made 
in the Chapter \ref{cha:tool}).

\subsection{Evaluation by the formal literature}

The BigCloneBench dataset \citep{bigclonebench} is a collection of Java code with a selection of code duplications pairs categorized by the authors, these
code duplications pairs are not separed solely if they are duplications, the pairs are also categorized by the code duplication
types presented in Subsection \ref{subsec:types}.

Liut et al presented a sample method to evaluate the their code duplications detection method, in which they sample 20,000 pairs
from each of the code duplication types, with the addition of 20,000 nonduplication pairs to act as the negative samples
\citep{tailor}. To a fairness evaluation, we use the same sample method as them to evaluate our tool.

The proposed tool does not differentiate the code duplication types and infers duplications in the function level instead of the 
code file level, which is the standard in the state-of-art code duplication methods, we were required to adjust the form 
we compute metrics of the evalution. To accomplish this, we will define that every pair of functions in which the similarity metric
is equal or superior to a threshold X will be considered a duplication, and the pair of code files that the functions are contained
are marked as duplication. If on this method, a code duplication pair is corretly indentified, it will count as a correct to its 
code duplication types, if a incorrect duplication pair is infered, it will count as incorrect to all code duplication types. Given
these definitions, we will compute the Recall metric presented in \ref{subsec:codemethods} for each of the code duplication types.

To understand the variation of meaning when we change the similarity threshold, we will eval our tool for the given variations of
the threshold X and discuss the results: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, 100\%.

\subsection{Evaluation by Empirical analyses}

\section{Ethcnographic study}

### 2024-06-28 conteudo/methods/methods.tex
add initial ethnography

@@
 literature, and a empirical analyses of a selected subset of functions in which the tool infers.
-With the question RQ1.1 answered, we will return our focus to the RQ1. To answer the RQ1, we opted to create a etchnographic study, 
+With the question RQ1.1 answered, we will return our focus to the RQ1. To answer the RQ1, we opted to create a ethnographic study, 
 using participant-observation methodology, we expect to mitigate a selected subset of duplicated functions by refactoring them using
@@
 
-\section{Ethcnographic study}
+Given the problems of the BigCLoneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail} and the limitations of it in our
+scope such as being a Java dataset while the Linux kernel are primarily build in the C programming language, we opted to, together
+with the formal litarature evaluation, perform a empirical study analyzing the codes in which our tool infered as duplications.
+
+The number of code lines and components in the Linux kernel makes it impractical to analyse every code duplication. To limit our
+scope, we will analyse solely code duplications in the AMD Display driver. This limitation is still does not reduce the code
+duplication to analyse, since in our preliminar tool execution showed in the Figure \ref{fig:relatory_ex}, there is over 20,000
+lines of code duplications on the AMD Display driver with similarity metric of 100\%.
+
+For conducting our empiricial analyses, we opted to do a random selection in the code duplications pairs infered by our tool. To
+do that, we will use the definition that if the similarity metric of a function pair is greater or equal to X, will be considered a
+duplication. To understand the optimal values of X we will analyse 10 functions pair randomic selected 
+that have similarity metric X for each of the variations of the parameter X, which are 
+30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, 100\%. 
+As it is not expected to have code duplications with similarity exactly a whole number, we will allow a
+variation of 1\% for greater or less to each of the variations of the parameter X.
+
+To do the random selection of functions pairs, we list the code duplications for each of the variations of X and execute 
+a random selection algorithm
+builtin in in the source code language of our tool. The exactly algorithm of the random selection, along with the selected 
+function pairs can be found in Appendix (CREATE APENDICE AND ADD HERE).
+
+When we analyses the selected functions pairs, we will judged empiricaly if they are a duplication or not, then we are not
+going further as
+categorizing the duplications by code duplication type. We intend to collect the data on this empirical analyses and compare it
+to the formal literature method, to undertand if our tool in fact detect duplicated functions and 
+the difference it makes when varying the minimum similarity parameter of our tool.
+
+\section{Ethnographic study}
+
+Ethnography is a research method recognized as a significant qualitative empirical approach 
+suited to understanding people and cultures, and their associated social and work practices \citep{bookethno}. 
+The central tenet of the ethnography approach is to enable research to undertand values, beliefs or ideas shared for a 
+ community under study from the members' point of view \citep{ethnosoft}. Thus, the ethnographer is required to act as
+a member of  the community, learning and observing deeply what people do and why they do.
+
+Sharp et al discuss the application of the ethnography method in the software engineering research, demonstrating the low adoption 
+of the method and its advantages to uncover what practioners do and why they do \citep{ethnosoft}. In our research context, if we
+mitigate code duplications by ourselves without iteracting with the Linux community, we cannot undertand if the linux maintainers 
+would accept our approach for code duplication and what concerns they would have. Thus, we resorted to a ethnography research, where
+we will act as a member of the Linux kernel, begin a contributor with focus on removing code duplications found.
+
+On this section, we will describe a generic approach pipeline to mitigate code duplication found in the Linux kernel. Then we
+will describe the methodology to select a subset of duplicated functions found by our tool and finally, describe the ethnographic
+methodology to interact with the Linux community.
+
+\subsection{Generic approach pipeline to mitigate code duplication}
+
+The C programming language's way to communicate between source code files is through creation of header files, which is
+a library specification \citep{Cbook}. As the Linux kernel is mainly build in C, we resorted to remove code duplications by
+creating a library, where the code artifacts, instead of using their duplications, will pass to use the library.
+
+When removing duplicated functions, our desired behavior is to this function to every portion of the code it exist, only allowed a
+unique version existent in the library created by us. For this reason, given our initial duplicated function pair, we will use 
+the function information functionality(This functionality is described in \ref{subsec:functioncommand}) of our tool to find all 
+places that this function is defined in the source code. Thus, we finish with a collection of duplicated function and the collection
+of code files in which these functions are defined into.
+
+Given the semantics of a library to containing the shared code by a collection of code files, we added a additional step of the 
+approach, which is searching for others functions contained in all of code files of the collection we are working with. 
+On this moment, we have a collection of code files, and a non-empty set of functions which are shared by all of the code files 
+in the collection.
+
+Finally, for each of the functions shared by all code files, we do a specific-case aproach to mitigate each one of them.
+If a function is shared perfectly by all the code files, we use a simply approach to only remove them and create the function in
+the source code file, if it is not shared perfectly, we look for refactoring methods which can be applied on this specific case.
+
+Finishing our mitigation method, if the mitigation ended up working in a initial step, we move it foward to the 
+Ethnographic method, iteracting with the Linux community to understand what are their thoughts on the mitigation. If the mitigation
+ended up as not working as expected, we will categorized it in two categories, the first one is that the refactoring requires a
+decision in which we, as not specialist and non-maintainers of the Linux kernel, cannot make it correctly, although
+we believe a specialist would be able to refactoring it. The second categorization is that we do not know if it is in fact possible
+to refactoring the duplication.
+
+The figure (CREATE FIGURE AND ADD HERE) ilustrates the approach pipeline described in this section.
+
+\subsection{Selection of the duplicated functions}
+
+\subsection{Ethnographic methodology}

### 2024-06-28 conteudo/methods/methods.tex
add random selection of ethnografy

@@
 
+\label{subsec:pipeline}
+
 The C programming language's way to communicate between source code files is through creation of header files, which is
@@
 
+As discussed in the empyrical analyses for the tool, it is impractical to evaluate every code duplication on Linux kernel. Thus,
+similarily to the empyrical analyses for the tool, we will limit the experiment scope to only the AMD Display driver and select
+a random sample of duplicated functions infered by our tool to initiate the mitigation approach defined in \ref{subsec:pipeline}. 
+We chose a sample size of 30 duplicated functions, as there are multiples studies demonstrated that a sample size of this size 
+have a statistical significance \citep{sample1,sample2}.
+
+On the the random selection of the duplicated functions, we deside to put a additional restriction on which functions pairs are 
+elegible to be selected. The restriction is that we will only select functions pairs with similarity metric equal 
+to 100\%, we made this choice because, as showed in the Figure \ref{fig:relatory_ex}, there are more then 20,000 lines codes 
+in the AMD Display driver in which the similarity metric reaches this value. On all function pairs that respect the restriction,
+we will choose then uniformly at random.
+
+The functions pairs selected at random following our methodoly can be consulted in the Table (CREATE TABLE AND ADD HERE).
+
+CREATE TABLE AND ADD HERE.
+
 \subsection{Ethnographic methodology}

### 2024-06-28 conteudo/methods/methods.tex
Add initial methods with all texts

@@
 
-Finishing our mitigation method, if the mitigation ended up working in a initial step, we move it foward to the 
+Finishing our mitigation method, if the mitigation ended up working in a initial step, we may move it foward to the 
 Ethnographic method, iteracting with the Linux community to understand what are their thoughts on the mitigation. If the mitigation
@@
 we believe a specialist would be able to refactoring it. The second categorization is that we do not know if it is in fact possible
-to refactoring the duplication.
+to refactoring the duplication. With or without a sucessful mitigation, we will store our mitigations on (ADD REP AND PUT HERE) 
+to research tranparency.
 
@@
 \subsection{Ethnographic methodology}
+
+When we make a code duplication mitigation following our approach pipeline described in \citep{subsec:pipeline}, we are apt
+to initiate our ethnographic research. We resorted to a participant-observation approach as our ethnographic methodology, where we 
+will act as contributors of the Linux kernel to the AMD Display driver, interacting with the maintainers of the AMD Display driver
+attempting to get our mitigations added to the Linux kernel.
+
+The Linux kernel communication is made through public mailing lists, where anyone can consult the discussions made in it.
+We plan to collect data artifacts through the AMD Display diver mainling list.
+The mainling list will register our iteractions with Linux community, with 
+their members sharing their opinions and suggestions to the mitigation we made, corroborating to the discussion about the 
+effectiveness of our proposed approach.
+
+Given the time limit of our research, we does not expect to be able to iteract with the Linux community throughtout successfully
+mitigation we made using the proposed approach, as it pass through factors we do not control, which is the time spent waiting 
+for Linux community's members interact with us.
+For this reason, we resorted to commit to do the participant-observation approach 
+in 5 of the successfully mitigation and use them for our discussion.

### 2024-07-11 conteudo/background/codeduplication.tex
add initial ethnography
Small revision of the text

@@
 programming language specific approches, which cannot be easily implemented in programming languagues that were not used in 
-their research \citep{litreview}. These are main concern for our research and for this reason we will use a generalist textual-based 
+their research \citep{litreview}. These are concerns for our research, for this reason we will use a generalist textual-based 
 approach based solely on Natural Languague Processing (NLP) text similarity detection methods, which we will introduce and compare
@@
 
-\label{subsec:codumethods}
+\label{subsec:codemethods}
 

### 2024-07-11 conteudo/background/nlpsimilarity.tex
Small revision of the text

@@
 There are multiple methods for building vector embeddings, such as bag-of-words, TF-IDF, LSI, Word2Vec \citep{gensimlivro}. 
-The recent works related to build vector embeddings uses Largue Language Models (LLM), for example the usage of BERT, ELMo 
+The recent works related to build vector embeddings uses Large Language Models (LLM), for example the usage of BERT, ELMo 
 and GPT \citep{llmsimilar}. The most popular distance function used by the literature are the cosine similarity 
@@
 
-It is a famous fact that LLM have high computational cost, one example of such is a experiment done by Reimers and Gurevych for 
+It is a famous fact that LLM have high computational cost, one example of such is a analyse done by Reimers and Gurevych for 
 finding the most similar pair in a collection of 10,000 sentences, that required about 65 hours with BERT \citep{bertsimilar}. This 
-is a similar magnitude of comparation pairs we want to do through our work and for this reason we chose to not use this kind of 
+is a similar magnitude of comparation pairs we want to do through our work, and for this reason we chose to not use this kind of 
 vector embedding method, but it is a valid solution for smaller codebases.

### 2024-07-11 conteudo/tool/architecture.tex
Small revision of the text

--- 2024-06-15 conteudo/tool/architecture.tex
@@
 
-The Setup step is composed of two components, the Function Breaker and the Code Duplication Finder. the pipeline of this step work as
-follow: the Function Parser receives the codebase we are interested, extract the functions of the codebase along with metadata and
-creates a new temporary codebase where it function becomes a new code file. The Code Duplication Finder iterate over every pair of 
-files in the temporary codebase computing the code duplications and save them in the Code Duplication Database, which is a text file
-that stores every code duplication as a triple <function1,function2,similarity>, where function1 and function2 are the functions that 
-are a duplication of each other, and similarity is the metric given by the code duplication method utilized, which in our case, it is
-the cossine similiratiy explained in \ref{sec:similarity}. Finally, the Query Responder consumes the temporary codebase and the 
-Code Duplication Database to extract code duplication related information as per user request. A diagram to ilustrate the pipeline
-can be found in the Figure (ADD FIGURE HERE). There is a explanation of how each component work below.
+The Setup step is composed of two components, the Function Breaker and the Code Duplication Finder. The pipeline of this step work as
+follow: the Function Parser receives the codebase the user are interested, extract the functions of the codebase along with metadata 
+and
+creates a new temporary codebase where the functions extracted becomes news code files. 
+The Code Duplication Finder iterate over every pair of 
+files in the temporary codebase checking if they are a code clone and if so, save them in the Code Duplication Database, 
+which is a text file
+that stores every code duplication as a triple \textbf{<function1,function2,similarity>}
+, where function1 and function2 are the functions that 
+are a duplication of each other, and similarity is the metric given by the code duplication detection method utilized, which in our case, 
+it is
+the cossine similarity explained in \ref{sec:similarity}. Finally, the Query Responder consumes the temporary codebase and the 
+Code Duplication Database to extract duplicated functions related information as per user request. A diagram to ilustrate the pipeline
+can be found in the Figure \ref{fig:diagram}. There is a explanation of how each component work through this section.
+
+\begin{figure}
+\includegraphics{diagrama_mestrado}
+\caption{The architecture diagram demonstrating the relationship between the components.}
+\label{fig:diagram}
+\end{figure}
 
@@
 The input codebase is expected to be a folder in the machine the tool is running into. All files in the codebase that is not a source
-code file from a programming language supported by the tool will be ignored.
+code file from a programming language supported by the tool will be ignored. The tool validates that a source file is supported through
+analyses of the file extension.
 
@@
 As a alternative approach, we can build the syntax tree \citep{compiler} of the programming language and use it to 
-extract the functions, which is a more complex task then our main approach. For the programming languages we support, we implemented
-it as follow:
+extract the functions, which is a more complex task then our main approach. For the programming languages that e support, 
+we implemented them as follow:
 
@@
 	\begin{item}
-		\textbf{C}:
+		\textbf{C}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in
+		the code context and some of code lines before the definition of the code block. By the C programming language's grammar, 
+		the functions are defined in a global context, with depth of 0. There are other elements with the same depth as functions,
+		we differentiate them by analysing if the first non-empty character before the curly brackets of the code block is a close 
+		parenthesis, since functions are the only elements that respect this pattern in the C programming language.
 	\end{item}
 	\begin{item}
-		\textbf{Java}:
+		\textbf{Java}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in
+		the code context and some of code lines before the definition of the code block. By the Java's grammar, 
+		the functions are defined inside classes. There are different depths which we can have functions, as Java accepts inner
+		classes. We differentiate functions of other elements by analysing if the first non-empty character before the curly brackets of the code block is a close 
+		parenthesis. There is a problem with this approach as respect the pattern we defined for functions such as for loop, 
+		and for this reason, it is required to apply other filters.
+		We chose to only analyse functions of depth 1, which ignores every others elements that respect are code blocks and have
+		a parentesis as the first non-empy character before the curly brackets. This choice has a counterpart, 
+		functions declared inside Inner Classes are not supported by our extractor. We made this choice
+		as there are not inner classes on the BigCloneBench \citep{bigclonebench}, which is the only codebase in Java explored in this work.
+
 	\end{item}
@@
 compilable code, which is expected to happen in the temporary codebase as it not stores complete code artifacts. Platis implemented
-this code cuplication detection method and distributed it to the public with a MIT license \citep{platistool} \citep{mitlicense}, 
+this code cuplication detection method and distributed it to the public with a MIT license \citep{platistool,mitlicense} ,
 we utilize his implementation as it has some nice to have components such as removing code comments, we do the appropriated changes
@@
 a codebase of $10000$ functions and supposing each relative path plus the function name have $50$ charactes, the Code Duplication 
-Database will be a file of size approximately $5000^2 \times 50 ~= 5$ Gigabytes, as it is not expected to a real large codebase 
+Database will be a file of size approximately $5000^2 \times 50 ~= 5$ Gigabytes, as it is not expected in a real large codebase 
 to the majority of pairs of functions be a duplication of each other, this number can be reduced dramastically.

### 2024-07-11 conteudo/tool/functionalities.tex
Add base figures know in a acceptable manner, 20/06
add initial ethnography
Small revision of the text

@@
 
-\section{Tool functionalites}
+\section{Tool functionalities}
 \label{subsec:func}
@@
 We proposes three main functionalities for the user in the tool, which are named duplication explorer, function information
-and duplication relatory. Each of the functionalities recevies their adequated parameters to execute, with a exception
+and duplication relatory. Each of the functionalities receives their adequated parameters to execute, with a exception
 of a optional parameter shared by all of them, which is minimum similarity per query. The minimum similarity per query is different
 from the siminum similarity presented in \ref{subsec:setup}, as it enables a volatile change in the minimum similarity
-interested in a query specific context. There is a explanation of each of the main functionalities throughout this section.
+interested in a query specific context. There is a explanation of each of the main functionalities through this section.
 
@@
 
-The duplication explorer is the principal functionality of our tool, the functionality purpose is to expose for the user the code duplication
-pairs found by our tool. We implement optional filters methods to enable more complex queries, which are described below. A example
-of the functionality usage can be found in the figure \ref{fig:explorer_ex}.
+The duplication explorer is the principal functionality of our tool, the functionality purpose is to expose for the user the 
+duplicated functions pairs found by our tool. We implement optional filters methods to enable more complex queries, 
+which are described below. A example of the functionality usage can be found in the figure \ref{fig:explorer_ex}.
 
@@
 		given by the user as a parameter. We say that a function matches a pattern if the string formed by the concatenation of
-		the relative path of the code file plus the function name. The user can define if both functions or at least one of the 
+		the relative path of the code file plus the function name contains the pattern as a substring. 
+		The user can define if both functions or at least one of the 
 		functions in the pair need to match the pattern by given a specific parameter.
@@
 \begin{figure}
-\includegraphics{explorer_example}
+\includegraphics[scale=0.5]{explorer_example}
 \caption{Explorer Example}
@@
 
-The Function information purpose it to give a more in depth information about a specific function. The functionality inform the user
-about the relative path, the function name and the lines the function exist in the code file that the function is contained for the
-function that the user is interested and every function that is a duplication of it. As there ca nbe multiple functions with the 
-same name in a codebase, we distinguish then by the user providing a pattern string and we use the first function find that match
-the pattern. We say that a function matches a pattern if the string formed by the concatenation of the relative path of the 
-code file plus the function name. 
+\label{subsec:functioncommand}
+
+The Function information purpose it to give a more in depth information about a specific function. The functionality 
+receives a function the user is interested in and then inform the user
+about the relative path, the function name and the lines the function is defined for the function the user is interested and for
+every function that is a duplication of the given function.
+There can be multiple functions with the 
+same name in a codebase, to distinguish them, we receive a pattern string from the user, and  we use the first function find that match
+the given pattern. We say that a function matches a pattern if the string formed by the concatenation of the relative path of the 
+code file plus the function name contains the pattern as a substring. 
 A example of the functionality usage can be found in the figure \ref{fig:function_ex}.
@@
 \begin{figure}
-\includegraphics{function_example}
+\includegraphics[scale=0.5]{function_example}
 \caption{Function Example}
@@
 of how much lines of duplicated code is found per folder in the codebase and present it to the user in a reasonable format. To extract
-the information, we use the Code Duplication Database to list the code duplications, and the temporary code to know how much lines of
+the information, we use the Code Duplication Database to list the duplicated functions, and the temporary code to know how much lines of
 duplicated code contains in each of the functions. (DO I TALK ABOUT THE TRIE USAGE TO MAINTAIN THE HIERARCHICAL TREE STRUCTURE? I 
@@
 \begin{figure}
-\includegraphics{relatory_example}
+\includegraphics[scale=0.5]{relatory_example}
 \caption{Function Example}

### 2024-07-11 conteudo/tool/tool.tex
Small revision of the text

@@
 Our proposed tool is a Command Line Interface (CLI) software, the tool purpose is to enable developers to query 
-information about code clone duplication in a function level, by function level we mean to find functions in a codebase that 
-are code duplications. The tool is separated in two steps. The 
+information about code clone duplication in a function level, which means to find pairs of functions in a codebase
+that where the functions in the pair are code clones of each other. The tool is separated in two steps. The 
 first one, denoted as Setup, is executed only once for each codebase and it is responsible for processing the

### 2024-07-11 conteudo/background/linux.tex
Add more info required
do some things missed

@@
 made to the project are formated as patches, which are text documents describing differences between two different 
-versions of a source tree \citep{melissa}.
+versions of a source tree  The patches are sent and reviewed through the Linux mailing lists \citep{melissa}.
 
-AMD Display Driver is a linux's subsystem responsible for the management of AMD GPU's. This subsystem have a problem 
-with code clone duplication, which are acknowledged by the subsystem's maintainters. For this reason, we selected this
-subsystem as the primary focus for refactoring the detected code clone duplication by our tool.
-(MAYBE CHANGE TO: For this reason, we selected this subsystem as the primary focus for our empirical analysis in our research).
+Mailing lists are the official means of communication of the Linux community, through the mailing lists the community members 
+send and review
+contributions to the kernel, discuss topics related to the Linux such as which new features to add, et cetera. 
+As there are multiples
+subsystems in the Linux kernel, there are multiples mailing lists in the Linux to organize the discussion made. Every subsystem 
+have their own mailing list to concentrate the discussion related to the subsystem \citep{melissa}. The iteractions made in the
+mailing lists are documented and stored in the Linux kernel Lore \citep{linuxlore}, which serves as a repository of the
+discussion made in the Linux kernel to be consulted by anyone interested.
+
+One the the Linux kernel subsystem is the AMD Display Driver, a subsystem responsible for implement the drivers required to enable
+AMD GPU's work properly in the Linux environment. The maintainers of this subsystem reported a empirical problem they have in
+the subsystem, which is a significant amount of duplicated code in the subsystem, a problem that hinders the maintainment of the
+subsystem. This reported problem is the starting point that initiated our research, making the AMD Display Driver our principal
+objecte of study to undertand how to mitigate duplicated code in the Linux kernel.
+
+(I WANT TO CHANGE THE AMD DISPLAY paragraph, WHY DO I NEED TO CHANGE IT? I SHOULD ANOTATE WHAT I WANTED, I MADE A CHANGE
+ TO MAKE IT BETTER BUT NOT SURE IF THERE IS SOME INFORMATION I WANTED TO PUT IN IT)

### 2024-07-11 conteudo/methods/methods.tex
Add a comment to approach later
do some things missed

@@
 
-On this section, we will describe a generic approach pipeline to mitigate code duplication found in the Linux kernel. Then we
+On this section, we will describe a generic systematic approach to mitigate code duplication found in the Linux kernel. Then we
 will describe the methodology to select a subset of duplicated functions found by our tool and finally, describe the ethnographic
@@
 
-\subsection{Generic approach pipeline to mitigate code duplication}
+\subsection{Generic systematic approach to mitigate code duplication}
 
@@
 
-The figure (CREATE FIGURE AND ADD HERE) ilustrates the approach pipeline described in this section.
+The figure (CREATE FIGURE AND ADD HERE) ilustrates the systematic approach described in this section.
 
@@
 
+(MAYBE I SHOULD ADD A RESTRICTION ABOUT IGNORING AUTOMATED GENERATED CODE, such as DML folder)
+
 The functions pairs selected at random following our methodoly can be consulted in the Table (CREATE TABLE AND ADD HERE).
@@
 
-When we make a code duplication mitigation following our approach pipeline described in \citep{subsec:pipeline}, we are apt
+When we make a code duplication mitigation following our systematic approach described in \ref{subsec:pipeline}, we are apt
 to initiate our ethnographic research. We resorted to a participant-observation approach as our ethnographic methodology, where we 

### 2024-07-16 conteudo/introduction/intro.tex
Fix grammar errors by grammarly free tier

--- 2024-06-26 conteudo/introduction/intro.tex
@@
 
-Representing 3.77\% of the personal desktops and 70,69\% of the smartphones worldwide, the Linux kernel 
-is a open source project providing value to billions of people in their daily lives. 
-The magnitude that the project has reached over the years makes the addition of features, as well as handling bugs and security 
-vunerabilities, potentially affect a significant number of people, demonstrating the importance of maintaining the 
-kernel.
+Representing 3.77\% of personal desktops and 70,69\% of smartphones worldwide, the Linux kernel is an open-source 
+project providing value to billions of people daily. The magnitude that the project has reached over the years makes 
+the addition of features, as well as handling bugs and security vulnerabilities, potentially affect a significant 
+number of people, demonstrating the importance of maintaining the kernel.
 
-The maintainment of the Linux kernel is a complex task, as the project contains more then 19 million lines of code and have more
-then thirteen thousand contributors along the lifespan of the project \citep{linuxquantity}. Given the magnitude of the project,
-it is unrealistic to expect that all contributors followed the best programming practices and principles throughout every 
-contributions they made to the project. It is expected to find a good amount of bad quality code artifacts in the source code, 
-which could result in hindering or compromising the maintenance of this project, as well as the addition of new features for it.
+Maintaining the Linux kernel is a complex task, as the project contains more than 19 million 
+lines of code and has more than thirteen thousand contributors along the project's lifespan
+\citep{linuxquantity}.
+Given the magnitude of the project, it is unrealistic to expect that all contributors 
+followed the best programming practices and principles throughout every contribution they made to the project.
+It is expected to find a good amount of bad-quality code artifacts in the source code, 
+which could result in hindering or compromising the maintenance of this project, as well as the 
+addition of new features for it.
 
-Device driver, as stated by Madieu, are a "piece of software whose aim is to control and manage a particular hardware device, hence 
-the name device driver" \citep{driverdef}. Device drivers are major part of the Linux kernel, representing 66 \% of the source code and
-responsible for the management of crucial element such as keyboards, mouses and GPU \citep{marcelo}. The moment the kernel addopts
-to support new hardwares, it requires to make alteration of existent device drivers or the creation of new ones, 
-this makes the maintainment of the devices drivers a important duty of the Linux community.
+Device drivers, as stated by Madieu, are a "piece of software whose aim is to control and manage
+a particular hardware device, hence the name device driver"
+\citep{driverdef}. 
+Device drivers are a major part of the Linux kernel, representing 66 \% of the source code, and 
+are responsible for the management of crucial elements such as keyboards, mice, and GPU
+\citep{marcelo}. 
+The moment the kernel chooses to support new hardware, it is required to make alterations to existent 
+device drivers or the creation of new ones, this makes the maintainment of the device drivers 
+an important duty of the Linux community.
 
-The AMD Display driver of the Linux kernel is driver responsible to enable AMD GPU's to operate correctly in desktop linux
-environments, which is a important duty, given that AMD GPU represented 19 \% of the personal GPU market in 2023 \citep{gpumarket}.
-When we iteracted with the maintainers of the AMD Display driver, they shared with us their pain with one of the characteristics
-bad quality code artifacts, that is duplicated code artifacts, which hinders the maintainment of the driver.
+The AMD Display driver of the Linux kernel is the driver responsible for enabling AMD GPUs to operate 
+correctly in desktop Linux environments, which is an important duty, given that AMD GPU represented 19 \% 
+of the personal GPU market in 2023  \citep{gpumarket} . 
+When we interacted with the maintainers of the AMD Display driver, they shared with us their pain with one 
+of the characteristics of bad quality code artifacts, which are duplicated code artifacts, which hinder 
+the maintainment of the driver.
 
-Searching for solutions to the AMD Display driver problem with duplicated code artifacts, we were able to find tools that given 
-two code artifacts, infer if they are duplication of each other, and the formal literature in the code duplication mainly focused 
-in optimizing the accuracy of these tools in this given task. These solutions do not approach our problem, as we are not able to give them a codebase
-and they returns to us the code duplication in the codebase, and they do not guide us to how mitigate the code duplications after
-we detected them, requiring us to search for others alternatives.
+Searching for solutions to the AMD Display driver problem with duplicated code artifacts, we were 
+able to find tools that gave two code artifacts, infer if they are duplications of each other, and the 
+formal literature in the code duplication mainly focused on optimizing the accuracy of these tools in this 
+given task. These solutions do not address our problem, as we are not able to give them a codebase and 
+they return to us the code duplication in the codebase, and they do not guide us to how to mitigate the code 
+duplications after we detected them, requiring us to search for others alternatives.
 
-Given the unsuccessful results on our serach in the formal literature, we started to search for solutions in the gray literature, 
-but we ended up without success. In the gray literature, the best solutions we found were primitive solutions that finds every pair
-of code files that are duplications. As the AMD Display driver is a driver created in 20XX (TODO HERE) and maintained until today,
-the code 
-files that contains duplicated code normally also have specific code that is only implemented in this file. Thus, 
-a solution that finds duplication in a more deeply context is necessary to correct find duplicatons in the Linux kernel context.
+Given the unsuccessful results of our search in the formal literature, we started to search for solutions in the 
+gray literature, but we ended up without success. In the gray literature, the best solutions we found were 
+primitive solutions that find every pair of code files that are duplications. 
+The AMD Display driver was created in 20XX (TODO HERE) and maintained until today. The code files that 
+contain duplicated code normally also have specific code that is only implemented in this file. 
+Thus, a solution that finds duplication in a deeper context is necessary to find correct duplications 
+in the Linux kernel context.
 
@@
 
-The code duplication problem outlined in Section \ref{sec:problem} cannot be resolved by the solutions presented in the formal 
-literature nor the gray literature, we intend to propose an approach to indentify and mitigate the code duplications in the 
-Linux Kernel. For that, we propose a tool capable to indentify code duplications in the kernel, and we mitigate code duplications in the kernel. 
-On this research, we expect to extract patterns of software engineers techniques, such as refactoring methods, which we commonly 
-used to mitigate the code duplications, thus, serving as guidance to future contributors to approach the code duplications problem 
-in the given context. To properly investigate our proposed approach, we developed the research questions below to guie our research.
+The code duplication problem outlined in the Section above cannot be resolved by the solutions 
+presented in the formal literature or the gray literature, we intend to propose an approach to 
+identify and mitigate the code duplications in the Linux Kernel. For that, we propose a tool capable 
+of identifying duplicated functions in the kernel, and we intend to mitigate code duplications in the 
+Linux Kernel.
 
-\textbf{RQ1:} How effective is our approach in mitigating duplicate code in the Linux kernel?
+In this research, we expect to extract patterns of software engineers' techniques, such as refactoring methods, 
+which we commonly use to mitigate code duplications, thus, serving as guidance to future contributors to 
+approach the code duplication problem in the given context. To properly investigate our proposed approach,
+we developed our primary research question below to guide our research.
 
-\textbf{RQ1:} How effective is our approach in identifying duplicate code in the Linux kernel?
+\textbf{RQ1:} How effective is our approach in identifying and mitigating duplicate code in the Linux kernel?
 
-This research is divided into two phases. On the \textbf{Phase I}, we will propose, build and validate a tool capable to indentify code 
-duplications in function level for the C programming language, we expects this proposed tool to indentify code duplications
-in the Linux kernel, as in our preliminary investigation the code files in the kernel is commonly a union of duplicated and 
-nonduplicated functions. To validate the tool, we will do a triangulation of results by evaluate our tool in the code duplication
-databases used in the formal literature and (DO NOT KNOW IF I REALLY WILL DO THIS) doing a empirical analyses in a set
-of randomly selected codes of the AMD Display driver which our proposed tool claims that is a duplications. 
-The detailed explication of how we will validate our tool is demonstrated in the Chapter \ref{cha:method}.
+Given the complexity of the proposed question, we created a supplementary research question to verify midway 
+results, which is related to the performance of our proposed tool in the Linux kernel, the supplementary research question
+is defined as:
 
-Given the first phase completion, we would have a validated tool to indentify code duplications in the Linux kernel 
-context and we can move on to how mitigate these duplications, which comprehend the second phase of our research. On the 
-\textbf{Phase 2}, we intend to discover ways to mitigate the code duplication detected in the linux kernel. 
-To achieve our objective, we designed a etchnographic approach, which enables us to interact with the linux community
-to collect artifacts validating or refuting our mitigation. The detailed explication of the etchnographic approach is demonstrated
-in the Chatper \ref{cha:method}.
+\textbf{RQ1.1:} How accurate is the proposed tool for detecting duplicate functions?
+
+This research is divided into two phases. In \textbf{Phase I}, we will propose, build, and validate a tool capable 
+of identifying code duplications at the function level for the C programming language. We expect this proposed tool 
+to identify code duplications in the Linux kernel, as in our preliminary investigation of the code files in the kernel 
+is commonly a union of duplicated and nonduplicated functions. To validate the tool, we will do a triangulation of results 
+by evaluating our tool in the code duplication databases used in the formal literature and doing an empirical analysis 
+in a set of randomly selected codes of the AMD Display driver which our proposed tool claims is a duplication. 
+A detailed explication of how we will validate our tool is demonstrated in Chapter \ref{cha:method}.
+
+Given the first phase completion, we would have a validated tool to identify code duplications in the Linux kernel 
+context and, we can move on to how to mitigate these duplications, which comprehend the second phase of our research.
+In \textbf{Phase 2}, we intend to discover ways to mitigate the code duplication detected in the Linux kernel. 
+To achieve our objective, we designed an ethnographic approach, which enables us to interact with the Linux community 
+to collect artifacts validating or refuting our mitigation.
 
 (POINTS TO CONSIDER: I do not like much this introduction, but is the best I could do)
-(I DO NOT HAVE MORE A PLACE TO FIT THIS PARAGRAPH) An additional usefulness of our proposed approach, which we do not verify in our research, is to give a new method to people become 
+(I DO NOT HAVE MORE A PLACE TO FIT THIS PARAGRAPH) An additional usefulness of our proposed approach, which we do not verify
+in our research, is to give a new method to people become 
 contributors to the Linux kernel. The process of starting to contribute to the Linux kernel can be unintuitive, which can cause 
 people to distance themselves. Our approach provides to the community a new form of newcomers to start contributing by removing 
-simple code duplications, while they learn the whole process of being a Linux contribuitor.
-
-\section**{Research Design}
+simple code duplications, while they learn the whole process of being a Linux contributor.
 
 \section**{Manuscript Structure}
+
+This manuscript consists of four more chapters.
+Chapter \ref{cha:back} presents the literature overview for code duplication detection (main definitions, 
+current approaches in the literature), a brief description of the Linux kernel, and a review of refactoring methods 
+used throughout this research.
+Chapter \ref{cha:tool} presents our proposed tool to detect code duplication in the Linux kernel, describing all 
+the main components.
+Chapter \ref{cha:method} describes the research methods selected to guide our work.
+Chapter \ref{cha:results} presents the current research stage and the work plan, with the addition of 
+preliminary results obtained.

### 2024-07-16 conteudo/background/linux.tex
Fix grammar by grammarly free tier

--- 2024-06-11 conteudo/background/linux.tex
@@
 
-The Linux kernel is an Unix-like operating system published by Linus Torvalds in 1992 as a free software and
-it started becoming popular in late 1990s \citep{linuxbook}. Against popular belief, linux is not a full 
-operating system as it do not includes applications such as filesystem utilities, windowing systems. The "Linux" 
-distributions know to the public are in fact GNU/Linux operating systems \citep{gnuref}.
+The Linux kernel is a Unix-like operating system published by Linus Torvalds in 1992 as a free software, and
+it started becoming popular in the late 1990s \citep{linuxbook}. Against popular belief, Linux is not a full 
+operating system as it does not include applications such as filesystem utilities or windowing systems. The "Linux" 
+distributions known to the public are, in fact, GNU/Linux operating systems \citep{gnuref}.
 
-TEXT ABOUT LINUX NUMBER SCOPE ETC.
+The Linux kernel is monolithic and organized as a set of subsystems such as process scheduler, memory management, 
+device driver infrastructure, networking, and filesystems \citep{melissa}.
+Each subsystem usually has a maintainer or a group of maintainers.  Maintainers are people responsible 
+for managing and accepting contributions to the subsystem they care for \citep{melissa}.
 
-The Linux project is organized as a set of subsystem such as process scheduler, memory management, 
-device driver infrastruture, networking and filesystems \citep{melissa}.
-Each subsystem usually have a maintainer or a group of maintainers, which are people responsible 
-for managing and accepting contributions into the subsystem they care of it \citep{melissa}.
+The contributions to Linux are coordinated with the usage of the Git source management tool.
+The contributions to the project are formatted as patches, which are text documents describing differences 
+between two different versions of a source tree The patches are sent and reviewed through the 
+Linux mailing lists \citep{melissa}.
 
-The contributions to the Linux are coordinate with the usage of the Git source management tool. The contributions 
-made to the project are formated as patches, which are text documents describing differences between two different 
-versions of a source tree \citep{melissa}.
+Mailing lists are the official means of communication of the Linux community, through the mailing lists 
+the community members send and review contributions to the kernel, discuss topics related to Linux such as 
+which new features to add, et cetera. As there are multiple subsystems in the Linux kernel, there are 
+multiple mailing lists in the Linux to organize the discussion made. Every subsystem has its mailing 
+list to concentrate the discussion related to the subsystem \citep{melissa}. 
+The interactions made in the mailing lists are documented and stored in the Linux kernel Lore
+\citep{linuxlore}, which serves as a repository of the discussion made in the Linux kernel to be
+consulted by anyone interested.
 
-TALK ABOUT HOW THE TOOL ENABLE A NEW WAY TO START CONTRITUTING TO THE KERNEL
+One of the Linux kernel subsystems is the AMD Display Driver, a subsystem responsible for 
+implementing the drivers required to enable AMD GPU to work properly in the Linux environment. 
+The maintainers of this subsystem reported an empirical problem they have in the subsystem, which is 
+a significant amount of duplicated code in the subsystem, which hinders the maintenance of the subsystem.
+This reported problem is the starting point that initiated our research, making the AMD Display Driver our 
+principal object of study to understand how to mitigate duplicated code in the Linux kernel.
 
-AMD Display Driver is a linux's subsystem responsible for the management of AMD GPU's. This subsystem have a problem 
-with code clone duplication acknowledged by the subsystem's maintainters. For this reason, we selected this
-subsystem as the primary focus for our empirical analysis in our research.
+(I WANT TO CHANGE THE AMD DISPLAY paragraph, WHY DO I NEED TO CHANGE IT? I SHOULD ANOTATE WHAT I WANTED, I MADE A CHANGE
+ TO MAKE IT BETTER BUT NOT SURE IF THERE IS SOME INFORMATION I WANTED TO PUT IN IT, 
+(MISSING AMD DISPLAY DRIVER MAILING LISTS, I WILL FIND IT LATTER)

### 2024-07-16 conteudo/background/codeduplication.tex
Fix grammar error with grammarly free tier, except bigclonebench

@@
 
-Code Duplication Detection is a field in computer science studies that gets the attention of researchers 
+Code Duplication Detection is a field in computer science studies that has the attention of researchers 
 back to 1988 \citep{firstman}.
-The occurrence of Code Duplication is a harmful artifact to have in a software, affecting software related 
-tasks such as code readability,
-introdution of bugs, etc.   \citep{harmone}. 
-On most of the cases, the occurrence of code duplication tends to create a more unstable software then the 
-nonduplicate code counterpart.   \citep{harmtwo}
+The occurrence of Code Duplication is a harmful artifact to have in software, affecting software-related 
+tasks such as code readability, introduction of bugs, etc \citep{harmone}. 
+In most cases, the occurrence of code duplication tends to create more unstable software 
+than the nonduplicate code counterpart \citep{harmtwo} .
 
-Through the years, the subject of studies about Code Duplication Detection branched out mainly in two fields of study,
-Code Clone Detection and Code Plagiarism Detection,
-with the former one focused on the technical aspect while the later one also introduces an social aspect on the field of research
-\citep{litreview}. For our research we are only interested on the Code Clone Detection branch given the nature of our work.
+Through the years, the subject of studies related to Code Duplication Detection branched out mainly in 
+two fields of study, Code Clone Detection and Code Plagiarism Detection, with the former focused on the 
+technical aspect while the latter also introduced a social aspect of the field of research
+\citep{litreview}. 
+In this work, we are only interested in the Code Clone Detection branch given the nature of our work.
 
@@
 
-The categorization of two code artifacts as a duplication of each other is a subjective task (REFERENCE HERE). 
-To mitigate this human aspect on the given problem, the literature classified code duplication on the scale of changes between 
-the code artifacts in four main types of code clone duplication \citep{litreview}. Given two code artifacts F and G, 
-the types of Code Duplication are described as follow \citep{litreview}:
+The categorization of two artifacts as a duplication of each other can open the question of what is considered 
+a duplication. To approach this question, the literature classified code duplication on the scale of the 
+changes between the code artifacts, creating a classification of four types of code clone 
+duplications \citep{litreview}. 
+When comparing the two code artifacts, F and G, the four types of code duplication can be described 
+as follows \citep{litreview}:
 
@@
 	\begin{item}
-		\textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names, white spaces 
-		and any other kind of irrelevant elements. (ADD FIGURE) shows a typical example of Type-1 Code 
+		\textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names,
+		white spaces, and any other kinds of irrelevant elements. 
+		(ADD FIGURE) shows a typical example of Type-1 Code 
 		Duplication \citep{litreview}. 
@@
 	\begin{item}
-		\textbf{Type-2:} The differences between F and G is the same as the type-1, with the addition of considering 
-		addition and deletion of redudant codes. (ADD FIGURE) shows a typical example of Type-2 Code Duplication \citep{litreview}. 
+		\textbf{Type-2:} The differences between F and G are the same as type-1, with the addition of 
+		considering the addition and deletion of redundant codes.
+		(ADD FIGURE) shows a typical example of Type-2 Code Duplication \citep{litreview}. 
 	\end{item}
 	\begin{item}
-		\textbf{Type-3:} The difference between F and G is the same as the type-2, with the addition of considering 
-		the reorder of code blocks as well as statements within code blocks. (ADD FIGURE) shows a typical example of 
+		\textbf{Type-3:} The difference between F and G is the same as type-2, with the addition of considering 
+		the reorder of code blocks, as well as statements within code blocks. (ADD FIGURE) shows a typical example of 
 		Type-3 Code Duplication \citep{litreview}. 
@@
 	\begin{item}
-		\textbf{Type-4:} The difference between F and G is the same as the type-3, with the addition of considering 
-		changes on data sctrucure, the order of operands/operators in expresssions, or replacing part of codes with 
-		equivalent composition. (ADD FIGURE) shows a typical example of Type-4 Code Duplication \citep{litreview}. 
+		\textbf{Type-4:} The difference between F and G is the same as type-3, with the addition of considering 
+		changes in data structures, the order of operands/operators in expressions, or replacing part of codes 
+		with equivalent composition.
+		(ADD FIGURE) shows a typical example of Type-4 Code Duplication \citep{litreview}. 
 	\end{item}
@@
 
-Given the passage of time, the literature developed many techniques and methods to approach the code clone detection problem. 
-For better understanding and analyze of these approachs, the literature divided the researches done in
-five main methodologies given the main nature of their work \citep{litreview} , which are:
+Given the passage of time, the literature developed many techniques and methods to approach the code clone 
+detection problem. For a better understanding and analysis of these approaches, the literature divided the 
+research done into five methodologies given the nature of their work  \citep{litreview}, which are:
 
 \begin{itemize}
-	\item \textbf{Textual-based approaches:} The literature review made by Chen \citep{litreview}  states that this methodology
-	is the first to be investigated by the literature. This line of methodology is based on see code as a text artifact and apply
-	textual approachs for text similarity. One example of this methodology applied is the work of Roy and Cordy \citep{textexample}, 
+	\item \textbf{Textual-based approaches:}  The literature review made by Chen \citep{litreview}  
+	states that this methodology is the first to be investigated by the literature. This methodology line 
+	is based on seeing code as a text artifact and applying textual approaches for text similarity. One 
+	example of this methodology applied is the work of Roy and Cordy
+	\citep{textexample},
 	which uses parsing and grammar techniques to detect code clones.
 
-	\item \textbf{Token-based approaches:} This methodoly tries to approach the problem by seeing the code in a similar way to
-	the lexical analysis of the compiling process \citep{litreview} . That means doing work in the line of recognizing constants, keywords and other 
-	specific programming languages tokens, mapping variables to <pair,value> representations to become naming-independent, etc. One example of 
-	this methodology applied  is the work of Toomey et al. \citep{tokenexample} to detect code clones through hashed token sequences.
+	\item \textbf{Token-based approaches:} This method tries to approach the problem by seeing the code in a
+	similar way to the lexical analysis of the compiling process \citep{litreview}. 
+	That means doing work along the lines of recognizing constants, keywords, and other specific programming 
+	language tokens, mapping variables to representations to become naming-independent, etc. One example 
+	of this methodology applied is the work of Toomey et al. \citep{tokenexample} 
+	to detect code clones through hashed token sequences.
 
-	\item \textbf{Tree-based approaches:} While the token-based approach uses the lexical analysis of the compiling process, the 
-	tree-based approach uses the syntax tree, the structure that  stores the semantic information of a code artifact in respect of the 
-	programming language \citep{compiler}. One example of research that uses this information to detect code clones is the one done by
-	Chilowicz and Duris \citep{treeexample}.
+	\item \textbf{Tree-based approaches:} While the token-based approach uses the lexical analysis of the 
+	compiling process, the tree-based approach uses the syntax tree. The syntax tree is a data structure that 
+	stores the semantic information of a code artifact concerning the programming language \citep{compiler}. 
+	One example of research that uses this information to detect code clones is done by Chilowicz and Duris
+	\citep{treeexample}.
 
-	\item \textbf{Metric-based approaches:} This methodology approches the problem by extracting metrics from the code artifact and use 
-	those metrics as a embedding representation of the artifact to measure similarity \citep{litreview}, similarly to the famous  
-	Word2Vec work done by Mikolov et al \citep{wordtovec}. Examples of metrics can be program size, number of variables, number of 
-	access to memory, etc. One example of this methodology applied is the work done by Kaur and Sharma. \citep{metricexample}
+	\item \textbf{Metric-based approaches:} This methodology approaches the problem by extracting metrics 
+	from the code artifact and using those metrics as an embedding representation of the artifact to measure 
+	similarity \citep{litreview}, similar to the famous Word2Vec work done by Mikolov et al.
+	\citep{wordtovec}. Examples of metrics can be program size, number of variables, number of access to
+	memory, etc. One example of this methodology applied is the work done by Kaur and Sharm
+	\citep{metricexample}.
 
-	\item \textbf{Graph-based approaches:} Similarly to the Tree-based approaches, this methotodology depends heavily on an intermediate
-	program representation, called program dependence graph (PDG) \citep{prodg}. The representation store information 
-	about program's control, statement and data dependencies. One example of usage of the program dependence graph is 
-	the recently work done by Liu et al. \citep{tailor}.
+	\item \textbf{Graph-based approaches:} Similarly to the Tree-based approaches, this methodology depends 
+	heavily on an intermediate program representation, called program dependence graph (PDG) \citep{prodg}. 
+	The representation stores information about the program’s control, statement, and data dependencies. 
+	One example of this methodology applied is the recent work done by Liu et al. \citep{tailor},
+	one of the state-of-the-art works in the field.
+
 \end{itemize}
 
-The state-of-art methods for code clone detection usually have high computational time and memory space usage, as well being
-programming language specific approches, which cannot be easily implemented in programming languagues that were not used in 
-their research \citep{litreview}. These are concerns for our research, for this reason we will use a generalist textual-based 
-approach based solely on Natural Languague Processing (NLP) text similarity detection methods, which we will introduce and compare
-with the literature approches later on this document.
+The state-of-the-art methods for code clone detection usually have high computational time and memory space 
+usage, as well being programming language-specific approaches, which cannot be easily implemented in programming 
+languages that were not used in their research \citep{litreview}. 
+These are concerns for our research, and for this reason, we will use a generalist textual-based approach 
+based solely on Natural Language Processing (NLP) text similarity detection methods, which we will introduce 
+in Chapter \ref{cha:tool} and evaluate this approach in this work.
 

### 2024-07-17 conteudo/background/nlpsimilarity.tex
Fix grammar errors with grammarly free tier

--- 2024-06-07 conteudo/background/nlpsimilarity.tex
@@
 \section{Text similarity detection methods}
+\label{sec:similarity}.
 
-The standard methodology for text similarity detection are based on building vector embeddings of
-the texts and apply a distance function in these embeddings to measure the similarity between them. 
-There are multiple methods for building vector embeddings, such as bag-of-words, TF-IDF, LSI, Word2Vec \citep{gensimlivro}. 
-The recent works related to build vector embeddings uses Largue Language Models (LLM), for example the usage of BERT, ELMo 
-and GPT \citep{llmsimilar}. The most popular distance function used by the literature are the cosine similarity 
-defined by the formula \citep{cosineref}:
+The standard methodology for text similarity detection is grounded on building vector embeddings of 
+the texts and applying a distance function in these embeddings to measure the similarity between them. 
+There are multiple vector embedding methods proposed by the literature, such as bag-of-words, TF-IDF, 
+LSI, and Word2Vec \citep{gensimlivro}.
+Recent works related to building vector embeddings use Large Language Models (LLM), for example,
+the work of Ethayarajh, which uses BERT, ELMo, and GPT \citep{llmsimilar}. 
+The most popular distance function used by the literature is the cosine similarity defined by the formula
+\citep{cosineref}:
 
@@
 
-Where $A$ and $B$ are the vector embeddings of the two texts. One of the advantages of the cosine similarity is that $SC(A,B)$ 
-is a number between $0$ and $1$, where $SC(A,B)$ values more close to the number $1$ means that two vectors are more similar.
+Where $A$ and $B$ are the vector embeddings of the two texts, an advantage of the cosine similarity is that $SC(A,B)$ is
+a number between $0$ and $1$, where $SC(A,B)$ values closer to the number $1$ means that two vectors are more similar, 
+it is usual to compare the cosine similarity to approximation to a probability value, transforming the values between 
+$0$ and $1$ to a percentage between $0 \% $ and $100\%$, through our research, we will use the percentage form.
 
-It is a famous fact that LLM have high computational cost, one example of such is a experiment done by Reimers and Gurevych for 
-finding the most similar pair in a collection of 10,000 sentences, that required about 65 hours with BERT \citep{bertsimilar}. This 
-is a similar magnitude of comparation pairs we want to do through our work and for this reason we chose to not use this kind of 
-vector embedding method, but it is a valid solution for smaller codebases.
+It is a famous fact that LLM has a high computational cost. An example of the high computational cost is an analysis 
+done by Reimers and Gurevych to find the most similar pair in a collection of 10,000 sentences, 
+which required about 65 hours with BERT \citep{bertsimilar}. 
+We want to compare a similar magnitude of pairs in our work, so we decided not to use LLM for our vector embedding 
+method, even though it is a valid solution for smaller codebases.
 
@@
 
-Gensim is a open source python library \citep{gensim} which the owners declare themselves as "The fastest library 
-for training of vector embeddings - Python or otherwise. The core algorithms in Gensim use battle-hardened, 
-highly optimized and parallelized C routines." \citep{gensimsite}. For the velocity of gensim we chose to use this library
-for doing our vector embeddings.
+Gensim is an open-source Python library \citep{gensim} which the owners declare themselves as "The fastest library 
+for training of vector embeddings - Python or otherwise. The core algorithms in Gensim use battle-hardened, highly 
+optimized and parallelized C routines." \citep{gensimsite}. For the velocity of Gensim, we chose to use this library 
+for our vector embedding method.
 
-In this work we will not explore more alternatives for vector embeddings. We do not want to optimize the computational 
-cost or the code clone detection to the state-of-art as they are not the main objectives of our research.
+In this work, we will not explore more alternatives for vector embeddings. We do not want to optimize the
+computational cost or the code clone detection to the state of the art as they are not the main objectives 
+of our research.

### 2024-07-17 conteudo/background/refactor.tex
Fix grammar errors with grammarly free tier

@@
 
-Fowler defines refactoring in his book as "a controlled technique for imporving the design of an existing code base. Its
-essence is applying a series of small behavior-preserving transformations, each of which too small to be worth doing. However
-the cumulative effect of each of these transformations is quite significant." \citep{refactorbook}. The elimination of detected 
-code clone duplication is a refactoring task that necessitates the application of refactoring methods. 
-Fowler presents multiples refactoring methods such as parameterize method, replace parameter with explicit methods and 
-replace parameter with method \citep{refactorbook}. There is a briefly explanation of the methods used in our research below.
+Fowler defines refactoring in his book as "a controlled technique for improving the design of an existing code base. 
+Its essence is applying a series of small behavior-preserving transformations, each of which is too small to be worth 
+doing. However, the cumulative effect of each of these transformations is quite significant." \citep{refactorbook}. 
+The elimination of detected code clone duplication is a refactoring task that necessitates the 
+application of refactoring methods. Fowler presents multiple refactoring methods such as the 
+\textit{parameterize method}, the \textit{replace parameter with explicit methods} and 
+the \textit{replace parameter with method} \citep{refactorbook}. 
+There is a brief explanation of the methods used in our research below.
 

### 2024-07-17 conteudo/tool/architecture.tex
Fix grammar erros with grammarly free tier

--- 2024-06-20 conteudo/tool/architecture.tex
@@
 
-The Setup step is composed of two components, the Function Breaker and the Code Duplication Finder. the pipeline of this step work as
-follow: the Function Parser receives the codebase we are interested, extract the functions of the codebase along with metadata and
-creates a new temporary codebase where it function becomes a new code file. The Code Duplication Finder iterate over every pair of 
-files in the temporary codebase computing the code duplications and save them in the Code Duplication Database, which is a text file
-that stores every code duplication as a triple <function1,function2,similarity>, where function1 and function2 are the functions that 
-are a duplication of each other, and similarity is the metric given by the code duplication method utilized, which in our case, it is
-the cossine similarity explained in \ref{sec:similarity}. Finally, the Query Responder consumes the temporary codebase and the 
-Code Duplication Database to extract code duplication related information as per user request. A diagram to ilustrate the pipeline
-can be found in the Figure \ref{fig:diagram}. There is a explanation of how each component work below.
+The Setup step consists of two components, the Function Breaker and the Code Duplication Finder. 
+The pipeline of this step works as follows: the Function Parser receives the codebase the user 
+is interested in, extracts the functions of the codebase along with metadata, and creates a new 
+temporary codebase where the functions extracted become news code files. The Code Duplication Finder 
+iterates over every pair of files in the temporary codebase, checks if they are code clones, and if 
+so, saves them in the Code Duplication Database, which is a text file that stores every code duplication 
+as a triple \textbf{<function1,function2,similarity>}, where function1 and function2 are the functions 
+that are a duplication of each other, the similarity is the metric given by the code duplication detection 
+method utilized, which in our case, it is the cosine similarity explained in \ref{sec:similarity}. 
+Finally, the Query Responder consumes the temporary codebase and the Code Duplication Database to 
+extract duplicated functions-related information per user request. Figure \ref{fig:diagram} illustrates 
+the tool pipeline. There is an explanation of how each component works through this section.
 
@@
 \includegraphics{diagrama_mestrado}
-\caption{The architecture diagram demonstrating the relationship between the components.}
+\caption{The architecture diagram demonstrating the relationship between the tool components.}
 \label{fig:diagram}
@@
 
-The input codebase is expected to be a folder in the machine the tool is running into. All files in the codebase that is not a source
-code file from a programming language supported by the tool will be ignored.
+The input codebase is a folder in the machine the tool is running into. All files in the codebase 
+that are not source code files from a supported programming language by the tool are ignored. The 
+tool validates the support for a source file through analysis of its file extension.
 
@@
 
-The Function Parser receives the input codebase and transforms it into the temporary codebase. The Function Parser iterates through
-every source code file from the programming language it support and use a specific programming languague's function 
-extractor to extract
-every function in the file, for each function extracted we create two new source code files and a metadata file in the temporary codebase
-which is represented by the pair \textbf{<source code file, function extracted>}, which represents the source code file the 
-function is extracted from and the 
-proper function. The first new source code file contains the function's body from the function it represents, while the other
-new source code file contains the function's signature, the metadata file 
-contains extra relevant information such as function name, the line the function signature starts in the original souce code 
-file and the line the function's body ends in the original source code file. The programming language we support at the moment 
-is C and Java. Can be found a example of a transform of a source code in the Figure (CREATE APENDIX AND ADD HERE).
+The Function Parser receives the input codebase and transforms it into the temporary codebase. The 
+Function Parser iterates through every source code file from the programming language it supports 
+and uses a specific programming language function extractor to extract every function in the file. 
+For each function extracted, we create two new source code files and a metadata file in the temporary 
+codebase, represented by the pair,  the source code file of the function, and the proper function. 
+The first new source code file contains the function’s body from the function it represents, while 
+the other new source code file contains the function’s signature. The metadata file contains 
+additional relevant information about the function, such as the function name, the line where the 
+function signature starts in the source code file, and the line where the function’s body ends in 
+the source code file. The programming language we support at the moment is C and Java. Can be found 
+an example of a transformation of a source code in Figure (CREATE FIGURE AND ADD HERE).
 
-About the programming languague's function extractor, we approached this problem in a way that any person comfortable with a language
-can adapt our approach to their specific programming language. Usually, the code blocks can be easily extracted from 
-source code files, independently if the code blocks are defined by curly brackets or by indentation, and it is possible to
-infer if it represents a function's body given how deeply it is in the code context and a few lines before the start 
-of the code block. We followed exactly this approach to implement the function extractor for the languagues we implemented. 
-As a alternative approach, we can build the syntax tree \citep{compiler} of the programming language and use it to 
-extract the functions, which is a more complex task then our main approach. For the programming languages we support, we implemented
-it as follow:
+About the programming language’s function extractor, we approached this problem in a way that any 
+person comfortable with a language can adapt our approach to their specific programming language. 
+Usually, the code blocks can be easily extracted from source code files, independently if the code 
+blocks are defined by curly brackets or by indentation, and it is possible to infer if it represents 
+a function’s body given how deeply it is in the code context and a few lines before the start of the 
+code block. We followed this approach to implement the function extractor for the supported languages. 
+As an alternative approach, we can build the syntax tree \citep{compiler} 
+of the programming language and use it to extract the functions, which is a more complex task than 
+our approach. For the programming languages that we support, we implemented them as follows:
 
@@
 	\begin{item}
-		\textbf{C}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in
-		the code context and som lines of code before the definition of the code block. By the C programming language's grammar, 
-		the functions are defined in a global context, with depth of 0. There are other elements with the same depth as functions,
-		we differentiate them by analysing if the first non-empty character before the curly brackets of the code block is a close 
-		parenthesis, since functions are the only elements that respect this pattern in the C programming language.
+		\textbf{C}: Given a source code file, we extract every code block from it, maintaining the depth 
+		of the code block in the code context and some of the code lines before the definition of the 
+		code block. According to the C programming language’s grammar, functions are defined in a 
+		global context at depth 0. There are other elements with the same depth as functions, and we 
+		differentiate them by analyzing if the first non-empty character before the curly brackets of 
+		the code block is a closing parenthesis. This pattern uniquely identifies functions 
+		within C code.
+
 	\end{item}
 	\begin{item}
-		\textbf{Java}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in
-		the code context and som lines of code before the definition of the code block. By the Java's grammar, 
-		the functions are defined inside classes. There are different depths which we can have functions, as Java accepts inner
-		classes. We differentiate functions of other elements by analysing if the first non-empty character before the curly brackets of the code block is a close 
-		parenthesis, since there are others elements which also respect this pattern such as for loop, we would need to apply other 
-		filters. We chose to only analyse functions of depth 1, which ignores every others elements that respect the previously
-		filters, but as counterpart, functions declared inside Inner Classes are not supported by our extractor. We made this choice
-		as there are not inner classes on the codebases explored in this work.
+		\textbf{Java}:  Given a source code file, we extract every code block from it, maintaining the 
+		depth of the code block in the code context and some of the code lines before the definition 
+		of the code block. According to Java’s grammar, the functions are defined inside classes. 
+		Functions can exist in different depths, as Java accepts inner classes. We differentiate 
+		functions of other elements by analyzing if the first non-empty character before the curly 
+		brackets of the code block is a close parenthesis. There is a problem with this approach in 
+		Java, as there are other elements that respect this pattern, such as for loop, thus requiring 
+		more filters. We chose to analyze functions of depth 1, which ignores every other element that 
+		is a code block and has a parenthesis as the first non-empty character before the curly brackets. 
+		The counterpart to this choice is that the extractor does not find functions declared inside 
+		inner classes. The counterpart is not a problem in this work as there are no inner classes 
+		on the BigCloneBench \citep{bigclonebench}, the only codebase in Java explored in this work.
 
@@
 
-The temporary codebase is a transformation of the input codebase done by the Function Parser, every function in the input codebase
-is represented in the temporary codebase as three files. There are description of each file below and an example of each can found
-in the figure (CREATE FIGURE AND ADD HERE).
+The temporary codebase is a transformation of the input codebase done by the Function Parser. 
+Each function in the input codebase is represented in the temporary codebase as three files.
+Descriptions of these files are provided below, and an example of each can be found in 
+the figure (CREATE FIGURE AND ADD HERE).
 
@@
 	\begin{item}
-		\textbf{Info file:} This file contains metadata of the function. The information this file contains is:
-		the function name, the file name that has the function, the relative path of the file related to the input codebase, 
-		the line that the function's signatures starts, the line that the functions body starts, the line that the function's body 
-		ends and if there is a line break between the end of the function's signature and the start of the function's body.
-
+		\textbf{Info file:} This file contains metadata of the function. The information this 
+		file contains is the function name, the file that has the function, the relative path 
+		of the file related to the input codebase, the line that the function’s signatures 
+		start, the line that the function body initiates, the line that the function’s body 
+		ends and if there is a line break between the end of the function’s signature and the 
+		start of the function’s body.
 	\end{item}
@@
 
-The Code Duplication Finder iterates through every pair of source code files in the temporary codebase, which in this context represents
-functions of the input codebase, and for every pair of files, we execute a code duplication detection method that computes a metric
-that measures of how similar the pair of files are, which we call similarity. Finally, if the similarity is greater of equal then the
-minimum similarity threshold (a parameter passed by the user when it executes the Setup step), we store this pair of files along with
-it similarity in the Code Duplication Database.
+The Code Duplication Finder iterates through every pair of source code files in the temporary 
+codebase, which in this context represents functions of the input codebase and, for every pair 
+of files, we execute a code duplication detection method that computes a metric that measures 
+of how similar the pair of files are, which we call similarity. Finally, if the similarity is 
+greater or equal to the minimum similarity threshold (a parameter passed by the user when it 
+executes the Setup step), we store this pair of functions along with its similarity in the 
+Code Duplication Database.
 
-For the code duplication detection method, we visualize the source code files as texts and execute the TF-IDF vector embedding method
-implemented by the Gensim library \citep{gensim} and compute the cosine similarity as the similarity metric. We chose this method 
-given the alegated Gensim statements about itself performance, being a programming language independent method and the nonnecessity of 
-compilable code, which is expected to happen in the temporary codebase as it not stores complete code artifacts. Platis implemented
-this code cuplication detection method and distributed it to the public with a MIT license \citep{platistool} \citep{mitlicense}, 
-we utilize his implementation as it has some nice to have components such as removing code comments, we do the appropriated changes
-in the input and output of his implementation to respect the expected format of our tool. 
-It is feasible to change the duplication detection method in a manner similar to what we did with Platis's implementation.
+For the code duplication detection method, we visualize the source code files as texts and 
+execute the TF-IDF vector embedding method implemented by the Gensim library \citep{gensim}
+and compute the cosine similarity as the similarity metric. We chose this method given the
+alleged Gensim statements about its performance, being a programming language-independent 
+method, and the nonnecessity of compilable code, which is expected to happen in the temporary 
+codebase as it does not store complete code artifacts. Platis implemented this code 
+duplication detection method and distributed it to the public with an MIT license
+\citep{platistool,mitlicense}.
+We utilized his implementation as it has some useful components, such as removing code 
+comments. We made the appropriate changes in the input and output of his implementation 
+to respect the expected format of our tool. It is feasible to change the duplication 
+detection method similarly to what we did with Platis’s implementation.
 
@@
 
-The Code Duplication Database is a text file that contains the duplicated pairs of functions found by the Code Duplication Finder, 
-the first line of the file contains a number that represents the number of the duplicated pairs, after that, there is 
-a line of each duplicated pairs that contains the path of the functions in the temporary codebase and the similarity metric of the 
-pair. A example of the Code Duplication Database can be found in the figure (ADD FIGURE HERE).
+The Code Duplication Database is a text file that contains the duplicated pairs of 
+functions found by the Code Duplication Finder. The first line of the file contains 
+a number that represents the number of the duplicated pairs. After that, there is a 
+line for each duplicated pair containing the path of the functions in the temporary codebase 
+and the similarity metric of the pair. An example of the Code Duplication Database can be 
+found in the figure (ADD FIGURE HERE).
 
@@
 
-The Setup step is a procedure executed by the user once per codebase. The Setup step receives the input codebase and the minimum
-similarity metric value to a pair of functions to be considered a duplication, then it executes the Function Parser and the Code 
-Duplication Finder, to generate the temporary codebase and the Code Duplication Database, which will be later used by the Query
-Responder step. The value of the minimum similarity metric is to reduce the Code Duplication Database, which optimize the
-memory usage and the computational cost of the Query Responder step. For example, if this parameter does not exist and we work in 
-a codebase of $10000$ functions and supposing each relative path plus the function name have $50$ charactes, the Code Duplication 
-Database will be a file of size approximately $5000^2 \times 50 ~= 5$ Gigabytes, as it is not expected to a real large codebase 
-to the majority of pairs of functions be a duplication of each other, this number can be reduced dramastically.
+The Setup step is a procedure executed by the user once per codebase. The Setup step takes 
+the input codebase and a minimum similarity metric value for function pairs considered 
+duplications. It then runs the Function Parser and the Code Duplication Finder to create 
+a temporary codebase and the Code Duplication Database. The value of the minimum similarity 
+metric is to reduce the Code Duplication Database, which optimizes the memory usage and the 
+computational cost of the Query Responder step. For example, if this parameter does not exist 
+and we work in a codebase of $10000$ functions and supposing each relative path plus the 
+function name has 50 characters, the Code Duplication Database will be a file of size 
+approximately $5000^2 \times 50 ~= 5$ Gigabytes, as it is not expected in large codebase 
+to the majority of pairs of functions be a duplication of each other, 
+the size file can be reduced dramatically.
 
@@
 
-The Query Responder step is the component that the user executes multiples times to consult information about code duplication in
-a codebase that was previously preprocessed by the Setup step. The Query Responder step consumes the temporary codebase and the 
-Code Duplication Database created by the Setup step. This component implements all the tool's functionalities exposed 
-for the user, which are described in the \ref{subsec:func}.
+The Query Responder step is the component the user executes multiple times to consult information 
+about duplicated functions in a codebase processed by the Setup step. The Query Responder step 
+consumes the temporary codebase and the Code Duplication Database created by the Setup step. 
+This component implements all the tool’s functionalities exposed to the user, described in 
+\ref{subsec:func}.

### 2024-07-17 conteudo/tool/tool.tex
Fix grammar erros with grammarly free tier

--- 2024-06-14 conteudo/tool/tool.tex
@@
 
-Our proposed tool is a Command Line Interface (CLI) software, the tool purpose is to enable developers to query 
-information about code clone duplication in a function level, by function level we mean to find functions in a codebase that 
-are code duplications. The tool is separated in two steps. The 
-first one, denoted as Setup, is executed only once for each codebase and it is responsible for processing the
-codebase doing tasks such as extracting functions of the code files 
-and searching for code clone duplications. The second step, denoted as Query Responder, is
-responsible to execute the tool functionalities using the information obtained in the Setup step. More information about the 
-Setup and Query Responder behavior can be found in \ref{subsec:architecture} and in \ref{subsec:func}.
+Our proposed tool is a Command Line Interface (CLI) software. Its purpose is to enable developers to query 
+information about code clone duplication at the function level, which means finding pairs of functions in a codebase 
+that are code clones of each other.
+
+The tool operates in two steps. The first step, called Setup, is executed only 
+once for each codebase and is responsible for processing the codebase by extracting functions from code files and 
+searching for code clone duplications. The second step, called Query Responder, executes the tool's functionalities 
+using the information obtained in the Setup step. More information about the Setup and Query Responder behavior can 
+be found in sections \ref{subsec:architecture}  and \ref{subsec:func}.

### 2024-07-17 conteudo/tool/functionalities.tex
Fix grammar erros with grammarly free tier

@@
 
-We proposes three main functionalities for the user in the tool, which are named duplication explorer, function information
-and duplication relatory. Each of the functionalities receives their adequated parameters to execute, with a exception
-of a optional parameter shared by all of them, which is minimum similarity per query. The minimum similarity per query is different
-from the siminum similarity presented in \ref{subsec:setup}, as it enables a volatile change in the minimum similarity
-interested in a query specific context. There is a explanation of each of the main functionalities through this section.
+We propose three main functionalities for the user in the tool, called duplication explorer,
+function information, and duplication relatory. Each of the functionalities receives its 
+adequated parameters to execute, except an optional parameter shared by all of them, which 
+is minimum similarity per query. The minimum similarity per query is not the same as the 
+minimum similarity presented in \ref{subsec:setup}, as it enables a volatile change in the minimum 
+similarity interest in a query-specific context. There is an explanation of each of the 
+main functionalities in this section.
 
@@
 
-The duplication explorer is the principal functionality of our tool, the functionality purpose is to expose for the user the 
-duplicated functions pairs found by our tool. We implement optional filters methods to enable more complex queries, 
-which are described below. A example of the functionality usage can be found in the figure \ref{fig:explorer_ex}.
+The duplication explorer is the principal functionality of our tool. The functionality's
+purpose is to expose the user to the duplicated function pairs found by our tool. We 
+implement optional filter methods to enable more complex queries, as described below. 
+Figure \ref{fig:explorer_ex} demonstrates the functionality usage.
 
@@
 	\begin{item}
-		\textbf{Sorting filter:} This filter enables the user to sort the resuls by similarity of the functions pairs or 
-		by the number of duplicated lines. By default the tool sorts by similarity and the user can pass a sort parameter to change
+		\textbf{Sorting filter:} This filter enables the user to sort the results by the 
+		similarity of the function pairs or by the number of duplicated lines. By default, 
+		the tool sorts by similarity, and the user can pass a sort parameter to change 
 		the sorting method.
@@
 	\begin{item}
-		\textbf{Limiter filter:} This filter receives a number parameter named limiter by the user, which limits the number of results
-		showed to the user by this given number.
+		\textbf{Limiter filter:} This filter receives a number parameter named limiter 
+		by the user, which limits the number of results shown to the user by this given number.
 	\end{item}
@@
 	\begin{item}
-		\textbf{Pattern filter:} This filter make the explorer ignore every pair of duplicated functions that does not match a pattern
-		given by the user as a parameter. We say that a function matches a pattern if the string formed by the concatenation of
-		the relative path of the code file plus the function name contains the pattern as a substring. 
-		The user can define if both functions or at least one of the 
-		functions in the pair need to match the pattern by given a specific parameter.
+		\textbf{Pattern filter:} This filter makes the explorer ignore every pair 
+		of duplicated functions that do not match a pattern given by the user as a parameter.
+		We say that a function matches a pattern if the string formed by the concatenation 
+		of the relative path of the code file plus the function name contains the pattern 
+		as a substring. The user can define if both functions or at least one of the 
+		functions in the pair need to match the pattern by giving a specific parameter.
 	\end{item}
@@
 
-The Function information purpose it to give a more in depth information about a specific function. The functionality 
-receives a function the user is interested in and then inform the user
-about the relative path, the function name and the lines the function is defined for the function the user is interested and for
-every function that is a duplication of the given function.
-There can be multiple functions with the 
-same name in a codebase, to distinguish them, we receive a pattern string from the user, and  we use the first function find that match
-the given pattern. We say that a function matches a pattern if the string formed by the concatenation of the relative path of the 
-code file plus the function name contains the pattern as a substring. 
-A example of the functionality usage can be found in the figure \ref{fig:function_ex}.
+The Function information purpose is to give more in-depth information about a specific 
+function. The functionality receives a function the user is interested in and then 
+informs the user about the relative path, the function name, and the lines the function 
+is defined for the function the user is interested in and for every function that is a 
+duplication of the given function. There can be multiple functions with the same name 
+in a codebase. To distinguish them, we receive a pattern string from the user, and we 
+use the first function to find one that matches the given pattern. We say that a 
+function matches a pattern if the string formed by the concatenation of the relative 
+path of the code file plus the function name contains the pattern as a substring. 
+Figure \ref{fig:function_ex} demonstrates the functionality usage.
 
@@
 
-The duplication relatory purpose is to give a overview of the input codebase. The functionality extract the information
-of how much lines of duplicated code is found per folder in the codebase and present it to the user in a reasonable format. To extract
-the information, we use the Code Duplication Database to list the duplicated functions, and the temporary code to know how much lines of
-duplicated code contains in each of the functions. (DO I TALK ABOUT THE TRIE USAGE TO MAINTAIN THE HIERARCHICAL TREE STRUCTURE? I 
-THINK NOT). A example of the functionality usage can be found in the figure \ref{fig:relatory_ex}.
+The duplication relatory purpose is to give an overview of the input codebase. 
+The functionality extracts how many lines of duplicated code per folder in the 
+codebase and presents it to the user in a reasonable format. To extract the 
+information, we use the Code Duplication Database to list the duplicated functions 
+and the temporary code to know how many lines the duplication functions have. 
+Figure \ref{fig:relatory_ex} demonstrates the functionality usage. 
+(DO I TALK ABOUT THE TRIE USAGE TO MAINTAIN THE HIERARCHICAL TREE STRUCTURE? I THINK NOT)
 

### 2024-07-17 conteudo/tool/architecture.tex
Fix typo

@@
 For each function extracted, we create two new source code files and a metadata file in the temporary 
-codebase, represented by the pair,  the source code file of the function, and the proper function. 
+codebase, represented by the pair \textbf{<file name, function name>},  the source code file of the function, and the proper function. 
 The first new source code file contains the function’s body from the function it represents, while 

### 2024-07-18 conteudo/background/codeduplication.tex
Add small missing things in background

@@
 
-\subsection{Methodology to evaluate code duplication methods}
+\subsection{Evaluating code duplication methods}
 
@@
 
-Given the necessity to evaluate and compare approches for code clone detection, it was developed and widely-adopted two datasets: 
-OJ-Clone, a dataset collected from a pedagogical online judge system \citep{ojclone}, and BigCloneBench, a dataset collected from over 
-25,000 Java software systems \citep{bigclonebench}. The existence of only two spreaded datasets is a open concern in literature as 
-the limited scope of OJ-Clone and the problems of BigCloneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail}
+Given the necessity to evaluate and compare approaches for code clone detection, it was developed and widely 
+adopted: OJ-Clone, a dataset collected from a pedagogical online judge system \citep{ojclone},
+and BigCloneBench, a dataset collected from over 25,000 Java software systems \citep{bigclonebench}. 
+The existence of only two spread datasets is an open concern in literature as the limited scope of OJ-Clone 
+and the problems of BigCloneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail}. The BigCloneBench
+is the dataset used by most of the literature works, given the limited pedagogical scope of the OJ-Clone.
 
-(ADD INFORMATION ABOUT THE MORE SEPARATION OF CLONE TYPES THE BIGCLONEBENCH makes)
+To evaluate the code clone detection approaches against the datasets, the works compute metrics using 
+the predictions made related to the instances of the dataset. There are four possible outcomes when an 
+instance is predicted by a code clone detection approach, which are:
+
+\begin{itemize}
+	\begin{item}
+		\textit{True positivity (TP)}: This outcome happens when the approach predicts that the instance 
+		is a code clone and the instance is a code clone.
+	\end{item}
+	\begin{item}
+		\textit{False positivity (FP)}: This outcome happens when the approach predicts that the instance 
+		is a code clone, but the instance is not a code clone.
+	\end{item}
+		\begin{item}
+		\textit{True negative (TN)}: This outcome happens when the approach predicts that the instance 
+		is not a code clone and the instance is not a code clone.
+	\end{item}
+	\begin{item}
+		\textit{False negative (FN)}: This outcome happens when the approach predicts that the instance 
+		is not a code clone, but the instance is a code clone.
+	\end{item}
+\end{itemize}
+
+Given the number of TP, FP, TN, and FN that resulted after an approach evaluation against a dataset,
+it is possible to compute classical metrics such as Precision, Recall, and F-Score \citep{recall}.
+In this work, we use the Recall metric as the code clone detection approaches usually make available 
+this metric to compare themselves in contrast with the others. The Recall metric is defined by the 
+following formula \citep{recall}:
+
+$$Recall = \frac{\sum TP}{\sum TP + \sum FN }$$
+
+Where $\sum TP$ is the number of true positives in the evaluation and $\sum FN$ is the number of 
+false negatives. The Recall metric is a number between $0$ and $1$, where a value closer to $1$ is 
+better, and a value closer to $0$ is worse.

### 2024-07-18 conteudo/background/linux.tex
Add small missing things in background

@@
 
-One of the Linux kernel subsystems is the AMD Display Driver, a subsystem responsible for 
-implementing the drivers required to enable AMD GPU to work properly in the Linux environment. 
-The maintainers of this subsystem reported an empirical problem they have in the subsystem, which is 
-a significant amount of duplicated code in the subsystem, which hinders the maintenance of the subsystem.
-This reported problem is the starting point that initiated our research, making the AMD Display Driver our 
-principal object of study to understand how to mitigate duplicated code in the Linux kernel.
+One of the Linux kernel subsystems is the AMD Display Driver, a subsystem responsible for implementing 
+the drivers required to enable AMD GPU to work correctly in the Linux environment. The maintainers of 
+this subsystem reported an empirical problem they have in the subsystem, which is a significant amount 
+of duplicated code in the subsystem, which hinders the maintenance of the subsystem. The reported issue 
+is the starting point that initiated our research, making the AMD Display Driver our principal object 
+of study to understand how to mitigate duplicated code in the Linux kernel. The AMD Display Driver 
+contains X code lines, Y functions/structs, and Z code files, representing a significant amount of 
+the Linux kernel. As with any other Linux subsystem, the discussions related to the AMD Display Driver 
+are made in its proper mailing list (FIND MAILING LIST AND CITE HERE).
 

### 2024-07-18 conteudo/methods/methods.tex
Fix grammar errors with grammarly free tier

--- 2024-06-28 conteudo/methods/methods.tex
@@
 
-To answer our proposed research questions, we opted to divide our research. Firstly, we will approach the supplementary research 
-RQ1.1, which will serve as a background to our research question RQ1. To answer the RQ1.1, we propose a multiple-method investigation,
-by validating our proposed tool in the BigCloneBench databaset \citep{bigclonebench}, which is the standard method used by the
-literature, and a empirical analyses of a selected subset of functions in which the tool infers.
-With the question RQ1.1 answered, we will return our focus to the RQ1. To answer the RQ1, we opted to create a ethnographic study, 
-using participant-observation methodology, we expect to mitigate a selected subset of duplicated functions by refactoring them using
-a designed sistematic approach, while iteracting with the Linux community to collect feedback of the mitigation made by us.
+To answer our proposed research questions, we opted to divide our research. Firstly, we will approach 
+the supplementary research RQ1.1, which will serve as a background to our research question RQ1. 
+To answer the RQ1.1, we propose a multiple-method investigation by validating our proposed tool in 
+the BigCloneBench dataset \citep{bigclonebench}, which is the standard method used by the literature, 
+and an empirical analysis of a selected subset of functions in which the tool infers. With the question 
+RQ1.1 answered, we will return our focus to the RQ1. To answer the RQ1, we opted to create an 
+ethnographic study using participant-observation methodology. We expect to mitigate a selected subset 
+of duplicated functions by refactoring them using a designed systematic approach while interacting 
+with the Linux community to collect feedback on the mitigation made by us.
 
@@
 
-The multiple-method investigation to verify the efficiency of our tool in a methodical way is a method in which we consider 
-trustworthy.
-The first method used is the evaluation of the proposed tool in the BigCloneBench database \citep{bigclonebench}, one of the 
-standard code duplication databases used by validation of tool presented in the formal literature. 
-As presented by Krinke and Ragkhitwetsagul, the BigCloneBench is not a perfect dataset \citep{bigfail}, and for that reason we opted to do
-a additional evalution, in which we will do a empirical analyses on a select subset of functions of the AMD Display driver in which
-our tool infers it is a duplication. To a concrete evaluatiion, in both of the methods we will apply a  variation of the similarity 
-metric (The similarity metric is a metric measuring the similarity of two code duplications, a proper explanation is made 
-in the Chapter \ref{cha:tool}).
+The multiple-method investigation to verify the efficiency of our tool is a method that we 
+consider trustworthy. The first method used is to evaluate the proposed tool in the BigCloneBench 
+database \citep{bigclonebench} one of the standard code duplication databases used to validate 
+the tool presented in the formal literature. As presented by Krinke and Ragkhitwetsagul, the 
+BigCloneBench is not a perfect dataset \citep{bigfail}. For that reason, we opted 
+to do an additional evaluation, in which we will do an empirical analysis on a select subset of 
+functions of the AMD Display driver, which our tool infers is a duplication. For a concrete evaluation,
+in both methods, we will apply a variation of the similarity metric (The similarity metric is a 
+metric that measures the similarity of two code duplications; a proper explanation is made
+in Chapter \ref{cha:tool}).
 
@@
 
-The BigCloneBench dataset \citep{bigclonebench} is a collection of Java code with a selection of code duplications pairs categorized by the authors, these
-code duplications pairs are not separed solely if they are duplications, the pairs are also categorized by the code duplication
-types presented in Subsection \ref{subsec:types}.
+The BigCloneBench dataset \citep{bigclonebench} is a collection of Java code with a selection of
+code duplication pairs categorized by the authors. These code duplication pairs are labeled by 
+the code clone type described in \ref{subsec:types}.
 
-Liut et al presented a sample method to evaluate the their code duplications detection method, in which they sample 20,000 pairs
-from each of the code duplication types, with the addition of 20,000 nonduplication pairs to act as the negative samples
-\citep{tailor}. To a fairness evaluation, we use the same sample method as them to evaluate our tool.
+Liut et al. presented a sample method to evaluate their code duplication detection method, 
+in which they sample 20,000 pairs from each of the code duplication types with the addition 
+of 20,000 nonduplication pairs to act as the negative samples \citep{tailor}. 
+For a fair evaluation, we use the same sample method as them to evaluate our tool.
 
-The proposed tool does not differentiate the code duplication types and infers duplications in the function level instead of the 
-code file level, which is the standard in the state-of-art code duplication methods, we were required to adjust the form 
-we compute metrics of the evalution. To accomplish this, we will define that every pair of functions in which the similarity metric
-is equal or superior to a threshold X will be considered a duplication, and the pair of code files that the functions are contained
-are marked as duplication. If on this method, a code duplication pair is corretly indentified, it will count as a correct to its 
-code duplication types, if a incorrect duplication pair is infered, it will count as incorrect to all code duplication types. Given
-these definitions, we will compute the Recall metric presented in \ref{subsec:codemethods} for each of the code duplication types.
+The proposed tool does not differentiate the code duplication types and infers duplications 
+at the function level instead of the code file level, which is the standard in state-of-the-art 
+methods. We were required to adjust the form we computed metrics of the evaluation. To accomplish 
+this, we will define that every pair of functions in which the similarity metric is equal or 
+superior to a threshold X will be considered a duplication, and the pair of code files in which 
+the functions exist are labeled as duplication. If, on this method, a code duplication pair is 
+correctly identified, it will count as correct to its code duplication types. If an incorrect 
+duplication pair is inferred as a duplication, it will count as incorrect to all code duplication 
+types. Given these definitions, we will compute the Recall metric presented in \ref{subsec:codemethods}
+for each code duplication type.
 
-To understand the variation of meaning when we change the similarity threshold, we will eval our tool for the given variations of
-the threshold X and discuss the results: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, 100\%.
+To understand the variation of meaning when we change the similarity threshold, we will evaluate 
+our tool for the given variations of the threshold X and discuss the results: 
+30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, 100\%.
 
@@
 
-Given the problems of the BigCLoneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail} and the limitations of it in our
-scope such as being a Java dataset while the Linux kernel are primarily build in the C programming language, we opted to, together
-with the formal litarature evaluation, perform a empirical study analyzing the codes in which our tool infered as duplications.
+Given the problems of the BigCLoneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail}
+and the limitations of it in our scope, such as being a Java dataset while the Linux kernel 
+is predominantly built in the C programming language, we opted to, together with the formal 
+literature evaluation, perform an empirical study analyzing the codes in which our tool inferred 
+as duplications.
 
-The number of code lines and components in the Linux kernel makes it impractical to analyse every code duplication. To limit our
-scope, we will analyse solely code duplications in the AMD Display driver. This limitation is still does not reduce the code
-duplication to analyse, since in our preliminar tool execution showed in the Figure \ref{fig:relatory_ex}, there is over 20,000
-lines of code duplications on the AMD Display driver with similarity metric of 100\%.
+The number of code lines and components in the Linux kernel makes it impractical to analyze every 
+code duplication. To limit our scope, we will analyze solely code duplications in the AMD Display 
+driver. This limitation still does not reduce the code duplication to analyze. In our preliminary 
+tool execution shown in Figure \ref{fig:relatory_ex}, there are over 20,000 lines of code 
+duplications on the AMD Display driver with a similarity metric of 100\%.
 
-For conducting our empiricial analyses, we opted to do a random selection in the code duplications pairs infered by our tool. To
-do that, we will use the definition that if the similarity metric of a function pair is greater or equal to X, will be considered a
-duplication. To understand the optimal values of X we will analyse 10 functions pair randomic selected 
-that have similarity metric X for each of the variations of the parameter X, which are 
+For our empirical analyses, we opted to make a random selection in the code duplication pairs 
+inferred by our tool. To do that, we will use the definition that if the similarity metric of a 
+function pair is greater or equal to X is considered a duplication. To understand the optimal 
+values of X, we will analyze 10 functions pair random selected that have similarity metric X 
+for each of the variations of the parameter X, which are 
 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, 100\%. 
-As it is not expected to have code duplications with similarity exactly a whole number, we will allow a
-variation of 1\% for greater or less to each of the variations of the parameter X.
+As it is unexpected to have code duplications with similarity equals a whole number, we will 
+allow a variation of 1\% for greater or less to each of the variations of the parameter X.
 
-To do the random selection of functions pairs, we list the code duplications for each of the variations of X and execute 
-a random selection algorithm
-builtin in in the source code language of our tool. The exactly algorithm of the random selection, along with the selected 
-function pairs can be found in Appendix (CREATE APENDICE AND ADD HERE).
+To make the random selection of function pairs, we list the code duplications for each of the 
+X variations and execute a random selection algorithm built into the source code language of 
+our tool. The exact algorithm of the random selection, with the selected function pairs, can 
+be found in the Appendix (CREATE Appendix AND ADD HERE).
 
-When we analyses the selected functions pairs, we will judged empiricaly if they are a duplication or not, then we are not
-going further as
-categorizing the duplications by code duplication type. We intend to collect the data on this empirical analyses and compare it
-to the formal literature method, to undertand if our tool in fact detect duplicated functions and 
-the difference it makes when varying the minimum similarity parameter of our tool.
+When we analyze the selected function pairs, we will empirically judge if they are duplications,
+which means we are not going deeper as categorizing the duplications by code duplication type. 
+We intend to collect the data on this empirical analysis and compare it to the formal literature 
+method to understand if our tool detects duplicated functions and the difference it makes when 
+varying the minimum similarity parameter of our tool.
 
@@
 
-Ethnography is a research method recognized as a significant qualitative empirical approach 
-suited to understanding people and cultures, and their associated social and work practices \citep{bookethno}. 
-The central tenet of the ethnography approach is to enable research to undertand values, beliefs or ideas shared for a 
- community under study from the members' point of view \citep{ethnosoft}. Thus, the ethnographer is required to act as
-a member of  the community, learning and observing deeply what people do and why they do.
+Ethnography is a research method recognized as a significant qualitative empirical approach to 
+understanding people, their cultures, and their associated social and work practices \citep{bookethno}.
+The central tenet of the ethnography approach is to enable researchers to understand values, 
+beliefs, or ideas shared by a community under study from the members’ point of view 
+\citep{ethnosoft}. Thus, the ethnographer acts as a community member, learning and observing 
+what people do and why they do it.
 
-Sharp et al discuss the application of the ethnography method in the software engineering research, demonstrating the low adoption 
-of the method and its advantages to uncover what practioners do and why they do \citep{ethnosoft}. In our research context, if we
-mitigate code duplications by ourselves without iteracting with the Linux community, we cannot undertand if the linux maintainers 
-would accept our approach for code duplication and what concerns they would have. Thus, we resorted to a ethnography research, where
-we will act as a member of the Linux kernel, begin a contributor with focus on removing code duplications found.
+Sharp et al. discuss the application of the ethnography method in software engineering research, 
+demonstrating the low adoption of the method and its advantages in uncovering what practitioners 
+do and why they do it \citep{ethnosoft}. In our research context, if we mitigate code 
+duplications by ourselves without interacting with the Linux community, we cannot understand if 
+the Linux maintainers would accept our approach for code duplication and what concerns they 
+would have. Thus, we resorted to ethnography research, where we will act as a member of the 
+Linux kernel and act as a contributor with a focus on removing code duplications found.
 
-On this section, we will describe a generic approach pipeline to mitigate code duplication found in the Linux kernel. Then we
-will describe the methodology to select a subset of duplicated functions found by our tool and finally, describe the ethnographic
-methodology to interact with the Linux community.
+In this section, we will outline a systematic approach to mitigate code duplication in the 
+Linux kernel. Next, we will detail the methodology for selecting a subset of duplicated functions 
+identified by our tool. Finally, we will describe the ethnographic methodology for interacting 
+with the Linux community.
 
-\subsection{Generic approach pipeline to mitigate code duplication}
+\subsection{Systematic approach to mitigate code duplication}
 
@@
 
-The C programming language's way to communicate between source code files is through creation of header files, which is
-a library specification \citep{Cbook}. As the Linux kernel is mainly build in C, we resorted to remove code duplications by
-creating a library, where the code artifacts, instead of using their duplications, will pass to use the library.
+The C programming language’s way to communicate between source code files is through creating 
+header files, which is a library specification \citep{Cbook}. As the Linux kernel 
+is predominantly built in C, we remove code duplications by creating a library where the code 
+artifacts will use the library instead of their duplications.
 
-When removing duplicated functions, our desired behavior is to this function to every portion of the code it exist, only allowed a
-unique version existent in the library created by us. For this reason, given our initial duplicated function pair, we will use 
-the function information functionality(This functionality is described in \ref{subsec:functioncommand}) of our tool to find all 
-places that this function is defined in the source code. Thus, we finish with a collection of duplicated function and the collection
-of code files in which these functions are defined into.
+When removing duplicated functions, our desired behavior is to remove the functions in every 
+portion of the code in which the functions exist, only allowing a unique version to exist in 
+the library created. For this reason, given our initial duplicated function pair, we will use 
+the function information functionality(Described in \ref{subsec:functioncommand}) to find all
+places where this function exists in the source code. Thus, we finish with a collection of 
+duplicated functions and the code files in which these functions exist.
 
-Given the semantics of a library to containing the shared code by a collection of code files, we added a additional step of the 
-approach, which is searching for others functions contained in all of code files of the collection we are working with. 
-On this moment, we have a collection of code files, and a non-empty set of functions which are shared by all of the code files 
-in the collection.
+Given the semantics of a library containing the shared code by a collection of code files, 
+we added a step of the approach, which is searching for other functions contained in all of 
+the code files of the collection we are working with. Thus, we have a collection of code files 
+and a non-empty set of functions shared by all the code files in the collection.
 
-Finally, for each of the functions shared by all code files, we do a specific-case aproach to mitigate each one of them.
-If a function is shared perfectly by all the code files, we use a simply approach to only remove them and create the function in
-the source code file, if it is not shared perfectly, we look for refactoring methods which can be applied on this specific case.
+Finally, for each of the functions shared by all code files, we make a specific-case approach 
+to mitigate each one of them. If a function is shared by all the code files without 
+modifications, we use a simple approach to only remove them and create the function in the 
+source code file. If it is shared with modifications, we look for refactoring methods applicable 
+to this specific case.
 
-Finishing our mitigation method, if the mitigation ended up working in a initial step, we move it foward to the 
-Ethnographic method, iteracting with the Linux community to understand what are their thoughts on the mitigation. If the mitigation
-ended up as not working as expected, we will categorized it in two categories, the first one is that the refactoring requires a
-decision in which we, as not specialist and non-maintainers of the Linux kernel, cannot make it correctly, although
-we believe a specialist would be able to refactoring it. The second categorization is that we do not know if it is in fact possible
-to refactoring the duplication.
+Finishing our mitigation method, if the mitigation ends up working in an initial step, we may 
+move it forward to the Ethnographic method, interacting with the Linux community to understand 
+their thoughts on the mitigation. If the mitigation ends up not working as expected, we will 
+categorize it in two categories, the first one is that the refactoring requires a decision in 
+which we, as not specialists and non-maintainers of the Linux kernel, cannot make it correctly,
+although we believe a specialist would be able to refactor it. The second categorization is 
+that we do not know if it is possible to refactor the duplication. With or without successful 
+mitigation, we will store our mitigations on (ADD REP AND PUT HERE) to research transparency.
 
-The figure (CREATE FIGURE AND ADD HERE) ilustrates the approach pipeline described in this section.
+The figure (CREATE FIGURE AND ADD HERE) illustrates the systematic approach described in this section.
 
@@
 
-As discussed in the empyrical analyses for the tool, it is impractical to evaluate every code duplication on Linux kernel. Thus,
-similarily to the empyrical analyses for the tool, we will limit the experiment scope to only the AMD Display driver and select
-a random sample of duplicated functions infered by our tool to initiate the mitigation approach defined in \ref{subsec:pipeline}. 
-We chose a sample size of 30 duplicated functions, as there are multiples studies demonstrated that a sample size of this size 
-have a statistical significance \citep{sample1,sample2}.
+As discussed in the empirical analyses for the tool, it is impractical to evaluate every code 
+duplication on the Linux kernel. Thus, similarly to the empirical analyses for the tool, we will 
+limit the experiment scope to only the AMD Display driver and select a random sample of duplicated 
+functions inferred by our tool to initiate the mitigation approach defined in \ref{subsec:pipeline}. 
+We chose a sample size of 30 duplicated functions, as multiple studies demonstrated that a sample 
+size of this size has statistical significance \citep{sample1,sample2}.
 
-On the the random selection of the duplicated functions, we deside to put a additional restriction on which functions pairs are 
-elegible to be selected. The restriction is that we will only select functions pairs with similarity metric equal 
-to 100\%, we made this choice because, as showed in the Figure \ref{fig:relatory_ex}, there are more then 20,000 lines codes 
-in the AMD Display driver in which the similarity metric reaches this value. On all function pairs that respect the restriction,
-we will choose then uniformly at random.
+On the random selection of the duplicated functions, we decided to put an additional restriction 
+on which function pairs are eligible to be selected. The restriction is that we will choose function 
+pairs with a similarity metric equal to 100\%. We made this choice because, as shown in \ref{fig:relatory_ex},
+there are more than 20,000 lines of code in the AMD Display driver in which the similarity metric 
+reaches this value. On all function pairs that respect the restriction, we will choose them 
+uniformly at random.
 
-The functions pairs selected at random following our methodoly can be consulted in the Table (CREATE TABLE AND ADD HERE).
+(MAYBE I SHOULD ADD A RESTRICTION ABOUT IGNORING AUTOMATED GENERATED CODE, such as DML folder)
+
+The function pairs selected randomly following our methodology can be consulted in the Table
+(CREATE TABLE AND ADD HERE).
 
@@
 \subsection{Ethnographic methodology}
+
+When we make a code duplication mitigation following our systematic approach described in 
+\ref{subsec:pipeline}, we are apt to initiate our ethnographic research. We resorted to a
+participant-observation approach as our ethnographic methodology, where we will act as 
+contributors of the Linux kernel to the AMD Display driver, interacting with the maintainers 
+of the AMD Display driver attempting to get our mitigations added to the Linux kernel.
+
+The Linux kernel communication is made through public mailing lists, where anyone can consult 
+the discussions. We plan to collect data artifacts through the AMD Display driver mailing list. 
+The mailing list will register our interactions with the Linux community, with their members 
+sharing their opinions and suggestions for the mitigation we made, corroborating the discussion 
+about the effectiveness of our proposed approach.
+
+Given the time limit of our research, we do not expect to be able to interact with the Linux 
+community throughout the successful mitigation we made using the proposed approach, as it passes 
+through factors we do not control, which is the time spent waiting for Linux community’s members 
+interact with us. For this reason, we resorted to committing to do the participant-observation 
+approach in 5 of the successful mitigations and use them for our discussion.

### 2024-07-18 conteudo/methods/methods.tex
Reformulate init of the chapter

@@
 
-To answer our proposed research questions, we opted to divide our research. Firstly, we will approach 
-the supplementary research RQ1.1, which will serve as a background to our research question RQ1. 
-To answer the RQ1.1, we propose a multiple-method investigation by validating our proposed tool in 
-the BigCloneBench dataset \citep{bigclonebench}, which is the standard method used by the literature, 
-and an empirical analysis of a selected subset of functions in which the tool infers. With the question 
-RQ1.1 answered, we will return our focus to the RQ1. To answer the RQ1, we opted to create an 
-ethnographic study using participant-observation methodology. We expect to mitigate a selected subset 
-of duplicated functions by refactoring them using a designed systematic approach while interacting 
-with the Linux community to collect feedback on the mitigation made by us.
+To answer our proposed research questions, we opted to do a multiple-method research 
+where we covered the research question with different methodologies. To approach the RQ1, 
+we opted to create an ethnographic study using participant-observation methodology where 
+we plan to mitigate a selected subset of duplicated functions by refactoring them using 
+a designed systematic approach while interacting with the Linux community to collect 
+feedback on the mitigation made by us.
 
-\section{Multiple-method Investigation}
+To answer RQ1, which covers our main objective to mitigate code duplication in the Linux 
+kernel, we are first required to be able to identify them. For this reason, we developed 
+a tool and need to understand if the tool can identify code duplications in the Linux kernel, 
+which covers the supplementary research question RQ1.1 for the RQ1.
 
-The multiple-method investigation to verify the efficiency of our tool is a method that we 
-consider trustworthy. The first method used is to evaluate the proposed tool in the BigCloneBench 
-database \citep{bigclonebench} one of the standard code duplication databases used to validate 
-the tool presented in the formal literature. As presented by Krinke and Ragkhitwetsagul, the 
-BigCloneBench is not a perfect dataset \citep{bigfail}. For that reason, we opted 
-to do an additional evaluation, in which we will do an empirical analysis on a select subset of 
-functions of the AMD Display driver, which our tool infers is a duplication. For a concrete evaluation,
-in both methods, we will apply a variation of the similarity metric (The similarity metric is a 
-metric that measures the similarity of two code duplications; a proper explanation is made
-in Chapter \ref{cha:tool}).
+To answer RQ1.1, we applied two independent submethods and performed a triangulation of 
+the results. The first submethod is to evaluate the proposed tool similarly to the literature, 
+by comparing it against the BigCloneBench dataset \citep{bigclonebench}. The second submethod
+for answering RQ1.1 is to do an empirical analysis on a selected subset of functions of the 
+AMD Display driver that our tool infers is a duplication. With the two submethods completed, 
+we will do a triangulation of results analyzing the findings. We opted for the triangulation 
+of the two submethods because the formal literature method alone is limited and not fully 
+trustworthy \citep{bigfail, litreview}.
 
-\subsection{Evaluation by the formal literature}
+\subsection{Evaluation of the tool by the formal literature}
 
@@
 
-\subsection{Evaluation by Empirical analyses}
+\subsection{Evaluation of the tool by Empirical analyses}
 

### 2024-07-19 conteudo/introduction/intro.tex
Try to change multiple-method as the whole methodology

@@
 
-\textbf{RQ1:} How effective is our approach in identifying and mitigating duplicate code in the Linux kernel?
+\textbf{RQ1:} How effective is our approach in mitigating duplicate code in the Linux kernel?
 
@@
 
-This research is divided into two phases. In \textbf{Phase I}, we will propose, build, and validate a tool capable 
+To address our research question, we do a multiple-method research utilizing different methodologies throughout 
+the study. We structured this work into two phases, each with distinct research methods.
+
+In \textbf{Phase I}, we will propose, build, and validate a tool capable 
 of identifying code duplications at the function level for the C programming language. We expect this proposed tool 
@@
 
+Figure (CREATE FIGURE AND ADD HERE) sumamarizes the research steps proposed in this work.
+
 \section**{Manuscript Structure}

### 2024-07-26 conteudo/background/codeduplication.tex
Add Type Figures

@@
 		\textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names,
-		white spaces, and any other kinds of irrelevant elements. 
-		(ADD FIGURE) shows a typical example of Type-1 Code 
-		Duplication \citep{litreview}. 
+		white spaces, and any other kinds of irrelevant elements. Figure
+		\ref{fig:type1} shows a typical example of Type-1 clone duplication 
+		\citep{litreview}. 
 	\end{item}
@@
 		\textbf{Type-2:} The differences between F and G are the same as type-1, with the addition of 
-		considering the addition and deletion of redundant codes.
-		(ADD FIGURE) shows a typical example of Type-2 Code Duplication \citep{litreview}. 
+		considering the addition and deletion of redundant codes. Figure
+		\ref{fig:type2} shows a typical example of Type-2 code clone duplication  \citep{litreview}. 
 	\end{item}
@@
 		\textbf{Type-3:} The difference between F and G is the same as type-2, with the addition of considering 
-		the reorder of code blocks, as well as statements within code blocks. (ADD FIGURE) shows a typical example of 
-		Type-3 Code Duplication \citep{litreview}. 
+		the reorder of code blocks, as well as statements within code blocks. Figure
+		\ref{fig:type3} shows a typical example of 
+		Type-3 code clone duplication \citep{litreview}. 
 	\end{item}
@@
 		changes in data structures, the order of operands/operators in expressions, or replacing part of codes 
-		with equivalent composition.
-		(ADD FIGURE) shows a typical example of Type-4 Code Duplication \citep{litreview}. 
+		with equivalent composition. Figure
+		\ref{fig:type4} shows a typical example of Type-4 code clone duplication \citep{litreview}. 
 	\end{item}
@@
 \end{itemize}
+
+\begin{figure}
+\includegraphics[scale=0.38]{types/type1}
+\caption{Example of Type-1 code clone duplication.}
+The example is based on the example presented by CHEN et al. \citep{litreview}. 
+\label{fig:type1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.38]{types/type2}
+\caption{Example of Type-2 code clone duplication.}
+The example is based on the example presented by CHEN et al. \citep{litreview}. 
+\label{fig:type2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.38]{types/type3}
+\caption{Example of Type-3 code clone duplication.}
+The example is based on the example presented by CHEN et al. \citep{litreview}. 
+\label{fig:type3}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.38]{types/type4}
+\caption{Example of Type-4 code clone duplication.}
+The example is based on the example presented by CHEN et al. \citep{litreview}. 
+\label{fig:type4}
+\end{figure}
 

### 2024-07-26 conteudo/tool/functionalities.tex
do figures desc

@@
 \includegraphics[scale=0.5]{explorer_example}
-\caption{Explorer Example}
+\caption{Example of the \textit{duplication explorer} functionality of the proposed tool in action.}
 \label{fig:explorer_ex}
@@
 \includegraphics[scale=0.5]{function_example}
-\caption{Function Example}
+\caption{Example of the \textit{function information} functionality of the proposed tool in action.}
 \label{fig:function_ex}
@@
 \includegraphics[scale=0.5]{relatory_example}
-\caption{Function Example}
+\caption{Example of the \textit{duplication relatory} functionality X of the proposed tool in action.}
 \label{fig:relatory_ex}

### 2024-07-26 conteudo/introduction/intro.tex
add design diagram
adjust research figure and intro tex

@@
 
+\begin{figure}
+\includegraphics[scale=0.7]{research_design}
+\caption{Diagram of the research design.}
+\label{fig:reDesign}
+\end{figure}
+
 Given the first phase completion, we would have a validated tool to identify code duplications in the Linux kernel 
@@
 
-(POINTS TO CONSIDER: I do not like much this introduction, but is the best I could do)
-(I DO NOT HAVE MORE A PLACE TO FIT THIS PARAGRAPH) An additional usefulness of our proposed approach, which we do not verify
-in our research, is to give a new method to people become 
-contributors to the Linux kernel. The process of starting to contribute to the Linux kernel can be unintuitive, which can cause 
-people to distance themselves. Our approach provides to the community a new form of newcomers to start contributing by removing 
-simple code duplications, while they learn the whole process of being a Linux contributor.
-
-Figure (CREATE FIGURE AND ADD HERE) sumamarizes the research steps proposed in this work.
+Figure \ref{fig:reDesign} sumamarizes the research steps proposed in this work.
 

### 2024-07-27 figuras/example_function_parser/file_name.c
add initial code artifacts for example

@@ -1,0 +1,53 @@
+#include <stdio.h>
+
+//two equal sum functions with different logic
+int sum_of_two_elements(int left_hand, int right_hand){
+	return left_hand + right_hand;
+}
+
+int sum(int a, int b){
+	int ret = a+b;
+	return ret;
+}
+
+//tree functions that find the sum of elements of an array
+int sum_of_array(int array[], int n){
+	int ret = 0;
+	for(int i = 0; i < n; i++){
+		ret += array[i];
+	}
+	return ret;
+}
+
+int sum_of_array2(int n, int array[]){
+	int ret = 0;
+	for(int i = 0; i < n; i++){
+		ret += array[i];
+	}
+	return ret;
+}
+
+int sum_of_vector(int vector[], int size_of_vector)
+{
+	int ret = 0;
+	for(int i = size_of_vector-1; i >= 0; i--){
+		ret += vector[i];
+	}
+	return ret;
+}
+
+
+int main(int argc, char *argv[]){
+
+	int v[5];
+	v[0] = 1;
+	v[1] = 2;
+	v[2] = 3;
+	v[3] = 4;
+	v[4] = 5;
+	int n = 5;
+	int C = sum_of_array(v,5);
+	int D = sum_of_array2(5,v);
+	int E = sum_of_vector(v,5);
+	return 0;
+}

### 2024-07-27 conteudo/tool/architecture.tex
add transformation fig

@@
 \includegraphics{diagrama_mestrado}
-\caption{The architecture diagram demonstrating the relationship between the tool components.}
+\caption{Architecture diagram demonstrating the relationship between the tool components.}
 \label{fig:diagram}
@@
 function signature starts in the source code file, and the line where the function’s body ends in 
-the source code file. The programming language we support at the moment is C and Java. Can be found 
-an example of a transformation of a source code in Figure (CREATE FIGURE AND ADD HERE).
+the source code file. The programming language we support at the moment is C and Java. 
+Figure \ref{fig:transform} illustrates an example of a function transformation into the 
+temporary codebase.
+
+\begin{figure}
+\includegraphics[scale=0.3]{example_function_parser/example_parser}
+\caption{Transformation of a function into the temporary codebase}
+\label{fig:transform}
+\end{figure}
 
@@
 Descriptions of these files are provided below, and an example of each can be found in 
-the figure (CREATE FIGURE AND ADD HERE).
+the figure \ref{fig:transform}.
 

### 2024-07-27 conteudo/methods/methods.tex
Remove commentary about systematic fig, I do not want it anymore
remove commentary about add random selection algo, it is stupid

@@
 X variations and execute a random selection algorithm built into the source code language of 
-our tool. The exact algorithm of the random selection, with the selected function pairs, can 
+our tool. The random selected function selected function pairs can 
 be found in the Appendix (CREATE Appendix AND ADD HERE).
@@
 
-The figure (CREATE FIGURE AND ADD HERE) illustrates the systematic approach described in this section.
-
 \subsection{Selection of the duplicated functions}

### 2024-07-29 conteudo/preliminar/table_empirical.tex
create tables of tool

\begin{table}
\begin{tabular}{ | c | c | c | c | c | c | m{8em} | }

\hline

\textbf{Similarity range} & \textbf{T1} & \textbf{T2} & \textbf{T3} & \textbf{T4}
& \textbf{Negative} & \textbf{Percentage} \\ \hline 
99\% - 100\% & 9 & 0 & 0 & 1 & 0 & \textbf{100\%} \\ \hline
89\% - 91\% & 0 & 8 & 0 & 1 & 1 & \textbf{90\%} \\ \hline
79\% - 81\% & 0 & 3 & 2 & 3 & 2 & \textbf{80\%} \\ \hline
69\% - 71\% & 0 & 3 & 1 & 1 & 5 & \textbf{50\%} \\ \hline
59\% - 61\% & 0 & 0 & 0 & 2 & 8 & \textbf{20\%} \\ \hline
49\% - 51\% & 0 & 0 & 0 & 0 & 0 & \textbf{0\%} \\ \hline
39\% - 41\% & 0 & 0 & 0 & 0 & 0 & \textbf{0\%} \\ \hline
29\% - 31\% & 0 & 0 & 0 & 0 & 10 & \textbf{0\%} \\ \hline

\hline

\end{tabular}
\label{tab:emp}
\caption{Results of the proposed tool in the empirical analyses in the AMD Display driver}

The results were obtained by randomly selecting ten functions in the AMD Display driver for 
each similarity range and analyzing if they are duplications and which duplication type.

\end{table}

### 2024-07-30 conteudo/eval/eval.tex
Add initial tool eval

On this Chapter, we exposes our experiment results bout the proposed tool performance following
the methology described in Chapter \ref{cha:method}. Firstly, we shows the results obtained by our tool
in the BigCloneBench dataset\citep{bigclonebench}, which is the standard form of the formal literature 
\citep{litreview}. Secondly, we shows the results obtained by our tool on the empirical analyses. 
Finaly, we do a triangulation of data analysing our finds on the tool performance.

\subsection{Tool results on the BigCloneBench dataset}

\input{conteudo/eval/table_bigclone}

\subsection{Tool results on the empirical analyses}

\input{conteudo/eval/table_empirical}

The results were obtained by randomly selecting ten functions in the AMD Display driver for 
each similarity range and analyzing if they are duplications and which duplication type.

\subsection{Discussion}

### 2024-07-30 conteudo/background/codeduplication.tex
add t3 split on big

@@
 	\begin{item}
-		\textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names,
+		\textbf{Type-1(T1):} The differences between F and G are in the context of revision of comments, variable names,
 		white spaces, and any other kinds of irrelevant elements. Figure
@@
 	\begin{item}
-		\textbf{Type-2:} The differences between F and G are the same as type-1, with the addition of 
+		\textbf{Type-2(T2):} The differences between F and G are the same as type-1, with the addition of 
 		considering the addition and deletion of redundant codes. Figure
@@
 	\begin{item}
-		\textbf{Type-3:} The difference between F and G is the same as type-2, with the addition of considering 
+		\textbf{Type-3(T3):} The difference between F and G is the same as type-2, with the addition of considering 
 		the reorder of code blocks, as well as statements within code blocks. Figure
@@
 	\begin{item}
-		\textbf{Type-4:} The difference between F and G is the same as type-3, with the addition of considering 
+		\textbf{Type-4(T4):} The difference between F and G is the same as type-3, with the addition of considering 
 		changes in data structures, the order of operands/operators in expressions, or replacing part of codes 
@@
 
-To evaluate the code clone detection approaches against the datasets, the works compute metrics using 
+The BigCloneBench dataset does an additional split on the code clone duplications types for the Type-3,
+where the dataset separated the Type-3 into three types, which are Weak Type-3 (WT3), Medium Type-3 (MT3), 
+Strong Type-3 (ST3), with the WT3 being closer to the Type-4 and ST3 being closer to the Type-2 \citep{bigclonebench}.
+Finally,  the BigCloneBench dataset unifies the WT3 with the T4, resulting in 6 possible outcomes of a code pair, 
+which are Type-1, Type-2, Strong Type-3, Medium Type-3, Weak Type-3/Type-4, and not a duplication \citep{bigclonebench}.
+
+To evaluate the code clone detection approaches against the datasets, the researchs compute metrics using 
 the predictions made related to the instances of the dataset. There are four possible outcomes when an 

### 2024-07-30 conteudo/eval/eval.tex
add more initial eval

@@
 
+As described in section \ref{sec:metbig}. For the evaluation of our proposed tool in the
+BigCLoneBench dataset  \citep{bigclonebench}, we define that every pair of functions in which the 
+similarity metric is equal or superior to a threshold X will be considered a duplication, and the 
+pair of code files in which the functions exist are labeled as duplication. If, on this method, 
+a code duplication pair is correctly identified, it will count as correct to its code duplication 
+types. If an incorrect duplication pair is inferred as a duplication, it will count as incorrect 
+to all code duplication types. Table \ref{tab:bigclone} shows the results obtained by our tool in the 
+BigCloneBench dataset \citep{bigclonebench}.
+
+The results obtained in Table \ref{tab:bigclone} indicate that our tool is a pessimist. There 
+is no negative sample inferred as duplication even for a minimum similarity threshold of 30\%, 
+on the other hand, it is bad for identifying code duplications of complex code duplication types, 
+that is Medium Type-3 and Weak Type-3/Type-4. As mentioned in Chapter \ref{cha:tool}, if a 
+better result is required on these complex code duplication types, it is possible to change 
+the code duplication detection method, as it is designed to be a replaceable component in our 
+tool, given the limitations.
+
 \subsection{Tool results on the empirical analyses}

### 2024-07-31 conteudo/methods/methods.tex
add more initial eval
adjust methods

@@
 
-\subsection{Evaluation of the tool by the formal literature}
+\section{Evaluation of the tool by the formal literature}
+
+\label{sec:metbig}
 
@@
 
-\subsection{Evaluation of the tool by Empirical analyses}
+\section{Evaluation of the tool by Empirical analyses}
+
+\label{sec:metemp}
 
@@
 code duplication. To limit our scope, we will analyze solely code duplications in the AMD Display 
-driver. This limitation still does not reduce the code duplication to analyze. In our preliminary 
+driver. This limitation still does not reduce the code duplication to analyze, in our preliminary 
 tool execution shown in Figure \ref{fig:relatory_ex}, there are over 20,000 lines of code 
@@
 
-When we analyze the selected function pairs, we will empirically judge if they are duplications,
-which means we are not going deeper as categorizing the duplications by code duplication type. 
+When we analyze the selected function pairs, we will empirically judge if they are duplications. 
 We intend to collect the data on this empirical analysis and compare it to the formal literature 

### 2024-07-31 conteudo/eval/eval.tex
add base empirical results

@@
 
-The results obtained in Table \ref{tab:bigclone} indicate that our tool is a pessimist. There 
+The results presented in Table \ref{tab:bigclone} indicate that our tool is a pessimist. There 
 is no negative sample inferred as duplication even for a minimum similarity threshold of 30\%, 
@@
 
+As described in section \ref{sec:metemp} For the evaluation of our proposed tool by our empirical 
+analyses, we randomly selected ten function pairs for a variation of ranges of the similarity 
+metric computed by our tool. After that, we analyzed if they were duplications, and if yes, 
+we judged the duplication type. Table \ref{tab:emp} presents the results obtained by our tool 
+in the empirical analyses.
+
 \input{conteudo/eval/table_empirical}
 
-The results were obtained by randomly selecting ten functions in the AMD Display driver for 
-each similarity range and analyzing if they are duplications and which duplication type.
+The results presented in Table \ref{tab:emp} indicate that as the similarity metric reduces, 
+there is a significant increase in function pairs that are not duplications in the Linux kernel. 
+For the ranges 79\%-81\% of similarity or greater, we have a high percentage of code duplications. 
+With 69\%-71\% of similarity, the rate goes down to 50\%. With 59\%-61\%, the rate decreases to 20\%.
+With 49\%-51\% similarity or lower, we did not find any code duplications.
 

### 2024-07-31 conteudo/eval/eval.tex
do triangulation

@@
 
-\subsection{Discussion}
+\subsection{Triangulation of results}
+
+Both the BigCloneBench and the empirical method results point out that our tool has a good 
+result for detecting Type-1 and Type-2 code clone duplications and has a poor performance 
+in detecting the other types. Using a similarity metric between 80\% and 100\% shows good 
+results in both methods.
+
+There is a discrepancy in the results of non-code duplications inferred as duplications 
+between the two methods. In the BigCloneBench results, we did not find any negative samples,
+while in the empirical analyses, we found a higher percentage of negative samples. We have 
+two possible reasons for this discrepancy.
+
+The first reason is the limitations and problems of the BigCloneBench presented by Krinke 
+and Ragkhitwetsagul \citep{bigfail}. The second reason is that the AMD Display driver 
+is a system where each code artifact shares a meaning. While the BigCloneBench is a 
+collection of self-contained code artifacts without sharing meaning with each, as our 
+tool uses a text-based code clone detection method, it is expected to have poorer 
+performance in the AMD Display driver than in the BigCloneBench dataset. We do not know 
+if the state-of-the-art code duplication detection has the same problem and is obscured 
+by the evaluation in the BigCloneBench dataset.

### 2024-07-31 conteudo/apendices/emp-lista.tex
format structure and remove artifacts from example

\chapter{Random selected samples}

THIS APPENDIX IS FOR LIST THE RANDOM SELECTED FUNCTIONS PAIRS IN THE 
EMPIRICIAL ANALYSES OF THE TOOL. I will put it here later.

### 2024-07-31 conteudo/preliminar/results.tex
create tables of tool
add initial preliminar

@@
 
-Base Content of the Preliminar Results of the research. TODO.
+We finished Phase 1 of our research, where we developed our proposed tool and 
+evaluated the tool on the BigCloneBench dataset and with our empirical analyses. 
+We discussed the results found to understand that the tool is capable of identifying 
+code clone duplications in the Linux kernel. With these results, we answered RQ1.1 
+(\textit{How accurate is the proposed tool for detecting duplicate functions?}).
+
+Related to our RQ1
+(\textit{How effective is our approach in mitigating duplicate code in the Linux kernel?}),
+we intend to approach our ethnographic study presented in \ref{sec:meteth} as the next steps 
+in our work. As an early validation to see the potential of this research, we conducted the 
+proposed systematic approach to mitigate code duplication on two function pairs. Table
+(CREATE TABLE AND ADD HERE) shows the early results found.
+
+CREATE TABLE AND ADD HERE
+
+The first function pair requires an expert/maintainer on the AMD Display driver to finalize 
+the refactorization. The reason for this is that the duplicated code artifacts use 
+copy-pasted code but import different configuration files. The AMD Display driver 
+configurations files design requires the code that uses it to import the exact configuration 
+in the compiler time without being generic. For this reason, the refactorization requires 
+a change in the form the AMD Display driver deals with configuration files, which is a more 
+complex task that requires an expert/maintainer on the AMD Display driver.
+
+The second function pair refactorization was successful. At the moment of writing this 
+report, we are sending the refactorization as a patch to the AMD Display driver to know 
+the opinion of the maintainers of the driver.
+
+\section{Next steps}
+
+As shown in our proposed schedule (Figure CREATE FIGURE AND ADD HERE), as we advance, 
+we intend to do the proper ethnographic study, starting by completing our systematic 
+approach to mitigate code duplications, and with this mitigation on hands we intend 
+to process to send them to the AMD Display driver mailing list to understand the 
+AMD Display driver maintainers about the mitigations made by us.

### 2024-07-31 conteudo/background/codeduplication.tex
space on T

@@
 	\begin{item}
-		\textbf{Type-1(T1):} The differences between F and G are in the context of revision of comments, variable names,
+		\textbf{Type-1 (T1):} The differences between F and G are in the context of revision of comments, variable names,
 		white spaces, and any other kinds of irrelevant elements. Figure
@@
 	\begin{item}
-		\textbf{Type-2(T2):} The differences between F and G are the same as type-1, with the addition of 
+		\textbf{Type-2 (T2):} The differences between F and G are the same as type-1, with the addition of 
 		considering the addition and deletion of redundant codes. Figure
@@
 	\begin{item}
-		\textbf{Type-3(T3):} The difference between F and G is the same as type-2, with the addition of considering 
+		\textbf{Type-3 (T3):} The difference between F and G is the same as type-2, with the addition of considering 
 		the reorder of code blocks, as well as statements within code blocks. Figure
@@
 	\begin{item}
-		\textbf{Type-4(T4):} The difference between F and G is the same as type-3, with the addition of considering 
+		\textbf{Type-4 (T4):} The difference between F and G is the same as type-3, with the addition of considering 
 		changes in data structures, the order of operands/operators in expressions, or replacing part of codes 

### 2024-08-03 conteudo/introduction/intro.tex
add the eval chap on
remove bullshit

@@
 primitive solutions that find every pair of code files that are duplications. 
-The AMD Display driver was created in 20XX (TODO HERE) and maintained until today. The code files that 
+The code files on the AMD Display driver that 
 contain duplicated code normally also have specific code that is only implemented in this file. 
@@
 
-This manuscript consists of four more chapters.
+This manuscript consists of five more chapters.
 Chapter \ref{cha:back} presents the literature overview for code duplication detection (main definitions, 
@@
 Chapter \ref{cha:method} describes the research methods selected to guide our work.
+Chapter \ref{cha:eval} shows the results obtained of the research methods to evaluate our proposed tool.
 Chapter \ref{cha:results} presents the current research stage and the work plan, with the addition of 

### 2024-08-03 conteudo/apendices/cem.tex
create appendix

\section{Function pairs with similarity between 99\% and 100\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/gpio/dcn20/hw\_factory\_dcn20.c

\textbf{Function name 1:} define\_ddc\_registers

\textbf{File 2:}a dc/gpio/dcn21/hw\_factory\_dcn21.c

\textbf{Function name 2:} define\_ddc\_registers

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/gpio/dcn10/hw\_factory\_dcn10.c

\textbf{Function name 1:} dal\_hw\_factory\_dcn10\_init

\textbf{File 2:} dc/gpio/dcn21/hw\_factory\_dcn21.c

\textbf{Function name 2:} dal\_hw\_factory\_dcn21\_init

\textbf{Similarity:} 99.10\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 3 }

\

\textbf{File 1:} dmub/src/dmub\_dcn32.c

\textbf{Function name 1:} dmub\_dcn32\_is\_supported

\textbf{File 2:} dmub/src/dmub\_dcn35.c

\textbf{Function name 2:} dmub\_dcn35\_is\_supported

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/bios/dce112/command\_table\_helper2\_dce112.c

\textbf{Function name 1:} clock\_source\_id\_to\_atom\_phy\_clk\_src\_id

\textbf{File 2:} dc/bios/dce60/command\_table\_helper\_dce60.c

\textbf{Function name 2:} clock\_source\_id\_to\_atom\_phy\_clk\_src\_id

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/resource/dcn314/dcn314\_resource.c

\textbf{Function name 1:} dcn31\_opp\_create

\textbf{File 2:} dc/resource/dcn315/dcn315\_resource.c

\textbf{Function name 2:} dcn31\_opp\_create

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/resource/dcn314/dcn314\_resource.c

\textbf{Function name 1:} dcn31\_apg\_create

\textbf{File 2:} dc/resource/dcn316/dcn316\_resource.c

\textbf{Function name 2:} dcn31\_apg\_create

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 7 }

\

\textbf{File 1:} dmub/src/dmub\_dcn20.c

\textbf{Function name 1:} dmub\_dcn20\_set\_outbox0\_rptr

\textbf{File 2:} dmub/src/dmub\_dcn35.c

\textbf{Function name 2:} dmub\_dcn35\_set\_outbox0\_rptr

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/resource/dcn301/dcn301\_resource.c

\textbf{Function name 1:} dcn301\_dpp\_destroy

\textbf{File 2:} dc/resource/dcn31/dcn31\_resource.c

\textbf{Function name 2:} dcn31\_dpp\_destroy

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/resource/dce110/dce110\_resource.c

\textbf{Function name 1:} dce110\_transform\_destroy

\textbf{File 2:} dc/resource/dce80/dce80\_resource.c

\textbf{Function name 2:} dce80\_transform\_destroy

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/resource/dcn32/dcn32\_resource.c

\textbf{Function name 1:} dcn32\_clock\_source\_create

\textbf{File 2:} dc/resource/dcn321/dcn321\_resource.c

\textbf{Function name 2:} dcn321\_clock\_source\_create

\textbf{Similarity:} 100\%

\textbf{Duplication type:} Type-1

### 2024-08-03 conteudo/apendices/cinquenta.tex
create appendix

\section{Function pairs with similarity between 49\% and 51\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/hwss/dcn10/dcn10\_hwseq.c

\textbf{Function name 1:} dcn10\_init\_pipes

\textbf{File 2:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 2:} dcn20\_program\_front\_end\_for\_ctx

\textbf{Similarity:} 50.27\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/bios/command\_table\_helper.c

\textbf{Function name 1:} dal\_cmd\_table\_helper\_encoder\_id\_to\_atom

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm.c

\textbf{Function name 2:} to\_drm\_connector\_type

\textbf{Similarity:} 50.51\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 3 }

\

\textbf{File 1:} dmub/src/dmub\_dcn32.c

\textbf{Function name 1:} dmub\_dcn32\_get\_inbox1\_wptr

\textbf{File 2:} dmub/src/dmub\_dcn31.c

\textbf{Function name 2:} dmub\_dcn31\_set\_inbox1\_wptr

\textbf{Similarity:} 49.64\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/dcn30/dcn30\_mpc.c

\textbf{Function name 1:} mpc3\_program\_luta

\textbf{File 2:} dc/dpp/dcn10/dcn10\_dpp\_cm.c

\textbf{Function name 2:} dpp1\_cm\_program\_regamma\_lutb\_settings

\textbf{Similarity:} 50.96\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/pg/dcn35/dcn35\_pg\_cntl.c

\textbf{Function name 1:} pg\_cntl35\_io\_clk\_status

\textbf{File 2:} dc/pg/dcn35/dcn35\_pg\_cntl.c

\textbf{Function name 2:} dcn\_pg\_cntl\_destroy

\textbf{Similarity:} 49.52\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/dml/dcn314/display\_rq\_dlg\_calc\_314.c

\textbf{Function name 1:} is\_dual\_plane

\textbf{File 2:} dc/hubbub/dcn30/dcn30\_hubbub.c

\textbf{Function name 2:} convert\_and\_clamp

\textbf{Similarity:} 49.20\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/dce/dce\_stream\_encoder.c

\textbf{Function name 1:} dce110\_stream\_encoder\_stop\_hdmi\_info\_packets

\textbf{File 2:} dc/dio/dcn30/dcn30\_dio\_stream\_encoder.c

\textbf{Function name 2:} enc3\_stream\_encoder\_stop\_hdmi\_info\_packets

\textbf{Similarity:} 50.33\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/dml/dcn32/dcn32\_fpu.c

\textbf{Function name 1:} dcn32\_calculate\_wm\_and\_dlg\_fpu

\textbf{File 2:} dc/dml/dcn30/dcn30\_fpu.c

\textbf{Function name 2:} dcn30\_fpu\_update\_soc\_for\_wm\_a

\textbf{Similarity:} 50.81\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/resource/dcn301/dcn301\_resource.c

\textbf{Function name 1:} dcn301\_dpp\_create

\textbf{File 2:} dc/resource/dcn321/dcn321\_resource.c

\textbf{Function name 2:} dcn321\_dpp\_destroy

\textbf{Similarity:} 50.39\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/gpio/dce60/hw\_translate\_dce60.c

\textbf{Function name 1:} offset\_to\_id

\textbf{File 2:} dc/irq/dcn302/irq\_service\_dcn302.c

\textbf{Function name 2:} to\_dal\_irq\_source\_dcn302

\textbf{Similarity:} 49.33\%

\textbf{Duplication type:} Negative

### 2024-08-03 conteudo/apendices/emp-lista.tex
create appendix

@@
 
-THIS APPENDIX IS FOR LIST THE RANDOM SELECTED FUNCTIONS PAIRS IN THE 
-EMPIRICIAL ANALYSES OF THE TOOL. I will put it here later.
+\label{app:emp}
+
+This appendix includes the list of function pairs used to evaluate the tool, as 
+described in section \ref{sec:metemp}. 
+The results are available in Table \ref{tab:emp}. 
+The files indicate the relative path to the AMD Display driver root directory.
+
+\input{conteudo/apendices/cem}
+
+\input{conteudo/apendices/noventa}
+
+\input{conteudo/apendices/oitenta}
+
+\input{conteudo/apendices/setenta}
+
+\input{conteudo/apendices/sesenta}
+
+\input{conteudo/apendices/cinquenta}
+
+\input{conteudo/apendices/quarenta}
+
+\input{conteudo/apendices/trinta}

### 2024-08-03 conteudo/apendices/noventa.tex
create appendix

\section{Function pairs with similarity between 89\% and 91\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/dce/dce\_opp.c

\textbf{Function name 1:} dce110\_opp\_program\_fmt

\textbf{File 2:} dc/dce/dce\_opp.c

\textbf{Function name 2:} dce60\_opp\_program\_fmt

\textbf{Similarity:} 89.59\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/dpp/dcn10/dcn10\_dpp\_cm.c

\textbf{Function name 1:} dpp1\_cm\_get\_gamut\_remap

\textbf{File 2:} dc/dpp/dcn20/dcn20\_dpp\_cm.c

\textbf{Function name 2:} dpp2\_cm\_get\_gamut\_remap

\textbf{Similarity:} 90.20\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/resource/dcn30/dcn30\_resource.c

\textbf{Function name 1:} dcn30\_hubbub\_create

\textbf{File 2:} dc/resource/dcn316/dcn316\_resource.c

\textbf{Function name 2:} dcn31\_hubbub\_create

\textbf{Similarity:} 89.02\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/resource/dce110/dce110\_resource.c

\textbf{Function name 1:} dce110\_destroy\_resource\_pool

\textbf{File 2:} dc/resource/dce120/dce120\_resource.c

\textbf{Function name 2:} dce120\_destroy\_resource\_pool

\textbf{Similarity:} 90.39\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/resource/dce120/dce120\_resource.c

\textbf{Function name 1:} dce120\_clock\_source\_create

\textbf{File 2:} dc/resource/dcn301/dcn301\_resource.c

\textbf{Function name 2:} dcn301\_clock\_source\_create

\textbf{Similarity:} 90.88\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/dml/dcn32/dcn32\_fpu.c

\textbf{Function name 1:} dcn32\_enough\_pipes\_for\_subvp

\textbf{File 2:} dc/dml2/dml2\_mall\_phantom.c

\textbf{Function name 2:} enough\_pipes\_for\_subvp

\textbf{Similarity:} 90.74\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/pg/dcn35/dcn35\_pg\_cntl.c

\textbf{Function name 1:} pg\_cntl35\_io\_clk\_status

\textbf{File 2:} dc/pg/dcn35/dcn35\_pg\_cntl.c

\textbf{Function name 2:} pg\_cntl35\_mem\_status

\textbf{Similarity:} 90.74\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 1:} dcn20\_disable\_writeback

\textbf{File 2:} dc/hwss/dcn30/dcn30\_hwseq.c

\textbf{Function name 2:} dcn30\_disable\_writeback

\textbf{Similarity:} 89.32\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/dpp/dcn20/dcn20\_dpp\_cm.c

\textbf{Function name 1:} dpp20\_program\_shaper\_luta\_settings

\textbf{File 2:} dc/mpc/dcn32/dcn32\_mpc.c

\textbf{Function name 2:} mpc32\_program\_shaper\_lutb\_settings

\textbf{Similarity:} 89.34\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/clk\_mgr/dcn31/dcn31\_smu.c

\textbf{Function name 1:} dcn31\_smu\_enable\_phy\_refclk\_pwrdwn

\textbf{File 2:} dc/clk\_mgr/dcn314/dcn314\_smu.c

\textbf{Function name 2:} dcn314\_smu\_enable\_phy\_refclk\_pwrdwn

\textbf{Similarity:} 89.03\%

\textbf{Duplication type:} Type-2

### 2024-08-03 conteudo/apendices/quarenta.tex
create appendix

\section{Function pairs with similarity between 39\% and 41\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/dc\_dmub\_srv.c

\textbf{Function name 1:} dc\_dmub\_srv\_is\_restore\_required

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm.c

\textbf{Function name 2:} dm\_dmub\_hw\_resume

\textbf{Similarity:} 39.92\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/hwss/dce110/dce110\_hwseq.c

\textbf{Function name 1:} dce110\_program\_front\_end\_for\_pipe

\textbf{File 2:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 2:} dcn20\_enable\_stream

\textbf{Similarity:} 39.76\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/core/dc\_resource.c

\textbf{Function name 1:} dc\_resource\_find\_first\_free\_pll

\textbf{File 2:} dc/core/dc\_resource.c

\textbf{Function name 2:} find\_free\_hpo\_dp\_link\_enc

\textbf{Similarity:} 40.64\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/hwss/dcn10/dcn10\_hwseq.c

\textbf{Function name 1:} dcn10\_reset\_back\_end\_for\_pipe

\textbf{File 2:} dc/resource/dcn20/dcn20\_resource.c

\textbf{Function name 2:} dcn20\_add\_dsc\_to\_stream\_resource

\textbf{Similarity:} 39.00\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/hubp/dcn20/dcn20\_hubp.c

\textbf{Function name 1:} hubp2\_read\_state\_common

\textbf{File 2:} dc/hubbub/dcn21/dcn21\_hubbub.c

\textbf{Function name 2:} hubbub21\_wm\_read\_state

\textbf{Similarity:} 40.60\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 1:} get\_phyd32clk\_src

\textbf{File 2:} dc/dml/calcs/dcn\_calcs.c

\textbf{Function name 2:} tl\_sw\_mode\_to\_bw\_defs

\textbf{Similarity:} 39.00\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 1:} dcn20\_init\_sys\_ctx

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm.c

\textbf{Function name 2:} mmhub\_read\_system\_context

\textbf{Similarity:} 39.73\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/clk\_mgr/dcn31/dcn31\_clk\_mgr.c

\textbf{Function name 1:} dcn31\_update\_clocks

\textbf{File 2:} dc/clk\_mgr/dcn30/dcn30\_clk\_mgr.c

\textbf{Function name 2:} dcn3\_init\_clocks

\textbf{Similarity:} 39.77\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} modules/power/power\_helpers.c

\textbf{Function name 1:} dmub\_init\_abm\_config

\textbf{File 2:} dc/hwss/dce110/dce110\_hwseq.c

\textbf{Function name 2:} dce110\_is\_hpo\_enabled

\textbf{Similarity:} 39.96\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} modules/hdcp/hdcp\_psp.c

\textbf{Function name 1:} mod\_hdcp\_add\_display\_to\_topology

\textbf{File 2:} modules/hdcp/hdcp\_ddc.c

\textbf{Function name 2:} mod\_hdcp\_read\_ksvlist

\textbf{Similarity:} 39.89\%

\textbf{Duplication type:} Negative

### 2024-08-03 conteudo/apendices/sesenta.tex
create appendix

\section{Function pairs with similarity between 59\% and 61\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/hwss/dcn30/dcn30\_hwseq.c

\textbf{Function name 1:} dcn30\_set\_avmute

\textbf{File 2:} dc/hwss/dcn10/dcn10\_hwseq.c

\textbf{Function name 2:} dcn10\_wait\_for\_pending\_cleared

\textbf{Similarity:} 59.99\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/irq/dcn30/irq\_service\_dcn30.c

\textbf{Function name 1:} to\_dal\_irq\_source\_dcn30

\textbf{File 2:} dc/link/link\_detection.c

\textbf{Function name 2:} get\_basic\_signal\_type

\textbf{Similarity:} 60.97\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/irq/dce60/irq\_service\_dce60.c

\textbf{Function name 1:} hpd\_ack

\textbf{File 2:} dc/irq/irq\_service.c

\textbf{Function name 2:} dal\_irq\_service\_set\_generic

\textbf{Similarity:} 60.92\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/dml/dcn20/dcn20\_fpu.c

\textbf{Function name 1:} dcn20\_populate\_dml\_pipes\_from\_context

\textbf{File 2:} dc/resource/dcn314/dcn314\_resource.c

\textbf{Function name 2:} dcn314\_populate\_dml\_pipes\_from\_context

\textbf{Similarity:} 59.03\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/dccg/dcn401/dcn401\_dccg.c

\textbf{Function name 1:} dccg401\_set\_physymclk

\textbf{File 2:} dc/dccg/dcn35/dcn35\_dccg.c

\textbf{Function name 2:} dccg35\_set\_physymclk

\textbf{Similarity:} 60.43\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/hwss/dcn20/dcn20\_hwseq.c

\textbf{Function name 1:} dcn20\_enable\_plane

\textbf{File 2:} dc/hwss/dcn10/dcn10\_hwseq.c

\textbf{Function name 2:} dcn10\_enable\_plane

\textbf{Similarity:} 59.46\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/dce/dce\_stream\_encoder.c

\textbf{Function name 1:} dce110\_stream\_encoder\_lvds\_set\_stream\_attribute

\textbf{File 2:} dc/dio/dcn10/dcn10\_stream\_encoder.c

\textbf{Function name 2:} enc1\_stream\_encoder\_dvi\_set\_stream\_attribute

\textbf{Similarity:} 60.67\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/dml/dcn31/dcn31\_fpu.c

\textbf{Function name 1:} dcn31\_calculate\_wm\_and\_dlg\_fp

\textbf{File 2:} dc/dml/dcn20/dcn20\_fpu.c

\textbf{Function name 2:} dcn21\_calculate\_wm

\textbf{Similarity:} 59.46\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/resource/dcn10/dcn10\_resource.c

\textbf{Function name 1:} dcn10\_resource\_destruct

\textbf{File 2:} dc/resource/dcn30/dcn30\_resource.c

\textbf{Function name 2:} dcn30\_resource\_construct

\textbf{Similarity:} 60.64\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/clk\_mgr/dcn30/dcn30\_clk\_mgr.c

\textbf{Function name 1:} dcn3\_is\_smu\_present

\textbf{File 2:} dc/clk\_mgr/dcn316/dcn316\_clk\_mgr.c

\textbf{Function name 2:} dcn316\_enable\_pme\_wa

\textbf{Similarity:} 60.54\%

\textbf{Duplication type:} Negative

### 2024-08-03 conteudo/apendices/trinta.tex
create appendix

\section{Function pairs with similarity between 29\% and 31\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/core/dc\_state.c

\textbf{Function name 1:} dc\_state\_add\_all\_planes\_for\_stream

\textbf{File 2:} dc/resource/dcn20/dcn20\_resource.c

\textbf{Function name 2:} dcn20\_add\_dsc\_to\_stream\_resource

\textbf{Similarity:} 29.14\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/gpio/dcn21/hw\_factory\_dcn21.c

\textbf{Function name 1:} define\_ddc\_registers

\textbf{File 2:} dc/link/protocols/link\_ddc.c

\textbf{Function name 2:} ddc\_service\_destruct

\textbf{Similarity:} 30.89\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/core/dc\_resource.c

\textbf{Function name 1:} resource\_find\_free\_secondary\_pipe\_legacy

\textbf{File 2:} dc/dml/dcn32/dcn32\_fpu.c

\textbf{Function name 2:} update\_pipe\_slice\_table\_with\_split\_flags

\textbf{Similarity:} 29.61\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/dml2/dml21/src/dml2\_core/dml2\_core\_dcn4.c

\textbf{Function name 1:} lookup\_uclk\_dpm\_index\_by\_freq

\textbf{File 2:} dc/dml2/dml2\_dc\_resource\_mgmt.c

\textbf{Function name 2:} is\_pipe\_in\_candidate\_array

\textbf{Similarity:} 30.65\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/resource/dce112/dce112\_resource.c

\textbf{Function name 1:} dce112\_resource\_destruct

\textbf{File 2:} dc/resource/dcn20/dcn20\_resource.c

\textbf{Function name 2:} dcn20\_acquire\_dsc

\textbf{Similarity:} 30.99\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 6 }

\

\textbf{File 1:} modules/hdcp/hdcp\_ddc.c

\textbf{Function name 1:} mod\_hdcp\_write\_no\_stored\_km

\textbf{File 2:} modules/hdcp/hdcp\_ddc.c

\textbf{Function name 2:} mod\_hdcp\_write\_ainfo

\textbf{Similarity:} 29.82\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/dml2/dml2\_mall\_phantom.c

\textbf{Function name 1:} all\_pipes\_have\_stream\_and\_plane

\textbf{File 2:} dc/basics/dce\_calcs.c

\textbf{Function name 2:} all\_displays\_in\_sync

\textbf{Similarity:} 30.59\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/resource/dcn302/dcn302\_resource.c

\textbf{Function name 1:} dcn302\_resource\_destruct

\textbf{File 2:} dc/core/dc.c

\textbf{Function name 2:} dc\_resume

\textbf{Similarity:} 29.97\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/optc/dcn314/dcn314\_optc.c

\textbf{Function name 1:} optc314\_enable\_crtc

\textbf{File 2:} dc/optc/dcn32/dcn32\_optc.c

\textbf{Function name 2:} optc32\_disable\_phantom\_otg

\textbf{Similarity:} 29.97\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/irq/irq\_service.c

\textbf{Function name 1:} dal\_irq\_service\_ack

\textbf{File 2:} dc/irq/irq\_service.c

\textbf{Function name 2:} find\_irq\_source\_info

\textbf{Similarity:} 30.65\%

\textbf{Duplication type:} Negative

### 2024-08-03 conteudo/methods/methods.tex
add initial preliminar
add refer to repository
create appendix

@@
 our tool. The random selected function selected function pairs can 
-be found in the Appendix (CREATE Appendix AND ADD HERE).
+be found in the Appendix \ref{app:emp}.
 
@@
 \section{Ethnographic study}
+
+\label{sec:meteth}
 
@@
 that we do not know if it is possible to refactor the duplication. With or without successful 
-mitigation, we will store our mitigations on (ADD REP AND PUT HERE) to research transparency.
+mitigation, we will store our mitigations on \url{https://github.com/LipArcanjo/master-artifacts} 
+to research transparency.
 

### 2024-08-03 conteudo/methods/methods.tex
change table to appendix

@@
 
-(MAYBE I SHOULD ADD A RESTRICTION ABOUT IGNORING AUTOMATED GENERATED CODE, such as DML folder)
-
-The function pairs selected randomly following our methodology can be consulted in the Table
-(CREATE TABLE AND ADD HERE).
-
-CREATE TABLE AND ADD HERE.
+The function pairs selected randomly following our methodology can be consulted in the Appendix
+(CREATE Appendix AND ADD HERE).
 

### 2024-08-04 conteudo/background/refactor.tex
add refactor methods

@@
 
-\subsection{Parameterize method}
+\paragraph{Parameterize method}
 
-TODO. I EXPECT TO FILL THIS SECTION AS THE WORK EVOLVES.
+\
+
+The Parameterize method consists of replacing multiple similar codes with a singular function. 
+The function created implements the different behaviors between similar codes by using parameters, 
+which tell which behavior the function should use \citep{refactorbook}.
+Figure (CREATE FIGURE AND ADD HERE) demonstrates an example of the Parameterize method usage.
+
+\paragraph{Extract method}
+
+\
+
+The Extract method consists of splitting a complex code artifact into multiple simple methods/functions. 
+One of the advantages of this method is enabling the created functions to be used multiple times through 
+the code project. In contrast, with only the complex code artifact, it is not always possible to reuse 
+it entirely, thus creating a new code duplication \citep{refactorbook}. 
+Figure (CREATE FIGURE AND ADD HERE) demonstrates an example of the Extract method usage.

### 2024-08-04 conteudo/background/refactor.tex
add refactor examples

@@
 which tell which behavior the function should use \citep{refactorbook}.
-Figure (CREATE FIGURE AND ADD HERE) demonstrates an example of the Parameterize method usage.
+Figure \ref{fig:parameter} demonstrates an example of the Parameterize method usage.
+
+\begin{figure}
+\includegraphics[scale=0.36]{refactor/parameterize}
+\caption{Example of Parameterize method usage.}
+\label{fig:parameter}
+\end{figure}
 
@@
 it entirely, thus creating a new code duplication \citep{refactorbook}. 
-Figure (CREATE FIGURE AND ADD HERE) demonstrates an example of the Extract method usage.
+Figure \ref{fig:extract} demonstrates an example of the Extract method usage.
+
+\begin{figure}
+\includegraphics[scale=0.36]{refactor/extract}
+\caption{Example of Extract method usage.}
+\label{fig:extract}
+\end{figure}

### 2024-08-04 figuras/refactor/extract.cpp

@@ -1,0 +1,38 @@
+void sort_and_print(vector<int> a){
+	int n = a.size();
+	for(int i = 0; i < n; i++){
+		for(int j = 0; j < n-1; j++){
+			if(a[j] > a[j+1]){
+				swap(a[j],a[j+1]);
+			}
+		}
+	}
+	for(int i = 0; i < n; i++){
+		printf("%d ",&a[i]);
+	}
+}
+
+vector<int> sort_increasing(vector<int> a){
+	for(int i = 0; i < n; i++){
+		for(int j = 0; j < n-1; j++){
+			if(a[j] > a[j+1]){
+				swap(a[j],a[j+1]);
+			}
+		}
+	}
+	return a;
+}
+
+void print(vector<int> a){
+	int n = a.size();
+	for(int i = 0; i < n; i++){
+		printf("%d ",&a[i]);
+	}
+}
+
+void sort_and_print(vector<int> a){
+	a = sort_inscreasing(a);
+	print(a);
+}
+
+

### 2024-08-04 figuras/refactor/parameterize.cpp

@@ -1,0 +1,38 @@
+vector<int> sort_increasing(vector<int> a){
+	int n = a.size();
+	for(int i = 0; i < n; i++){
+		for(int j = 0; j < n-1; j++){
+			if(a[j] > a[j+1]){
+				swap(a[j],a[j+1]);
+			}
+		}
+	}
+	return a;
+}
+
+vector<int> sort_decreasing(vector<int> a){
+	int n = a.size();
+	for(int i = 0; i < n; i++){
+		for(int j = 0; j < n-1; j++){
+			if(a[j] < a[j+1]){
+				swap(a[j],a[j+1]);
+			}
+		}
+	}
+	return a;
+}
+
+vector<int> sort(
+	vector<int> a, 
+	function<bool(int,int)> cmp
+){
+	int n = a.size();
+	for(int i = 0; i < n; i++){
+		for(int j = 0; j < n-1; j++){
+			if(cmp(a[j],a[j+1]){
+				swap(a[j],a[j+1]);
+			}
+		}
+	}
+	return a;
+}

### 2024-08-04 conteudo/preliminar/results.tex
add table
add base cro because of error

@@
 proposed systematic approach to mitigate code duplication on two function pairs. Table
-(CREATE TABLE AND ADD HERE) shows the early results found.
+\ref{tab:patch} shows the early results found.
 
-CREATE TABLE AND ADD HERE
+\input{conteudo/preliminar/table_patch}
 
@@
 
-As shown in our proposed schedule (Figure CREATE FIGURE AND ADD HERE), as we advance, 
+As shown in our proposed schedule \ref{fig:cronograma}, as we advance, 
 we intend to do the proper ethnographic study, starting by completing our systematic 
@@
 AMD Display driver maintainers about the mitigations made by us.
+
+\input{conteudo/preliminar/cronograma}

### 2024-09-09 conteudo/resumoabstract.tex
create resumo and remove appendix as it is unnecesary at the moment

%!TeX root=../tese.tex
%("dica" para o editor de texto: este arquivo é parte de um documento maior)
% para saber mais: https://tex.stackexchange.com/q/78101

% As palavras-chave são obrigatórias, em português e em inglês, e devem ser
% definidas antes do resumo/abstract. Acrescente quantas forem necessárias.
\palavraschave{Linux, Duplicação de código, ferramenta}

\keywords{Linux,Code duplication,Tool}

% O resumo é obrigatório, em português e inglês. Estes comandos também
% geram automaticamente a referência para o próprio documento, conforme
% as normas sugeridas da USP.
\resumo{

O kernel Linux é um projeto de código aberto amplamente utilizado, alimentando 3,77\% 
dos desktops pessoais e 70,69\% dos smartphones em todo o mundo. Dada a sua extensa 
base de usuários, qualquer adição de recursos, correção de bugs ou tratamento de 
vulnerabilidades de segurança pode impactar milhões de pessoas. Manter o kernel é 
uma tarefa enorme, envolvendo mais de 19 milhões de linhas de código e contribuições 
de mais de mil desenvolvedores. Devido à escala do projeto, nem todas as contribuições 
seguem as melhores práticas, resultando em artefatos de código de baixa qualidade que 
podem complicar a manutenção e o desenvolvimento de novos recursos. Um desses artefatos 
é a duplicação de código, que é o foco desta pesquisa. As ferramentas existentes 
para detectar duplicação de código normalmente comparam dois artefatos de código para 
determinar se são duplicatas. No entanto, essas ferramentas não são adequadas para 
identificar duplicações em bases de código em grande escala, como o kernel Linux, nem 
oferecem orientações sobre como resolver as duplicações detectadas. Apesar das 
pesquisas tanto na literatura formal quanto na cinza, as soluções existentes não 
conseguem resolver os desafios específicos do kernel Linux. Para abordar esse tema, 
esta pesquisa foca em dois objetivos principais. Primeiro, propõe uma nova ferramenta 
para identificar duplicações de código dentro do kernel Linux. Segundo, analisa as 
duplicações identificadas para extrair padrões e métodos de refatoração, oferecendo 
orientações para futuros contribuintes.

}

\abstract{

The Linux kernel is a widely used open-source project, powering 3.77\% of personal desktops 
and 70.69\% of smartphones worldwide. Given its extensive user base, any addition of features, 
bug fixes, or handling of security vulnerabilities can impact millions. Maintaining the kernel 
is an enormous task, involving more than 19 million lines of code and contributions from over 
thirteen thousand developers. Due to the scale of the project, not all contributions adhere to best 
practices, resulting in poor-quality code artifacts that may complicate maintenance and feature 
development. One such artifact is code duplication, which is the focus of this research. 
Existing tools for detecting code duplication typically compare two code artifacts to determine 
if they are duplicates. However, these tools are not well-suited for identifying duplications 
in large-scale codebases like the Linux kernel, nor do they guide on resolving detected 
duplications. Despite searches in both formal and gray literature, existing solutions fall 
short of addressing the specific challenges of the Linux kernel. To address this topic, this 
research focuses on two key goals. First, it proposes a new tool to identify code duplications 
within the Linux kernel. Second, it analyzes the identified duplications to extract patterns 
and refactoring methods, offering guidance to future contributors.

}

### 2024-09-10 conteudo/introduction/intro.tex
fazendo as mudancas de acordo com as sugestoes

@@
 
-Representing 3.77\% of personal desktops and 70,69\% of smartphones worldwide, the Linux kernel is an open-source 
+Representing 96,3\% of the top one million web servers and 85\% of smartphones worldwide, the Linux kernel is an open-source 
 project providing value to billions of people daily. The magnitude that the project has reached over the years makes 
 the addition of features, as well as handling bugs and security vulnerabilities, potentially affect a significant 
-number of people, demonstrating the importance of maintaining the kernel.
+number of people, demonstrating the importance of maintaining the kernel \citep{linuxdata}.
 

### 2024-09-10 conteudo/resumoabstract.tex
fazendo as mudancas de acordo com as sugestoes

@@
 
-O kernel Linux é um projeto de código aberto amplamente utilizado, alimentando 3,77\% 
-dos desktops pessoais e 70,69\% dos smartphones em todo o mundo. Dada a sua extensa 
+O kernel Linux é um projeto de software livre amplamente utilizado, alimentando 96,3\% 
+dos um milhão maiores servidores web e 85\% dos smartphones em todo o mundo. Dada a sua extensa 
 base de usuários, qualquer adição de recursos, correção de bugs ou tratamento de 
@@
 duplicações identificadas para extrair padrões e métodos de refatoração, oferecendo 
-orientações para futuros contribuintes.
+orientações para futuros contribuidores.
 
@@
 
-The Linux kernel is a widely used open-source project, powering 3.77\% of personal desktops 
-and 70.69\% of smartphones worldwide. Given its extensive user base, any addition of features, 
+The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, powering
+96.3\% of the top one million web servers 
+and 85\% of smartphones worldwide. Given its extensive user base, any addition of features, 
 bug fixes, or handling of security vulnerabilities can impact millions. Maintaining the kernel 

### 2024-10-28 conteudo/introduction/intro.tex
Reviewing introduction

--- 2024-08-03 conteudo/introduction/intro.tex
@@
 
-Representing 3.77\% of personal desktops and 70,69\% of smartphones worldwide, the Linux kernel is an open-source 
-project providing value to billions of people daily. The magnitude that the project has reached over the years makes 
-the addition of features, as well as handling bugs and security vulnerabilities, potentially affect a significant 
-number of people, demonstrating the importance of maintaining the kernel.
+The Linux kernel is a Free/Libre/Open-Source project providing value to
+billions of people daily, representing 96.3\% of the top one million web
+servers and 85\% of smartphones worldwide. The magnitude that the project has
+reached over the years makes the addition of features, as well as handling bugs
+and security vulnerabilities, potentially affect a significant number of
+people, demonstrating the importance of maintaining the kernel
+\citep{linuxdata}.
 
-Maintaining the Linux kernel is a complex task, as the project contains more than 19 million 
-lines of code and has more than thirteen thousand contributors along the project's lifespan
-\citep{linuxquantity}.
-Given the magnitude of the project, it is unrealistic to expect that all contributors 
-followed the best programming practices and principles throughout every contribution they made to the project.
-It is expected to find a good amount of bad-quality code artifacts in the source code, 
-which could result in hindering or compromising the maintenance of this project, as well as the 
-addition of new features for it.
+Maintaining the Linux kernel is a complex task, as the project contains more
+than 19 million lines of code and has had over thirteen thousand contributors
+throughout its lifespan \citep{linuxquantity}. Given the magnitude of the Linux
+kernel project, it is unrealistic to expect that all contributors have
+consistently followed the best programming practices and principles in every
+contribution made to the project. Consequently, it is expected that a
+significant amount of low-quality code artifacts will be found in the source
+code, which could hinder or compromise the maintenance of the project and the
+addition of new features.
 
-Device drivers, as stated by Madieu, are a "piece of software whose aim is to control and manage
-a particular hardware device, hence the name device driver"
-\citep{driverdef}. 
-Device drivers are a major part of the Linux kernel, representing 66 \% of the source code, and 
-are responsible for the management of crucial elements such as keyboards, mice, and GPU
-\citep{marcelo}. 
-The moment the kernel chooses to support new hardware, it is required to make alterations to existent 
-device drivers or the creation of new ones, this makes the maintainment of the device drivers 
-an important duty of the Linux community.
+As stated by \citet{driverdef}, device drivers are a ``piece of software whose
+aim is to control and manage a particular hardware device, hence the name
+device driver.'' Device drivers are a major part of the Linux kernel,
+representing 66\% of the source code, and are responsible for managing
+essential elements such as keyboards, mice, and GPUs \citep{marcelo}. When the
+kernel opts to support new hardware, it necessitates alterations to existing
+device drivers or the creation of new ones, making maintaining device drivers a
+critical duty of the Linux community.
 
-The AMD Display driver of the Linux kernel is the driver responsible for enabling AMD GPUs to operate 
-correctly in desktop Linux environments, which is an important duty, given that AMD GPU represented 19 \% 
-of the personal GPU market in 2023  \citep{gpumarket} . 
-When we interacted with the maintainers of the AMD Display driver, they shared with us their pain with one 
-of the characteristics of bad quality code artifacts, which are duplicated code artifacts, which hinder 
-the maintainment of the driver.
+The AMD Display driver in the Linux kernel is responsible for enabling AMD GPUs
+to operate correctly in desktop Linux environments, which is an essential task,
+given that AMD GPUs represented 19\% of the personal GPU market in 2023
+\citep{gpumarket}. When we interacted with the maintainers of the AMD Display
+driver, they shared with us their challenges regarding one of the
+characteristics of low-quality code artifacts: duplicated code, which hampers
+the maintenance of the driver.
 
-Searching for solutions to the AMD Display driver problem with duplicated code artifacts, we were 
-able to find tools that gave two code artifacts, infer if they are duplications of each other, and the 
-formal literature in the code duplication mainly focused on optimizing the accuracy of these tools in this 
-given task. These solutions do not address our problem, as we are not able to give them a codebase and 
-they return to us the code duplication in the codebase, and they do not guide us to how to mitigate the code 
-duplications after we detected them, requiring us to search for others alternatives.
+In searching for solutions to the AMD Display driver problem related to
+duplicated code artifacts, we found tools capable of determining, given two
+code artifacts, whether they are duplications of each other. However, the
+formal literature on code duplication mainly focuses on optimizing the accuracy
+of these tools for this specific task. These solutions do not address our
+problem, as we cannot provide a complete codebase and have the tools return the
+code duplications within it. Moreover, they do not guide us on how to mitigate
+the code duplications after detection, necessitating the exploration of
+alternative solutions.
 
-Given the unsuccessful results of our search in the formal literature, we started to search for solutions in the 
-gray literature, but we ended up without success. In the gray literature, the best solutions we found were 
-primitive solutions that find every pair of code files that are duplications. 
-The code files on the AMD Display driver that 
-contain duplicated code normally also have specific code that is only implemented in this file. 
-Thus, a solution that finds duplication in a deeper context is necessary to find correct duplications 
-in the Linux kernel context.
+Given the unsuccessful results of our search in the formal literature, we
+turned to gray literature for solutions but were also met with limited success.
+The best solutions we found in the gray literature were primitive, merely
+identifying every pair of code files that are duplications. The code files in
+the AMD Display driver containing duplicated code typically have specific code
+unique to each file. Therefore, a solution that identifies duplication in a
+deeper context is necessary to find accurate duplications within the Linux
+kernel context.
 
@@
 
-The code duplication problem outlined in the Section above cannot be resolved by the solutions 
-presented in the formal literature or the gray literature, we intend to propose an approach to 
-identify and mitigate the code duplications in the Linux Kernel. For that, we propose a tool capable 
-of identifying duplicated functions in the kernel, and we intend to mitigate code duplications in the 
-Linux Kernel.
+According to our studies and observations, the code duplication problem
+outlined in the section above cannot be resolved by the solutions presented in
+the formal or gray literature. To address this gap, we propose an approach to
+identify and mitigate code duplications in the Linux kernel. Specifically, we
+aim to develop a tool capable of detecting duplicated functions within the
+kernel and establish strategies for mitigating these duplications.
 
-In this research, we expect to extract patterns of software engineers' techniques, such as refactoring methods, 
-which we commonly use to mitigate code duplications, thus, serving as guidance to future contributors to 
-approach the code duplication problem in the given context. To properly investigate our proposed approach,
-we developed our primary research question below to guide our research.
+In this research, we expect to extract patterns of software engineering
+techniques, such as refactoring methods commonly used to mitigate code
+duplications. These patterns will guide future contributors in tackling the
+code duplication problem in the given context. To properly investigate our
+proposed approach, we developed the following primary research question to
+guide our study:
 
-\textbf{RQ1:} How effective is our approach in mitigating duplicate code in the Linux kernel?
+\textbf{RQ1:} What are effective approaches for identifying duplicate code in the
+Linux kernel?
 
-Given the complexity of the proposed question, we created a supplementary research question to verify midway 
-results, which is related to the performance of our proposed tool in the Linux kernel, the supplementary research question
-is defined as:
+Given the complexity of the primary question, we formulated a supplementary
+research question to assess midway results related to the performance of our
+proposed tool in the Linux kernel. The supplementary research question is
+defined as follows:
 
-\textbf{RQ1.1:} How accurate is the proposed tool for detecting duplicate functions?
+\textbf{RQ1.1:} How accurately does an approach detect duplicated functions
+within the Linux kernel?
 
-To address our research question, we do a multiple-method research utilizing different methodologies throughout 
-the study. We structured this work into two phases, each with distinct research methods.
+We adopt a multi-method research design utilizing various methodologies
+throughout the study to address our research questions. We structured this work
+into two phases, each employing distinct research methods.
 
-In \textbf{Phase I}, we will propose, build, and validate a tool capable 
-of identifying code duplications at the function level for the C programming language. We expect this proposed tool 
-to identify code duplications in the Linux kernel, as in our preliminary investigation of the code files in the kernel 
-is commonly a union of duplicated and nonduplicated functions. To validate the tool, we will do a triangulation of results 
-by evaluating our tool in the code duplication databases used in the formal literature and doing an empirical analysis 
-in a set of randomly selected codes of the AMD Display driver which our proposed tool claims is a duplication. 
-A detailed explication of how we will validate our tool is demonstrated in Chapter \ref{cha:method}.
+In \textbf{Phase I}, we will propose, develop, and validate a tool capable of
+identifying code duplications at the function level for the C programming
+language. We expect this tool to effectively identify code duplications in the
+Linux kernel, as our preliminary investigation suggests that code files in the
+kernel often contain a mix of duplicated and non-duplicated functions. To
+validate the tool, we will perform a triangulation of results by evaluating our
+tool against code duplication databases used in the formal literature and by
+conducting an empirical analysis on a randomly selected set of code files from
+the AMD Display driver, which our tool identifies as containing duplications.
+Chapter \ref{cha:method} provides a detailed explanation of our validation
+process.
 
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[scale=0.7]{research_design}
@@
 
-Given the first phase completion, we would have a validated tool to identify code duplications in the Linux kernel 
-context and, we can move on to how to mitigate these duplications, which comprehend the second phase of our research.
-In \textbf{Phase 2}, we intend to discover ways to mitigate the code duplication detected in the Linux kernel. 
-To achieve our objective, we designed an ethnographic approach, which enables us to interact with the Linux community 
-to collect artifacts validating or refuting our mitigation.
+Upon completing the first phase, we will have a validated tool for identifying
+code duplications within the Linux kernel context. We can then proceed to the
+second phase of our research, which focuses on mitigating these duplications.
+In \textbf{Phase II}, we aim to discover practical strategies to address the
+code duplications detected in the Linux kernel. To achieve this objective, we
+have designed an ethnographic approach that allows us to interact with the
+Linux community and collect artifacts that validate or refute our proposed
+mitigation strategies.
 
-Figure \ref{fig:reDesign} sumamarizes the research steps proposed in this work.
+Figure \ref{fig:reDesign} summarizes the research steps proposed in this work.
 
@@
 
-This manuscript consists of five more chapters.
-Chapter \ref{cha:back} presents the literature overview for code duplication detection (main definitions, 
-current approaches in the literature), a brief description of the Linux kernel, and a review of refactoring methods 
-used throughout this research.
-Chapter \ref{cha:tool} presents our proposed tool to detect code duplication in the Linux kernel, describing all 
-the main components.
-Chapter \ref{cha:method} describes the research methods selected to guide our work.
-Chapter \ref{cha:eval} shows the results obtained of the research methods to evaluate our proposed tool.
-Chapter \ref{cha:results} presents the current research stage and the work plan, with the addition of 
-preliminary results obtained.
+This manuscript consists of five more chapters. Chapter \ref{cha:back} presents
+the literature overview for code duplication detection (main definitions,
+current approaches in the literature), a brief description of the Linux kernel,
+and a review of refactoring methods used throughout this research. Chapter
+\ref{cha:tool} presents our proposed tool to detect code duplication in the
+Linux kernel, describing all the main components.  Chapter \ref{cha:method}
+describes the research methods selected to guide our work. Chapter
+\ref{cha:eval} shows the results obtained from the research methods to evaluate
+our proposed tool. Chapter \ref{cha:results} presents the current research
+stage and the work plan, with the addition of preliminary results obtained.

### 2024-11-08 conteudo/background/linux.tex
create tables of tool
Reviewing background

--- 2024-07-11 conteudo/background/linux.tex
@@
 \en
+
+\todo[inline]{Adicionar um "preambulo" antes da seções do capítulo}
 
@@
 
-The Linux kernel is an Unix-like operating system published by Linus Torvalds in 1992 as a free software and
-it started becoming popular in late 1990s \citep{linuxbook}. Against popular belief, linux is not a full 
-operating system as it do not includes applications such as filesystem utilities, windowing systems. The "Linux" 
-distributions know to the public are in fact GNU/Linux operating systems \citep{gnuref}.
+The Linux kernel, a Unix-like operating system kernel, was published by Linus
+Torvalds in 1992 as free software. It began gaining popularity in the late
+1990s \citep{linuxbook}. Contrary to popular belief, Linux itself is not a
+complete operating system, as it lacks components like filesystem utilities and
+windowing systems. The so-called ``Linux distributions'' that are commonly
+recognized by the public are actually GNU/Linux operating systems, which
+include various essential applications alongside the Linux kernel
+\citep{gnuref}.
 
-The Linux kernel is monolithic organized as a set of subsystem such as process scheduler, memory management, 
-device driver infrastruture, networking and filesystems \citep{melissa}.
-Each subsystem usually have a maintainer or a group of maintainers, which are people responsible 
-for managing and accepting contributions into the subsystem they care of \citep{melissa}.
+The Linux kernel is monolithic and organized into a set of subsystems, such as
+the process scheduler, memory management, device driver infrastructure,
+networking, and filesystems \citep{melissa}. Each subsystem typically has a
+dedicated maintainer or a team of maintainers responsible for overseeing its
+development and managing contributions \citep{melissa}.
 
-The contributions to the Linux are coordinate with the usage of the Git source management tool. The contributions 
-made to the project are formated as patches, which are text documents describing differences between two different 
-versions of a source tree  The patches are sent and reviewed through the Linux mailing lists \citep{melissa}.
+Development contributions to the Linux kernel are coordinated using the Git
+version control system. These contributions are formatted as patches—text
+documents that outline the differences between two versions of the source code.
+These patches are then submitted and reviewed via the Linux mailing lists
+\citep{melissa}.
 
-Mailing lists are the official means of communication of the Linux community, through the mailing lists the community members 
-send and review
-contributions to the kernel, discuss topics related to the Linux such as which new features to add, et cetera. 
-As there are multiples
-subsystems in the Linux kernel, there are multiples mailing lists in the Linux to organize the discussion made. Every subsystem 
-have their own mailing list to concentrate the discussion related to the subsystem \citep{melissa}. The iteractions made in the
-mailing lists are documented and stored in the Linux kernel Lore \citep{linuxlore}, which serves as a repository of the
-discussion made in the Linux kernel to be consulted by anyone interested.
+Mailing lists are the primary means of communication within the Linux
+community. Members of the community use them to submit and review patches,
+discuss Linux-related topics, and debate potential new features. Given the
+kernel's extensive modularity, each subsystem has its mailing list, allowing
+discussions to be organized around specific areas of development
+\citep{melissa}. All interactions on these mailing lists are documented and
+archived in the Linux Kernel Lore \citep{linuxlore}, which serves as a
+repository for kernel discussions and can be accessed by anyone interested.
 
-One the the Linux kernel subsystem is the AMD Display Driver, a subsystem responsible for implement the drivers required to enable
-AMD GPU's work properly in the Linux environment. The maintainers of this subsystem reported a empirical problem they have in
-the subsystem, which is a significant amount of duplicated code in the subsystem, a problem that hinders the maintainment of the
-subsystem. This reported problem is the starting point that initiated our research, making the AMD Display Driver our principal
-objecte of study to undertand how to mitigate duplicated code in the Linux kernel.
-
-(I WANT TO CHANGE THE AMD DISPLAY paragraph, WHY DO I NEED TO CHANGE IT? I SHOULD ANOTATE WHAT I WANTED, I MADE A CHANGE
- TO MAKE IT BETTER BUT NOT SURE IF THERE IS SOME INFORMATION I WANTED TO PUT IN IT)
+One of the subsystems within the Linux kernel is the AMD Display Driver, which
+is responsible for implementing the drivers needed for AMD GPUs to function
+correctly in a Linux environment. The maintainers of this subsystem have
+reported a practical issue: a significant amount of duplicated code, which
+complicates the maintenance of this subsystem. This problem serves as the
+starting point for our research, with the AMD Display Driver as our primary
+object of study as we aim to understand and mitigate code duplication in the
+Linux kernel. The AMD Display Driver subsystem contains X lines of code, Y
+functions/structs, and Z code files, making it a considerable component within
+the Linux kernel. Like other subsystems, discussions about the AMD Display
+Driver are conducted on its designated mailing list (FIND MAILING LIST AND CITE
+HERE).

### 2024-11-08 conteudo/background/nlpsimilarity.tex
Reviewing background

@@
 
-\section{Text similarity detection methods}
-\label{sec:similarity}.
+\section{Text Similarity Detection Methods}
+\label{sec:similarity}
 
-The standard methodology for text similarity detection is grounded on building vector embeddings of 
-the texts and applying a distance function in these embeddings to measure the similarity between them. 
-There are multiple vector embedding methods proposed by the literature, such as bag-of-words, TF-IDF, 
-LSI, and Word2Vec \citep{gensimlivro}.
-Recent works related to building vector embeddings use Large Language Models (LLM), for example,
-the work of Ethayarajh, which uses BERT, ELMo, and GPT \citep{llmsimilar}. 
-The most popular distance function used by the literature is the cosine similarity defined by the formula
+The standard methodology for text similarity detection is based on building
+vector embeddings of texts and applying a distance function to these embeddings
+to measure the similarity between them. The literature proposes multiple vector
+embedding methods, such as bag-of-words, TF-IDF, LSI, and Word2Vec
+\citep{gensimlivro}. Recent works related to vector embedding construction
+utilize Large Language Models (LLMs), as exemplified by Ethayarajh's work,
+which employs BERT, ELMo, and GPT \citep{llmsimilar}. The most popular distance
+function used in the literature is cosine similarity, defined by the formula
 \citep{cosineref}:
@@
 
-Where $A$ and $B$ are the vector embeddings of the two texts, an advantage of the cosine similarity is that $SC(A,B)$ is
-a number between $0$ and $1$, where $SC(A,B)$ values closer to the number $1$ means that two vectors are more similar, 
-it is usual to compare the cosine similarity to approximation to a probability value, transforming the values between 
-$0$ and $1$ to a percentage between $0 \% $ and $100\%$, through our research, we will use the percentage form.
+where $A$ and $B$ are the vector embeddings of the two texts. An advantage of
+cosine similarity is that $SC(A,B)$ produces a number between $0$ and $1$,
+where values closer to $1$ indicate higher similarity. It is common to
+interpret cosine similarity as an approximation of probability, transforming
+values between $0$ and $1$ into percentages between $0\%$ and $100\%$.
+Throughout our research, we will use this percentage form.
 
-It is a famous fact that LLM has a high computational cost. An example of the high computational cost is an analysis 
-done by Reimers and Gurevych to find the most similar pair in a collection of 10,000 sentences, 
-which required about 65 hours with BERT \citep{bertsimilar}. 
-We want to compare a similar magnitude of pairs in our work, so we decided not to use LLM for our vector embedding 
-method, even though it is a valid solution for smaller codebases.
+It is well-known that LLMs have a high computational cost. For instance, an
+analysis by Reimers and Gurevych found that finding the most similar pair in a
+collection of 10,000 sentences required about 65 hours with BERT
+\citep{bertsimilar}. Since we aim to compare a similar magnitude of pairs in
+our work, we decided not to use LLMs for our vector embedding method, even
+though they may be suitable for smaller codebases.
 
@@
 
-Gensim is an open-source Python library \citep{gensim} which the owners declare themselves as "The fastest library 
-for training of vector embeddings - Python or otherwise. The core algorithms in Gensim use battle-hardened, highly 
-optimized and parallelized C routines." \citep{gensimsite}. For the velocity of Gensim, we chose to use this library 
-for our vector embedding method.
+Gensim is an open-source Python library \citep{gensim}, which its creators
+describe as "The fastest library for training of vector embeddings - Python or
+otherwise. The core algorithms in Gensim use battle-hardened, highly optimized,
+and parallelized C routines." \citep{gensimsite}. Due to Gensim's speed, we
+chose this library for our vector embedding method.
 
-In this work, we will not explore more alternatives for vector embeddings. We do not want to optimize the
-computational cost or the code clone detection to the state of the art as they are not the main objectives 
-of our research.
+In this work, we do not explore further alternatives for vector embeddings, as
+optimizing computational cost or achieving state-of-the-art code clone
+detection is not the primary objective of our research.

### 2024-11-09 conteudo/introduction/intro.tex
Removing research questions

@@
 
-According to our studies and observations, the code duplication problem
-outlined in the section above cannot be resolved by the solutions presented in
-the formal or gray literature. To address this gap, we propose an approach to
-identify and mitigate code duplications in the Linux kernel. Specifically, we
-aim to develop a tool capable of detecting duplicated functions within the
-kernel and establish strategies for mitigating these duplications.
-
-In this research, we expect to extract patterns of software engineering
-techniques, such as refactoring methods commonly used to mitigate code
-duplications. These patterns will guide future contributors in tackling the
-code duplication problem in the given context. To properly investigate our
-proposed approach, we developed the following primary research question to
-guide our study:
-
-\textbf{RQ1:} What are effective approaches for identifying duplicate code in the
-Linux kernel?
-
-Given the complexity of the primary question, we formulated a supplementary
-research question to assess midway results related to the performance of our
-proposed tool in the Linux kernel. The supplementary research question is
-defined as follows:
-
-\textbf{RQ1.1:} How accurately does an approach detect duplicated functions
-within the Linux kernel?
-
-We adopt a multi-method research design utilizing various methodologies
-throughout the study to address our research questions. We structured this work
-into two phases, each employing distinct research methods.
-
-In \textbf{Phase I}, we will propose, develop, and validate a tool capable of
-identifying code duplications at the function level for the C programming
-language. We expect this tool to effectively identify code duplications in the
-Linux kernel, as our preliminary investigation suggests that code files in the
-kernel often contain a mix of duplicated and non-duplicated functions. To
-validate the tool, we will perform a triangulation of results by evaluating our
-tool against code duplication databases used in the formal literature and by
-conducting an empirical analysis on a randomly selected set of code files from
-the AMD Display driver, which our tool identifies as containing duplications.
-Chapter \ref{cha:method} provides a detailed explanation of our validation
-process.
+Our studies and observations reveal that the code duplication problem described
+in the previous section cannot be fully addressed by existing tools in
+either the formal or gray literature. To fill this gap, we propose an approach
+to identify and mitigate code duplications specifically within the Linux
+kernel. Our approach includes the development of a tool capable of detecting
+duplicated functions within the kernel and establishing strategies for
+addressing these duplications.
 
@@
 
-Upon completing the first phase, we will have a validated tool for identifying
-code duplications within the Linux kernel context. We can then proceed to the
-second phase of our research, which focuses on mitigating these duplications.
-In \textbf{Phase II}, we aim to discover practical strategies to address the
-code duplications detected in the Linux kernel. To achieve this objective, we
-have designed an ethnographic approach that allows us to interact with the
-Linux community and collect artifacts that validate or refute our proposed
-mitigation strategies.
+This research aims to identify patterns of software engineering techniques,
+such as refactoring methods commonly used to address code duplications. These
+patterns will serve as a guide for future contributors in mitigating code
+duplication issues within this context. To investigate our approach, we use the
+initial version of our tool to identify duplicate code in the Linux kernel,
+specifically targeting parts of the AMD Display driver code. We will then
+refactor portions of this code and submit patches to maintainers, providing
+evidence of the tool's accuracy in detecting duplicate functions.
 
-Figure \ref{fig:reDesign} summarizes the research steps proposed in this work.
+Our research design follows a multi-method approach, utilizing various
+methodologies to answer our research questions. We have structured the study
+into two phases, each employing different research methods.
+
+In \textbf{Phase I}, we propose, develop, and validate a tool capable of
+identifying code duplications at the function level within C codebases. We
+expect this tool to effectively detect code duplications in the Linux kernel,
+as preliminary investigations suggest that kernel code files often contain a
+mix of duplicated and unique functions. To validate the tool, we will use a
+triangulation method, comparing its results with established code duplication
+databases referenced in the literature, and conducting an empirical analysis on
+a randomly selected set of files from the AMD Display driver, which our tool
+flags as containing duplications. A detailed explanation of our validation
+process is provided in Chapter \ref{cha:method}.
+
+Upon completing the first phase, we anticipate having a validated tool for
+identifying code duplications within the Linux kernel context. We will then
+proceed to the second phase, which focuses on strategies for mitigating these
+duplications. In \textbf{Phase II}, our objective is to identify practical
+approaches to address the code duplications detected in the Linux kernel. To
+achieve this, we have designed an ethnographic approach that enables us to
+interact with the Linux development community and gather artifacts that either
+support or challenge our proposed mitigation strategies. Figure
+\ref{fig:reDesign} summarizes the research steps proposed in this study.
 

### 2024-11-09 conteudo/tool/tool.tex
add ref to github implementation of the tool
Reviewing tool proposal - part 1

@@
 
-Our proposed tool is a Command Line Interface (CLI) software. Its purpose is to enable developers to query 
-information about code clone duplication at the function level, which means finding pairs of functions in a codebase 
-that are code clones of each other.
+Our proposed tool is a Command Line Interface (CLI) application designed to
+help developers identify code clone duplication at the function level.
+Specifically, it enables the detection of pairs of functions within a codebase
+that are clones of each other.
 
-The tool operates in two steps. The first step, called Setup, is executed only 
-once for each codebase and is responsible for processing the codebase by extracting functions from code files and 
-searching for code clone duplications. The second step, called Query Responder, executes the tool's functionalities 
-using the information obtained in the Setup step. More information about the Setup and Query Responder behavior can 
-be found in sections \ref{subsec:architecture}  and \ref{subsec:func}.
+The tool operates in two main steps. The first step, called *Setup*, is
+executed only once for each codebase. This step processes the codebase by
+extracting functions from code files and identifying code clone duplications.
+The second step, called *Query Responder*, utilizes the information obtained in
+the Setup phase to execute the tool’s functionalities as needed. Detailed
+descriptions of the Setup and Query Responder processes can be found in
+Sections \ref{subsec:architecture} and \ref{subsec:func}.
+
+The tool has been implemented and is available for reference at
+\url{https://github.com/LipArcanjo/test-duplication-code}.

### 2024-11-09 conteudo/tool/architecture.tex
remove commentary about fig of duplication codebase
Reviewing tool proposal - part 2

@@
 
-The Setup step consists of two components, the Function Breaker and the Code Duplication Finder. 
-The pipeline of this step works as follows: the Function Parser receives the codebase the user 
-is interested in, extracts the functions of the codebase along with metadata, and creates a new 
-temporary codebase where the functions extracted become news code files. The Code Duplication Finder 
-iterates over every pair of files in the temporary codebase, checks if they are code clones, and if 
-so, saves them in the Code Duplication Database, which is a text file that stores every code duplication 
-as a triple \textbf{<function1,function2,similarity>}, where function1 and function2 are the functions 
-that are a duplication of each other, the similarity is the metric given by the code duplication detection 
-method utilized, which in our case, it is the cosine similarity explained in \ref{sec:similarity}. 
-Finally, the Query Responder consumes the temporary codebase and the Code Duplication Database to 
-extract duplicated functions-related information per user request. Figure \ref{fig:diagram} illustrates 
-the tool pipeline. There is an explanation of how each component works through this section.
+The Setup step consists of the Function Breaker and the Code Duplication Finder.
+The pipeline of this step works as follows: the Function Parser receives the codebase the user is interested in, extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files.
+The Code Duplication Finder iterates over every pair of files in the temporary codebase, checks if they are code clones, and, if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple \textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized, which in our case is the cosine similarity explained in \ref{sec:similarity}.
+Finally, the Query Responder consumes the temporary codebase and the Code Duplication Database to extract duplicated functions-related information per user request.
+Figure \ref{fig:diagram} illustrates the tool pipeline.
+An explanation of how each component works is provided in this section.
 
@@
 
-The input codebase is a folder in the machine the tool is running into. All files in the codebase 
-that are not source code files from a supported programming language by the tool are ignored. The 
-tool validates the support for a source file through analysis of its file extension.
+The input codebase is a folder on the machine running the tool.
+All files in the codebase that are not source code files from a programming language supported by the tool are ignored.
+The tool validates support for a source file by analyzing its file extension.
 
@@
 
-The Function Parser receives the input codebase and transforms it into the temporary codebase. The 
-Function Parser iterates through every source code file from the programming language it supports 
-and uses a specific programming language function extractor to extract every function in the file. 
-For each function extracted, we create two new source code files and a metadata file in the temporary 
-codebase, represented by the pair \textbf{<file name, function name>},  the source code file of the function, and the proper function. 
-The first new source code file contains the function’s body from the function it represents, while 
-the other new source code file contains the function’s signature. The metadata file contains 
-additional relevant information about the function, such as the function name, the line where the 
-function signature starts in the source code file, and the line where the function’s body ends in 
-the source code file. The programming language we support at the moment is C and Java. 
-Figure \ref{fig:transform} illustrates an example of a function transformation into the 
-temporary codebase.
+The Function Parser receives the input codebase and transforms it into the temporary codebase.
+The Function Parser iterates through every source code file from the programming language it supports and uses a specific programming language function extractor to extract every function in the file.
+For each function extracted, two new source code files and a metadata file are created in the temporary codebase, represented by the pair \textbf{<file name, function name>}, the source code file of the function, and the proper function.
+The first new source code file contains the function's body from the function it represents, while the other new source code file contains the function's signature.
+The metadata file contains additional relevant information about the function, such as the function name, the line where the function signature starts in the source code file, and the line where the function's body ends in the source code file.
+The programming languages supported at the moment are C and Java.
+Figure \ref{fig:transform} illustrates an example of a function transformation into the temporary codebase.
 
@@
 
-About the programming language’s function extractor, we approached this problem in a way that any 
-person comfortable with a language can adapt our approach to their specific programming language. 
-Usually, the code blocks can be easily extracted from source code files, independently if the code 
-blocks are defined by curly brackets or by indentation, and it is possible to infer if it represents 
-a function’s body given how deeply it is in the code context and a few lines before the start of the 
-code block. We followed this approach to implement the function extractor for the supported languages. 
-As an alternative approach, we can build the syntax tree \citep{compiler} 
-of the programming language and use it to extract the functions, which is a more complex task than 
-our approach. For the programming languages that we support, we implemented them as follows:
+Regarding the programming language's function extractor, we approached this problem so that anyone comfortable with a language can adapt our approach to their specific programming language.
+Usually, code blocks can be easily extracted from source code files, independently if the code blocks are defined by curly brackets or by indentation, and it is possible to infer if they represent a function's body given their depth in the code context and a few lines before the start of the code block.
+We followed this approach to implement the function extractor for the supported languages.
+As an alternative approach, we can build the syntax tree \citep{compiler} of the programming language and use it to extract the functions, which is a more complex task than our approach.
+For the programming languages that we support, we implemented them as follows:
 
 \begin{itemize}
-	\begin{item}
-		\textbf{C}: Given a source code file, we extract every code block from it, maintaining the depth 
-		of the code block in the code context and some of the code lines before the definition of the 
-		code block. According to the C programming language’s grammar, functions are defined in a 
-		global context at depth 0. There are other elements with the same depth as functions, and we 
-		differentiate them by analyzing if the first non-empty character before the curly brackets of 
-		the code block is a closing parenthesis. This pattern uniquely identifies functions 
-		within C code.
+	\item \textbf{C}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in the code context and some of the code lines before the definition of the code block.
+According to the C programming language's grammar, functions are defined in a global context at depth 0.
+There are other elements with the same depth as functions, and we differentiate them by analyzing if the first non-empty character before the curly brackets of the code block is a closing parenthesis.
+This pattern uniquely identifies functions within C code.
 
-	\end{item}
-	\begin{item}
-		\textbf{Java}:  Given a source code file, we extract every code block from it, maintaining the 
-		depth of the code block in the code context and some of the code lines before the definition 
-		of the code block. According to Java’s grammar, the functions are defined inside classes. 
-		Functions can exist in different depths, as Java accepts inner classes. We differentiate 
-		functions of other elements by analyzing if the first non-empty character before the curly 
-		brackets of the code block is a close parenthesis. There is a problem with this approach in 
-		Java, as there are other elements that respect this pattern, such as for loop, thus requiring 
-		more filters. We chose to analyze functions of depth 1, which ignores every other element that 
-		is a code block and has a parenthesis as the first non-empty character before the curly brackets. 
-		The counterpart to this choice is that the extractor does not find functions declared inside 
-		inner classes. The counterpart is not a problem in this work as there are no inner classes 
-		on the BigCloneBench \citep{bigclonebench}, the only codebase in Java explored in this work.
-
-	\end{item}
+	\item \textbf{Java}:  Given a source code file, we extract every code block from it, maintaining the depth of the code block in the code context and some of the code lines before the definition of the code block.
+According to Java's grammar, functions are defined inside classes.
+Functions can exist at different depths, as Java accepts inner classes.
+We differentiate functions from other elements by analyzing if the first non-empty character before the curly brackets of the code block is a closing parenthesis.
+A challenge with this approach in Java is that other elements, such as for-loops, can follow this pattern, necessitating additional filters.
+We chose to analyze functions at depth 1, which ignores any other element that is a code block and has a parenthesis as the first non-empty character before the curly brackets.
+This choice means the extractor does not find functions declared inside inner classes, but this limitation is not an issue in this work, as there are no inner classes in BigCloneBench \citep{bigclonebench}, the only Java codebase explored in this work.
 \end{itemize}
@@
 
-The temporary codebase is a transformation of the input codebase done by the Function Parser. 
+The temporary codebase is a transformation of the input codebase performed by the Function Parser.
 Each function in the input codebase is represented in the temporary codebase as three files.
-Descriptions of these files are provided below, and an example of each can be found in 
-the figure \ref{fig:transform}.
+Descriptions of these files are provided below, with examples shown in Figure \ref{fig:transform}.
 
 \begin{itemize}
-	\begin{item}
-		\textbf{Header file}: This file contains the signature of the function it represents;
-	\end{item}
-	\begin{item}
-		\textbf{Source file:} This file contains the body of the function it represents;
-	\end{item}
-	\begin{item}
-		\textbf{Info file:} This file contains metadata of the function. The information this 
-		file contains is the function name, the file that has the function, the relative path 
-		of the file related to the input codebase, the line that the function’s signatures 
-		start, the line that the function body initiates, the line that the function’s body 
-		ends and if there is a line break between the end of the function’s signature and the 
-		start of the function’s body.
-	\end{item}
+	\item \textbf{Header file}: This file contains the signature of the function it represents.
+	\item \textbf{Source file}: This file contains the body of the function it represents.
+	\item \textbf{Info file}: This file contains metadata about the function, including the function name, the file that contains the function, the file's relative path to the input codebase, the line where the function's signature starts, the line where the function body begins, the line where the function's body ends, and whether there is a line break between the end of the function's signature and the start of the function's body.
 \end{itemize}
@@
 
-The Code Duplication Finder iterates through every pair of source code files in the temporary 
-codebase, which in this context represents functions of the input codebase and, for every pair 
-of files, we execute a code duplication detection method that computes a metric that measures 
-of how similar the pair of files are, which we call similarity. Finally, if the similarity is 
-greater or equal to the minimum similarity threshold (a parameter passed by the user when it 
-executes the Setup step), we store this pair of functions along with its similarity in the 
-Code Duplication Database.
+The Code Duplication Finder iterates through every pair of source code files in the temporary codebase, representing functions from the input codebase.
+For each pair of files, we execute a code duplication detection method that computes a metric measuring how similar the pair of files is, which we refer to as similarity.
+If the similarity is greater than or equal to the minimum similarity threshold (a parameter provided by the user during the Setup step), we store this pair of functions along with its similarity in the Code Duplication Database.
 
-For the code duplication detection method, we visualize the source code files as texts and 
-execute the TF-IDF vector embedding method implemented by the Gensim library \citep{gensim}
-and compute the cosine similarity as the similarity metric. We chose this method given the
-alleged Gensim statements about its performance, being a programming language-independent 
-method, and the nonnecessity of compilable code, which is expected to happen in the temporary 
-codebase as it does not store complete code artifacts. Platis implemented this code 
-duplication detection method and distributed it to the public with an MIT license
-\citep{platistool,mitlicense}.
-We utilized his implementation as it has some useful components, such as removing code 
-comments. We made the appropriate changes in the input and output of his implementation 
-to respect the expected format of our tool. It is feasible to change the duplication 
-detection method similarly to what we did with Platis’s implementation.
+For the code duplication detection method, we treat the source code files as text and apply the TF-IDF vector embedding method implemented by the Gensim library \citep{gensim}, then compute cosine similarity as the similarity metric.
+This method was chosen for its claimed performance, programming language independence, and the fact that it does not require compilable code, which is expected in the temporary codebase, as it does not contain complete code artifacts.
+\cite{platistool} implemented this code duplication detection method and released it under the MIT license \citep{mitlicense}.
+We adapted his implementation to fit our tool's expected input and output formats.
+Switching the duplication detection method is feasible, similar to how we adapted Plates' implementation.
 
@@
 
-The Code Duplication Database is a text file that contains the duplicated pairs of 
-functions found by the Code Duplication Finder. The first line of the file contains 
-a number that represents the number of the duplicated pairs. After that, there is a 
-line for each duplicated pair containing the path of the functions in the temporary codebase 
-and the similarity metric of the pair. An example of the Code Duplication Database can be 
-found in the figure (ADD FIGURE HERE).
+The Code Duplication Database is a text file that contains the duplicated pairs of functions found by the Code Duplication Finder.
+The first line of the file specifies the number of duplicated pairs.
+Following this, each line lists a duplicated pair, including the paths of the functions in the temporary codebase and the similarity metric of the pair.
 
 \subsection{Setup step}
-
 \label{subsec:setup}
 
-The Setup step is a procedure executed by the user once per codebase. The Setup step takes 
-the input codebase and a minimum similarity metric value for function pairs considered 
-duplications. It then runs the Function Parser and the Code Duplication Finder to create 
-a temporary codebase and the Code Duplication Database. The value of the minimum similarity 
-metric is to reduce the Code Duplication Database, which optimizes the memory usage and the 
-computational cost of the Query Responder step. For example, if this parameter does not exist 
-and we work in a codebase of $10000$ functions and supposing each relative path plus the 
-function name has 50 characters, the Code Duplication Database will be a file of size 
-approximately $5000^2 \times 50 ~= 5$ Gigabytes, as it is not expected in large codebase 
-to the majority of pairs of functions be a duplication of each other, 
-the size file can be reduced dramatically.
+The Setup step is a procedure executed by the user once per codebase.
+The Setup step takes the input codebase and a minimum similarity metric value for function pairs considered duplicates.
+It then runs the Function Parser and the Code Duplication Finder to create a temporary codebase and the Code Duplication Database.
+Setting a minimum similarity metric reduces the Code Duplication Database size, optimizing memory usage and the computational cost of the Query Responder step.
+For example, if this parameter does not exist and we work with a codebase of $10000$ functions, each with a 50-character relative path and function name, the Code Duplication Database would reach approximately $5000^2 \times 50 ~= 5$ gigabytes.
+As most function pairs are unlikely to be duplicated in large codebases, this file size can be significantly reduced.
 
@@
 
-The Query Responder step is the component the user executes multiple times to consult information 
-about duplicated functions in a codebase processed by the Setup step. The Query Responder step 
-consumes the temporary codebase and the Code Duplication Database created by the Setup step. 
-This component implements all the tool’s functionalities exposed to the user, described in 
-\ref{subsec:func}.
+The Query Responder step is the component the user executes multiple times to consult information about duplicated functions in a codebase processed by the Setup step.
+The Query Responder step utilizes the temporary codebase and the Code Duplication Database as inputs to answer the user's request, such as querying the total number of duplicated functions in a specific input codebase or listing duplicated functions above a certain similarity threshold.
+This component was developed to be extensible, allowing other queries to be easily added.
+Currently, the Query Responder step supports the following queries:
+
+\begin{itemize}
+	\item Number of duplicated functions above a similarity threshold.
+	\item Number of duplicated functions equal or above a similarity threshold.
+	\item List of duplicated functions above a similarity threshold.
+	\item List of duplicated functions equal to or above a similarity threshold.
+\end{itemize}

### 2024-11-09 conteudo/tool/architecture.tex
Reviewing tool proposal - part 3

@@
 The pipeline of this step works as follows: the Function Parser receives the codebase the user is interested in, extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files.
-The Code Duplication Finder iterates over every pair of files in the temporary codebase, checks if they are code clones, and, if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple \textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized, which in our case is the cosine similarity explained in \ref{sec:similarity}.
+The Code Duplication Finder iterates over every pair of files in the temporary codebase, checks if they are code clones, and, if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple \textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized, which in our case is the cosine similarity explained in Section \ref{sec:similarity}.
 Finally, the Query Responder consumes the temporary codebase and the Code Duplication Database to extract duplicated functions-related information per user request.

### 2024-11-09 conteudo/methods/methods.tex
create resumo and remove appendix as it is unnecesary at the moment
Reviewing research desing

@@
-\en
+\todo[inline]{Depois, veja melhor o tempo verbal ...
+tem coisas que coloquei no presente, porque ainda está fazendo ou irá fazer; tem outras coisas já planejadas e feitas que o texto ficou no passado ...
+veja se está fluindo bem e como deixar mais uniforme o tempo verbal aqui e nas outras partes do texto, tentando deixar claro que já está feito e o que será feito depois}
 
-To answer our proposed research questions, we opted to do a multiple-method research 
-where we covered the research question with different methodologies. To approach the RQ1, 
-we opted to create an ethnographic study using participant-observation methodology where 
-we plan to mitigate a selected subset of duplicated functions by refactoring them using 
-a designed systematic approach while interacting with the Linux community to collect 
-feedback on the mitigation made by us.
+We employ a multi-method research approach, using different methodologies to achieve our goals.
+Our first choice is to conduct an ethnographic study using a participant-observation methodology.
+In this study, we aim to mitigate a subset of duplicated functions by refactoring them systematically while interacting with the Linux community to gather feedback on our mitigation efforts.
+To address our main objective of mitigating code duplication in the Linux kernel, we first needed the ability to identify these duplications accurately.
+Thus, we developed a tool and assessed its capability to detect code duplications in the Linux kernel.
 
-To answer RQ1, which covers our main objective to mitigate code duplication in the Linux 
-kernel, we are first required to be able to identify them. For this reason, we developed 
-a tool and need to understand if the tool can identify code duplications in the Linux kernel, 
-which covers the supplementary research question RQ1.1 for the RQ1.
+We apply two independent methods and perform triangulation on the results.
+The first method evaluates the proposed tool using an approach from the literature, comparing it against the BigCloneBench dataset \citep{bigclonebench}.
+The second method is an empirical analysis of a subset of functions within the AMD Display driver that our tool identified as duplicates.
+After completing these two methods, we will triangulate the findings to analyze the results comprehensively.
+We chose triangulation because relying solely on the literature-based method has limitations and may not be entirely reliable \citep{bigfail, litreview}.
 
-To answer RQ1.1, we applied two independent submethods and performed a triangulation of 
-the results. The first submethod is to evaluate the proposed tool similarly to the literature, 
-by comparing it against the BigCloneBench dataset \citep{bigclonebench}. The second submethod
-for answering RQ1.1 is to do an empirical analysis on a selected subset of functions of the 
-AMD Display driver that our tool infers is a duplication. With the two submethods completed, 
-we will do a triangulation of results analyzing the findings. We opted for the triangulation 
-of the two submethods because the formal literature method alone is limited and not fully 
-trustworthy \citep{bigfail, litreview}.
-
-\section{Evaluation of the tool by the formal literature}
+\section{Evaluation of the Tool by Formal Literature}
 
@@
 
-The BigCloneBench dataset \citep{bigclonebench} is a collection of Java code with a selection of
-code duplication pairs categorized by the authors. These code duplication pairs are labeled by 
-the code clone type described in \ref{subsec:types}.
+The BigCloneBench dataset \citep{bigclonebench} is a Java code collection containing pairs of code duplications categorized by clone type, as defined in Section \ref{subsec:types}.
+%
+Following the methodology presented by \citep{tailor}, we sampled 20,000 pairs from each clone type, adding 20,000 non-duplicate pairs as negative samples.
+We applied the same sampling approach to our tool to ensure a fair evaluation.
 
-Liut et al. presented a sample method to evaluate their code duplication detection method, 
-in which they sample 20,000 pairs from each of the code duplication types with the addition 
-of 20,000 nonduplication pairs to act as the negative samples \citep{tailor}. 
-For a fair evaluation, we use the same sample method as them to evaluate our tool.
+Unlike traditional methods, our tool does not distinguish between clone types and identifies duplications at the function level rather than the file level, which is typical in state-of-the-art tools.
+Therefore, we adapted the metric calculation method for evaluation purposes.
+Specifically, we considered every pair of functions with a similarity metric equal to or greater than a threshold \(X\) as duplicates and marked the corresponding file pairs.
+A correctly identified duplication pair counts as accurate for its clone type, while an incorrectly inferred pair is considered incorrect across all clone types.
+With these definitions, we computed the Recall metric from Section \ref{subsec:codemethods} for each clone type.
+%
+To understand the impact of varying the similarity threshold, we evaluated our tool using different threshold values: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%.
+We then analyzed and discussed the results.
 
-The proposed tool does not differentiate the code duplication types and infers duplications 
-at the function level instead of the code file level, which is the standard in state-of-the-art 
-methods. We were required to adjust the form we computed metrics of the evaluation. To accomplish 
-this, we will define that every pair of functions in which the similarity metric is equal or 
-superior to a threshold X will be considered a duplication, and the pair of code files in which 
-the functions exist are labeled as duplication. If, on this method, a code duplication pair is 
-correctly identified, it will count as correct to its code duplication types. If an incorrect 
-duplication pair is inferred as a duplication, it will count as incorrect to all code duplication 
-types. Given these definitions, we will compute the Recall metric presented in \ref{subsec:codemethods}
-for each code duplication type.
-
-To understand the variation of meaning when we change the similarity threshold, we will evaluate 
-our tool for the given variations of the threshold X and discuss the results: 
-30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, 100\%.
-
-\section{Evaluation of the tool by Empirical analyses}
+\section{Evaluation of the Tool by Empirical Analysis}
 
@@
 
-Given the problems of the BigCLoneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail}
-and the limitations of it in our scope, such as being a Java dataset while the Linux kernel 
-is predominantly built in the C programming language, we opted to, together with the formal 
-literature evaluation, perform an empirical study analyzing the codes in which our tool inferred 
-as duplications.
+Due to the limitations of the BigCloneBench dataset, as noted by \cite{bigfail}, and its lack of suitability for our context (Java dataset vs.
+C-based Linux kernel), we complemented the formal evaluation with an empirical study of the tool inferred duplications.
+%
+Given the vast number of lines and components in the Linux kernel, analyzing every duplication is impractical.
+We focused our scope on the AMD Display driver, which, as shown in Figure \ref{fig:relatory_ex}, contains over 20,000 lines of duplicated code at a similarity metric of 100\%.
 
-The number of code lines and components in the Linux kernel makes it impractical to analyze every 
-code duplication. To limit our scope, we will analyze solely code duplications in the AMD Display 
-driver. This limitation still does not reduce the code duplication to analyze, in our preliminary 
-tool execution shown in Figure \ref{fig:relatory_ex}, there are over 20,000 lines of code 
-duplications on the AMD Display driver with a similarity metric of 100\%.
+In this empirical analysis, we randomly sampled function pairs identified by the tool as duplicates.
+For each similarity threshold \(X\) (30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%), we randomly selected ten function pairs with a similarity close to \(X\), allowing for a 1\% deviation.
+We implemented the random selection using the tool's source code.
 
-For our empirical analyses, we opted to make a random selection in the code duplication pairs 
-inferred by our tool. To do that, we will use the definition that if the similarity metric of a 
-function pair is greater or equal to X is considered a duplication. To understand the optimal 
-values of X, we will analyze 10 functions pair random selected that have similarity metric X 
-for each of the variations of the parameter X, which are 
-30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, 100\%. 
-As it is unexpected to have code duplications with similarity equals a whole number, we will 
-allow a variation of 1\% for greater or less to each of the variations of the parameter X.
+Appendix \ref{app:emp} lists the randomly selected function pairs.
+We empirically evaluated these pairs to determine if they were duplications, comparing these results with the literature-based method to assess our tool's effectiveness and understand the impact of the minimum similarity parameter.
 
-To make the random selection of function pairs, we list the code duplications for each of the 
-X variations and execute a random selection algorithm built into the source code language of 
-our tool. The random selected function selected function pairs can 
-be found in the Appendix \ref{app:emp}.
-
-When we analyze the selected function pairs, we will empirically judge if they are duplications. 
-We intend to collect the data on this empirical analysis and compare it to the formal literature 
-method to understand if our tool detects duplicated functions and the difference it makes when 
-varying the minimum similarity parameter of our tool.
-
-\section{Ethnographic study}
+\section{Ethnographic Study}
 
@@
 
-Ethnography is a research method recognized as a significant qualitative empirical approach to 
-understanding people, their cultures, and their associated social and work practices \citep{bookethno}.
-The central tenet of the ethnography approach is to enable researchers to understand values, 
-beliefs, or ideas shared by a community under study from the members’ point of view 
-\citep{ethnosoft}. Thus, the ethnographer acts as a community member, learning and observing 
-what people do and why they do it.
+Ethnography is a recognized qualitative empirical method for understanding people, their cultures, and their work practices \citep{bookethno}.
+It allows researchers to gain insights into values, beliefs, and ideas from the perspective of community members \citep{ethnosoft}.
+In our research, refactoring code duplications without community interaction would prevent us from understanding whether Linux maintainers would accept our approach or identify their concerns.
+Thus, we used ethnography, positioning ourselves as Linux kernel contributors focused on removing identified code duplications.
 
-Sharp et al. discuss the application of the ethnography method in software engineering research, 
-demonstrating the low adoption of the method and its advantages in uncovering what practitioners 
-do and why they do it \citep{ethnosoft}. In our research context, if we mitigate code 
-duplications by ourselves without interacting with the Linux community, we cannot understand if 
-the Linux maintainers would accept our approach for code duplication and what concerns they 
-would have. Thus, we resorted to ethnography research, where we will act as a member of the 
-Linux kernel and act as a contributor with a focus on removing code duplications found.
+This section outlines our systematic approach for mitigating code duplication in the Linux kernel, details the methodology for selecting duplicated functions, and describes the ethnographic methodology for engaging with the Linux community.
 
-In this section, we will outline a systematic approach to mitigate code duplication in the 
-Linux kernel. Next, we will detail the methodology for selecting a subset of duplicated functions 
-identified by our tool. Finally, we will describe the ethnographic methodology for interacting 
-with the Linux community.
-
-\subsection{Systematic approach to mitigate code duplication}
-
+\subsection{Systematic Approach to Mitigate Code Duplication}
 \label{subsec:pipeline}
 
-The C programming language’s way to communicate between source code files is through creating 
-header files, which is a library specification \citep{Cbook}. As the Linux kernel 
-is predominantly built in C, we remove code duplications by creating a library where the code 
-artifacts will use the library instead of their duplications.
+In C programming, communication between source files is achieved by creating header files that specify libraries \citep{Cbook}.
+Since the Linux kernel is primarily written in C, we eliminated code duplications by consolidating duplicated code into a single library, which would replace instances of duplication across the codebase.
+%
+For each duplicated function pair, we used the \textit{function information} functionality (described in Section \ref{subsec:functioncommand}) to locate all instances of the function in the code.
+This strategy resulted in a collection of duplicated functions and corresponding code files.
 
-When removing duplicated functions, our desired behavior is to remove the functions in every 
-portion of the code in which the functions exist, only allowing a unique version to exist in 
-the library created. For this reason, given our initial duplicated function pair, we will use 
-the function information functionality(Described in \ref{subsec:functioncommand}) to find all
-places where this function exists in the source code. Thus, we finish with a collection of 
-duplicated functions and the code files in which these functions exist.
+To identify shared code more effectively, we extended this approach to search for other common functions across all collection files.
+We then applied specific refactoring methods to each shared function.
+If the functions were identical across files, we removed the duplicates and created a single function in the library.
+If modifications existed, we applied case-specific refactoring.
 
-Given the semantics of a library containing the shared code by a collection of code files, 
-we added a step of the approach, which is searching for other functions contained in all of 
-the code files of the collection we are working with. Thus, we have a collection of code files 
-and a non-empty set of functions shared by all the code files in the collection.
+Successful mitigations were submitted to the Linux community to gauge their feedback.
+If unsuccessful, we categorized the cases: some required specialist decisions beyond our expertise, and others were potentially non-refactorable.
+All mitigations were documented for research transparency at \url{https://github.com/LipArcanjo/master-artifacts}.
 
-Finally, for each of the functions shared by all code files, we make a specific-case approach 
-to mitigate each one of them. If a function is shared by all the code files without 
-modifications, we use a simple approach to only remove them and create the function in the 
-source code file. If it is shared with modifications, we look for refactoring methods applicable 
-to this specific case.
+\subsection{Selection of Duplicated Functions}
 
-Finishing our mitigation method, if the mitigation ends up working in an initial step, we may 
-move it forward to the Ethnographic method, interacting with the Linux community to understand 
-their thoughts on the mitigation. If the mitigation ends up not working as expected, we will 
-categorize it in two categories, the first one is that the refactoring requires a decision in 
-which we, as not specialists and non-maintainers of the Linux kernel, cannot make it correctly,
-although we believe a specialist would be able to refactor it. The second categorization is 
-that we do not know if it is possible to refactor the duplication. With or without successful 
-mitigation, we will store our mitigations on \url{https://github.com/LipArcanjo/master-artifacts} 
-to research transparency.
+As with the empirical analysis, evaluating every code duplication in the Linux kernel is impractical.
+Thus, we limited the experiment to the AMD Display driver and used random sampling of duplicated functions inferred by our tool to initiate the mitigation approach in Section \ref{subsec:pipeline}.
+We chose a sample size of 30 duplicated functions, as studies indicate this size provides statistical significance \citep{sample1, sample2}.
 
-\subsection{Selection of the duplicated functions}
+For random selection, we restricted eligibility to function pairs with a similarity metric of 100\%. With this criterion, we randomly chose functions from eligible pairs uniformly.
 
-As discussed in the empirical analyses for the tool, it is impractical to evaluate every code 
-duplication on the Linux kernel. Thus, similarly to the empirical analyses for the tool, we will 
-limit the experiment scope to only the AMD Display driver and select a random sample of duplicated 
-functions inferred by our tool to initiate the mitigation approach defined in \ref{subsec:pipeline}. 
-We chose a sample size of 30 duplicated functions, as multiple studies demonstrated that a sample 
-size of this size has statistical significance \citep{sample1,sample2}.
+\subsection{Ethnographic Methodology}
 
-On the random selection of the duplicated functions, we decided to put an additional restriction 
-on which function pairs are eligible to be selected. The restriction is that we will choose function 
-pairs with a similarity metric equal to 100\%. We made this choice because, as shown in \ref{fig:relatory_ex},
-there are more than 20,000 lines of code in the AMD Display driver in which the similarity metric 
-reaches this value. On all function pairs that respect the restriction, we will choose them 
-uniformly at random.
+We conduct ethnographic research upon completing a code duplication mitigation following the systematic approach in Section \ref{subsec:pipeline}.
+We adopt a participant-observation methodology, contributing to the Linux kernel for the AMD Display driver and interacting with maintainers to incorporate our mitigations.
 
-The function pairs selected randomly following our methodology can be consulted in the Appendix
-(CREATE Appendix AND ADD HERE).
+Linux kernel communication occurs via public mailing lists, where discussions are accessible to all.
+We collect data artifacts from the AMD Display driver mailing list, which records community feedback on our mitigation efforts, providing valuable insights into the proposed approach effectiveness.
 
-\subsection{Ethnographic methodology}
-
-When we make a code duplication mitigation following our systematic approach described in 
-\ref{subsec:pipeline}, we are apt to initiate our ethnographic research. We resorted to a
-participant-observation approach as our ethnographic methodology, where we will act as 
-contributors of the Linux kernel to the AMD Display driver, interacting with the maintainers 
-of the AMD Display driver attempting to get our mitigations added to the Linux kernel.
-
-The Linux kernel communication is made through public mailing lists, where anyone can consult 
-the discussions. We plan to collect data artifacts through the AMD Display driver mailing list. 
-The mailing list will register our interactions with the Linux community, with their members 
-sharing their opinions and suggestions for the mitigation we made, corroborating the discussion 
-about the effectiveness of our proposed approach.
-
-Given the time limit of our research, we do not expect to be able to interact with the Linux 
-community throughout the successful mitigation we made using the proposed approach, as it passes 
-through factors we do not control, which is the time spent waiting for Linux community’s members 
-interact with us. For this reason, we resorted to committing to do the participant-observation 
-approach in 5 of the successful mitigations and use them for our discussion.
+Given our research timeline, we do not expect continuous interactions with the Linux community throughout all mitigations, as response time varies.
+Thus, we committed to applying the participant-observation approach for three to five successful mitigations and using these for our final discussion.

### 2024-11-09 conteudo/eval/eval.tex
Reviewing the final chapter and applying other adjustments

--- 2024-07-30 conteudo/eval/eval.tex
@@
-On this Chapter, we exposes our experiment results bout the proposed tool performance following
-the methology described in Chapter \ref{cha:method}. Firstly, we shows the results obtained by our tool
-in the BigCloneBench dataset\citep{bigclonebench}, which is the standard form of the formal literature 
-\citep{litreview}. Secondly, we shows the results obtained by our tool on the empirical analyses. 
-Finaly, we do a triangulation of data analysing our finds on the tool performance.
+\en
 
-\subsection{Tool results on the BigCloneBench dataset}
+\todo[inline]{Juntei os capítulos; veja se consegue colocar algo aqui no início; não sei se final remarks é o melhor nome para a segunda secao aqui}
+
+\section{Tool Evaluation}
+\label{sec:eval}
+
+In this section, we present the test results of our tool performance, following the methodology described in Chapter \ref{cha:method}.
+First, we provide the results obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}, which serves as a standard reference in the literature \citep{litreview}.
+Second, we discuss the results from our empirical analysis.
+Finally, we triangulate the data, analyzing our findings regarding the tool performance.
+
+\subsection{Tool Results on the BigCloneBench Dataset}
 
@@
 
-As described in section \ref{sec:metbig}. For the evaluation of our proposed tool in the
-BigCLoneBench dataset  \citep{bigclonebench}, we define that every pair of functions in which the 
-similarity metric is equal or superior to a threshold X will be considered a duplication, and the 
-pair of code files in which the functions exist are labeled as duplication. If, on this method, 
-a code duplication pair is correctly identified, it will count as correct to its code duplication 
-types. If an incorrect duplication pair is inferred as a duplication, it will count as incorrect 
-to all code duplication types. Table \ref{tab:bigclone} shows the results obtained by our tool in the 
-BigCloneBench dataset \citep{bigclonebench}.
+As described in Section \ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset \citep{bigclonebench}, we define that every pair of functions with a similarity metric equal to or greater than a threshold \( X \) is considered a duplication.
+The pair of code files containing these functions are then labeled as duplications.
+If our tool correctly identifies a code duplication pair, it is counted as correct for the respective code duplication type.
+If it infers a duplication pair incorrectly, it is counted as incorrect for all code duplication types.
+Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}.
 
-The results obtained in Table \ref{tab:bigclone} indicate that our tool is a pessimist. There 
-is no negative sample inferred as duplication even for a minimum similarity threshold of 30\%, 
-on the other hand, it is bad for identifying code duplications of complex code duplication types, 
-that is Medium Type-3 and Weak Type-3/Type-4. As mentioned in Chapter \ref{cha:tool}, if a 
-better result is required on these complex code duplication types, it is possible to change 
-the code duplication detection method, as it is designed to be a replaceable component in our 
-tool, given the limitations.
+The results in Table \ref{tab:bigclone} indicate that our tool is conservative in its predictions.
+No negative sample was inferred as a duplication, even with a minimum similarity threshold of 30\%.
+However, it performs poorly identifying complex duplication types, such as Medium Type-3 and Weak Type-3/Type-4.
+As discussed in Chapter \ref{cha:tool}, if improved performance on these complex types is needed, our tool allows replacing the current duplication detection method due to its modular design.
 
-\subsection{Tool results on the empirical analyses}
+\subsection{Tool Results on Empirical Analysis}
+
+As outlined in Section \ref{sec:metemp}, for the empirical evaluation of our tool, we randomly selected ten function pairs across various similarity metric ranges computed by our tool.
+We then examined each pair to determine whether they were duplications and, if so, categorized the duplication type.
+Table \ref{tab:emp} presents the results from our empirical analysis.
 
@@
 
-The results were obtained by randomly selecting ten functions in the AMD Display driver for 
-each similarity range and analyzing if they are duplications and which duplication type.
+The results in Table \ref{tab:emp} show that as the similarity metric decreases, there is a notable increase in function pairs that are not duplications within the Linux kernel.
+For similarity ranges of 79\%-81\% or higher, there is a high percentage of code duplications.
+At 69\%-71\% similarity, the duplication rate drops to 50%.
+At 59\%-61%, it further decreases to 20%, and below 49\%-51%, no code duplications were found.
 
-\subsection{Discussion}
+\subsection{Triangulation of Results}
+
+The BigCloneBench results and our empirical methods indicate that our tool performs well in detecting Type-1 and Type-2 code clone duplications but struggles with more complex types.
+A similarity threshold between 80\% and 100\% yields favorable results in both methods.
+
+However, there is a discrepancy in the detection of non-duplication pairs between the two methods.
+In the BigCloneBench evaluation, we found no false positives (negative samples inferred as duplications), while the empirical analysis revealed a higher percentage of false positives.
+We propose two potential explanations for this discrepancy.
+
+The first reason may be the limitations and known issues with BigCloneBench, as highlighted by Krinke and Ragkhitwetsagul \citep{bigfail}.
+The second reason could relate to the nature of the AMD Display driver, where each code artifact shares semantic meaning.
+In contrast, BigCloneBench comprises self-contained code artifacts without shared semantics.
+Given that our tool relies on a text-based code clone detection approach, it may naturally perform worse on the AMD Display driver than on the BigCloneBench dataset.
+We cannot determine if state-of-the-art code duplication detection tools face similar issues obscured by evaluations limited to the BigCloneBench dataset.

### 2024-11-10 conteudo/background/codeduplication.tex

--- 2024-07-26 conteudo/background/codeduplication.tex
@@
 
-Code Duplication Detection is a field in computer science studies that has the attention of researchers 
-back to 1988 \citep{firstman}.
-The occurrence of Code Duplication is a harmful artifact to have in software, affecting software-related 
-tasks such as code readability, introduction of bugs, etc \citep{harmone}. 
-In most cases, the occurrence of code duplication tends to create more unstable software 
-than the nonduplicate code counterpart \citep{harmtwo} .
+Code Duplication Detection is a field in computer science that has attracted
+researchers' attention since 1988 \citep{firstman}. The presence of code
+duplication is considered a harmful artifact in software, as it negatively
+affects tasks such as code readability and increases the likelihood of
+introducing bugs \citep{harmone}. In most cases, code duplication tends to make
+software more unstable compared to its non-duplicated counterpart
+\citep{harmtwo}.
 
-Through the years, the subject of studies related to Code Duplication Detection branched out mainly in 
-two fields of study, Code Clone Detection and Code Plagiarism Detection, with the former focused on the 
-technical aspect while the latter also introduced a social aspect of the field of research
-\citep{litreview}. 
-In this work, we are only interested in the Code Clone Detection branch given the nature of our work.
+Over the years, research on Code Duplication Detection has branched into two
+main areas: Code Clone Detection and Code Plagiarism Detection. The former
+focuses on technical aspects, while the latter incorporates social dimensions
+into the research field \citep{litreview}. In this work, we focus solely on the
+Code Clone Detection branch due to the nature of our study.
 
-\subsection{Types of code clone duplication}
+\subsection{Types of Code Clone Duplication}
 
@@
 
-The categorization of two artifacts as a duplication of each other can open the question of what is considered 
-a duplication. To approach this question, the literature classified code duplication on the scale of the 
-changes between the code artifacts, creating a classification of four types of code clone 
-duplications \citep{litreview}. 
-When comparing the two code artifacts, F and G, the four types of code duplication can be described 
-as follows \citep{litreview}:
+The classification of two code artifacts as duplicates raises the question of
+what constitutes duplication. To address this, the literature categorizes code
+duplication based on the extent of differences between code artifacts,
+resulting in four types of code clone duplications \citep{litreview}. When
+comparing two code artifacts, F and G, the four types of code duplication can
+be described as follows \citep{litreview}:
 
@@
     \begin{item}
-		\textbf{Type-1:} The differences between F and G are in the context of revision of comments, variable names,
-		white spaces, and any other kinds of irrelevant elements. Figure
-		\ref{fig:type1} shows a typical example of Type-1 clone duplication 
-		\citep{litreview}. 
+	\textbf{Type-1 (T1):} The differences between F and G are limited to
+	    changes in comments, variable names, white spaces, and other
+	    irrelevant elements. Figure \ref{fig:type1} shows a typical example
+	    of Type-1 clone duplication \citep{litreview}.
     \end{item}
     \begin{item}
-		\textbf{Type-2:} The differences between F and G are the same as type-1, with the addition of 
-		considering the addition and deletion of redundant codes. Figure
-		\ref{fig:type2} shows a typical example of Type-2 code clone duplication  \citep{litreview}. 
+	\textbf{Type-2 (T2):} The differences between F and G include those
+	    from Type-1, along with the addition and deletion of redundant
+	    code. Figure \ref{fig:type2} shows a typical example of Type-2 code
+	    clone duplication \citep{litreview}.
     \end{item}
     \begin{item}
-		\textbf{Type-3:} The difference between F and G is the same as type-2, with the addition of considering 
-		the reorder of code blocks, as well as statements within code blocks. Figure
-		\ref{fig:type3} shows a typical example of 
-		Type-3 code clone duplication \citep{litreview}. 
+	\textbf{Type-3 (T3):} The differences between F and G include those
+	    from Type-2, as well as reordering code blocks and statements
+	    within code blocks. Figure \ref{fig:type3} shows a typical example
+	    of Type-3 code clone duplication \citep{litreview}.
     \end{item}
     \begin{item}
-		\textbf{Type-4:} The difference between F and G is the same as type-3, with the addition of considering 
-		changes in data structures, the order of operands/operators in expressions, or replacing part of codes 
-		with equivalent composition. Figure
-		\ref{fig:type4} shows a typical example of Type-4 code clone duplication \citep{litreview}. 
+	\textbf{Type-4 (T4):} The differences between F and G include those
+	    from Type-3, along with changes in data structures, the order of
+	    operands/operators in expressions, or the replacement of code parts
+	    with equivalent compositions. Figure \ref{fig:type4} shows a
+	    typical example of Type-4 code clone duplication \citep{litreview}.
     \end{item}
-
 \end{itemize}
@@
 \caption{Example of Type-1 code clone duplication.}
-The example is based on the example presented by CHEN et al. \citep{litreview}. 
+The example is based on the work presented by CHEN et al. \citep{litreview}.
 \label{fig:type1}
@@
 \caption{Example of Type-2 code clone duplication.}
-The example is based on the example presented by CHEN et al. \citep{litreview}. 
+The example is based on the work presented by CHEN et al. \citep{litreview}.
 \label{fig:type2}
@@
 \caption{Example of Type-3 code clone duplication.}
-The example is based on the example presented by CHEN et al. \citep{litreview}. 
+The example is based on the work presented by CHEN et al. \citep{litreview}.
 \label{fig:type3}
@@
 \caption{Example of Type-4 code clone duplication.}
-The example is based on the example presented by CHEN et al. \citep{litreview}. 
+The example is based on the work presented by CHEN et al. \citep{litreview}.
 \label{fig:type4}
@@
 
-\subsection{Literature approches for code clone detection}
+\subsection{Literature Approaches for Code Clone Detection}
 
-Given the passage of time, the literature developed many techniques and methods to approach the code clone 
-detection problem. For a better understanding and analysis of these approaches, the literature divided the 
-research done into five methodologies given the nature of their work  \citep{litreview}, which are:
+Over time, various techniques and methods have been developed in the literature
+to address the code clone detection problem. For better analysis and
+understanding, these approaches are categorized into five methodologies based
+on their characteristics \citep{litreview}:
 
 \begin{itemize}
-	\item \textbf{Textual-based approaches:}  The literature review made by Chen \citep{litreview}  
-	states that this methodology is the first to be investigated by the literature. This methodology line 
-	is based on seeing code as a text artifact and applying textual approaches for text similarity. One 
-	example of this methodology applied is the work of Roy and Cordy
-	\citep{textexample},
+    \item \textbf{Textual-based approaches:} According to Chen’s literature 
+    review \citep{litreview}, this is the earliest methodology explored in 
+    the field. It treats code as a text artifact and applies textual similarity
+    techniques. An example is the work of Roy and Cordy \citep{textexample}, 
     which uses parsing and grammar techniques to detect code clones.
 
-	\item \textbf{Token-based approaches:} This method tries to approach the problem by seeing the code in a
-	similar way to the lexical analysis of the compiling process \citep{litreview}. 
-	That means doing work along the lines of recognizing constants, keywords, and other specific programming 
-	language tokens, mapping variables to representations to become naming-independent, etc. One example 
-	of this methodology applied is the work of Toomey et al. \citep{tokenexample} 
-	to detect code clones through hashed token sequences.
+    \item \textbf{Token-based approaches:} This method tackles the problem 
+    similarly to lexical analysis in the compilation process \citep{litreview}. 
+    It involves identifying programming language tokens such as constants and 
+    keywords, mapping variables to independent representations, etc. An example
+    is the work of Toomey et al. \citep{tokenexample}, which detects code 
+    clones through hashed token sequences.
 
-	\item \textbf{Tree-based approaches:} While the token-based approach uses the lexical analysis of the 
-	compiling process, the tree-based approach uses the syntax tree. The syntax tree is a data structure that 
-	stores the semantic information of a code artifact concerning the programming language \citep{compiler}. 
-	One example of research that uses this information to detect code clones is done by Chilowicz and Duris
-	\citep{treeexample}.
+    \item \textbf{Tree-based approaches:} While token-based approaches utilize
+     lexical analysis, tree-based approaches rely on syntax trees, which store
+     semantic information of code artifacts based on the programming language
+     \citep{compiler}. An example of this approach is the work by Chilowicz 
+     and Duris \citep{treeexample}.
 
-	\item \textbf{Metric-based approaches:} This methodology approaches the problem by extracting metrics 
-	from the code artifact and using those metrics as an embedding representation of the artifact to measure 
-	similarity \citep{litreview}, similar to the famous Word2Vec work done by Mikolov et al.
-	\citep{wordtovec}. Examples of metrics can be program size, number of variables, number of access to
-	memory, etc. One example of this methodology applied is the work done by Kaur and Sharm
-	\citep{metricexample}.
+    \item \textbf{Metric-based approaches:} This methodology uses extracted 
+    metrics from code artifacts as embedding representations to measure 
+    similarity \citep{litreview}, akin to the Word2Vec model by Mikolov et al. 
+    \citep{wordtovec}. Metrics can include program size, number of variables, 
+    memory access counts, etc. An example is the work by Kaur and Sharma \citep{metricexample}.
 
-	\item \textbf{Graph-based approaches:} Similarly to the Tree-based approaches, this methodology depends 
-	heavily on an intermediate program representation, called program dependence graph (PDG) \citep{prodg}. 
-	The representation stores information about the program’s control, statement, and data dependencies. 
-	One example of this methodology applied is the recent work done by Liu et al. \citep{tailor},
-	one of the state-of-the-art works in the field.
-
+    \item \textbf{Graph-based approaches:} Similar to tree-based approaches, 
+    this methodology depends on intermediate program representations, 
+    specifically the program dependence graph (PDG), which stores information 
+    on control, statements, and data dependencies \citep{prodg}. A recent 
+    example is the work by Liu et al. \citep{tailor}, which represents one of 
+    the state-of-the-art works in the field.
 \end{itemize}
 
-The state-of-the-art methods for code clone detection usually have high computational time and memory space 
-usage, as well being programming language-specific approaches, which cannot be easily implemented in programming 
-languages that were not used in their research \citep{litreview}. 
-These are concerns for our research, and for this reason, we will use a generalist textual-based approach 
-based solely on Natural Language Processing (NLP) text similarity detection methods, which we will introduce 
-in Chapter \ref{cha:tool} and evaluate this approach in this work.
+State-of-the-art code clone detection methods typically require substantial
+computational time and memory, and many are language-specific, limiting their
+applicability across different programming languages \citep{litreview}. Given
+these constraints, we opted for a generalist textual-based approach using
+Natural Language Processing (NLP) text similarity methods, which we will
+introduce in Chapter \ref{cha:tool} and evaluate in this work.
 
-\subsection{Evaluating code duplication methods}
+\subsection{Evaluating Code Duplication Methods}
 
@@
 
-Given the necessity to evaluate and compare approaches for code clone detection, it was developed and widely 
-adopted: OJ-Clone, a dataset collected from a pedagogical online judge system \citep{ojclone},
-and BigCloneBench, a dataset collected from over 25,000 Java software systems \citep{bigclonebench}. 
-The existence of only two spread datasets is an open concern in literature as the limited scope of OJ-Clone 
-and the problems of BigCloneBench presented by Krinke and Ragkhitwetsagul \citep{bigfail}. The BigCloneBench
-is the dataset used by most of the literature works, given the limited pedagogical scope of the OJ-Clone.
+To evaluate and compare code clone detection methods, two widely adopted
+datasets are available: OJ-Clone, collected from a pedagogical online judge
+system \citep{ojclone}, and BigCloneBench, compiled from over 25,000 Java
+software systems \citep{bigclonebench}. However, literature has expressed
+concerns over the limited scope of OJ-Clone and the issues with BigCloneBench
+identified by Krinke and Ragkhitwetsagul \citep{bigfail}. Despite these
+limitations, BigCloneBench is commonly used due to the limited pedagogical
+scope of OJ-Clone.
 
-To evaluate the code clone detection approaches against the datasets, the works compute metrics using 
-the predictions made related to the instances of the dataset. There are four possible outcomes when an 
-instance is predicted by a code clone detection approach, which are:
+BigCloneBench further divides Type-3 duplications into three categories—Weak
+Type-3 (WT3), Medium Type-3 (MT3), and Strong Type-3 (ST3)—where WT3 is closer
+to Type-4 and ST3 is closer to Type-2 \citep{bigclonebench}. This results in
+six possible outcomes for a code pair: Type-1, Type-2, Strong Type-3, Medium
+Type-3, Weak Type-3/Type-4, and non-duplicate \citep{bigclonebench}.
+
+To assess code clone detection methods against these datasets, researchers
+calculate metrics based on the model’s predictions. There are four possible
+outcomes for each prediction:
 
@@
     \begin{item}
-		\textit{True positivity (TP)}: This outcome happens when the approach predicts that the instance 
-		is a code clone and the instance is a code clone.
+        \textit{True Positivity (TP):} The model correctly identifies an instance as a code clone.
     \end{item}
     \begin{item}
-		\textit{False positivity (FP)}: This outcome happens when the approach predicts that the instance 
-		is a code clone, but the instance is not a code clone.
+        \textit{False Positivity (FP):} The model incorrectly identifies a non-clone instance as a code clone.
     \end{item}
     \begin{item}
-		\textit{True negative (TN)}: This outcome happens when the approach predicts that the instance 
-		is not a code clone and the instance is not a code clone.
+        \textit{True Negative (TN):} The model correctly identifies an instance as non-clone.
     \end{item}
     \begin{item}
-		\textit{False negative (FN)}: This outcome happens when the approach predicts that the instance 
-		is not a code clone, but the instance is a code clone.
+        \textit{False Negative (FN):} The model incorrectly identifies a clone instance as non-clone.
     \end{item}
@@
 
-Given the number of TP, FP, TN, and FN that resulted after an approach evaluation against a dataset,
-it is possible to compute classical metrics such as Precision, Recall, and F-Score \citep{recall}.
-In this work, we use the Recall metric as the code clone detection approaches usually make available 
-this metric to compare themselves in contrast with the others. The Recall metric is defined by the 
-following formula \citep{recall}:
+Using the counts of TP, FP, TN, and FN, standard metrics such as Precision,
+Recall, and F-Score can be calculated \citep{recall}. In this work, we focus on
+the Recall metric, as it is commonly reported in literature for comparing code
+clone detection approaches. Recall is defined as follows \citep{recall}:
 
@@
 
-Where $\sum TP$ is the number of true positives in the evaluation and $\sum FN$ is the number of 
-false negatives. The Recall metric is a number between $0$ and $1$, where a value closer to $1$ is 
-better, and a value closer to $0$ is worse.
+Where $\sum TP$ is the total number of true positives, and $\sum FN$ is the
+total number of false negatives. Recall values range between 0 and 1, with
+values closer to 1 indicating better performance.

### 2024-11-10 conteudo/introduction/intro.tex

@@
 \en
-
-\section**{Problem outline}
-\label{sec:problem}
 
@@
 
-This research aims to identify patterns of software engineering techniques,
-such as refactoring methods commonly used to address code duplications. These
-patterns will serve as a guide for future contributors in mitigating code
-duplication issues within this context. To investigate our approach, we use the
-initial version of our tool to identify duplicate code in the Linux kernel,
-specifically targeting parts of the AMD Display driver code. We will then
-refactor portions of this code and submit patches to maintainers, providing
-evidence of the tool's accuracy in detecting duplicate functions.
-
-Our research design follows a multi-method approach, utilizing various
-methodologies to answer our research questions. We have structured the study
-into two phases, each employing different research methods.
+This research aims to identify patterns of software engineering techniques, such as refactoring methods commonly used to address code duplications.
+These patterns will serve as a guide for future contributors in mitigating code duplication issues within this context.
+To investigate our approach, we are developing a tool to identify duplicate code in the Linux kernel, explicitly targeting parts of the AMD Display driver code.
+We are then refactoring portions of this code and submitting patches to maintainers, providing evidence of the tool's accuracy in detecting duplicate functions.
+Our research design follows a multi-method approach, structured into two phases, each employing different research methods.
 
@@
 describes the research methods selected to guide our work. Chapter
-\ref{cha:eval} shows the results obtained from the research methods to evaluate
-our proposed tool. Chapter \ref{cha:results} presents the current research
-stage and the work plan, with the addition of preliminary results obtained.
+\ref{cha:results} shows the first results obtained from the research methods to evaluate
+our proposed tool and presents the current research stage and the work plan,
+with the addition of a discussion about our preliminary results.

### 2024-11-10 conteudo/preliminar/results.tex

--- 2024-06-20 conteudo/preliminar/results.tex
@@
 
-Base Content of the Preliminar Results of the research. TODO.
+\section{Final Remarks}
+
+We have completed Phase 1 of our research, during which we developed our proposed tool and evaluated it using the BigCloneBench dataset and through empirical analyses.
+The results indicate our tool can identify code clone duplications within the Linux kernel.
+
+As the next step, we plan to conduct an ethnographic study as outlined in Section \ref{sec:meteth}.
+To preliminarily assess the potential impact of this research, we applied our systematic approach for mitigating code duplication to two pairs of functions.
+Table \ref{tab:patch} presents the initial results obtained from this process.
+
+\input{conteudo/preliminar/table_patch}
+
+In the first function pair, completing the refactoring requires the expertise of a maintainer for the AMD Display driver.
+This observation is because duplicated code artifacts involve copy-pasted code that imports different configuration files.
+The design of the AMD Display driver configuration files necessitates importing specific configurations at compile time, which prevents a generic approach.
+Consequently, refactoring this pair requires changes in how the AMD Display driver manages configuration files -- a complex task that demands input from an expert or maintainer.
+
+The refactoring of the second function pair, however, was successful. We submitted this refactoring as a patch to the AMD Display driver to obtain feedback from the maintainers.
+\todo[inline]{Atualizar ... tentar colocar algo do feedback que recebeu da revisão do patch}
+
+\input{conteudo/preliminar/cronograma}
+\todo[inline]{Atualizar o cronograma, alocando tempo para evolucao da ferramenta}
+
+Finally, as indicated in our proposed schedule in Table \ref{fig:cronograma}, our upcoming tasks include conducting the complete ethnographic study.
+This step will start with completing our systematic approach to code duplication mitigation.
+Once we have finalized the mitigations, we intend to submit them to the AMD Display driver mailing list to gather insights and feedback from the maintainers regarding the changes we implemented.
+
+\todo[inline]{colocar aqui também algo da evolucao da ferramenta que sera feito até o final do mestrado}

### 2024-11-10 conteudo/tool/functionalities.tex
remove comment about trie, would be a waste and too deep

--- 2024-07-17 conteudo/tool/functionalities.tex
@@
 
-\section{Tool functionalities}
+\section{Tool Functionalities}
 \label{subsec:func}
 
-We propose three main functionalities for the user in the tool, called duplication explorer,
-function information, and duplication relatory. Each of the functionalities receives its 
-adequated parameters to execute, except an optional parameter shared by all of them, which 
-is minimum similarity per query. The minimum similarity per query is not the same as the 
-minimum similarity presented in \ref{subsec:setup}, as it enables a volatile change in the minimum 
-similarity interest in a query-specific context. There is an explanation of each of the 
-main functionalities in this section.
+We propose three main functionalities for the user in the tool, called the \textbf{Duplication Explorer}, \textbf{Function Information}, and \textbf{Duplication Report}.
+Each functionality accepts specific parameters to perform its operations, with an optional parameter shared by all: \textbf{minimum similarity per query}.
+This minimum similarity per query is distinct from the minimum similarity setting introduced in Section \ref{subsec:setup}, as it enables a dynamic adjustment of the similarity threshold within a query-specific context.
+Below is an explanation of each of these main functionalities.
 
-\subsection{Duplication explorer}
+\subsection{Duplication Explorer}
 
-The duplication explorer is the principal functionality of our tool. The functionality's
-purpose is to expose the user to the duplicated function pairs found by our tool. We 
-implement optional filter methods to enable more complex queries, as described below. 
-Figure \ref{fig:explorer_ex} demonstrates the functionality usage.
-
-\begin{itemize}
-	\begin{item}
-		\textbf{Sorting filter:} This filter enables the user to sort the results by the 
-		similarity of the function pairs or by the number of duplicated lines. By default, 
-		the tool sorts by similarity, and the user can pass a sort parameter to change 
-		the sorting method.
-	\end{item}
-
-	\begin{item}
-		\textbf{Limiter filter:} This filter receives a number parameter named limiter 
-		by the user, which limits the number of results shown to the user by this given number.
-	\end{item}
-
-	\begin{item}
-		\textbf{Pattern filter:} This filter makes the explorer ignore every pair 
-		of duplicated functions that do not match a pattern given by the user as a parameter.
-		We say that a function matches a pattern if the string formed by the concatenation 
-		of the relative path of the code file plus the function name contains the pattern 
-		as a substring. The user can define if both functions or at least one of the 
-		functions in the pair need to match the pattern by giving a specific parameter.
-	\end{item}
-\end{itemize}
+The Duplication Explorer is the primary functionality of our tool, designed to present the user with pairs of duplicated functions identified by the tool.
+We implement optional filters to facilitate more complex queries, as described below.
+Figure \ref{fig:explorer_ex} demonstrates an example of this functionality in use.
 
@@
 \includegraphics[scale=0.5]{explorer_example}
-\caption{Explorer Example}
+\caption{Example of the \textit{Duplication Explorer} functionality of the proposed tool in action.}
 \label{fig:explorer_ex}
@@
 
-\subsection{Function information}
+\begin{itemize}
+	\item \textbf{Sorting Filter:} This filter allows the user to sort the results by the similarity score of the function pairs or by the number of duplicated lines.
+By default, the tool sorts results by similarity, but the user can pass a sorting parameter to switch the sorting method.
 
+	\item \textbf{Limiter Filter:} This filter accepts a numerical parameter, called a limiter, which restricts the number of results displayed to the user to the specified number.
+
+	\item \textbf{Pattern Filter:} This filter allows the Duplication Explorer to ignore duplicated function pairs that do not match a user-defined pattern.
+A function matches a pattern if the string is formed by concatenating the relative path of the code file with the function name, which contains the pattern as a substring.
+The user can specify whether both functions in the pair must match the pattern or if matching one of the functions is sufficient.
+\end{itemize}
+
+\subsection{Function Information}
 \label{subsec:functioncommand}
 
-The Function information purpose is to give more in-depth information about a specific 
-function. The functionality receives a function the user is interested in and then 
-informs the user about the relative path, the function name, and the lines the function 
-is defined for the function the user is interested in and for every function that is a 
-duplication of the given function. There can be multiple functions with the same name 
-in a codebase. To distinguish them, we receive a pattern string from the user, and we 
-use the first function to find one that matches the given pattern. We say that a 
-function matches a pattern if the string formed by the concatenation of the relative 
-path of the code file plus the function name contains the pattern as a substring. 
-Figure \ref{fig:function_ex} demonstrates the functionality usage.
+The purpose of the Function Information functionality is to provide detailed information about a specific function.
+The functionality receives a target function from the user and returns information such as the relative path, function name, and line numbers where the function is defined.
+Additionally, it provides similar information for every function that duplicates the given function.
 
@@
 \includegraphics[scale=0.5]{function_example}
-\caption{Function Example}
+\caption{Example of the \textit{Function Information} functionality of the proposed tool in action.}
 \label{fig:function_ex}
@@
 
-\subsection{Duplication relatory}
+Since multiple functions with the same name may exist in a codebase, the user can provide a pattern string to differentiate them.
+A function matches a pattern if the concatenated string of the relative path and function name contains the pattern as a substring.
+Figure \ref{fig:function_ex} demonstrates this functionality in action.
 
-The duplication relatory purpose is to give an overview of the input codebase. 
-The functionality extracts how many lines of duplicated code per folder in the 
-codebase and presents it to the user in a reasonable format. To extract the 
-information, we use the Code Duplication Database to list the duplicated functions 
-and the temporary code to know how many lines the duplication functions have. 
-Figure \ref{fig:relatory_ex} demonstrates the functionality usage. 
-(DO I TALK ABOUT THE TRIE USAGE TO MAINTAIN THE HIERARCHICAL TREE STRUCTURE? I THINK NOT)
+\subsection{Duplication Report}
+
+The purpose of the Duplication Report functionality is to provide an overview of duplicated code within the input codebase.
+This functionality calculates the number of duplicated lines per folder in the codebase and presents the information to the user in a readable format.
 
@@
 \includegraphics[scale=0.5]{relatory_example}
-\caption{Function Example}
+\caption{Example of the \textit{Duplication Report} functionality of the proposed tool in action.}
 \label{fig:relatory_ex}
 \end{figure}
+
+The tool utilizes the Code Duplication Database to list duplicated functions and the temporary codebase to determine the line count for these functions to generate this report.
+Figure \ref{fig:relatory_ex} demonstrates an example of this functionality in action.

### 2024-11-10 conteudo/tool/tool.tex

@@
 
-The tool operates in two main steps. The first step, called *Setup*, is
+The tool operates in two main steps. The first step, called \textbf{Setup}, is
 executed only once for each codebase. This step processes the codebase by
 extracting functions from code files and identifying code clone duplications.
-The second step, called *Query Responder*, utilizes the information obtained in
-the Setup phase to execute the tool’s functionalities as needed. Detailed
+The second step, called \textbf{Query Responder}, utilizes the information obtained in
+the Setup phase to execute the tool functionalities as needed. Detailed
 descriptions of the Setup and Query Responder processes can be found in

### 2024-11-10 conteudo/background/linux.tex
Reviewing the final chapter and applying other adjustments
fix data from amd size

@@
 object of study as we aim to understand and mitigate code duplication in the
-Linux kernel. The AMD Display Driver subsystem contains X lines of code, Y
-functions/structs, and Z code files, making it a considerable component within
-the Linux kernel. Like other subsystems, discussions about the AMD Display
-Driver are conducted on its designated mailing list (FIND MAILING LIST AND CITE
-HERE).
+Linux kernel. The AMD Display Driver subsystem contains more then 391000 lines 
+of code and more then 1089 code files
+\footnote{ 
+Data extracted directly from the Linux repository with the help of the cloc tool.
+\citep{cloc}
+}, 
+making it a considerable component within the Linux kernel. Like other 
+subsystems, discussions about the AMD Display Driver are conducted on its 
+designated mailing list.
+
+\todo[inline]{FIND MAILING LIST AND CITE HERE}

### 2024-11-11 conteudo/background/linux.tex
changes on linux

@@
 subsystems, discussions about the AMD Display Driver are conducted on its 
-designated mailing list.
+designated mailing list, which can be accessed here: 
+\url{https://lists.freedesktop.org/mailman/listinfo/amd-gfx}.
 
-\todo[inline]{FIND MAILING LIST AND CITE HERE}
+\todo[inline]{FIND MAILING LIST AND CITE HERE (DONE?) }

### 2024-11-11 conteudo/preliminar/results.tex
Add tool in cronogram
add commentary about tool improv

@@
 \input{conteudo/preliminar/cronograma}
-\todo[inline]{Atualizar o cronograma, alocando tempo para evolucao da ferramenta}
+\todo[inline]{Atualizar o cronograma, alocando tempo para evolucao da ferramenta (DONE?) }
 
-Finally, as indicated in our proposed schedule in Table \ref{fig:cronograma}, our upcoming tasks include conducting the complete ethnographic study.
-This step will start with completing our systematic approach to code duplication mitigation.
-Once we have finalized the mitigations, we intend to submit them to the AMD Display driver mailing list to gather insights and feedback from the maintainers regarding the changes we implemented.
+Finally, as indicated in our proposed schedule in Table \ref{fig:cronograma}, 
+our upcoming tasks include conducting the complete ethnographic study.
+This step will start with completing our systematic approach to code 
+duplication mitigation. Once we have finalized the mitigations, we intend 
+to submit them to the AMD Display driver mailing list to gather insights 
+and feedback from the maintainers regarding the changes we implemented.
 
-\todo[inline]{colocar aqui também algo da evolucao da ferramenta que sera feito até o final do mestrado}
+Our upcoming tasks also include implementing improvements to our proposed tool. 
+While the tool’s current state allowed us to conduct the experiments described in 
+Chapter \ref{cha:method},
+it is not yet adequately user-friendly or manageable for external users. 
+To achieve our complementary goal of providing a solution for code duplication 
+in the kernel, we plan to improve the documentation, refine the code structure, 
+and address any issues we encounter along the way, providing the Linux community 
+with a way to tackle the issue.
+
+\todo[inline]{colocar aqui também algo da evolucao da ferramenta que sera feito até o final do mestrado (DONE?) }

### 2024-11-11 conteudo/eval/eval.tex
add initial text of preliminar

@@
 \todo[inline]{Juntei os capítulos; veja se consegue colocar algo aqui no início; não sei se final remarks é o melhor nome para a segunda secao aqui}
+
+To finalize this report, we present the current state of our research. As of the 
+time of writing, we have finished \textbf{Phase I} of this research, 
+which includes the implementation of 
+the proposed tool and the evaluation of the tool through both formal literature 
+and empirical analyses. This chapter begins with the presentation of the results 
+from the tool evaluation. Next, we present a preliminary experiment we are currently 
+conducting to assess the viability of our Ethnographic Study. Finally, we outline 
+the next steps in this research.
 

### 2024-11-11 conteudo/background/linux.tex
add little preambulo to introduce chapter

@@
 \todo[inline]{Adicionar um "preambulo" antes da seções do capítulo}
+
+This chapter presents an overview of fundamental topics needed to understand 
+the research, such as Linux Kernel, refactoring, types of code duplication, 
+code duplication detection methods presented in the literature and ways to 
+evaluating these methods. This foundation will make it easier to grasp the 
+concepts discussed in the research.
 

### 2024-11-12 conteudo/preliminar/results.tex
add initial feedback information

@@
 
-The refactoring of the second function pair, however, was successful. We submitted this refactoring as a patch to the AMD Display driver to obtain feedback from the maintainers.
+The refactoring of the second function pair, however, was successful. 
+We submitted this refactoring as a patch to the AMD Display driver to obtain feedback 
+from the maintainers.
+
+The process of sending this patch did not yield satisfactory results in terms of 
+interaction with the AMD Display driver maintainers. We initially submitted the 
+patch on August 9th, 2024, but did not receive any replies. As a result, we had 
+to resend it on October 9th, 2024, and finally received an initial response on 
+November 3rd, 2024, from our point of contact at the AMD Display driver. 
+The response time from our interactions with the AMD Display driver is making 
+us reconsider the viability of our initial plans.
+
+Regarding the patch content, only minor changes were requested to align with code 
+style and best practices that were initially unknown to us. More details of the 
+patch can be found in Appendix (CREATE APPENDIX AND ADD HERE).
+
 \todo[inline]{Atualizar ... tentar colocar algo do feedback que recebeu da revisão do patch}

### 2024-11-12 conteudo/preliminar/results.tex
add appendix for feedback
fix typo

@@
 
-The process of sending this patch did not yield satisfactory results in terms of 
+The process of sending the patch did not yield satisfactory results in terms of 
 interaction with the AMD Display driver maintainers. We initially submitted the 
@@
 style and best practices that were initially unknown to us. More details of the 
-patch can be found in Appendix (CREATE APPENDIX AND ADD HERE).
+patch can be found in Appendix \ref{app:feedback}.
 

### 2024-11-13 conteudo/apendices/gray.tex
add appendix of gray lit

\chapter{Gray Literature tools explored}

\label{app:gray}

This appendix contains the the list of gray literature tools 
explored in this initial exploratory research mentioned
in the Introduction. To test this tools, we created a simple 
toy codebase with multiples duplicated code in C.

\paragraph{Copy/Paste Detector (CPD)}

\

\textbf{Is free to use:} Yes.

We installed and tested it. It is a pretty primitive tool 
with little practical usage, and our experience using it 
was not good. It works by finding textual matches between files.

Access from \url{https://pmd.github.io/pmd/pmd_userdocs_cpd}.

\paragraph{CppDepend}

\

\textbf{Is free to use:} No. There is a free trial periody.

Tools with paywalls do not meet our requirements, 
so we did not investigate the tool further for this reason.

Access from 
\url{https://cppdepend.com/blog/tracking-the-hidden-duplicate-code-in-a-c-code-base/}.

\paragraph{Duplo}

\

\textbf{Is free to use:} Yes.

We installed and tested it. It is a pretty primitive tool 
with little practical usage, and our experience using it was good. 
It works by finding textual matches between files.

Access from
\url{https://github.com/dlidstrom/Duplo}.

\paragraph{Duploq}

\

\textbf{Is free to use:} Yes.

It is a front-end application for Duplo only, designed to 
enable a more intuitive and configurable user experience. 
We did not test it further, as it has the same capabilities 
as Duplo itself.

Access from
\url{https://github.com/duploq/duploq}

\paragraph{Simian}

\

\textbf{Is free to use:} No. 
There is a license for academic research and education use.

Tools with paywalls do not meet our requirements, 
so we did not investigate the tool further for this reason.

Access from
\url{https://simian.quandarypeak.com/}.

\paragraph{Code Climate}

\

\textbf{Is free to use:} No. 
There is a license for open source use.

Tools with paywalls do not meet our requirements, so this 
tool was initially skipped. Out of curiosity, we tried it 
ourselves, and it did not work correctly with our toy
dataset of code duplications.

Access from
\url{https://codeclimate.com/}.

\paragraph{Visual Detection of Duplicated Code}

\

\textbf{Is free to use:} In principle, yes.

We found this tool in a paper and found it interesting. 
Unfortunately, we were not able to find an executable 
version of this tool for testing.

Access from
\url{https://www.researchgate.net/publication/2866183_Visual_Detection_of_Duplicated_Code}.

\paragraph{duplicate-code-detection-tool}

\

\textbf{Is free to use:} Yes.

It is the only tool we found that is not based on textual 
matching. It looks pretty good for small codebases but 
performs poorly when used in complex codebases. This tool 
provided significant inspiration for this work, as 
mentioned in Chapter \ref{cha:tool}.

Access from
\url{https://github.com/platisd/duplicate-code-detection-tool}

### 2024-11-14 conteudo/methods/methods.tex
fix some verbal times, i believe it is correct now

@@
 
-We apply two independent methods and perform triangulation on the results.
-The first method evaluates the proposed tool using an approach from the literature, comparing it against the BigCloneBench dataset \citep{bigclonebench}.
-The second method is an empirical analysis of a subset of functions within the AMD Display driver that our tool identified as duplicates.
-After completing these two methods, we will triangulate the findings to analyze the results comprehensively.
-We chose triangulation because relying solely on the literature-based method has limitations and may not be entirely reliable \citep{bigfail, litreview}.
+We applied two independent methods and perform triangulation on the results. 
+The first method evaluates the proposed tool using an approach from the literature, 
+comparing it against the BigCloneBench dataset \citep{bigclonebench}. 
+The second method is an empirical analysis of a subset of functions within the 
+AMD Display driver that our tool identified as duplicates. After completing these two 
+methods, we triangulated the findings to analyze the results comprehensively. 
+We chose triangulation because relying solely on the literature-based method has 
+limitations and may not be entirely reliable \citep{bigfail, litreview}.
 
@@
 
-Ethnography is a recognized qualitative empirical method for understanding people, their cultures, and their work practices \citep{bookethno}.
-It allows researchers to gain insights into values, beliefs, and ideas from the perspective of community members \citep{ethnosoft}.
-In our research, refactoring code duplications without community interaction would prevent us from understanding whether Linux maintainers would accept our approach or identify their concerns.
-Thus, we used ethnography, positioning ourselves as Linux kernel contributors focused on removing identified code duplications.
+Ethnography is a recognized qualitative empirical method for understanding people, 
+their cultures, and their work practices \citep{bookethno}. It allows researchers 
+to gain insights into values, beliefs, and ideas from the perspective of community 
+members \citep{ethnosoft}. In our research, refactoring code duplications without 
+community interaction would prevent us from understanding whether Linux maintainers 
+would accept our approach or identify their concerns. Thus, we will use ethnography, 
+positioning ourselves as Linux kernel contributors focused on removing identified 
+code duplications.
 
@@
 As with the empirical analysis, evaluating every code duplication in the Linux kernel is impractical. 
-Thus, we limited the experiment to the AMD Display driver and used random sampling of duplicated functions inferred by our tool to initiate the mitigation approach in Section \ref{subsec:pipeline}.
-We chose a sample size of 30 duplicated functions, as studies indicate this size provides statistical significance \citep{sample1, sample2}.
+Thus, we will limit the experiment to the AMD Display driver and used random sampling of duplicated functions
+inferred by our tool to initiate the mitigation approach in Section \ref{subsec:pipeline}. 
+We will choose a sample size of 30 duplicated functions, as studies indicate this size provides 
+statistical significance \citep{sample1, sample2}.
 
-For random selection, we restricted eligibility to function pairs with a similarity metric of 100\%. With this criterion, we randomly chose functions from eligible pairs uniformly.
+For random selection, we will restrict the eligibility to function pairs with a similarity metric of 100\%. 
+With this criterion, we randomly chose functions from eligible pairs uniformly.
 

### 2024-11-14 conteudo/tool/tool.tex
add arkanjo as tool name

@@
 
-Our proposed tool is a Command Line Interface (CLI) application designed to
+Arkanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
 help developers identify code clone duplication at the function level.
@@
 
-The tool has been implemented and is available for reference at
-\url{https://github.com/LipArcanjo/test-duplication-code}.
+The Arkanjo tool has been implemented and is available for reference at
+\url{https://github.com/LipArcanjo/arkanjo}.

### 2024-11-14 conteudo/introduction/intro.tex
choose to put in appendix the list of gray
add appendix of gray lit
add arkanjo and change proposed tool to arkanjo in multiples places

@@
 kernel context. 
+The list of gray literature tools examined in this exploratory research 
+is available in Appendix A \ref{app:gray}.
 
 \section**{Research Design}
+
+\label{sec:intresearch}
 
@@
 and a review of refactoring methods used throughout this research. Chapter
-\ref{cha:tool} presents our proposed tool to detect code duplication in the
+\ref{cha:tool} presents Arkanjo, our proposed tool to detect code duplication in the
 Linux kernel, describing all the main components.  Chapter \ref{cha:method}
@@
 \ref{cha:results} shows the first results obtained from the research methods to evaluate
-our proposed tool and presents the current research stage and the work plan,
+our tool and presents the current research stage and the work plan,
 with the addition of a discussion about our preliminary results.

### 2024-11-14 conteudo/tool/functionalities.tex
add arkanjo and change proposed tool to arkanjo in multiples places

@@
 \includegraphics[scale=0.5]{explorer_example}
-\caption{Example of the \textit{Duplication Explorer} functionality of the proposed tool in action.}
+\caption{Example of the \textit{Duplication Explorer} functionality of the Arkanjo tool in action.}
 \label{fig:explorer_ex}
@@
 \includegraphics[scale=0.5]{function_example}
-\caption{Example of the \textit{Function Information} functionality of the proposed tool in action.}
+\caption{Example of the \textit{Function Information} functionality of the Arkanjo tool in action.}
 \label{fig:function_ex}
@@
 \includegraphics[scale=0.5]{relatory_example}
-\caption{Example of the \textit{Duplication Report} functionality of the proposed tool in action.}
+\caption{Example of the \textit{Duplication Report} functionality of the Arkanjo tool in action.}
 \label{fig:relatory_ex}

### 2024-11-15 conteudo/eval/eval.tex
fix missing end of paragraph

@@
 
-The results in Table \ref{tab:emp} show that as the similarity metric decreases, there is a notable increase in function pairs that are not duplications within the Linux kernel.
-For similarity ranges of 79\%-81\% or higher, there is a high percentage of code duplications.
-At 69\%-71\% similarity, the duplication rate drops to 50%.
-At 59\%-61%, it further decreases to 20%, and below 49\%-51%, no code duplications were found.
+The results in Table \ref{tab:emp} show that as the similarity metric decreases, 
+there is a notable increase in function pairs that are not duplications within 
+the Linux kernel. For similarity ranges of 79\%-81\% or higher, there is a high 
+percentage of code duplications. At 69\%-71\% similarity, the duplication rate 
+drops to 50\%. At 59\%-61\%, it further decreases to 20\%, and below 49\%-51\%, 
+no code duplications were found.
 

### 2024-11-15 conteudo/methods/methods.tex
remove TODO's

@@
-\todo[inline]{Depois, veja melhor o tempo verbal ...
-tem coisas que coloquei no presente, porque ainda está fazendo ou irá fazer; tem outras coisas já planejadas e feitas que o texto ficou no passado ...
-veja se está fluindo bem e como deixar mais uniforme o tempo verbal aqui e nas outras partes do texto, tentando deixar claro que já está feito e o que será feito depois}
-
 We employ a multi-method research approach, using different methodologies to achieve our goals.

### 2024-11-15 conteudo/eval/eval.tex
remove TODO's
reforce that we do not expect good results in T3/T4 types

@@
 \en
-
-\todo[inline]{Juntei os capítulos; veja se consegue colocar algo aqui no início; não sei se final remarks é o melhor nome para a segunda secao aqui}
 
@@
 
-In this section, we present the test results of our tool performance, following the methodology described in Chapter \ref{cha:method}.
-First, we provide the results obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}, which serves as a standard reference in the literature \citep{litreview}.
-Second, we discuss the results from our empirical analysis.
-Finally, we triangulate the data, analyzing our findings regarding the tool performance.
-
-\subsection{Tool Results on the BigCloneBench Dataset}
+In this section, we present the test results of our tool performance, following the 
+methodology described in Chapter \ref{cha:method}. First, we provide the results 
+obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}, which 
+serves as a standard reference in the literature \citep{litreview}. Second, we 
+discuss the results from our empirical analysis. Finally, we triangulate the data, 
+analyzing our findings regarding the tool performance. 
+As a reminder, we only expect our tool to yield good results for Type-1 and Type-2,
+as mentioned in \ref{subsec:code_finder}.
 

### 2024-11-15 conteudo/apendices/gray.tex
fix typo
change to related tools for some parts

@@
-\chapter{Gray Literature tools explored}
+\chapter{Related tools explored}
 
@@
 
-This appendix contains the the list of gray literature tools 
+This appendix contains the the list of related tools 
 explored in this initial exploratory research mentioned
@@
 ourselves, and it did not work correctly with our toy
-dataset of code duplications.
+codebase of code duplications.
 

### 2024-11-15 conteudo/eval/eval.tex
change to related tools for some parts

@@
 which includes the implementation of 
-the proposed tool and the evaluation of the tool through both formal literature 
-and empirical analyses. This chapter begins with the presentation of the results 
-from the tool evaluation. Next, we present a preliminary experiment we are currently 
-conducting to assess the viability of our Ethnographic Study. Finally, we outline 
-the next steps in this research.
+the proposed tool and the evaluation of the tool through both a method from the 
+literature and an empirical analyses. This chapter begins with the presentation of 
+the results from the tool evaluation. Next, we present a preliminary experiment 
+we are currently conducting to assess the viability of our Ethnographic Study. 
+Finally, we outline the next steps in this research.
 

### 2024-11-15 conteudo/introduction/intro.tex
change to related tools for some parts

@@
 Given the unsuccessful results of our search in the formal literature, we
-turned to gray literature for solutions but were also met with limited success.
-The best solutions we found in the gray literature were primitive, merely
+turned to explore related tools not present in literature for solutions, but were also met 
+with limited success.
+The best solutions we found were primitive, merely
 identifying every pair of code files that are duplications. The code files in
@@
 kernel context. 
-The list of gray literature tools examined in this exploratory research 
-is available in Appendix A \ref{app:gray}.
+The list of related tools examined in this exploratory research 
+is available in Appendix \ref{app:gray}.
 

### 2024-11-15 conteudo/introduction/intro.tex
fix formal literature

@@
 code artifacts, whether they are duplications of each other. However, the
-formal literature on code duplication mainly focuses on optimizing the accuracy
+literature on code duplication mainly focuses on optimizing the accuracy
 of these tools for this specific task. These solutions do not address our
@@
 
-Given the unsuccessful results of our search in the formal literature, we
+Given the unsuccessful results of our search in the literature, we
 turned to explore related tools not present in literature for solutions, but were also met 
@@
 \begin{figure}[h]
-\includegraphics[scale=0.7]{research_design}
+\includegraphics[scale=0.46]{research_design}
 \caption{Diagram of the research design.}
@@
 triangulation method, comparing its results with established code duplication
-databases referenced in the literature, and conducting an empirical analysis on
+database referenced in the literature, and conducting an empirical analysis on
 a randomly selected set of files from the AMD Display driver, which our tool

### 2024-11-15 conteudo/methods/methods.tex
fix formal literature

@@
 
-\section{Evaluation of the Tool by Formal Literature}
+\section{Evaluation of the Tool by the BigCloneBench dataset}
 
@@
 
-Due to the limitations of the BigCloneBench dataset, as noted by \cite{bigfail}, and its lack of suitability for our context (Java dataset vs.
-C-based Linux kernel), we complemented the formal evaluation with an empirical study of the tool inferred duplications.
+Due to the limitations of the BigCloneBench dataset, as noted by \cite{bigfail}, and its lack of suitability 
+for our context (Java dataset vs. C-based Linux kernel), we complemented the BigCloneBench evaluation with 
+an empirical study of the tool inferred duplications.
+
 %

### 2024-11-16 conteudo/preliminar/results.tex
remove TODO's
redo patch info

--- 2024-11-11 conteudo/preliminar/results.tex
@@
 
-The refactoring of the second function pair, however, was successful. We submitted this refactoring as a patch to the AMD Display driver to obtain feedback from the maintainers.
-\todo[inline]{Atualizar ... tentar colocar algo do feedback que recebeu da revisão do patch}
+The refactoring of the second function pair, however, was successful. 
+We submitted this refactoring as a patch to the AMD Display driver to obtain feedback 
+from the maintainers.
+
+The process of sending the patch was more time consuming then the expected . We 
+initially submitted the patch on August 9th, 2024, but did not receive any 
+replies. As a result, we had to resend it on October 9th, 2024, and finally 
+received an initial response on 
+November 3rd, 2024. 
+While talking to our point of contact for the AMD Display driver, we discovered 
+new information about the driver's processes. Some of the code in the driver is 
+shared across all operating systems that support the implemented GPUs. This creates 
+a challenging process within AMD to format and submit changes across the supported 
+systems while simultaneously applying measures to mitigate errors.
+Additionally, we discovered that duplicated code isn’t seen as negatively by AMD 
+display driver developers as it is in the literature. One example shared with us 
+is that duplicated code makes GPU driver code more independent, letting developers 
+make changes to a GPU without worrying about testing compatibility with other GPUs, 
+which helps save a lot of time and effort.
+
+Regarding the patch content, only minor changes were requested to align with code 
+style and best practices that were initially unknown to us. More details of the 
+patch can be found in Appendix \ref{app:feedback}.
 
 \input{conteudo/preliminar/cronograma}
-\todo[inline]{Atualizar o cronograma, alocando tempo para evolucao da ferramenta (DONE?) }
 
@@
 with a way to tackle the issue.
-
-\todo[inline]{colocar aqui também algo da evolucao da ferramenta que sera feito até o final do mestrado (DONE?) }

### 2024-11-18 conteudo/introduction/intro.tex
gray -> grey

@@
 in the previous section cannot be fully addressed by existing tools in
-either the formal or gray literature. To fill this gap, we propose an approach
+either the formal or grey literature. To fill this gap, we propose an approach
 to identify and mitigate code duplications specifically within the Linux

### 2024-11-18 conteudo/preliminar/results.tex
minor changes on the feedback process

--- 2024-11-12 conteudo/preliminar/results.tex
@@
 The refactoring of the second function pair, however, was successful.
-We submitted this refactoring as a patch to the AMD Display driver to obtain feedback 
-from the maintainers.
+We submitted this refactoring as a patch to the AMD Display driver to obtain feedback from the maintainers.
+The process of sending the patch took more time than expected.
+We initially submitted the patch on August 9th, 2024, but we did not receive replies.
+As a result, we had to resend it on October 9th, 2024 and received an initial response on November 3rd, 2024.
 
-The process of sending the patch did not yield satisfactory results in terms of 
-interaction with the AMD Display driver maintainers. We initially submitted the 
-patch on August 9th, 2024, but did not receive any replies. As a result, we had 
-to resend it on October 9th, 2024, and finally received an initial response on 
-November 3rd, 2024, from our point of contact at the AMD Display driver. 
-The response time from our interactions with the AMD Display driver is making 
-us reconsider the viability of our initial plans.
+We understood new details about the driver evolution processes in discussions with our partner from the AMD Display driver.
+Some of the code in the driver is shared across all operating systems that support the implemented GPUs.
+This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors.
+
+Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering.
+One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others.
+This approach helps save significant time and effort.
 
@@
 
-\todo[inline]{Atualizar ... tentar colocar algo do feedback que recebeu da revisão do patch}
-
 \input{conteudo/preliminar/cronograma}
-\todo[inline]{Atualizar o cronograma, alocando tempo para evolucao da ferramenta (DONE?) }
 

### 2025-04-24 conteudo/background/linux.tex
remove TODO's
add comments on todos for the dissertation text

@@
 \en
-
-\todo[inline]{Adicionar um "preambulo" antes da seções do capítulo}
 
@@
 
-\todo[inline]{FIND MAILING LIST AND CITE HERE (DONE?) }
+\todo[inline]{Require an introduction about IIO drivers.}
+
+\todo[inline]{Not exactly on this section, but I need to mention GIT, as one of the similarity methods is based on the GIT command.}

### 2025-04-24 conteudo/background/refactor.tex
add comments on todos for the dissertation text

--- 2024-07-17 conteudo/background/refactor.tex
@@
 
-Fowler defines refactoring in his book as "a controlled technique for improving the design of an existing code base. 
-Its essence is applying a series of small behavior-preserving transformations, each of which is too small to be worth 
-doing. However, the cumulative effect of each of these transformations is quite significant." \citep{refactorbook}. 
-The elimination of detected code clone duplication is a refactoring task that necessitates the 
-application of refactoring methods. Fowler presents multiple refactoring methods such as the 
-\textit{parameterize method}, the \textit{replace parameter with explicit methods} and 
-the \textit{replace parameter with method} \citep{refactorbook}. 
-There is a brief explanation of the methods used in our research below.
+\todo[inline]{Reconsider if I need this function anymore.}
 
-\subsection{Parameterize method}
+Fowler defines refactoring as "a controlled technique for improving the design
+of an existing code base. Its essence is applying a series of small
+behavior-preserving transformations, each of which is too small to be worth
+doing. However, the cumulative effect of each of these transformations is quite
+significant" \citep{refactorbook}. The elimination of detected code clones and
+duplication is a refactoring task that requires the application of refactoring
+methods. Fowler presents several refactoring techniques, including
+\textit{parameterize method}, \textit{replace parameter with explicit methods},
+and \textit{replace parameter with method} \citep{refactorbook}. Below is a
+brief explanation of the methods used in our research.
 
-TODO. I EXPECT TO FILL THIS SECTION AS THE WORK EVOLVES.
+\begin{figure}[ht]
+\includegraphics[scale=0.36]{refactor/parameterize}
+\caption{Example of Parameterize Method usage.}
+\label{fig:parameter}
+\end{figure}
+
+\begin{figure}[ht]
+\includegraphics[scale=0.36]{refactor/extract}
+\caption{Example of Extract Method usage.}
+\label{fig:extract}
+\end{figure}
+
+\paragraph{Parameterize Method.}
+
+The Parameterize Method involves replacing multiple similar code blocks with a
+single function. This function incorporates parameters to account for the
+different behaviors across similar code snippets, allowing it to adjust
+functionality based on the parameter values \citep{refactorbook}. Figure
+\ref{fig:parameter} illustrates an example of the Parameterize Method.
+
+\paragraph{Extract Method.}
+
+The Extract Method entails breaking down a complex code artifact into multiple,
+simpler methods or functions. One of the primary advantages of this method is
+that the newly created functions can be reused throughout the project. By
+contrast, a complex code artifact might not be entirely reusable, potentially
+resulting in further code duplication \citep{refactorbook}. Figure
+\ref{fig:extract} demonstrates an example of the Extract Method.

### 2025-04-24 conteudo/eval/eval.tex
add comments on todos for the dissertation text

@@
 \en
+
+\todo[inline]{As it is primarly results, everything should be remade, the structure of the files for this section is bad.}
+
+\todo[inline]{Remember the typos on the tables. }
+
+\todo[inline]{Idea: The research methods should be 3 phases, one of the tool creation/testing, one for the my experiences submiting a path, and the last for the experiences of the students on the course.
+Maybe split in four to split the tool creation for the tool testing.}
 

### 2025-04-24 conteudo/methods/methods.tex
add comments on todos for the dissertation text

@@
 limitations and may not be entirely reliable \citep{bigfail, litreview}.
+
+\todo[inline]{The focus of the above section should be a little more on the tool, the introduction about mitigating duplications is good enough.
+A little about the fact that we will test on IIO later maybe is good.
+}
+
+\todo[inline]{I SHOULD mention that the test on the tool were made using the gensim similarity method and not the git method.}
 
@@
 
+\todo[inline]{I just mention the problem, maybe a more explictly mention of the problem here or on the background would be good.}
+
 Due to the limitations of the BigCloneBench dataset, as noted by \cite{bigfail}, and its lack of suitability 
@@
 
+\todo[inline]{Take care: Big focus on AMD display driver and the tone of the writing supports the focus. }
+
 \section{Ethnographic Study}
@@
 \label{sec:meteth}
+
+\todo[inline]{THE WHOLE SECTION NEEDS TO BE REMADE.}
 

### 2025-04-24 conteudo/tool/tool.tex
add comments on todos for the dissertation text

@@
 \url{https://github.com/LipArcanjo/arkanjo}.
+
+\todo[inline]{I need to rename the Setup as Preprocessor step.}
+
+\todo[inline]{Update the Function parser description of the C parser and Java parser}
+
+\todo[inline]{Add the other similarity method added on the tool using git for the Code Duplication Finder.}
+
+\todo[inline]{Update diagram and functionalites pictures (do not need sudo anymore!).}
+
+\todo[inline]{Big maybes: Describe more how the functionalites are implemented, as the focus shifted for the tool as of now.}

### 2025-04-27 conteudo/tool/architecture.tex
fix architecture diagram
reforce that we do not expect good results in T3/T4 types
update some things on the tool

--- 2024-11-09 conteudo/tool/architecture.tex
@@
 
-The Setup step consists of the Function Breaker and the Code Duplication Finder.
-The pipeline of this step works as follows: the Function Parser receives the codebase the user is interested in, extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files.
-The Code Duplication Finder iterates over every pair of files in the temporary codebase, checks if they are code clones, and, if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple \textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized, which in our case is the cosine similarity explained in \ref{sec:similarity}.
-Finally, the Query Responder consumes the temporary codebase and the Code Duplication Database to extract duplicated functions-related information per user request.
-Figure \ref{fig:diagram} illustrates the tool pipeline.
-An explanation of how each component works is provided in this section.
+The Preprocessor contains two main components, the Function Breaker and the Code Duplication Finder. 
+The flow of how the Preprocessor works is as follow: Function Parser receives the codebase the user is interested in,
+extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files. 
+The Code Duplication Finder iterates over every pair of files in the Temporary Codebase, checks if they are code clones, and, 
+if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple 
+\textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates 
+of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized in the Code Duplication Finder. ,
+
+The Query Responder consumes the Temporary Codebase and the Code Duplication Database to extract duplicated functions-related information per user request.
+
+Figure \ref{fig:diagram} illustrates the tool architecture. An explanation of how each component works is provided in this section.
 
 \begin{figure}
-\includegraphics{diagrama_mestrado}
+\includegraphics[scale=0.5]{diagrama_mestrado}
 \caption{Architecture diagram demonstrating the relationship between the tool components.}
@@
 
-\begin{itemize}
-	\item \textbf{C}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in the code context and some of the code lines before the definition of the code block.
-According to the C programming language's grammar, functions are defined in a global context at depth 0.
-There are other elements with the same depth as functions, and we differentiate them by analyzing if the first non-empty character before the curly brackets of the code block is a closing parenthesis.
-This pattern uniquely identifies functions within C code.
+\subsubsection{C Programming Language}
 
-	\item \textbf{Java}:  Given a source code file, we extract every code block from it, maintaining the depth of the code block in the code context and some of the code lines before the definition of the code block.
-According to Java's grammar, functions are defined inside classes.
-Functions can exist at different depths, as Java accepts inner classes.
-We differentiate functions from other elements by analyzing if the first non-empty character before the curly brackets of the code block is a closing parenthesis.
-A challenge with this approach in Java is that other elements, such as for-loops, can follow this pattern, necessitating additional filters.
-We chose to analyze functions at depth 1, which ignores any other element that is a code block and has a parenthesis as the first non-empty character before the curly brackets.
-This choice means the extractor does not find functions declared inside inner classes, but this limitation is not an issue in this work, as there are no inner classes in BigCloneBench \citep{bigclonebench}, the only Java codebase explored in this work.
-\end{itemize}
+For a source file in the C programming language, we extract the code blocks with depth 0, as per the programming 
+language's grammar, functions are defined in a global context at depth 0. To extract the code blocks, we firstly need to 
+identify the portions of code that are comments or \textit{\#define} directives and we mark them as irrelevant, after that, 
+it is simple work of finding the curly brackets that are in depth 0, and with that we find the body of the functions.
+
+To extract the header of the functions, we start at the open curly brackets and go back in the source file consireding 
+the programming language grammar  and the portions of code that have been marked as irrelevant because of comments and 
+\textit{\#define} directives.
+
+There are other elements in depth 0 that are not functions, we analyze if the code blocks is of a function at the same 
+we try to extract the header of the code block.
+
+\subsubsection{Java Programming Language}
+
+For the Java programming language, we have decided to implemente a simpler approach that does not support 
+the programming language completely, as the only Java codebase explored in this work is the BigCloneBench \citep{bigclonebench} and 
+the approach implemented cover the needs of this codebase.
+
+For a source file in Java, we extract the code blocks with depth 1 and some code lines before the definition of the code block. 
+Then, we check if the code blocks are functions by analyzing if the first non-empty character before the curly brackets of 
+the code block is a closing parenthesis.
+
+As per Java's grammar, functions are defined inside classes and can exist at different depths, as Java accepts inner classes, 
+that means that this approach does not work with inner classes. The challenge to implement a more generic approach 
+is that other elements of depth 2 or deeper, such as for-loops, can follow the closing parenthesis pattern, necessitating 
+a better analysis to different these elements. The BigCloneBench \citep{bigclonebench} does not contains inner classes.
 
@@
 
-The Code Duplication Finder iterates through every pair of source code files in the temporary codebase, representing functions from the input codebase.
-For each pair of files, we execute a code duplication detection method that computes a metric measuring how similar the pair of files is, which we refer to as similarity.
-If the similarity is greater than or equal to the minimum similarity threshold (a parameter provided by the user during the Setup step), we store this pair of functions along with its similarity in the Code Duplication Database.
+\label{subsec:code_finder}
 
-For the code duplication detection method, we treat the source code files as text and apply the TF-IDF vector embedding method implemented by the Gensim library \citep{gensim}, then compute cosine similarity as the similarity metric.
-This method was chosen for its claimed performance, programming language independence, and the fact that it does not require compilable code, which is expected in the temporary codebase, as it does not contain complete code artifacts.
-\cite{platistool} implemented this code duplication detection method and released it under the MIT license \citep{mitlicense}.
-We adapted his implementation to fit our tool's expected input and output formats.
-Switching the duplication detection method is feasible, similar to how we adapted Plates' implementation.
+The Code Duplication Finder iterates through every pair of source code files in the temporary codebase, representing 
+functions from the input codebase. For each pair of files, we execute a code duplication detection method that computes 
+a metric measuring how similar the pair of files is, which we refer to as similarity. 
+If the similarity is greater than or equal to the minimum similarity threshold 
+(a parameter provided by the user during the Setup step), we store this pair of functions along with its 
+similarity in the Code Duplication Database.
+
+We implemented two methods for code duplication detection in the tool, in which the user can choose what to use.
+The first method is based in text similarity, the second is based in simpler method based on the number of equal lines.
+The experiments and tests on this research were done using the text similarity method, the purpose of the 
+second method is to demonstrate that the tool can have multiple duplication detection methods.
+
+\subsubsection{Text Similarity Method}
+
+For the text similarity method, we treat the source code files 
+as text and apply the TF-IDF vector embedding method implemented by the 
+Gensim library \citep{gensim}, then compute cosine similarity as the 
+similarity metric.
+
+This method was chosen for its claimed performance, 
+programming language independence, and the fact that it does not require 
+compilable code, which is expected in the temporary codebase, as it does 
+not contain complete code artifacts. We expect our method to yield good 
+results for Type-1 and Type-2 code duplication types while performing 
+poorly for the other types. It is possible to change the code duplication 
+detection method to one of the state-of-the-art methods if needed.
+
+\cite{platistool} implemented this code duplication detection method and 
+released it under the MIT license \citep{mitlicense}. We adapted his 
+implementation to fit our tool's expected input and output formats. 
+Switching the duplication detection method is feasible, similar to how 
+we adapted Plates' implementation.
+
+\subsubsection{Difference Method}
+
+For the difference method, 
+we implemented a method considering the number of exactly equal lines between two functions.
+For two functions \textit{function1} and \textit{function2}, let $NL_1$ be the number of lines in \textit{function1},
+$NL_2$ be the number of lines in \textit{function2}, and $EQ$ be the number of equal lines between \textit{function1} and
+\textit{function2}, we compute the similarity with the formula:
+
+$$\frac{2 \cdot EQ}{NL_1 + NL_2}$$
+
+This code duplication method is considerably simpler and more explainable then the text similarity method, 
+as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
+the \textit{diff command} built-in in the Linux environment \citep{diffcommand}.
+
+The difference method was not tested on the evaluation of the tool and other phases of this research. 
+The method was implemented to show that tool can have multiple duplication detection methods.
 
@@
 
-\subsection{Setup step}
+\subsection{Preprocessor}
 \label{subsec:setup}
 
-The Setup step is a procedure executed by the user once per codebase.
-The Setup step takes the input codebase and a minimum similarity metric value for function pairs considered duplicates.
+The Preprocessor is a procedure executed by the user once per codebase. 
+The Preprocessor takes the input codebase and a minimum similarity metric value for function pairs considered duplicates. 
 It then runs the Function Parser and the Code Duplication Finder to create a temporary codebase and the Code Duplication Database. 
-Setting a minimum similarity metric reduces the Code Duplication Database size, optimizing memory usage and the computational cost of the Query Responder step.
-For example, if this parameter does not exist and we work with a codebase of $10000$ functions, each with a 50-character relative path and function name, the Code Duplication Database would reach approximately $5000^2 \times 50 ~= 5$ gigabytes.
+Setting a minimum similarity metric reduces the Code Duplication Database size, optimizing memory usage and the computational 
+cost of the Query Responder step. For example, if this parameter does not exist and we work with a codebase of $10000$ functions, 
+each with a 50-character relative path and function name, the Code Duplication Database would reach approximately $5000^2 \times 50 ~= 5$ gigabytes. 
 As most function pairs are unlikely to be duplicated in large codebases, this file size can be significantly reduced.
 
-\subsection{Query Responder step}
+\subsection{Query Responder}
+
+\todo[inline]{I NEED TO CHANGE HERE}
 
@@
 \begin{itemize}
-	\item Number of duplicated functions above a similarity threshold.
+	\item Number of dupliddcated functions above a similarity threshold.
 	\item Number of duplicated functions equal or above a similarity threshold.

### 2025-04-27 conteudo/tool/tool.tex
update some things on the tool

@@
 
-The tool operates in two main steps. The first step, called \textbf{Setup}, is
-executed only once for each codebase. This step processes the codebase by
-extracting functions from code files and identifying code clone duplications.
-The second step, called \textbf{Query Responder}, utilizes the information obtained in
-the Setup phase to execute the tool functionalities as needed. Detailed
-descriptions of the Setup and Query Responder processes can be found in
+The tool operates in two main parts, the \textbf{Preprocessor} and the \textbf{Query Responder}. 
+The \textbf{Preprocessor} is responsible to perform the heavy computations to find the code duplications 
+accross a codebase and produce artifact to consulte information related to the duplications in a structured form.
+The \textbf{Query Responder} consumes the artifacts produced by the Preprocessor to execute the tool functionalites as requested.
+This choice enable the tool to concentrate the heavy and slows parts that can be execute only once per codebase, 
+thus allowing a fast performance for multiple queries related to a single codebase.
+More detailed descriptions of the Setup and Query Responder processes can be found in
 Sections \ref{subsec:architecture} and \ref{subsec:func}.

### 2025-04-27 conteudo/tool/architecture.tex
remove setup and change arc diagram

@@
 
-The Query Responder consumes the Temporary Codebase and the Code Duplication Database to extract duplicated functions-related information per user request.
+The Query Responder consumes the Temporary Codebase and the Code Duplication Database to extract duplicated 
+functions-related information per user request. If the user executes the Query Responder without executing the 
+Preprocessor, the Query Responder calls the Preprocessor to create the required artifacts.
 
@@
 \begin{figure}
-\includegraphics[scale=0.5]{diagrama_mestrado}
+\includegraphics[scale=0.4]{diagrama_mestrado}
 \caption{Architecture diagram demonstrating the relationship between the tool components.}
@@
 If the similarity is greater than or equal to the minimum similarity threshold 
-(a parameter provided by the user during the Setup step), we store this pair of functions along with its 
+(a parameter provided by the user during the Preprocessor), we store this pair of functions along with its 
 similarity in the Code Duplication Database.
@@
 
-The Query Responder step is the component the user executes multiple times to consult information about duplicated functions in a codebase processed by the Setup step.
+The Query Responder step is the component the user executes multiple times to consult information about duplicated functions in a codebase processed by the Preprocessor.
 The Query Responder step utilizes the temporary codebase and the Code Duplication Database as inputs to answer the user's request, such as querying the total number of duplicated functions in a specific input codebase or listing duplicated functions above a certain similarity threshold.

### 2025-04-27 conteudo/introduction/intro.tex
add comments on todos for the dissertation text
some changes to better reflect the change of the focus of the research

@@
 
+\todo[inline]{Require updated information about the linux kernel. Required from qualification exam.}
+
 In searching for solutions to the AMD Display driver problem related to
@@
 
+\todo[inline]{Maybe talk about the tests on IIO software, as the test is not tested only on AMD Display driver anymore. Maybe mention the tool explicitly.}
+
 \section**{Research Design}
@@
 \label{sec:intresearch}
+
+\todo[inline]{THIS SECTION NEEDS TO BE REMADE FROM ZERO.}
 
@@
 
+\todo[inline]{Focus more on the tool here.}
+
 \begin{figure}[h]
@@
 
-This research aims to identify patterns of software engineering techniques, such as refactoring methods commonly used to address code duplications.
-These patterns will serve as a guide for future contributors in mitigating code duplication issues within this context.
-To investigate our approach, we are developing a tool to identify duplicate code in the Linux kernel, explicitly targeting parts of the AMD Display driver code.
-We are then refactoring portions of this code and submitting patches to maintainers, providing evidence of the tool's accuracy in detecting duplicate functions.
-Our research design follows a multi-method approach, structured into two phases, each employing different research methods.
+This research aims to identify patterns of software engineering techniques, 
+such as refactoring methods commonly used to address code duplications. 
+These patterns will serve as a guide for future contributors in mitigating 
+code duplication issues within this context. To investigate our approach, 
+we are developing a tool to identify duplicate code in the Linux kernel, 
+explicitly targeting parts of the AMD Display driver code and the IIO driver. We are then refactoring 
+portions of this code and submitting patches to maintainers, providing evidence 
+of the tool's accuracy in detecting duplicate functions. Our research design follows 
+a multi-method approach, structured into two phases, each employing different research methods.
 
@@
 identifying code duplications within the Linux kernel context. We will then
-proceed to the second phase, which focuses on strategies for mitigating these
-duplications. In \textbf{Phase II}, our objective is to identify practical
-approaches to address the code duplications detected in the Linux kernel. To
-achieve this, we have designed an ethnographic approach that enables us to
+proceed to the second phase, which focuses on mitigating these
+duplications. In \textbf{Phase II}, our objective is to verify if we can 
+practically mitigate the duplication finds by the developed tool. 
+To achieve this, we have designed an ethnographic approach that enables us to
 interact with the Linux development community and gather artifacts that either

### 2025-04-27 conteudo/methods/methods.tex
some changes to better reflect the change of the focus of the research

@@
-We employ a multi-method research approach, using different methodologies to achieve our goals.
-Our first choice is to conduct an ethnographic study using a participant-observation methodology.
-In this study, we aim to mitigate a subset of duplicated functions by refactoring them systematically while interacting with the Linux community to gather feedback on our mitigation efforts.
-To address our main objective of mitigating code duplication in the Linux kernel, we first needed the ability to identify these duplications accurately.
-Thus, we developed a tool and assessed its capability to detect code duplications in the Linux kernel.
+We employ a multi-method research approach, using different methodologies to achieve our goals
+to approach code duplication in some of the Linux subsystems. 
+To validate if the Arkanjo tool is capable of finding code duplications, 
+applied two independent methods and perform triangulation on the results.
 
-We applied two independent methods and perform triangulation on the results. 
 The first method evaluates the proposed tool using an approach from the literature, 
@@
 We chose triangulation because relying solely on the literature-based method has 
-limitations and may not be entirely reliable \citep{bigfail, litreview}.
+limitations and may not be entirely reliable \citep{bigfail, litreview}. As a reminder, 
+all research approches were done using the text similarity method as the 
+code dupication method of the tool.
+
+Given the validation of the tool's capabilities, we proceed to undertand the viability 
+to mitigate the code duplication finds by the tool. 
+To underdand the viability, our first choice is to conduct an ethnographic study using 
+a participant-observation methodology. 
+In this study, we aim to mitigate a subset of duplicated functions by refactoring them 
+while interacting with the Linux community to gather feedback on our mitigation efforts.
+
+The ethnographic study happened in two phases. 
+In the first phase, the author of this research explored the process of mitigating
+a code duplication found in the AMD Display Driver and sent it as a PATCH to iteract with the
+Linux community and get their feedback. On a second phase, we proposed to students on the 
+course Free Software Development at University São Paulo to mitigate the code duplications found 
+by the tool in the IIO and AMD Display Driver and pass the process of send the mitigations as PATCH.
+(DOUBLE CHECK IF THE STUDENTS HAVE SENT THE PATCHES).
+
+I
 

### 2025-04-30 conteudo/methods/methods.tex
add work 29/04

@@
 
-\todo[inline]{THE WHOLE SECTION NEEDS TO BE REMADE.}
-
 Ethnography is a recognized qualitative empirical method for understanding people, 
@@
 
-This section outlines our systematic approach for mitigating code duplication in the Linux kernel, details the methodology for selecting duplicated functions, and describes the ethnographic methodology for engaging with the Linux community.
+This section outlines our approach for mitigatin code duplication in the Linux kernel
+and describes the ethnographic methodology for engagin with the Linux community. Firstly, 
+we present the participant observation research method explored by the author. Then, 
+we present the methodolody of the experiment done with the students.
 
-\subsection{Systematic Approach to Mitigate Code Duplication}
-\label{subsec:pipeline}
+\subsection{Participant observation method}
 
-In C programming, communication between source files is achieved by creating header files that specify libraries \citep{Cbook}.
-Since the Linux kernel is primarily written in C, we eliminated code duplications by consolidating duplicated code into a single library, which would replace instances of duplication across the codebase.
-%
-For each duplicated function pair, we used the \textit{function information} functionality (described in Section \ref{subsec:functioncommand}) to locate all instances of the function in the code.
-This strategy resulted in a collection of duplicated functions and corresponding code files.
+\subsection{Research with students}
 
-To identify shared code more effectively, we extended this approach to search for other common functions across all collection files.
-We then applied specific refactoring methods to each shared function.
-If the functions were identical across files, we removed the duplicates and created a single function in the library.
-If modifications existed, we applied case-specific refactoring.
+\todo[inline]{Rsearch with students does not look good.}
 
-Successful mitigations were submitted to the Linux community to gauge their feedback.
-If unsuccessful, we categorized the cases: some required specialist decisions beyond our expertise, and others were potentially non-refactorable.
-All mitigations were documented for research transparency at \url{https://github.com/LipArcanjo/master-artifacts}.
+In the Free Software Development course at University of São Paulo, ministred by the advisor 
+of this work at the first semester of 2025, we have proposed to the students the possibility
+of mitigating duplications in the IIO subsystem and the AMD display driver found with the Arkanjo tool.
+The course had a total of X students, and of that, X students being undergraduate students
+and Y begin graduate students.
 
-\subsection{Selection of Duplicated Functions}
+On the course, the students are asked to form groups of one, two or three students and together 
+make a contribution to a open source software. To auxiliate the students on this task, the 
+professor and teaching assistants prepares multiple alternatives on how to contribute to open source projects, 
+including simple contributions as updating documentation or code style fixes. In this context, 
+we have proposed two alternatives for the students related to the tool.
 
-As with the empirical analysis, evaluating every code duplication in the Linux kernel is impractical. 
-Thus, we will limit the experiment to the AMD Display driver and used random sampling of duplicated functions
-inferred by our tool to initiate the mitigation approach in Section \ref{subsec:pipeline}. 
-We will choose a sample size of 30 duplicated functions, as studies indicate this size provides 
-statistical significance \citep{sample1, sample2}.
+For the first alternative, one of the teaching assistants used the Arkanjo tool in the IIO subsystem and
+listed interesting duplications that the groups can approach, and it is up to the students on how
+to refactor the code to mitigate the duplications and pass through the process to send patches to the IIO 
+maintainers to review. For the second alternative, we propose a similar experience then the 
+participant observation method, where the students pass through the whole process of executing the Arkanjo tool
+in the AMD display driver, analyze the duplications found, create a patch to mitigate a duplication and sent
+the patch to the AMD display driver's maintainers.
 
-For random selection, we will restrict the eligibility to function pairs with a similarity metric of 100\%. 
-With this criterion, we randomly chose functions from eligible pairs uniformly.
+From the students enrolled in the in the course, X opted for the first alternative and 1 opted for the second
+alternative, forming Y groups working on the first alternative and 1 group working on the second alternative. 
+These number represent X\% of the students choosing for approching duplications them the other contribution 
+alternatives, including more simpler and easy to be done.  It is worth highlighting that none of the students
+that choose to mitigate duplications had made a contribution to the Linux kernel before,
+meaning that the patch to mitigate the duplication being the first tentative to contribute to the kernel as
+newcomers.
 
-\subsection{Ethnographic Methodology}
+The groups have been request to document the experience and approaches to refactor the duplications and the
+experience of submiting a patch to the Linux kernel in blogs. We had analyzed these blogs to undertand 
+patterns of refactor used for mitigating the duplications, the experience on send the patches and the 
+opinions of the subsystem maintainers on the mitigations. Appendix X list the blogs and patches made by the 
+student.
 
-We conduct ethnographic research upon completing a code duplication mitigation following the systematic approach in Section \ref{subsec:pipeline}.
-We adopt a participant-observation methodology, contributing to the Linux kernel for the AMD Display driver and interacting with maintainers to incorporate our mitigations.
-
-Linux kernel communication occurs via public mailing lists, where discussions are accessible to all.
-We collect data artifacts from the AMD Display driver mailing list, which records community feedback on our mitigation efforts, providing valuable insights into the proposed approach effectiveness.
-
-Given our research timeline, we do not expect continuous interactions with the Linux community throughout all mitigations, as response time varies.
-Thus, we committed to applying the participant-observation approach for three to five successful mitigations and using these for our final discussion.
+For the teaching assistant and the student that has executed the Arkanjo tool to explore duplications, 
+we have asked them to answer some questions related to the experience of using the tool and what could be
+improved. The question asked can be found below and the answers can be found in Appendix Y.

### 2025-05-03 conteudo/methods/methods.tex
add some work from 03/05

@@
 course Free Software Development at University São Paulo to mitigate the code duplications found 
-by the tool in the IIO and AMD Display Driver and pass the process of send the mitigations as PATCH.
+by the tool in the IIO subsystems and the AMD Display Driver and pass the process of send the mitigations as PATCH.
 (DOUBLE CHECK IF THE STUDENTS HAVE SENT THE PATCHES).
-
-I
 
@@
 
-%
 Given the vast number of lines and components in the Linux kernel, analyzing every duplication is impractical.
@@
 would accept our approach or identify their concerns. Thus, we will use ethnography, 
-positioning ourselves as Linux kernel contributors focused on removing identified 
-code duplications.
+positioning ourselves and students as Linux kernel contributors focused on removing 
+identified  code duplications.
 
-This section outlines our approach for mitigatin code duplication in the Linux kernel
+This section outlines our approach for mitigating code duplication in the Linux kernel
 and describes the ethnographic methodology for engagin with the Linux community. Firstly, 
@@
 
+Given the tool evaluations that demonstrate the tool capabilites of finding duplications, 
+we proceeded to undertand if the duplications found are possible to be mitigate and
+the kernel subsystem's maintainers opinions about the duplications and their respective mitigations.
+To accomplish this objective, we initialy proposed a participant-observation experiment, 
+where we, people that have never contribute to the Linux kernel, 
+act as contributors to the AMD Display driver and collect artifacts of our experience in the process.
+
+As a first step of this experiment, we executed the tool in the AMD Display driver codebase and 
+manually analysed the biggest duplications in terms of number of lines and raised to 
+two duplicated function pairs that we judged interesting to try to mitigate. 
+We have choose to approach the biggest duplications as a proof of concept that 
+we can make significant impacts. The two duplicated function pairs are shown in Table X.
+
+Given the two duplicated function pairs, we observed that the files that the functions exist contains 
+multiple other duplications. Thus, we proposed a simple systematic approach to mitigate all the 
+functions duplicated in the context and not only the duplicated function pairs. 
+The systematic approach is defined below.
+
+After we approached the duplications with the systematic approach. For the duplicated function
+pairs that had successfully being mitigate the duplication, we sent patches to the AMD Display driver
+maillist to validate the mitigations and see the maintainers opinion, while documentaion the process 
+and our impressions.
+
+\subsubsection{Systematic Approach to Mitigate Code Duplication}
+
+In C programming, communication between source files is achieved by creating header files that specify 
+libraries \citep{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
+by consolidating duplicated code into a single library, which would replace instances of 
+duplication across the codebase.
+
+For each duplicated function pair, for each function in the pair, we used the \textit{function information} functionality 
+(described in Section \ref{subsec:functioncommand}) to locate all duplications of the function in the codebase, 
+as it may have been duplicated more times then the two occurences of the duplicated function pair. 
+This strategy resulted in a collection of duplicated functions and corresponding code files.
+
+To identify shared code more effectively, we extended this approach to search for other common 
+functions across all collection files. We then applied specific refactoring methods to each shared 
+function. If the functions were identical across files, we removed the duplicates and created a single 
+function in the library. If modifications existed, we applied case-specific refactoring.
+
 \subsection{Research with students}
 
-\todo[inline]{Rsearch with students does not look good.}
+\todo[inline]{Research with students does not look good.}
+
+Posterior of the initial participant observation that we made, we decided to expand the experiment 
+by proposing to students mitigate duplication found by the tool and send patches to the kernel community.
 
@@
 For the first alternative, one of the teaching assistants used the Arkanjo tool in the IIO subsystem and
-listed interesting duplications that the groups can approach, and it is up to the students on how
+listed interesting duplicated functions pair that the groups can simple mitigate this specific duplication pair, 
+and it is up to the students on how
 to refactor the code to mitigate the duplications and pass through the process to send patches to the IIO 
@@
 
-For the teaching assistant and the student that has executed the Arkanjo tool to explore duplications, 
+For the teaching assistant and the student that executed the Arkanjo tool to explore duplications, 
 we have asked them to answer some questions related to the experience of using the tool and what could be

### 2025-05-03 conteudo/results/eval.tex
reorginaze results files

--- 2024-11-15 conteudo/eval/eval.tex
@@
-\en
-
-To finalize this report, we present the current state of our research. As of the 
-time of writing, we have finished \textbf{Phase I} of this research, 
-which includes the implementation of 
-the proposed tool and the evaluation of the tool through both formal literature 
-and empirical analyses. This chapter begins with the presentation of the results 
-from the tool evaluation. Next, we present a preliminary experiment we are currently 
-conducting to assess the viability of our Ethnographic Study. Finally, we outline 
-the next steps in this research.
-
 \section{Tool Evaluation}
@@
 
-\input{conteudo/eval/table_bigclone}
+\input{conteudo/results/table_bigclone}
 
@@
 
-\input{conteudo/eval/table_empirical}
+\input{conteudo/results/table_empirical}
 

### 2025-05-03 conteudo/results/results.tex
reorginaze results files

--- 2025-04-24 conteudo/eval/eval.tex
@@
 
-\section{Tool Evaluation}
-\label{sec:eval}
+\input{conteudo/results/eval}
 
-In this section, we present the test results of our tool performance, following the 
-methodology described in Chapter \ref{cha:method}. First, we provide the results 
-obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}, which 
-serves as a standard reference in the literature \citep{litreview}. Second, we 
-discuss the results from our empirical analysis. Finally, we triangulate the data, 
-analyzing our findings regarding the tool performance. 
-As a reminder, we only expect our tool to yield good results for Type-1 and Type-2,
-as mentioned in \ref{subsec:code_finder}.
-
-\input{conteudo/eval/table_bigclone}
-
-As described in Section \ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset \citep{bigclonebench}, we define that every pair of functions with a similarity metric equal to or greater than a threshold \( X \) is considered a duplication.
-The pair of code files containing these functions are then labeled as duplications.
-If our tool correctly identifies a code duplication pair, it is counted as correct for the respective code duplication type.
-If it infers a duplication pair incorrectly, it is counted as incorrect for all code duplication types.
-Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}.
-
-The results in Table \ref{tab:bigclone} indicate that our tool is conservative in its predictions.
-No negative sample was inferred as a duplication, even with a minimum similarity threshold of 30\%.
-However, it performs poorly identifying complex duplication types, such as Medium Type-3 and Weak Type-3/Type-4.
-As discussed in Chapter \ref{cha:tool}, if improved performance on these complex types is needed, our tool allows replacing the current duplication detection method due to its modular design.
-
-\subsection{Tool Results on Empirical Analysis}
-
-As outlined in Section \ref{sec:metemp}, for the empirical evaluation of our tool, we randomly selected ten function pairs across various similarity metric ranges computed by our tool.
-We then examined each pair to determine whether they were duplications and, if so, categorized the duplication type.
-Table \ref{tab:emp} presents the results from our empirical analysis.
-
-\input{conteudo/eval/table_empirical}
-
-The results in Table \ref{tab:emp} show that as the similarity metric decreases, 
-there is a notable increase in function pairs that are not duplications within 
-the Linux kernel. For similarity ranges of 79\%-81\% or higher, there is a high 
-percentage of code duplications. At 69\%-71\% similarity, the duplication rate 
-drops to 50\%. At 59\%-61\%, it further decreases to 20\%, and below 49\%-51\%, 
-no code duplications were found.
-
-\subsection{Triangulation of Results}
-
-The BigCloneBench results and our empirical methods indicate that our tool performs well in detecting Type-1 and Type-2 code clone duplications but struggles with more complex types.
-A similarity threshold between 80\% and 100\% yields favorable results in both methods.
-
-However, there is a discrepancy in the detection of non-duplication pairs between the two methods.
-In the BigCloneBench evaluation, we found no false positives (negative samples inferred as duplications), while the empirical analysis revealed a higher percentage of false positives.
-We propose two potential explanations for this discrepancy.
-
-The first reason may be the limitations and known issues with BigCloneBench, as highlighted by Krinke and Ragkhitwetsagul \citep{bigfail}.
-The second reason could relate to the nature of the AMD Display driver, where each code artifact shares semantic meaning.
-In contrast, BigCloneBench comprises self-contained code artifacts without shared semantics.
-Given that our tool relies on a text-based code clone detection approach, it may naturally perform worse on the AMD Display driver than on the BigCloneBench dataset.
-We cannot determine if state-of-the-art code duplication detection tools face similar issues obscured by evaluations limited to the BigCloneBench dataset.
+\input{conteudo/results/participant}

### 2025-05-04 conteudo/background/linux.tex
add diff information

@@
 
-\todo[inline]{Not exactly on this section, but I need to mention GIT, as one of the similarity methods is based on the GIT command.}
+\subsection{Diff command}
+
+The \textit{diff} command is a command line in Linux that allows the user to compare 
+two files line by line \citep{diffcommand}. There are multiple forms 
+that the results of the command can be shown, one example of the command being executed 
+can be found in Figure \ref{fig:diff}. With the command results, it is easy to visualize and extract 
+the number of lines that are equal and different between the two compared files.
+
+\begin{figure}
+\includegraphics[scale=0.3]{diff_command}
+\caption{Example of \textit{diff} command executed}
+File 1 content is shown in red and file 2 content is shown in green. \textit{-} 
+represents a line that only exist in file 1, \textit{+} represents a line that only exist in file 2,
+\textit{!} represents a line that exist in both file but has modifications between them, and empty on the
+start of a line represents that the line exist in both files without any modifications.
+\label{fig:diff}
+\end{figure}
+
+We use thhe \textit{diff} command as the base of one of the code duplication detection methods 
+on the tool proposed on this work, although it is not extensive explored as the text similarity
+method for code duplication detection.

### 2025-05-11 conteudo/background/linux.tex
add sowrk 11/05

@@
 
-One of the subsystems within the Linux kernel is the AMD Display Driver, which
+One of the subsystems within the Linux kernel is the AMD Display driver, which
 is responsible for implementing the drivers needed for AMD GPUs to function
@@
 object of study as we aim to understand and mitigate code duplication in the
-Linux kernel. The AMD Display Driver subsystem contains more then 391000 lines 
+Linux kernel. The AMD Display driver subsystem contains more then 391000 lines 
 of code and more then 1089 code files
@@
 making it a considerable component within the Linux kernel. Like other 
-subsystems, discussions about the AMD Display Driver are conducted on its 
+subsystems, discussions about the AMD Display driver are conducted on its 
 designated mailing list, which can be accessed here: 
@@
 
-\todo[inline]{Require an introduction about IIO drivers.}
+Another subsytem within the Linux kernel explored in this work is the Industrial
+I/O subsytem (IIO). From the Linux kernel documentation, ''The main purpose of the 
+Industrial I/O subsystem (IIO) is to provide support for devices that in some sense 
+perform either analog-to-digital conversion (ADC) or digital-to-analog conversion (DAC) or both``\citep{iiodoc}.
+The IIO subsystem contains more then X lines of code and more then Y code files 
+\footnote{
+Data extracted directly from the Linux repository with the help of the clol tool. \citep{cloc}
+}.
+We opted to include the IIO subsystem in our experiments to ensure our findings are not limited to a single subsystem. 
+Additionally, we have an established point of contact within the IIO subsystem to facilitate this work.
+. Like other subsystems, discussions about the IIO subsytem are conducted on its 
+designated mailing list, which can be found here: 
+\url{https://subspace.kernel.org/vger.kernel.org.html}.
 

### 2025-05-11 conteudo/introduction/intro.tex
add sowrk 11/05

@@
 
-\todo[inline]{Maybe talk about the tests on IIO software, as the test is not tested only on AMD Display driver anymore. Maybe mention the tool explicitly.}
-
 \section**{Research Design}
@@
 \label{sec:intresearch}
-
-\todo[inline]{THIS SECTION NEEDS TO BE REMADE FROM ZERO.}
 
@@
 kernel. Our approach includes the development of a tool capable of detecting
-duplicated functions within the kernel and establishing strategies for
-addressing these duplications.
-
-\todo[inline]{Focus more on the tool here.}
+duplicated functions and establishing strategies for
+addressing these duplications. We tested the approach in the context of two
+kernel components, the AMD Display Driver the the Industrial IO subsystem.
 
@@
 code duplication issues within this context. To investigate our approach, 
-we are developing a tool to identify duplicate code in the Linux kernel, 
-explicitly targeting parts of the AMD Display driver code and the IIO driver. We are then refactoring 
+we are developing a tool to identify duplicate code. We are then refactoring 
 portions of this code and submitting patches to maintainers, providing evidence 
@@
 
-\section**{Manuscript Structure}
+\section**{Thesis Structure}
 
@@
 the literature overview for code duplication detection (main definitions,
-current approaches in the literature), a brief description of the Linux kernel,
+current approaches in the literature), a brief description of the Linux kernel and the components
+explored in this work, 
 and a review of refactoring methods used throughout this research. Chapter
-\ref{cha:tool} presents Arkanjo, our proposed tool to detect code duplication in the
-Linux kernel, describing all the main components.  Chapter \ref{cha:method}
+\ref{cha:tool} presents ArKanjo, our proposed tool to detect code duplication, 
+describing all the main components.  Chapter \ref{cha:method}
 describes the research methods selected to guide our work. Chapter
-\ref{cha:results} shows the first results obtained from the research methods to evaluate
-our tool and presents the current research stage and the work plan,
-with the addition of a discussion about our preliminary results.
+\ref{cha:results} shows the results we had through our work, from the research methods to 
+evaluate our tool and the ethnographic studies. Chapter \ref{cha:conclusion}udes this research.

### 2025-05-11 conteudo/results/participant.tex
reorginaze results files
add sowrk 11/05

--- 2024-11-18 conteudo/preliminar/results.tex
@@
 
-\section{Final Remarks}
+\section{Participant Observation Experiment}
 
-We have completed Phase 1 of our research, during which we developed our proposed tool and evaluated it using the BigCloneBench dataset and through empirical analyses.
+We have completed Phase 1 of ourdd research, during which we developed our proposed tool and evaluated it using the BigCloneBench dataset and through empirical analyses.
 The results indicate our tool can identify code clone duplications within the Linux kernel.
@@
 
-\input{conteudo/preliminar/table_patch}
+\input{conteudo/results/table_patch}
 
@@
 
-\input{conteudo/preliminar/cronograma}
+\input{conteudo/results/cronograma}
 

### 2025-05-11 conteudo/results/results.tex
add sowrk 11/05

@@
 
-\todo[inline]{As it is primarly results, everything should be remade, the structure of the files for this section is bad.}
-
-\todo[inline]{Remember the typos on the tables. }
-
-\todo[inline]{Idea: The research methods should be 3 phases, one of the tool creation/testing, one for the my experiences submiting a path, and the last for the experiences of the students on the course.
-Maybe split in four to split the tool creation for the tool testing.}
-
-To finalize this report, we present the current state of our research. As of the 
-time of writing, we have finished \textbf{Phase I} of this research, 
-which includes the implementation of 
-the proposed tool and the evaluation of the tool through both a method from the 
-literature and an empirical analyses. This chapter begins with the presentation of 
-the results from the tool evaluation. Next, we present a preliminary experiment 
-we are currently conducting to assess the viability of our Ethnographic Study. 
-Finally, we outline the next steps in this research.
+In this chapter we present and analyze the results found in the experiments and 
+methods done through our research. Following the cronology of the research, we 
+start by presenting the ArKanjo tool's evaluation through both a method from the literature
+and an empirical analyses. Then, we present and discuss the experience of the initial participant
+observation methodolody, where we mitigate a code duplication on the AMD Display driver and sent it as a patch.
+Finally, we show and discuss our discoveries on the experiment done with students at the course in University of 
+São Paulo, where we propose to students to mitigate duplications found by ArKanjo tool at IIO subsystem and the
+AMD Display driver and document their founds, opinions and experiences.
 
@@
 \input{conteudo/results/participant}
+
+\input{conteudo/results/students}

### 2025-05-11 conteudo/tool/tool.tex
remove setup and change arc diagram
add sowrk 11/05

@@
 thus allowing a fast performance for multiple queries related to a single codebase.
-More detailed descriptions of the Setup and Query Responder processes can be found in
+More detailed descriptions of the Preprocessor and Query Responder can be found in
 Sections \ref{subsec:architecture} and \ref{subsec:func}.
@@
 
-\todo[inline]{I need to rename the Setup as Preprocessor step.}
-
-\todo[inline]{Update the Function parser description of the C parser and Java parser}
-
-\todo[inline]{Add the other similarity method added on the tool using git for the Code Duplication Finder.}
-
-\todo[inline]{Update diagram and functionalites pictures (do not need sudo anymore!).}
-
 \todo[inline]{Big maybes: Describe more how the functionalites are implemented, as the focus shifted for the tool as of now.}

### 2025-05-18 conteudo/methods/methods.tex
add sowrk 11/05
Add work on 17/05

@@
 (DOUBLE CHECK IF THE STUDENTS HAVE SENT THE PATCHES).
-
-\todo[inline]{The focus of the above section should be a little more on the tool, the introduction about mitigating duplications is good enough.
-A little about the fact that we will test on IIO later maybe is good.
-}
 
@@
 \subsection{Participant observation method}
+\label{subsec:partmethod}
 
@@
 We have choose to approach the biggest duplications as a proof of concept that 
-we can make significant impacts. The two duplicated function pairs are shown in Table X.
+we can make significant impacts. The two duplicated function pairs chosen are:
+
+\begin{itemize}
+\item \textbf{Duplicated function pair 1}
+\begin{itemize}
+\item Function named \textit{offset\_to\_id} in file \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c}.
+\item Function \textit{offset\_to\_id} in file \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
+\end{itemize}
+\item \textbf{Duplicated function pair 2}
+%\begin{itemize}
+\item Function named \textit{phy\_id\_to\_atom} in file \textit{dc/bios/dce110/command\_table\_helper\_dce110.c}.
+\item Function named \textit{phy\_id\_to\_atom} in file \textit{dc/bios/dce60/command\_table\_helper\_dce60.c}.
+%\end{itemize}
+\end{itemize}
 
@@
 \subsubsection{Systematic Approach to Mitigate Code Duplication}
+\label{subsubsec:systematic}
 

### 2025-05-18 conteudo/results/participant.tex
Add work on 17/05

@@
 
-We have completed Phase 1 of ourdd research, during which we developed our proposed tool and evaluated it using the BigCloneBench dataset and through empirical analyses.
-The results indicate our tool can identify code clone duplications within the Linux kernel.
+For our initial participant observation, we attempted to mitigate two code duplications found by ArKanjo tool
+in the AMD Display driver applying our systematic approach presented in \ref{subsubsec:systematic}. 
+As presented in section \ref{sec:meteth}, we opted for applying the systematic approach in the functions 
+\textit{offset\_to\_id} and \textit{phy\_id\_to\_atom} in the folders \textit{dc/gpio} and \text{dc/bios} respectively.
 
-As the next step, we plan to conduct an ethnographic study as outlined in Section \ref{sec:meteth}.
-To preliminarily assess the potential impact of this research, we applied our systematic approach for mitigating code duplication to two pairs of functions.
-Table \ref{tab:patch} presents the initial results obtained from this process.
+Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
+\textit{phy\_id\_to\_atom} will be referred to as mitigation 2. Table \ref{tab:patch} summarizes the impact of the mitigations
+in terms of files and lines changed, and the refactoring method used. We present our experience and learning for each mitigation 
+in the next sections.
 
 \input{conteudo/results/table_patch}
+
+\subsection{Mitigation 1}
+
+On the process of the systematic approach, we found that the \textit{offset\_to\_id} function exist in 10 files, enclosing
+multiple GPU architectures. The duplicated functions are not exactly equals, as there are some minimal specific logics applied
+to each GPU architecture. To address this issue, we applied the extract method to break the function in smaller parts and the 
+parameterize method to address the specific logics for the GPU's.
+
+The base idea on how we would mitigate the duplications found were done, but we could not complete the refactoring on a state 
+capable to submit as a patch to the driver. 
+We found that the configuration files on the AMD Display driver are designed to be imported at compilation time
+using \textit{\#define} macros. The configuration files design choice makes the refactoring of duplications on generic 
+approachs trickier, as refactoring code that depends on these files requires significant refactor on the configuration files
+design and deep knowledge of the codebase, which us as first-timer contributors do not have. Thus, we opted to not continue 
+on investigating this mitigation.
+
+\subsection{Mitigation 2}
+
+\subsection{OLD}
 
@@
 patch can be found in Appendix \ref{app:feedback}.
-
-\input{conteudo/results/cronograma}
-
-Finally, as indicated in our proposed schedule in Table \ref{fig:cronograma}, 
-our upcoming tasks include conducting the complete ethnographic study.
-This step will start with completing our systematic approach to code 
-duplication mitigation. Once we have finalized the mitigations, we intend 
-to submit them to the AMD Display driver mailing list to gather insights 
-and feedback from the maintainers regarding the changes we implemented.
-
-Our upcoming tasks also include implementing improvements to our proposed tool. 
-While the tool’s current state allowed us to conduct the experiments described in 
-Chapter \ref{cha:method},
-it is not yet adequately user-friendly or manageable for external users. 
-To achieve our complementary goal of providing a solution for code duplication 
-in the kernel, we plan to improve the documentation, refine the code structure, 
-and address any issues we encounter along the way, providing the Linux community 
-with a way to tackle the issue.

### 2025-05-18 conteudo/results/participant.tex
add some work of 18/05

@@
 
-On the process of the systematic approach, we found that the \textit{offset\_to\_id} function exist in 10 files, enclosing
+On the process of the systematic approach, we found that the \textit{offset\_to\_id} function exist in code 10 files, enclosing
 multiple GPU architectures. The duplicated functions are not exactly equals, as there are some minimal specific logics applied
@@
 
-The base idea on how we would mitigate the duplications found were done, but we could not complete the refactoring on a state 
+We reached a point with a scratch on how we would mitigate the duplications found were done, but we could not complete the refactoring on a state 
 capable to submit as a patch to the driver. 
@@
 
+On the process of the systematic approach, we found that the \textit{phy\_id\_to\_atom} function exist in 5 code files, and 
+there are two other functions that are duplicated on those 5 code files, which . All functions across the files were
+exactly equal, thus there was no need to apply refactor methods presented in literature. The refactoring work was resumed
+to creating a generic library and fixing the compilation targets.
+
+The code files touched in the second mitigation does 
+not depend on configuration files, thus we have not faced the same issues on the first mitigation. 
+The refactoring of the code to mitigate duplications reached a point we judged that was good enough to submit as a patch
+to contribute to the AMD Display driver code quality. Thus we moved to send the refactoring to the driver's maintainers 
+as patch, documenting or findings and experiences on the process.
+
+\todo[inline]{A image of the timeline of the patch can be good to not have only test here}
+
+The experience on submiting a patch was not as we expected.
+The overall process took 7 months and 16 days, which is a time significant slower then we expected.
+We initially submitted the patch on Auguest 9th, 2024, but we did not receive replies at first. Thus, we had
+to to resend it on October 9th 2024. We received an initial response on November 3rd, 2024 asked for minimal changes, 
+we submitted a second version with the changes request on November 16th, 2024. We did not receive replies
+at first and need to resend the patch on February 7th, 2025. 
+We received a reply on February 11th 2025 asking for additional changes. 
+We sent an third version on February 24th 2025 and the patch got accepted and integrated in the kernel codebase in 
+February 25th 2025.
+
+We discussed with our point of contact at the AMD Display driver to understand the reason of the slow process
+to submit a patch. 
+We learned that some of the code in the driver is shared across all operating systems that 
+support the implemented GPUs. 
+This fact creates a complex process within AMD to format and submit changes across the supported systems while 
+simultaneously implementing measures to mitigate errors.
+Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, 
+as we understand it is a good practice in software engineering. One example shared with us is that duplicated 
+code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without 
+needing to test compatibility with others. This approach helps save significant time and effort.
+
+Regarding the patch changes requested for the second version, 
+only minor changes were requested to align with code style, license correct use and best practices that
+were initially unkown to use. For the change requested on the third version, we were simple asked to 
+move the functions created in the generic library to a existent file, instead of creating a new one.
+More details of the patch can be found in Appendix \ref{app:feedback} and the patch submitted in 
+the linux codebase can be see here: \url{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}.
+
+The initial plan of this research was to approach and sent a large number of mitigations to
+collect a broad number of artifacts from the driver community, but the slow process to send a patch
+makes this plan unviable. Thus, we opted to redirect this research to work with students and refactoring duplicated function
+pairs found by the tool directly in the IIO and the AMD Display driver, without passing through the whole process 
+of the systematic approach proposed.
+This alternative allowed us to collect artifacts from the students experiences and findings, and not be 
+to dependable of the iteraction with the kernel community.
+A additional advantage is to accelerate the iteraction with the kernel community
+because we are not applying the whole systematic, meaning that the code changes became smaller.
+
 \subsection{OLD}
@@
 This observation is because duplicated code artifacts involve copy-pasted code that imports different configuration files.
-The design of the AMD Display driver configuration files necessitates importing specific configurations at compile time, which prevents a generic approach.
-Consequently, refactoring this pair requires changes in how the AMD Display driver manages configuration files -- a complex task that demands input from an expert or maintainer.
+The design of the AMD Display driver configuration files necessitates importing specific configurations at compile time, 
+which prevents a generic approach. Consequently, refactoring this pair requires changes in how the AMD Display driver 
+manages configuration files -- a complex task that demands input from an expert or maintainer.
 
-The refactoring of the second function pair, however, was successful.
-We submitted this refactoring as a patch to the AMD Display driver to obtain feedback from the maintainers.
-The process of sending the patch took more time than expected.
-We initially submitted the patch on August 9th, 2024, but we did not receive replies.
-As a result, we had to resend it on October 9th, 2024 and received an initial response on November 3rd, 2024.
+The refactoring of the second function pair, however, was successful. We submitted this refactoring as a patch to the AMD 
+Display driver to obtain feedback from the maintainers. The process of sending the patch took more time than expected.
+We initially submitted the patch on August 9th, 2024, but we did not receive replies. As a result, we had to resend 
+it on October 9th, 2024 and received an initial response on November 3rd, 2024.
 
 We understood new details about the driver evolution processes in discussions with our partner from the AMD Display driver.
-Some of the code in the driver is shared across all operating systems that support the implemented GPUs.
-This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors.
-
-Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering.
-One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others.
-This approach helps save significant time and effort.
 

### 2025-05-19 conteudo/results/table_students.tex
add some work 18/05

\begin{table}
\begin{tabular}{ | c | c | c | c | m{6em} | }

\hline

\textbf{Mitigation} & \textbf{Subsystem} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
\\ \hline

1 & IIO & 0 & 0 & None \\ \hline
2 & IIO & 0 & 0 & None \\ \hline
3 & IIO & 0 & 0 & None \\ \hline
4 & IIO & 0 & 0 & None \\ \hline
5 & IIO & 0 & 0 & None \\ \hline
6 & IIO & 0 & 0 & None \\ \hline
7 & IIO & 0 & 0 & None \\ \hline
8 & IIO & 0 & 0 & None \\ \hline
9 & IIO & 0 & 0 & None \\ \hline
10 & IIO & 0 & 0 & None \\ \hline
11 & IIO & 0 & 0 & None \\ \hline
12 & AMD & 0 & 0 & None \\ \hline

\hline
\end{tabular}
\caption{Refactor methods used and metrics of the students's mitigations.}
\label{tab:stu}
\end{table}

### 2025-05-19 conteudo/results/students.tex
add students

\section{Students}

% As presented in chapter \ref{cha:methods}, we propose to students to mitigate duplications
% found by the ArKanjo tool and send patches to the linux subsystems maintainers, while
% documenting their experiences and learning through the process. 
% 
% For the students that worked
% in the IIO subsytems, we gave them duplicated function pairs found by the tool and their 
% responsability was to think of how to mitigate the duplication and sent the patch to the 
% maintainers. The student that worked in the AMD Display driver have the additional work
% to use the tool to find the duplications.
% 
% Appendix (CREATE APPENDIX AND ADD HERE X) has references
% to the artifacts discussed in this section, including the duplication function pairs that students 
% mitigate, the students's reports, the patch students sent to the linux kernel, and the 
% answers of the questionary asked to the student that worked in the AMD Display driver and the 
% teaching assistant that used the tool to find duplications in the IIO.
% 
% \input{conteudo/results/table_students}
% 
% Table (CREATE TABLE AND ADD HERE X) summarizes the mitigation approachs of the students in terms 
% of lines changed and refactor methods used. BLABLABLA
% 
% 
% \subsection{Questionary Answers}
% 
% BLABLABLA

### 2025-05-20 conteudo/results/students.tex
add some work

@@
 
-% As presented in chapter \ref{cha:methods}, we propose to students to mitigate duplications
-% found by the ArKanjo tool and send patches to the linux subsystems maintainers, while
-% documenting their experiences and learning through the process. 
-% 
-% For the students that worked
-% in the IIO subsytems, we gave them duplicated function pairs found by the tool and their 
-% responsability was to think of how to mitigate the duplication and sent the patch to the 
-% maintainers. The student that worked in the AMD Display driver have the additional work
-% to use the tool to find the duplications.
-% 
-% Appendix (CREATE APPENDIX AND ADD HERE X) has references
-% to the artifacts discussed in this section, including the duplication function pairs that students 
-% mitigate, the students's reports, the patch students sent to the linux kernel, and the 
-% answers of the questionary asked to the student that worked in the AMD Display driver and the 
-% teaching assistant that used the tool to find duplications in the IIO.
-% 
-% \input{conteudo/results/table_students}
-% 
-% Table (CREATE TABLE AND ADD HERE X) summarizes the mitigation approachs of the students in terms 
-% of lines changed and refactor methods used. BLABLABLA
-% 
-% 
-% \subsection{Questionary Answers}
-% 
-% BLABLABLA
+As presented in chapter \ref{cha:method}, we propose to students to mitigate duplications
+found by the ArKanjo tool and send patches to the linux subsystems maintainers, while
+documenting their experiences and learning through the process.
+
+For the students that worked
+in the IIO subsytems, we gave them duplicated function pairs found by the tool and their 
+responsability was to think of how to mitigate the duplication and sent the patch to the 
+maintainers. The student that worked in the AMD Display driver have the additional work
+to use the tool to find the duplications.
+
+Appendix (CREATE APPENDIX AND ADD HERE X) has references
+to the artifacts discussed in this section, including the duplication function pairs that students 
+mitigate, the students's reports, the patch students sent to the linux kernel, and the 
+answers of the questionary asked to the student that worked in the AMD Display driver and the 
+teaching assistant that used the tool to find duplications in the IIO.
+
+\input{conteudo/results/table_students}
+
+Table (CREATE TABLE AND ADD HERE X) summarizes the mitigation approachs of the students in terms 
+of lines changed and refactor methods used. BLABLABLA
+
+\subsection{Questionary Answers}
+
+BLABLABLA

### 2025-05-25 conteudo/apendices/students-artifacts.tex
add some work from 24-05

\chapter{Students Artifacts}

\label{app:stuart}

This appendix includes the artifacts produced by the students at University of São Paulo.

\paragraph{Group 1}

\

\textbf{Blog 1:} \url{https://linux.ime.usp.br/~juliacalixto/software-livre.html}

\textbf{Blog 2:} \url{https://linux.ime.usp.br/~soueuabianca/post2.html}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250426194135.244774-1-biancalvao@gmail.com/}

\paragraph{Group 2}

\

\textbf{Blog 1:} \url{https://mikhael-abdallah.github.io/software-livre/posts/aula-06/}

\textbf{Blog 2:} \url{https://augustomb02.github.io/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/all/20250429093923.347370-1-mikhael_abdallah@usp.br/}

\paragraph{Group 3}

\

\textbf{Blog 1:} \url{https://beavct.github.io/}

\textbf{Blog 2:} \url{https://gabriela-victor-github-io.vercel.app/blog/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250424002144.23260-1-beatrizvianacosta16@gmail.com/}

\paragraph{Group 4}

\

\textbf{Blog 1:} \url{https://linux.ime.usp.br/~aschwarz}

\textbf{Blog 2:} \url{https://whip-foxtrot-b57.notion.site/MAC0470-Desenvolvimento-de-Software-Livre-2025-1dcde8837d}

\textbf{Blog 3:} \url{https://www.linux.ime.usp.br/~artpilone/sl_blog.html}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250421145534.91146-1-arthurpilone@usp.br/}

\paragraph{Group 5}

\

\textbf{Blog 1:} \url{https://gust-vaz.github.io/}

\textbf{Blog 2:} \url{https://xivor.github.io/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250426153307.59689da0@jic23-huawei/T/#t}

\paragraph{Group 6}

\

\textbf{Blog 1:} \url{https://gabriellimmaa.github.io/}

\textbf{Blog 2:} \url{https://vitormarquesr.github.io/blog/}

\textbf{PATCH in Linux lore:} Students have not provided the link.

\paragraph{Group 7}

\

\textbf{Blog 1:} \url{https://isacaselli.github.io/}

\textbf{Blog 2:} \url{https://rodrigocmichelassi.github.io/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250428132551.176788-1-bellacaselli20@gmail.com/}

\paragraph{Group 8}

\

\textbf{Blog 1:} \url{https://linux.ime.usp.br/~nattan/mac470/mac470.html}

\textbf{PATCH in Linux lore:} Students have not provided the link.

\paragraph{Group 9}

\

\textbf{Blog 1:} \url{https://andrewijano.github.io/mac5856/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250423170526.15143-1-andrew.lopes@alumni.usp.br/}

\paragraph{Group 10}

\

\textbf{Blog 1:} \url{https://linux.ime.usp.br/~apolo/mac0470/}

\textbf{Blog 2:} \url{https://ncbor.github.io/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250428034222.318898-1-gabrielgeraldinosouza@gmail.com/T/#u}

\paragraph{Group 11}

\

\textbf{Blog 1:} \url{https://o-carneiro.github.io/}

\textbf{Blog 2:} \url{https://fernandoglima.github.io/}

\textbf{Blog 3:} \url{https://eijiuchiyama.github.io/softwarelivre/}

\textbf{PATCH in Linux lore:} Students have not provided the link.

\paragraph{Group 12}

\

\textbf{Blog 1:} \url{https://saguileran.github.io/blog/2025/free-software-development/}

\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/amd-gfx/f78f68ba-80e0-47ca-97a2-37a15d076e68@amd.com/T/#t}

### 2025-05-25 conteudo/methods/methods.tex
add some work 18/05
add some work from 24-05

@@
 of mitigating duplications in the IIO subsystem and the AMD display driver found with the Arkanjo tool.
-The course had a total of X students, and of that, X students being undergraduate students
+The course had a total of 26 students, and of that, X students being undergraduate students
 and Y begin graduate students.
@@
 
-From the students enrolled in the in the course, X opted for the first alternative and 1 opted for the second
-alternative, forming Y groups working on the first alternative and 1 group working on the second alternative. 
-These number represent X\% of the students choosing for approching duplications them the other contribution 
-alternatives, including more simpler and easy to be done.  It is worth highlighting that none of the students
+From the students enrolled in the in the course, 25 students opted for the first alternative and 1 student
+opted for the second alternative, forming 11 groups working on the first alternative and 1 group 
+working on the second alternative. It is worth highlighting that none of the students
 that choose to mitigate duplications had made a contribution to the Linux kernel before,
@@
 patterns of refactor used for mitigating the duplications, the experience on send the patches and the 
-opinions of the subsystem maintainers on the mitigations. Appendix X list the blogs and patches made by the 
-student.
+opinions of the subsystem maintainers on the mitigations.
 
@@
 we have asked them to answer some questions related to the experience of using the tool and what could be
-improved. The question asked can be found below and the answers can be found in Appendix Y.
+improved. The questions asked is the following:

### 2025-05-25 conteudo/results/participant.tex
add some work from 24-05

@@
 A additional advantage is to accelerate the iteraction with the kernel community
-because we are not applying the whole systematic, meaning that the code changes became smaller.
-
-\subsection{OLD}
-
-In the first function pair, completing the refactoring requires the expertise of a maintainer for the AMD Display driver. 
-This observation is because duplicated code artifacts involve copy-pasted code that imports different configuration files.
-The design of the AMD Display driver configuration files necessitates importing specific configurations at compile time, 
-which prevents a generic approach. Consequently, refactoring this pair requires changes in how the AMD Display driver 
-manages configuration files -- a complex task that demands input from an expert or maintainer.
-
-The refactoring of the second function pair, however, was successful. We submitted this refactoring as a patch to the AMD 
-Display driver to obtain feedback from the maintainers. The process of sending the patch took more time than expected.
-We initially submitted the patch on August 9th, 2024, but we did not receive replies. As a result, we had to resend 
-it on October 9th, 2024 and received an initial response on November 3rd, 2024.
-
-We understood new details about the driver evolution processes in discussions with our partner from the AMD Display driver.
-
-Regarding the patch content, only minor changes were requested to align with code 
-style and best practices that were initially unknown to us. More details of the 
-patch can be found in Appendix \ref{app:feedback}.
+because we are not applying the systematic approach, which translate to smaller code changes.

### 2025-05-25 conteudo/results/students.tex
add some work from 24-05

@@
 
-Appendix (CREATE APPENDIX AND ADD HERE X) has references
+Appendix \ref{app:stuart} has references
 to the artifacts discussed in this section, including the duplication function pairs that students 
@@
 
-Table (CREATE TABLE AND ADD HERE X) summarizes the mitigation approachs of the students in terms 
-of lines changed and refactor methods used. BLABLABLA
+Table \ref{tab:stu} summarizes the mitigation approachs of students's groups in terms 
+of lines changed and refactor methods used. A total of seven groups applied the parameterize refactor method,
+and two groups applied the extractor method on their approaches to mitigate the duplications. 
+The other three groups does not applied any refactor methods presented in the literature. 
+We discuss the groups's approaches and the subsystem maintainers opinions on the patches that students
+sent.
+
+Groups one and four have worked in similar duplications, where there are two functions returning a boolean mirroring
+the results. That means, when the first function returns true, the second function returns false, and when the first
+functions returns false, the second function return false. The group one have a simple approach and the maintainers 
+approved the changes quickly. The group four attempted to mitigate the duplication with macros, which is a method 
+found in the code file they were working. The maintainers disagree with the proposed approach, arguing 
+that the proposed changes decreases the code readability and asked for a approach similar to the group one. 
+The group four submitted the patch after coding the approach proposed by the maintainers.
+
+Group three, six and ten created a struct to encapsulate the parameters to mitigate the duplication when 
+applying the parameterize method. The group three mitigation was send to the maintainers and the approved 
+without significant comments, while the patch from group ten was reject by the same maintainer that approved
+the patch from group three. The maintainer justification to reject the patch is that group ten refactored 
+code files that has simple context, threfore adding a layer of abstraction to remove a duplication 
+deacrease the code readability without clear gains on the codebase quality.
+
+Group seven and eight used the extract method to approach the mitigations. The group eight send the 
+mitigations to the maintainers and the code changes was approved with minor requested changes
+about code style and smaller errors fix. For the group seven mitigation, the maintainers reject 
+their patch because the code changes adds layers of abstractions on the code, turning it more complex
+and harder to read. The maintainers proposed to the group seven to remove the duplicated functions 
+and insert the duplicated code multiple times were the functions were being called. 
+This suggestion we could think that it is a bad practice for code quality, as adds more complexity and 
+responsibality to a bigger function, it is view as the best choice for removing layers of abstraction 
+and making the code easier to read.
+
+Group nine and eleven faced similar issue as us, where configurations files
+designed as macros complicates to approach the mitigations. Group nine applied the parameterize method
+to not need to know the configurations in the generic function, which is reasonable as the duplication
+approach only depends on one configuration value. Group eleven worked with multiple configuration values,
+and to refactor the duplication they proposed to create a bigger macro that encapsulates the other macros,
+which is a clever solution but adds abstraction and complexity on macros, which makes the code harder
+to debug and read. We do not get the feedback of the maintainers on the group eleven approach to know their 
+opinions.
+
+Group twelve is the student that executed the tool and pass throught the process of findings duplications.
+The student found a duplication in the driver that happens accross eighteen code files, creating a patch
+of removing four hundreds lines of duplicated code. 
+The student send the mitigations to the maintainers and their feedback pointed to fix warning in the code, 
+without arguing about removing the duplications.
+
+Analysing the students approaches for the duplications and the maintainers feedback, 
+we saw that people without previous experience in the linux kernel development could approach 
+duplications found by the ArKanjo tool and become contributors to the kernel. We observed that 
+not all duplications are view as code with bad quality, with the maintainers analysing the trade off
+between what is the purpose of the duplicated code, the code readability, and the levels of abstraction 
+added to address the duplications.
 

### 2025-05-25 conteudo/results/table_students.tex
add some work from 24-05

@@
 
-\textbf{Mitigation} & \textbf{Subsystem} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
+\textbf{Group} & \textbf{Subsystem} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
 \\ \hline
 
-1 & IIO & 0 & 0 & None \\ \hline
-2 & IIO & 0 & 0 & None \\ \hline
-3 & IIO & 0 & 0 & None \\ \hline
-4 & IIO & 0 & 0 & None \\ \hline
-5 & IIO & 0 & 0 & None \\ \hline
-6 & IIO & 0 & 0 & None \\ \hline
-7 & IIO & 0 & 0 & None \\ \hline
-8 & IIO & 0 & 0 & None \\ \hline
-9 & IIO & 0 & 0 & None \\ \hline
-10 & IIO & 0 & 0 & None \\ \hline
-11 & IIO & 0 & 0 & None \\ \hline
-12 & AMD & 0 & 0 & None \\ \hline
+1 & IIO & 7 & 14 & None \\ \hline
+2 & IIO & 23 & 24 & Parameterize method \\ \hline
+3 & IIO & 12 & 30 & Parameterize method \\ \hline
+4 & IIO & 2 & 7 & None \\ \hline
+5 & IIO & 17 & 70 & Parameterize method \\ \hline
+6 & IIO & - & - & Parameterize method \\ \hline
+7 & IIO & 21 & 22  & None \\ \hline
+8 & IIO & - & - & None \\ \hline
+9 & IIO & 33 & 52 & Parameterize method \\ \hline
+10 & IIO & 32 & 58 & Parameterize method \\ \hline
+11 & IIO & - & - & Parameterize method \\ \hline
+12 & AMD & 89 & 489 & None \\ \hline
 

### 2025-05-25 conteudo/methods/methods.tex
fix small typos on text

@@
 to approach code duplication in some of the Linux subsystems. 
-To validate if the Arkanjo tool is capable of finding code duplications, 
+To validate if the ArKanjo tool is capable of finding code duplications, 
 applied two independent methods and perform triangulation on the results.
@@
 by the tool in the IIO subsystems and the AMD Display Driver and pass the process of send the mitigations as PATCH.
-(DOUBLE CHECK IF THE STUDENTS HAVE SENT THE PATCHES).
-
-\todo[inline]{I SHOULD mention that the test on the tool were made using the gensim similarity method and not the git method.}
 
@@
 
-\todo[inline]{I just mention the problem, maybe a more explictly mention of the problem here or on the background would be good.}
-
 Due to the limitations of the BigCloneBench dataset, as noted by \cite{bigfail}, and its lack of suitability 
@@
 We empirically evaluated these pairs to determine if they were duplications, comparing these results with the literature-based method to assess our tool's effectiveness and understand the impact of the minimum similarity parameter.
-
-\todo[inline]{Take care: Big focus on AMD display driver and the tone of the writing supports the focus. }
 
@@
 
-\subsection{Research with students}
-
-\todo[inline]{Research with students does not look good.}
+\subsection{Study with University Students}
 
@@
 of this work at the first semester of 2025, we have proposed to the students the possibility
-of mitigating duplications in the IIO subsystem and the AMD display driver found with the Arkanjo tool.
+of mitigating duplications in the IIO subsystem and the AMD display driver found with the ArKanjo tool.
 The course had a total of 26 students, and of that, X students being undergraduate students
@@
 
-For the first alternative, one of the teaching assistants used the Arkanjo tool in the IIO subsystem and
+For the first alternative, one of the teaching assistants used the ArKanjo tool in the IIO subsystem and
 listed interesting duplicated functions pair that the groups can simple mitigate this specific duplication pair, 
@@
 maintainers to review. For the second alternative, we propose a similar experience then the 
-participant observation method, where the students pass through the whole process of executing the Arkanjo tool
+participant observation method, where the students pass through the whole process of executing the ArKanjo tool
 in the AMD display driver, analyze the duplications found, create a patch to mitigate a duplication and sent
@@
 
-For the teaching assistant and the student that executed the Arkanjo tool to explore duplications, 
+For the teaching assistant and the student that executed the ArKanjo tool to explore duplications, 
 we have asked them to answer some questions related to the experience of using the tool and what could be

### 2025-05-25 conteudo/tool/functionalities.tex
change tool figures
fix small typos on text

@@
 \begin{figure}
-\includegraphics[scale=0.5]{explorer_example}
-\caption{Example of the \textit{Duplication Explorer} functionality of the Arkanjo tool in action.}
+\includegraphics[scale=0.30]{explorer_example}
+\caption{Example of the \textit{Duplication Explorer} functionality of the ArKanjo tool in action.}
 \label{fig:explorer_ex}
@@
 \begin{figure}
-\includegraphics[scale=0.5]{function_example}
-\caption{Example of the \textit{Function Information} functionality of the Arkanjo tool in action.}
+\includegraphics[scale=0.3]{function_example}
+\caption{Example of the \textit{Function Information} functionality of the ArKanjo tool in action.}
 \label{fig:function_ex}
@@
 \begin{figure}
-\includegraphics[scale=0.5]{relatory_example}
-\caption{Example of the \textit{Duplication Report} functionality of the Arkanjo tool in action.}
+\includegraphics[scale=0.3]{relatory_example}
+\caption{Example of the \textit{Duplication Report} functionality of the ArKanjo tool in action.}
 \label{fig:relatory_ex}

### 2025-05-25 conteudo/tool/tool.tex
fix small typos on text

@@
 
-Arkanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
+ArKanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
 help developers identify code clone duplication at the function level.
@@
 
-The Arkanjo tool has been implemented and is available for reference at
+The ArKanjo tool has been implemented and is available for reference at
 \url{https://github.com/LipArcanjo/arkanjo}.

### 2025-05-25 conteudo/apendices/survey.tex
add appendix with initial survey

\chapter{Survey Results}

\label{app:survey}

This appendix includes the questions asked in the survey for the teaching assistant and the
student that used the ArKanjo tool. Teaching assistant's answers are showed as answer one and 
the student's answers are showed as answer two.

\paragraph{Question 1}

\

\textbf{Question:} How was your experience with the installation process of the tool? 
If there were any difficulties during the installation, please indicate them. 
Please rate the tool from one to five in regard to this aspect.

\textbf{Answer 1:} At first I've tried to install it on Arch Linux
(which is not officially supported by ArKanjo), and had some trouble to find and install 
the corresponding dependencies. Then, I've decided to run it on an Ubuntu container and 
had no problems with the installation. Score: 5 out of 5.

\textbf{Answer 2:}

\paragraph{Question 2}

\

\textbf{Question:}How was your experience with the documentation and using the tool in general? 
Do you feel there were any missing or unclear parts in the documentation? If so, which ones?
Rate the tool  from one to five related to the documentation.

\textbf{Answer 1:} There were some misleading information about the usage of some flags, 
but Luan clarified it all and committed to correct the documentation. These problems, 
regarding similarity percentage for queries, didn't cause significant issues to use ArKanjo, 
but limited my workflow a bit. Score: 4 out of 5.

\textbf{Answer 2:}

\paragraph{Question 3}

\

\textbf{Question:}Regarding the preprocessing step to find duplicates, where you pass the codebase. 
Did you experience any difficulties with this process? How was the user experience? 
Was it clear enough what the process was doing and why there might be a delay in execution?
Rate the tool  from one to five related to the preprocessing step.

\textbf{Answer 1:} Running the preprocessor was intuitive and the documentation covered 
it all correctly. My only complaint is that I was only able to preprocess with the gensim method. 
The diff strategy kept running for hours and didn't complete when I gave up. Score: 4 out of 5.

\textbf{Answer 2:}

\paragraph{Question 4}

\

\textbf{Question:} Regarding executing the commands implemented in the tool, did you 
have difficulty understanding what the tool's output represented and how to execute the commands?
What is your opinion on the formatting and presentation of the output? Would you change anything 
in the output presentation that you believe would improve the experience?
Rate the tool  from one to five related to executing and understanding the output of the commands.

\textbf{Answer 1:} The command outputs were very intuitive and easy to analyze. My only complaint 
is that some of it is colored and this is problematic to output directly onto a file for further analysis, 
since the color chars are preserved. I would suggest a new flag that decides if output is highlighted 
with colors, and either default it to no-color or having a clear documentation about the output colors.
Score: 4 out of 5.

\textbf{Anyyswer 2:}

\paragraph{Question 5}

\

\textbf{Question:} Did the tool meet your expectations? Do you believe it effectively 
fulfills its purpose, that is, detecting code duplications?

\textbf{Answer 1:} It did fill my expectations. I've collected a bunch of code duplications with ArKanjo,
presented the most promissing results to other people and most of them resulted in merged patches to the 
IIO and DRM subsystems.

\textbf{Anyyswer 2:}

\paragraph{Question 6}

\

\textbf{Question:} Based on your current knowledge of the tool, do you have any ideas 
for new features that you consider relevant or interesting?

\textbf{Answer 1:} I would suggest polishing the diff heuristic to find duplications and/or add new ones. 
Setting output colors with flags would be a good addition. Flags to highlight duplications that happen on 
a same file or directory would also be good.

\textbf{Anyyswer 2:}

\paragraph{Question 7}

\

\textbf{Question:} Please provide any additional comments or insights you would like to 
share that have not been covered in the previous questions.

\textbf{Answer 1:} For some decentralized codebases, where a directory is maintained by different people 
and open source communities, such as the Linux Kernel, it would be interesting to be able to filter 
duplications that happen on a same file or directory. When using ArKanjo, I had to use external scripting 
and AI to do this, and it would be great if ArKanjo had built-in features for output analysis.
Otherwise, great tool that helped the improvement of code quality and reduce technical debt.

\textbf{Answer 2:}

### 2025-05-25 conteudo/methods/methods.tex
add appendix with initial survey
add paragraph about contrib and remove some todo

@@
 we have asked them to answer some questions related to the experience of using the tool and what could be
-improved. The questions asked is the following:
+improved. The survey with the users's answers can be found in Appendix \ref{app:survey}.
+
+On a second stage of the Free Software Development course, we propose to students to contribute directly 
+with the ArKanjo tool development without help from the tool creator, to verify if the tool can be a 
+free software where multiple people can contribute. The tool's issue approached by the students where 
+of our knowledge our pointed to us in the survey feedback. A total of three students forming two groups 
+worked in direct contributions, a not a very expressive number compared with the initiate experiment.

### 2025-05-26 conteudo/results/students.tex
fix small typos on text
finish work of 25-05

@@
-\section{Students}
+\section{Study with University Students}
 
@@
 
-\subsection{Questionary Answers}
+\subsection{Survey Results Analysis}
 
 BLABLABLA
+
+\subsection{Students Contribution to the Tool}
+
+For the two groups opted to contribute directly to the ArKanjo tool development, 
+we listed multiple issues existed in the tool. The first group opted for working 
+in expand the tool suport for multiple operation systems. Some libraries used in the 
+tool development only had support for the Ubuntu distribution, and the first group was 
+responsible to migrate these libraries to libraries that have support to multiple operation
+systems. This issue was of our knowledge and pointed by the teaching assistant in the survey, as 
+he faced the issue while using the tool.
+
+The second group opted for working in a webpage for the tool documentation and support 
+for light and dark mode. The colored output presented in the figures in Chapter \ref{cha:tool} 
+were hard coded at first, does not considering the user terminal colors, 
+thus giving a bad user experience. The website is helpfull to 
+faciliate people to contribute or use the tool. We did not know how to approach the light and 
+dark mode at first and have not needed to look on how
+to fix the issue as the group search and found a solution on his own.
+
+The students successfully delivered the issues fix proposed. We found interesting to see 
+people able to contribute to the tool on issues that even us did not know how to approach.
+The issues helps the user experience and facilitates other researches using the ArKanjo tool 
+that we may explore in future works. Appendix \ref{app:contribution} has references to the
+students contributions, the issue list proposed to the students and figures of the 
+documentation webpage.

### 2025-05-27 conteudo/apendices/survey.tex
add full answers of survey

@@
 
-\textbf{Answer 2:}
+\textbf{Answer 2:} It was clear. The installation process was easy, just copy and paste. Score: 5 out of 5.
 
@@
 
-\textbf{Answer 2:}
+\textbf{Answer 2:} The readme is direct and explains how to use the tool. An example would be nice,
+something like a video or gi. It also would be good to have more detailed documentation
+or hints for the use process. Score: 4 out of 5.
 
@@
 
-\textbf{Answer 2:}
+\textbf{Answer 2:} I experienced a few errors. The first was related to disk space. I selected a path with a
+lot of data, a huge code base, so it takes many minutes and disk space. In some cases,
+the tool stops with some error but fortunately saves the process before stopping.
+Score: 4 out of 5.
 
@@
 
-\textbf{Anyyswer 2:}
+\textbf{Anyyswer 2:} The results are clear and text colors help to understand them better. The final results are
+displayed but not saved, it would be nice to also save them for future reference. Score: 4 out 5.
 
@@
 
-\textbf{Anyyswer 2:}
+\textbf{Anyyswer 2:} Yes, I detected many duplicates, it was fast and simple to use.
 
@@
 
+\begin{itemize}
+
+\item Multiple search, enables the possibility to have several tmp files and to select one
+of them.
+
+\item Different methods to find duplicates.
+
+\item Define a threshold to the number of repeated lines, for example, find duplicates
+with more than 90\% of similarity and more than 20 lines repeated.
+
+\end{itemize}
+
 \paragraph{Question 7}
@@
 
-\textbf{Answer 2:}
+\textbf{Answer 2:} It was a great tool and a good experience.

### 2025-05-27 conteudo/results/students.tex
add work 26-05

@@
 
-BLABLABLA
+We asked for the teaching assistant and the student that used the ArKanjo to answer a survey with some 
+question related to the tool experience, such as installation, documentation, facility of use, new ideas
+to improve the tool, etc. Appendix \ref{app:survey} has the complete survey with the answers.
+
+Related to the tool installation, the teaching assistant nor the students faced issues installing on the
+Debian distributions, although the teaching assistant attempted to install in the Arch Linux distribution and
+failed to install on this linux distribution.  About the tool documentation, the teaching assistant pointed errors in 
+the documentation that worsed his experience. The student asked for more examples 
+and tutorials on using the tool, in text or video formats.
+
+On the ArKanjo tool preprocessing step, the student reported problems of disk space usage and crash errors,
+which we knew that could happen as the tool iterates and can save information about every pair of functions in
+a codebase, as we did not implemented any mechanism to control the memory usage. The teaching assistant reported 
+to only be able to execute the preprocessor with the \textit{gensim} similarity method, and the \textit{diff}
+method kepted running for hours without finishing. This was a unexpected behaviour to us, as the \textit{gensim}
+is a text similarity method while the \textit{diff} method is a simple counter of common lines with a operation system
+default command. The teaching assistant report corroborate the \textit{gensim} efficiency.
+
+Related to the functionalites of the tool, both the student assistant and the student 
+reported a missing of a mechanism to save the tool results on a separated file instead of the terminal output.
+The tool by default outputted with colors as show in the section \ref{subsec:func}, which complicates to redirect
+the terminal output to a file.
+
+We asked the overall impression of the tool and if met their expectations and to give a score from one 
+to five to installation, documentation, preprocessing, and functionalites. 
+Both the teaching assistant and the student gave a score of at least 4 to each mentioned point,
+thought that the tool fulfill their expectations and have a overall good experience.
+
+Finally, the teaching assistant the the student reported features ideas to the tool that may be interesting, which 
+is as follow:
+
+\begin{itemize}
+
+\item Enable the possibility to have several codebases preprocessor artifacts and to select one of them.
+
+\item Addition of different methods to find duplication.
+
+\item Addition of more filters to search duplication, such as , find duplicates with
+more than 90\% of similarity and more than 20 lines repeated.
+
+\item Addition of options to highligh and filter duplications that happen on same file or directory.
+
+\item Addition of support to save the tool output in files without manual preprocessing of the colored output.
+
+\end{itemize}
+
+With the survey feedback, we understood the the users had a good experience and found value in the ArKanjo tool,
+and provided critical points that could be improved in the tool. The most critical issues reported had been fixed,
+such as documentation errors and support for multiple linux distribution and different operation system. 
+The less critical issues became a open point that can be addressed in future works.
 

### 2025-06-01 conteudo/background/linux.tex
add ref to ubuntu and linux

@@
 recognized by the public are actually GNU/Linux operating systems, which
-include various essential applications alongside the Linux kernel
-\citep{gnuref}.
+include various essential applications alongside the Linux kernel,
+such as system
+\citep{gnuref}. Examples of Linux distributions are Ubuntu \citep{ubuntu} 
+and Arch Linux \citep{archlinux}.
 

### 2025-06-01 conteudo/introduction/intro.tex
add updated linux information

@@
 Maintaining the Linux kernel is a complex task, as the project contains more
-than 19 million lines of code and has had over thirteen thousand contributors
-throughout its lifespan \citep{linuxquantity}. Given the magnitude of the Linux
+than 28 million lines of code and has had over twenty thousand contributors
+throughout its lifespan
+\footnote{Data directly extracted from Linux github repository \citep{linuxrepo}.
+Counted  28783390 lines of code and 20699 number of contributors in May 29th, 2025.
+Number of contributors counted as number of unique full name listed as author of a patch.}.
+Given the magnitude of the Linux
 kernel project, it is unrealistic to expect that all contributors have
@@
 the maintenance of the driver.
-
-\todo[inline]{Require updated information about the linux kernel. Required from qualification exam.}
 

### 2025-06-01 conteudo/results/table_students.tex
some work

@@
 6 & IIO & - & - & Parameterize method \\ \hline
-7 & IIO & 21 & 22  & None \\ \hline
-8 & IIO & - & - & None \\ \hline
+7 & IIO & 21 & 22  & Extract method \\ \hline
+8 & IIO & - & - & Extract method \\ \hline
 9 & IIO & 33 & 52 & Parameterize method \\ \hline

### 2025-06-01 conteudo/conclusion/conclusion.tex
add initial conclusion

The Linux Kernel's massive scale, with over 28 million lines of code 
and more than twenty thousand contributors, presents significant maintenance 
challenges; code duplication persists as a notable issue within its codebase, 
potentially hindering its evolution and patching processes. 
Prevailing academic and existing free software approaches to duplication 
detection often lack the scalability for comprehensive codebase analysis or 
offer limited functionality unsuitable for the kernel's complexity, 
generally failing to deliver actionable results or guidance for mitigation 
within its specific context.

In response to these limitations, this research addressed the challenge of 
code duplication within the Linux Kernel through the development of ArKanjo, 
command-line tool for Linux kernel maintenance designed to detect and 
analyze function-level duplications. Released under the MIT license, 
the architectural design of ArKanjo, incorporating a Preprocessor and a 
Query Responder, separates computationally intensive analysis from efficient 
querying for duplications within large codebases. 
Validation procedures, employing both the BigCloneBench dataset and empirical 
analysis conducted on the AMD Display driver, indicated ArKanjo's proficient 
performance in identifying Type-1 and Type-2 duplications, notably when 
similarity thresholds are configured between 80\% and 100\%.

The ethnographic studies, which involved participant observation and a project 
with university students, provided a realistic view of both the opportunities 
and the considerable challenges in mitigating code duplications found by ArKanjo. 
The tool was effectively used by first-time contributors to the Linux kernel to 
identify duplications and prepare patches. While this research included a
successfully merged patch by the author, and successful submissions from many 
student groups, a notable portion of student efforts 
encountered significant hurdles. These difficulties often stemmed from maintainer 
feedback where proposed changes, though reducing duplication, were rejected or 
required substantial rework due to concerns about code readability, added abstraction, 
or the specific context of the code. Technical complexities, such as C macros in 
configuration files, and lengthy patch review times also contributed to the challenges. 
Furthermore, the studies confirmed that maintainers do not always view code duplication 
as inherently negative, especially when it may improve clarity or facilitate 
independent hardware support.

In its entirety, this research contributes the ArKanjo tool as an operational 
method for detecting code duplication, and importantly, provides insights into 
the process of managing such duplications within the Linux kernel subsystems 
studied. The results show that while ArKanjo can effectively identify 
duplicated functions, the path from detection to an accepted mitigation within 
the kernel involves careful consideration of the specific code, established 
community practices, and the maintainers' perspectives on code quality and readability.

\section{Future Work}

The work presented in this thesis suggests several directions for future investigation. 
Enhancements to the ArKanjo tool could be pursued, such as investigating a multi-stage 
detection strategy. This strategy might employ a computationally lighter method for an 
initial, broad pass to identify candidate duplications, followed by more 
resource-intensive, state-of-the-art techniques for refined analysis and improved 
detection of complex clone types. Such an approach could complement efforts for 
improved resource management during its preprocessing phase for very large codebases.
The tool's language support could also be broadened by
developing parsers for additional programming languages, building upon its adaptable
architecture. Beyond the tool itself, future research could extend the empirical
analysis of code duplication and mitigation strategies to a broader range of Linux kernel 
subsystems or other large-scale C projects, to determine the wider 
applicability of the observations made.

### 2025-06-01 conteudo/introduction/intro.tex
add some changes to not be to broad

@@
 to identify and mitigate code duplications specifically within the Linux
-kernel. Our approach includes the development of a tool capable of detecting
+kernel subsystems, focuding our study on the AMD Display Driver and the Industril IO subsytem. 
+Our approach includes the development of a tool capable of detecting
 duplicated functions and establishing strategies for
@@
 identifying code duplications at the function level within C codebases. We
-expect this tool to effectively detect code duplications in the Linux kernel,
+expect this tool to effectively detect code duplications in the targeted subsytems of the Linux kernel,
 as preliminary investigations suggest that kernel code files often contain a
@@
 Upon completing the first phase, we anticipate having a validated tool for
-identifying code duplications within the Linux kernel context. We will then
+identifying code duplications within the context of the Linux kernel subsytems studied. 
+We then
 proceed to the second phase, which focuses on mitigating these

### 2025-06-01 conteudo/methods/methods.tex
add some changes to not be to broad

@@
 This section outlines our approach for mitigating code duplication in the Linux kernel
-and describes the ethnographic methodology for engagin with the Linux community. Firstly, 
+and describes the ethnographic methodology for engaging with the Linux community. Firstly, 
 we present the participant observation research method explored by the author. Then, 

### 2025-06-01 conteudo/results/eval.tex
Add work on 17/05
add some changes to not be to broad

@@
 as mentioned in \ref{subsec:code_finder}.
+
+\section{Tool Results on BigCloneBench}
 
@@
 there is a notable increase in function pairs that are not duplications within 
-the Linux kernel. For similarity ranges of 79\%-81\% or higher, there is a high 
+the within the context of the AMD Display driver (the Linux kernel component studied here). 
+For similarity ranges of 79\%-81\% or higher, there is a high 
 percentage of code duplications. At 69\%-71\% similarity, the duplication rate 

### 2025-06-01 conteudo/resumoabstract.tex
gray -> grey
add some changes to not be to broad

@@
 conseguem resolver os desafios específicos do kernel Linux. Para abordar esse tema, 
-esta pesquisa foca em dois objetivos principais. Primeiro, propõe uma nova ferramenta 
-para identificar duplicações de código dentro do kernel Linux. Segundo, analisa as 
+esta pesquisa foca em dois objetivos principais. 
+Primeiro, propõe uma nova ferramenta projetada para identificar duplicações de código dentro 
+do kernel Linux, com sua aplicação prática e eficácia avaliadas em subsistemas chave: o AMD 
+Display Driver e o subsistema Industrial I/O.
+Segundo, analisa as 
 duplicações identificadas para extrair padrões e métodos de refatoração, oferecendo 
@@
 in large-scale codebases like the Linux kernel, nor do they guide on resolving detected 
-duplications. Despite searches in both formal and gray literature, existing solutions fall 
+duplications. Despite searches in both formal and grey literature, existing solutions fall 
 short of addressing the specific challenges of the Linux kernel. To address this topic, this 
-research focuses on two key goals. First, it proposes a new tool to identify code duplications 
-within the Linux kernel. Second, it analyzes the identified duplications to extract patterns 
+research focuses on two key goals. 
+First, it proposes a new tool designed for identifying code duplications within the Linux kernel, 
+with its practical application and effectiveness evaluated in key subsystems: the AMD Display Driver 
+and the Industrial I/O subsystem.
+Second, it analyzes the identified duplications to extract patterns 
 and refactoring methods, offering guidance to future contributors.

### 2025-06-01 conteudo/background/linux.tex
fix grammar with grammarly

@@
 This chapter presents an overview of fundamental topics needed to understand 
-the research, such as Linux Kernel, refactoring, types of code duplication, 
+the research, such as the Linux Kernel, refactoring, types of code duplication, 
 code duplication detection methods presented in the literature and ways to 
-evaluating these methods. This foundation will make it easier to grasp the 
+evaluate these methods. This foundation will make it easier to grasp the 
 concepts discussed in the research.
@@
 
-The Linux kernel is monolithic and organized into a set of subsystems, such as
+The Linux kernel is monolithic and organized into subsystems, such as
 the process scheduler, memory management, device driver infrastructure,
@@
 version control system. These contributions are formatted as patches—text
-documents that outline the differences between two versions of the source code.
+documents that outline the differences between two source code versions.
 These patches are then submitted and reviewed via the Linux mailing lists
@@
 complicates the maintenance of this subsystem. This problem serves as the
-starting point for our research, with the AMD Display Driver as our primary
+starting point for our research, with the AMD Display driver as our primary
 object of study as we aim to understand and mitigate code duplication in the
@@
 
-Another subsytem within the Linux kernel explored in this work is the Industrial
-I/O subsytem (IIO). From the Linux kernel documentation, ''The main purpose of the 
-Industrial I/O subsystem (IIO) is to provide support for devices that in some sense 
-perform either analog-to-digital conversion (ADC) or digital-to-analog conversion (DAC) or both``\citep{iiodoc}.
-The IIO subsystem contains more then X lines of code and more then Y code files 
+Another subsystem within the Linux kernel explored in this work is the
+Industrial I/O subsystem (IIO). According to the Linux kernel documentation, 
+''The main purpose of the Industrial I/O subsystem (IIO) is to provide support 
+for devices that in some sense perform either analog-to-digital conversion (ADC) 
+or digital-to-analog conversion (DAC) or both''\citep{iiodoc}. The IIO subsystem 
+contains more than X lines of code and more than Y code files
 \footnote{
@@
 }.
-We opted to include the IIO subsystem in our experiments to ensure our findings are not limited to a single subsystem. 
-Additionally, we have an established point of contact within the IIO subsystem to facilitate this work.
-. Like other subsystems, discussions about the IIO subsytem are conducted on its 
+We opted to include the IIO subsystem in our experiments to ensure our findings 
+are not limited to one subsystem. Additionally, we have an established point of
+contact within the IIO subsystem to facilitate this work.
+Like other subsystems, discussions about the IIO subsytem are conducted on its 
 designated mailing list, which can be found here: 
@@
 The \textit{diff} command is a command line in Linux that allows the user to compare 
-two files line by line \citep{diffcommand}. There are multiple forms 
-that the results of the command can be shown, one example of the command being executed 
-can be found in Figure \ref{fig:diff}. With the command results, it is easy to visualize and extract 
-the number of lines that are equal and different between the two compared files.
+two files line by line \citep{diffcommand}. There are multiple forms in which the command 
+results can be shown to the user, and Figure \ref{fig:diff} shows one of them. 
+With the command results, it is easy to visualize and extract the number of lines that 
+are equal and different between the two compared files.
 
 \begin{figure}
-\includegraphics[scale=0.3]{diff_command}
+\includegraphics[scale=0.25]{diff_command}
 \caption{Example of \textit{diff} command executed}
-File 1 content is shown in red and file 2 content is shown in green. \textit{-} 
-represents a line that only exist in file 1, \textit{+} represents a line that only exist in file 2,
-\textit{!} represents a line that exist in both file but has modifications between them, and empty on the
-start of a line represents that the line exist in both files without any modifications.
+File 1's content is shown in red, and file 2's content is shown in green. \textit{-} 
+represents a line that only exists in file 1, \textit{+} represents a line that only 
+exists in file 2,\textit{!} represents a line that exists in both files but has modifications 
+between them, and empty at the start of a line represents that the line exists in 
+both files without any modifications.
 \label{fig:diff}
@@
 
-We use thhe \textit{diff} command as the base of one of the code duplication detection methods 
-on the tool proposed on this work, although it is not extensive explored as the text similarity
-method for code duplication detection.
+We use the \textit{diff} command as the base of one of the code duplication detection 
+methods on the tool proposed in this work, although it is not explored as the text 
+similarity method for code duplication detection.

### 2025-06-01 conteudo/introduction/intro.tex
fix grammar with grammarly

@@
 
-Our studies and observations reveal that the code duplication problem described
-in the previous section cannot be fully addressed by existing tools in
-either the formal or grey literature. To fill this gap, we propose an approach
-to identify and mitigate code duplications specifically within the Linux
-kernel subsystems, focuding our study on the AMD Display Driver and the Industril IO subsytem. 
-Our approach includes the development of a tool capable of detecting
-duplicated functions and establishing strategies for
-addressing these duplications. We tested the approach in the context of two
-kernel components, the AMD Display Driver the the Industrial IO subsystem.
+Our studies and observations reveal that existing tools cannot fully address the code duplication 
+problem described in the previous section in the formal or grey literature. To fill this gap,
+we propose an approach to identify and mitigate code duplications, specifically within the
+Linux kernel subsystems, focusing our study on the AMD Display driver and the Industrial 
+IO subsystem. Our approach includes the development of a tool capable of detecting duplicated 
+functions and establishing strategies for addressing these duplications. We tested the approach 
+in the context of two kernel components, the AMD Display driver and the Industrial IO subsystem.
 
@@
 code duplication issues within this context. To investigate our approach, 
-we are developing a tool to identify duplicate code. We are then refactoring 
-portions of this code and submitting patches to maintainers, providing evidence 
-of the tool's accuracy in detecting duplicate functions. Our research design follows 
-a multi-method approach, structured into two phases, each employing different research methods.
+we are developing a tool to identify duplicate code. We then refactor portions 
+of this code and submit patches to maintainers, providing evidence of the tool’s 
+accuracy in detecting duplicate functions. Our research design follows a 
+multi-method approach, structured into two phases, each employing different research methods.
 
 In \textbf{Phase I}, we propose, develop, and validate a tool capable of
-identifying code duplications at the function level within C codebases. We
-expect this tool to effectively detect code duplications in the targeted subsytems of the Linux kernel,
-as preliminary investigations suggest that kernel code files often contain a
-mix of duplicated and unique functions. To validate the tool, we will use a
-triangulation method, comparing its results with established code duplication
-database referenced in the literature, and conducting an empirical analysis on
-a randomly selected set of files from the AMD Display driver, which our tool
-flags as containing duplications. A detailed explanation of our validation
-process is provided in Chapter \ref{cha:method}.
+identifying code duplications at the function level within C codebases.
+We expect this tool to effectively detect code duplications in the targeted subsystems 
+of the Linux kernel, as preliminary investigations suggest that kernel code files often 
+contain a mix of duplicated and unique functions. We will use a triangulation method to 
+validate the tool, comparing its results with an established code duplication database 
+referenced in the literature and conducting an empirical analysis on a randomly selected 
+set of files from the AMD Display driver, which our tool flags as containing duplications.
+A detailed explanation of our validation process is provided in Chapter \ref{cha:method}.
 
-Upon completing the first phase, we anticipate having a validated tool for
-identifying code duplications within the context of the Linux kernel subsytems studied. 
-We then
-proceed to the second phase, which focuses on mitigating these
-duplications. In \textbf{Phase II}, our objective is to verify if we can 
-practically mitigate the duplication finds by the developed tool. 
-To achieve this, we have designed an ethnographic approach that enables us to
-interact with the Linux development community and gather artifacts that either
-support or challenge our proposed mitigation strategies. Figure
-\ref{fig:reDesign} summarizes the research steps proposed in this study.
+Upon completing the first phase, we anticipate having a validated tool for identifying 
+code duplications within the context of the Linux kernel subsystems studied. We then 
+proceed to the second phase, which focuses on mitigating these duplications. 
+In \textbf{Phase II}, we aim to verify if we can mitigate the duplication found by 
+the developed tool. To achieve this, we have designed an ethnographic approach that 
+enables us to interact with the Linux development community and gather artifacts that 
+either support or challenge our proposed mitigation strategies. Figure \ref{fig:reDesign} 
+summarizes the research steps proposed in this study.
 
@@
 \ref{cha:results} shows the results we had through our work, from the research methods to 
-evaluate our tool and the ethnographic studies. Chapter \ref{cha:conclusion}udes this research.
+evaluate our tool and the ethnographic studies. Chapter \ref{cha:conclusion} concludes this research.

### 2025-06-01 conteudo/methods/methods.tex
fix grammar with grammarly

--- 2025-05-25 conteudo/methods/methods.tex
@@
-We employ a multi-method research approach, using different methodologies to achieve our goals
-to approach code duplication in some of the Linux subsystems. 
-To validate if the ArKanjo tool is capable of finding code duplications, 
-applied two independent methods and perform triangulation on the results.
+We employ a multi-method research approach, using different methodologies to achieve our 
+goals to approach code duplication in some of the Linux subsystems. To validate if the 
+ArKanjo tool can find code duplications, we applied two independent methods and performed 
+triangulation on the results.
 
@@
 We chose triangulation because relying solely on the literature-based method has 
-limitations and may not be entirely reliable \citep{bigfail, litreview}. As a reminder, 
-all research approches were done using the text similarity method as the 
-code dupication method of the tool.
+limitations and may may be unreliable \citep{bigfail, litreview}. 
+As a reminder, all research approaches were done using the text similarity method 
+as the code duplication method of the tool.
 
-Given the validation of the tool's capabilities, we proceed to undertand the viability 
-to mitigate the code duplication finds by the tool. 
-To underdand the viability, our first choice is to conduct an ethnographic study using 
-a participant-observation methodology. 
-In this study, we aim to mitigate a subset of duplicated functions by refactoring them 
-while interacting with the Linux community to gather feedback on our mitigation efforts.
+Given the validation of the tool’s capabilities, we proceed to understand the viability 
+of mitigating the code duplication found by the tool. We conducted an ethnographic study 
+using a participant-observation methodology to understand its viability. In this study, 
+we aim to mitigate a subset of duplicated functions by refactoring them while interacting 
+with the Linux community to gather feedback on our mitigation efforts.
 
-The ethnographic study happened in two phases. 
-In the first phase, the author of this research explored the process of mitigating
-a code duplication found in the AMD Display Driver and sent it as a PATCH to iteract with the
-Linux community and get their feedback. On a second phase, we proposed to students on the 
-course Free Software Development at University São Paulo to mitigate the code duplications found 
-by the tool in the IIO subsystems and the AMD Display Driver and pass the process of send the mitigations as PATCH.
+The ethnographic study had two phases. In the first phase, the author explored the process 
+of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to 
+interact with the Linux community and get their feedback. In the second phase, we proposed 
+to students in the course Free Software Development at the University of São Paulo to mitigate 
+the code duplications found by the tool in the IIO subsystems and the AMD Display driver and 
+pass the process of sending the mitigations as PATCH.
 
@@
 
-This section outlines our approach for mitigating code duplication in the Linux kernel
-and describes the ethnographic methodology for engagin with the Linux community. Firstly, 
-we present the participant observation research method explored by the author. Then, 
-we present the methodolody of the experiment done with the students.
+This section outlines our approach to mitigating code duplication in the Linux kernel 
+and describes the ethnographic methodology for engaging with the Linux community. 
+Firstly, we present the participant observation research method explored by the author. 
+Then, we present the methodology of the experiment done with the students.
 
@@
 
-Given the tool evaluations that demonstrate the tool capabilites of finding duplications, 
-we proceeded to undertand if the duplications found are possible to be mitigate and
-the kernel subsystem's maintainers opinions about the duplications and their respective mitigations.
-To accomplish this objective, we initialy proposed a participant-observation experiment, 
-where we, people that have never contribute to the Linux kernel, 
-act as contributors to the AMD Display driver and collect artifacts of our experience in the process.
+Given the tool evaluations demonstrating the tool's capabilities of finding duplications, 
+we proceeded to understand if the duplications found can be mitigated and the kernel 
+subsystem's maintainers' opinions about the duplications and their respective mitigations. 
+To accomplish this objective, we initially proposed a participant-observation experiment, 
+where we, people who have never contributed to the Linux kernel, act as contributors to 
+the AMD Display driver and collect artifacts of our experience in the process.
 
-As a first step of this experiment, we executed the tool in the AMD Display driver codebase and 
-manually analysed the biggest duplications in terms of number of lines and raised to 
-two duplicated function pairs that we judged interesting to try to mitigate. 
-We have choose to approach the biggest duplications as a proof of concept that 
-we can make significant impacts. The two duplicated function pairs chosen are:
+As a first step of this experiment, we executed the tool in the AMD Display driver codebase. 
+We manually analyzed the biggest duplications regarding the number of lines and raised them 
+to two duplicated function pairs that we judged interesting to try to mitigate. We have 
+chosen to approach the biggest duplications as proof of concept so that we can make 
+significant impacts. The two duplicated function pairs chosen are:
 
@@
 
-Given the two duplicated function pairs, we observed that the files that the functions exist contains 
-multiple other duplications. Thus, we proposed a simple systematic approach to mitigate all the 
-functions duplicated in the context and not only the duplicated function pairs. 
+Given the two duplicated function pairs, we observed that the files in the functions 
+contain multiple other duplications. Thus, we proposed a simple systematic approach 
+to mitigate all the functions duplicated in the context, not just the duplicated 
+function pairs. 
+After we approached the duplications with the systematic approach, for the duplicated 
+function pairs that had successfully mitigated the duplication, we sent patches to the 
+AMD Display driver email list to validate the mitigations and see the maintainer's opinion 
+while documenting the process and our impressions.
 The systematic approach is defined below.
-
-After we approached the duplications with the systematic approach. For the duplicated function
-pairs that had successfully being mitigate the duplication, we sent patches to the AMD Display driver
-maillist to validate the mitigations and see the maintainers opinion, while documentaion the process 
-and our impressions.
 
@@
 (described in Section \ref{subsec:functioncommand}) to locate all duplications of the function in the codebase, 
-as it may have been duplicated more times then the two occurences of the duplicated function pair. 
+as it may have been duplicated more times than the two occurrences of the duplicated function pair. 
 This strategy resulted in a collection of duplicated functions and corresponding code files.
@@
 
-Posterior of the initial participant observation that we made, we decided to expand the experiment 
-by proposing to students mitigate duplication found by the tool and send patches to the kernel community.
+Based on our initial participant observation, we decided to expand the experiment by proposing 
+to students how to mitigate duplication found by the tool and send patches to the kernel community.
 
-In the Free Software Development course at University of São Paulo, ministred by the advisor 
-of this work at the first semester of 2025, we have proposed to the students the possibility
-of mitigating duplications in the IIO subsystem and the AMD display driver found with the ArKanjo tool.
-The course had a total of 26 students, and of that, X students being undergraduate students
-and Y begin graduate students.
+In the Free Software Development course at the University of São Paulo, taught by the advisor of 
+this work in the first semester of 2025, we proposed to the students the possibility of mitigating 
+duplications in the IIO subsystem and the AMD display driver found with the ArKanjo tool. 
+The course had 26 students, of whom X were undergraduate students and Y were graduate students.
 
-On the course, the students are asked to form groups of one, two or three students and together 
-make a contribution to a open source software. To auxiliate the students on this task, the 
-professor and teaching assistants prepares multiple alternatives on how to contribute to open source projects, 
-including simple contributions as updating documentation or code style fixes. In this context, 
-we have proposed two alternatives for the students related to the tool.
+In the course, the students are asked to form groups of one, two, or three students and contribute 
+to free software together. To assist the students in this task, the professor and teaching assistants 
+prepare multiple alternatives for contributing to free software projects, including simple contributions 
+such as updating documentation or code style fixes. In this context, we have proposed two options 
+for the students related to the tool.
 
-For the first alternative, one of the teaching assistants used the ArKanjo tool in the IIO subsystem and
-listed interesting duplicated functions pair that the groups can simple mitigate this specific duplication pair, 
-and it is up to the students on how
-to refactor the code to mitigate the duplications and pass through the process to send patches to the IIO 
-maintainers to review. For the second alternative, we propose a similar experience then the 
-participant observation method, where the students pass through the whole process of executing the ArKanjo tool
-in the AMD display driver, analyze the duplications found, create a patch to mitigate a duplication and sent
-the patch to the AMD display driver's maintainers.
+For the first alternative, one of the teaching assistants used the ArKanjo tool in the IIO subsystem 
+and listed interesting duplicated function pairs that the groups could mitigate this specific 
+duplication pair. It is up to the students to refactor the code to mitigate the duplications and 
+pass through the process of sending patches to the IIO maintainers for review. For the second 
+alternative, we propose a similar experience to the participant observation method, where the 
+students pass through the whole process of executing the ArKanjo tool in the AMD display driver, 
+analyze the duplications found, create a patch to mitigate a duplication, and send the patch to the 
+AMD display driver’s maintainers.
 
-From the students enrolled in the in the course, 25 students opted for the first alternative and 1 student
-opted for the second alternative, forming 11 groups working on the first alternative and 1 group 
-working on the second alternative. It is worth highlighting that none of the students
-that choose to mitigate duplications had made a contribution to the Linux kernel before,
-meaning that the patch to mitigate the duplication being the first tentative to contribute to the kernel as
-newcomers.
+From the students enrolled in the course, 25 students opted for the first alternative, and one 
+student opted for the second alternative, forming 11 groups working on the first alternative 
+and 1 group working on the second alternative. It is worth highlighting that none of the students 
+who chose to mitigate duplications had contributed to the Linux kernel before, meaning that the 
+patch to mitigate the duplications was the first tentative to contribute to the kernel as newcomers.
 
-The groups have been request to document the experience and approaches to refactor the duplications and the
-experience of submiting a patch to the Linux kernel in blogs. We had analyzed these blogs to undertand 
-patterns of refactor used for mitigating the duplications, the experience on send the patches and the 
-opinions of the subsystem maintainers on the mitigations.
+The groups have been requesting that we document the experience and approaches to refactoring the 
+duplications and the experience of submitting a patch to the Linux kernel in blogs. We analyzed 
+these blogs to understand the patterns of refactoring used for mitigating the duplications, the 
+experience of sending the patches, and the opinions of the subsystem maintainers on the mitigations.
 
-For the teaching assistant and the student that executed the ArKanjo tool to explore duplications, 
-we have asked them to answer some questions related to the experience of using the tool and what could be
-improved. The questions asked is the following:
+The teaching assistant and the student who executed the ArKanjo tool to explore duplications were 
+asked to answer some questions related to their experience using the tool and what could be improved. 
+The survey with the users' answers can be found in Appendix \ref{app:survey}.
+
+In the second stage of the Free Software Development course, we propose that students contribute 
+directly to the ArKanjo tool development without help from the tool creator to verify whether the 
+tool can be free software to which multiple people can contribute. The students approached the tool’s 
+issue, and to our knowledge, they pointed to us in the survey feedback. Three students forming two 
+groups worked in direct contributions, which was not a very expressive number compared with the 
+initial experiment.

### 2025-06-01 conteudo/results/participant.tex
some work
fix grammar with grammarly

--- 2025-05-18 conteudo/results/participant.tex
@@
 
-For our initial participant observation, we attempted to mitigate two code duplications found by ArKanjo tool
-in the AMD Display driver applying our systematic approach presented in \ref{subsubsec:systematic}. 
-As presented in section \ref{sec:meteth}, we opted for applying the systematic approach in the functions 
-\textit{offset\_to\_id} and \textit{phy\_id\_to\_atom} in the folders \textit{dc/gpio} and \text{dc/bios} respectively.
+For our initial participant observation, we attempted to mitigate two code 
+duplications found by the ArKanjo tool in the AMD Display driver by applying 
+our systematic approach presented in \ref{subsubsec:systematic}. 
+As presented in section \ref{sec:meteth},
+we opted to apply the systematic approach in the functions
+\textit{offset\_to\_id} and \textit{phy\_id\_to\_atom} in the folders 
+\textit{dc/gpio} and \text{dc/bios}, respectively.
 
 Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
-\textit{phy\_id\_to\_atom} will be referred to as mitigation 2. Table \ref{tab:patch} summarizes the impact of the mitigations
-in terms of files and lines changed, and the refactoring method used. We present our experience and learning for each mitigation 
-in the next sections.
+\textit{phy\_id\_to\_atom} will be referred to as mitigation 2. 
+Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
+files and lines changed and the refactoring method used. In the next sections, 
+we present our experience and learning for each mitigation.
 
@@
 
-On the process of the systematic approach, we found that the \textit{offset\_to\_id} function exist in 10 files, enclosing
-multiple GPU architectures. The duplicated functions are not exactly equals, as there are some minimal specific logics applied
-to each GPU architecture. To address this issue, we applied the extract method to break the function in smaller parts and the 
-parameterize method to address the specific logics for the GPU's.
+In the systematic approach, we found that the \textit{offset\_to\_id} function 
+exists in code 10 files, enclosing multiple GPU architectures. The duplicated 
+functions are not exactly equal, as some minimal specific logic is applied to 
+each GPU architecture. To address this issue, we used the extract method to 
+break the function into smaller parts and the parameterized method to address 
+the specific logic for the GPUs.
 
-The base idea on how we would mitigate the duplications found were done, but we could not complete the refactoring on a state 
-capable to submit as a patch to the driver. 
-We found that the configuration files on the AMD Display driver are designed to be imported at compilation time
-using \textit{\#define} macros. The configuration files design choice makes the refactoring of duplications on generic 
-approachs trickier, as refactoring code that depends on these files requires significant refactor on the configuration files
-design and deep knowledge of the codebase, which us as first-timer contributors do not have. Thus, we opted to not continue 
-on investigating this mitigation.
+We reached a point with a scratch on how to mitigate the duplications found, 
+but we could not complete the refactoring on a state capable of submitting as 
+a patch to the driver. We found that the configuration files on the AMD Display 
+driver are designed to be imported at compilation time using \textit{\#define} macros. 
+The configuration files design choice makes the refactoring of duplications on 
+generic approaches trickier, as refactoring code that depends on these files 
+requires significant refactoring on the configuration files design and deep 
+knowledge of the codebase, which we as first-timer contributors, do not have. 
+Thus, we opted not to continue investigating this mitigation.
 
@@
 
-\subsection{OLD}
+Using the systematic approach, we found that the \textit{phy\_id\_to\_atom}
+function exists in five code files, and two other functions are duplicated 
+in those five code files. All functions across the files were exactly equal, 
+so there was no need to apply the refactoring methods presented in the literature. 
+The refactoring work was resumed to create a generic library and fix the compilation targets.
 
-In the first function pair, completing the refactoring requires the expertise of a maintainer for the AMD Display driver.
-This observation is because duplicated code artifacts involve copy-pasted code that imports different configuration files.
-The design of the AMD Display driver configuration files necessitates importing specific configurations at compile time, which prevents a generic approach.
-Consequently, refactoring this pair requires changes in how the AMD Display driver manages configuration files -- a complex task that demands input from an expert or maintainer.
+The code files in the second mitigation do not depend on configuration files. Thus, 
+we did not face the same issues during the first mitigation. Refactoring the code 
+to mitigate duplications reached a point we judged was good enough to submit as a 
+patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
+the refactoring to the driver’s maintainers as a patch, documenting our findings and experiences 
+on the process.
 
-The refactoring of the second function pair, however, was successful.
-We submitted this refactoring as a patch to the AMD Display driver to obtain feedback from the maintainers.
-The process of sending the patch took more time than expected.
-We initially submitted the patch on August 9th, 2024, but we did not receive replies.
-As a result, we had to resend it on October 9th, 2024 and received an initial response on November 3rd, 2024.
+\begin{figure}
+\includegraphics[scale=0.45]{timeline_patch}
+\caption{Timeline of the PATCH submission process}
+\label{fig:timeline}
+\end{figure}
 
-We understood new details about the driver evolution processes in discussions with our partner from the AMD Display driver.
-Some of the code in the driver is shared across all operating systems that support the implemented GPUs.
-This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors.
+The experience of submitting a patch was not as we expected. The process took 7 months 
+and 16 days, which is significantly slower than expected. We initially submitted the 
+patch on August 9th, 2024, but we did not receive replies. Thus, we had to resend it 
+on October 9th, 2024. We received an initial response on November 3rd, 2024, asking for 
+minimal changes; we submitted a second version with the changes request on November 16th, 2024. 
+We did not receive replies at first and need to resend the patch on February 7th, 2025. 
+We received a reply on February 11th, 2025, asking for additional changes. We sent a third 
+version on February 24th, 2025, and the patch was accepted and integrated into the kernel 
+codebase on February 25th, 2025. Figure \ref{fig:timeline} illustrates the timeline of 
+the patch submission.
 
-Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering.
-One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others.
+We discussed with our point of contact at the AMD Display driver to understand the reason 
+for the slow process of submitting a patch. We learned that some of the code in the driver 
+is shared across all operating systems that support the implemented GPUs. This fact creates 
+a complex process within AMD to format and submit changes across the supported systems while 
+simultaneously implementing measures to mitigate errors. Additionally, we comprehended that 
+AMD Display driver developers can not view duplicated code negatively, as we understand it 
+is a good practice in software engineering. One example shared with us is that duplicated 
+code enhances the independence of GPU driver code, allowing developers to make changes to a 
+specific GPU without needing to test compatibility with others. 
 This approach helps save significant time and effort.
 
-Regarding the patch content, only minor changes were requested to align with code 
-style and best practices that were initially unknown to us. More details of the 
-patch can be found in Appendix \ref{app:feedback}.
+Regarding the patch changes requested for the second version, the maintainers only asked 
+for minor changes to align with code style, license correct use, and best practices that 
+were initially unknown to use. For the change requested on the third version, the maintainers 
+asked us to move the functions created in the generic library to an existing file instead of 
+creating a new one. More details of the patch can be found in Appendix \ref{app:feedback}, 
+and the patch submitted in the Linux codebase can be seen here:
+\url{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}.
+
+The initial plan of this research was to approach and send many mitigations to collect 
+many artifacts from the driver community. Still, the slow process of sending a patch 
+makes this plan unviable. Thus, we opted to redirect this research to work with students 
+and refactoring duplicated function pairs found by the tool directly in the IIO and the 
+AMD Display driver without passing through the whole process of the systematic approach 
+proposed. This alternative allowed us to collect artifacts from the students' experiences 
+and findings and not be too dependent on the interaction with the kernel community. 
+Another advantage is accelerating the interaction with the kernel community because we are 
+not applying the systematic approach, which translates to more minor code changes.

### 2025-06-01 conteudo/results/results.tex
fix grammar with grammarly

@@
 
-In this chapter we present and analyze the results found in the experiments and 
-methods done through our research. Following the cronology of the research, we 
-start by presenting the ArKanjo tool's evaluation through both a method from the literature
-and an empirical analyses. Then, we present and discuss the experience of the initial participant
-observation methodolody, where we mitigate a code duplication on the AMD Display driver and sent it as a patch.
-Finally, we show and discuss our discoveries on the experiment done with students at the course in University of 
-São Paulo, where we propose to students to mitigate duplications found by ArKanjo tool at IIO subsystem and the
-AMD Display driver and document their founds, opinions and experiences.
+In this chapter, we present and analyze the results of the experiments and methods 
+used in our research. Following the chronology of the study, we start by presenting the 
+evaluation of the ArKanjo tool through both a method from the literature and an empirical 
+analysis. Then, we present and discuss the experience of the initial participant 
+observation methodology, where we mitigate a code duplication on the AMD Display Driver 
+and send it as a patch. Finally, we show and discuss our discoveries on the experiment 
+done with students at the University of São Paulo, where we propose to students to mitigate 
+duplications found by the ArKanjo tool at IIO subsystem and the AMD Display Driver, and 
+document their founds, opinions, and experiences.
 

### 2025-06-01 conteudo/results/students.tex
fix grammar with grammarly

--- 2025-05-25 conteudo/results/students.tex
@@
-\section{Students}
+\section{Study with University Students}
 
-As presented in chapter \ref{cha:method}, we propose to students to mitigate duplications
-found by the ArKanjo tool and send patches to the linux subsystems maintainers, while
-documenting their experiences and learning through the process.
+As presented in Chapter \ref{cha:method}, we propose that students mitigate duplications 
+found by the ArKanjo tool, send patches to the Linux subsystem maintainers, and document 
+their experiences and learnings.
 
-For the students that worked
-in the IIO subsytems, we gave them duplicated function pairs found by the tool and their 
-responsability was to think of how to mitigate the duplication and sent the patch to the 
-maintainers. The student that worked in the AMD Display driver have the additional work
-to use the tool to find the duplications.
+For the students who worked in the IIO subsystems, we gave them duplicated function pairs 
+found by the tool, and their responsibility was to think of how to mitigate the duplication 
+and send the patch to the maintainers. The student that worked in the AMD Display driver 
+have the additional work to use the tool to find the duplications.
 
-Appendix \ref{app:stuart} has references
-to the artifacts discussed in this section, including the duplication function pairs that students 
-mitigate, the students's reports, the patch students sent to the linux kernel, and the 
-answers of the questionary asked to the student that worked in the AMD Display driver and the 
-teaching assistant that used the tool to find duplications in the IIO.
+Appendix \ref{app:stuart} has references to the artifacts discussed in this section, 
+including the duplication function pairs that students mitigate, the students’ reports, 
+the patch students sent to the Linux kernel, and the answers to the questionnaire asked 
+of the students who worked in the AMD Display driver and the teaching assistant that used 
+the tool to find duplications in the IIO.
 
@@
 
-Table \ref{tab:stu} summarizes the mitigation approachs of students's groups in terms 
-of lines changed and refactor methods used. A total of seven groups applied the parameterize refactor method,
-and two groups applied the extractor method on their approaches to mitigate the duplications. 
-The other three groups does not applied any refactor methods presented in the literature. 
-We discuss the groups's approaches and the subsystem maintainers opinions on the patches that students
-sent.
+Table \ref{tab:stu} summarizes students' groups' mitigation approaches regarding lines 
+changed and refactor methods used. Seven groups applied the parameterized refactor method, 
+and two applied the extractor method to their approaches to mitigate the duplications. 
+The other three groups do not use any refactor methods presented in the literature. 
+We discuss the groups' approaches and the subsystem maintainers' opinions on the patches 
+that students sent.
 
-Groups one and four have worked in similar duplications, where there are two functions returning a boolean mirroring
-the results. That means, when the first function returns true, the second function returns false, and when the first
-functions returns false, the second function return false. The group one have a simple approach and the maintainers 
-approved the changes quickly. The group four attempted to mitigate the duplication with macros, which is a method 
-found in the code file they were working. The maintainers disagree with the proposed approach, arguing 
-that the proposed changes decreases the code readability and asked for a approach similar to the group one. 
-The group four submitted the patch after coding the approach proposed by the maintainers.
+Groups one and four have worked in similar duplications, where two functions return a 
+boolean mirroring the results. That means when the first function returns true, the second 
+function returns false, and when the first function returns false, the second function 
+returns false. Group one had a simple approach, and the maintainers approved the changes quickly. 
+Group four attempted to mitigate the duplication with macros, a method found in the code file 
+they were working on. The maintainers disagreed with the proposed approach, arguing that the 
+proposed changes decrease the code readability, and asked for an approach similar to group one. 
+Group four submitted the patch after coding the approach proposed by the maintainers.
 
-Group three, six and ten created a struct to encapsulate the parameters to mitigate the duplication when 
-applying the parameterize method. The group three mitigation was send to the maintainers and the approved 
-without significant comments, while the patch from group ten was reject by the same maintainer that approved
-the patch from group three. The maintainer justification to reject the patch is that group ten refactored 
-code files that has simple context, threfore adding a layer of abstraction to remove a duplication 
-deacrease the code readability without clear gains on the codebase quality.
+Groups three, six, and ten created a struct to encapsulate the parameters and mitigate duplication 
+when applying the parameterized method. Group three sent the mitigation to the maintainers and 
+approved it without significant comments. In contrast, the patch from group ten was rejected by 
+the same maintainer who approved the patch from group three. The maintainer justification to reject 
+the patch is that group ten refactored code files that has simple context, threfore adding a layer 
+of abstraction to remove a duplication deacrease the code readability without clear gains on 
+the codebase quality.
 
-Group seven and eight used the extract method to approach the mitigations. The group eight send the 
-mitigations to the maintainers and the code changes was approved with minor requested changes
-about code style and smaller errors fix. For the group seven mitigation, the maintainers reject 
-their patch because the code changes adds layers of abstractions on the code, turning it more complex
-and harder to read. The maintainers proposed to the group seven to remove the duplicated functions 
-and insert the duplicated code multiple times were the functions were being called. 
-This suggestion we could think that it is a bad practice for code quality, as adds more complexity and 
-responsibality to a bigger function, it is view as the best choice for removing layers of abstraction 
-and making the code easier to read.
+Groups seven and eight used the extract method to approach the mitigations. Group eight sent the 
+mitigations to the maintainers, and the code changes were approved, with minor requested changes 
+in code style and smaller errors fixed. For group seven mitigation, the maintainers reject their 
+patch because the code changes add layers of abstractions to the code, making it more complex and 
+challenging to read. The maintainers proposed that group seven remove the duplicated functions and 
+insert the duplicated code multiple times where the functions were being called. This suggestion 
+may be bad practice for code quality, as it adds more complexity and responsibility to a bigger 
+function; it is viewed as the best choice for removing layers of abstraction and making the 
+code easier to read.
 
-Group nine and eleven faced similar issue as us, where configurations files
-designed as macros complicates to approach the mitigations. Group nine applied the parameterize method
-to not need to know the configurations in the generic function, which is reasonable as the duplication
-approach only depends on one configuration value. Group eleven worked with multiple configuration values,
-and to refactor the duplication they proposed to create a bigger macro that encapsulates the other macros,
-which is a clever solution but adds abstraction and complexity on macros, which makes the code harder
-to debug and read. We do not get the feedback of the maintainers on the group eleven approach to know their 
-opinions.
+Groups nine and eleven faced similar issues to those in Mitigaiton 1 on our participant observation 
+experiment, where configuration files designed as macros complicated the approach to the mitigations. 
+Group nine applied the parameterized method so as not to need to know the configurations in the 
+generic function, which is reasonable as the duplication approach only depends on one configuration 
+value. Group eleven worked with multiple configuration values, and to refactor the duplication, 
+they proposed to create a bigger macro that encapsulates the other macros, which is a clever 
+solution but adds abstraction and complexity to macros, which makes the code harder to debug and read.
+We do not get the feedback of the maintainers on the group eleven approach to know their opinions.
 
-Group twelve is the student that executed the tool and pass throught the process of findings duplications.
-The student found a duplication in the driver that happens accross eighteen code files, creating a patch
-of removing four hundreds lines of duplicated code. 
-The student send the mitigations to the maintainers and their feedback pointed to fix warning in the code, 
-without arguing about removing the duplications.
+Group twelve is the student who executed the tool and passed through the process of finding 
+duplications. The student found a duplication in the driver that happened across eighteen code files, 
+creating a patch to remove four hundred lines of duplicated code. The student sent the mitigations 
+to the maintainers, and their feedback pointed to fixing the warning in the code without arguing 
+about removing the duplications.
 
-Analysing the students approaches for the duplications and the maintainers feedback, 
-we saw that people without previous experience in the linux kernel development could approach 
-duplications found by the ArKanjo tool and become contributors to the kernel. We observed that 
-not all duplications are view as code with bad quality, with the maintainers analysing the trade off
-between what is the purpose of the duplicated code, the code readability, and the levels of abstraction 
-added to address the duplications.
+Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that 
+people without previous experience in Linux kernel development could approach duplications found 
+by the ArKanjo tool and become contributors to the kernel. We observed that not all duplications 
+are viewed as code with bad quality, with the maintainers analyzing the trade-off between the 
+purpose of the duplicated code, code readability, and the levels of abstraction added to 
+address the duplications.
 
-\subsection{Questionary Answers}
+\subsection{Survey Results Analysis}
 
-BLABLABLA
+We asked the teaching assistant and the students who used the ArKanjo to answer a survey with 
+some questions related to the tool experience, such as installation, documentation, facility of 
+use, new ideas to improve the tool, etc. Appendix \ref{app:survey} has the complete survey 
+with the answers.
+
+Related to the tool installation, neither the teaching assistant nor the students faced issues 
+installing on the Debian distributions. However, the teaching assistant attempted to install it 
+in the Arch Linux distribution and failed to install it on this Linux distribution. 
+Regarding the tool's documentation, the teaching assistant pointed out errors that worsened his 
+experience. The student asked for more examples and tutorials on using the tool in text or 
+video formats.
+
+On the ArKanjo tool preprocessing step, the student reported problems of disk space usage and 
+crash errors, which we knew could happen as the tool iterates and can save information about 
+every pair of functions in a codebase, as we did not implement any mechanism to control the 
+memory usage. The teaching assistant reported only being able to execute the preprocessor with 
+the text similarity method, and the diff method kept running for hours without finishing. 
+The teaching assistant's report was unexpected behavior for us, as the text similarity method 
+initially looked more complex; in contrast, the \textit{diff} method is a simple counter of equal lines 
+with an operation system default command. The teaching assistant report corroborates the 
+gensim efficiency \citep{gensim}.
+
+Related to the tool's functionalities, both the student assistant and the student reported 
+a missing mechanism to save the tool results on a separate file instead of the terminal output. 
+The tool, by default, outputted with colors, as shown in section \ref{subsec:func}, which complicates 
+redirecting the terminal output to a file.
+
+We asked for the overall impression of the tool and if it met their expectations. We also asked 
+them to give a score of one to five for installation, documentation, preprocessing, and functionalities. 
+Both the teaching assistant and the student scored at least four on each mentioned point, thought that 
+the tool fulfilled their expectations, and had an overall good experience. Finally, the teaching assistant 
+the the student reported features ideas to the tool that may be interesting, which 
+are showed below.
+
+\paragraph{List of Possible Improvements}
+
+\begin{itemize}
+
+\item Enable the possibility of having several codebases of preprocessor artifacts and selecting one of them.
+
+\item Add different methods to find duplication.
+
+\item Adding more filters to search for duplicates, such as finding duplicates with more than 90\% 
+similarity and more than 20 lines repeated.
+
+\item Add options to highlight and filter duplications on the same file or directory.
+
+\item Addition of support to save the tool output in files without manual preprocessing of the colored output.
+
+\end{itemize}
+
+With the survey feedback, we understood that users had a good experience and found value in the 
+ArKanjo tool. They also provided critical points that could be improved. The most critical 
+reported issues, such as documentation errors and a lack of support for multiple Linux distributions 
+and different operating systems, have been fixed. Less critical issues remain open points that 
+can be addressed in future work.
+
+\subsection{Students Contribution to the Tool}
+
+We listed multiple issues in the tool for the two groups that opted to contribute directly 
+to developing the ArKanjo tool. The first group opted to work on expanding the tool's support 
+for multiple operating systems. Some libraries used in the tool development only supported the 
+Ubuntu distribution, and the first group was responsible for migrating these libraries to 
+libraries with support for multiple operating systems. This issue was of our knowledge and 
+was pointed out by the teaching assistant in the survey, as he reported the problem while 
+using the tool.
+
+The second group opted to work on a webpage for the tool documentation and support for 
+light and dark modes. The colored output presented in the figures in Chapter \ref{cha:tool} was 
+hard coded at first, not considering the user terminal colors, thus giving a bad user experience. 
+The website helps facilitate people's contribution or use of the tool. We did not know how to 
+approach the light and dark modes at first and did not need to look at how to fix the issue as 
+the group searched and found a solution.
+
+The students successfully delivered the issues proposed. We found it interesting to see 
+people able to contribute to the tool on the problems that we did not even know how to approach.
+The issues help the user experience and facilitate other research using the ArKanjo tool that 
+we may explore in future works. Appendix \ref{app:contribution} references the 
+students' contributions, the issue list proposed to the students, and figures from 
+the documentation webpage.

### 2025-06-01 conteudo/tool/architecture.tex
fix grammar with grammarly

@@
 
-For a source file in the C programming language, we extract the code blocks with depth 0, as per the programming 
-language's grammar, functions are defined in a global context at depth 0. To extract the code blocks, we firstly need to 
-identify the portions of code that are comments or \textit{\#define} directives and we mark them as irrelevant, after that, 
-it is simple work of finding the curly brackets that are in depth 0, and with that we find the body of the functions.
+For a source file in the C programming language, we extract the code blocks with depth 0; 
+as per the programming language’s grammar, functions are defined in a global context at depth 0. 
+To extract the code blocks, we first need to identify the portions of code that are comments or 
+\textit{\#define} directives and mark them as irrelevant. After that, it is simple work to find 
+the curly brackets in depth 0, and with that, we find the body of the functions.
 
-To extract the header of the functions, we start at the open curly brackets and go back in the source file consireding 
-the programming language grammar  and the portions of code that have been marked as irrelevant because of comments and 
-\textit{\#define} directives.
-
-There are other elements in depth 0 that are not functions, we analyze if the code blocks is of a function at the same 
-we try to extract the header of the code block.
+To extract the header of the functions, we start at the open curly brackets and go back in the 
+source file considering the programming language grammar and the portions of code that have been 
+marked as irrelevant because of comments and \textit{\#define} directives.
+There are other elements in depth zero that are not functions. We analyze whether the code blocks 
+are functions at the same time that we extract the header of the code block.
 
@@
 
-For the Java programming language, we have decided to implemente a simpler approach that does not support 
-the programming language completely, as the only Java codebase explored in this work is the BigCloneBench \citep{bigclonebench} and 
-the approach implemented cover the needs of this codebase.
+For the Java programming language, we have decided to implement a more straightforward approach 
+that does not support the programming language completely, as the only Java codebase explored in 
+this work is the BigCloneBench \citep{bigclonebench}, and the approach implemented covers the 
+needs of this codebase.
 
-For a source file in Java, we extract the code blocks with depth 1 and some code lines before the definition of the code block. 
-Then, we check if the code blocks are functions by analyzing if the first non-empty character before the curly brackets of 
-the code block is a closing parenthesis.
+For a source file in Java, we extract the code blocks with depth one and some code lines before 
+defining them. Then, we check if the code blocks are functions by analyzing whether the first 
+non-empty character before the curly brackets of the code block is a closing parenthesis.
 
-As per Java's grammar, functions are defined inside classes and can exist at different depths, as Java accepts inner classes, 
-that means that this approach does not work with inner classes. The challenge to implement a more generic approach 
-is that other elements of depth 2 or deeper, such as for-loops, can follow the closing parenthesis pattern, necessitating 
-a better analysis to different these elements. The BigCloneBench \citep{bigclonebench} does not contains inner classes.
+As per Java’s grammar, functions are defined inside classes and can exist at different depths. 
+As Java accepts inner classes, this approach does not work with inner classes. The challenge 
+to implementing a more generic approach is that other elements of depth two or deeper, such 
+as for-loops, can follow the closing parenthesis pattern, necessitating a better analysis of 
+these elements. The BigCloneBench \citep{bigclonebench} does not contain inner classes.
 
@@
 
-We implemented two methods for code duplication detection in the tool, in which the user can choose what to use.
-The first method is based in text similarity, the second is based in simpler method based on the number of equal lines.
-The experiments and tests on this research were done using the text similarity method, the purpose of the 
-second method is to demonstrate that the tool can have multiple duplication detection methods.
+We implemented two methods for code duplication detection in the tool, which the user can 
+choose to use. The first method is based on text similarity, and the second is simpler and 
+based on the number of equal lines. The experiments and tests in this research were done 
+using the text similarity method. The purpose of the second method is to demonstrate that 
+the tool can have multiple duplication detection methods.
 
@@
 For the difference method, 
-we implemented a method considering the number of exactly equal lines between two functions.
+we implemented a method that considered the number of exactly equal lines between two functions.
 For two functions \textit{function1} and \textit{function2}, let $NL_1$ be the number of lines in \textit{function1},
@@
 
-This code duplication method is considerably simpler and more explainable then the text similarity method, 
+This code duplication method is considerably simpler and more explainable than the text similarity method, 
 as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
-the \textit{diff command} built-in in the Linux environment \citep{diffcommand}.
+the \textit{diff} command built-in in the Linux environment \citep{diffcommand}.
 
-The difference method was not tested on the evaluation of the tool and other phases of this research. 
-The method was implemented to show that tool can have multiple duplication detection methods.
+We did not test the Diff method during the tool evaluation and other phases of this research. 
+We implemented the technique to show that a tool can have multiple duplication detection methods.
 

### 2025-06-01 conteudo/tool/tool.tex
add paragraph about contrib and remove some todo
fix grammar with grammarly

@@
 
-The tool operates in two main parts, the \textbf{Preprocessor} and the \textbf{Query Responder}. 
-The \textbf{Preprocessor} is responsible to perform the heavy computations to find the code duplications 
-accross a codebase and produce artifact to consulte information related to the duplications in a structured form.
-The \textbf{Query Responder} consumes the artifacts produced by the Preprocessor to execute the tool functionalites as requested.
-This choice enable the tool to concentrate the heavy and slows parts that can be execute only once per codebase, 
+The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
+The \textbf{Preprocessor} is responsible for performing heavy computations to find the code 
+duplications across a codebase and produce artifacts to consult information related to the 
+duplications in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
+by the Preprocessor to execute the tool functionalities as requested. This choice enables the 
+tool to concentrate on the heavy and slow parts that can be executed only once per codebase, 
 thus allowing a fast performance for multiple queries related to a single codebase.
-More detailed descriptions of the Preprocessor and Query Responder can be found in
-Sections \ref{subsec:architecture} and \ref{subsec:func}.
 
@@
 \url{https://github.com/LipArcanjo/arkanjo}.
-
-\todo[inline]{Big maybes: Describe more how the functionalites are implemented, as the focus shifted for the tool as of now.}

### 2025-06-01 conteudo/methods/methods.tex
fix IIO subsystem

@@
 to students in the course Free Software Development at the University of São Paulo to mitigate 
-the code duplications found by the tool in the IIO subsystems and the AMD Display driver and 
+the code duplications found by the tool in the IIO subsystem and the AMD Display driver and 
 pass the process of sending the mitigations as PATCH.

### 2025-06-01 conteudo/results/participant.tex
fix IIO subsystem

@@
 makes this plan unviable. Thus, we opted to redirect this research to work with students 
-and refactoring duplicated function pairs found by the tool directly in the IIO and the 
+and refactoring duplicated function pairs found by the tool directly in the IIO subsystem and the 
 AMD Display driver without passing through the whole process of the systematic approach 

### 2025-06-01 conteudo/results/students.tex
fix IIO subsystem

@@
 
-For the students who worked in the IIO subsystems, we gave them duplicated function pairs 
+For the students who worked in the IIO subsystem, we gave them duplicated function pairs 
 found by the tool, and their responsibility was to think of how to mitigate the duplication 
@@
 of the students who worked in the AMD Display driver and the teaching assistant that used 
-the tool to find duplications in the IIO.
+the tool to find duplications in the IIO subsystem.
 

### 2025-06-05 conteudo/background/linux.tex
some work from 04-06

@@
 \includegraphics[scale=0.25]{diff_command}
-\caption{Example of \textit{diff} command executed}
+\caption[Example of \textit{diff} command executed]{
+Example of \textit{diff} command executed
+\footnote{
 File 1's content is shown in red, and file 2's content is shown in green. \textit{-} 
@@
 both files without any modifications.
+}
+}
 \label{fig:diff}

### 2025-06-05 conteudo/methods/methods.tex
some work from 04-06

@@
 We chose triangulation because relying solely on the literature-based method has 
-limitations and may may be unreliable \citep{bigfail, litreview}. 
+limitations and may be unreliable \citep{bigfail, litreview}. 
 As a reminder, all research approaches were done using the text similarity method 
-as the code duplication method of the tool.
+as the code duplication detection method of the tool.
 
@@
 \item Function named \textit{offset\_to\_id} in file \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c}.
-\item Function \textit{offset\_to\_id} in file \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
+\item Function named \textit{offset\_to\_id} in file \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
 \end{itemize}
@@
 (described in Section \ref{subsec:functioncommand}) to locate all duplications of the function in the codebase, 
-as it may have been duplicated more times than the two occurrences of the duplicated function pair. 
+as it may have been duplicated more times than the two occurrences contained in the duplicated function pair. 
 This strategy resulted in a collection of duplicated functions and corresponding code files.
@@
 For the first alternative, one of the teaching assistants used the ArKanjo tool in the IIO subsystem 
-and listed interesting duplicated function pairs that the groups could mitigate this specific 
-duplication pair. It is up to the students to refactor the code to mitigate the duplications and 
-pass through the process of sending patches to the IIO maintainers for review. For the second 
-alternative, we propose a similar experience to the participant observation method, where the 
-students pass through the whole process of executing the ArKanjo tool in the AMD display driver, 
-analyze the duplications found, create a patch to mitigate a duplication, and send the patch to the 
-AMD display driver’s maintainers.
+and listed duplicated function pairs found using the text similarity duplication detection method with 
+a 90\% similarity threshold. It is then up to the students to refactor the code to mitigate these 
+duplications and go through the process of sending patches to the IIO maintainers for review.
+
+For the second alternative, we propose a similar experience to the participant observation method. 
+In this approach, students go through the entire process of executing the ArKanjo tool in the 
+AMD display driver, analyzing the duplications found, creating a patch to mitigate a duplication, 
+and sending the patch to the AMD display driver’s maintainers.
 

### 2025-06-05 conteudo/results/students.tex
some work from 04-06

@@
 including the duplication function pairs that students mitigate, the students’ reports, 
-the patch students sent to the Linux kernel, and the answers to the questionnaire asked 
-of the students who worked in the AMD Display driver and the teaching assistant that used 
+the patch students sent to the Linux kernel, and the answers to the survey asked 
+of the student who worked in the AMD Display driver and the teaching assistant that used 
 the tool to find duplications in the IIO subsystem.
@@
 The other three groups do not use any refactor methods presented in the literature.
-We discuss the groups' approaches and the subsystem maintainers' opinions on the patches 
-that students sent.
 
@@
 of abstraction to remove a duplication deacrease the code readability without clear gains on 
-the codebase quality.
+the codebase quality. Figure \ref{fig:group_10_reply} shows the maintainers's feedback for group ten.
+
+\begin{figure}
+\includegraphics[scale=0.8]{group_10_reply}
+\caption[Maintainers' Feedback on Group Ten's Patch]{
+Maintainers' Feedback on Group Ten's Patch
+\footnote{Image obtained from Linux Lore, which is publicly provided at:
+\url{https://lore.kernel.org/linux-iio/20250428034222.318898-1-gabrielgeraldinosouza@gmail.com/T/\#u}}
+}
+\label{fig:group_10_reply}
+\end{figure}
 
@@
 function; it is viewed as the best choice for removing layers of abstraction and making the 
-code easier to read.
+code easier to read. Figure \ref{fig:group_7_reply} shows the maintainers's feedback for group seven.
+
+\begin{figure}
+\includegraphics[scale=0.8]{group_7_reply}
+\caption[Maintainers' Feedback on Group Seven's Patch]{
+Maintainers' Feedback on Group Seven's Patch
+\footnote{Image obtained from Linux Lore, which is publicly provided at:
+\url{https://lore.kernel.org/linux-iio/20250428132551.176788-1-bellacaselli20@gmail.com/}}
+}
+\label{fig:group_7_reply}
+\end{figure}
 

### 2025-06-05 conteudo/results/table_students.tex
some work from 04-06

--- 2025-05-25 conteudo/results/table_students.tex
@@
 6 & IIO & - & - & Parameterize method \\ \hline
-7 & IIO & 21 & 22  & None \\ \hline
-8 & IIO & - & - & None \\ \hline
+7 & IIO & 21 & 22  & Extract method \\ \hline
+8 & IIO & 20 & 16 & Extract method \\ \hline
 9 & IIO & 33 & 52 & Parameterize method \\ \hline
@@
 \end{tabular}
+Lines with \textit{-} represents groups that does not provided the Linux Lore URL of the PATCH sent to consult.
 \caption{Refactor methods used and metrics of the students's mitigations.}

### 2025-06-05 conteudo/tool/architecture.tex
I think this is fine

@@
 
-\todo[inline]{I NEED TO CHANGE HERE}
-
-The Query Responder step is the component the user executes multiple times to consult information about duplicated functions in a codebase processed by the Preprocessor.
-The Query Responder step utilizes the temporary codebase and the Code Duplication Database as inputs to answer the user's request, such as querying the total number of duplicated functions in a specific input codebase or listing duplicated functions above a certain similarity threshold.
-This component was developed to be extensible, allowing other queries to be easily added.
-Currently, the Query Responder step supports the following queries:
-
-\begin{itemize}
-	\item Number of dupliddcated functions above a similarity threshold.
-	\item Number of duplicated functions equal or above a similarity threshold.
-	\item List of duplicated functions above a similarity threshold.
-	\item List of duplicated functions equal to or above a similarity threshold.
-\end{itemize}
+The Query Responder step is the component the user executes multiple times to consult information 
+about duplicated functions in a codebase processed by the Preprocessor. 
+It utilizes the Temporary Codebase and the Code Duplication Database as inputs to answer user requests, 
+such as querying for duplicated functions above a certain similarity threshold or provide counts of such functions.
+This component was developed to be extensible, allowing other query functionalities to be easily added. 
+The specific functionalities provided to the user by the Query Responder, which include the 
+Duplication Explorer, Function Information, and Duplication Report, are described in Section \ref{subsec:func}.

### 2025-06-09 paper/paper.tex
add paper artifacts

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  %\graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage{graphicx}
  \graphicspath{{figuras/},{fig},{logos/},{img/},{images/},{imagens/}}
  % declare the path(s) where your graphic files are
  % \graphicspath{{./eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps, .png, .jpeg}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{ArKanjo: a tool for detecting function-level Code Duplication in the Linux Kernel}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Luan Icaro Pinto Arcanjo}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil 05508-220\\
Email: }
\and
\IEEEauthorblockN{Paulo Meirelles}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil 05508-220\\
Email: paulormm@ime.usp.br}
\and
\IEEEauthorblockN{David Tadokoro}
\IEEEauthorblockA{
Universidade de São Paulo\\
São Paulo, SP Brazil 05508-220\\
Email: davidbtadokoro@usp.br}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Linux kernel’s massive scale (+28 M LoC, +20 K contributors) presents unique 
maintenance challenges. Surprisingly, Code duplication remains a persistent issue 
in the kernel’s codebase, which could hinder its evolution and patching. Academic 
approaches often focus on pairwise comparison of code artifacts, not directly applied 
for comprehensive codebase analyses. Other existing free software tools explored in 
practice frequently suffer from limited functionality, such as primitive textual 
matching, prove too narrow in scope, or fail to deliver effective results on complex, 
large-scale codebases. Existing solutions generally fail to address the Linux kernel’s 
specific needs: (1) scalability to handle its size, (2) actionable results for 
developers, and (3) integration with kernel development workflows. 
This paper presents ArKanjo, a novel command-line tool for Linux kernel maintenance 
designed to detect and analyze function-level duplications. Released under the MIT 
license, ArKanjo employs a two-stage architecture consisting of a Preprocessor and 
a Query Responder that separates computationally intensive analysis from efficient 
querying for duplications within large codebases. Pivotal advantages of ArKanjo over 
existing solutions include: (1) optimization for C codebases with kernel-specific 
patterns; (2) preprocessing that enables rapid queries without redundant analysis; 
and (3) prioritization of duplicates that impact maintainability, such as copied 
buggy logic. We evaluate Arkanjo against real-world duplication cases in recent 
kernel versions, demonstrating its effectiveness in identifying problematic clones 
that generic tools often overlook. By identifying well-defined, manageable 
duplication instances, ArKanjo effectively lowers the barrier for new contributors, 
a capability evidenced by its role in guiding students to make their first code 
improvements to the kernel. ArKanjo offers immediate value to kernel maintainers 
and serves as a replicable model for clone detection in other large-scale free 
software projects.
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}

The Linux kernel is a foundational Free/Libre/Open Source Software (FLOSS) project, 
critical to a significant portion of the world's digital infrastructure. Maintaining 
the kernel is an enormous undertaking, involving more than 28 million lines of code 
and contributions from over twenty thousand developers. Within this context, code 
duplication persists as a significant challenge, a known harmful practice that 
negatively affects code readability and increases the likelihood of introducing bugs 
(Hordijk et al., 2009; Hotta et al., 2012). This issue is particularly acute in the 
kernel's device drivers, which comprise over 66\% of the source code (Schmitt, 2022). 
For instance, maintainers of the AMD Display driver have specifically highlighted code 
duplication as a significant impediment to their work.

The detection of code duplication, or code clones, has been a subject of research for 
decades (Jankowitz, 1988). The literature provides a standard taxonomy, classifying 
clones into four types based on their degree of similarity, from identical copies 
(Type-1) to semantically equivalent but textually different fragments (Type-4) (Chen et al., 2022). 
Various detection methodologies have been developed, including textual, token-based, 
tree-based, and graph-based approaches (Chen et al., 2022). These have culminated in 
state-of-the-art techniques, such as the graph-based work by Liu et al. (2023). 
However, the primary focus of such academic work remains on determining if a given 
pair of code artifacts are duplicates, rather than providing a scalable method to scan 
an entire codebase for actionable results. Conversely, existing free software tools explored 
in practice often lack this sophistication, such as relying on primitive textual matching 
that is insufficient for the complexity and scale of the kernel.

To fill this gap, this research proposes a new approach embodied in ArKanjo, a tool 
designed specifically to identify and facilitate the mitigation of function-level code 
duplications within the Linux kernel. To test the tool, we employ multiple analytical 
methods and ethnographic studies.

\section{ArKanjo Tool}

ArKanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
help developers identify code clone duplication at the function level. Released under the
MIT license, ArKanjo is available at the url: \url{https://github.com/LipArcanjo/arkanjo}

\subsection{Architecture}

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{diagrama_mestrado}
\caption{Architecture diagram demonstrating the relationship between the tool components}
\label{fig:diagram}
\end{figure}

The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} is responsible for performing heavy computations to find the code 
duplications across a codebase and produce artifacts to consult information related to the 
duplications in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
by the Preprocessor to execute the tool functionalities as requested. This choice enables the 
tool to concentrate on the heavy and slow parts that can be executed only once per codebase, 
thus allowing a fast performance for multiple queries related to a single codebase.

The Preprocessor contains two main components, the Function Breaker and the Code Duplication Finder. 
The flow of how the Preprocessor works is as follow: Function Parser receives the codebase the user is interested in,
extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files. 
The Code Duplication Finder iterates over every pair of files in the Temporary Codebase, checks if they are code clones, and, 
if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple 
\textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates 
of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized in the Code Duplication Finder.
Figure \ref{fig:diagram} illustrates the tool architecture.

The Query Responder consumes the Temporary Codebase and the Code Duplication Database to extract duplicated 
functions-related information per user request. If the user executes the Query Responder without executing the 
Preprocessor, the Query Responder calls the Preprocessor to create the required artifacts.

The Function Parser receives the input codebase and transforms it into the temporary
codebase. The Function Parser iterates through every source code file from the programming 
language it supports and uses a specific programming language function extractor
to extract every function in the file. For each function extracted, two new source code files
and a metadata file are created in the temporary codebase, represented by the pair <file
name, function name>, the source code file of the function, and the proper function.
The first new source code file contains the function’s body from the function it represents,
while the other new source code file contains the function’s signature. The metadata file
contains additional relevant information about the function, such as the function name,
the line where the function signature starts in the source code file, and the line where the
function’s body ends in the source code file. The programming languages supported at
the moment are C and Java, although Java has limited support.

The Code Duplication Finder iterates through every pair of source code files in the
temporary codebase, representing functions from the input codebase. For each pair of
files, we execute a code duplication detection method that computes a metric measuring
how similar the pair of files is, which we refer to as similarity. If the similarity is greater
than or equal to the minimum similarity threshold (a parameter provided by the user
during the Preprocessor), we store this pair of functions along with its similarity in the
Code Duplication Database. 
We implemented two methods for code duplication detection in the tool, which the user
can choose to use. The first method is based on text similarity, and the second is simpler
and based on the number of equal lines. The experiments and tests in this research were
done using the text similarity method.

\subsubsection{Code Duplication Methods Used}

For the text similarity method, we treat the source code files as text and apply the
TF-IDF vector embedding method implemented by the Gensim library (ADD REF),
then compute cosine similarity as the similarity metric. This method was chosen for 
its claimed performance, programming language independence, and the fact that it does 
not require compilable code, which is expected in the
temporary codebase, as it does not contain complete code artifacts.
(SHOULD I ADD THE FOLLOWING? THERE IS LITTLE SPACE:) 
Plates implemented this code duplication detection method and released it under
the MIT license(ADD REF). We adapted his implementation to fit our tool’s expected
input and output formats.

For the difference method, 
we implemented a method that considered the number of exactly equal lines between two functions.
For two functions \textit{function1} and \textit{function2}, we compute the similarity as the 
ratio of duplicated lines between the two functions by the total number of lines in both functions.
This code duplication method is considerably simpler and more explainable than the text similarity method, 
as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
the \textit{diff} command built-in in the Linux environment (ADD REF).

\subsection{Functionalities}

We propose three main functionalities for the user in the tool, called the Duplication
Explorer, Function Information, and Duplication Report. Each functionality accepts
specific parameters to perform its operations.

The Duplication Explorer is the primary functionality of our tool, designed to present
the user with pairs of duplicated functions identified by the tool. We implement optional
filters to facilitate more complex queries. Figure \ref{fig:explorer_ex} demonstrates an
example of this functionality in use.

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{explorer_example}
\caption{Example of the \textit{Duplication Explorer} functionality of the ArKanjo tool.}
\label{fig:explorer_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{function_example}
\caption{Example of the \textit{Function Information} functionality of the ArKanjo tool.}
\label{fig:function_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3in]{relatory_example}
\caption{Example of the \textit{Duplication Report} functionality of the ArKanjo tool.}
\label{fig:relatory_ex}
\end{figure}

The Function Information functionality provides detailed information about a specific function. 
The functionality receives a target function from the user
and returns information such as the relative path, function name, and line numbers where
the function is defined. Additionally, it provides similar information for every function
that duplicates the given function. Figure \ref{fig:function_ex} demonstrates an
example of this functionality in use.

The Duplication Report functionality provides an overview of
duplicated code within the input codebase. This functionality calculates the number of
duplicated lines per folder in the codebase and presents the information to the user in
a readable format. Figure \ref{fig:relatory_ex} demonstrates an example of this functionality in use.

\section{Methods}

To validate the ArKanjo tool capabilities to find code duplications within the 
text similarity detection method used, we applied two independent
methods. The first method evaluates the proposed tool using an approach from the literature,
comparing it against the BigCloneBench dataset (Svajlenko et al., 2014). The second
method is an empirical analysis of a subset of functions within the AMD Display driver
that our tool identified as duplicates.

To validate the tool against the BigCloneBench dataset (Svajlenko et al., 2014), we followed
the methodology presentedy by by (Liu et al., 2023), we sampled 20,000 pairs from each clone
type, adding 20,000 non-duplicate pairs as negative samples. We applied the same sampling
approach to our tool to ensure a fair evaluation.
Unlike state-of-art methods, our tool does not distinguish between clone types and
identifies duplications at the function level rather than the file level, which is typical in
state-of-the-art tools. Therefore, we adapted the metric calculation method for evaluation
purposes. Specifically, we considered every pair of functions with a similarity metric
equal to or greater than a threshold 𝑋 as duplicates and marked the corresponding file
pairs. A correctly identified duplication pair counts as accurate for its clone type, while
an incorrectly inferred pair is considered incorrect across all clone types.
To understand the impact of varying the similarity threshold, we evaluated our tool using
different threshold values: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%. We then analyzed
and discussed the results.

For the empirical analysis, we randomly sampled function pairs identified by the tool as
duplicates. For each similarity threshold X (30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%) 
we randomly selected ten function pairs with a similarity close to X, allowing for a 1\%
deviation.

To assess if the duplications found by the ArKanjo tool were actionable, the research conducted 
a multi-part ethnographic study. In this study, the author and 12 student groups from the
Free Software Development course at University of São Paulo acted as new contributors 
to the Linux kernel. 
They used the tool to identify duplications within the Industrial I/O (IIO) subsytem and AMD Display 
driver subsystems, refactored the code to address the issues, and submitted their proposed 
fixes as patches to the community maintainers. Throughout this process, the students documented 
their refactoring approaches and their experiences interacting with the kernel community.

\section{Results}

\subsection{Tool Evaluation}

\begin{table}[!t]

\centering

\renewcommand{\arraystretch}{1.3}
\begin{tabular}{ | m{13mm} | c | c | c | c | m{6mm} | m{6mm} | }

\hline

\textbf{Similarity Threshould} & \textbf{T1} & \textbf{T2} & ST3 & MT3
& WT3/ T4 & \textbf{False} \\ \hline

100\% & 100\% & 5\% & 6\% & 0\% & 0\% & 100\% \\ \hline
90\% & 100\% & 85\% & 26\% & 0\% & 0\% & 100\% \\ \hline
80\% & 100\% & 87\% & 37\% & 1\% & 0\% & 100\% \\ \hline
70\% & 100\% & 88\% & 44\% & 2\% & 0\% & 100\% \\ \hline
60\% & 100\% & 89\% & 49\% & 4\% & 0\% & 100\% \\ \hline
50\% & 100\% & 90\% & 64\% & 6\% & 0\% & 100\% \\ \hline
40\% & 100\% & 90\% & 68\% & 9\% & 0\% & 100\% \\ \hline
30\% & 100\% & 98\% & 74\% & 13\% & 0\% & 100\% \\ \hline

\hline
\end{tabular}
\caption{Recall of the ArKanjo tool in the BigCloneBench dataset varying minimum similarity threshould.}
\label{tab:bigclone}
\end{table}

\begin{table}[!t]

\centering

\begin{tabular}{ | m{15mm} | c | c | c | c | m{6mm} | m{10mm} | }

\hline

\textbf{Similarity range} & \textbf{T1} & \textbf{T2} & T3 & T4
& \textbf{False} & \textbf{Success Percentage} \\ \hline
99\% - 100\% & 9 & 0 & 0 & 1 & 0 & \textbf{100\%} \\ \hline
89\% - 91\% & 0 & 8 & 0 & 1 & 1 & \textbf{90\%} \\ \hline
79\% - 81\% & 0 & 3 & 2 & 3 & 2 & \textbf{80\%} \\ \hline
69\% - 71\% & 0 & 3 & 1 & 1 & 5 & \textbf{50\%} \\ \hline
59\% - 61\% & 0 & 0 & 0 & 2 & 8 & \textbf{20\%} \\ \hline
49\% - 51\% & 0 & 0 & 0 & 0 & 10 & \textbf{0\%} \\ \hline
39\% - 41\% & 0 & 0 & 0 & 0 & 10 & \textbf{0\%} \\ \hline
29\% - 31\% & 0 & 0 & 0 & 0 & 10 & \textbf{0\%} \\ \hline

\hline

\end{tabular}
\caption{Results of the ArKanjo tool in the empirical analyses in the AMD Display driver}
\label{tab:emp}

\end{table}

Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCLoneBench dataset
and table \ref{tab:emp} shows the results obtained in the empricial analysis.
The BigCloneBench results and our empirical methods indicate that our tool performs
well in detecting Type-1 and Type-2 code clone duplications and can thus reveal propagated 
issues like copied buggy logic, but it struggles with more
complex types. A similarity threshold between 80\% and 100\% yields favorable results
in both methods.

However, there is a discrepancy in the detection of non-duplication pairs between
the two methods. In the BigCloneBench evaluation, we found no false positives (negative
samples inferred as duplications), while the empirical analysis revealed a higher percentage
of false positives. We propose two potential explanations for this discrepancy.

The first reason may be the limitations and known issues with BigCloneBench, as
highlighted by Krinke and Ragkhitwetsagul (Krinke and Ragkhitwetsagul, 2022). The
second reason could relate to the nature of the AMD Display driver, where each code artifact
shares semantic meaning. In contrast, BigCloneBench comprises self-contained code
artifacts without shared semantics. Given that our tool relies on a text-based code clone
detection approach, it may naturally perform worse on the AMD Display driver than on the
BigCloneBench dataset.

\subsection{Ethnographic Study}

Using the ArKanjo tool, the author and 12 student groups identified and refactored duplicated 
code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work 
as patches to the community maintainers. This effort resulted in 13 total patch submissions: 
one from the author to the AMD Display driver and 12 from the student groups, with 11 targeting 
the IIO subsystem and one for the AMD Display driver. The author's patch, which removed over 
500 lines of duplicated code, was successfully accepted and merged into the kernel after a 
lengthy review process.

The student efforts showed that newcomers could use the tool to make effective contributions. 
Of the 12 student patches, three were accepted, three were refused, and six required follow-up 
fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 
Parameterize Method (ADD REF) and two used the Extract Method (ADD REF), which are simple refactoring methods. 
This result corroborates that ArKanjo finds actionable duplications and that newcomers can 
successfully contribute patches to the kernel.

While this research included a successfully merged patch by the author and students,
a notable portion of student efforts encountered
significant hurdles. These difficulties often stemmed from maintainer feedback where
proposed changes, though reducing duplication, were rejected or required substantial
rework due to concerns about code readability, added abstraction, or the specific context
of the code. Technical complexities, such as C macros in configuration files, and lengthy
patch review times also contributed to the challenges. 
This highlights that while ArKanjo effectively identifies duplications, a successful 
contribution involves a trade-off between code deduplication and other factors like the 
specific code's context and maintainer priorities.

\section{Conclusion}

In conclusion, this paper presented ArKanjo, a novel tool designed to effectively detect 
and facilitate the refactoring of function-level code duplication within the Linux kernel. 
Evaluations demonstrated that ArKanjo is highly effective at identifying Type-1 and Type-2 
clones, which are often indicative of propagated issues like copied buggy logic. The practical 
value of the tool was confirmed through an ethnographic study where the author and student 
groups used ArKanjo to identify actionable duplications. This effort led to 13 patch 
submissions, including one from the author that removed over 500 lines of code and was merged 
into the kernel, alongside three successful student contributions. While some refactoring 
efforts were declined due to maintainer priorities, ArKanjo successfully lowers the barrier 
for new contributors to make meaningful improvements. The tool provides immediate value for 
kernel maintenance and serves as a model for clone detection in other large-scale projects.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

% conference papers do not normally have an appendix

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bibliografia}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

 \bibitem{IEEEhowto:kopka}
 H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% that's all folks
\end{document}

### 2025-06-11 paper/paper.tex
add references

@@
 negatively affects code readability and increases the likelihood of introducing bugs 
-(Hordijk et al., 2009; Hotta et al., 2012). This issue is particularly acute in the 
-kernel's device drivers, which comprise over 66\% of the source code (Schmitt, 2022). 
+\cite{harmone,harmtwo}. This issue is particularly acute in the 
+kernel's device drivers, which comprise over 66\% of the source code \cite{marcelo}. 
 For instance, maintainers of the AMD Display driver have specifically highlighted code 
@@
 clones into four types based on their degree of similarity, from identical copies 
-(Type-1) to semantically equivalent but textually different fragments (Type-4) (Chen et al., 2022). 
+(Type-1) to semantically equivalent but textually different fragments (Type-4) \cite{litreview}. 
 Various detection methodologies have been developed, including textual, token-based, 
-tree-based, and graph-based approaches (Chen et al., 2022). These have culminated in 
-state-of-the-art techniques, such as the graph-based work by Liu et al. (2023). 
+tree-based, and graph-based approaches \cite{litreview}. These have culminated in 
+state-of-the-art techniques, such as the graph-based work by Liu et al \cite{tailor}. 
 However, the primary focus of such academic work remains on determining if a given 
@@
 For the text similarity method, we treat the source code files as text and apply the
-TF-IDF vector embedding method implemented by the Gensim library (ADD REF),
+TF-IDF vector embedding method implemented by the Gensim library \cite{gensim},
 then compute cosine similarity as the similarity metric. This method was chosen for 
@@
 temporary codebase, as it does not contain complete code artifacts.
-(SHOULD I ADD THE FOLLOWING? THERE IS LITTLE SPACE:) 
-Plates implemented this code duplication detection method and released it under
-the MIT license(ADD REF). We adapted his implementation to fit our tool’s expected
-input and output formats.
 
@@
 as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
-the \textit{diff} command built-in in the Linux environment (ADD REF).
+the \textit{diff} command built-in in the Linux environment.
 
@@
 
-To validate the tool against the BigCloneBench dataset (Svajlenko et al., 2014), we followed
-the methodology presentedy by by (Liu et al., 2023), we sampled 20,000 pairs from each clone
+To validate the tool against the BigCloneBench dataset \cite{bigclonebench}, we followed
+the methodology presentedy by Liu et al. \cite{tailor}, we sampled 20,000 pairs from each clone
 type, adding 20,000 non-duplicate pairs as negative samples. We applied the same sampling
@@
 purposes. Specifically, we considered every pair of functions with a similarity metric
-equal to or greater than a threshold 𝑋 as duplicates and marked the corresponding file
+equal to or greater than a threshold X as duplicates and marked the corresponding file
 pairs. A correctly identified duplication pair counts as accurate for its clone type, while
@@
 The first reason may be the limitations and known issues with BigCloneBench, as
-highlighted by Krinke and Ragkhitwetsagul (Krinke and Ragkhitwetsagul, 2022). The
+highlighted by Krinke and Ragkhitwetsagul \cite{bigfail}. The
 second reason could relate to the nature of the AMD Display driver, where each code artifact
@@
 fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 
-Parameterize Method (ADD REF) and two used the Extract Method (ADD REF), which are simple refactoring methods. 
+Parameterize Method and two used the Extract Method \cite{refactorbook}, which are simple refactoring methods. 
 This result corroborates that ArKanjo finds actionable duplications and that newcomers can 
@@
 % use section* for acknowledgment
-\section*{Acknowledgment}
-
-The authors would like to thank...
 
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtran}
 % argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bibliografia}
+\bibliography{IEEEabrv,referencias}
 %
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-
- \bibitem{IEEEhowto:kopka}
- H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
+% \begin{thebibliography}{1}
+%  
+%  \bibitem{IEEEhowto:kopka}
+%  H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
+%    0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+%  
+% \end{thebibliography}
 

### 2025-06-12 conteudo/methods/methods.tex
add more numerical results from students

@@
 Firstly, we present the participant observation research method explored by the author. 
-Then, we present the methodology of the experiment done with the students.
+Then, we present the methodology of the non-participant observation analyze of patches 
+sent by the students.
 
@@
 subsystem's maintainers' opinions about the duplications and their respective mitigations. 
-To accomplish this objective, we initially proposed a participant-observation experiment, 
+To accomplish this objective, we proposed a participant-observation experiment, 
 where we, people who have never contributed to the Linux kernel, act as contributors to 
@@
 
-As a first step of this experiment, we executed the tool in the AMD Display driver codebase. 
+We executed the tool in the AMD Display driver codebase. 
 We manually analyzed the biggest duplications regarding the number of lines and raised them 
@@
 
-\subsection{Study with University Students}
+\subsection{Non-participant Observation Study with University Students}
 
@@
 groups worked in direct contributions, which was not a very expressive number compared with the 
-initial experiment.
+initial stage of the course.

### 2025-06-12 conteudo/results/participant.tex
add more numerical results from students

@@
 
-For our initial participant observation, we attempted to mitigate two code 
+For our participant observation, we attempted to mitigate two code 
 duplications found by the ArKanjo tool in the AMD Display driver by applying 

### 2025-06-12 conteudo/results/results.tex
add more numerical results from students

@@
 observation methodology, where we mitigate a code duplication on the AMD Display Driver 
-and send it as a patch. Finally, we show and discuss our discoveries on the experiment 
-done with students at the University of São Paulo, where we propose to students to mitigate 
-duplications found by the ArKanjo tool at IIO subsystem and the AMD Display Driver, and 
-document their founds, opinions, and experiences.
+and send it as a patch. Finally, we show and discuss our discoveries of the non-participant observation
+of the students approaches to mitigate duplications and patches's discussions.
 

### 2025-06-12 conteudo/results/students.tex
add more numerical results from students

@@
 
-As presented in Chapter \ref{cha:method}, we propose that students mitigate duplications 
-found by the ArKanjo tool, send patches to the Linux subsystem maintainers, and document 
-their experiences and learnings.
+As presented in Chapter \ref{cha:method}. It was proposed to students, as one of the activities 
+of the Free Software course, that they make the first-time contribution to the Linux kernel to 
+mitigate duplication found by the ArKanjo tool while sending patches to the Linux subsystem 
+maintainers and documenting their experiences and learnings.
 
-For the students who worked in the IIO subsystem, we gave them duplicated function pairs 
+For the students who worked in the IIO subsystem, the teaching assistant gave them duplicated 
+function pairs 
 found by the tool, and their responsibility was to think of how to mitigate the duplication 
@@
 
-Table \ref{tab:stu} summarizes students' groups' mitigation approaches regarding lines 
-changed and refactor methods used. Seven groups applied the parameterized refactor method, 
+Table \ref{tab:stu} summarizes students' groups' mitigation approaches. Seven groups applied 
+the parameterized refactor method, 
 and two applied the extractor method to their approaches to mitigate the duplications. 
 The other three groups do not use any refactor methods presented in the literature.  
+Five groups got their patches accepted and merged in Linux; four groups got their patches rejected, 
+and one group is still working on the patch submission, being in the fifth version of the patch at 
+the time this document was written. Two groups decided not to continue sending the patches after 
+they made initial versions.
 

### 2025-06-12 conteudo/results/table_students.tex
add more numerical results from students

--- 2025-05-25 conteudo/results/table_students.tex
@@
 \begin{table}
-\begin{tabular}{ | c | c | c | c | m{6em} | }
+\begin{tabular}{ | c | c | c | m{6em} | }
 
@@
 
-\textbf{Group} & \textbf{Subsystem} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
+\textbf{Group} & \textbf{Subsystem} & \textbf{Patch Status} & \textbf{Refactoring methods used} 
 \\ \hline
 
-1 & IIO & 7 & 14 & None \\ \hline
-2 & IIO & 23 & 24 & Parameterize method \\ \hline
-3 & IIO & 12 & 30 & Parameterize method \\ \hline
-4 & IIO & 2 & 7 & None \\ \hline
-5 & IIO & 17 & 70 & Parameterize method \\ \hline
-6 & IIO & - & - & Parameterize method \\ \hline
-7 & IIO & 21 & 22  & None \\ \hline
-8 & IIO & - & - & None \\ \hline
-9 & IIO & 33 & 52 & Parameterize method \\ \hline
-10 & IIO & 32 & 58 & Parameterize method \\ \hline
-11 & IIO & - & - & Parameterize method \\ \hline
-12 & AMD & 89 & 489 & None \\ \hline
+1 & IIO & Rejected & None \\ \hline
+2 & IIO & Rejected & Parameterize method \\ \hline
+3 & IIO & Accepted & Parameterize method \\ \hline
+4 & IIO & Accepted & None \\ \hline
+5 & IIO & Accepted  & Parameterize method \\ \hline
+6 & IIO & Students stopped working on it & Parameterize method \\ \hline
+7 & IIO & Accepted  & Extract method \\ \hline
+8 & IIO & Rejected & Extract method \\ \hline
+9 & IIO & Work in progress & Parameterize method \\ \hline
+10 & IIO & Rejected & Parameterize method \\ \hline
+11 & IIO & Students stopped working on it & Parameterize method \\ \hline
+12 & AMD & Accepted & None \\ \hline
 

### 2025-06-12 conteudo/methods/methods.tex
change to more on the tone the students

@@
 
-Based on our initial participant observation, we decided to expand the experiment by proposing 
-to students how to mitigate duplication found by the tool and send patches to the kernel community.
+A core dynamic of the Free Software Development course at the University of São Paulo is to 
+have students make a practical contribution to a free software project. In the first semester of 
+2025, the professor teaching the course, who is also the advisor of this work, identified an 
+opportunity to leverage the ArKanjo tool within this existing pedagogical structure. As a 
+contribution option, the professor proposed that students mitigate code duplications found in the 
+Linux kernel. This initiative was driven by the professor as a practical course activity, and 
+the author of this thesis was not involved in using the tool to find duplications nor in developing 
+the approaches to mitigate them. The tasks were carried out entirely by the course staff 
+and the students.
 
-In the Free Software Development course at the University of São Paulo, taught by the advisor of 
-this work in the first semester of 2025, we proposed to the students the possibility of mitigating 
-duplications in the IIO subsystem and the AMD display driver found with the ArKanjo tool. 
-The course had 26 students, of whom X were undergraduate students and Y were graduate students.
+The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of 
+one to three. To assist students in this task, the professor and teaching assistants prepare multiple 
+alternatives for contributing to free software projects, including simpler contributions such as 
+updating documentation or fixing code style issues. In this context, two new options related to the 
+ArKanjo tool were proposed alongside the others.
 
-In the course, the students are asked to form groups of one, two, or three students and contribute 
-to free software together. To assist the students in this task, the professor and teaching assistants 
-prepare multiple alternatives for contributing to free software projects, including simple contributions 
-such as updating documentation or code style fixes. In this context, we have proposed two options 
-for the students related to the tool.
+For the first ArKanjo-related alternative, a teaching assistant ran the tool on the IIO subsystem, 
+creating a list of duplicated function pairs found with a 90\% similarity threshold. The students' 
+task was to refactor the code to mitigate these specific duplications and then navigate the process 
+of submitting their patches to the IIO maintainers for review.
 
-For the first alternative, one of the teaching assistants used the ArKanjo tool in the IIO subsystem 
-and listed duplicated function pairs found using the text similarity duplication detection method with 
-a 90\% similarity threshold. It is then up to the students to refactor the code to mitigate these 
-duplications and go through the process of sending patches to the IIO maintainers for review.
+For the second alternative, students were offered an experience similar to our initial participant 
+observation. In this model, the students were responsible for the entire workflow: executing the 
+ArKanjo tool on the AMD display driver, independently analyzing the results to identify a duplication 
+to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.
 
-For the second alternative, we propose a similar experience to the participant observation method. 
-In this approach, students go through the entire process of executing the ArKanjo tool in the 
-AMD display driver, analyzing the duplications found, creating a patch to mitigate a duplication, 
-and sending the patch to the AMD display driver’s maintainers.
+Of the students who chose to work on the ArKanjo-related tasks, 25 (20 undergraduate and 5 graduate) 
+formed 11 groups to pursue the first alternative. One graduate student opted for the second alternative, 
+forming a single group. Notably, none of these students had prior experience contributing to the 
+Linux kernel, making this their first attempt to submit a patch as newcomers.
 
-From the students enrolled in the course, 25 students opted for the first alternative, and one 
-student opted for the second alternative, forming 11 groups working on the first alternative 
-and 1 group working on the second alternative. It is worth highlighting that none of the students 
-who chose to mitigate duplications had contributed to the Linux kernel before, meaning that the 
-patch to mitigate the duplications was the first tentative to contribute to the kernel as newcomers.
-
-The groups have been requesting that we document the experience and approaches to refactoring the 
+The groups were requested to document their experience and approaches to refactoring the 
 duplications and the experience of submitting a patch to the Linux kernel in blogs. We analyzed 
@@
 issue, and to our knowledge, they pointed to us in the survey feedback. Three students forming two 
-groups worked in direct contributions, which was not a very expressive number compared with the 
-initial stage of the course.
+groups worked in direct contributions.

### 2025-06-13 conteudo/methods/methods.tex
add me

@@
 creating a list of duplicated function pairs found with a 90\% similarity threshold. The students' 
-task was to refactor the code to mitigate these specific duplications and then navigate the process 
+task was to refactor the code to mitigate these specific duplications and then pass through the process 
 of submitting their patches to the IIO maintainers for review.

### 2025-06-13 conteudo/resumoabstract.tex
change abstract

@@
 pesquisas tanto na literatura formal quanto na cinza, as soluções existentes não 
-conseguem resolver os desafios específicos do kernel Linux. Para abordar esse tema, 
-esta pesquisa foca em dois objetivos principais. 
-Primeiro, propõe uma nova ferramenta projetada para identificar duplicações de código dentro 
-do kernel Linux, com sua aplicação prática e eficácia avaliadas em subsistemas chave: o AMD 
-Display Driver e o subsistema Industrial I/O.
-Segundo, analisa as 
-duplicações identificadas para extrair padrões e métodos de refatoração, oferecendo 
-orientações para futuros contribuidores.
+conseguem resolver os desafios específicos do kernel Linux. 
+Esta pesquisa apresenta o ArKanjo, uma ferramenta de linha de comando para a manutenção 
+do kernel Linux, projetada para detectar e analisar duplicações em nível de função. 
+Lançado sob a licença MIT, o ArKanjo emprega uma arquitetura de dois estágios que consiste 
+em um Preprocessor e um Query Responder, que separa a análise computacionalmente intensiva 
+da consulta eficiente por duplicações em grandes bases de código. Em segundo lugar, ele 
+analisa as duplicações identificadas para extrair padrões e métodos de refatoração, oferecendo 
+orientação a futuros contribuidores. Avaliamos o Arkanjo em casos de duplicação do mundo 
+real em versões recentes do kernel, demonstrando sua eficácia na identificação de clones 
+problemáticos que ferramentas genéricas frequentemente ignoram. Ao identificar instâncias 
+de duplicação bem definidas e gerenciáveis, o ArKanjo reduz efetivamente a barreira para 
+novos contribuidores, uma capacidade evidenciada por seu papel em orientar estudantes a 
+fazerem suas primeiras melhorias de código no kernel.
 
@@
 bug fixes, or handling of security vulnerabilities can impact millions. Maintaining the kernel 
-is an enormous task, involving more than 19 million lines of code and contributions from over 
-thirteen thousand developers. Due to the scale of the project, not all contributions adhere to best 
+is an enormous task, involving more than 28 million lines of code and contributions from over 
+twenty thousand developers. Due to the scale of the project, not all contributions adhere to best 
 practices, resulting in poor-quality code artifacts that may complicate maintenance and feature 
@@
 duplications. Despite searches in both formal and grey literature, existing solutions fall 
-short of addressing the specific challenges of the Linux kernel. To address this topic, this 
-research focuses on two key goals. 
-First, it proposes a new tool designed for identifying code duplications within the Linux kernel, 
-with its practical application and effectiveness evaluated in key subsystems: the AMD Display Driver 
-and the Industrial I/O subsystem.
+short of addressing the specific challenges of the Linux kernel.
+This research presents ArKanjo  , a command-line tool for Linux kernel maintenance designed to detect and
+analyze function-level duplications. Released under the MIT
+license, ArKanjo employs a two-stage architecture consisting
+of a Preprocessor and a Query Responder that separates
+computationally intensive analysis from efficient querying for duplications within large codebases.
 Second, it analyzes the identified duplications to extract patterns 
-and refactoring methods, offering guidance to future contributors.
+and refactoring methods, offering guidance to future contributors. We evaluate Arkanjo 
+against realworld duplication cases in recent kernel versions, demonstrating
+its effectiveness in identifying problematic clones that generic
+tools often overlook. By identifying well-defined, manageable
+duplication instances, ArKanjo effectively lowers the barrier
+for new contributors, a capability evidenced by its role in guiding students to make their first 
+code improvements to the kernel.
 

### 2025-06-13 conteudo/introduction/intro.tex
update research image
update intro to fit more with text and keep the style of intro of paper

@@
 
-The Linux kernel is a Free/Libre/Open-Source project providing value to
-billions of people daily, representing 96.3\% of the top one million web
-servers and 85\% of smartphones worldwide. The magnitude that the project has
-reached over the years makes the addition of features, as well as handling bugs
-and security vulnerabilities, potentially affect a significant number of
-people, demonstrating the importance of maintaining the kernel
-\citep{linuxdata}.
-
-Maintaining the Linux kernel is a complex task, as the project contains more
-than 28 million lines of code and has had over twenty thousand contributors
-throughout its lifespan
+The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, 
+powering 96.3\% of the top one million web servers and 85\% of smartphones worldwide. 
+Maintaining the kernel is an enormous task, involving more than 28 million lines of 
+code and contributions from over twenty thousand developers
 \footnote{Data directly extracted from Linux github repository \citep{linuxrepo}.
@@
 Number of contributors counted as number of unique full name listed as author of a patch.}.
-Given the magnitude of the Linux
-kernel project, it is unrealistic to expect that all contributors have
-consistently followed the best programming practices and principles in every
-contribution made to the project. Consequently, it is expected that a
-significant amount of low-quality code artifacts will be found in the source
-code, which could hinder or compromise the maintenance of the project and the
-addition of new features.
+Due to the scale of the project, not all contributions adhere to best practices, resulting 
+in poor-quality code artifacts that may complicate maintenance and future feature development.
 
-As stated by \citet{driverdef}, device drivers are a ``piece of software whose
-aim is to control and manage a particular hardware device, hence the name
-device driver.'' Device drivers are a major part of the Linux kernel,
-representing 66\% of the source code, and are responsible for managing
-essential elements such as keyboards, mice, and GPUs \citep{marcelo}. When the
-kernel opts to support new hardware, it necessitates alterations to existing
-device drivers or the creation of new ones, making maintaining device drivers a
-critical duty of the Linux community.
+One such artifact is code duplication, which is the focus of this research. The motivation for 
+this work began with a practical problem rooted in device drivers, which are a major part 
+of the kernel, representing 66\% of the source code\citep{marcelo}. We interacted with the 
+maintainers of the AMD Display driver, an essential component for the 19\% of personal GPUs 
+manufactured by AMD in 2023\citep{gpumarket}. They shared their challenges with a significant 
+amount of duplicated code that hampers the driver's maintenance. In searching for a solution, 
+we found that existing tools are not well-suited for identifying duplications in large-scale 
+codebases like the Linux kernel, nor do they offer guidance on how to resolve the duplications 
+they find. Despite searching both formal and grey literature, existing solutions fell short of 
+addressing these specific challenges. The list of related tools examined in this exploratory 
+research is available in Appendix \ref{app:gray}.
 
-The AMD Display driver in the Linux kernel is responsible for enabling AMD GPUs
-to operate correctly in desktop Linux environments, which is an essential task,
-given that AMD GPUs represented 19\% of the personal GPU market in 2023
-\citep{gpumarket}. When we interacted with the maintainers of the AMD Display
-driver, they shared with us their challenges regarding one of the
-characteristics of low-quality code artifacts: duplicated code, which hampers
-the maintenance of the driver.
+This research presents ArKanjo, a command-line tool designed to detect and analyze function-level 
+duplications in the Linux kernel. Released under the MIT license, ArKanjo uses a two-stage 
+architecture with a Preprocessor and a Query Responder. This design separates the heavy 
+computational analysis from the fast querying of duplication results, making it effective for 
+large codebases. We validated the tool by comparing its results with the BigCloneBench dataset
+\citep{bigclonebench} and by conducting an empirical analysis on a randomly selected set of 
+files from the AMD Display driver.
 
-In searching for solutions to the AMD Display driver problem related to
-duplicated code artifacts, we found tools capable of determining, given two
-code artifacts, whether they are duplications of each other. However, the
-literature on code duplication mainly focuses on optimizing the accuracy
-of these tools for this specific task. These solutions do not address our
-problem, as we cannot provide a complete codebase and have the tools return the
-code duplications within it. Moreover, they do not guide us on how to mitigate
-the code duplications after detection, necessitating the exploration of
-alternative solutions.
-
-Given the unsuccessful results of our search in the literature, we
-turned to explore related tools not present in literature for solutions, but were also met 
-with limited success.
-The best solutions we found were primitive, merely
-identifying every pair of code files that are duplications. The code files in
-the AMD Display driver containing duplicated code typically have specific code
-unique to each file. Therefore, a solution that identifies duplication in a
-deeper context is necessary to find accurate duplications within the Linux
-kernel context. 
-The list of related tools examined in this exploratory research 
-is available in Appendix \ref{app:gray}.
+Beyond detection, we also investigated if the duplications found by ArKanjo could be mitigated.
+This was achieved through an ethnographic approach, which included a participant observation 
+study where we attempted to mitigate duplications, and a non-participant observation study of 
+university students tasked with fixing duplications found by the tool. These studies showed that 
+ArKanjo can effectively lower the barrier for new contributors, as evidenced by its role in 
+guiding students to make their first code improvements to the kernel.
 
@@
 
-Our studies and observations reveal that existing tools cannot fully address the code duplication 
-problem described in the previous section in the formal or grey literature. To fill this gap,
-we propose an approach to identify and mitigate code duplications, specifically within the
-Linux kernel subsystems, focusing our study on the AMD Display driver and the Industrial 
-IO subsystem. Our approach includes the development of a tool capable of detecting duplicated 
-functions and establishing strategies for addressing these duplications. We tested the approach 
-in the context of two kernel components, the AMD Display driver and the Industrial IO subsystem.
+To address the code duplication problem within the Linux kernel, this research employs a 
+multi-method approach designed to bridge the gap between detection and practical mitigation. 
+Our approach, illustrated in Figure \ref{fig:reDesign}, is structured into two interconnected phases that progress 
+from tool development to an in-depth study of its application within the kernel’s development community. 
+The overall objective is to produce not only a functional tool but also actionable insights into 
+the patterns and practices surrounding code duplication and its resolution.
 
 \begin{figure}[h]
-\includegraphics[scale=0.46]{research_design}
+\includegraphics[scale=0.9]{research_design}
 \caption{Diagram of the research design.}
@@
 
-This research aims to identify patterns of software engineering techniques, 
-such as refactoring methods commonly used to address code duplications. 
-These patterns will serve as a guide for future contributors in mitigating 
-code duplication issues within this context. To investigate our approach, 
-we are developing a tool to identify duplicate code. We then refactor portions 
-of this code and submit patches to maintainers, providing evidence of the tool’s 
-accuracy in detecting duplicate functions. Our research design follows a 
-multi-method approach, structured into two phases, each employing different research methods.
+The initial phase of our research focused on the development and validation of ArKanjo, a 
+command-line tool capable of identifying code duplication at the function level.  The 
+tool's accuracy was rigorously validated by evaluating its performance against the 
+standard BigCloneBench dataset \citep{bigclonebench} and by conducting an empirical analysis 
+of its results within the AMD Display driver.  This foundational work ensured we had a 
+reliable instrument to proceed with the practical investigation of code duplication 
+in our target subsystems.
 
-In \textbf{Phase I}, we propose, develop, and validate a tool capable of
-identifying code duplications at the function level within C codebases.
-We expect this tool to effectively detect code duplications in the targeted subsystems 
-of the Linux kernel, as preliminary investigations suggest that kernel code files often 
-contain a mix of duplicated and unique functions. We will use a triangulation method to 
-validate the tool, comparing its results with an established code duplication database 
-referenced in the literature and conducting an empirical analysis on a randomly selected 
-set of files from the AMD Display driver, which our tool flags as containing duplications.
-A detailed explanation of our validation process is provided in Chapter \ref{cha:method}.
+With a validated tool, the research then moved from detection to practice, investigating 
+the real-world challenges of mitigating the identified duplications. This was accomplished 
+through an ethnographic study designed to engage with the kernel community and understand 
+their perspectives on code quality and refactoring.  This study had two components: a 
+participant-observation experiment where the author submitted refactoring patches to the 
+AMD Display driver, and a non-participant observation of university students using ArKanjo 
+to contribute fixes to both the AMD Display and Industrial I/O subsystems.
 
-Upon completing the first phase, we anticipate having a validated tool for identifying 
-code duplications within the context of the Linux kernel subsystems studied. We then 
-proceed to the second phase, which focuses on mitigating these duplications. 
-In \textbf{Phase II}, we aim to verify if we can mitigate the duplication found by 
-the developed tool. To achieve this, we have designed an ethnographic approach that 
-enables us to interact with the Linux development community and gather artifacts that 
-either support or challenge our proposed mitigation strategies. Figure \ref{fig:reDesign} 
-summarizes the research steps proposed in this study.
+By combining tool 
+development with direct community interaction, this research design allows for a comprehensive 
+understanding of the technical and social factors that influence the mitigation of duplicated 
+code in a large-scale open-source project.
 

### 2025-06-17 paper/paper.tex
udpate number on paper
add accent on my name
update my email

@@
 % affiliations
-\author{\IEEEauthorblockN{Luan Icaro Pinto Arcanjo}
+\author{\IEEEauthorblockN{Luan Ícaro Pinto Arcanjo}
 \IEEEauthorblockA{
@@
 São Paulo, SP Brazil 05508-220\\
-Email: }
+Email: luanicaro@usp.br}
 \and
@@
 The student efforts showed that newcomers could use the tool to make effective contributions. 
-Of the 12 student patches, three were accepted, three were refused, and six required follow-up 
+Of the 12 student patches, five were accepted, four were refused, and three required follow-up 
 fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 

### 2025-06-17 paper/paper.tex
Updating author's order and removing comments from the template
Reviewing the Abstract

@@
-%% bare_conf.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
+\documentclass[conference]{IEEEtran}
+\usepackage{cite}
+\usepackage{array}
+\usepackage{url}
 
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-\documentclass[conference]{IEEEtran}
-% Some Computer Society conferences also require the compsoc mode option,
-% but others use the standard conference format.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
 \ifCLASSINFOpdf
@@
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at: 
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
 
@@
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
+
 \title{ArKanjo: a tool for detecting function-level Code Duplication in the Linux Kernel}
@@
 \and
+\IEEEauthorblockN{David Tadokoro}
+\IEEEauthorblockA{
+Universidade de São Paulo\\
+São Paulo, SP Brazil 05508-220\\
+Email: davidbtadokoro@usp.br}
+\and
 \IEEEauthorblockN{Paulo Meirelles}
@@
 Email: paulormm@ime.usp.br}
-\and
-\IEEEauthorblockN{David Tadokoro}
-\IEEEauthorblockA{
-Universidade de São Paulo\\
-São Paulo, SP Brazil 05508-220\\
-Email: davidbtadokoro@usp.br}
+
 }
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3}, 
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
 
@@
 The Linux kernel’s massive scale (+28 M LoC, +20 K contributors) presents unique 
-maintenance challenges. Surprisingly, Code duplication remains a persistent issue 
-in the kernel’s codebase, which could hinder its evolution and patching. Academic 
-approaches often focus on pairwise comparison of code artifacts, not directly applied 
-for comprehensive codebase analyses. Other existing free software tools explored in 
+maintenance challenges. Surprisingly, code duplication remains a persistent issue 
+in the kernel’s codebase, which can hinder its evolution and patching. Academic 
+approaches often focus on pairwise comparison of code artifacts, which are not directly applied 
+to comprehensive codebase analyses. Other existing free software tools explored in 
 practice frequently suffer from limited functionality, such as primitive textual 
@@
 developers, and (3) integration with kernel development workflows.
-This paper presents ArKanjo, a novel command-line tool for Linux kernel maintenance 
+%
+This paper presents \textit{ArKanjo}, a novel command-line tool for Linux kernel maintenance 
 designed to detect and analyze function-level duplications. Released under the MIT 
@@
 and (3) prioritization of duplicates that impact maintainability, such as copied 
-buggy logic. We evaluate Arkanjo against real-world duplication cases in recent 
+buggy logic. 
+%
+We evaluate ArKanjo against real-world duplication cases in recent 
 kernel versions, demonstrating its effectiveness in identifying problematic clones 
@@
 software projects.
+
 \end{abstract}
@@
 
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a 
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively. 
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgment
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
 \bibliographystyle{IEEEtran}
@@
 \bibliography{IEEEabrv,referencias}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-% \begin{thebibliography}{1}
-%  
-%  \bibitem{IEEEhowto:kopka}
-%  H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-%    0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-%  
-% \end{thebibliography}
 
-% that's all folks
 \end{document}

### 2025-06-17 paper/paper.tex
Reviewing the Introduction section

@@
 \usepackage{url}
+\usepackage{todonotes}
 
@@
 
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 
-The Linux kernel is a foundational Free/Libre/Open Source Software (FLOSS) project, 
-critical to a significant portion of the world's digital infrastructure. Maintaining 
-the kernel is an enormous undertaking, involving more than 28 million lines of code 
+The Linux kernel is a foundational Free Software project 
+critical to a significant portion of the world digital infrastructure. Maintaining 
+the kernel is an enormous undertaking involving over 28 million lines of code 
 and contributions from over twenty thousand developers. Within this context, code 
 duplication persists as a significant challenge, a known harmful practice that 
-negatively affects code readability and increases the likelihood of introducing bugs 
-\cite{harmone,harmtwo}. This issue is particularly acute in the 
-kernel's device drivers, which comprise over 66\% of the source code \cite{marcelo}. 
+negatively affects code readability and increases the likelihood of introducing
+bugs~\cite{harmone,harmtwo}. This issue is particularly acute in the 
+kernel device drivers, which comprise over 66\% of the source code~\cite{marcelo}. 
 For instance, maintainers of the AMD Display driver have specifically highlighted code 
@@
 
-The detection of code duplication, or code clones, has been a subject of research for 
+\todo[inline]{colocar no bibtex e usar cite para (Jankowitz, 1988)}
+The detection of code duplication, or code clones, has been a research subject for 
 decades (Jankowitz, 1988). The literature provides a standard taxonomy, classifying 
 clones into four types based on their degree of similarity, from identical copies 
-(Type-1) to semantically equivalent but textually different fragments (Type-4) \cite{litreview}. 
+(Type-1) to semantically equivalent but textually different fragments (Type-4)~\cite{litreview}. 
 Various detection methodologies have been developed, including textual, token-based, 
-tree-based, and graph-based approaches \cite{litreview}. These have culminated in 
-state-of-the-art techniques, such as the graph-based work by Liu et al \cite{tailor}. 
-However, the primary focus of such academic work remains on determining if a given 
-pair of code artifacts are duplicates, rather than providing a scalable method to scan 
+tree-based, and graph-based approaches~\cite{litreview}. These have culminated in 
+state-of-the-art techniques, such as the graph-based work by Liu et al.~\cite{tailor}. 	
+However, the primary focus of such academic work remains on determining whether a given 
+pair of code artifacts are duplicates rather than providing a scalable method to scan 
 an entire codebase for actionable results. Conversely, existing free software tools explored 
-in practice often lack this sophistication, such as relying on primitive textual matching 
+in practice often lack this sophistication, relying instead on primitive textual matching 
 that is insufficient for the complexity and scale of the kernel.
 
-To fill this gap, this research proposes a new approach embodied in ArKanjo, a tool 
+To fill this gap, this research proposes a new approach embodied in \textit{ArKanjo}, a tool 
 designed specifically to identify and facilitate the mitigation of function-level code 

### 2025-06-17 paper/paper.tex
Adding the Acknowledgment section
Reviewing the ArKanjo Tool section

@@
 Universidade de São Paulo\\
-São Paulo, SP Brazil 05508-220\\
+São Paulo, SP Brazil\\
 Email: luanicaro@usp.br}
@@
 Universidade de São Paulo\\
-São Paulo, SP Brazil 05508-220\\
+São Paulo, SP Brazil\\
 Email: davidbtadokoro@usp.br}
@@
 Universidade de São Paulo\\
-São Paulo, SP Brazil 05508-220\\
+São Paulo, SP Brazil\\
 Email: paulormm@ime.usp.br}
@@
 \begin{abstract}
-The Linux kernel’s massive scale (+28 M LoC, +20 K contributors) presents unique 
+The Linux kernel massive scale (+28 M LoC, +20 K contributors) presents unique 
 maintenance challenges. Surprisingly, code duplication remains a persistent issue 
-in the kernel’s codebase, which can hinder its evolution and patching. Academic 
+in the kernel codebase, which can hinder its evolution and patching. Academic 
 approaches often focus on pairwise comparison of code artifacts, which are not directly applied 
@@
 matching, prove too narrow in scope, or fail to deliver effective results on complex, 
-large-scale codebases. Existing solutions generally fail to address the Linux kernel’s 
+large-scale codebases. Existing solutions generally fail to address the Linux kernel 
 specific needs: (1) scalability to handle its size, (2) actionable results for 
@@
 
-The Linux kernel is a foundational Free Software project 
+The Linux kernel is a foundational free software project 
 critical to a significant portion of the world digital infrastructure. Maintaining 
@@
 help developers identify code clone duplication at the function level. Released under the
-MIT license, ArKanjo is available at the url: \url{https://github.com/LipArcanjo/arkanjo}
+MIT license, ArKanjo is available at \textit{\url{github.com/LipArcanjo/arkanjo}}.
 
@@
 \centering
-\includegraphics[width=3in]{diagrama_mestrado}
+\includegraphics[width=3.5in]{diagrama_mestrado}
 \caption{Architecture diagram demonstrating the relationship between the tool components}
@@
 The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
-The \textbf{Preprocessor} is responsible for performing heavy computations to find the code 
-duplications across a codebase and produce artifacts to consult information related to the 
-duplications in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
-by the Preprocessor to execute the tool functionalities as requested. This choice enables the 
-tool to concentrate on the heavy and slow parts that can be executed only once per codebase, 
-thus allowing a fast performance for multiple queries related to a single codebase.
+The \textbf{Preprocessor} performs heavy computations to find code 
+duplications across a codebase and produces artifacts to store duplication-related information 
+in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
+by the Preprocessor to execute the tool functionalities as requested. This design allows the 
+tool to perform the heavy and time-consuming steps only once per codebase, enabling fast performance 
+for multiple queries on the same codebase.
 
-The Preprocessor contains two main components, the Function Breaker and the Code Duplication Finder. 
-The flow of how the Preprocessor works is as follow: Function Parser receives the codebase the user is interested in,
-extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files. 
-The Code Duplication Finder iterates over every pair of files in the Temporary Codebase, checks if they are code clones, and, 
-if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple 
-\textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates 
-of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized in the Code Duplication Finder.
-Figure \ref{fig:diagram} illustrates the tool architecture.
+The Preprocessor contains two main components: the Function Breaker and the Code Duplication Finder. 
+The workflow of the Preprocessor is as follows: the Function Parser receives the input codebase,
+extracts the functions along with metadata, and creates a temporary codebase where each extracted 
+function becomes a new code file. The Code Duplication Finder then iterates over every pair of files 
+in the temporary codebase, checks whether they are code clones, and, if so, stores the result in 
+the Code Duplication Database. This database is a text file that records each code duplication as a triple 
+\textbf{<function1, function2, similarity>}, where \textbf{function1} and \textbf{function2} are the 
+duplicated functions, and \textbf{similarity} is the score returned by the duplication detection method 
+used in the Code Duplication Finder. Figure~\ref{fig:diagram} illustrates the tool architecture.
 
-The Query Responder consumes the Temporary Codebase and the Code Duplication Database to extract duplicated 
-functions-related information per user request. If the user executes the Query Responder without executing the 
-Preprocessor, the Query Responder calls the Preprocessor to create the required artifacts.
+The Query Responder uses the temporary codebase and the Code Duplication Database to extract 
+information about duplicated functions based on the user's request. If the user executes the 
+Query Responder without previously executing the Preprocessor, the Query Responder automatically 
+calls the Preprocessor to generate the required artifacts.
 
 The Function Parser receives the input codebase and transforms it into the temporary
-codebase. The Function Parser iterates through every source code file from the programming 
-language it supports and uses a specific programming language function extractor
-to extract every function in the file. For each function extracted, two new source code files
-and a metadata file are created in the temporary codebase, represented by the pair <file
-name, function name>, the source code file of the function, and the proper function.
-The first new source code file contains the function’s body from the function it represents,
-while the other new source code file contains the function’s signature. The metadata file
-contains additional relevant information about the function, such as the function name,
-the line where the function signature starts in the source code file, and the line where the
-function’s body ends in the source code file. The programming languages supported at
-the moment are C and Java, although Java has limited support.
+codebase. It iterates through every source code file written in a supported programming 
+language and uses a language-specific extractor to isolate each function. For each extracted 
+function, two new source code files and a metadata file are created in the temporary codebase, 
+represented by the pair \textbf{<file name, function name>}, the source code of the function, 
+and the function metadata. One of the new files contains the function body, and the other 
+contains the function signature. The metadata file includes additional relevant information,
+such as the function name, the line where the signature starts, and the line where the function 
+body ends. Currently, the supported programming languages are C and Java, although Java support is limited.
 
 The Code Duplication Finder iterates through every pair of source code files in the
-temporary codebase, representing functions from the input codebase. For each pair of
-files, we execute a code duplication detection method that computes a metric measuring
-how similar the pair of files is, which we refer to as similarity. If the similarity is greater
-than or equal to the minimum similarity threshold (a parameter provided by the user
-during the Preprocessor), we store this pair of functions along with its similarity in the
-Code Duplication Database. 
-We implemented two methods for code duplication detection in the tool, which the user
-can choose to use. The first method is based on text similarity, and the second is simpler
-and based on the number of equal lines. The experiments and tests in this research were
-done using the text similarity method.
+temporary codebase, each representing a function from the input codebase. For each pair,
+a code duplication detection method computes a similarity score. If the similarity is greater 
+than or equal to the minimum threshold (provided by the user during preprocessing), the pair 
+is stored in the Code Duplication Database along with the similarity score.
+
+We implemented two duplication detection methods in the tool, which users can choose 
+between. The first method is based on text similarity, and the second is a more straightforward approach 
+based on the number of identical lines. The experiments and tests in this research were 
+conducted using the text similarity method.
 
@@
 
-For the text similarity method, we treat the source code files as text and apply the
-TF-IDF vector embedding method implemented by the Gensim library \cite{gensim},
-then compute cosine similarity as the similarity metric. This method was chosen for 
-its claimed performance, programming language independence, and the fact that it does 
-not require compilable code, which is expected in the
-temporary codebase, as it does not contain complete code artifacts.
+For the text similarity method, we treat the source code files as plain text and apply the
+TF-IDF vector embedding method implemented by the Gensim library~\cite{gensim},
+then compute cosine similarity as the similarity metric. We chose this method for 
+its reported performance, programming language independence, and the fact that it does 
+not require compilable code, an important aspect given that the
+temporary codebase does not contain complete code artifacts.
 
-For the difference method, 
-we implemented a method that considered the number of exactly equal lines between two functions.
-For two functions \textit{function1} and \textit{function2}, we compute the similarity as the 
-ratio of duplicated lines between the two functions by the total number of lines in both functions.
-This code duplication method is considerably simpler and more explainable than the text similarity method, 
-as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
-the \textit{diff} command built-in in the Linux environment.
+We implemented a more straightforward approach for the difference method,
+considering the number of exactly equal lines between two functions.
+For two functions, \textit{function1} and \textit{function2}, we compute the similarity as the 
+ratio of duplicated lines to the total number of lines across both functions.
+This method is considerably simpler and more explainable than the text similarity method, 
+The metric is defined by the proportion of common lines between the two functions. 
+We use the \textit{diff} command built into the Linux environment to compute the number of equal lines.
 
@@
 
-We propose three main functionalities for the user in the tool, called the Duplication
+We propose three main functionalities for the user in the tool, called Duplication
 Explorer, Function Information, and Duplication Report. Each functionality accepts
@@
 the user with pairs of duplicated functions identified by the tool. We implement optional
-filters to facilitate more complex queries. Figure \ref{fig:explorer_ex} demonstrates an
+filters to support more complex queries. Figure~\ref{fig:explorer_ex} demonstrates an
 example of this functionality in use.
@@
 The Function Information functionality provides detailed information about a specific function. 
-The functionality receives a target function from the user
+It receives a target function from the user
 and returns information such as the relative path, function name, and line numbers where
 the function is defined. Additionally, it provides similar information for every function
-that duplicates the given function. Figure \ref{fig:function_ex} demonstrates an
+that duplicates the given function. Figure~\ref{fig:function_ex} demonstrates an
 example of this functionality in use.
@@
 The Duplication Report functionality provides an overview of
-duplicated code within the input codebase. This functionality calculates the number of
+duplicated code within the input codebase. It calculates the number of
 duplicated lines per folder in the codebase and presents the information to the user in
-a readable format. Figure \ref{fig:relatory_ex} demonstrates an example of this functionality in use.
+a readable format. Figure~\ref{fig:relatory_ex} demonstrates an example of this functionality in use.
 
@@
 
+\section*{Acknowledgment}
+
+This study was financed, in part, by  the University of
+São Paulo/USP (Proc. 22.1.9345.1.2), the São Paulo Research Foundation/FAPESP
+and the São Paulo State Data Analysis System Foundation/SEADE (Proc. 2023/18026-8), Brazil.
+
 \bibliographystyle{IEEEtran}

### 2025-06-17 paper/paper.tex
Reviewing the Methods section

@@
 
-To validate the ArKanjo tool capabilities to find code duplications within the 
-text similarity detection method used, we applied two independent
-methods. The first method evaluates the proposed tool using an approach from the literature,
+\todo[inline]{colocar no bibtex e usar cite para (Svajlenko et al., 2014)}
+To validate the capabilities of the ArKanjo tool in finding code duplications using the 
+text similarity detection method, we applied two independent
+evaluation approaches. The first method assesses the tool using a literature-based approach,
 comparing it against the BigCloneBench dataset (Svajlenko et al., 2014). The second
-method is an empirical analysis of a subset of functions within the AMD Display driver
+method consists of an empirical analysis of a subset of functions within the AMD Display driver
 that our tool identified as duplicates.
@@
 To validate the tool against the BigCloneBench dataset \cite{bigclonebench}, we followed
-the methodology presentedy by Liu et al. \cite{tailor}, we sampled 20,000 pairs from each clone
-type, adding 20,000 non-duplicate pairs as negative samples. We applied the same sampling
-approach to our tool to ensure a fair evaluation.
-Unlike state-of-art methods, our tool does not distinguish between clone types and
+the methodology presented by Liu et al.~\cite{tailor}. We sampled 20{,}000 pairs from each clone
+type and added 20{,}000 non-duplicate pairs as negative samples. We applied the same sampling
+strategy to our tool to ensure a fair comparison.
+Unlike state-of-the-art methods, our tool does not distinguish between clone types and
 identifies duplications at the function level rather than the file level, which is typical in
@@
 purposes. Specifically, we considered every pair of functions with a similarity metric
-equal to or greater than a threshold X as duplicates and marked the corresponding file
-pairs. A correctly identified duplication pair counts as accurate for its clone type, while
+equal to or greater than a threshold $X$ as duplicates and marked the corresponding file
+pairs. A correctly identified duplication pair is counted as accurate for its clone type, while
 an incorrectly inferred pair is considered incorrect across all clone types.
@@
 For the empirical analysis, we randomly sampled function pairs identified by the tool as
-duplicates. For each similarity threshold X (30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%) 
-we randomly selected ten function pairs with a similarity close to X, allowing for a 1\%
+duplicates. For each similarity threshold $X$ (30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%), 
+we randomly selected ten function pairs with a similarity close to $X$, allowing for a 1\%
 deviation.
 
-To assess if the duplications found by the ArKanjo tool were actionable, the research conducted 
-a multi-part ethnographic study. In this study, the author and 12 student groups from the
-Free Software Development course at University of São Paulo acted as new contributors 
+To assess whether the duplications found by the ArKanjo tool were actionable, this research conducted 
+a multi-part ethnographic study. The author and 12 student groups from the
+Free Software Development course at the University of São Paulo acted as new contributors 
 to the Linux kernel. 
-They used the tool to identify duplications within the Industrial I/O (IIO) subsytem and AMD Display 
-driver subsystems, refactored the code to address the issues, and submitted their proposed 
-fixes as patches to the community maintainers. Throughout this process, the students documented 
-their refactoring approaches and their experiences interacting with the kernel community.
+They used the tool to identify duplications within the Industrial I/O (IIO) subsystem and the AMD Display 
+driver subsystem refactored the code to address the issues, and submitted their proposed 
+fixes as patches to the community maintainers. The students documented 
+their refactoring approaches and experiences interacting with the kernel community throughout this process.
 

### 2025-06-18 paper/paper.tex
resolve todos

@@
 
-\todo[inline]{colocar no bibtex e usar cite para (Jankowitz, 1988)}
 The detection of code duplication, or code clones, has been a research subject for 
-decades (Jankowitz, 1988). The literature provides a standard taxonomy, classifying 
+decades~\cite{firstman}. The literature provides a standard taxonomy, classifying 
 clones into four types based on their degree of similarity, from identical copies 
@@
 
-\todo[inline]{colocar no bibtex e usar cite para (Svajlenko et al., 2014)}
 To validate the capabilities of the ArKanjo tool in finding code duplications using the 
@@
 evaluation approaches. The first method assesses the tool using a literature-based approach,
-comparing it against the BigCloneBench dataset (Svajlenko et al., 2014). The second
+comparing it against the BigCloneBench dataset~\cite{bigclonebench}. The second
 method consists of an empirical analysis of a subset of functions within the AMD Display driver
@@
 
-To validate the tool against the BigCloneBench dataset \cite{bigclonebench}, we followed
+To validate the tool against the BigCloneBench dataset~\cite{bigclonebench}, we followed
 the methodology presented by Liu et al.~\cite{tailor}. We sampled 20{,}000 pairs from each clone

### 2025-06-18 paper/paper.tex
Reviewing the results and conclusion sections, and other minor adjustments

@@
 \IEEEauthorblockA{
-Universidade de São Paulo\\
-São Paulo, SP Brazil\\
-Email: luanicaro@usp.br}
+Universidade de São Paulo, Brazil\\
+luanicaro@usp.br}
 \and
@@
 \IEEEauthorblockA{
-Universidade de São Paulo\\
-São Paulo, SP Brazil\\
-Email: davidbtadokoro@usp.br}
+Universidade de São Paulo, Brazil\\
+davidbtadokoro@usp.br}
 \and
@@
 \IEEEauthorblockA{
-Universidade de São Paulo\\
-São Paulo, SP Brazil\\
-Email: paulormm@ime.usp.br}
+Universidade de São Paulo, Brazil\\
+paulormm@ime.usp.br}
 
@@
 \includegraphics[width=3.5in]{diagrama_mestrado}
-\caption{Architecture diagram demonstrating the relationship between the tool components}
+\caption{Architecture diagram with the tool components}
 \label{fig:diagram}
@@
 
-The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
+The tool operates in two main parts, as illustrates in Figure~\ref{fig:diagram}: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
 The \textbf{Preprocessor} performs heavy computations to find code 
@@
 duplicated functions, and \textbf{similarity} is the score returned by the duplication detection method 
-used in the Code Duplication Finder. Figure~\ref{fig:diagram} illustrates the tool architecture.
+used in the Code Duplication Finder.
 
 The Query Responder uses the temporary codebase and the Code Duplication Database to extract 
-information about duplicated functions based on the user's request. If the user executes the 
-Query Responder without previously executing the Preprocessor, the Query Responder automatically 
+information about duplicated functions based on the user's request. If the user executes it
+without previously executing the Preprocessor, the Query Responder automatically 
 calls the Preprocessor to generate the required artifacts.
@@
 \centering
-\includegraphics[width=3in]{explorer_example}
+\includegraphics[width=3.5in]{explorer_example}
 \caption{Example of the \textit{Duplication Explorer} functionality of the ArKanjo tool.}
@@
 \centering
-\includegraphics[width=3in]{function_example}
+\includegraphics[width=3.5in]{function_example}
 \caption{Example of the \textit{Function Information} functionality of the ArKanjo tool.}
@@
 \centering
-\includegraphics[width=3in]{relatory_example}
+\includegraphics[width=3.5in]{relatory_example}
 \caption{Example of the \textit{Duplication Report} functionality of the ArKanjo tool.}
@@
 
+In this section, we assess the ArKanjo tool effectiveness in detecting code duplications through benchmark comparison and empirical analysis.
+We also describe findings from an ethnographic study involving real contributions to the Linux kernel, highlighting the tool practical utility and limitations in developer workflows.
+
 \subsection{Tool Evaluation}
@@
 \centering
-
+\caption{Recall of the ArKanjo tool in the BigCloneBench dataset.}
 \renewcommand{\arraystretch}{1.3}
@@
 \end{tabular}
-\caption{Recall of the ArKanjo tool in the BigCloneBench dataset varying minimum similarity threshould.}
+
 \label{tab:bigclone}
@@
 \centering
-
+\caption{Results of the ArKanjo tool analyzing the AMD Display driver.}
 \begin{tabular}{ | m{15mm} | c | c | c | c | m{6mm} | m{10mm} | }
@@
 \end{tabular}
-\caption{Results of the ArKanjo tool in the empirical analyses in the AMD Display driver}
+
 \label{tab:emp}
@@
 
-Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCLoneBench dataset
-and table \ref{tab:emp} shows the results obtained in the empricial analysis.
-The BigCloneBench results and our empirical methods indicate that our tool performs
-well in detecting Type-1 and Type-2 code clone duplications and can thus reveal propagated 
-issues like copied buggy logic, but it struggles with more
-complex types. A similarity threshold between 80\% and 100\% yields favorable results
-in both methods.
+Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCloneBench dataset,
+and Table \ref{tab:emp} presents the results from the empirical analysis.
+The outcomes from both the BigCloneBench evaluation and our empirical method indicate that
+our tool performs well in detecting Type-1 and Type-2 code clone duplications,
+which are sufficient to reveal propagated issues like copied buggy logic.
+However, it struggles with more complex clone types. A similarity threshold between 80\% and 100\% 
+yields favorable results in both methods.
 
-However, there is a discrepancy in the detection of non-duplication pairs between
-the two methods. In the BigCloneBench evaluation, we found no false positives (negative
-samples inferred as duplications), while the empirical analysis revealed a higher percentage
+Nevertheless, a discrepancy emerges when detecting non-duplicate pairs between
+the two methods. In the BigCloneBench evaluation, we found no false positives (i.e., negative
+samples inferred as duplications), whereas the empirical analysis revealed a higher percentage
 of false positives. We propose two potential explanations for this discrepancy.
 
-The first reason may be the limitations and known issues with BigCloneBench, as
+The first explanation concerns the limitations and known issues with BigCloneBench, as
 highlighted by Krinke and Ragkhitwetsagul \cite{bigfail}. The
-second reason could relate to the nature of the AMD Display driver, where each code artifact
-shares semantic meaning. In contrast, BigCloneBench comprises self-contained code
-artifacts without shared semantics. Given that our tool relies on a text-based code clone
+second explanation relates to the nature of the AMD Display driver, where code artifacts
+often share semantic meaning. In contrast, BigCloneBench comprises self-contained code
+artifacts with minimal or no shared semantics. Since our tool relies on a text-based code clone
 detection approach, it may naturally perform worse on the AMD Display driver than on the
@@
 
-Using the ArKanjo tool, the author and 12 student groups identified and refactored duplicated 
+Using the ArKanjo tool, the main author and 12 student groups identified and refactored duplicated 
 code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work 
-as patches to the community maintainers. This effort resulted in 13 total patch submissions: 
-one from the author to the AMD Display driver and 12 from the student groups, with 11 targeting 
-the IIO subsystem and one for the AMD Display driver. The author's patch, which removed over 
-500 lines of duplicated code, was successfully accepted and merged into the kernel after a 
-lengthy review process.
+as patches to the community maintainers. This effort resulted in 13 patch submissions: 
+One was from the main author to the AMD Display driver, and 12 were from the student groups, with 11 targeting 
+the IIO subsystem and one aimed at the AMD Display driver. The author's patch was successfully accepted and merged into the kernel after a 
+lengthy review process, removing over 
+500 lines of duplicated code.
 
-The student efforts showed that newcomers could use the tool to make effective contributions. 
-Of the 12 student patches, five were accepted, four were refused, and three required follow-up 
+The student efforts demonstrated that newcomers could use the tool to contribute effectively. 
+Of the 12 student patches, five were accepted, four were rejected, and three required follow-up 
 fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 
-Parameterize Method and two used the Extract Method \cite{refactorbook}, which are simple refactoring methods. 
-This result corroborates that ArKanjo finds actionable duplications and that newcomers can 
-successfully contribute patches to the kernel.
+\textit{Parameterize Method}, and two used the \textit{Extract Method} \cite{refactorbook}, both of which are 
+straightforward refactoring strategies. These results support the conclusion that ArKanjo identifies 
+actionable duplications and that newcomers can successfully contribute patches to the kernel.
 
-While this research included a successfully merged patch by the author and students,
-a notable portion of student efforts encountered
-significant hurdles. These difficulties often stemmed from maintainer feedback where
-proposed changes, though reducing duplication, were rejected or required substantial
-rework due to concerns about code readability, added abstraction, or the specific context
-of the code. Technical complexities, such as C macros in configuration files, and lengthy
-patch review times also contributed to the challenges. 
-This highlights that while ArKanjo effectively identifies duplications, a successful 
-contribution involves a trade-off between code deduplication and other factors like the 
-specific code's context and maintainer priorities.
+While this research includes successfully merged patches by the main author and student groups,
+a significant portion of the student contributions encountered notable challenges. These difficulties 
+often stemmed from maintainer feedback indicating that, although the proposed changes reduced duplication, 
+they were rejected or required substantial revision due to concerns about code readability, increased abstraction, 
+or the specific context of the code. Additional challenges included technical complexities, such as C macros in 
+configuration files and extended patch review timelines.
+
+These findings highlight that while ArKanjo effectively identifies code duplications, a successful 
+contribution also requires navigating trade-offs between deduplication and other factors, such as 
+contextual appropriateness and maintainer expectations.
 
@@
 
-In conclusion, this paper presented ArKanjo, a novel tool designed to effectively detect 
+This paper presented ArKanjo, a novel tool designed to detect effectively 
 and facilitate the refactoring of function-level code duplication within the Linux kernel. 
 Evaluations demonstrated that ArKanjo is highly effective at identifying Type-1 and Type-2 
-clones, which are often indicative of propagated issues like copied buggy logic. The practical 
-value of the tool was confirmed through an ethnographic study where the author and student 
+clones, which are often indicative of propagated issues such as copied buggy logic. We validated the practical 
+value of the tool through an ethnographic study in which the main author and student 
 groups used ArKanjo to identify actionable duplications. This effort led to 13 patch 
-submissions, including one from the author that removed over 500 lines of code and was merged 
-into the kernel, alongside three successful student contributions. While some refactoring 
-efforts were declined due to maintainer priorities, ArKanjo successfully lowers the barrier 
+submissions, including one from the main author that removed over 500 lines of code and merged 
+into the kernel, alongside five successful student contributions at this moment. While some refactoring 
+efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 
 for new contributors to make meaningful improvements. The tool provides immediate value for 
@@
 
-This study was financed, in part, by  the University of
-São Paulo/USP (Proc. 22.1.9345.1.2), the São Paulo Research Foundation/FAPESP
-and the São Paulo State Data Analysis System Foundation/SEADE (Proc. 2023/18026-8), Brazil.
+This study was financed, in part, by CAPES (Finance Code 001), the University of São Paulo – USP (Proc.
+22.1.9345.1.2), the São Paulo Research Foundation – FAPESP and the São Paulo State Data Analysis System Foundation – SEADE (Proc.
+2023/18026-8), Brazil.
 

### 2025-06-18 paper/paper.tex
Adding footnote about Free Software, and other minor adjustments

@@
 % affiliations
-\author{\IEEEauthorblockN{Luan Ícaro Pinto Arcanjo}
+\author{\IEEEauthorblockN{Luan Arcanjo}
 \IEEEauthorblockA{
@@
 approaches often focus on pairwise comparison of code artifacts, which are not directly applied 
-to comprehensive codebase analyses. Other existing free software tools explored in 
+to comprehensive codebase analyses. Other existing free
+software\footnote{In this paper, we use the term \textit{Free Software}
+interchangeably with \textit{Open Source Software}.} tools explored in 
 practice frequently suffer from limited functionality, such as primitive textual 
@@
 designed specifically to identify and facilitate the mitigation of function-level code 
-duplications within the Linux kernel. To test the tool, we employ multiple analytical 
-methods and ethnographic studies.
+duplications within the Linux kernel. We employ multiple analytical 
+methods and ethnographic studies to test it.
 
@@
 the user with pairs of duplicated functions identified by the tool. We implement optional
-filters to support more complex queries. Figure~\ref{fig:explorer_ex} demonstrates an
-example of this functionality in use.
+filters to support more complex queries. Figure~\ref{fig:explorer_ex}
+demonstrates an example of this functionality in use.
 
@@
 \includegraphics[width=3.5in]{explorer_example}
-\caption{Example of the \textit{Duplication Explorer} functionality of the ArKanjo tool.}
+\caption{Example of the \textit{Duplication Explorer} functionality.}
 \label{fig:explorer_ex}
@@
 \includegraphics[width=3.5in]{function_example}
-\caption{Example of the \textit{Function Information} functionality of the ArKanjo tool.}
+\caption{Example of the \textit{Function Information} functionality.}
 \label{fig:function_ex}
@@
 \includegraphics[width=3.5in]{relatory_example}
-\caption{Example of the \textit{Duplication Report} functionality of the ArKanjo tool.}
+\caption{Example of the \textit{Duplication Report} functionality.}
 \label{fig:relatory_ex}
@@
 duplicated lines per folder in the codebase and presents the information to the user in
-a readable format. Figure~\ref{fig:relatory_ex} demonstrates an example of this functionality in use.
+a readable format, as demonstrated in Figure~\ref{fig:relatory_ex}.
 
@@
 
-Nevertheless, a discrepancy emerges when detecting non-duplicate pairs between
+A discrepancy emerges when detecting non-duplicate pairs between
 the two methods. In the BigCloneBench evaluation, we found no false positives (i.e., negative

### 2025-06-19 paper/paper.tex
Update paper: several adjustments and fix to compile with pdflatex command

@@
 \usepackage{url}
+\usepackage{graphicx}
 \usepackage{todonotes}
-
-\ifCLASSINFOpdf
-   \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  %\graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  \usepackage{graphicx}
-  \graphicspath{{figuras/},{fig},{logos/},{img/},{images/},{imagens/}}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{./eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  \DeclareGraphicsExtensions{.eps, .png, .jpeg}
-\fi
 
@@
 state-of-the-art techniques, such as the graph-based work by Liu et al.~\cite{tailor}.
+
 However, the primary focus of such academic work remains on determining whether a given 
@@
 an entire codebase for actionable results. Conversely, existing free software tools explored 
-in practice often lack this sophistication, relying instead on primitive textual matching 
+in practice, often lack this sophistication, relying instead on primitive textual matching 
 that is insufficient for the complexity and scale of the kernel.
@@
 
-\begin{figure}[!t]
+\begin{figure}[ht]
 \centering
-\includegraphics[width=3.5in]{diagrama_mestrado}
+\includegraphics[width=3.5in]{fig/diagrama_mestrado.png}
 \caption{Architecture diagram with the tool components}
@@
 
-The tool operates in two main parts, as illustrates in Figure~\ref{fig:diagram}: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
+The tool operates in two main parts, as illustrated in Figure~\ref{fig:diagram}: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
 The \textbf{Preprocessor} performs heavy computations to find code 
@@
 the Code Duplication Database. This database is a text file that records each code duplication as a triple 
-\textbf{<function1, function2, similarity>}, where \textbf{function1} and \textbf{function2} are the 
+\textbf{$<$function1, function2, similarity$>$}, where \textbf{function1} and \textbf{function2} are the 
 duplicated functions, and \textbf{similarity} is the score returned by the duplication detection method 
@@
 The Query Responder uses the temporary codebase and the Code Duplication Database to extract 
-information about duplicated functions based on the user's request. If the user executes it
+information about duplicated functions based on the user's request. If the user executes the Query Responder
 without previously executing the Preprocessor, the Query Responder automatically 
@@
 function, two new source code files and a metadata file are created in the temporary codebase, 
-represented by the pair \textbf{<file name, function name>}, the source code of the function, 
+represented by the pair \textbf{$<$file name, function name$>$}, the source code of the function, 
 and the function metadata. One of the new files contains the function body, and the other 
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=3.5in]{fig/explorer_example.png}
+\caption{Example of the \textit{Duplication Explorer} functionality.}
+\label{fig:explorer_ex}
+\end{figure}
+
 The Duplication Explorer is the primary functionality of our tool, designed to present
@@
 
-\begin{figure}[!t]
+\begin{figure}[ht]
 \centering
-\includegraphics[width=3.5in]{explorer_example}
-\caption{Example of the \textit{Duplication Explorer} functionality.}
-\label{fig:explorer_ex}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=3.5in]{function_example}
+\includegraphics[width=3.5in]{fig/function_example.png}
 \caption{Example of the \textit{Function Information} functionality.}
 \label{fig:function_ex}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[width=3.5in]{relatory_example}
-\caption{Example of the \textit{Duplication Report} functionality.}
-\label{fig:relatory_ex}
 \end{figure}
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=4in]{fig/relatory_example.png}
+\caption{Example of the \textit{Duplication Report} functionality.}
+\label{fig:relatory_ex}
+\end{figure}
+
 The Duplication Report functionality provides an overview of
@@
 
-To validate the capabilities of the ArKanjo tool in finding code duplications using the 
-text similarity detection method, we applied two independent
-evaluation approaches. The first method assesses the tool using a literature-based approach,
+We applied two independent evaluation approaches to validate the capabilities of the ArKanjo tool in finding code duplications using the text similarity detection method.
+The first method assesses the tool using a literature-based approach,
 comparing it against the BigCloneBench dataset~\cite{bigclonebench}. The second
@@
 strategy to our tool to ensure a fair comparison.
+
 Unlike state-of-the-art methods, our tool does not distinguish between clone types and
@@
 
-To assess whether the duplications found by the ArKanjo tool were actionable, this research conducted 
-a multi-part ethnographic study. The author and 12 student groups from the
-Free Software Development course at the University of São Paulo acted as new contributors 
-to the Linux kernel. 
-They used the tool to identify duplications within the Industrial I/O (IIO) subsystem and the AMD Display 
-driver subsystem refactored the code to address the issues, and submitted their proposed 
-fixes as patches to the community maintainers. The students documented 
-their refactoring approaches and experiences interacting with the kernel community throughout this process.
+We conducted a multi-part ethnographic study to assess whether the duplications found by the ArKanjo tool were actionable.
+In the second semester of 2024 and the first semester of 2025, respectively, the main author, conducting a first validation, and 12 student groups (19 undergraduate and 6 graduate students) from the Free Software Development course at the University of São Paulo acted as new contributors to the Linux kernel in a broader experiment, used the tool to identify duplications within the Industrial I/O (IIO) subsystem and the AMD Display driver subsystem refactored the code to address the issues, and submitted their proposed fixes as patches to the community maintainers.
+The students documented their refactoring approaches and experiences interacting with the kernel community throughout this process.
 
@@
 
-\begin{table}[!t]
+\begin{table}[ht]
 
@@
 
-A discrepancy emerges when detecting non-duplicate pairs between
+Nevertheless, a discrepancy emerges when detecting non-duplicate pairs between
 the two methods. In the BigCloneBench evaluation, we found no false positives (i.e., negative
@@
 The first explanation concerns the limitations and known issues with BigCloneBench, as
-highlighted by Krinke and Ragkhitwetsagul \cite{bigfail}. The
+highlighted by Krinke and Ragkhitwetsagul~\cite{bigfail}. The
 second explanation relates to the nature of the AMD Display driver, where code artifacts
@@
 
-Using the ArKanjo tool, the main author and 12 student groups identified and refactored duplicated 
-code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work 
+Using the ArKanjo tool, the main author and 26 students (divided into 12 groups) identified and refactored duplicated code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work
 as patches to the community maintainers. This effort resulted in 13 patch submissions: 
 One was from the main author to the AMD Display driver, and 12 were from the student groups, with 11 targeting 
-the IIO subsystem and one aimed at the AMD Display driver. The author's patch was successfully accepted and merged into the kernel after a 
-lengthy review process, removing over 
-500 lines of duplicated code.
+the IIO subsystem and one aimed at the AMD Display driver.
+After a lengthy review process, the main author's patch was successfully accepted and merged into the kernel, removing over 500 lines of duplicated code.
 
@@
 
+\newpage
 \bibliographystyle{IEEEtran}

### 2025-06-19 paper/paper.tex
fix the number of graduate students

@@
 We conducted a multi-part ethnographic study to assess whether the duplications found by the ArKanjo tool were actionable.
-In the second semester of 2024 and the first semester of 2025, respectively, the main author, conducting a first validation, and 12 student groups (19 undergraduate and 6 graduate students) from the Free Software Development course at the University of São Paulo acted as new contributors to the Linux kernel in a broader experiment, used the tool to identify duplications within the Industrial I/O (IIO) subsystem and the AMD Display driver subsystem refactored the code to address the issues, and submitted their proposed fixes as patches to the community maintainers.
+In the second semester of 2024 and the first semester of 2025, respectively, the main author, conducting a first validation, and 12 student groups (19 undergraduate and 7 graduate students) from the Free Software Development course at the University of São Paulo acted as new contributors to the Linux kernel in a broader experiment, used the tool to identify duplications within the Industrial I/O (IIO) subsystem and the AMD Display driver subsystem refactored the code to address the issues, and submitted their proposed fixes as patches to the community maintainers.
 The students documented their refactoring approaches and experiences interacting with the kernel community throughout this process.

### 2025-06-22 conteudo/introduction/intro.tex
Update intro.tex

@@
 ArKanjo can effectively lower the barrier for new contributors, as evidenced by its role in 
-guiding students to make their first code improvements to the kernel.
+guiding students to make their first code contributions to the kernel.
 

### 2025-06-22 conteudo/methods/methods.tex
update number of students on the tesis

@@
 
-Of the students who chose to work on the ArKanjo-related tasks, 25 (20 undergraduate and 5 graduate) 
+Of the students who chose to work on the ArKanjo-related tasks, 26 (19 undergraduate and 7 graduate) 
 formed 11 groups to pursue the first alternative. One graduate student opted for the second alternative, 

### 2025-06-25 conteudo/results/students.tex
Fix typo

@@
 
-Groups nine and eleven faced similar issues to those in Mitigaiton 1 on our participant observation 
+Groups nine and eleven faced similar issues to those in Mitigation 1 on our participant observation 
 experiment, where configuration files designed as macros complicated the approach to the mitigations. 

### 2025-06-25 conteudo/results/students.tex
Update students.tex

@@
 the same maintainer who approved the patch from group three. The maintainer justification to reject 
-the patch is that group ten refactored code files that has simple context, threfore adding a layer 
+the patch is that group ten refactored code files that has simple context, therefore adding a layer 
 of abstraction to remove a duplication deacrease the code readability without clear gains on 

### 2025-06-25 conteudo/results/students.tex
Update students.tex

@@
 experiment, where configuration files designed as macros complicated the approach to the mitigations. 
-Group nine applied the parameterized method so as not to need to know the configurations in the 
+Group nine applied the parameterized method to not need to know the configurations in the 
 generic function, which is reasonable as the duplication approach only depends on one configuration 

### 2025-07-01 paper/paper.tex
Fix the simplest points from revision
It is missing:
- the url.
- the textfile with the accepted patches.
- information about speed.
- future works (maybe)

@@
 querying for duplications within large codebases. Pivotal advantages of ArKanjo over 
-existing solutions include: (1) optimization for C codebases with kernel-specific 
-patterns; (2) preprocessing that enables rapid queries without redundant analysis; 
-and (3) prioritization of duplicates that impact maintainability, such as copied 
+existing solutions include: (1) preprocessing that enables rapid queries without redundant analysis; 
+and (2) prioritization of duplicates that impact maintainability, such as copied 
 buggy logic. 
@@
 represented by the pair \textbf{$<$file name, function name$>$}, the source code of the function, 
-and the function metadata. One of the new files contains the function body, and the other 
+and the function metadata (in practice, this is a duplication of the original codebase, 
+structured as a parsed representation). One of the new files contains the function body, and the other 
 contains the function signature. The metadata file includes additional relevant information,
@@
 
-Unlike state-of-the-art methods, our tool does not distinguish between clone types and
-identifies duplications at the function level rather than the file level, which is typical in
-state-of-the-art tools. Therefore, we adapted the metric calculation method for evaluation
+Unlike state-of-the-art methods, our tool does not distinguish between clone types. 
+Additionally, it identifies duplications at the function level, whereas most existing 
+tools operate at the file level.
+Therefore, we adapted the metric calculation method for evaluation
 purposes. Specifically, we considered every pair of functions with a similarity metric
@@
 
-Using the ArKanjo tool, the main author and 26 students (divided into 12 groups) identified and refactored duplicated code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work
-as patches to the community maintainers. This effort resulted in 13 patch submissions: 
-One was from the main author to the AMD Display driver, and 12 were from the student groups, with 11 targeting 
-the IIO subsystem and one aimed at the AMD Display driver.
-After a lengthy review process, the main author's patch was successfully accepted and merged into the kernel, removing over 500 lines of duplicated code.
+Using the ArKanjo tool, the main author and 23 students (divided into 11 groups) identified and refactored 
+duplicated code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work 
+as patches to the community maintainers. This effort resulted in 12 patch submissions: 
+One was from the main author to the AMD Display driver, and 11 were from the student groups, with 10 targeting 
+the IIO subsystem and one aimed at the AMD Display driver. After a lengthy review process, the main author's patch 
+was successfully accepted and merged into the kernel, removing over 500 lines of duplicated code.
 
 The student efforts demonstrated that newcomers could use the tool to contribute effectively. 
-Of the 12 student patches, five were accepted, four were rejected, and three required follow-up 
+Of the 11 student patches, five were accepted, four were rejected, and three required follow-up 
 fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 
@@
 value of the tool through an ethnographic study in which the main author and student 
-groups used ArKanjo to identify actionable duplications. This effort led to 13 patch 
-submissions, including one from the main author that removed over 500 lines of code and merged 
-into the kernel, alongside five successful student contributions at this moment. While some refactoring 
+groups used ArKanjo to identify actionable duplications. This effort led to 13 patch submissions, 
+including one from the main author and five from student groups that were successfully merged into 
+the kernel, collectively removing 991 lines of duplicated code. While some refactoring 
 efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 

### 2025-07-02 paper/paper.tex
maybe final version

@@
 \usepackage{todonotes}
+\usepackage{hyperref}
 
@@
 help developers identify code clone duplication at the function level. Released under the
-MIT license, ArKanjo is available at \textit{\url{github.com/LipArcanjo/arkanjo}}.
+MIT license, ArKanjo is available at 
+\textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.
 
@@
 \textbf{Similarity range} & \textbf{T1} & \textbf{T2} & T3 & T4
-& \textbf{False} & \textbf{Success Percentage} \\ \hline
+& \textbf{False} & \textbf{Success Ratio} \\ \hline
 99\% - 100\% & 9 & 0 & 0 & 1 & 0 & \textbf{100\%} \\ \hline
@@
 
+%Although a formal performance benchmark was not conducted, we recorded the tool's performance 
+%on a machine equipped with a Ryzen 5700x processor and 32GB of RAM. The Preprocessor's 
+%worst-case execution time across five runs was 8 minutes and 56 seconds on the AMD Display driver 
+%and 9 minutes and 41 seconds on the IIO subsystem. In contrast, the Query Responder processed 
+%requests in under two seconds, demonstrating the efficiency of the two-stage architecture 
+%described in this paper.
+
+While a formal performance benchmark was not conducted, we measured the Preprocessor's execution 
+time on a machine equipped with a Ryzen 5700X processor and 32GB of RAM. Across five runs on the 
+AMD Display driver, the worst-case execution time for the Preprocessor was 8 minutes and 56 seconds. 
+On the IIO subsystem, the worst-case time over five runs was 9 minutes and 41 seconds. 
+In contrast, the Query Responder processed requests in under two seconds in both cases, 
+demonstrating the efficiency of the two-stage architecture.
+
 \subsection{Ethnographic Study}
@@
 the IIO subsystem and one aimed at the AMD Display driver. After a lengthy review process, the main author's patch 
-was successfully accepted and merged into the kernel, removing over 500 lines of duplicated code.
+was successfully accepted and merged into the kernel.
 
 The student efforts demonstrated that newcomers could use the tool to contribute effectively. 
-Of the 11 student patches, five were accepted, four were rejected, and three required follow-up 
-fixes based on maintainer feedback. To remove the duplicated code, seven student groups used the 
-\textit{Parameterize Method}, and two used the \textit{Extract Method} \cite{refactorbook}, both of which are 
+Of the 11 student patches, 5 were accepted, 5 were rejected, and 1 required follow-up 
+fixes based on maintainer feedback. To remove the duplicated code, 7 student groups used the 
+\textit{Parameterize Method}, and 2 used the \textit{Extract Method} \cite{refactorbook}, both of which are 
 straightforward refactoring strategies. These results support the conclusion that ArKanjo identifies 
@@
 or the specific context of the code. Additional challenges included technical complexities, such as C macros in 
-configuration files and extended patch review timelines.
-
-These findings highlight that while ArKanjo effectively identifies code duplications, a successful 
+configuration files and extended patch review timelines. These findings highlight that while ArKanjo 
+effectively identifies code duplications, a successful 
 contribution also requires navigating trade-offs between deduplication and other factors, such as 
@@
 including one from the main author and five from student groups that were successfully merged into 
-the kernel, collectively removing 991 lines of duplicated code. While some refactoring 
+the kernel
+\footnote{Reference for the accepted patches:  
+\textit{\href{https://github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}{github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}}.
+}, 
+collectively removing 991 lines of duplicated code. While some refactoring 
 efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 

### 2025-07-03 paper/paper.tex
Minor adjustments

@@
 matching, prove too narrow in scope, or fail to deliver effective results on complex, 
-large-scale codebases. Existing solutions generally fail to address the Linux kernel 
-specific needs: (1) scalability to handle its size, (2) actionable results for 
+large-scale codebases. Existing solutions generally fail to address the Linux kernel-specific needs: (1) scalability to handle its size, (2) actionable results for 
 developers, and (3) integration with kernel development workflows.
@@
 negatively affects code readability and increases the likelihood of introducing
-bugs~\cite{harmone,harmtwo}. This issue is particularly acute in the 
-kernel device drivers, which comprise over 66\% of the source code~\cite{marcelo}. 
-For instance, maintainers of the AMD Display driver have specifically highlighted code 
-duplication as a significant impediment to their work.
+bugs~\cite{harmone,harmtwo}. This issue is particularly acute in kernel device drivers, which comprise over 66\% of the source code~\cite{marcelo}. 
+For instance, maintainers of the AMD Display driver have specifically highlighted code duplication as a significant impediment to their work.
 
-The detection of code duplication, or code clones, has been a research subject for 
+Detecting code duplication, or code clones, has been a research subject for 
 decades~\cite{firstman}. The literature provides a standard taxonomy, classifying 
@@
 
-However, the primary focus of such academic work remains on determining whether a given 
-pair of code artifacts are duplicates rather than providing a scalable method to scan 
+However, the primary focus of such academic work remains on determining whether a given pair of code artifacts is a duplicate rather than providing a scalable method to scan 
 an entire codebase for actionable results. Conversely, existing free software tools explored 
@@
 The Preprocessor contains two main components: the Function Breaker and the Code Duplication Finder. 
-The workflow of the Preprocessor is as follows: the Function Parser receives the input codebase,
-extracts the functions along with metadata, and creates a temporary codebase where each extracted 
-function becomes a new code file. The Code Duplication Finder then iterates over every pair of files 
+The workflow of the Preprocessor is as follows: the Function Parser receives the input codebase, extracts the functions along with metadata, and creates a temporary codebase where each extracted function becomes a new code file.
+The Code Duplication Finder then iterates over every pair of files
 in the temporary codebase, checks whether they are code clones, and, if so, stores the result in 
@@
 For the text similarity method, we treat the source code files as plain text and apply the
-TF-IDF vector embedding method implemented by the Gensim library~\cite{gensim},
-then compute cosine similarity as the similarity metric. We chose this method for 
-its reported performance, programming language independence, and the fact that it does 
-not require compilable code, an important aspect given that the
-temporary codebase does not contain complete code artifacts.
+TF-IDF vector embedding method implemented by the Gensim library~\cite{gensim}, then compute cosine similarity as the similarity metric.
+We chose this method for its reported performance, programming language independence, and the fact that it does
+not require compilable code, an important aspect given that the temporary codebase does not contain complete code artifacts.
 
@@
 ratio of duplicated lines to the total number of lines across both functions.
-This method is considerably simpler and more explainable than the text similarity method, 
+This method is considerably simpler and more explainable than the text similarity method. 
 The metric is defined by the proportion of common lines between the two functions. 
@@
 
-We propose three main functionalities for the user in the tool, called Duplication
-Explorer, Function Information, and Duplication Report. Each functionality accepts
+We propose three main functionalities for the user in the tool: Duplication Explorer, Function Information, and Duplication Report. Each functionality accepts
 specific parameters to perform its operations.
@@
 the user with pairs of duplicated functions identified by the tool. We implement optional
-filters to support more complex queries. Figure~\ref{fig:explorer_ex}
-demonstrates an example of this functionality in use.
+filters to support more complex queries. Figure~\ref{fig:explorer_ex} demonstrates an example of this functionality in use.
 
@@
 Unlike state-of-the-art methods, our tool does not distinguish between clone types. 
-Additionally, it identifies duplications at the function level, whereas most existing 
-tools operate at the file level.
+Additionally, it identifies duplications at the function level, whereas most existing tools operate at the file level.
 Therefore, we adapted the metric calculation method for evaluation
@@
 
-%Although a formal performance benchmark was not conducted, we recorded the tool's performance 
-%on a machine equipped with a Ryzen 5700x processor and 32GB of RAM. The Preprocessor's 
-%worst-case execution time across five runs was 8 minutes and 56 seconds on the AMD Display driver 
-%and 9 minutes and 41 seconds on the IIO subsystem. In contrast, the Query Responder processed 
-%requests in under two seconds, demonstrating the efficiency of the two-stage architecture 
-%described in this paper.
-
-While a formal performance benchmark was not conducted, we measured the Preprocessor's execution 
-time on a machine equipped with a Ryzen 5700X processor and 32GB of RAM. Across five runs on the 
+While we did not conduct a formal performance benchmark, we measured the Preprocessor's execution 
+time on a machine equipped with a Ryzen 5700X processor and 32GB of RAM. Across 5 runs on the 
 AMD Display driver, the worst-case execution time for the Preprocessor was 8 minutes and 56 seconds. 
-On the IIO subsystem, the worst-case time over five runs was 9 minutes and 41 seconds. 
-In contrast, the Query Responder processed requests in under two seconds in both cases, 
+On the IIO subsystem, the worst-case time over 5 runs was 9 minutes and 41 seconds. 
+In contrast, the Query Responder processed requests in under 2 seconds in both cases, 
 demonstrating the efficiency of the two-stage architecture.
@@
 
-The student efforts demonstrated that newcomers could use the tool to contribute effectively. 
-Of the 11 student patches, 5 were accepted, 5 were rejected, and 1 required follow-up 
+The students' efforts demonstrated that newcomers could use the tool to contribute effectively. 
+Of the 11 students' patches, 5 were accepted, 5 were rejected, and 1 required follow-up 
 fixes based on maintainer feedback. To remove the duplicated code, 7 student groups used the 
 \textit{Parameterize Method}, and 2 used the \textit{Extract Method} \cite{refactorbook}, both of which are 
-straightforward refactoring strategies. These results support the conclusion that ArKanjo identifies 
+straightforward refactoring strategies. These results support that ArKanjo identifies 
 actionable duplications and that newcomers can successfully contribute patches to the kernel.
@@
 While this research includes successfully merged patches by the main author and student groups,
-a significant portion of the student contributions encountered notable challenges. These difficulties 
+a significant portion of the students' contributions encountered notable challenges. These difficulties 
 often stemmed from maintainer feedback indicating that, although the proposed changes reduced duplication, 
@@
 
-This paper presented ArKanjo, a novel tool designed to detect effectively 
-and facilitate the refactoring of function-level code duplication within the Linux kernel. 
+This paper presented ArKanjo, a novel tool designed to detect effectively and facilitate the refactoring of function-level code duplication within the Linux kernel. 
 Evaluations demonstrated that ArKanjo is highly effective at identifying Type-1 and Type-2 
 clones, which are often indicative of propagated issues such as copied buggy logic. We validated the practical 
-value of the tool through an ethnographic study in which the main author and student 
-groups used ArKanjo to identify actionable duplications. This effort led to 13 patch submissions, 
-including one from the main author and five from student groups that were successfully merged into 
-the kernel
+value of the tool through an ethnographic study in which the main author and student groups used ArKanjo to identify actionable duplications.
+This effort led to 13 patch submissions, including 1 from the main author and 5 from student groups that were successfully merged into the kernel
 \footnote{Reference for the accepted patches:  
 \textit{\href{https://github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}{github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}}.
-}, 
-collectively removing 991 lines of duplicated code. While some refactoring 
-efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 
+}, collectively removing 991 lines of duplicated code. While some refactoring efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 
 for new contributors to make meaningful improvements. The tool provides immediate value for 
@@
 This study was financed, in part, by CAPES (Finance Code 001), the University of São Paulo – USP (Proc.
-22.1.9345.1.2), the São Paulo Research Foundation – FAPESP and the São Paulo State Data Analysis System Foundation – SEADE (Proc.
+22.1.9345.1.2) and the São Paulo Research Foundation – FAPESP with the São Paulo State Data Analysis System Foundation – SEADE (Proc.
 2023/18026-8), Brazil.

### 2025-07-07 conteudo/methods/methods.tex
Add todo note about a typo
update with the papers updates and revisions

--- 2025-06-12 conteudo/methods/methods.tex
@@
 
-Given the tool evaluations demonstrating the tool's capabilities of finding duplications, 
-we proceeded to understand if the duplications found can be mitigated and the kernel 
-subsystem's maintainers' opinions about the duplications and their respective mitigations. 
-To accomplish this objective, we proposed a participant-observation experiment, 
-where we, people who have never contributed to the Linux kernel, act as contributors to 
-the AMD Display driver and collect artifacts of our experience in the process.
+To investigate whether we can use the ArKanjo tool to find ways to contribute to the 
+Linux kernel, we conducted a participant observation experiment in the second semester 
+of 2024, where we acted as first-time contributors to the AMD Display driver, collecting 
+artifacts of our experience in the process.
 
-We executed the tool in the AMD Display driver codebase. 
-We manually analyzed the biggest duplications regarding the number of lines and raised them 
-to two duplicated function pairs that we judged interesting to try to mitigate. We have 
-chosen to approach the biggest duplications as proof of concept so that we can make 
-significant impacts. The two duplicated function pairs chosen are:
+In the AMD Display driver codebase, we executed the ArKanjo tool. 
+We manually analyzed the biggest duplications regarding the number of lines and found 
+two duplicated function pairs that we judged interesting to try to mitigate. We 
+have chosen to approach the biggest duplications as proof of concept so that we 
+can make significant impacts.
 
-\begin{itemize}
-\item \textbf{Duplicated function pair 1}
-\begin{itemize}
-\item Function named \textit{offset\_to\_id} in file \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c}.
-\item Function named \textit{offset\_to\_id} in file \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
-\end{itemize}
-\item \textbf{Duplicated function pair 2}
-%\begin{itemize}
-\item Function named \textit{phy\_id\_to\_atom} in file \textit{dc/bios/dce110/command\_table\_helper\_dce110.c}.
-\item Function named \textit{phy\_id\_to\_atom} in file \textit{dc/bios/dce60/command\_table\_helper\_dce60.c}.
-%\end{itemize}
-\end{itemize}
+The first duplicated function pair chosen is function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
+The second duplicated function pair is function \textit{phy\_id\_to\_atom} in file
+\textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
+Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
+\textit{phy\_id\_to\_atom} will be referred to as mitigation 2.
 
@@
 function pairs. 
-After we approached the duplications with the systematic approach, for the duplicated 
+After we approached the duplications with the systematic strategy, for the duplicated 
 function pairs that had successfully mitigated the duplication, we sent patches to the 
@@
 In C programming, communication between source files is achieved by creating header files that specify 
-libraries \citep{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
+libraries~\cite{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
 by consolidating duplicated code into a single library, which would replace instances of 
@@
 
-For each duplicated function pair, for each function in the pair, we used the \textit{function information} functionality 
-(described in Section \ref{subsec:functioncommand}) to locate all duplications of the function in the codebase, 
-as it may have been duplicated more times than the two occurrences contained in the duplicated function pair. 
+For each duplicated function pair, for each function in the pair, we used the \texttt{anonymous-tool} to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair.
 This strategy resulted in a collection of duplicated functions and corresponding code files.
@@
 
-A core dynamic of the Free Software Development course at the University of São Paulo is to 
+A core dynamic of the course (concurrently for undergraduate and postgraduate studies) is to 
 have students make a practical contribution to a free software project. In the first semester of 
 2025, the professor teaching the course, who is also the advisor of this work, identified an 
-opportunity to leverage the ArKanjo tool within this existing pedagogical structure. As a 
-contribution option, the professor proposed that students mitigate code duplications found in the 
-Linux kernel. This initiative was driven by the professor as a practical course activity, and 
-the author of this thesis was not involved in using the tool to find duplications nor in developing 
-the approaches to mitigate them. The tasks were carried out entirely by the course staff 
-and the students.
+opportunity to leverage the ArKanjo tool within this existing pedagogical structure. 
+As a contribution option, the professor proposed that students mitigate code duplications found 
+in the Linux kernel. The professor drove this initiative as a practical course activity, 
+and the author of this thesis was not involved in using the tool to identify duplications or 
+in developing approaches to mitigate them. The tasks were carried out entirely by the course 
+staff and the students.
 
 The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of
-one to three. To assist students in this task, the professor and teaching assistants prepare multiple 
+two or three members. To assist students in this task, the professor and teaching assistants prepare multiple
 alternatives for contributing to free software projects, including simpler contributions such as
 updating documentation or fixing code style issues. In this context, two new options related to the
-ArKanjo tool were proposed alongside the others.
+\texttt{anonymous-tool} were proposed alongside the others.
 
-For the first ArKanjo-related alternative, a teaching assistant ran the tool on the IIO subsystem, 
-creating a list of duplicated function pairs found with a 90\% similarity threshold. The students' 
-task was to refactor the code to mitigate these specific duplications and then navigate the process 
-of submitting their patches to the IIO maintainers for review.
+For the first ArKanjo-related alternative, a teaching assistant ran the tool on the IIO subsystem twice, 
+creating two lists of duplicated function pairs, one with a 100\% similarity threshold and the other 
+with a 90\% similarity threshold. Then, the following process happened to both lists:
+
+\begin{enumerate}
+    \item The teaching assistant filtered the lists to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different devices may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
+    \item The teacher assistant removed duplications that were too short and would not represent a significant change for a patch application;
+    \item The teacher assistant manually ranked code duplications and marked some duplications with insights based on personal opinion and perceptions over the remaining entries.
+This was done to guide students so that they could approach the technical debt with a greater chance of having their contribution merged into the code.
+\end{enumerate}
+
+The students' task was to select a recommended entry from the list, refactor the code to eliminate 
+these specific duplications, and then submit their patches to the IIO maintainers for review.
 
@@
 observation. In this model, the students were responsible for the entire workflow: executing the
-ArKanjo tool on the AMD display driver, independently analyzing the results to identify a duplication 
+ArKanjo on the AMD display driver, independently analyzing the results to identify a duplication
 to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.
 
-Of the students who chose to work on the ArKanjo-related tasks, 25 (20 undergraduate and 5 graduate) 
-formed 11 groups to pursue the first alternative. One graduate student opted for the second alternative, 
-forming a single group. Notably, none of these students had prior experience contributing to the 
-Linux kernel, making this their first attempt to submit a patch as newcomers.
+Of the students who chose to work on the \texttt{anonymous-tool}-related tasks, 
+23 (16 undergraduate and 7 graduate)
+formed 11 groups to pursue the first alternative. One graduate student also opted for the second 
+alternative. Notably, none of these students had prior experience contributing to the Linux kernel, 
+making this their first attempt to submit a patch as newcomers to the project.
 
-The groups were requested to document their experience and approaches to refactoring the 
-duplications and the experience of submitting a patch to the Linux kernel in blogs. We analyzed 
-these blogs to understand the patterns of refactoring used for mitigating the duplications, the 
-experience of sending the patches, and the opinions of the subsystem maintainers on the mitigations.
-
-The teaching assistant and the student who executed the ArKanjo tool to explore duplications were 
-asked to answer some questions related to their experience using the tool and what could be improved. 
-The survey with the users' answers can be found in Appendix \ref{app:survey}.
-
-In the second stage of the Free Software Development course, we propose that students contribute 
-directly to the ArKanjo tool development without help from the tool creator to verify whether the 
-tool can be free software to which multiple people can contribute. The students approached the tool’s 
-issue, and to our knowledge, they pointed to us in the survey feedback. Three students forming two 
-groups worked in direct contributions.
+The groups were requested to document their experiences and approaches to refactoring the duplications,
+as well as their experiences submitting a patch to the Linux kernel, in blogs. We analyzed these 
+blogs to understand the refactoring patterns used to mitigate duplications, the experience of 
+sending patches, and the opinions of subsystem maintainers on the mitigations.

### 2025-07-07 conteudo/results/participant.tex
update with the papers updates and revisions

@@
 As presented in section \ref{sec:meteth},
-we opted to apply the systematic approach in the functions
-\textit{offset\_to\_id} and \textit{phy\_id\_to\_atom} in the folders 
-\textit{dc/gpio} and \text{dc/bios}, respectively.
+we opted to apply the systematic approach in the mitigations.
 
-Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
-\textit{phy\_id\_to\_atom} will be referred to as mitigation 2. 
 Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
@@
 In the systematic approach, we found that the \textit{offset\_to\_id} function
-exists in code 10 files, enclosing multiple GPU architectures. The duplicated 
+exists in 10 code files, enclosing multiple GPU architectures. The duplicated
 functions are not exactly equal, as some minimal specific logic is applied to 
@@
 
-We reached a point with a scratch on how to mitigate the duplications found, 
-but we could not complete the refactoring on a state capable of submitting as 
-a patch to the driver. We found that the configuration files on the AMD Display 
+We reached a point where we had a scratch on how to mitigate the duplications found, 
+but we could not complete the refactoring to a state capable of submitting as a 
+patch to the driver. We found that the configuration files on the AMD Display 
 driver are designed to be imported at compilation time using \textit{\#define} macros. 
-The configuration files design choice makes the refactoring of duplications on 
-generic approaches trickier, as refactoring code that depends on these files 
-requires significant refactoring on the configuration files design and deep 
-knowledge of the codebase, which we as first-timer contributors, do not have. 
-Thus, we opted not to continue investigating this mitigation.
+The configuration files' design choice makes refactoring duplications on generic 
+approaches trickier, as refactoring code that depends on these files requires 
+significant refactoring of the configuration files' design and deep knowledge 
+of the codebase, which we, as first-time contributors, do not have. Thus, we opted 
+not to continue investigating this mitigation.
 
@@
 
-Using the systematic approach, we found that the \textit{phy\_id\_to\_atom}
-function exists in five code files, and two other functions are duplicated 
+Using the systematic approach, we found that the \textit{phy\_id\_to\_atom} function 
+exists in five code files, and two other functions are duplicated
 in those five code files. All functions across the files were exactly equal,
 so there was no need to apply the refactoring methods presented in the literature.
-The refactoring work was resumed to create a generic library and fix the compilation targets.
+The refactoring was resumed to create a generic library and fix the compilation targets.
 
@@
 patch to contribute to the AMD Display driver code quality. Thus, we moved to send
-the refactoring to the driver’s maintainers as a patch, documenting our findings and experiences 
+the refactoring to the driver maintainers as a patch, documenting our findings and experiences
 on the process.
@@
 
-The experience of submitting a patch was not as we expected. The process took 7 months 
-and 16 days, which is significantly slower than expected. We initially submitted the 
-patch on August 9th, 2024, but we did not receive replies. Thus, we had to resend it 
-on October 9th, 2024. We received an initial response on November 3rd, 2024, asking for 
-minimal changes; we submitted a second version with the changes request on November 16th, 2024. 
-We did not receive replies at first and need to resend the patch on February 7th, 2025. 
-We received a reply on February 11th, 2025, asking for additional changes. We sent a third 
-version on February 24th, 2025, and the patch was accepted and integrated into the kernel 
-codebase on February 25th, 2025. Figure \ref{fig:timeline} illustrates the timeline of 
-the patch submission.
+The experience of submitting a patch was not as we expected. The process took 7 
+months and 16 days, significantly slower than expected. We initially submitted 
+the patch on August 9, 2024, but we did not receive replies. Thus, we had to 
+resend it on October 9, 2024. We received an initial response on November 3, 2024, 
+requesting minimal changes. We submitted a second version with the requested 
+changes on November 16, 2024. We did not receive replies initially and will 
+need to resend the patch on February 7, 2025. We received a reply on 
+February 11, 2025, asking for additional changes. We sent a third version on 
+February 24, 2025, and the patch was accepted and integrated into the kernel 
+codebase on February 25, 2025. Figure \ref{fig:timeline} illustrates the 
+timeline of the patch submission.
 
-We discussed with our point of contact at the AMD Display driver to understand the reason 
-for the slow process of submitting a patch. We learned that some of the code in the driver 
-is shared across all operating systems that support the implemented GPUs. This fact creates 
-a complex process within AMD to format and submit changes across the supported systems while 
-simultaneously implementing measures to mitigate errors. Additionally, we comprehended that 
-AMD Display driver developers can not view duplicated code negatively, as we understand it 
-is a good practice in software engineering. One example shared with us is that duplicated 
-code enhances the independence of GPU driver code, allowing developers to make changes to a 
-specific GPU without needing to test compatibility with others. 
-This approach helps save significant time and effort.
+After the initial reply from the maintainers, we observed that the review process 
+became more responsive. We hypothesize that once a patch receives initial attention, 
+it is more closely tracked by maintainers. Another contributing factor to the initial 
+delay could be the maintainers' workload at the time of the first submission.
 
-Regarding the patch changes requested for the second version, the maintainers only asked 
-for minor changes to align with code style, license correct use, and best practices that 
-were initially unknown to use. For the change requested on the third version, the maintainers 
-asked us to move the functions created in the generic library to an existing file instead of 
-creating a new one. More details of the patch can be found in Appendix \ref{app:feedback}, 
-and the patch submitted in the Linux codebase can be seen here:
-\url{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}.
+We discussed the Display driver with our point of contact at AMD to understand the 
+reason for the slow patch submission process. We learned that some of the code in 
+the driver is shared across all operating systems that support the implemented GPUs. 
+This fact creates a complex process within AMD to format and submit changes across 
+the supported systems while simultaneously implementing measures to mitigate errors. 
+Additionally, we comprehended that AMD Display driver developers can not view 
+duplicated code negatively, as we understand it is a good practice in software engineering. 
+One example shared with us is that duplicated code enhances the independence of 
+GPU driver code, allowing developers to make changes to a specific GPU without 
+needing to test compatibility with others. This approach helps save a 
+significant amount of time and effort.
+
+Regarding the patch changes requested for the second version, the maintainers 
+only asked for minor changes to align with code style, correct license use, 
+and best practices that were initially unknown. For the change requested on 
+the third version, the maintainers asked us to move the functions created in 
+the generic library to an existing file instead of creating a new one.
+\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.
 

### 2025-07-07 conteudo/results/students.tex
update with the papers updates and revisions

@@
 maintainers and documenting their experiences and learnings.
-
-For the students who worked in the IIO subsystem, the teaching assistant gave them duplicated 
-function pairs 
-found by the tool, and their responsibility was to think of how to mitigate the duplication 
-and send the patch to the maintainers. The student that worked in the AMD Display driver 
-have the additional work to use the tool to find the duplications.
 
@@
 
-Table \ref{tab:stu} summarizes students' groups' mitigation approaches. Seven groups applied 
-the parameterized refactor method, 
-and two applied the extractor method to their approaches to mitigate the duplications. 
-The other three groups do not use any refactor methods presented in the literature.  
-Five groups got their patches accepted and merged in Linux; four groups got their patches rejected, 
-and one group is still working on the patch submission, being in the fifth version of the patch at 
-the time this document was written. Two groups decided not to continue sending the patches after 
-they made initial versions.
+Table \ref{tab:stu} summarizes the student groups' mitigation approaches. It contains 
+the information about the chosen subsystem and the similarity threshold for groups that 
+opted for the first approach. Each entry also has the patch status, refactoring method, 
+and code diff corresponding to the final patch versions before it was dropped or accepted 
+into the code base. The code diff for a patchset corresponds to the sum of added and removed 
+lines of each of its patches.
 
-Groups one and four have worked in similar duplications, where two functions return a 
-boolean mirroring the results. That means when the first function returns true, the second 
-function returns false, and when the first function returns false, the second function 
-returns false. Group one had a simple approach, and the maintainers approved the changes quickly. 
-Group four attempted to mitigate the duplication with macros, a method found in the code file 
-they were working on. The maintainers disagreed with the proposed approach, arguing that the 
-proposed changes decrease the code readability, and asked for an approach similar to group one. 
-Group four submitted the patch after coding the approach proposed by the maintainers.
+Seven groups applied the parameterized refactor method, two groups used the extractor method, and one group applied the inline method to their approaches to mitigate their duplications.
+Four patches do not contain a refactoring technique defined in the literature (undefined).
+Six groups had their patches accepted and fully merged into Linux; four groups had their patches dropped; one group has its patchset partially applied and is still working on patch submission, currently in the sixth version of the patch when this document was written, but we can also consider this case as accepted.
 
-Groups three, six, and ten created a struct to encapsulate the parameters and mitigate duplication 
-when applying the parameterized method. Group three sent the mitigation to the maintainers and 
-approved it without significant comments. In contrast, the patch from group ten was rejected by 
-the same maintainer who approved the patch from group three. The maintainer justification to reject 
-the patch is that group ten refactored code files that has simple context, therefore adding a layer 
-of abstraction to remove a duplication deacrease the code readability without clear gains on 
-the codebase quality. Figure \ref{fig:group_10_reply} shows the maintainers's feedback for group ten.
+Groups 1 and 4 have worked in similar duplications, where two functions are involved: one to determine if a register is volatile and the other to determine if it is writable.
+These are two functions that return opposite Boolean results.
+Group 1 initially addressed the problem using the extract method, while Group 2 used macros.
+After feedback from maintainers, both solutions converged to keep one of the functions while changing the content of the other to return a negated query of the former function.
+However, by the end of the feedback cycle, maintainers considered the patch from Group 1 irrelevant, while the patch from Group 4 was accepted.
+The reason is that Group 1 has chosen to apply the refactoring to a device that contains a register that is both read-only and non-volatile.
+Thus, the technical debt selected by Group 1 was not merely a duplication but an incorrect implementation of the method.
 
-\begin{figure}
-\includegraphics[scale=0.8]{group_10_reply}
-\caption[Maintainers' Feedback on Group Ten's Patch]{
-Maintainers' Feedback on Group Ten's Patch
-\footnote{Image obtained from Linux Lore, which is publicly provided at:
-\url{https://lore.kernel.org/linux-iio/20250428034222.318898-1-gabrielgeraldinosouza@gmail.com/T/\#u}}
-}
-\label{fig:group_10_reply}
-\end{figure}
+Groups 3, 6, and 10 created a struct to encapsulate the parameters and mitigate duplication when applying the parameterized method.
+The patches from Groups 6 and 10 were rejected by the same maintainer who approved the patch from Group 3.
+The maintainer's justification for rejecting the patch from Group 6 is that it proposes to fix duplication while adding more lines than removing, and is not convincing enough in other aspects, such as readability.
+The reason for rejecting the patch from Group 10 is that they refactored code files with simple context; therefore, adding a layer of abstraction to remove duplication decreases code readability without clear gains in codebase quality.
+The maintainers' feedback for Group 10 was:
+\begin{quote}
 
-Groups seven and eight used the extract method to approach the mitigations. Group eight sent the 
-mitigations to the maintainers, and the code changes were approved, with minor requested changes 
-in code style and smaller errors fixed. For group seven mitigation, the maintainers reject their 
-patch because the code changes add layers of abstractions to the code, making it more complex and 
-challenging to read. The maintainers proposed that group seven remove the duplicated functions and 
-insert the duplicated code multiple times where the functions were being called. This suggestion 
-may be bad practice for code quality, as it adds more complexity and responsibility to a bigger 
-function; it is viewed as the best choice for removing layers of abstraction and making the 
-code easier to read. Figure \ref{fig:group_7_reply} shows the maintainers's feedback for group seven.
+\textit{``A few comments inline, but overall I'm not sure the code reduction is sufficient to cover the resulting loss of readability.
+Sometimes a switch is simply clearer than a partial look-up table.''}
 
-\begin{figure}
-\includegraphics[scale=0.8]{group_7_reply}
-\caption[Maintainers' Feedback on Group Seven's Patch]{
-Maintainers' Feedback on Group Seven's Patch
-\footnote{Image obtained from Linux Lore, which is publicly provided at:
-\url{https://lore.kernel.org/linux-iio/20250428132551.176788-1-bellacaselli20@gmail.com/}}
-}
-\label{fig:group_7_reply}
-\end{figure}
+\textit{``So nice idea, but I'm not seeing it as being a good move for this particular code.
+I'd be slightly interested to see the optimized output of the two approaches, but this is far from a high-performance path, so we care a lot more about readability here.''}
+\end{quote}
 
-Groups nine and eleven faced similar issues to those in Mitigation 1 on our participant observation 
-experiment, where configuration files designed as macros complicated the approach to the mitigations. 
-Group nine applied the parameterized method to not need to know the configurations in the 
-generic function, which is reasonable as the duplication approach only depends on one configuration 
-value. Group eleven worked with multiple configuration values, and to refactor the duplication, 
-they proposed to create a bigger macro that encapsulates the other macros, which is a clever 
-solution but adds abstraction and complexity to macros, which makes the code harder to debug and read.
-We do not get the feedback of the maintainers on the group eleven approach to know their opinions.
+Groups 7 and 8 used the extract method to approach the mitigations.
+Group 8 sent the mitigations to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed.
+For Group 7 mitigation, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function.
+After the feedback cycle, maintainers accepted the patch in version 3.
 
-Group twelve is the student who executed the tool and passed through the process of finding 
-duplications. The student found a duplication in the driver that happened across eighteen code files, 
-creating a patch to remove four hundred lines of duplicated code. The student sent the mitigations 
-to the maintainers, and their feedback pointed to fixing the warning in the code without arguing 
-about removing the duplications.
+Group 9 initially approached the selected duplication with a parameterized method, but after entering the feedback cycle, it was revealed that the problem represented a much more complex technical debt.
+The patch is still under development, currently at version 6, and has been separated into a 4-patch patch set, where half of them have been applied, and the remaining entries are receiving suggestions.
+Fortunately, maintainers are receptive and provide constant feedback; the contribution is expected to be fully merged if the students do not abandon the current work.
 
-Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that 
-people without previous experience in Linux kernel development could approach duplications found 
-by the ArKanjo tool and become contributors to the kernel. We observed that not all duplications 
-are viewed as code with bad quality, with the maintainers analyzing the trade-off between the 
-purpose of the duplicated code, code readability, and the levels of abstraction added to 
-address the duplications.
+Group 11 is the student who executed the tool and completed the process of identifying duplications.
+The student identified a duplication in the driver that occurred across eighteen code files, creating a patch to remove approximately four hundred lines of duplicated code.
+The student sent the mitigations to the maintainers, and their feedback pointed to fixing the warning in the code without arguing about removing the duplications.
+After addressing the issue, the patch was applied.
+This student also contributed to IIO, tackling a code duplication with inline method refactoring.
+This one, however, was received with more skepticism by the maintainer, and the student dropped the contribution because he did not think the feedback instructions were clear enough.
+
+Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach duplications found by the \texttt{anonymous-tool} and become contributors to the kernel.
+We observed that not all duplications
+are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code, code readability, and the levels of abstraction added to address the duplications.
+
+\subsection{Statistics of the contributions}
+
+\todo[inline]{Is it ok to add it here in the text? It was Spessoto that writed this section entirely and I reviewed it}
+
+The eleven groups, which have submitted 11 contributions to the IIO subsystem, have collectively 
+requested the insertion of 379 lines and the deletion of 605 lines in the subsystem, 
+representing an average of 34.45 lines added and 55 lines removed per contribution.
+
+Considering the sum of fully merged contributions, we have a code diff value of +51/-143, 
+with an average of +12.75/-35.75. This ratio is better than dropped contributions, which 
+have a total of +179/-219 with an average of +29.83/-36.5.
+
+It is also possible to notice a difference based on the similarity threshold used to 
+detect duplications. The proposed patches to address 100\%-similarity duplications 
+summed up to +46/-90 (average of +9.2/-18). For a 90\% value, the result was a total 
+of +333/-515 with an average of +55.5/-85.83. The 90\% value enabled ArKanjo to identify 
+less strict duplication cases, and, therefore, code segments with a greater extension, 
+compared to the 100\% similarity entries, which were shorter segments, causing a 
+difference in the average number of modified lines.
+
+Finally, the AMD contribution from Group 11, which involves the complete use of the ArKanjo toolchain, 
+got a completely different outcome. With a total of 90 lines inserted and 489 lines 
+removed, there are some points to be considered for the success and high modification 
+extension of the contribution:
+
+\begin{itemize}
+    \item The AMD Display subsystem has some self-reported problems regarding overall technical debt;
+    \item The AMD Display subsystem has a more centralized code structure and maintainer hierarchy, enabling contributors to tackle code duplications across a broader scope and multiple files, which happened to Group 11's contribution;
+    \item The complete control over the process of finding code duplication gave Group 11 the opportunity to identify and reason better over the duplications found by \texttt{anonymous-tool}, enabling a more strategic approach to searching for promising patch possibilities.
+\end{itemize}
 

### 2025-07-07 conteudo/tool/architecture.tex
update with the papers updates and revisions

@@
 
-The Function Parser receives the input codebase and transforms it into the temporary codebase.
-The Function Parser iterates through every source code file from the programming language it supports and uses a specific programming language function extractor to extract every function in the file.
-For each function extracted, two new source code files and a metadata file are created in the temporary codebase, represented by the pair \textbf{<file name, function name>}, the source code file of the function, and the proper function.
-The first new source code file contains the function's body from the function it represents, while the other new source code file contains the function's signature.
-The metadata file contains additional relevant information about the function, such as the function name, the line where the function signature starts in the source code file, and the line where the function's body ends in the source code file.
-The programming languages supported at the moment are C and Java.
-Figure \ref{fig:transform} illustrates an example of a function transformation into the temporary codebase.
+The Function Parser receives the input codebase and transforms it into the temporary
+codebase. It iterates through every source code file written in a supported programming 
+language and uses a language-specific extractor to isolate each function. For each extracted 
+function, two new source code files and a metadata file are created in the temporary codebase, 
+represented by the pair \textbf{$<$file name, function name$>$}, the source code of the function, 
+and the function metadata (in practice, this is a duplication of the original codebase, 
+structured as a parsed representation). One of the new files contains the function body, and the other 
+contains the function signature. The metadata file includes additional relevant information,
+such as the function name, the line where the signature starts, and the line where the function 
+body ends. Currently, the supported programming languages are C and Java, although Java support is limited.
 

### 2025-07-13 conteudo/resumoabstract.tex
fix typo on abstract

@@
 analisa as duplicações identificadas para extrair padrões e métodos de refatoração, oferecendo 
-orientação a futuros contribuidores. Avaliamos o Arkanjo em casos de duplicação do mundo 
+orientação a futuros contribuidores. Avaliamos o ArKanjo em casos de duplicação do mundo 
 real em versões recentes do kernel, demonstrando sua eficácia na identificação de clones 
@@
 Second, it analyzes the identified duplications to extract patterns 
-and refactoring methods, offering guidance to future contributors. We evaluate Arkanjo 
-against realworld duplication cases in recent kernel versions, demonstrating
+and refactoring methods, offering guidance to future contributors. We evaluate ArKanjo 
+against real world duplication cases in recent kernel versions, demonstrating
 its effectiveness in identifying problematic clones that generic

### 2025-07-13 conteudo/apendices/students-artifacts.tex
some work from 04-06
update students apendix

@@
 
+\textbf{Accepted PATCH (v1):} \url{https://lore.kernel.org/linux-iio/20250424002144.23260-1-beatrizvianacosta16@gmail.com/}
+
 \paragraph{Group 4}
@@
 
+\textbf{Accepted PATCH (v3):} \url{https://lore.kernel.org/linux-iio/20250426163311.546500bc@jic23-huawei/}
+
 \paragraph{Group 5}
@@
 \textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250426153307.59689da0@jic23-huawei/T/#t}
+
+\textbf{Accepted PATCH (v1):} \url{https://lore.kernel.org/linux-iio/20250426153307.59689da0@jic23-huawei/T/#t}
 
@@
 
+\textbf{Accepted PATCH (v3):} \url{https://lore.kernel.org/linux-iio/20250531160807.39e54096@jic23-huawei/}
+
 \paragraph{Group 8}
@@
 
-\textbf{PATCH in Linux lore:} Students have not provided the link.
+\textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/all/20250422191114.32832-1-nattanferreira58@gmail.com/#t}
+
+\textbf{Accepted PATCH (v4):} \url{https://lore.kernel.org/all/20250614124401.32b17eda@jic23-huawei/}
 
@@
 \textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/linux-iio/20250423170526.15143-1-andrew.lopes@alumni.usp.br/}
+
+\textbf{Accepted PATCH (partially accepted on v6):} \url{https://lore.kernel.org/linux-iio/20250621183843.2f8bcb48@jic23-huawei/}
 
@@
 
-\textbf{Blog 1:} \url{https://o-carneiro.github.io/}
-
-\textbf{Blog 2:} \url{https://fernandoglima.github.io/}
-
-\textbf{Blog 3:} \url{https://eijiuchiyama.github.io/softwarelivre/}
-
-\textbf{PATCH in Linux lore:} Students have not provided the link.
-
-\paragraph{Group 12}
-
-\
-
 \textbf{Blog 1:} \url{https://saguileran.github.io/blog/2025/free-software-development/}
@@
 \textbf{PATCH in Linux lore:} \url{https://lore.kernel.org/amd-gfx/f78f68ba-80e0-47ca-97a2-37a15d076e68@amd.com/T/#t}
+
+\textbf{Accepted PATCH (v2):} \url{https://lore.kernel.org/amd-gfx/CADAW+GDN-4Lwd87h793csWF6r5cgqpgrpEPzRyDV9+rfkNPmTA@mail.gmail.com/}

### 2025-07-13 conteudo/apendices/gray.tex
update links on apendix to fix mergin issues

@@
 
-Access from \url{https://pmd.github.io/pmd/pmd_userdocs_cpd}.
+Access from \textit{\seqsplit{pmd.github.io/pmd/pmd\_userdocs\_cpd}}.
 
@@
 Access from 
-\url{https://cppdepend.com/blog/tracking-the-hidden-duplicate-code-in-a-c-code-base/}.
+\textit{\seqsplit{cppdepend.com/blog/tracking-the-hidden-duplicate-code-in-a-c-code-base/}}.
 
@@
 Access from
-\url{https://github.com/dlidstrom/Duplo}.
+\textit{\seqsplit{github.com/dlidstrom/Duplo}}.
 
@@
 Access from
-\url{https://github.com/duploq/duploq}
+\textit{\seqsplit{github.com/duploq/duploq}}.
 
@@
 Access from
-\url{https://simian.quandarypeak.com/}.
+\textit{\seqsplit{simian.quandarypeak.com/}}.
 
@@
 Access from
-\url{https://codeclimate.com/}.
+\textit{\seqsplit{codeclimate.com/}}.
 
@@
 Access from
-\url{https://www.researchgate.net/publication/2866183_Visual_Detection_of_Duplicated_Code}.
+\textit{\seqsplit{www.researchgate.net/publication/2866183\_Visual\_Detection\_of\_Duplicated\_Code}}.
 
@@
 Access from
-\url{https://github.com/platisd/duplicate-code-detection-tool}
+\textit{\seqsplit{github.com/platisd/duplicate-code-detection-tool}}.

### 2025-07-14 conteudo/background/refactor.tex
add sowrk 11/05
add paragraph about contrib and remove some todo
add inline method to refactoring

@@
 \section{Refactoring}
-
-\todo[inline]{Reconsider if I need this function anymore.}
 
@@
 \ref{fig:extract} demonstrates an example of the Extract Method.
+
+\paragraph{Inline Method.}
+
+The inline method involves replacing a function call with the entire body of 
+the function, which is copied to where the function is called. One of the 
+advantages of this method is reducing the number of unnecessary methods and 
+abstraction in the code, making the code more straightforward
+\citep{refactorbook}.

### 2025-07-24 conteudo/background/codeduplication.tex
another grammar typos fix

@@
 Code Duplication Detection is a field in computer science that has attracted
-researchers' attention since 1988 \citep{firstman}. The presence of code
+researchers' attention since 1988 \citep{firstman}. Code
 duplication is considered a harmful artifact in software, as it negatively
@@
 introducing bugs \citep{harmone}. In most cases, code duplication tends to make
-software more unstable compared to its non-duplicated counterpart
+software more unstable than its non-duplicated counterpart
 \citep{harmtwo}.
@@
 
-The classification of two code artifacts as duplicates raises the question of
+Classifying two code artifacts as duplicates raises the question of
 what constitutes duplication. To address this, the literature categorizes code
@@
 Recall, and F-Score can be calculated \citep{recall}. In this work, we focus on
-the Recall metric, as it is commonly reported in literature for comparing code
+the Recall metric, which is commonly reported in literature for comparing code
 clone detection approaches. Recall is defined as follows \citep{recall}:

### 2025-07-24 conteudo/background/linux.tex
add lines from IIO
Update typo on IIO metrics
another grammar typos fix

@@
 the research, such as the Linux Kernel, refactoring, types of code duplication, 
-code duplication detection methods presented in the literature and ways to 
+code duplication detection methods presented in the literature, and ways to 
 evaluate these methods. This foundation will make it easier to grasp the 
@@
 Torvalds in 1992 as free software. It began gaining popularity in the late
-1990s \citep{linuxbook}. Contrary to popular belief, Linux itself is not a
+1990s \citep{linuxbook}. Contrary to popular belief, Linux is not a
 complete operating system, as it lacks components like filesystem utilities and
-windowing systems. The so-called ``Linux distributions'' that are commonly
-recognized by the public are actually GNU/Linux operating systems, which
+windowing systems. The so-called ``Linux distributions'' commonly
+recognized by the public are GNU/Linux operating systems, which
 include various essential applications alongside the Linux kernel,
@@
 Mailing lists are the primary means of communication within the Linux
-community. Members of the community use them to submit and review patches,
+community. Community members use them to submit and review patches,
 discuss Linux-related topics, and debate potential new features. Given the
@@
 object of study as we aim to understand and mitigate code duplication in the
-Linux kernel. The AMD Display driver subsystem contains more then 391000 lines 
-of code and more then 1089 code files
+Linux kernel. The AMD Display driver subsystem contains more than 391000 lines 
+of code and more than 1089 code files
 \footnote{ 
@@
 or digital-to-analog conversion (DAC) or both''\citep{iiodoc}. The IIO subsystem 
-contains more than X lines of code and more than Y code files
+contains more than 281772 lines of code and more than 755 code files
 \footnote{
@@
 contact within the IIO subsystem to facilitate this work.
-Like other subsystems, discussions about the IIO subsytem are conducted on its 
+Like other subsystems, discussions about the IIO subsystem are conducted on its 
 designated mailing list, which can be found here: 
@@
 methods on the tool proposed in this work, although it is not explored as the text 
-similarity method for code duplication detection.
+similarity method chosen for code duplication detection.

### 2025-07-24 conteudo/background/nlpsimilarity.tex
another grammar typos fix

@@
 vector embeddings of texts and applying a distance function to these embeddings
-to measure the similarity between them. The literature proposes multiple vector
+to measure their similarity. The literature proposes multiple vector
 embedding methods, such as bag-of-words, TF-IDF, LSI, and Word2Vec
@@
 
-where $A$ and $B$ are the vector embeddings of the two texts. An advantage of
+Where $A$ and $B$ are the vector embeddings of the two texts. An advantage of
 cosine similarity is that $SC(A,B)$ produces a number between $0$ and $1$,
@@
 collection of 10,000 sentences required about 65 hours with BERT
-\citep{bertsimilar}. Since we aim to compare a similar magnitude of pairs in
+\citep{bertsimilar}. Since we aim to compare pairs of similar magnitude in
 our work, we decided not to use LLMs for our vector embedding method, even

### 2025-07-24 conteudo/conclusion/conclusion.tex
another grammar typos fix

@@
 challenges; code duplication persists as a notable issue within its codebase, 
-potentially hindering its evolution and patching processes. 
-Prevailing academic and existing free software approaches to duplication 
-detection often lack the scalability for comprehensive codebase analysis or 
-offer limited functionality unsuitable for the kernel's complexity, 
-generally failing to deliver actionable results or guidance for mitigation 
-within its specific context.
+potentially hindering its evolution and patching processes. Prevailing academic 
+and existing free software approaches to duplication detection often lack the 
+scalability for comprehensive codebase analysis or offer limited functionality 
+unsuitable for the kernel's complexity, generally failing to deliver actionable 
+results or guidance for mitigation within its specific context.
 
-In response to these limitations, this research addressed the challenge of 
-code duplication within the Linux Kernel through the development of ArKanjo, 
-command-line tool for Linux kernel maintenance designed to detect and 
-analyze function-level duplications. Released under the MIT license, 
-the architectural design of ArKanjo, incorporating a Preprocessor and a 
-Query Responder, separates computationally intensive analysis from efficient 
-querying for duplications within large codebases. 
-Validation procedures, employing both the BigCloneBench dataset and empirical 
-analysis conducted on the AMD Display driver, indicated ArKanjo's proficient 
-performance in identifying Type-1 and Type-2 duplications, notably when 
-similarity thresholds are configured between 80\% and 100\%.
+In response to these limitations, this research addressed the challenge of code 
+duplication within the Linux Kernel by developing ArKanjo, a command-line tool 
+for Linux kernel maintenance designed to detect and analyze function-level duplications. 
+Released under the MIT license, the architectural design of ArKanjo, incorporating a 
+Preprocessor and a Query Responder, separates computationally intensive analysis from 
+efficient querying for duplications within large codebases. Validation procedures, 
+employing both the BigCloneBench dataset and empirical analysis conducted on the AMD 
+Display driver, indicated ArKanjo's proficient performance in identifying Type-1 and 
+Type-2 duplications, notably when similarity thresholds are configured between 80\% and 100\%.
 
-The ethnographic studies, which involved participant observation and a project 
-with university students, provided a realistic view of both the opportunities 
-and the considerable challenges in mitigating code duplications found by ArKanjo. 
-The tool was effectively used by first-time contributors to the Linux kernel to 
-identify duplications and prepare patches. While this research included a
-successfully merged patch by the author, and successful submissions from many 
-student groups, a notable portion of student efforts 
-encountered significant hurdles. These difficulties often stemmed from maintainer 
-feedback where proposed changes, though reducing duplication, were rejected or 
-required substantial rework due to concerns about code readability, added abstraction, 
-or the specific context of the code. Technical complexities, such as C macros in 
-configuration files, and lengthy patch review times also contributed to the challenges. 
-Furthermore, the studies confirmed that maintainers do not always view code duplication 
-as inherently negative, especially when it may improve clarity or facilitate 
-independent hardware support.
+The ethnographic studies, which involved participant observation and a project with university 
+students, provided a realistic view of the opportunities and the considerable challenges in 
+mitigating code duplications found by ArKanjo. The tool was effectively used by first-time 
+contributors to the Linux kernel to 
+identify duplications and prepare patches. While this research included a successfully merged 
+patch by the author, and successful submissions from many student groups, a notable portion of 
+student efforts encountered significant hurdles. These difficulties often stemmed from maintainer 
+feedback where proposed changes, though reducing duplication, were rejected or required 
+substantial rework due to concerns about code readability, added abstraction, or the specific 
+context of the code. Technical complexities, such as C macros in configuration files, and lengthy 
+patch review times also contributed to the challenges. Furthermore, the studies confirmed that 
+maintainers do not always view code duplication as inherently harmful, especially when it may 
+improve clarity or facilitate independent hardware support.
 
-In its entirety, this research contributes the ArKanjo tool as an operational 
-method for detecting code duplication, and importantly, provides insights into 
-the process of managing such duplications within the Linux kernel subsystems 
-studied. The results show that while ArKanjo can effectively identify 
-duplicated functions, the path from detection to an accepted mitigation within 
-the kernel involves careful consideration of the specific code, established 
-community practices, and the maintainers' perspectives on code quality and readability.
+This research contributes to the ArKanjo tool as an operational method for detecting code 
+duplication and, importantly, provides insights into the process of managing such duplications 
+within the Linux kernel subsystems studied. The results show that while ArKanjo can effectively 
+identify duplicated functions, the path from detection to an accepted mitigation within the kernel 
+involves careful consideration of the specific code, established community practices, and the 
+maintainers' perspectives on code quality and readability.
 
@@
 
-The work presented in this thesis suggests several directions for future investigation. 
-Enhancements to the ArKanjo tool could be pursued, such as investigating a multi-stage 
-detection strategy. This strategy might employ a computationally lighter method for an 
-initial, broad pass to identify candidate duplications, followed by more 
-resource-intensive, state-of-the-art techniques for refined analysis and improved 
-detection of complex clone types. Such an approach could complement efforts for 
-improved resource management during its preprocessing phase for very large codebases.
-The tool's language support could also be broadened by
-developing parsers for additional programming languages, building upon its adaptable
-architecture. Beyond the tool itself, future research could extend the empirical
-analysis of code duplication and mitigation strategies to a broader range of Linux kernel 
-subsystems or other large-scale C projects, to determine the wider 
-applicability of the observations made.
+The work presented in this thesis suggests several directions for future investigation. Enhancements 
+to the ArKanjo tool could be pursued, such as investigating a multi-stage detection strategy. 
+This strategy might employ a computationally lighter method for an initial, broad pass to identify 
+candidate duplications, followed by more resource-intensive, state-of-the-art techniques for refined 
+analysis and improved detection of complex clone types. Such an approach could complement efforts for 
+improved resource management during its preprocessing phase for large codebases. The tool's language 
+support could be broadened by developing parsers for additional programming languages, building upon 
+its adaptable architecture. Beyond the tool itself, future research could extend the empirical analysis 
+of code duplication and mitigation strategies to a broader range of Linux kernel subsystems or other 
+large-scale C projects, to determine the broader applicability of the observations made.

### 2025-07-24 conteudo/introduction/intro.tex
another grammar typos fix

@@
 \footnote{Data directly extracted from Linux github repository \citep{linuxrepo}.
-Counted  28783390 lines of code and 20699 number of contributors in May 29th, 2025.
-Number of contributors counted as number of unique full name listed as author of a patch.}.
+Counted  28783390 lines of code and 20699 contributors on May 29th, 2025.
+Number of contributors counted as unique full names listed as a patch's author.}.
 Due to the scale of the project, not all contributions adhere to best practices, resulting 
@@
 One such artifact is code duplication, which is the focus of this research. The motivation for 
-this work began with a practical problem rooted in device drivers, which are a major part 
+this work began with a practical problem rooted in device drivers, which are a significant part 
 of the kernel, representing 66\% of the source code\citep{marcelo}. We interacted with the 
@@
 we found that existing tools are not well-suited for identifying duplications in large-scale 
-codebases like the Linux kernel, nor do they offer guidance on how to resolve the duplications 
-they find. Despite searching both formal and grey literature, existing solutions fell short of 
+codebases like the Linux kernel, nor do they offer guidance on resolving the duplications 
+they find. Despite searching formal and grey literature, existing solutions failed to address
 addressing these specific challenges. The list of related tools examined in this exploratory 
@@
 To address the code duplication problem within the Linux kernel, this research employs a 
-multi-method approach designed to bridge the gap between detection and practical mitigation. 
+multi-method approach to bridge the gap between detection and practical mitigation. 
 Our approach, illustrated in Figure \ref{fig:reDesign}, is structured into two interconnected phases that progress 
 from tool development to an in-depth study of its application within the kernel’s development community. 
-The overall objective is to produce not only a functional tool but also actionable insights into 
+The overall objective is to produce a functional tool and actionable insights into 
 the patterns and practices surrounding code duplication and its resolution.
@@
 
-The initial phase of our research focused on the development and validation of ArKanjo, a 
-command-line tool capable of identifying code duplication at the function level.  The 
-tool's accuracy was rigorously validated by evaluating its performance against the 
+The initial phase of our research focused on developing and validating ArKanjo, a
+command-line tool capable of identifying code duplication at the function level. 
+We rigorously validated the tool's accuracy by evaluating its performance against the
 standard BigCloneBench dataset \citep{bigclonebench} and by conducting an empirical analysis
@@
 
-With a validated tool, the research then moved from detection to practice, investigating 
-the real-world challenges of mitigating the identified duplications. This was accomplished 
-through an ethnographic study designed to engage with the kernel community and understand 
-their perspectives on code quality and refactoring.  This study had two components: a 
+With a validated tool, the research moved from detection to practice, investigating the 
+real-world challenges of mitigating the identified duplications. This was accomplished 
+through an ethnographic study to engage with the kernel community and understand their 
+perspectives on code quality and refactoring. This study had two components: a 
 participant-observation experiment where the author submitted refactoring patches to the 

### 2025-07-24 conteudo/methods/methods.tex
another grammar typos fix

@@
 We employ a multi-method research approach, using different methodologies to achieve our 
-goals to approach code duplication in some of the Linux subsystems. To validate if the 
+goals of approaching code duplication in some of the Linux subsystems. To validate if the 
 ArKanjo tool can find code duplications, we applied two independent methods and performed 
@@
 limitations and may be unreliable \citep{bigfail, litreview}. 
-As a reminder, all research approaches were done using the text similarity method 
+As a reminder, we have done all research approaches using the text similarity method 
 as the code duplication detection method of the tool.
@@
 the code duplications found by the tool in the IIO subsystem and the AMD Display driver and 
-pass the process of sending the mitigations as PATCH.
+pass the process of sending the mitigations as a PATCH.
 
@@
 for our context (Java dataset vs. C-based Linux kernel), we complemented the BigCloneBench evaluation with 
-an empirical study of the tool inferred duplications.
+an empirical study of the tool's inferred duplications.
 
@@
 Firstly, we present the participant observation research method explored by the author. 
-Then, we present the methodology of the non-participant observation analyze of patches 
+Then, we present the methodology of the non-participant observation analysis of patches 
 sent by the students.

### 2025-07-24 conteudo/results/eval.tex
another grammar typos fix

--- 2025-05-03 conteudo/results/eval.tex
@@
 
+\section{Tool Results on BigCloneBench}
+
 \input{conteudo/results/table_bigclone}
 
-As described in Section \ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset \citep{bigclonebench}, we define that every pair of functions with a similarity metric equal to or greater than a threshold \( X \) is considered a duplication.
-The pair of code files containing these functions are then labeled as duplications.
-If our tool correctly identifies a code duplication pair, it is counted as correct for the respective code duplication type.
-If it infers a duplication pair incorrectly, it is counted as incorrect for all code duplication types.
+As described in Section \ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset \citep{bigclonebench}, we define 
+that every pair of functions with a similarity metric equal to or greater than a threshold \( X \) is considered a duplication. 
+The pair of code files containing these functions is then labeled as duplications. If our tool correctly 
+identifies a code duplication pair, it is counted as correct for the respective code duplication type. If it infers a 
+duplication pair incorrectly, it is counted as incorrect for all code duplication types.
 Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}.
@@
 there is a notable increase in function pairs that are not duplications within 
-the Linux kernel. For similarity ranges of 79\%-81\% or higher, there is a high 
+the context of the AMD Display driver (the Linux kernel component studied here). 
+For similarity ranges of 79\%-81\% or higher, there is a high 
 percentage of code duplications. At 69\%-71\% similarity, the duplication rate 

### 2025-07-24 conteudo/results/students.tex
remove todo as I need to submit the text
another grammar typos fix

@@
 As presented in Chapter \ref{cha:method}. It was proposed to students, as one of the activities 
-of the Free Software course, that they make the first-time contribution to the Linux kernel to 
+of the Free Software course, they make their first contribution to the Linux kernel to 
 mitigate duplication found by the ArKanjo tool while sending patches to the Linux subsystem 
@@
 
-\todo[inline]{Is it ok to add it here in the text? It was Spessoto that writed this section entirely and I reviewed it}
-
 The eleven groups, which have submitted 11 contributions to the IIO subsystem, have collectively 
@@
 We asked the teaching assistant and the students who used the ArKanjo to answer a survey with 
-some questions related to the tool experience, such as installation, documentation, facility of 
+some questions related to the tool experience, such as installation, documentation, ease of 
 use, new ideas to improve the tool, etc. Appendix \ref{app:survey} has the complete survey 
@@
 Related to the tool's functionalities, both the student assistant and the student reported 
-a missing mechanism to save the tool results on a separate file instead of the terminal output. 
-The tool, by default, outputted with colors, as shown in section \ref{subsec:func}, which complicates 
+a missing mechanism to save the tool results in a separate file instead of the terminal output. 
+The tool, by default, outputs with colors, as shown in section \ref{subsec:func}, which complicates 
 redirecting the terminal output to a file.
@@
 the tool fulfilled their expectations, and had an overall good experience. Finally, the teaching assistant 
-the the student reported features ideas to the tool that may be interesting, which 
-are showed below.
+and the student reported features to the tool that may be interesting, which 
+are shown below.
 
@@
 ArKanjo tool. They also provided critical points that could be improved. The most critical 
-reported issues, such as documentation errors and a lack of support for multiple Linux distributions 
-and different operating systems, have been fixed. Less critical issues remain open points that 
+reported issues have been fixed, such as documentation errors and a lack of support for multiple Linux distributions 
+and different operating systems. Less critical issues remain open points that 
 can be addressed in future work.
@@
 light and dark modes. The colored output presented in the figures in Chapter \ref{cha:tool} was 
-hard coded at first, not considering the user terminal colors, thus giving a bad user experience. 
+hard-coded at first, not considering the user terminal colors, thus giving a bad user experience. 
 The website helps facilitate people's contribution or use of the tool. We did not know how to 

### 2025-07-24 conteudo/resumoabstract.tex
update keywords
another grammar typos fix

@@
 % definidas antes do resumo/abstract. Acrescente quantas forem necessárias.
-\palavraschave{Linux, Duplicação de código, ferramenta}
+\palavraschave{Software livre, kernel Linux, duplicação de código, refatoração, manutenção de software}
 
-\keywords{Linux,Code duplication,Tool}
+\keywords{Free software,Linux kernel, Code duplication, Refactoring, code maintenance}
 
@@
 bug fixes, or handling of security vulnerabilities can impact millions. Maintaining the kernel 
-is an enormous task, involving more than 28 million lines of code and contributions from over 
+is an enormous task involving over 28 million lines of code and contributions from over 
 twenty thousand developers. Due to the scale of the project, not all contributions adhere to best 
@@
 and refactoring methods, offering guidance to future contributors. We evaluate ArKanjo 
-against real world duplication cases in recent kernel versions, demonstrating
+against real-world duplication cases in recent kernel versions, demonstrating
 its effectiveness in identifying problematic clones that generic

### 2025-07-24 conteudo/tool/architecture.tex
another grammar typos fix

@@
 The Preprocessor contains two main components, the Function Breaker and the Code Duplication Finder. 
-The flow of how the Preprocessor works is as follow: Function Parser receives the codebase the user is interested in,
+The flow of how the Preprocessor works is as follows: Function Parser receives the codebase the user is interested in,
 extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files. 
@@
 
-Regarding the programming language's function extractor, we approached this problem so that anyone comfortable with a language can adapt our approach to their specific programming language.
-Usually, code blocks can be easily extracted from source code files, independently if the code blocks are defined by curly brackets or by indentation, and it is possible to infer if they represent a function's body given their depth in the code context and a few lines before the start of the code block.
-We followed this approach to implement the function extractor for the supported languages.
-As an alternative approach, we can build the syntax tree \citep{compiler} of the programming language and use it to extract the functions, which is a more complex task than our approach.
-For the programming languages that we support, we implemented them as follows:
+Regarding the programming language's function extractor, we approached this problem so that anyone 
+comfortable with a language can adapt our approach to their specific programming language. 
+Usually, code blocks can be easily extracted from source code files, independently of whether the 
+code blocks are defined by curly brackets or by indentation, and it is possible to infer if 
+they represent a function's body given their depth in the code context and a few lines before the 
+start of the code block. We followed this approach to implement the function extractor for the 
+supported languages. As an alternative approach, we can build the syntax tree \citep{compiler} 
+of the programming language and use it to extract the functions, which is a more complex task than 
+our approach. For the programming languages that we support, we implemented them as follows:
 
@@
 
-For a source file in the C programming language, we extract the code blocks with depth 0; 
-as per the programming language’s grammar, functions are defined in a global context at depth 0. 
+For a source file in the C programming language, we extract the code blocks with depth 0. 
+As per the programming language’s grammar, functions are defined in a global context at depth 0. 
 To extract the code blocks, we first need to identify the portions of code that are comments or 
 \textit{\#define} directives and mark them as irrelevant. After that, it is simple work to find 
-the curly brackets in depth 0, and with that, we find the body of the functions.
+the curly brackets in depth 0; with that, we find the body of the functions.
 
 To extract the header of the functions, we start at the open curly brackets and go back in the 
-source file considering the programming language grammar and the portions of code that have been 
+source file, considering the programming language grammar and the portions of code 
 marked as irrelevant because of comments and \textit{\#define} directives.
-There are other elements in depth zero that are not functions. We analyze whether the code blocks 
+Other elements in depth zero are not functions; we analyze whether the code blocks 
 are functions at the same time that we extract the header of the code block.
@@
 
-For a source file in Java, we extract the code blocks with depth one and some code lines before 
+We extract the code blocks with depth one and some code lines for a source file in Java before 
 defining them. Then, we check if the code blocks are functions by analyzing whether the first 
@@
 If the similarity is greater than or equal to the minimum similarity threshold 
-(a parameter provided by the user during the Preprocessor), we store this pair of functions along with its 
+(a parameter provided by the user during the Preprocessor). We store this pair of functions and its 
 similarity in the Code Duplication Database.
@@
 We implemented two methods for code duplication detection in the tool, which the user can 
-choose to use. The first method is based on text similarity, and the second is simpler and 
+use. The first method is based on text similarity, and the second is simpler and 
 based on the number of equal lines. The experiments and tests in this research were done 
@@
 
-This method was chosen for its claimed performance, 
+We chose this method for its claimed performance, 
 programming language independence, and the fact that it does not require 
@@
 results for Type-1 and Type-2 code duplication types while performing 
-poorly for the other types. It is possible to change the code duplication 
-detection method to one of the state-of-the-art methods if needed.
+poorly for others. Changing the code duplication
+detection method to one of the state-of-the-art methods is possible if needed.
 
@@
 This code duplication method is considerably simpler and more explainable than the text similarity method, 
-as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
-the \textit{diff} command built-in in the Linux environment \citep{diffcommand}.
+as the metric is the ratio of common lines between the two functions. 
+We use the \textit{diff} command built-in the Linux environment \citep{diffcommand}
+to compute the number of equal lines.
 
@@
 
-The Query Responder step is the component the user executes multiple times to consult information 
+The Query Responder step is the component that the user executes multiple times to consult information 
 about duplicated functions in a codebase processed by the Preprocessor. 
 It utilizes the Temporary Codebase and the Code Duplication Database as inputs to answer user requests, 
-such as querying for duplicated functions above a certain similarity threshold or provide counts of such functions.
+such as querying for duplicated functions above a certain similarity threshold or providing counts of such functions.
 This component was developed to be extensible, allowing other query functionalities to be easily added. 

### 2025-08-02 conteudo/introduction/intro.tex
Intro: minor adjustments

@@
 code and contributions from over twenty thousand developers
-\footnote{Data directly extracted from Linux github repository \citep{linuxrepo}.
+\footnote{Data directly extracted from Linux github repository~\citep{linuxrepo}.
 Counted  28783390 lines of code and 20699 contributors on May 29th, 2025.
-Number of contributors counted as unique full names listed as a patch's author.}.
+Number of contributors counted as unique full names listed as a patch author.}.
 Due to the scale of the project, not all contributions adhere to best practices, resulting 
@@
 this work began with a practical problem rooted in device drivers, which are a significant part 
-of the kernel, representing 66\% of the source code\citep{marcelo}. We interacted with the 
+of the kernel, representing 66\% of the source code~\citep{marcelo}. We interacted with the 
 maintainers of the AMD Display driver, an essential component for the 19\% of personal GPUs 
-manufactured by AMD in 2023\citep{gpumarket}. They shared their challenges with a significant 
-amount of duplicated code that hampers the driver's maintenance. In searching for a solution, 
+manufactured by AMD in 2023~\citep{gpumarket}. They shared their challenges with a significant 
+amount of duplicated code that hampers the driver maintenance. In searching for a solution, 
 we found that existing tools are not well-suited for identifying duplications in large-scale 
@@
 addressing these specific challenges. The list of related tools examined in this exploratory 
-research is available in Appendix \ref{app:gray}.
+research is available in Appendix~\ref{app:gray}.
 
@@
 computational analysis from the fast querying of duplication results, making it effective for 
-large codebases. We validated the tool by comparing its results with the BigCloneBench dataset
-\citep{bigclonebench} and by conducting an empirical analysis on a randomly selected set of 
+large codebases. We validated the tool by comparing its results with the BigCloneBench
+dataset~\citep{bigclonebench} and by conducting an empirical analysis on a randomly selected set of 
 files from the AMD Display driver.
@@
 multi-method approach to bridge the gap between detection and practical mitigation. 
-Our approach, illustrated in Figure \ref{fig:reDesign}, is structured into two interconnected phases that progress 
-from tool development to an in-depth study of its application within the kernel’s development community. 
+Our approach, illustrated in Figure~\ref{fig:reDesign}, is structured into two interconnected phases that progress 
+from tool development to an in-depth study of its application within the kernel development community. 
 The overall objective is to produce a functional tool and actionable insights into 
@@
 command-line tool capable of identifying code duplication at the function level. 
-We rigorously validated the tool's accuracy by evaluating its performance against the
-standard BigCloneBench dataset \citep{bigclonebench} and by conducting an empirical analysis
+We rigorously validated the tool accuracy by evaluating its performance against the
+standard BigCloneBench dataset~\citep{bigclonebench} and by conducting an empirical analysis
 of its results within the AMD Display driver. This foundational work ensured we had a
@@
 understanding of the technical and social factors that influence the mitigation of duplicated 
-code in a large-scale open-source project.
+code in a large-scale FLOSS project.
 
@@
 
-This manuscript consists of five more chapters. Chapter \ref{cha:back} presents
+This manuscript consists of five more chapters. Chapter~\ref{cha:back} presents
 the literature overview for code duplication detection (main definitions,
@@
 explored in this work, 
-and a review of refactoring methods used throughout this research. Chapter
-\ref{cha:tool} presents ArKanjo, our proposed tool to detect code duplication, 
-describing all the main components.  Chapter \ref{cha:method}
-describes the research methods selected to guide our work. Chapter
-\ref{cha:results} shows the results we had through our work, from the research methods to 
-evaluate our tool and the ethnographic studies. Chapter \ref{cha:conclusion} concludes this research.
+and a review of refactoring methods used throughout this research.
+Chapter~\ref{cha:tool} presents ArKanjo, our proposed tool to detect code duplication, 
+describing all the main components. Chapter~\ref{cha:method}
+describes the research methods selected to guide our work.
+Chapter~\ref{cha:results} shows the results we had through our work, from the research methods to 
+evaluate our tool and the ethnographic studies. Chapter~\ref{cha:conclusion} concludes this research.

### 2025-08-03 conteudo/introduction/intro.tex
update the points for abstract and intro

@@
 
-This research presents ArKanjo, a command-line tool designed to detect and analyze function-level 
-duplications in the Linux kernel. Released under the MIT license, ArKanjo uses a two-stage 
+This work presents an approach to address the code duplication issue in the Linux kernel; to 
+support this approach, we developed ArKanjo, a command-line tool designed to detect and analyze 
+function-level duplications. Released under the MIT license, ArKanjo uses a two-stage 
 architecture with a Preprocessor and a Query Responder. This design separates the heavy 

### 2025-08-03 conteudo/background/nlpsimilarity.tex
change to FLOSS, i am not sure if its right

@@
 
-Gensim is an open-source Python library \citep{gensim}, which its creators
+Gensim is an FLOSS Python library \citep{gensim}, which its creators
 describe as "The fastest library for training of vector embeddings - Python or

### 2025-08-03 conteudo/introduction/intro.tex
remove numbers on intro too

@@
 The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project,
-powering 96.3\% of the top one million web servers and 85\% of smartphones worldwide. 
+powering the majority of webservers and smartphones worldwide.
 Maintaining the kernel is an enormous task, involving more than 28 million lines of 

### 2025-08-05 conteudo/background/codeduplication.tex
Review background

@@
 Code Duplication Detection is a field in computer science that has attracted
-researchers' attention since 1988 \citep{firstman}. Code
+researchers' attention since 1988~\citep{firstman}. Code
 duplication is considered a harmful artifact in software, as it negatively
 affects tasks such as code readability and increases the likelihood of
-introducing bugs \citep{harmone}. In most cases, code duplication tends to make
-software more unstable than its non-duplicated counterpart
-\citep{harmtwo}.
+introducing bugs~\citep{harmone}. In most cases, code duplication tends to make
+software more unstable than its non-duplicated counterpart~\citep{harmtwo}.
 
@@
 focuses on technical aspects, while the latter incorporates social dimensions
-into the research field \citep{litreview}. In this work, we focus solely on the
+into the research field~\citep{litreview}. In this work, we focus solely on the
 Code Clone Detection branch due to the nature of our study.
 
+\subsection{Diff command}
+
+The \textit{diff} command is a command line in GNU/Linux that allows the user to compare 
+two files line by line~\citep{diffcommand}. There are multiple forms in which the command 
+results can be shown to the user, and Figure \ref{fig:diff} shows one of them. 
+With the command results, it is easy to visualize and extract the number of lines that 
+are equal and different between the two compared files.
+
+\begin{figure}
+\includegraphics[scale=0.2]{diff_command}
+\caption[Example of \textit{diff} command executed]{
+Example of \textit{diff} command executed
+\footnote{
+File 1 content is shown in red, and File 2 content is shown in green. \textit{-} 
+represents a line that only exists in file 1, \textit{+} represents a line that only 
+exists in file 2,\textit{!} represents a line that exists in both files but has modifications 
+between them, and empty at the start of a line represents that the line exists in 
+both files without any modifications.
+}
+}
+\label{fig:diff}
+\end{figure}
+
+We use the \textit{diff} command as the base of one of the code duplication detection 
+methods on the tool proposed in this work, although it is not explored as the text 
+similarity method chosen for code duplication detection.
+
 \subsection{Types of Code Clone Duplication}
-
 \label{subsec:types}
@@
 duplication based on the extent of differences between code artifacts,
-resulting in four types of code clone duplications \citep{litreview}. When
+resulting in four types of code clone duplications~\citep{litreview}. When
 comparing two code artifacts, F and G, the four types of code duplication can
-be described as follows \citep{litreview}:
+be described as follows~\citep{litreview}:
 
@@
 	    changes in comments, variable names, white spaces, and other
-	    irrelevant elements. Figure \ref{fig:type1} shows a typical example
-	    of Type-1 clone duplication \citep{litreview}.
+	    irrelevant elements. Figure~\ref{fig:type1} shows a typical example
+	    of Type-1 clone duplication~\citep{litreview}.
     \end{item}
@@
 	    from Type-1, along with the addition and deletion of redundant
-	    code. Figure \ref{fig:type2} shows a typical example of Type-2 code
-	    clone duplication \citep{litreview}.
+	    code. Figure~\ref{fig:type2} shows a typical example of Type-2 code
+	    clone duplication~\citep{litreview}.
     \end{item}
@@
 	    from Type-2, as well as reordering code blocks and statements
-	    within code blocks. Figure \ref{fig:type3} shows a typical example
-	    of Type-3 code clone duplication \citep{litreview}.
+	    within code blocks. Figure~\ref{fig:type3} shows a typical example
+	    of Type-3 code clone duplication~\citep{litreview}.
     \end{item}
@@
 	    operands/operators in expressions, or the replacement of code parts
-	    with equivalent compositions. Figure \ref{fig:type4} shows a
-	    typical example of Type-4 code clone duplication \citep{litreview}.
+	    with equivalent compositions. Figure~\ref{fig:type4} shows a
+	    typical example of Type-4 code clone duplication~\citep{litreview}.
     \end{item}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[scale=0.38]{types/type1}
-\caption{Example of Type-1 code clone duplication.}
-The example is based on the work presented by CHEN et al. \citep{litreview}.
+\caption{Example of Type-1 code clone duplication~\citep{litreview}.}
 \label{fig:type1}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[scale=0.38]{types/type2}
-\caption{Example of Type-2 code clone duplication.}
-The example is based on the work presented by CHEN et al. \citep{litreview}.
+\caption{Example of Type-2 code clone duplication~\citep{litreview}.}
 \label{fig:type2}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[scale=0.38]{types/type3}
-\caption{Example of Type-3 code clone duplication.}
-The example is based on the work presented by CHEN et al. \citep{litreview}.
+\caption{Example of Type-3 code clone duplication~\citep{litreview}.}
 \label{fig:type3}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[scale=0.38]{types/type4}
-\caption{Example of Type-4 code clone duplication.}
-The example is based on the work presented by CHEN et al. \citep{litreview}.
+\caption{Example of Type-4 code clone duplication~\citep{litreview}.}
 \label{fig:type4}
@@
     the field. It treats code as a text artifact and applies textual similarity
-    techniques. An example is the work of Roy and Cordy \citep{textexample}, 
+    techniques. An example is the work of~\cite{textexample}, 
     which uses parsing and grammar techniques to detect code clones.
@@
     keywords, mapping variables to independent representations, etc. An example
-    is the work of Toomey et al. \citep{tokenexample}, which detects code 
+    is the work of~\cite{tokenexample}, which detects code 
     clones through hashed token sequences.
@@
      lexical analysis, tree-based approaches rely on syntax trees, which store
-     semantic information of code artifacts based on the programming language
-     \citep{compiler}. An example of this approach is the work by Chilowicz 
-     and Duris \citep{treeexample}.
+     semantic information of code artifacts based on the programming
+     language~\citep{compiler}. An example of this approach is the work of~\cite{treeexample}.
 
@@
     metrics from code artifacts as embedding representations to measure 
-    similarity \citep{litreview}, akin to the Word2Vec model by Mikolov et al. 
-    \citep{wordtovec}. Metrics can include program size, number of variables, 
-    memory access counts, etc. An example is the work by Kaur and Sharma \citep{metricexample}.
+    similarity~\citep{litreview}, akin to the Word2Vec model by~\cite{wordtovec}.
+    Metrics can include program size, number of variables, 
+    memory access counts, etc. An example is the work of~\cite{metricexample}.
 
@@
     specifically the program dependence graph (PDG), which stores information 
-    on control, statements, and data dependencies \citep{prodg}. A recent 
-    example is the work by Liu et al. \citep{tailor}, which represents one of 
+    on control, statements, and data dependencies~\citep{prodg}. A recent 
+    example is the work of~\cite{tailor}, which represents one of 
     the state-of-the-art works in the field.
@@
 computational time and memory, and many are language-specific, limiting their
-applicability across different programming languages \citep{litreview}. Given
+applicability across different programming languages~\citep{litreview}. Given
 these constraints, we opted for a generalist textual-based approach using
 Natural Language Processing (NLP) text similarity methods, which we will
-introduce in Chapter \ref{cha:tool} and evaluate in this work.
+introduce in Chapter~\ref{cha:tool} and evaluate in this work.
 
@@
 datasets are available: OJ-Clone, collected from a pedagogical online judge
-system \citep{ojclone}, and BigCloneBench, compiled from over 25,000 Java
-software systems \citep{bigclonebench}. However, literature has expressed
+system~\citep{ojclone}, and BigCloneBench, compiled from over 25,000 Java
+software systems~\citep{bigclonebench}. However, literature has expressed
 concerns over the limited scope of OJ-Clone and the issues with BigCloneBench
-identified by Krinke and Ragkhitwetsagul \citep{bigfail}. Despite these
+identified by~\cite{bigfail}. Despite these
 limitations, BigCloneBench is commonly used due to the limited pedagogical
@@
 
-BigCloneBench further divides Type-3 duplications into three categories—Weak
-Type-3 (WT3), Medium Type-3 (MT3), and Strong Type-3 (ST3)—where WT3 is closer
-to Type-4 and ST3 is closer to Type-2 \citep{bigclonebench}. This results in
+BigCloneBench further divides Type-3 duplications into three categories: Weak
+Type-3 (WT3), Medium Type-3 (MT3), and Strong Type-3 (ST3); where WT3 is closer
+to Type-4 and ST3 is closer to Type-2~\citep{bigclonebench}. This results in
 six possible outcomes for a code pair: Type-1, Type-2, Strong Type-3, Medium
-Type-3, Weak Type-3/Type-4, and non-duplicate \citep{bigclonebench}.
+Type-3, Weak Type-3/Type-4, and non-duplicate~\citep{bigclonebench}.
 
@@
 Using the counts of TP, FP, TN, and FN, standard metrics such as Precision,
-Recall, and F-Score can be calculated \citep{recall}. In this work, we focus on
+Recall, and F-Score can be calculated~\citep{recall}. In this work, we focus on
 the Recall metric, which is commonly reported in literature for comparing code
-clone detection approaches. Recall is defined as follows \citep{recall}:
+clone detection approaches. Recall is defined as follows~\citep{recall}:
 

### 2025-08-05 conteudo/background/linux.tex
Review background

@@
 \section{The Linux Kernel}
-
-The Linux kernel, a Unix-like operating system kernel, was published by Linus
-Torvalds in 1992 as free software. It began gaining popularity in the late
-1990s \citep{linuxbook}. Contrary to popular belief, Linux is not a
+Linus Torvalds released the initial version of the Linux kernel, a Unix-like operating system kernel, in 1991.
+It was initially released under a license of Torvalds' creation, but the Linux kernel was later relicensed under GNU General Public License version 2 (GPLv2) in 1992.
+%
+It began gaining popularity in the late
+1990s~\citep{linuxbook}. Contrary to popular belief, Linux is not a
 complete operating system, as it lacks components like filesystem utilities and
@@
 include various essential applications alongside the Linux kernel,
-such as system
-\citep{gnuref}. Examples of Linux distributions are Ubuntu \citep{ubuntu} 
-and Arch Linux \citep{archlinux}.
+such as system~\citep{gnuref}. Examples of Linux distributions are Ubuntu 
+and Arch Linux.
 
@@
 the process scheduler, memory management, device driver infrastructure,
-networking, and filesystems \citep{melissa}. Each subsystem typically has a
+networking, and filesystems~\citep{melissa}. Each subsystem typically has a
 dedicated maintainer or a team of maintainers responsible for overseeing its
-development and managing contributions \citep{melissa}.
+development and managing contributions~\citep{melissa}.
 
@@
 documents that outline the differences between two source code versions.
-These patches are then submitted and reviewed via the Linux mailing lists
-\citep{melissa}.
+These patches are then submitted and reviewed via the Linux mailing lists~\citep{melissa}.
 
@@
 discuss Linux-related topics, and debate potential new features. Given the
-kernel's extensive modularity, each subsystem has its mailing list, allowing
-discussions to be organized around specific areas of development
-\citep{melissa}. All interactions on these mailing lists are documented and
-archived in the Linux Kernel Lore \citep{linuxlore}, which serves as a
+kernel extensive modularity, each subsystem has its mailing list, allowing
+discussions to be organized around specific areas of development~\citep{melissa}. All interactions on these mailing lists are documented and
+archived in the Linux Kernel Lore\footnote{\url{https://lore.kernel.org}}, which serves as a
 repository for kernel discussions and can be accessed by anyone interested.
@@
 is responsible for implementing the drivers needed for AMD GPUs to function
-correctly in a Linux environment. The maintainers of this subsystem have
-reported a practical issue: a significant amount of duplicated code, which
-complicates the maintenance of this subsystem. This problem serves as the
+correctly in a Linux environment.
+%
+During the exploratory phase of this work, in a brainstorming conversation with the maintainers of this subsystem, they reported a practical issue: a significant amount of duplicated code, which complicates the maintenance of this subsystem.
+%
+This problem serves as the
 starting point for our research, with the AMD Display driver as our primary
 object of study as we aim to understand and mitigate code duplication in the
-Linux kernel. The AMD Display driver subsystem contains more than 391000 lines 
-of code and more than 1089 code files
+Linux kernel. The AMD Display driver subsystem contains more than 391,000 lines 
+of code and more than 1,089 code files
 \footnote{ 
-Data extracted directly from the Linux repository with the help of the cloc tool.
-\citep{cloc}
+Data extracted directly from the Linux repository with the help of the cloc tool~\citep{cloc}.
 }, 
@@
 Industrial I/O subsystem (IIO). According to the Linux kernel documentation, 
-''The main purpose of the Industrial I/O subsystem (IIO) is to provide support 
+``The main purpose of the Industrial I/O subsystem (IIO) is to provide support 
 for devices that in some sense perform either analog-to-digital conversion (ADC) 
-or digital-to-analog conversion (DAC) or both''\citep{iiodoc}. The IIO subsystem 
-contains more than 281772 lines of code and more than 755 code files
+or digital-to-analog conversion (DAC) or both''~\citep{iiodoc}. The IIO subsystem 
+contains more than 281,772 lines of code and more than 755 code files
 \footnote{
-Data extracted directly from the Linux repository with the help of the clol tool. \citep{cloc}
+Data extracted directly from the Linux repository with the help of the clol tool~\citep{cloc}.
 }.
@@
 \url{https://subspace.kernel.org/vger.kernel.org.html}.
-
-\subsection{Diff command}
-
-The \textit{diff} command is a command line in Linux that allows the user to compare 
-two files line by line \citep{diffcommand}. There are multiple forms in which the command 
-results can be shown to the user, and Figure \ref{fig:diff} shows one of them. 
-With the command results, it is easy to visualize and extract the number of lines that 
-are equal and different between the two compared files.
-
-\begin{figure}
-\includegraphics[scale=0.25]{diff_command}
-\caption[Example of \textit{diff} command executed]{
-Example of \textit{diff} command executed
-\footnote{
-File 1's content is shown in red, and file 2's content is shown in green. \textit{-} 
-represents a line that only exists in file 1, \textit{+} represents a line that only 
-exists in file 2,\textit{!} represents a line that exists in both files but has modifications 
-between them, and empty at the start of a line represents that the line exists in 
-both files without any modifications.
-}
-}
-\label{fig:diff}
-\end{figure}
-
-We use the \textit{diff} command as the base of one of the code duplication detection 
-methods on the tool proposed in this work, although it is not explored as the text 
-similarity method chosen for code duplication detection.

### 2025-08-05 conteudo/background/nlpsimilarity.tex
Review background

@@
 to measure their similarity. The literature proposes multiple vector
-embedding methods, such as bag-of-words, TF-IDF, LSI, and Word2Vec
-\citep{gensimlivro}. Recent works related to vector embedding construction
+embedding methods, such as bag-of-words, TF-IDF, LSI, and
+Word2Vec~\citep{gensimlivro}. Recent works related to vector embedding construction
 utilize Large Language Models (LLMs), as exemplified by Ethayarajh's work,
-which employs BERT, ELMo, and GPT \citep{llmsimilar}. The most popular distance
-function used in the literature is cosine similarity, defined by the formula
-\citep{cosineref}:
+which employs BERT, ELMo, and GPT~\citep{llmsimilar}. The most popular distance
+function used in the literature is cosine similarity, defined by the formula~\citep{cosineref}:
 
@@
 analysis by Reimers and Gurevych found that finding the most similar pair in a
-collection of 10,000 sentences required about 65 hours with BERT
-\citep{bertsimilar}. Since we aim to compare pairs of similar magnitude in
+collection of 10,000 sentences required about 65 hours with
+BERT~\citep{bertsimilar}. Since we aim to compare pairs of similar magnitude in
 our work, we decided not to use LLMs for our vector embedding method, even
@@
 
-Gensim is an FLOSS Python library \citep{gensim}, which its creators
-describe as "The fastest library for training of vector embeddings - Python or
+Gensim is a FLOSS Python library \citep{gensim}, which its creators
+describe as ``The fastest library for training of vector embeddings -- Python or
 otherwise. The core algorithms in Gensim use battle-hardened, highly optimized,
-and parallelized C routines." \citep{gensimsite}. Due to Gensim's speed, we
+and parallelized C routines''~\citep{gensimsite}. Due to Gensim speed, we
 chose this library for our vector embedding method.

### 2025-08-05 conteudo/tool/architecture.tex
Review tool

@@
 
-Regarding the programming language's function extractor, we approached this problem so that anyone 
+Regarding the programming language function extractor, we approached this problem so that anyone 
 comfortable with a language can adapt our approach to their specific programming language. 
@@
 code blocks are defined by curly brackets or by indentation, and it is possible to infer if 
-they represent a function's body given their depth in the code context and a few lines before the 
+they represent a function body given their depth in the code context and a few lines before the 
 start of the code block. We followed this approach to implement the function extractor for the 
-supported languages. As an alternative approach, we can build the syntax tree \citep{compiler} 
+supported languages. As an alternative approach, we can build the syntax tree~\citep{compiler} 
 of the programming language and use it to extract the functions, which is a more complex task than 
@@
 For a source file in the C programming language, we extract the code blocks with depth 0. 
-As per the programming language’s grammar, functions are defined in a global context at depth 0. 
+As per the programming language grammar, functions are defined in a global context at depth 0. 
 To extract the code blocks, we first need to identify the portions of code that are comments or 
@@
 that does not support the programming language completely, as the only Java codebase explored in 
-this work is the BigCloneBench \citep{bigclonebench}, and the approach implemented covers the 
+this work is the BigCloneBench~\citep{bigclonebench}, and the approach implemented covers the 
 needs of this codebase.
@@
 
-As per Java’s grammar, functions are defined inside classes and can exist at different depths. 
+As per Java grammar, functions are defined inside classes and can exist at different depths. 
 As Java accepts inner classes, this approach does not work with inner classes. The challenge 
@@
 as for-loops, can follow the closing parenthesis pattern, necessitating a better analysis of 
-these elements. The BigCloneBench \citep{bigclonebench} does not contain inner classes.
+these elements. The BigCloneBench~\citep{bigclonebench} does not contain inner classes.
 
@@
 Each function in the input codebase is represented in the temporary codebase as three files.
-Descriptions of these files are provided below, with examples shown in Figure \ref{fig:transform}.
+Descriptions of these files are provided below, with examples shown in Figure~\ref{fig:transform}.
 
@@
 	\item \textbf{Source file}: This file contains the body of the function it represents.
-	\item \textbf{Info file}: This file contains metadata about the function, including the function name, the file that contains the function, the file's relative path to the input codebase, the line where the function's signature starts, the line where the function body begins, the line where the function's body ends, and whether there is a line break between the end of the function's signature and the start of the function's body.
+	\item \textbf{Info file}: This file contains metadata about the function, including the function name, the file that contains the function, the file relative path to the input codebase, the line where the function signature starts, the line where the function body begins, the line where the function body ends, and whether there is a line break between the end of the function signature and the start of the function body.
 \end{itemize}
@@
 \subsection{Code Duplication Finder}
-
 \label{subsec:code_finder}
 
-The Code Duplication Finder iterates through every pair of source code files in the temporary codebase, representing 
-functions from the input codebase. For each pair of files, we execute a code duplication detection method that computes 
-a metric measuring how similar the pair of files is, which we refer to as similarity. 
+The Code Duplication Finder iterates through every pair of source code files in the temporary codebase, representing functions from the input codebase.
+For each pair of files, we execute a code duplication detection method that computes a metric measuring how similar the pair of files is, which we refer to as similarity.
 If the similarity is greater than or equal to the minimum similarity threshold 
@@
 as text and apply the TF-IDF vector embedding method implemented by the 
-Gensim library \citep{gensim}, then compute cosine similarity as the 
+Gensim library~\citep{gensim}, then compute cosine similarity as the 
 similarity metric.
@@
 \cite{platistool} implemented this code duplication detection method and 
-released it under the MIT license \citep{mitlicense}. We adapted his 
-implementation to fit our tool's expected input and output formats. 
+released it under the MIT license. We adapted his 
+implementation to fit our tool expected input and output formats. 
 Switching the duplication detection method is feasible, similar to how 
@@
 Setting a minimum similarity metric reduces the Code Duplication Database size, optimizing memory usage and the computational 
-cost of the Query Responder step. For example, if this parameter does not exist and we work with a codebase of $10000$ functions, 
-each with a 50-character relative path and function name, the Code Duplication Database would reach approximately $5000^2 \times 50 ~= 5$ gigabytes. 
+cost of the Query Responder step. For example, if this parameter does not exist and we work with a codebase of $10,000$ functions, 
+each with a 50-character relative path and function name, the Code Duplication Database would reach approximately $5,000^2 \times 50 ~= 5$ gigabytes. 
 As most function pairs are unlikely to be duplicated in large codebases, this file size can be significantly reduced.
@@
 The specific functionalities provided to the user by the Query Responder, which include the 
-Duplication Explorer, Function Information, and Duplication Report, are described in Section \ref{subsec:func}.
+Duplication Explorer, Function Information, and Duplication Report, are described in Section~\ref{subsec:func}.

### 2025-08-05 conteudo/methods/methods.tex
Review methods

@@
 The first method evaluates the proposed tool using an approach from the literature, 
-comparing it against the BigCloneBench dataset \citep{bigclonebench}. 
+comparing it against the BigCloneBench dataset~\citep{bigclonebench}. 
 The second method is an empirical analysis of a subset of functions within the 
@@
 We chose triangulation because relying solely on the literature-based method has 
-limitations and may be unreliable \citep{bigfail, litreview}. 
+limitations and may be unreliable~\citep{bigfail,litreview}. 
 As a reminder, we have done all research approaches using the text similarity method 
@@
 
-Given the validation of the tool’s capabilities, we proceed to understand the viability 
+Given the validation of the tool capabilities, we proceed to understand the viability 
 of mitigating the code duplication found by the tool. We conducted an ethnographic study 
@@
 of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to 
-interact with the Linux community and get their feedback. In the second phase, we proposed 
+interact with the Linux kernel community and get their feedback. In the second phase, we proposed 
 to students in the course Free Software Development at the University of São Paulo to mitigate 
-the code duplications found by the tool in the IIO subsystem and the AMD Display driver and 
+the code duplications found by the ArKanjo in the IIO subsystem and the AMD Display driver and 
 pass the process of sending the mitigations as a PATCH.
@@
 \section{Evaluation of the Tool by the BigCloneBench dataset}
-
 \label{sec:metbig}
 
-The BigCloneBench dataset \citep{bigclonebench} is a Java code collection containing pairs of code duplications categorized by clone type, as defined in Section \ref{subsec:types}.
+The BigCloneBench dataset~\citep{bigclonebench} is a Java code collection containing pairs of code duplications categorized by clone type, as defined in Section~\ref{subsec:types}.
 %
-Following the methodology presented by \citep{tailor}, we sampled 20,000 pairs from each clone type, adding 20,000 non-duplicate pairs as negative samples.
+Following the methodology presented by~\citep{tailor}, we sampled 20,000 pairs from each clone type, adding 20,000 non-duplicate pairs as negative samples.
 We applied the same sampling approach to our tool to ensure a fair evaluation.
@@
 A correctly identified duplication pair counts as accurate for its clone type, while an incorrectly inferred pair is considered incorrect across all clone types.
-With these definitions, we computed the Recall metric from Section \ref{subsec:codemethods} for each clone type.
+With these definitions, we computed the Recall metric from Section~\ref{subsec:codemethods} for each clone type.
 %
@@
 
-Due to the limitations of the BigCloneBench dataset, as noted by \cite{bigfail}, and its lack of suitability 
-for our context (Java dataset vs. C-based Linux kernel), we complemented the BigCloneBench evaluation with 
-an empirical study of the tool's inferred duplications.
+Due to the limitations of the BigCloneBench dataset, as noted by~\cite{bigfail}, and its lack of suitability for our context (Java dataset vs.
+C-based Linux kernel), we complemented the BigCloneBench evaluation with an empirical study of the tool inferred duplications.
 
@@
 For each similarity threshold \(X\) (30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%), we randomly selected ten function pairs with a similarity close to \(X\), allowing for a 1\% deviation.
-We implemented the random selection using the tool's source code.
+We implemented the random selection using the tool source code.
 
-Appendix \ref{app:emp} lists the randomly selected function pairs.
-We empirically evaluated these pairs to determine if they were duplications, comparing these results with the literature-based method to assess our tool's effectiveness and understand the impact of the minimum similarity parameter.
+Appendix~\ref{app:emp} lists the randomly selected function pairs.
+We empirically evaluated these pairs to determine if they were duplications, comparing these results with the literature-based method to assess our tool effectiveness and understand the impact of the minimum similarity parameter.
 
 \section{Ethnographic Study}
-
 \label{sec:meteth}
@@
 Ethnography is a recognized qualitative empirical method for understanding people, 
-their cultures, and their work practices \citep{bookethno}. It allows researchers 
+their cultures, and their work practices~\citep{bookethno}. It allows researchers 
 to gain insights into values, beliefs, and ideas from the perspective of community 
-members \citep{ethnosoft}. In our research, refactoring code duplications without 
+members~\citep{ethnosoft}. In our research, refactoring code duplications without 
 community interaction would prevent us from understanding whether Linux maintainers 
@@
 In C programming, communication between source files is achieved by creating header files that specify 
-libraries~\cite{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
-by consolidating duplicated code into a single library, which would replace instances of 
+libraries~\cite{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications by consolidating duplicated code into a single library, which would replace instances of 
 duplication across the codebase.
 
-For each duplicated function pair, for each function in the pair, we used the \texttt{anonymous-tool} to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair.
+For each duplicated function pair, for each function in the pair, we used the ArKanjo tool to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair.
 This strategy resulted in a collection of duplicated functions and corresponding code files.
@@
 A core dynamic of the course (concurrently for undergraduate and postgraduate studies) is to 
-have students make a practical contribution to a free software project. In the first semester of 
+have students make a practical contribution to a FLOSS project. In the first semester of 
 2025, the professor teaching the course, who is also the advisor of this work, identified an 
@@
 in the Linux kernel. The professor drove this initiative as a practical course activity, 
-and the author of this thesis was not involved in using the tool to identify duplications or 
+and the author of this work was not involved in using the tool to identify duplications or 
 in developing approaches to mitigate them. The tasks were carried out entirely by the course 
@@
 The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of
-two or three members. To assist students in this task, the professor and teaching assistants prepare multiple
-alternatives for contributing to free software projects, including simpler contributions such as
-updating documentation or fixing code style issues. In this context, two new options related to the
-\texttt{anonymous-tool} were proposed alongside the others.
+two or three members.
+To assist students in this task, the professor and teaching assistants prepare multiple alternatives for contributing to FLOSS projects, including simpler contributions such as updating documentation or fixing code style issues.
+In this context, two new options related to the
+Arkanjo tool were proposed alongside the others.
 
-For the first ArKanjo-related alternative, a teaching assistant ran the tool on the IIO subsystem twice, 
-creating two lists of duplicated function pairs, one with a 100\% similarity threshold and the other 
-with a 90\% similarity threshold. Then, the following process happened to both lists:
+For the first ArKanjo-related alternative, a teaching assistant ran the tool on the IIO subsystem twice, creating two lists of duplicated function pairs, one with a 100\% similarity threshold and the other with a 90\% similarity threshold.
+Then, the following process happened to both lists:
 
@@
 ArKanjo on the AMD display driver, independently analyzing the results to identify a duplication
-to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.
+to fix, creating a patch to mitigate it, and sending the patch to the driver maintainers.
 
-Of the students who chose to work on the \texttt{anonymous-tool}-related tasks, 
-23 (16 undergraduate and 7 graduate)
+Of the students who chose to work on the ArKanjo-related tasks, 23 (16 undergraduate and 7 graduate)
 formed 11 groups to pursue the first alternative. One graduate student also opted for the second 
@@
 
-The groups were requested to document their experiences and approaches to refactoring the duplications,
-as well as their experiences submitting a patch to the Linux kernel, in blogs. We analyzed these 
-blogs to understand the refactoring patterns used to mitigate duplications, the experience of 
-sending patches, and the opinions of subsystem maintainers on the mitigations.
+The groups were requested to document their experiences and approaches to refactoring the duplications, as well as their experiences submitting a patch to the Linux kernel, in blogs.
+We analyzed these blogs to understand the refactoring patterns used to mitigate duplications, the experience of sending patches, and the opinions of subsystem maintainers on the mitigations.

### 2025-08-05 conteudo/results/eval.tex
Review results

@@
 
-In this section, we present the test results of our tool performance, following the 
+In this section, we present the test results of our tool usage, following the 
 methodology described in Chapter \ref{cha:method}. First, we provide the results 
@@
 
-As described in Section \ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset \citep{bigclonebench}, we define 
+As described in Section~\ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset \citep{bigclonebench}, we define 
 that every pair of functions with a similarity metric equal to or greater than a threshold \( X \) is considered a duplication. 
@@
 
-As outlined in Section \ref{sec:metemp}, for the empirical evaluation of our tool, we randomly selected ten function pairs across various similarity metric ranges computed by our tool.
+As outlined in Section~\ref{sec:metemp}, for the empirical evaluation of our tool, we randomly selected ten function pairs across various similarity metric ranges computed by our tool.
 We then examined each pair to determine whether they were duplications and, if so, categorized the duplication type.

### 2025-08-05 conteudo/results/participant.tex
another grammar typos fix
Review results

@@
 duplications found by the ArKanjo tool in the AMD Display driver by applying 
-our systematic approach presented in \ref{subsubsec:systematic}. 
-As presented in section \ref{sec:meteth},
+our systematic approach presented in Section~\ref{subsubsec:systematic}. 
+As presented in Section~\ref{sec:meteth},
 we opted to apply the systematic approach in the mitigations.
@@
 Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
-files and lines changed and the refactoring method used. In the next sections, 
-we present our experience and learning for each mitigation.
+files and lines changed and the refactoring method used. The following sections
+present our experience and learning for each mitigation.
 
@@
 driver are designed to be imported at compilation time using \textit{\#define} macros. 
-The configuration files' design choice makes refactoring duplications on generic 
+The configuration files design choice makes refactoring duplications on generic 
 approaches trickier, as refactoring code that depends on these files requires 
-significant refactoring of the configuration files' design and deep knowledge 
+significant refactoring of the configuration files design and deep knowledge 
 of the codebase, which we, as first-time contributors, do not have. Thus, we opted 
@@
 the third version, the maintainers asked us to move the functions created in 
-the generic library to an existing file instead of creating a new one.
-\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.
+the generic library to an existing file instead of creating a new one\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.
 

### 2025-08-05 conteudo/results/students.tex
Review results

@@
 
-As presented in Chapter \ref{cha:method}. It was proposed to students, as one of the activities 
-of the Free Software course, they make their first contribution to the Linux kernel to 
-mitigate duplication found by the ArKanjo tool while sending patches to the Linux subsystem 
-maintainers and documenting their experiences and learnings.
+As presented in Chapter~\ref{cha:method}, we proposed an experiment to students, as one of the activities of the Free Software course at the University of São Paulo, they make their first contribution to the Linux kernel to mitigate duplication found by the ArKanjo tool while sending patches to the Linux subsystem maintainers and documenting their experiences and learnings.
 
-Appendix \ref{app:stuart} has references to the artifacts discussed in this section, 
+Appendix~\ref{app:stuart} has references to the artifacts discussed in this section, 
 including the duplication function pairs that students mitigate, the students’ reports, 
@@
 
-Table \ref{tab:stu} summarizes the student groups' mitigation approaches. It contains 
+Table~\ref{tab:stu} summarizes the student groups' mitigation approaches. It contains 
 the information about the chosen subsystem and the similarity threshold for groups that 
@@
 
-Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach duplications found by the \texttt{anonymous-tool} and become contributors to the kernel.
-We observed that not all duplications
-are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code, code readability, and the levels of abstraction added to address the duplications.
+Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach duplications found by the ArKanjo tool and become contributors to the kernel.
+We observed that not all duplications are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code, code readability, and the levels of abstraction added to address the duplications.
 
@@
     \item The AMD Display subsystem has a more centralized code structure and maintainer hierarchy, enabling contributors to tackle code duplications across a broader scope and multiple files, which happened to Group 11's contribution;
-    \item The complete control over the process of finding code duplication gave Group 11 the opportunity to identify and reason better over the duplications found by \texttt{anonymous-tool}, enabling a more strategic approach to searching for promising patch possibilities.
+    \item The complete control over the process of finding code duplication gave Group 11 the opportunity to identify and reason better over the duplications found by ArKanjo, enabling a more strategic approach to searching for promising patch possibilities.
 \end{itemize}
@@
 in the Arch Linux distribution and failed to install it on this Linux distribution. 
-Regarding the tool's documentation, the teaching assistant pointed out errors that worsened his 
+Regarding the tool documentation, the teaching assistant pointed out errors that worsened his 
 experience. The student asked for more examples and tutorials on using the tool in text or 
@@
 
-Related to the tool's functionalities, both the student assistant and the student reported 
+Related to the tool functionalities, both the student assistant and the student reported 
 a missing mechanism to save the tool results in a separate file instead of the terminal output. 
-The tool, by default, outputs with colors, as shown in section \ref{subsec:func}, which complicates 
+The tool, by default, outputs with colors, as shown in Section~\ref{subsec:func}, which complicates 
 redirecting the terminal output to a file.
@@
 We listed multiple issues in the tool for the two groups that opted to contribute directly 
-to developing the ArKanjo tool. The first group opted to work on expanding the tool's support 
+to developing the ArKanjo tool. The first group opted to work on expanding the tool support 
 for multiple operating systems. Some libraries used in the tool development only supported the 
@@
 The second group opted to work on a webpage for the tool documentation and support for 
-light and dark modes. The colored output presented in the figures in Chapter \ref{cha:tool} was 
+light and dark modes. The colored output presented in the figures in Chapter~\ref{cha:tool} was 
 hard-coded at first, not considering the user terminal colors, thus giving a bad user experience. 
-The website helps facilitate people's contribution or use of the tool. We did not know how to 
+The website helps facilitate people's contribution or use of the tool: \url{https://arkanjo-tool.github.io}. We did not know how to 
 approach the light and dark modes at first and did not need to look at how to fix the issue as 

### 2025-08-05 conteudo/conclusion/conclusion.tex
Review conclusion

@@
-The Linux Kernel's massive scale, with over 28 million lines of code 
+The Linux Kernel massive scale, with over 28 million lines of code 
 and more than twenty thousand contributors, presents significant maintenance 
@@
 potentially hindering its evolution and patching processes. Prevailing academic 
-and existing free software approaches to duplication detection often lack the 
+and existing FLOSS approaches to duplication detection often lack the 
 scalability for comprehensive codebase analysis or offer limited functionality 
-unsuitable for the kernel's complexity, generally failing to deliver actionable 
+unsuitable for the kernel complexity, generally failing to deliver actionable 
 results or guidance for mitigation within its specific context.
@@
 
-The ethnographic studies, which involved participant observation and a project with university 
+The ethnographic studies, which involved participant observation and an experiment with university 
 students, provided a realistic view of the opportunities and the considerable challenges in 
@@
 analysis and improved detection of complex clone types. Such an approach could complement efforts for 
-improved resource management during its preprocessing phase for large codebases. The tool's language 
+improved resource management during its preprocessing phase for large codebases. The tool language 
 support could be broadened by developing parsers for additional programming languages, building upon 

### 2025-08-06 conteudo/results/eval.tex
add ~ on ref and cite

@@
 In this section, we present the test results of our tool usage, following the 
-methodology described in Chapter \ref{cha:method}. First, we provide the results 
-obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}, which 
-serves as a standard reference in the literature \citep{litreview}. Second, we 
+methodology described in Chapter~\ref{cha:method}. First, we provide the results 
+obtained by our tool on the BigCloneBench dataset~\citep{bigclonebench}, which 
+serves as a standard reference in the literature~\citep{litreview}. Second, we 
 discuss the results from our empirical analysis. Finally, we triangulate the data, 
@@
 As a reminder, we only expect our tool to yield good results for Type-1 and Type-2,
-as mentioned in \ref{subsec:code_finder}.
+as mentioned in~\ref{subsec:code_finder}.
 
@@
 
-As described in Section~\ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset \citep{bigclonebench}, we define 
+As described in Section~\ref{sec:metbig}, for evaluating our tool on the BigCloneBench dataset~\citep{bigclonebench}, we define 
 that every pair of functions with a similarity metric equal to or greater than a threshold \( X \) is considered a duplication. 
@@
 duplication pair incorrectly, it is counted as incorrect for all code duplication types.
-Table \ref{tab:bigclone} shows the results obtained by our tool on the BigCloneBench dataset \citep{bigclonebench}.
+Table~\ref{tab:bigclone} shows the results obtained by our tool on the BigCloneBench dataset~\citep{bigclonebench}.
 
-The results in Table \ref{tab:bigclone} indicate that our tool is conservative in its predictions.
+The results in Table~\ref{tab:bigclone} indicate that our tool is conservative in its predictions.
 No negative sample was inferred as a duplication, even with a minimum similarity threshold of 30\%.
 However, it performs poorly identifying complex duplication types, such as Medium Type-3 and Weak Type-3/Type-4.
-As discussed in Chapter \ref{cha:tool}, if improved performance on these complex types is needed, our tool allows replacing the current duplication detection method due to its modular design.
+As discussed in Chapter~\ref{cha:tool}, if improved performance on these complex types is needed, our tool allows replacing the current duplication detection method due to its modular design.
 
@@
 We then examined each pair to determine whether they were duplications and, if so, categorized the duplication type.
-Table \ref{tab:emp} presents the results from our empirical analysis.
+Table~\ref{tab:emp} presents the results from our empirical analysis.
 
@@
 
-The results in Table \ref{tab:emp} show that as the similarity metric decreases, 
+The results in Table~\ref{tab:emp} show that as the similarity metric decreases, 
 there is a notable increase in function pairs that are not duplications within 

### 2025-08-06 conteudo/background/codeduplication.tex
fix some ~ before ref and cite and a wrong , on tool arch

@@
 two files line by line~\citep{diffcommand}. There are multiple forms in which the command 
-results can be shown to the user, and Figure \ref{fig:diff} shows one of them. 
+results can be shown to the user, and Figure~\ref{fig:diff} shows one of them. 
 With the command results, it is easy to visualize and extract the number of lines that 
@@
 understanding, these approaches are categorized into five methodologies based
-on their characteristics \citep{litreview}:
+on their characteristics~\citep{litreview}:
 
@@
     \item \textbf{Textual-based approaches:} According to Chen’s literature 
-    review \citep{litreview}, this is the earliest methodology explored in 
+    review~\citep{litreview}, this is the earliest methodology explored in 
     the field. It treats code as a text artifact and applies textual similarity
@@
     \item \textbf{Token-based approaches:} This method tackles the problem 
-    similarly to lexical analysis in the compilation process \citep{litreview}. 
+    similarly to lexical analysis in the compilation process~\citep{litreview}. 
     It involves identifying programming language tokens such as constants and 

### 2025-08-06 conteudo/results/students.tex
fix some ~ before ref and cite and a wrong , on tool arch

@@
 some questions related to the tool experience, such as installation, documentation, ease of 
-use, new ideas to improve the tool, etc. Appendix \ref{app:survey} has the complete survey 
+use, new ideas to improve the tool, etc. Appendix~\ref{app:survey} has the complete survey 
 with the answers.
@@
 with an operation system default command. The teaching assistant report corroborates the 
-gensim efficiency \citep{gensim}.
+gensim efficiency~\citep{gensim}.
 
@@
 The issues help the user experience and facilitate other research using the ArKanjo tool that 
-we may explore in future works. Appendix \ref{app:contribution} references the 
+we may explore in future works. Appendix~\ref{app:contribution} references the 
 students' contributions, the issue list proposed to the students, and figures from 

### 2025-10-06 conteudo/background/linux.tex
fix points from defense

@@
 include various essential applications alongside the Linux kernel,
-such as system~\citep{gnuref}. Examples of Linux distributions are Ubuntu 
+such as core utilities, shelss, and compilers~\citep{gnuref}. Examples of Linux distributions are Ubuntu 
 and Arch Linux.

### 2025-10-06 conteudo/results/eval.tex
fix points from defense

@@
 
-\section{Tool Results on BigCloneBench}
+\subsection{Tool Results on BigCloneBench}
 
@@
 We cannot determine if state-of-the-art code duplication detection tools face similar issues obscured by evaluations limited to the BigCloneBench dataset.
+
+\subsection{Performance Evaluation of the Tool}
+
+Our performance evaluation, conducted on a machine equipped with a Ryzen 5700X processor 
+and 32GB of RAM, highlights the efficiency of the tool's two-stage architecture. 
+The offline Preprocessor step recorded a worst-case execution time of 8 minutes and 56 
+seconds for the AMD Display driver and 9 minutes and 41 seconds for the IIO subsystem 
+across five runs. ⏱️ In sharp contrast, the online Query Responder demonstrated real-time 
+capabilities, processing user requests in under 2 seconds in all test cases. This significant 
+performance difference underscores the effectiveness of separating the time-intensive 
+analysis from the interactive querying phase.

### 2025-10-06 conteudo/results/participant.tex
fix some ~ before ref and cite and a wrong , on tool arch
fix points from defense

@@
 requesting minimal changes. We submitted a second version with the requested 
-changes on November 16, 2024. We did not receive replies initially and will 
-need to resend the patch on February 7, 2025. We received a reply on 
+changes on November 16, 2024. We did not receive replies initially and needed
+to resend the patch on February 7, 2025. We received a reply on 
 February 11, 2025, asking for additional changes. We sent a third version on 
 February 24, 2025, and the patch was accepted and integrated into the kernel 
-codebase on February 25, 2025. Figure \ref{fig:timeline} illustrates the 
+codebase on February 25, 2025. Figure~\ref{fig:timeline} illustrates the 
 timeline of the patch submission.
@@
 the supported systems while simultaneously implementing measures to mitigate errors. 
-Additionally, we comprehended that AMD Display driver developers can not view 
+Additionally, we observed that AMD Display driver developers can not view 
 duplicated code negatively, as we understand it is a good practice in software engineering. 

### 2025-10-06 conteudo/results/eval.tex
fix typo

--- 2025-08-06 conteudo/results/eval.tex
@@
 We cannot determine if state-of-the-art code duplication detection tools face similar issues obscured by evaluations limited to the BigCloneBench dataset.
+
+\subsection{Performance Evaluation of the Tool}
+
+Our performance evaluation on a Ryzen 5700X machine with 32GB of RAM shows the tool's efficiency. 
+During testing, the offline Preprocessor's longest run was 8 minutes and 56 seconds for the AMD Display driver 
+and 9 minutes and 41 seconds for the IIO subsystem. By contrast, the online Query Responder handled all user
+requests in less than 2 seconds. This result confirms the value of separating the time-consuming analysis 
+from the real-time query phase.

### 2025-10-16 paper/techdebt2026/techdebt.tex
Add dir to TechDebt 2026 paper
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

\documentclass[10pt,conference]{IEEEtran}
\usepackage{graphicx,url}
\usepackage{hyperref}
\usepackage{nicematrix}
\usepackage{array}
\usepackage[utf8]{inputenc}  
\usepackage{seqsplit}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{Mitigating Code Duplication in the Linux Kernel}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations

\author{\IEEEauthorblockN{Anonymous Authors}}

%\author{\IEEEauthorblockN{Luan Arcanjo, Marcelo Spessoto, David Tadokoro, Paulo Meirelles}
%\IEEEauthorblockA{
%Free Software Competence Center\\
%Institute of Mathematics, Statistics, and Computer Science
%Universidade de São Paulo, Brazil\\
%\{luanicaro,marcelomspessoto,davidbtadokoro\}@usp.br,paulormm@ime.usp.br}
%}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The vast scale and continuous evolution of the Linux kernel make its maintenance a complex undertaking, where code duplication remains a persistent challenge that can hinder development and introduce bugs. This paper presents a multimethod ethnographic study investigating the socio-technical dynamics of contributing to the Linux kernel by mitigating identified redundancies. To support this investigation, we developed \texttt{anonymous-tool}, a command-line utility for detecting function-level duplications, offering a concrete entry point for new contributors. Our ethnographic approach, including participant and non-participant observations, demonstrated that addressing duplications is viable for lowering the contribution barrier. This point is evidenced by the main author and 7 of 11 (64\%) student groups having their patches accepted into the kernel, collectively removing 991 lines of duplicated code. Moreover, the study also reveals a more complex reality beyond simple clone removal: an analysis of maintainer feedback on both accepted and rejected patches highlights a nuanced understanding of code quality within the Linux kernel community, where the benefits of eliminating duplication are carefully weighed against factors such as readability, the introduction of new abstractions, and the specific context of the code.
\end{abstract}

% no keywords

\IEEEpeerreviewmaketitle

\section{Introduction}
\label{sec:introduction}

Coordinating hundreds or thousands of developers working on a software product is complex. Creating similar or even completely duplicated code is a frequent and often unavoidable outcome. Excessive duplication can negatively impact the maintainability of a project. When duplications occur, any change to a repeated code segment typically requires manual replication across all other copies. Developers must, therefore, remain constantly vigilant to synchronize different parts of the codebase, which is a highly error-prone and seemingly simple responsibility. With many duplications, modifications become slow, complex, and fragile. Removing or refactoring duplicated code demands great care to preserve the original behavior. Even minor mistakes can introduce subtle state changes that may lead to serious problems in the future.

The Linux kernel is a foundational Free/Libre and Open Source Software (FLOSS) project, a vital component of the world's digital infrastructure. Its maintenance is a massive undertaking, encompassing over 28 million lines of code and involving contributions from over twenty thousand developers. Within this context, code duplication remains a significant challenge,  reducing the code readability and increasing the risk of introducing bugs~\cite{harmone,harmtwo}. This issue is particularly prominent in kernel device drivers, which account for over 66\% of the Linux kernel source code.

The detection of code duplication, or \textit{code clones}, has been a research topic for decades~\cite{firstman}. The literature offers a widely accepted taxonomy that classifies clones into four types based on their similarity, from exact copies (Type-1) to semantically equivalent but syntactically different fragments (Type-4)~\cite{litreview}. Several detection techniques have emerged, including textual, token-based, tree-based, and graph-based approaches~\cite{litreview}. These efforts have culminated in state-of-the-art techniques, such as the graph-based approach proposed by Liu et al.~\cite{tailor}.

To address the code duplication issue, we developed \texttt{anonymous-tool}, a command-line tool designed to support Linux kernel maintenance by detecting and analyzing function-level duplications. This paper presents a multimethod ethnographic study to validate the tool capabilities in identifying duplications and supporting contributions to the Linux kernel by mitigating redundant code. Therefore, the study investigates how the Linux kernel community perceives the code quality related to code clones. Our findings indicate that maintainers consider not only duplication but also factors such as readability and appropriate levels of abstraction regarding code maintainability.

Ethnography is a well-established qualitative research method used to understand people, their cultures, and their work practices~\cite{bookethno}. It enables researchers to gain insights into values, beliefs, and practices from the perspective of community members~\cite{ethnosoft}.  
In this study, we first adopted a participant observation approach, where the main author actively engaged in mitigating code duplications while documenting the process and outcomes. Later, we followed a non-participant observation involving students tasked with resolving duplications identified by the tool. These combined studies showed that \texttt{anonymous-tool} can collaborate to reduce the entry barrier for new contributors. The student groups and the main author used the tool to guide their initial contributions to the Linux kernel.

\section{The Linux Kernel}

The Linux kernel is a monolithic system organized into subsystems, such as the process scheduler, memory management, and device drivers. Each subsystem typically has a dedicated maintainer or a team of maintainers overseeing its development and managing contributions. The development process is coordinated through the Git version control system. Contributions are formatted as patches, text documents that outline the differences between two source code versions. These patches are then submitted to the relevant Linux mailing lists for public review and discussion.

This study focuses on two of these subsystems, selected to represent different contexts for contribution. We explore the AMD Display driver, chosen for its importance within the hardware ecosystem, which is responsible for enabling AMD GPU functionality in Linux. This subsystem is substantial, with over 391,000 lines of code across more than 1,089 files. We also investigate the Industrial I/O (IIO) subsystem, often considered an accessible entry point for new contributors. The IIO subsystem provides support for devices that perform analog-to-digital or digital-to-analog conversions\footnote{\href{https://kernel.org/doc/html/v4.14/driver-api/iio/intro.html}{kernel.org/doc/html/v4.14/driver-api/iio/intro.html}} and contains over 281,772 lines of code in more than 755 files.

\section{Anonymous-tool}

We developed \texttt{anonymous-tool}, a command-line tool designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation. Released under the MIT license, \texttt{anonymous-tool} employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results. The first stage, the Preprocessor, analyzes the entire codebase, parses C functions, and stores identified duplicates in a database. The second stage, the Query Responder, uses this database to retrieve information about the identified duplications quickly.
%The tool is available at \textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.

\texttt{Anonymous-tool} is designed for the specific challenges of the kernel, with a focus on scalability to handle its size and prioritizing duplications that impact maintainability. The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying function-level duplications to address. Figure~\ref{fig:diagrama} illustrates the tool's architecture. The tool uses a text similarity method that applies TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.275]{diagrama_mestrado}
\caption{Architecture diagram with the tool components
}
\label{fig:diagrama}
\end{figure}

\section{Participant Observation Study}
\label{sec:participant}

To investigate whether we can use the \texttt{anonymous-tool} tool to find ways to contribute to the Linux kernel, we conducted a participant observation experiment in the second semester of 2024, where we acted as first-time contributors to the AMD Display driver, collecting artifacts of our experience in the process.

In the AMD Display driver codebase, we executed the \texttt{anonymous-tool}. We manually analyzed the biggest duplications regarding the number of lines and found two duplicated function pairs that we judged interesting to try to mitigate. We have chosen to approach the biggest duplications as proof of concept so that we can make significant impacts.

The first duplicated function pair chosen is function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}. The second duplicated function pair is function \textit{phy\_id\_to\_atom} in file 
\textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
\textit{phy\_id\_to\_atom} will be referred to as mitigation 2.

Given the two duplicated function pairs, we observed that the files in the functions 
contain multiple other duplications. Thus, we proposed a simple systematic approach 
to mitigate all the functions duplicated in the context, not just the duplicated 
function pairs. 
After we approached the duplications with the systematic strategy, for the duplicated 
function pairs that had successfully mitigated the duplication, we sent patches to the 
AMD Display driver email list to validate the mitigations and see the maintainer's opinion 
while documenting the process and our impressions.
The systematic approach is defined below.

In C programming, communication between source files is achieved by creating header files that specify 
libraries~\cite{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
by consolidating duplicated code into a single library, which would replace instances of 
duplication across the codebase.

For each duplicated function pair, for each function in the pair, we used the \texttt{anonymous-tool} to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair. 
This strategy resulted in a collection of duplicated functions and corresponding code files.

To identify shared code more effectively, we extended this approach to search for other common 
functions across all collection files. We then applied specific refactoring methods to each shared 
function. If the functions were identical across files, we removed the duplicates and created a single 
function in the library. If modifications existed, we applied case-specific refactoring.

\subsection{Results}
\label{subsec:participantresults}

Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
files and lines changed and the refactoring method used. The following sections present our experience and learning for each mitigation.

\begin{table}[ht]
\centering
\caption{Refactor methods used and mitigation metrics in the participant-observation experiment.}
\begin{tabular}{ | m{12mm} | m{10mm} | m{10mm} | m{14mm} | m{20mm} | }

\hline

\textbf{Mitigation} & \textbf{Files changed} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
\\ \hline

1 & 13 & +224 & -753 & Parameterize Method, Extract Method  \\ \hline
2 & 8 & +114 & -520 & None \\ \hline

\hline
\end{tabular}

\label{tab:patch}
\end{table}

\subsubsection{Mitigation 1}

In the systematic approach, we found that the \textit{offset\_to\_id} function 
exists in 10 code files, enclosing multiple GPU architectures. The duplicated 
functions are not exactly equal, as some minimal specific logic is applied to each GPU architecture. To address this issue, we used the extract method to 
break the function into smaller parts and the parameterized method to address 
the specific logic for the GPUs.

We reached a point where we had a scratch on how to mitigate the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver. We found that the configuration files on the AMD Display driver are designed to be imported at compilation time using \textit{\#define} macros. The configuration files' design choice makes refactoring duplications on generic approaches trickier, as refactoring code that depends on these files requires significant refactoring of the configuration files' design and deep knowledge of the codebase, which we, as first-time contributors, do not have. Thus, we opted not to continue investigating this mitigation.

\subsubsection{Mitigation 2}

Using the systematic approach, we found that the \textit{phy\_id\_to\_atom} function exists in five code files, and two other functions are duplicated 
in those five code files. All functions across the files were exactly equal, 
so there was no need to apply the refactoring methods presented in the literature. 
The refactoring was resumed to create a generic library and fix the compilation targets.

The code files in the second mitigation do not depend on configuration files. Thus, 
we did not face the same issues during the first mitigation. Refactoring the code 
to mitigate duplications reached a point we judged was good enough to submit as a 
patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
the refactoring to the driver maintainers as a patch, documenting our findings and experiences 
on the process.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.325]{timeline_patch}
\caption{Timeline of the PATCH submission process}
\label{fig:timeline}
\end{figure}

The experience of submitting a patch was not as we expected. The process took 7 months and 16 days, significantly slower than expected. We initially submitted the patch on August 9, 2024, but we did not receive replies. Thus, we had to resend it on October 9, 2024. We received an initial response on November 3, 2024, requesting minimal changes. We submitted a second version with the requested changes on November 16, 2024. We did not receive replies initially and will need to resend the patch on February 7, 2025. We received a reply on February 11, 2025, asking for additional changes. We sent a third version on February 24, 2025, and the patch was accepted and integrated into the kernel codebase on February 25, 2025. Figure \ref{fig:timeline} illustrates the timeline of the patch submission.

After the initial reply from the maintainers, we observed that the review process became more responsive. We hypothesize that once a patch receives initial attention, it is more closely tracked by maintainers. Another contributing factor to the initial delay could be the maintainers' workload at the time of the first submission.

We discussed the Display driver with our point of contact at AMD to understand the reason for the slow patch submission process. We learned that some of the code in the driver is shared across all operating systems that support the implemented GPUs. This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors. Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering. One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others. This approach helps save a significant amount of time and effort.

Regarding the patch changes requested for the second version, the maintainers only asked for minor changes to align with code style, correct license use, and best practices that were initially unknown. For the change requested on the third version, the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.

\section{Non-participant Observation Study}

A core dynamic of the course (concurrently for undergraduate and postgraduate studies) is to have students make a practical contribution to a free software project. In the first semester of 2025, the professor teaching the course, who is also the advisor of this work, identified an opportunity to leverage the \texttt{anonymous-tool} tool within this existing pedagogical structure. As a contribution option, the professor proposed that students mitigate code duplications found in the Linux kernel. The professor drove this initiative as a practical course activity, and the author of this thesis was not involved in using the tool to identify duplications or in developing approaches to mitigate them. The tasks were carried out entirely by the course staff and the students.

The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of 
two or three members. To assist students in this task, the professor and teaching assistants prepare multiple 
alternatives for contributing to free software projects, including simpler contributions such as 
updating documentation or fixing code style issues. In this context, two new options related to the 
\texttt{anonymous-tool} were proposed alongside the others.

For the first anonymous-tool-related alternative, a teaching assistant ran the tool on the IIO subsystem twice, creating two lists of duplicated function pairs, one with a 100\% similarity threshold and the other with a 90\% similarity threshold. Then, the following process happened to both lists:

\begin{enumerate}
    \item The teaching assistant filtered the lists to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different devices may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
    \item The teacher assistant removed duplications that were too short and would not represent a significant change for a patch application;
    \item The teacher assistant manually ranked code duplications and marked some duplications with insights based on personal opinion and perceptions over the remaining entries. This was done to guide students so that they could approach the technical debt with a greater chance of having their contribution merged into the code.
\end{enumerate}

The students' task was to select a recommended entry from the list, refactor the code to eliminate these specific duplications, and then submit their patches to the IIO maintainers for review.

For the second alternative, students were offered an experience similar to our initial participant 
observation. In this model, the students were responsible for the entire workflow: executing the 
\texttt{anonymous-tool} on the AMD display driver, independently analyzing the results to identify a duplication 
to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.

Of the students who chose to work on the \texttt{anonymous-tool}-related tasks, 23 (16 undergraduate and 7 graduate) 
formed 11 groups to pursue the first alternative. One graduate student also opted for the second alternative. Notably, none of these students had prior experience contributing to the Linux kernel, making this their first attempt to submit a patch as newcomers to the project.
%TODO: ver na versão final incluir o patch do Lucas (Octavio e Fernando)

The groups were requested to document their experiences and approaches to refactoring the duplications, as well as their experiences submitting a patch to the Linux kernel, in blogs. We analyzed these blogs to understand the refactoring patterns used to mitigate duplications, the experience of sending patches, and the opinions of subsystem maintainers on the mitigations.

\subsection{Overview of the submitted patches}

Table \ref{tab:stu} summarizes the student groups' mitigation approaches. It contains the information about the chosen subsystem and the similarity threshold for groups that opted for the first approach. Each entry also has the patch status, refactoring method, and code diff corresponding to the final patch versions before it was dropped or accepted into the code base. The code diff for a patchset corresponds to the sum of added and removed lines of each of its patches.

\begin{table*}
\centering
\caption{Refactor methods used and metrics of the students' mitigations.}
\begin{tabular}{ | c | m{15mm} | m{20mm} | m{30mm} | m{30mm} | m{20mm} | }
\hline
\textbf{Group} & \textbf{Subsystem} & \textbf{Similarity} & \textbf{Patch Status} & \textbf{Refactoring Method} & \textbf{Code Diff} \\
\hline

1 & IIO & 100\% & Dropped (v2) & Undefined & +7/-14 \\ \hline
2 & IIO & 100\% & Dropped (v1) & Parameterize method & +23/-27 \\ \hline
3 & IIO & 100\% & Accepted (v1) & Parameterize method & +12/-30 \\ \hline
4 & IIO & 100\% & Accepted (v3) & Undefined & +2/-7 \\ \hline
5 & IIO & 90\% & Accepted (v1) & Parameterize method & +17/-70 \\ \hline
6 & IIO & 90\% & Dropped (v1) & Parameterize method & +95/-92 \\ \hline
7 & IIO & 90\% & Accepted (v3) & Extract method & +20/-36 \\ \hline
8 & IIO & 90\% & Accepted (v4) & Extract method & +20/-16 \\ \hline
9 & IIO & 90\% & Accepted (v6) -- Partially & Parameterize method & +149/-243 \\ \hline
10 & IIO & 90\% & Dropped (v1) & Undefined & +32/-58 \\ \hline
11 & IIO & 100\% & Dropped (v1) & Inline method & +2/-12 \\ 
   & AMD & 90\% & Accepted (v2) & Undefined & +90/-489 \\ \hline

\end{tabular}%

\label{tab:stu}
\end{table*}

Seven groups applied the parameterized refactor method, two groups used the extractor method, and one group applied the inline method to their approaches to mitigate their duplications. Four patches do not contain a refactoring technique defined in the literature (undefined). Six groups had their patches accepted and fully merged into Linux; four groups had their patches dropped; one group has its patchset partially applied and is still working on patch submission, currently in the sixth version of the patch when this document was written, but we can also consider this case as accepted.

Groups 1 and 4 have worked in similar duplications, where two functions are involved: one to determine if a register is volatile and the other to determine if it is writable. These are two functions that return opposite Boolean results. Group 1 initially addressed the problem using the extract method, while Group 2 used macros. After feedback from maintainers, both solutions converged to keep one of the functions while changing the content of the other to return a negated query of the former function. However, by the end of the feedback cycle, maintainers considered the patch from Group 1 irrelevant, while the patch from Group 4 was accepted. The reason is that Group 1 has chosen to apply the refactoring to a device that contains a register that is both read-only and non-volatile. Thus, the technical debt selected by Group 1 was not merely a duplication but an incorrect implementation of the method.

Groups 3, 6, and 10 created a struct to encapsulate the parameters and mitigate duplication when applying the parameterized method. The patches from Groups 6 and 10 were rejected by the same maintainer who approved the patch from Group 3. The maintainer's justification for rejecting the patch from Group 6 is that it proposes to fix duplication while adding more lines than removing, and is not convincing enough in other aspects, such as readability. The reason for rejecting the patch from Group 10 is that they refactored code files with simple context; therefore, adding a layer of abstraction to remove duplication decreases code readability without clear gains in codebase quality. The maintainers' feedback for Group 10 was:
\begin{quote}

\textit{``A few comments inline, but overall I'm not sure the code reduction is sufficient to cover the resulting loss of readability. Sometimes a switch is simply clearer than a partial look-up table.''}

\textit{``So nice idea, but I'm not seeing it as being a good move for this particular code. I'd be slightly interested to see the optimized output of the two approaches, but this is far from a high-performance path, so we care a lot more about readability here.''} 
\end{quote}

Interestingly, both groups addressed duplications across initialization functions for different devices, where code readability is vulnerable to refactoring methods. In contrast, Group 3 tackled the duplication of a simple iterative algorithm, which also contained a bug that could be inferred by inspecting both functions. In this case, the group refactored code from a more critical technical debt and took the opportunity to fix an unnoticed incorrect behavior.

%\begin{figure}
%\centering
%\includegraphics[scale=0.8]{group_10_reply}
%\caption[Maintainers' Feedback on Group 10 Patch]{
% Maintainers' Feedback on Group Ten's Patch
%}
%\label{fig:group_10_reply}
%\end{figure}

Groups 7 and 8 used the extract method to approach the mitigations. Group 8 sent the mitigations to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed. For Group 7 mitigation, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function. After the feedback cycle, maintainers accepted the patch in version 3.

Group 9 initially approached the selected duplication with a parameterized method, but after entering the feedback cycle, it was revealed that the problem represented a much more complex technical debt. The patch is still under development, currently at version 6, and has been separated into a 4-patch patch set, where half of them have been applied, and the remaining entries are receiving suggestions. Fortunately, maintainers are receptive and provide constant feedback; the contribution is expected to be fully merged if the students do not abandon the current work.

Group 11 is the student who executed the tool and completed the process of identifying duplications. The student identified a duplication in the driver that occurred across eighteen code files, creating a patch to remove approximately four hundred lines of duplicated code. The student sent the mitigations to the maintainers, and their feedback pointed to fixing the warning in the code without arguing about removing the duplications. After addressing the issue, the patch was applied. This student also contributed to IIO, tackling a code duplication with inline method refactoring. This one, however, was received with more skepticism by the maintainer, and the student dropped the contribution because he did not think the feedback instructions were clear enough.

Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach duplications found by the \texttt{anonymous-tool} and become contributors to the kernel. We observed that not all duplications 
are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code, code readability, and the levels of abstraction added to address the duplications.

\subsection{Statistics of the contributions}

The eleven groups, which have submitted 11 contributions to the IIO subsystem, have collectively requested the insertion of 379 lines and the deletion of 605 lines in the subsystem, representing an average of 34.45 lines added and 55 lines removed per contribution.

Considering the sum of fully merged contributions, we have a code diff value of +51/-143, with an average of +12.75/-35.75. This ratio is better than dropped contributions, which have a total of +179/-219 with an average of +29.83/-36.5.

It is also possible to notice a difference based on the similarity threshold used to detect duplications. The proposed patches to address 100\%-similarity duplications summed up to +46/-90 (average of +9.2/-18). For a 90\% value, the result was a total of +333/-515 with an average of +55.5/-85.83. The 90\% value enabled \texttt{anonymous-tool} to identify less strict duplication cases, and, therefore, code segments with a greater extension, compared to the 100\% similarity entries, which were shorter segments, causing a difference in the average number of modified lines.

Finally, the AMD contribution from Group 11, which involves the complete use of the \texttt{anonymous-tool} toolchain, got a completely different outcome. With a total of 90 lines inserted and 489 lines removed, there are some points to be considered for the success and high modification extension of the contribution:

\begin{itemize}
    \item The AMD Display subsystem has some self-reported problems regarding overall technical debt;
    \item The AMD Display subsystem has a more centralized code structure and maintainer hierarchy, enabling contributors to tackle code duplications across a broader scope and multiple files, which happened to Group 11's contribution;
    \item The complete control over the process of finding code duplication gave Group 11 the opportunity to identify and reason better over the duplications found by \texttt{anonymous-tool}, enabling a more strategic approach to searching for promising patch possibilities. 
\end{itemize}

\section{Concluding Remarks}

The ethnographic studies, which included participant observation and experimentation with students, offered a realistic perspective on the opportunities and significant challenges in mitigating code duplications found by the \texttt{anonymous-tool} tool. The tool was effectively used by first-time contributors to the Linux kernel to identify duplications and submit patches~\footnote{To avoid violating the double-anonymized review process, we will add the tool repository, patches, and artifact links in the final camera-ready version of the paper, if it is accepted for publication.}.

While this research included 1 successfully merged patch by the main author, and 7 out of 11 student groups had their patches accepted and merged into Linux, a portion of the students' efforts encountered hurdles: 4 student groups had their patches rejected.
%TODO: avaliar se não estamos sendo muito negativos ao destacar aqui as 4 rejeições.
Despite reducing duplication, these difficulties often stemmed from maintainer feedback, where proposed changes were rejected or required substantial rework. This was often due to concerns about code readability, added abstraction, or the specific context of the code. Technical complexities, such as using C macros in configuration files, also contributed to the challenges. Additionally, lengthy patch review times further compounded these difficulties.

Furthermore, the studies indicated that maintainers may not always view code duplication as inherently unfavorable, especially when it might improve clarity or facilitate independent hardware support (in the IIO subsystem matter). This suggests a nuanced understanding of code quality within the Linux kernel community, where practicality and maintainability in specific contexts can sometimes outweigh the general principle of eliminating all code duplication.

%\section*{Acknowledgments}

\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,references}

\end{document}

### 2025-10-16 paper/techdebt2026/techdebt.tex
version with 5 + 1 pages and new title
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 
-\title{Mitigating Code Duplication in the Linux Kernel}
+\title{Repeat Yourself? When Linux kernel maintainers balance Code Duplication and Readability}
 
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.325]{timeline_patch}
-\caption{Timeline of the PATCH submission process}
-\label{fig:timeline}
-\end{figure}
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[scale=0.325]{timeline_patch}
+%\caption{Timeline of the PATCH submission process}
+%\label{fig:timeline}
+%\end{figure}
 
-The experience of submitting a patch was not as we expected.
+%The experience of submitting a patch was not as we expected.
 The process took 7 months and 16 days, significantly slower than expected.
@@
 We sent a third version on February 24, 2025, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
-Figure \ref{fig:timeline} illustrates the timeline of the patch submission.
+%Figure \ref{fig:timeline} illustrates the timeline of the patch submission.
 
-After the initial reply from the maintainers, we observed that the review process became more responsive.
+%After the initial reply from the maintainers, we observed that the review process became more responsive.
 We hypothesize that once a patch receives initial attention, it is more closely tracked by maintainers.
@@
 
-We discussed the Display driver with our point of contact at AMD to understand the reason for the slow patch submission process.
-We learned that some of the code in the driver is shared across all operating systems that support the implemented GPUs.
+%We discussed the Display driver with our point of contact at AMD to understand the reason for the slow patch submission process. We learned that
+
+We submitted the initial patch: the maintainers only asked for minor changes to align with code style, correct license use, and best practices that were initially unknown.
+We submitted a second version: the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
+We sent a third version, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
+%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.
+
+During this contribution process, we learned that some of the code in the driver is shared across all operating systems that support the implemented GPUs.
 This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors.
@@
 This approach helps save a significant amount of time and effort.
-
-Regarding the patch changes requested for the second version, the maintainers only asked for minor changes to align with code style, correct license use, and best practices that were initially unknown.
-For the change requested on the third version, the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
-%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.
 
@@
 
-\begin{table*}
+\begin{table}
 \centering
 \caption{Refactor methods used and metrics of the students' mitigations.}
-\begin{tabular}{ | c | m{15mm} | m{20mm} | m{30mm} | m{30mm} | m{20mm} | }
+\begin{tabular}{ |c |c |c |c |c | c| }
 \hline
-\textbf{Group} & \textbf{Subsystem} & \textbf{Similarity} & \textbf{Patch Status} & \textbf{Refactoring Method} & \textbf{Code Diff} \\
+\textbf{G} & \textbf{SS} & \textbf{SML} & \textbf{Patch Status} & \textbf{Method} & \textbf{Diff} \\
 \hline
@@
 1 & IIO & 100\% & Dropped (v2) & Undefined & +7/-14 \\ \hline
-2 & IIO & 100\% & Dropped (v1) & Parameterize method & +23/-27 \\ \hline
-3 & IIO & 100\% & Accepted (v1) & Parameterize method & +12/-30 \\ \hline
+2 & IIO & 100\% & Dropped (v1) & Parameterize & +23/-27 \\ \hline
+3 & IIO & 100\% & Accepted (v1) & Parameterize & +12/-30 \\ \hline
 4 & IIO & 100\% & Accepted (v3) & Undefined & +2/-7 \\ \hline
-5 & IIO & 90\% & Accepted (v1) & Parameterize method & +17/-70 \\ \hline
-6 & IIO & 90\% & Dropped (v1) & Parameterize method & +95/-92 \\ \hline
-7 & IIO & 90\% & Accepted (v3) & Extract method & +20/-36 \\ \hline
-8 & IIO & 90\% & Accepted (v4) & Extract method & +20/-16 \\ \hline
-9 & IIO & 90\% & Accepted (v6) -- Partially & Parameterize method & +149/-243 \\ \hline
+5 & IIO & 90\% & Accepted (v1) & Parameterize & +17/-70 \\ \hline
+6 & IIO & 90\% & Dropped (v1) & Parameterize & +95/-92 \\ \hline
+7 & IIO & 90\% & Accepted (v3) & Extract & +20/-36 \\ \hline
+8 & IIO & 90\% & Accepted (v4) & Extract & +20/-16 \\ \hline
+9 & IIO & 90\% & Accepted* (v6) & Parameterize & +149/-243 \\ \hline
 10 & IIO & 90\% & Dropped (v1) & Undefined & +32/-58 \\ \hline
-11 & IIO & 100\% & Dropped (v1) & Inline method & +2/-12 \\ 
+11 & IIO & 100\% & Dropped (v1) & Inline & +2/-12 \\ 
    & AMD & 90\% & Accepted (v2) & Undefined & +90/-489 \\ \hline
@@
 \label{tab:stu}
-\end{table*}
+\end{table}
 
@@
 The ethnographic studies, which included participant observation and experimentation with students, offered a realistic perspective on the opportunities and significant challenges in mitigating code duplications found by the \texttt{anonymous-tool} tool.
-The tool was effectively used by first-time contributors to the Linux kernel to identify duplications and submit patches~\footnote{To avoid violating the double-anonymized review process, we will add the tool repository, patches, and artifact links in the final camera-ready version of the paper, if it is accepted for publication.}.
+The tool was effectively used by first-time contributors to the Linux kernel to identify duplications and submit patches.
+%TODO: disponibilizar artefato da mesma forma que mandamos para a ICSE Education Track~\footnote{To avoid violating the double-anonymized review process, we will add the tool repository, patches, and artifact links in the final camera-ready version of the paper, if it is accepted for publication.}
 

### 2025-10-20 paper/techdebt2026/techdebt.tex
techdebt2025: Updates abstract
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 Our ethnographic approach, including participant and non-participant observations, demonstrated that addressing duplications is viable for lowering the contribution barrier.
-This point is evidenced by the main author and 7 of 11 (64\%) student groups having their patches accepted into the kernel, collectively removing 991 lines of duplicated code.
-Moreover, the study also reveals a more complex reality beyond simple clone removal: an analysis of maintainer feedback on both accepted and rejected patches highlights a nuanced understanding of code quality within the Linux kernel community, where the benefits of eliminating duplication are carefully weighed against factors such as readability, the introduction of new abstractions, and the specific context of the code.
+This point is evidenced by the main author and 7 of 11 (64\%) student groups having accepted contributions in the Linux project, collectively removing 991 lines of duplicated code.
+Nevertheless, the study reveals a more intriguing and complex reality beyond mindlessly removing clones.
+An analysis of maintainer feedback on both accepted and rejected contributions highlights a nuanced understanding of technical debt due to code duplication within the Linux kernel community, where the benefits of eliminating duplications are carefully weighed against factors such as readability, the introduction of new abstractions, and the specific code context.
 \end{abstract}

### 2025-10-20 paper/techdebt2026/techdebt.tex
techdebt2025: reorganizes methods and results
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 
-\section{Anonymous-tool}
+\section{Methods and Tool}
 
+We employ a multi-method research approach. We developed and evaluated a proposed tool using an approach from the literature, comparing it against the BigCloneBench dataset~\cite{bigclonebench}.
+%
+Given the validation of the tool capabilities, we proceed to understand the viability 
+of mitigating the code duplication found by the tool. We conducted an ethnographic study 
+using a participant-observation methodology to understand its viability.
+%
+In the first phase, the main author explored the process of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to interact with the Linux kernel community and get their feedback.
+In the second phase, we proposed to newcomers to mitigate the code duplications found by the tool in the IIO subsystem and the AMD Display driver and pass the process of sending the mitigations as a PATCH.
+
+%Uniformizar este entendimento no texto:
+
+%Fase 1:Complete Participant
+%O pesquisador não estava observando outros desenvolvedores passarem pelo processo; ele se tornou o desenvolvedor.
+Ele se imergiu totalmente no papel de um contribuidor do kernel, enfrentando os mesmos desafios, normas sociais e barreiras técnicas que qualquer membro da comunidade enfrentaria.
+% O objetivo aqui é ganhar uma visão "êmica" (de dentro), entendendo a cultura e o processo por meio da experiência vivida.
+Isso é, por vezes, também chamado de autoetnografia.
+Ele não estava "fingindo", ele estava genuinamente tentando contribuir para entender as dores.
+
+%Fase 2: Participant-as-Observer
+%O papel do pesquisador (no caso pesquisadores) mudou fundamentalmente. Ele deixou de ser o ator principal (o contribuidor) e passou a ser o facilitador (mentor).
+%O foco dos dados coletados não é mais a autoexperiência do pesquisador, mas sim a experiência dos estudantes.
+%A participação ainda é alta (selecionar, direcionar, revisar), o que influencia ativamente o cenário.
+No entanto, a identidade principal perante os participantes (os estudantes) é a de pesquisador/mentor, e o objetivo é observar o processo deles.
+
+\subsection{Tool}
 We developed \texttt{anonymous-tool}, a command-line tool designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation.
-Released under the MIT license, \texttt{anonymous-tool} employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
+Released under the LGPL license, it employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
 The first stage, the Preprocessor, analyzes the entire codebase, parses C functions, and stores identified duplicates in a database.
@@
 The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying function-level duplications to address.
-Figure~\ref{fig:diagrama} illustrates the tool's architecture.
 The tool uses a text similarity method that applies TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.
 
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.275]{diagrama_mestrado}
-\caption{Architecture diagram with the tool components
-}
-\label{fig:diagrama}
-\end{figure}
-
-\section{Participant Observation Study}
+\subsection{Complete Participant Observation Study}
 \label{sec:participant}
@@
 We have chosen to approach the biggest duplications as proof of concept so that we can make significant impacts.
-
-The first duplicated function pair chosen is function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
-The second duplicated function pair is function \textit{phy\_id\_to\_atom} in file
-\textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
-Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
-\textit{phy\_id\_to\_atom} will be referred to as mitigation 2.
 
@@
 
-\subsection{Results}
-\label{subsec:participantresults}
-
-Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
-files and lines changed and the refactoring method used. The following sections present our experience and learning for each mitigation.
-
-\begin{table}[ht]
-\centering
-\caption{Refactor methods used and mitigation metrics in the participant-observation experiment.}
-\begin{tabular}{ | m{12mm} | m{10mm} | m{10mm} | m{14mm} | m{20mm} | }
-
-\hline
-
-\textbf{Mitigation} & \textbf{Files changed} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
-\\ \hline
-
-1 & 13 & +224 & -753 & Parameterize Method, Extract Method  \\ \hline
-2 & 8 & +114 & -520 & None \\ \hline
-
-\hline
-\end{tabular}
-
-\label{tab:patch}
-\end{table}
-
-\subsubsection{Mitigation 1}
-
-In the systematic approach, we found that the \textit{offset\_to\_id} function 
-exists in 10 code files, enclosing multiple GPU architectures. The duplicated 
-functions are not exactly equal, as some minimal specific logic is applied to each GPU architecture. To address this issue, we used the extract method to 
-break the function into smaller parts and the parameterized method to address 
-the specific logic for the GPUs.
-
-We reached a point where we had a scratch on how to mitigate the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver.
-We found that the configuration files on the AMD Display driver are designed to be imported at compilation time using \textit{\#define} macros.
-The configuration files' design choice makes refactoring duplications on generic approaches trickier, as refactoring code that depends on these files requires significant refactoring of the configuration files' design and deep knowledge of the codebase, which we, as first-time contributors, do not have.
-Thus, we opted not to continue investigating this mitigation.
-
-\subsubsection{Mitigation 2}
-
-Using the systematic approach, we found that the \textit{phy\_id\_to\_atom} function exists in five code files, and two other functions are duplicated 
-in those five code files. All functions across the files were exactly equal, 
-so there was no need to apply the refactoring methods presented in the literature. 
-The refactoring was resumed to create a generic library and fix the compilation targets.
-
-The code files in the second mitigation do not depend on configuration files. Thus, 
-we did not face the same issues during the first mitigation. Refactoring the code 
-to mitigate duplications reached a point we judged was good enough to submit as a 
-patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
-the refactoring to the driver maintainers as a patch, documenting our findings and experiences 
-on the process.
-
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[scale=0.325]{timeline_patch}
-%\caption{Timeline of the PATCH submission process}
-%\label{fig:timeline}
-%\end{figure}
-
-%The experience of submitting a patch was not as we expected.
-The process took 7 months and 16 days, significantly slower than expected.
-We initially submitted the patch on August 9, 2024, but we did not receive replies.
-Thus, we had to resend it on October 9, 2024.
-We received an initial response on November 3, 2024, requesting minimal changes.
-We submitted a second version with the requested changes on November 16, 2024.
-We did not receive replies initially and will need to resend the patch on February 7, 2025.
-We received a reply on February 11, 2025, asking for additional changes.
-We sent a third version on February 24, 2025, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
-%Figure \ref{fig:timeline} illustrates the timeline of the patch submission.
-
-%After the initial reply from the maintainers, we observed that the review process became more responsive.
-We hypothesize that once a patch receives initial attention, it is more closely tracked by maintainers.
-Another contributing factor to the initial delay could be the maintainers' workload at the time of the first submission.
-
-%We discussed the Display driver with our point of contact at AMD to understand the reason for the slow patch submission process. We learned that
-
-We submitted the initial patch: the maintainers only asked for minor changes to align with code style, correct license use, and best practices that were initially unknown.
-We submitted a second version: the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
-We sent a third version, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
-%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.
-
-During this contribution process, we learned that some of the code in the driver is shared across all operating systems that support the implemented GPUs.
-This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors.
-Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering.
-One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others.
-This approach helps save a significant amount of time and effort.
-
-\section{Non-participant Observation Study}
+\subsection{Participant-as-Observer Study}
 
@@
 
-\subsection{Overview of the submitted patches}
+\section{Results from Participant Observation}
+
+%Results part 1 - AMD/Luan
+%Ver como melhor iniciar.
+The first duplicated function pair chosen is function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
+The second duplicated function pair is function \textit{phy\_id\_to\_atom} in file
+\textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
+Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
+\textit{phy\_id\_to\_atom} will be referred to as mitigation 2.
+
+Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
+files and lines changed and the refactoring method used. The following sections present our experience and learning for each mitigation.
+
+\begin{table}[ht]
+\centering
+\caption{Refactor methods used and mitigation metrics in the participant-observation experiment.}
+\begin{tabular}{ | m{12mm} | m{10mm} | m{10mm} | m{14mm} | m{20mm} | }
+
+\hline
+
+\textbf{Mitigation} & \textbf{Files changed} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
+\\ \hline
+
+1 & 13 & +224 & -753 & Parameterize Method, Extract Method  \\ \hline
+2 & 8 & +114 & -520 & None \\ \hline
+
+\hline
+\end{tabular}
+
+\label{tab:patch}
+\end{table}
+
+Firt, we found that the \textit{offset\_to\_id} function 
+exists in 10 code files, enclosing multiple GPU architectures. The duplicated 
+functions are not exactly equal, as some minimal specific logic is applied to each GPU architecture. To address this issue, we used the extract method to 
+break the function into smaller parts and the parameterized method to address 
+the specific logic for the GPUs.
+
+We reached a point where we had a scratch on how to mitigate the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver.
+We found that the configuration files on the AMD Display driver are designed to be imported at compilation time using \textit{\#define} macros.
+The configuration files' design choice makes refactoring duplications on generic approaches trickier, as refactoring code that depends on these files requires significant refactoring of the configuration files' design and deep knowledge of the codebase, which we, as first-time contributors, do not have.
+Thus, we opted not to continue investigating this mitigation.
+
+Second, we found that the \textit{phy\_id\_to\_atom} function exists in five code files, and two other functions are duplicated 
+in those five code files. All functions across the files were exactly equal, 
+so there was no need to apply the refactoring methods presented in the literature. 
+The refactoring was resumed to create a generic library and fix the compilation targets.
+
+The code files in the second mitigation do not depend on configuration files. Thus, 
+we did not face the same issues during the first mitigation. Refactoring the code 
+to mitigate duplications reached a point we judged was good enough to submit as a 
+patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
+the refactoring to the driver maintainers as a patch, documenting our findings and experiences 
+on the process.
+
+We submitted the initial patch: the maintainers only asked for minor changes to align with code style, correct license use, and best practices that were initially unknown.
+We submitted a second version: the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
+We sent a third version, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
+%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.
+
+During this contribution process, we learned that some of the code in the driver is shared across all operating systems that support the implemented GPUs.
+This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors.
+Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering.
+One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others.
+This approach helps save a significant amount of time and effort.
+
+\section{Results from Non-Participant Observation}
 
@@
 
-%\begin{figure}
-%\centering
-%\includegraphics[scale=0.8]{group_10_reply}
-%\caption[Maintainers' Feedback on Group 10 Patch]{
-% Maintainers' Feedback on Group Ten's Patch
-%}
-%\label{fig:group_10_reply}
-%\end{figure}
-
 Groups 7 and 8 used the extract method to approach the mitigations.
@@
 
-\subsection{Statistics of the contributions}
+%\subsection{Statistics of the contributions}
 

### 2025-10-21 paper/techdebt2026/techdebt.tex
techdebt2025: organizes sections from Paulo and David's brainstorm
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 \usepackage{seqsplit}
+\usepackage{todonotes}
 
@@
 %
-In the first phase, the main author explored the process of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to interact with the Linux kernel community and get their feedback.
-In the second phase, we proposed to newcomers to mitigate the code duplications found by the tool in the IIO subsystem and the AMD Display driver and pass the process of sending the mitigations as a PATCH.
+In the first phase, we explored the process of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to interact with the Linux kernel community and get their feedback (\textit{complete participant observation})[REF].
+%TODO: aqui já podemos falar da disciplinas bem brevemente
+In the second phase, we proposed to students to mitigate the code duplications found by the tool in the IIO subsystem and the AMD Display driver and pass the process of sending the mitigations as a PATCH (\textit{participant-as-observer}).
+%TODO: Aqui vamos precisa cita que rapidamente a metoria para poder falar de Participant-as-Observer
 
@@
 \subsection{Tool}
+\todo[inline]{Luan: revisar e colocar a justificativa sobre a necessidade de desenvolver a ArKanjo para o Kernel; se der, colocar uma ou duas linhas sobre a validacao.}
 We developed \texttt{anonymous-tool}, a command-line tool designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation.
@@
 
-\section{Results from Participant Observation}
+\section{Results}
+%TODO: ver como inicar melhor aqui ...
 
-%Results part 1 - AMD/Luan
-%Ver como melhor iniciar.
-The first duplicated function pair chosen is function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
-The second duplicated function pair is function \textit{phy\_id\_to\_atom} in file
+In the \textbf{complete participant observation} phase, the first duplicated function pair chosen was function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
+The second duplicated function pair was function \textit{phy\_id\_to\_atom} in file
 \textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
@@
 \textit{phy\_id\_to\_atom} will be referred to as mitigation 2.
+
+%TODO: ver se realmente queremos usar mitigation (ou contribution é melhor)
 
@@
 
-During this contribution process, we learned that some of the code in the driver is shared across all operating systems that support the implemented GPUs.
+%TODO: deve ficar claro que isto aqui, mesmo que não dito explicitamente pelo mantenedores, é o resultado da Complete Participant Observation
+\noindent\textbf{Takeaway 1:} Some of the code in the driver is shared across all operating systems that support the implemented GPUs.
 This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors.
-Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering.
+Additionally, \textbf{we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering}.
 One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others.
 This approach helps save a significant amount of time and effort.
+\todo[inline]{Luan, temos alguma evidência nas suas trocas com os mantenedores sobre isso? ou foi algo que veio mais das conversas com o Siqueira?}
 
-\section{Results from Non-Participant Observation}
-
-Table \ref{tab:stu} summarizes the student groups' mitigation approaches.
+%TODO: ver se transicao estah boa; ou se eh melhor ter subsecoes
+Regarding our analysis from the \textbf{participant-as-Observer} phase, Table~\ref{tab:stu} summarizes the student groups' mitigation approaches.
 It contains the information about the chosen subsystem and the similarity threshold for groups that opted for the first approach.
@@
 
-\begin{table}
+\begin{table}[ht]
 \centering

### 2025-10-22 paper/techdebt2026/techdebt.tex
luan changes 22/10

@@
 It enables researchers to gain insights into values, beliefs, and practices from the perspective of community members~\cite{ethnosoft}.
-In this study, we first adopted a participant observation approach, where the main author actively engaged in mitigating code duplications while documenting the process and outcomes.
-Later, we followed a non-participant observation involving students tasked with resolving duplications identified by the tool.
+In this study, we first adopted a complete participant observation approach, where the main author actively engaged in mitigating code duplications while documenting the process and outcomes.
+Later, we conducted a partipant-as-observer study involving students tasked with resolving duplications identified by the tool.
 These combined studies showed that \texttt{anonymous-tool} can collaborate to reduce the entry barrier for new contributors.
@@
 
-We employ a multi-method research approach. We developed and evaluated a proposed tool using an approach from the literature, comparing it against the BigCloneBench dataset~\cite{bigclonebench}.
+We employ a multi-method research approach. We developed and evaluated a proposed tool using an approach from the literature, 
+comparing it against the BigCloneBench dataset~\cite{bigclonebench}.
 %
@@
 %TODO: aqui já podemos falar da disciplinas bem brevemente
-In the second phase, we proposed to students to mitigate the code duplications found by the tool in the IIO subsystem and the AMD Display driver and pass the process of sending the mitigations as a PATCH (\textit{participant-as-observer}).
+In the second phase, during an university course, we proposed to students to mitigate the code duplications found by the tool in the IIO subsystem and the AMD Display driver and pass the process of sending the mitigations as a PATCH (\textit{participant-as-observer}) as a course activity while we guided them through the process.
 %TODO: Aqui vamos precisa cita que rapidamente a metoria para poder falar de Participant-as-Observer
@@
 \subsection{Tool}
+
 \todo[inline]{Luan: revisar e colocar a justificativa sobre a necessidade de desenvolver a ArKanjo para o Kernel; se der, colocar uma ou duas linhas sobre a validacao.}
-We developed \texttt{anonymous-tool}, a command-line tool designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation.
-Released under the LGPL license, it employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
-The first stage, the Preprocessor, analyzes the entire codebase, parses C functions, and stores identified duplicates in a database.
-The second stage, the Query Responder, uses this database to retrieve information about the identified duplications quickly.
+
+Academic research on code duplication often focuses on pairwise comparison of code artifacts, which is not directly applicable to comprehensive codebase analyses.
+Other existing free software tools we explored in practice suffered from limited functionality.
+Thus, we developed \texttt{anonymous-tool}, a command-line tool designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation. 
+Released under the LGPL license, it employs a two-stage architecture that separates the computationally intensive task of 
+finding duplicates from the process of querying the results.
+The first stage that analyzes the entire codebase, parses C functions, and stores identified duplicates in a database.
+The second stage, uses this database to retrieve information about the identified duplications quickly.
 %The tool is available at \textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.
@@
 \texttt{Anonymous-tool} is designed for the specific challenges of the kernel, with a focus on scalability to handle its size and prioritizing duplications that impact maintainability. 
-The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying function-level duplications to address.
-The tool uses a text similarity method that applies TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.
+The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying 
+function-level duplications to address. The tool uses a text similarity method that applies TF-IDF vector embedding, 
+implemented with the Gensim library~\cite{gensim}, and computes cosine distance similarity to detect clones.
+
+Validation of the tool on the BigCloneBench dataset~\cite{bigclonebench} showed that it is capable of detecting the majority of Type-1 and Type-2 duplications (100\% of Type-1 and 85\% of Type-2 with a cosine similarity threshold of 0.9),
+which is sufficient for conducting our ethnographic studies.
 
@@
 
-A core dynamic of the course (concurrently for undergraduate and postgraduate studies) is to have students make a practical contribution to a free software project.
-In the first semester of 2025, the professor teaching the course, who is also the advisor of this work, identified an opportunity to leverage the \texttt{anonymous-tool} tool within this existing pedagogical structure.
-As a contribution option, the professor proposed that students mitigate code duplications found in the Linux kernel.
-The professor drove this initiative as a practical course activity, and the author of this thesis was not involved in using the tool to identify duplications or in developing approaches to mitigate them.
-The tasks were carried out entirely by the course staff and the students.
+A core dynamic of the university course where we conducted the participant-as-observer study is to have students make a practical contribution to a free software project.
+In the first semester of 2025, we identified an opportunity to leverage the anonymous-tool within this existing pedagogical structure.
+As a contribution option, we proposed that students mitigate code duplications found in the Linux kernel as one possible way to contribute to a free software project.
 
 The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of 
-two or three members. To assist students in this task, the professor and teaching assistants prepare multiple 
+two or three members. To assist students in this task, we prepare multiple 
 alternatives for contributing to free software projects, including simpler contributions such as 
-updating documentation or fixing code style issues. In this context, two new options related to the 
-\texttt{anonymous-tool} were proposed alongside the others.
+updating documentation or fixing code style issues. In this context, two new options related to migitating
+duplications in the Linux kernel were proposed alongside the others.
 
-For the first anonymous-tool-related alternative, a teaching assistant ran the tool on the IIO subsystem twice, creating two lists of duplicated function pairs, one with a 100\% similarity threshold and the other with a 90\% similarity threshold.
-Then, the following process happened to both lists:
+For the first anonymous-tool-related alternative, we ran the tool on the IIO subsystem, creating a list of duplicated function pair.
+Then, the following process happened to the list to prepare more actionable duplications to the students:
 
 \begin{enumerate}
-    \item The teaching assistant filtered the lists to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different devices may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
-    \item The teacher assistant removed duplications that were too short and would not represent a significant change for a patch application;
-    \item The teacher assistant manually ranked code duplications and marked some duplications with insights based on personal opinion and perceptions over the remaining entries.
+    \item We filtered the list to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different devices may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
+    \item We removed duplications that were too short and would not represent a significant change for a patch application;
+    \item We manually ranked code duplications and marked some duplications with insights based on personal opinion and perceptions over the remaining entries.
 This was done to guide students so that they could approach the technical debt with a greater chance of having their contribution merged into the code.
@@
 
-In the \textbf{complete participant observation} phase, the first duplicated function pair chosen was function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}.
-The second duplicated function pair was function \textit{phy\_id\_to\_atom} in file
-\textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
-Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
-\textit{phy\_id\_to\_atom} will be referred to as mitigation 2.
+In the \textbf{complete participant observation} study, we chose two function pair found on AMD Display Driver to approach and apply the systematic approach, 
+which we will denote as mitigation 1 and mitigation 2.
 
@@
 
-Firt, we found that the \textit{offset\_to\_id} function 
+Firt, we found that the duplicated function from mitigation 1
 exists in 10 code files, enclosing multiple GPU architectures. The duplicated 
@@
 
-Second, we found that the \textit{phy\_id\_to\_atom} function exists in five code files, and two other functions are duplicated 
-in those five code files. All functions across the files were exactly equal, 
+Second, we found that the duplicated function from mitigation 2 exists in five code files, 
+and two other functions are duplicated in those five code files. All functions across the files were exactly equal, 
 so there was no need to apply the refactoring methods presented in the literature. 
@@
 This approach helps save a significant amount of time and effort.
-\todo[inline]{Luan, temos alguma evidência nas suas trocas com os mantenedores sobre isso? ou foi algo que veio mais das conversas com o Siqueira?}
+\todo[inline]{Luan, temos alguma evidência nas suas trocas com os mantenedores sobre isso?
+ou foi algo que veio mais das conversas com o Siqueira?
+Resposta: Não, foi tirado apenas das conversas com o Siqueira, podemos remover ou tem o paper que o Gregório citou que podemos mencionar?
+Que menciona que duplicações podem ser bem vistas em drivers.}
 

### 2025-10-22 paper/techdebt2026/techdebt.tex
techdebt2025: adjustments and minor review
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 This paper presents a multimethod ethnographic study investigating the socio-technical dynamics of contributing to the Linux kernel by mitigating identified redundancies.
-To support this investigation, we developed \texttt{anonymous-tool}, a command-line utility for detecting function-level duplications, offering a concrete entry point for new contributors.
+To support this investigation, we developed a command-line utility tool for detecting function-level duplications, offering a concrete entry point for new contributors.
 Our ethnographic approach, including participant and non-participant observations, demonstrated that addressing duplications is viable for lowering the contribution barrier.
@@
 
-To address the code duplication issue, we developed \texttt{anonymous-tool}, a command-line tool designed to support Linux kernel maintenance by detecting and analyzing function-level duplications.
+To address the code duplication issue, we developed a command-line tool designed to support Linux kernel maintenance by detecting and analyzing function-level duplications.
 This paper presents a multimethod ethnographic study to validate the tool capabilities in identifying duplications and supporting contributions to the Linux kernel by mitigating redundant code.
@@
 Later, we conducted a partipant-as-observer study involving students tasked with resolving duplications identified by the tool.
-These combined studies showed that \texttt{anonymous-tool} can collaborate to reduce the entry barrier for new contributors.
-The student groups and the main author used the tool to guide their initial contributions to the Linux kernel.
+These combined studies showed that our tool can collaborate to reduce the entry barrier for new contributors.
+The student groups (and we) used the tool to guide their initial contributions to the Linux kernel.
 
@@
 
-\section{Methods and Tool}
+\section{Research Strategy}
 
-We employ a multi-method research approach. We developed and evaluated a proposed tool using an approach from the literature, 
-comparing it against the BigCloneBench dataset~\cite{bigclonebench}.
+We employ a multi-method research approach. We developed and evaluated a proposed tool using an approach from the literature~[X].
 %
 Given the validation of the tool capabilities, we proceed to understand the viability 
-of mitigating the code duplication found by the tool. We conducted an ethnographic study 
-using a participant-observation methodology to understand its viability.
+of reducing the code duplication found by the tool. We conducted an ethnographic study 
+using a participant-observation methodology to understand its viability~\cite{}.
 %
-In the first phase, we explored the process of mitigating a code duplication found in the AMD Display driver and sent it as a PATCH to interact with the Linux kernel community and get their feedback (\textit{complete participant observation})[REF].
-%TODO: aqui já podemos falar da disciplinas bem brevemente
-In the second phase, during an university course, we proposed to students to mitigate the code duplications found by the tool in the IIO subsystem and the AMD Display driver and pass the process of sending the mitigations as a PATCH (\textit{participant-as-observer}) as a course activity while we guided them through the process.
-%TODO: Aqui vamos precisa cita que rapidamente a metoria para poder falar de Participant-as-Observer
+In the first phase, we explored the process of reducing code duplication found in the AMD Display driver and sent it as a patch to interact with the Linux kernel community and get their feedback (\textit{complete participant observation}~\cite{}).
+%
+In the second phase, during a university course, we proposed to students to reduce the code duplications found by our tool in the IIO subsystem and the AMD Display driver and pass the process of sending the contribution as a patch (\textit{participant-as-observer}~\cite{}) as a course activity while we guided them through the process.
 
-%Uniformizar este entendimento no texto:
-
+%Uniformizar este entendimento abaixo ao longo do texto:
 %Fase 1:Complete Participant
@@
 
-\todo[inline]{Luan: revisar e colocar a justificativa sobre a necessidade de desenvolver a ArKanjo para o Kernel; se der, colocar uma ou duas linhas sobre a validacao.}
-
 Academic research on code duplication often focuses on pairwise comparison of code artifacts, which is not directly applicable to comprehensive codebase analyses.
-Other existing free software tools we explored in practice suffered from limited functionality.
-Thus, we developed \texttt{anonymous-tool}, a command-line tool designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation. 
-Released under the LGPL license, it employs a two-stage architecture that separates the computationally intensive task of 
-finding duplicates from the process of querying the results.
-The first stage that analyzes the entire codebase, parses C functions, and stores identified duplicates in a database.
-The second stage, uses this database to retrieve information about the identified duplications quickly.
+Other existing free software tools we explored in practice suffered from limited functionality~[X].
+Thus, we developed a command-line tool~[X] designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation. 
+Released under the LGPL license, it employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
+The first stage analyzes the entire codebase, parses C functions, and stores identified duplicates in a database.
+The second stage uses this database to quickly retrieve information about the identified duplications.
 %The tool is available at \textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.
 
-\texttt{Anonymous-tool} is designed for the specific challenges of the kernel, with a focus on scalability to handle its size and prioritizing duplications that impact maintainability. 
-The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying 
-function-level duplications to address. The tool uses a text similarity method that applies TF-IDF vector embedding, 
-implemented with the Gensim library~\cite{gensim}, and computes cosine distance similarity to detect clones.
+Our tool is designed for the specific challenges of the kernel, with a focus on scalability to handle its size and prioritizing duplications that impact maintainability. 
+The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying function-level duplications to address.
+The tool uses a text similarity method that applies TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine distance similarity to detect clones.
 
-Validation of the tool on the BigCloneBench dataset~\cite{bigclonebench} showed that it is capable of detecting the majority of Type-1 and Type-2 duplications (100\% of Type-1 and 85\% of Type-2 with a cosine similarity threshold of 0.9),
-which is sufficient for conducting our ethnographic studies.
+We compared our tool against the BigCloneBench dataset~\cite{bigclonebench}.
+This validation showed that it can detect the majority of Type-1 and Type-2 duplications (100\% of Type-1 and 85\% of Type-2 with a cosine similarity threshold of 0.9)~[X], demonstrating sufficient accuracy for conducting our ethnographic studies.
 
@@
 
-To investigate whether we can use the \texttt{anonymous-tool} tool to find ways to contribute to the Linux kernel, we conducted a participant observation experiment in the second semester of 2024, where we acted as first-time contributors to the AMD Display driver, collecting artifacts of our experience in the process.
+To investigate whether we can use our tool to find ways to contribute to the Linux kernel, we conducted a complete participant observation experiment, where we acted as first-time contributors to the AMD Display driver, collecting artifacts of our experience in the process.
 
-In the AMD Display driver codebase, we executed the \texttt{anonymous-tool}.
+In the AMD Display driver codebase, we executed our tool.
 We manually analyzed the biggest duplications regarding the number of lines and found two duplicated function pairs that we judged interesting to try to mitigate.
@@
 
-For each duplicated function pair, for each function in the pair, we used the \texttt{anonymous-tool} to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair.
+For each duplicated function pair, for each function in the pair, we used our tool to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair.
 This strategy resulted in a collection of duplicated functions and corresponding code files.
@@
 A core dynamic of the university course where we conducted the participant-as-observer study is to have students make a practical contribution to a free software project.
-In the first semester of 2025, we identified an opportunity to leverage the anonymous-tool within this existing pedagogical structure.
+In the first semester of 2025, we identified an opportunity to leverage our tool within this existing pedagogical structure.
 As a contribution option, we proposed that students mitigate code duplications found in the Linux kernel as one possible way to contribute to a free software project.
@@
 
-For the first anonymous-tool-related alternative, we ran the tool on the IIO subsystem, creating a list of duplicated function pair.
+For the first our tool-related alternative, we ran the tool on the IIO subsystem, creating a list of duplicated function pair.
 Then, the following process happened to the list to prepare more actionable duplications to the students:
@@
 For the second alternative, students were offered an experience similar to our initial participant 
-observation. In this model, the students were responsible for the entire workflow: executing the 
-\texttt{anonymous-tool} on the AMD display driver, independently analyzing the results to identify a duplication 
+observation. In this model, the students were responsible for the entire workflow: executing our tool on the AMD display driver, independently analyzing the results to identify a duplication 
 to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.
 
-Of the students who chose to work on the \texttt{anonymous-tool}-related tasks, 23 (16 undergraduate and 7 graduate) 
+Of the students who chose to work on our tool-related tasks, 23 (16 undergraduate and 7 graduate) 
 formed 11 groups to pursue the first alternative.
@@
 This approach helps save a significant amount of time and effort.
-\todo[inline]{Luan, temos alguma evidência nas suas trocas com os mantenedores sobre isso?
-ou foi algo que veio mais das conversas com o Siqueira?
-Resposta: Não, foi tirado apenas das conversas com o Siqueira, podemos remover ou tem o paper que o Gregório citou que podemos mencionar?
-Que menciona que duplicações podem ser bem vistas em drivers.}
+\todo[inline]{...Foi tirado apenas das conversas com o Siqueira; tem o paper que o Gregório citou que menciona que duplicações podem ser bem vistas em drivers; isto um resultado/conclusão da Complete Participant Observation; e vamos citar que confirmamos o que está no paper Cloning Considered Harmful (Luan, acha que consegue fechar essa parte assim?)}
 
@@
 
-Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach duplications found by the \texttt{anonymous-tool} and become contributors to the kernel.
+Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach duplications found by our tool and become contributors to the kernel.
 We observed that not all duplications
@@
 For a 90\% value, the result was a total of +333/-515 with an average of +55.5/-85.83.
-The 90\% value enabled \texttt{anonymous-tool} to identify less strict duplication cases, and, therefore, code segments with a greater extension, compared to the 100\% similarity entries, which were shorter segments, causing a difference in the average number of modified lines.
+The 90\% value enabled our tool to identify less strict duplication cases, and, therefore, code segments with a greater extension, compared to the 100\% similarity entries, which were shorter segments, causing a difference in the average number of modified lines.
 
-Finally, the AMD contribution from Group 11, which involves the complete use of the \texttt{anonymous-tool} toolchain, got a completely different outcome.
+Finally, the AMD contribution from Group 11, which involves the complete use of our developed tool toolchain, got a completely different outcome.
 With a total of 90 lines inserted and 489 lines removed, there are some points to be considered for the success and high modification extension of the contribution:
@@
     \item The AMD Display subsystem has a more centralized code structure and maintainer hierarchy, enabling contributors to tackle code duplications across a broader scope and multiple files, which happened to Group 11's contribution;
-    \item The complete control over the process of finding code duplication gave Group 11 the opportunity to identify and reason better over the duplications found by \texttt{anonymous-tool}, enabling a more strategic approach to searching for promising patch possibilities.
+    \item The complete control over the process of finding code duplication gave Group 11 the opportunity to identify and reason better over the duplications found by our tool, enabling a more strategic approach to searching for promising patch possibilities.
 \end{itemize}
@@
 
-The ethnographic studies, which included participant observation and experimentation with students, offered a realistic perspective on the opportunities and significant challenges in mitigating code duplications found by the \texttt{anonymous-tool} tool.
+The ethnographic studies, which included participant observation and experimentation with students, offered a realistic perspective on the opportunities and significant challenges in mitigating code duplications found by our tool.
 The tool was effectively used by first-time contributors to the Linux kernel to identify duplications and submit patches.

### 2025-10-22 paper/techdebt2026/techdebt.tex
techdebt2025: adjustments (from the authors' meeting)
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 
+%TODO: temos que citar o conceito de DRY aqui ... (Book: Pragmatic Programmer, The: From Journeyman to Master)
+% We feel that the only way to develop software reliably, and to make our developments easier to understand and maintain, is to follow what we call the DRY principle: Every piece of knowledge must have a single, unambiguous, authoritative representation within a system.
 The detection of code duplication, or \textit{code clones}, has been a research topic for decades~\cite{firstman}.
@@
 
-In the \textbf{complete participant observation} study, we chose two function pair found on AMD Display Driver to approach and apply the systematic approach, 
-which we will denote as mitigation 1 and mitigation 2.
+In the \textbf{complete participant observation} study, we chose two function pair found on AMD Display Driver to approach and apply the systematic approach ...
 
-%TODO: ver se realmente queremos usar mitigation (ou contribution é melhor)
-
-Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
-files and lines changed and the refactoring method used. The following sections present our experience and learning for each mitigation.
-
-\begin{table}[ht]
-\centering
-\caption{Refactor methods used and mitigation metrics in the participant-observation experiment.}
-\begin{tabular}{ | m{12mm} | m{10mm} | m{10mm} | m{14mm} | m{20mm} | }
-
-\hline
-
-\textbf{Mitigation} & \textbf{Files changed} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
-\\ \hline
-
-1 & 13 & +224 & -753 & Parameterize Method, Extract Method  \\ \hline
-2 & 8 & +114 & -520 & None \\ \hline
-
-\hline
-\end{tabular}
-
-\label{tab:patch}
-\end{table}
+%TODO (Luan): nao terah mitigation 1 e 2, apenas a contribuicao enviada mesmo.
 
@@
 %TODO: deve ficar claro que isto aqui, mesmo que não dito explicitamente pelo mantenedores, é o resultado da Complete Participant Observation
-\noindent\textbf{Takeaway 1:} Some of the code in the driver is shared across all operating systems that support the implemented GPUs.
+\noindent\textbf{Takeaway T1:} Some of the code in the driver is shared across all operating systems that support the implemented GPUs.
 This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors.
@@
 Regarding our analysis from the \textbf{participant-as-Observer} phase, Table~\ref{tab:stu} summarizes the student groups' mitigation approaches.
-It contains the information about the chosen subsystem and the similarity threshold for groups that opted for the first approach.
-Each entry also has the patch status, refactoring method, and code diff corresponding to the final patch versions before it was dropped or accepted into the code base.
+It contains the information about the chosen subsystem (SS) and the similarity threshold for groups (SML) that opted for the first approach.
+Each entry also has the patch status, and code diff corresponding to the final patch versions before it was dropped or accepted into the code base.
 The code diff for a patchset corresponds to the sum of added and removed lines of each of its patches.
@@
 \hline
-\textbf{G} & \textbf{SS} & \textbf{SML} & \textbf{Patch Status} & \textbf{Method} & \textbf{Diff} \\
+\textbf{GID} & \textbf{SS} & \textbf{SML} & \textbf{Patch Status} & \textbf{Takeaway} & \textbf{Diff} \\
 \hline
-
-1 & IIO & 100\% & Dropped (v2) & Undefined & +7/-14 \\ \hline
-2 & IIO & 100\% & Dropped (v1) & Parameterize & +23/-27 \\ \hline
-3 & IIO & 100\% & Accepted (v1) & Parameterize & +12/-30 \\ \hline
-4 & IIO & 100\% & Accepted (v3) & Undefined & +2/-7 \\ \hline
-5 & IIO & 90\% & Accepted (v1) & Parameterize & +17/-70 \\ \hline
-6 & IIO & 90\% & Dropped (v1) & Parameterize & +95/-92 \\ \hline
-7 & IIO & 90\% & Accepted (v3) & Extract & +20/-36 \\ \hline
-8 & IIO & 90\% & Accepted (v4) & Extract & +20/-16 \\ \hline
-9 & IIO & 90\% & Accepted* (v6) & Parameterize & +149/-243 \\ \hline
-10 & IIO & 90\% & Dropped (v1) & Undefined & +32/-58 \\ \hline
-11 & IIO & 100\% & Dropped (v1) & Inline & +2/-12 \\ 
-   & AMD & 90\% & Accepted (v2) & Undefined & +90/-489 \\ \hline
+0 & AMD & ???\% & Accepted (v3) & T1 & +114/-520 \\ \hline
+1 & IIO & 100\% & Dropped (v2) & ? & +7/-14 \\ \hline
+2 & IIO & 100\% & Dropped (v1) & ? & +23/-27 \\ \hline
+3 & IIO & 100\% & Accepted (v1) & ? & +12/-30 \\ \hline
+4 & IIO & 100\% & Accepted (v3) & ? & +2/-7 \\ \hline
+5 & IIO & 90\% & Accepted (v1) & ? & +17/-70 \\ \hline
+6 & IIO & 90\% & Dropped (v1) & ? & +95/-92 \\ \hline
+7 & IIO & 90\% & Accepted (v3) & ? & +20/-36 \\ \hline
+8 & IIO & 90\% & Accepted (v4) & ? & +20/-16 \\ \hline
+9 & IIO & 90\% & Accepted (v6) & ? & +149/-243 \\ \hline
+10 & IIO & 90\% & Dropped (v1) & ? & +32/-58 \\ \hline
+11 & IIO & 100\% & Dropped (v1) & ? & +2/-12 \\ 
+   & AMD & 90\% & Accepted (v2) & ? & +90/-489 \\ \hline
 
@@
 
+Regarding to refactor methods ....
+
+%TODO: nao descrever e discutir mais as estrategias de refatoracao por patch ...
 Seven groups applied the parameterized refactor method, two groups used the extractor method, and one group applied the inline method to their approaches to mitigate their duplications.

### 2025-10-23 paper/techdebt2026/techdebt.tex
techdebt2026: techdebt: add takeaways w/ example quotes
From the patchset analysis, we detected 4 takeaways beyond the one
already in place. For now, dump them along with quotes in a temporary
section "Takeaways (temporary)", just to gauge the size of adding those
boxes. Beyond adding the takeaways and quotes, fill the column
"Takeaway" from the table with the contributions.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 \usepackage{todonotes}
+
+%% Takeaway Boxes %%
+\usepackage[most]{tcolorbox}
+\tcbset{colback=gray!10, colframe=gray!50!black, boxrule=0.5pt,
+        left=1mm, right=1mm, top=1mm, bottom=1mm, sharp corners}
+
+\newenvironment{takeaway}[1]{%
+  \begin{tcolorbox}
+  \textbf{#1:} \itshape}{\end{tcolorbox}}
+%% Takeaway Boxes %%
 
@@
 0 & AMD & ???\% & Accepted (v3) & T1 & +114/-520 \\ \hline
-1 & IIO & 100\% & Dropped (v2) & ? & +7/-14 \\ \hline
-2 & IIO & 100\% & Dropped (v1) & ? & +23/-27 \\ \hline
-3 & IIO & 100\% & Accepted (v1) & ? & +12/-30 \\ \hline
-4 & IIO & 100\% & Accepted (v3) & ? & +2/-7 \\ \hline
-5 & IIO & 90\% & Accepted (v1) & ? & +17/-70 \\ \hline
-6 & IIO & 90\% & Dropped (v1) & ? & +95/-92 \\ \hline
-7 & IIO & 90\% & Accepted (v3) & ? & +20/-36 \\ \hline
-8 & IIO & 90\% & Accepted (v4) & ? & +20/-16 \\ \hline
-9 & IIO & 90\% & Accepted (v6) & ? & +149/-243 \\ \hline
-10 & IIO & 90\% & Dropped (v1) & ? & +32/-58 \\ \hline
-11 & IIO & 100\% & Dropped (v1) & ? & +2/-12 \\ 
-   & AMD & 90\% & Accepted (v2) & ? & +90/-489 \\ \hline
+1 & IIO & 100\% & Dropped (v2) & T2, T3 & +7/-14 \\ \hline
+2 & IIO & 100\% & Dropped (v1) & T2 & +23/-27 \\ \hline
+3 & IIO & 100\% & Accepted (v1) & None & +12/-30 \\ \hline
+4 & IIO & 100\% & Accepted (v3) & T2 & +2/-7 \\ \hline
+5 & IIO & 90\% & Accepted (v1) & None & +17/-70 \\ \hline
+6 & IIO & 90\% & Dropped (v1) & T4 & +95/-92 \\ \hline
+7 & IIO & 90\% & Accepted (v3) & None & +20/-36 \\ \hline
+8 & IIO & 90\% & Accepted (v4) & T5 & +20/-16 \\ \hline
+9 & IIO & 90\% & Accepted (v6) & T5 & +149/-243 \\ \hline
+10 & IIO & 90\% & Dropped (v1) & T2, T5 & +32/-58 \\ \hline
+11 & IIO & 100\% & Dropped (v1) & None & +2/-12 \\
+   & AMD & 90\% & Accepted (v2) & None & +90/-489 \\ \hline
 
@@
 
+%TODO REMOVE ME!!! Essa seção é só para fazer o dump e ver como fica o tamanho.
+\section{Takeaways (temporary)}
+
+%TODO achar uma forma mais visual de colocar as aspas
+\noindent
+\begin{footnotesize}
+  \textit{
+    Quote: ``I think the old code is more readable than hiding the values in a macro
+    even if it is duplicating a few lines of code.''
+  }
+\end{footnotesize}
+
+\begin{takeaway}{Takeaway T2 (Readability)}
+  Maintainers notably prioritized readability over removing duplications in
+  many different contexts. In their view, going against the \textbf{DRY
+  principle} was worth it if it resulted in code that is easier to read and
+  understand.
+\end{takeaway}
+
+\noindent
+\begin{footnotesize}
+\textit{
+    Quote: ``[...] such patches might not worth it since the proposed
+    improvement is very small (and questionable) while the upstreaming process
+    still requires some effort.''
+  }
+\end{footnotesize}
+
+\begin{takeaway}{Takeaway T3 (Integration Overhead)}
+  A maintainer suggested that the effort put into the contribution after it was
+  accepted (i.e., the process of propagating the change to upper maintainers)
+  was considered. In other words, maintainers may \textbf{evaluate the
+  trade-off between the impact of contributions (not only deduplications) and
+  the struggle to integrate them into the complex Linux project}.
+\end{takeaway}
+
+\noindent
+\begin{footnotesize}
+\textit{
+    Quote: ``Wrapping this up doesn't provide any real advantage, requiring as
+    it does the reviewer to look at this function AND where the value is set
+    rather than seeing them in one place.''
+  }
+\end{footnotesize}
+
+%TODO Na citação abaixo, trazida pelo Arthur, pensar se não vale colocar as
+%referências base do paper do Lanza: ``Developers need to mentally reconstruct
+%the role and behavior of a class from scattered fragments of code [8],
+%presented as one page of text at a time. This process is cognitively demanding
+%[9], slow, and error-prone, especially in large systems with many
+%interdependent components [10].''
+%
+% Referências do paper do Lanza para esse trecho:
+%[8] M.-A. Storey, “Theories, methods and tools in program comprehension:
+%Past, present and future,” in Proceedings of IWPC 2005. IEEE, 2005,
+%pp. 181–191.
+%[9] J. Sillito, G. C. Murphy, and K. De Volder, “Questions programmers
+%ask during software evolution tasks,” in Proceedings FSE 2006. ACM,
+%2006, pp. 23––34.
+%[10] G. C. Murphy, D. Notkin, and K. Sullivan, “Software reflexion models:
+%Bridging the gap between source and high-level models,” in Proceedings
+%of FSE 1995. ACM, 1995, pp. 18–28.
+\begin{takeaway}{Takeaway T4 (Cognitive Load)}
+  Duplications that made it possible to read the code without needing to jump
+  to the function definition (to see its behavior) or the function call (to see
+  its concrete parameters) were justifiable to reduce cognitive load when
+  parsing through the code~\cite{skylines-cognitive-load}.
+\end{takeaway}
+
+\noindent
+\begin{footnotesize}
+\textit{
+    Quote: ``Note that is not an appropriate change for the large reads though
+    as \texttt{spi\_write\_then\_read()} bounces all buffers and so would add a
+    copy to those high(ish) performance paths.''
+  }
+\end{footnotesize}
+
+\begin{takeaway}{Takeaway T5 (Performance)}
+  Given the low-level development context of Linux and the IIO subsystem, which
+  is usually related to devices with limited resources, we detected hints that
+  duplication may be forgivable if it improves performance.
+\end{takeaway}
+
 \section{Concluding Remarks}

### 2025-10-24 paper/techdebt2026/techdebt.tex
Shorten refactoring method description
Replace details about refactoring methods used with a brief comment,
as this is not relevant to this paper.

@@
 
-Regarding to refactor methods ....
-
-%TODO: nao descrever e discutir mais as estrategias de refatoracao por patch ...
-Seven groups applied the parameterized refactor method, two groups used the extractor method, and one group applied the inline method to their approaches to mitigate their duplications.
-Four patches do not contain a refactoring technique defined in the literature (undefined).
-Six groups had their patches accepted and fully merged into Linux; four groups had their patches dropped; one group has its patchset partially applied and is still working on patch submission, currently in the sixth version of the patch when this document was written, but we can also consider this case as accepted.
+We observed that most groups applied the parameterized refactor method while a few used the extractor method.
 

### 2025-10-24 paper/techdebt2026/techdebt.tex
techdebt2026: update title
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-10-24 paper/techdebt2026/techdebt.tex
@@
 
-\title{Repeat Yourself? When Linux kernel maintainers balance Code Duplication and Readability}
+\title{When Do Repeat Yourself?\\
+Voices from the Trenches of Linux Kernel Maintainers on Code Duplication}
 
@@
 In the AMD Display driver codebase, we executed our tool.
-We manually analyzed the biggest duplications regarding the number of lines and found two duplicated function pairs that we judged interesting to try to mitigate.
+We manually analyzed the biggest duplications regarding the number of lines and found one duplicated function pairs that we judged interesting to try to reduce the duplication.
 We have chosen to approach the biggest duplications as proof of concept so that we can make significant impacts.
 
-Given the two duplicated function pairs, we observed that the files in the functions 
+Given the duplicated function pair, we observed that the files in the functions
 contain multiple other duplications. Thus, we proposed a simple systematic approach 
@@
 function pairs. 
-After we approached the duplications with the systematic strategy, for the duplicated 
-function pairs that had successfully mitigated the duplication, we sent patches to the 
+After we approached the duplications with the systematic strategy, we sent patch to the 
 AMD Display driver email list to validate the mitigations and see the maintainer's opinion 
@@
 
-In the \textbf{complete participant observation} study, we chose two function pair found on AMD Display Driver to approach and apply the systematic approach ...
+In the \textbf{complete participant observation} study, we chose one function pair found on AMD Display Driver to approach and apply the systematic approach.
 
 %TODO (Luan): nao terah mitigation 1 e 2, apenas a contribuicao enviada mesmo.
-
-Firt, we found that the duplicated function from mitigation 1
-exists in 10 code files, enclosing multiple GPU architectures. The duplicated 
-functions are not exactly equal, as some minimal specific logic is applied to each GPU architecture. To address this issue, we used the extract method to 
-break the function into smaller parts and the parameterized method to address 
-the specific logic for the GPUs.
 
@@
 Thus, we opted not to continue investigating this mitigation.
-
-Second, we found that the duplicated function from mitigation 2 exists in five code files, 
+We found that the duplicated function from the code redution exists in five code files, 
 and two other functions are duplicated in those five code files. All functions across the files were exactly equal, 
@@
 
-The code files in the second mitigation do not depend on configuration files. Thus, 
-we did not face the same issues during the first mitigation. Refactoring the code 
-to mitigate duplications reached a point we judged was good enough to submit as a 
+Refactoring the code to reduce the duplications reached a point we judged was good enough to submit as a 
 patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
@@
 %TODO: deve ficar claro que isto aqui, mesmo que não dito explicitamente pelo mantenedores, é o resultado da Complete Participant Observation
-\noindent\textbf{Takeaway T1:} Some of the code in the driver is shared across all operating systems that support the implemented GPUs.
-This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors.
-Additionally, \textbf{we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering}.
-One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others.
-This approach helps save a significant amount of time and effort.
-\todo[inline]{...Foi tirado apenas das conversas com o Siqueira; tem o paper que o Gregório citou que menciona que duplicações podem ser bem vistas em drivers; isto um resultado/conclusão da Complete Participant Observation; e vamos citar que confirmamos o que está no paper Cloning Considered Harmful (Luan, acha que consegue fechar essa parte assim?)}
+\noindent\textbf{Takeaway T1:}
+\textbf{We came to understand that AMD Display driver developers do not view duplicated code negatively, 
+which seems to contradict standard software engineering practices.} 
+In our attempt to understand the reason for the delay in the review process, 
+we contacted a maintainer. It was mentioned to us that one reason for the delay 
+is that duplicated code enhances the independence of GPU driver code. This allows 
+developers to make changes to a specific GPU without needing to test compatibility 
+with others, which helps save a significant amount of time and effort. This example of 
+hardware variations, also mentioned by Kapser and Godfrey~\cite{cloneharm}, 
+corroborates our understanding that driver developers do not view duplicated code negatively by default.
 

### 2025-10-24 paper/techdebt2026/techdebt.tex
techdebt2026: update Abstract
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 The vast scale and continuous evolution of the Linux kernel make its maintenance a complex undertaking, where code duplication remains a persistent challenge that can hinder development and introduce bugs.
-This paper presents a multimethod ethnographic study investigating the socio-technical dynamics of contributing to the Linux kernel by mitigating identified redundancies.
-To support this investigation, we developed a command-line utility tool for detecting function-level duplications, offering a concrete entry point for new contributors.
-Our ethnographic approach, including participant and non-participant observations, demonstrated that addressing duplications is viable for lowering the contribution barrier.
-This point is evidenced by the main author and 7 of 11 (64\%) student groups having accepted contributions in the Linux project, collectively removing 991 lines of duplicated code.
+This paper presents a multimethod ethnographic study investigating the socio-technical dynamics of contributing to the Linux kernel by reducing identified code duplication.
+To support this study, we developed a command-line utility tool for detecting function-level duplications, offering a concrete entry point for new contributors.
+Our ethnographic approach, including complete participant observation and participant-as-observer analyses, demonstrated that addressing duplications is viable for lowering the contribution barrier.
+Our ethnographic approach, including complete participant observation and participant-as-observer analyses, demonstrated that addressing duplications is viable for lowering the contribution barrier.
+This point is evidenced by 8 of 13 (62\%) accepted contributions (patches) from 24 newcomers (16 undergraduate and 8 graduate students) in the Linux kernel project, collectively removing 1.397 lines of duplicated code.
 Nevertheless, the study reveals a more intriguing and complex reality beyond mindlessly removing clones.

### 2025-10-24 paper/techdebt2026/techdebt.tex
remove mitigation 1 and rephase takeway 1 a little
techdebt2026: techdebt: revise introduction
Make adjustments to the introduction. Among which:
- Add the DRY quote
- Make the introduction more concise
- Give more weight to the contrasting interactions with the maintainers
  in terms of deduplications.

Rename the highlight box macro also, as a collateral effect.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-\newenvironment{takeaway}[1]{%
+\newenvironment{highlight-box}[1]{%
   \begin{tcolorbox}
@@
 
-Coordinating hundreds or thousands of developers working on a software product is complex.
-Creating similar or even completely duplicated code is a frequent and often unavoidable outcome.
-Excessive duplication can negatively impact the maintainability of a project.
-When duplications occur, any change to a repeated code segment typically requires manual replication across all other copies.
-Developers must, therefore, remain constantly vigilant to synchronize different parts of the codebase, which is a highly error-prone and seemingly simple responsibility.
-With many duplications, modifications become slow, complex, and fragile.
-Removing or refactoring duplicated code demands great care to preserve the original behavior.
-Even minor mistakes can introduce subtle state changes that may lead to serious problems in the future.
+Hundreds or thousands of developers working together on a software product is a complex and demanding task.
+Similar or even exact copies are a frequent and often unavoidable outcome.
+Excessive duplication can negatively impact the maintainability of a project, as, when it occurs, any change to a repeated code segment (to fix a bug, for instance) requires replication across all other copies.
+Developers must, therefore, remain constantly vigilant to synchronize different parts of the codebase, which is a highly error-prone yet seemingly simple responsibility.
+Regarded as a major source of technical debt, removing or refactoring duplicated code demands great care to preserve the original behavior, with minor mistakes possibly leading to serious problems in the future.
+
+A widespread dogma among developers is the \textit{Don't Repeat Yourself} principle~\cite{pragmatic-programmer}, stated as follows.
+
+\begin{highlight-box}{DRY principle}
+  ``We feel that the only way to develop software reliably, and to make our
+  developments easier to understand and maintain, is to follow what we call the
+  DRY principle: Every piece of knowledge must have a single, unambiguous,
+  authoritative representation within a system''.
+\end{highlight-box}
 
 The Linux kernel is a foundational Free/Libre and Open Source Software (FLOSS) project, a vital component of the world's digital infrastructure.
-Its maintenance is a massive undertaking, encompassing over 28 million lines of code and involving contributions from over twenty thousand developers.
+Considering only lines of code (no comments or blank lines) of strictly programming languages, in version 6.17, the main repository has more than 27 million lines of code and involved contributions from over twenty thousand developers.
 Within this context, code duplication remains a significant challenge,  reducing the code readability and increasing the risk of introducing bugs~\cite{harmone,harmtwo}.
@@
 
-%TODO: temos que citar o conceito de DRY aqui ... (Book: Pragmatic Programmer, The: From Journeyman to Master)
-% We feel that the only way to develop software reliably, and to make our developments easier to understand and maintain, is to follow what we call the DRY principle: Every piece of knowledge must have a single, unambiguous, authoritative representation within a system.
 The detection of code duplication, or \textit{code clones}, has been a research topic for decades~\cite{firstman}.
@@
 
-To address the code duplication issue, we developed a command-line tool designed to support Linux kernel maintenance by detecting and analyzing function-level duplications.
-This paper presents a multimethod ethnographic study to validate the tool capabilities in identifying duplications and supporting contributions to the Linux kernel by mitigating redundant code.
-Therefore, the study investigates how the Linux kernel community perceives the code quality related to code clones.
-Our findings indicate that maintainers consider not only duplication but also factors such as readability and appropriate levels of abstraction regarding code maintainability.
+To address code duplication in the Linux kernel, we developed a command-line tool capable of detecting and analyzing function-level duplications.
+We conducted a multimethod ethnographic study to test the tool capabilities by identifying duplications and submitting contributions, directly or indirectly, to the Linux kernel, removing the redundant code.
+Through interactions with maintainers, this study gained insights into how the Linux community perceives the code quality related to code clones.
+Contrary to the DRY principle, our findings indicate that, in some contexts, maintainers can tolerate duplications for better readability, performance increases, or even to avoid integration overhead.\looseness=-1
 
-Ethnography is a well-established qualitative research method used to understand people, their cultures, and their work practices~\cite{bookethno}.
-It enables researchers to gain insights into values, beliefs, and practices from the perspective of community members~\cite{ethnosoft}.
-In this study, we first adopted a complete participant observation approach, where the main author actively engaged in mitigating code duplications while documenting the process and outcomes.
-Later, we conducted a partipant-as-observer study involving students tasked with resolving duplications identified by the tool.
-These combined studies showed that our tool can collaborate to reduce the entry barrier for new contributors.
-The student groups (and we) used the tool to guide their initial contributions to the Linux kernel.
+Ethnography is a well-established qualitative method for understanding people, their cultures, and work practices~\cite{bookethno}.
+It provides insights into community members’ values, beliefs, and practices~\cite{ethnosoft}.
+In this study, we first employed a complete participant observation approach, actively contributing by performing deduplications while documenting the process.
+We then conducted a participant-as-observer study with students resolving duplications identified by the tool.
+Beyond demonstrating that our tool lowered the entry barrier for new contributors, these studies yielded intriguing observations on kernel deduplications.
+This paper describes the aforementioned studies, along with our visions and reflections from diving into the development process.
 
@@
 
-\begin{takeaway}{Takeaway T2 (Readability)}
+\begin{highlight-box}{Takeaway T2 (Readability)}
   Maintainers notably prioritized readability over removing duplications in
@@
   understand.
-\end{takeaway}
+\end{highlight-box}
 
@@
 
-\begin{takeaway}{Takeaway T3 (Integration Overhead)}
+\begin{highlight-box}{Takeaway T3 (Integration Overhead)}
   A maintainer suggested that the effort put into the contribution after it was
@@
   the struggle to integrate them into the complex Linux project}.
-\end{takeaway}
+\end{highlight-box}
 
@@
 %of FSE 1995. ACM, 1995, pp. 18–28.
-\begin{takeaway}{Takeaway T4 (Cognitive Load)}
+\begin{highlight-box}{Takeaway T4 (Cognitive Load)}
   Duplications that made it possible to read the code without needing to jump
@@
   parsing through the code~\cite{skylines-cognitive-load}.
-\end{takeaway}
+\end{highlight-box}
 
@@
 
-\begin{takeaway}{Takeaway T5 (Performance)}
+\begin{highlight-box}{Takeaway T5 (Performance)}
   Given the low-level development context of Linux and the IIO subsystem, which
@@
   duplication may be forgivable if it improves performance.
-\end{takeaway}
+\end{highlight-box}
 

### 2025-10-24 paper/techdebt2026/techdebt.tex
techdebt2026: techdebt: correct the title
Add the "You" in "When Do You Repeat Yourself" that was lacking

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>
techdebt2026: techdebt: remove duplicated sentende in the abstract
The abstract contained a duplicated sentence, so remove it.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>
techdebt2026: techdebt: revise background section
Make small adjustments in the background section.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-\title{When Do Repeat Yourself?\\
+\title{When Do You Repeat Yourself?\\
 Voices from the Trenches of Linux Kernel Maintainers on Code Duplication}
@@
 To support this study, we developed a command-line utility tool for detecting function-level duplications, offering a concrete entry point for new contributors.
-Our ethnographic approach, including complete participant observation and participant-as-observer analyses, demonstrated that addressing duplications is viable for lowering the contribution barrier.
 Our ethnographic approach, including complete participant observation and participant-as-observer analyses, demonstrated that addressing duplications is viable for lowering the contribution barrier.
@@
 
-The Linux kernel is a monolithic system organized into subsystems, such as the process scheduler, memory management, and device drivers.
+The Linux kernel is a complex project organized into subsystems, such as the process scheduler, memory management, and device drivers.
 Each subsystem typically has a dedicated maintainer or a team of maintainers overseeing its development and managing contributions.
-The development process is coordinated through the Git version control system.
-Contributions are formatted as patches, text documents that outline the differences between two source code versions.
+The development process is coordinated through Git, and contributions are formatted as patches, text documents that outline the differences between two source code versions.
 These patches are then submitted to the relevant Linux mailing lists for public review and discussion.
@@
 This study focuses on two of these subsystems, selected to represent different contexts for contribution.
-We explore the AMD Display driver, chosen for its importance within the hardware ecosystem, which is responsible for enabling AMD GPU functionality in Linux.
+We explore the \textit{AMD Display drivers}, chosen for their importance within the hardware ecosystem, which is responsible for natively enabling AMD GPU functionality in Linux.
 This subsystem is substantial, with over 391,000 lines of code across more than 1,089 files.
-We also investigate the Industrial I/O (IIO) subsystem, often considered an accessible entry point for new contributors.
-The IIO subsystem provides support for devices that perform analog-to-digital or digital-to-analog conversions\footnote{\href{https://kernel.org/doc/html/v4.14/driver-api/iio/intro.html}{kernel.org/doc/html/v4.14/driver-api/iio/intro.html}} and contains over 281,772 lines of code in more than 755 files.
+We also investigate the \textit{Industrial I/O (IIO)} subsystem, often considered an accessible entry point for new contributors.
+The IIO subsystem provides support for devices that perform analog-to-digital or digital-to-analog conversions\footnote{\href{https://kernel.org/doc/html/v6.17/driver-api/iio/intro.html}{https://kernel.org/doc/html/v6.17/driver-api/iio/intro.html}} and contains over 281,772 lines of code in more than 755 files.
 

### 2025-10-24 paper/techdebt2026/techdebt.tex
techdebt2026: techdebt: revise Research Strategy section
Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-We employ a multi-method research approach. We developed and evaluated a proposed tool using an approach from the literature~[X].
-%
-Given the validation of the tool capabilities, we proceed to understand the viability 
-of reducing the code duplication found by the tool. We conducted an ethnographic study 
-using a participant-observation methodology to understand its viability~\cite{}.
-%
-In the first phase, we explored the process of reducing code duplication found in the AMD Display driver and sent it as a patch to interact with the Linux kernel community and get their feedback (\textit{complete participant observation}~\cite{}).
-%
-In the second phase, during a university course, we proposed to students to reduce the code duplications found by our tool in the IIO subsystem and the AMD Display driver and pass the process of sending the contribution as a patch (\textit{participant-as-observer}~\cite{}) as a course activity while we guided them through the process.
+We employed a multi-method research approach.
+We developed and evaluated a proposed tool using a method from the literature~[X].
+Given this validation, we proceeded to understand the viability of reducing duplications found by the tool with an ethnographic study using a participant observation methodology~\cite{}.
+In the first phase, we interacted with the Linux kernel community and got their feedback (\textit{complete participant observation}~\cite{}) through sending deduplication patches to the AMD Display driver.
+In the second phase, during a university course, we proposed a similar activity of sending patches to AMD Display or IIO to students.
+At the same time, we guided and mentored them throughout the process (\textit{participant-as-observer}~\cite{}).
 
@@
 Our tool is designed for the specific challenges of the kernel, with a focus on scalability to handle its size and prioritizing duplications that impact maintainability.
-The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying function-level duplications to address.
 The tool uses a text similarity method that applies TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine distance similarity to detect clones.
@@
 
-To investigate whether we can use our tool to find ways to contribute to the Linux kernel, we conducted a complete participant observation experiment, where we acted as first-time contributors to the AMD Display driver, collecting artifacts of our experience in the process.
+To investigate whether we could use our tool to find ways to contribute to the Linux kernel, we conducted a complete participant observation experiment, where we acted as first-time contributors to the AMD Display driver, collecting artifacts of our experience in the process.
 
-In the AMD Display driver codebase, we executed our tool.
-We manually analyzed the biggest duplications regarding the number of lines and found one duplicated function pairs that we judged interesting to try to reduce the duplication.
-We have chosen to approach the biggest duplications as proof of concept so that we can make significant impacts.
+We executed our tool in the AMD Display driver codebase.
+We manually analyzed the biggest duplications regarding the number of lines, finding one pair that we judged promising to try to deduplicate.
+We approached big duplications as proof of concept to gauge the tool's potential for enabling significant impacts.
 
-Given the duplicated function pair, we observed that the files in the functions
-contain multiple other duplications. Thus, we proposed a simple systematic approach 
-to mitigate all the functions duplicated in the context, not just the duplicated 
-function pairs. 
-After we approached the duplications with the systematic strategy, we sent patch to the 
-AMD Display driver email list to validate the mitigations and see the maintainer's opinion 
-while documenting the process and our impressions.
-The systematic approach is defined below.
+Given the duplicated function pair, we observed that the files in the functions contained multiple other duplications.
+Thus, we proposed a simple systematic approach to mitigate all the functions duplicated in the context, not just the duplicated function pairs.
+After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions.
 
-In C programming, communication between source files is achieved by creating header files that specify 
-libraries~\cite{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
-by consolidating duplicated code into a single library, which would replace instances of 
-duplication across the codebase.
-
-For each duplicated function pair, for each function in the pair, we used our tool to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair.
+In C programming, communication between source files is achieved by creating header files that specify libraries~\cite{Cbook}.
+Since the Linux kernel is primarily written in C, our systematic strategy consisted of eliminating code duplications through consolidating the duplicated code into a single library, which would replace instances of duplication across the codebase.
+For each duplicated function pair, for each function in the pair, we used our tool to locate all duplications of the function in the codebase, as it could have been duplicated more times than the two occurrences detected initially.
 This strategy resulted in a collection of duplicated functions and corresponding code files.
 
-To identify shared code more effectively, we extended this approach to search for other common 
-functions across all collection files. We then applied specific refactoring methods to each shared 
-function. If the functions were identical across files, we removed the duplicates and created a single 
-function in the library. If modifications existed, we applied case-specific refactoring.
+To identify shared code more effectively, we extended this approach to search for other common functions across all collection files.
+We then applied specific refactoring methods to each shared function.
+If the functions were identical across files, we removed the duplicates and created a single function in the library.
+If modifications existed, we applied case-specific refactoring.
 
-\subsection{Participant-as-Observer Study}
+\subsection{Participant-as-Observer Observation Study}
 
-A core dynamic of the university course where we conducted the participant-as-observer study is to have students make a practical contribution to a free software project.
-In the first semester of 2025, we identified an opportunity to leverage our tool within this existing pedagogical structure.
-As a contribution option, we proposed that students mitigate code duplications found in the Linux kernel as one possible way to contribute to a free software project.
+A core dynamic of the university course where we conducted the participant-as-observer study was to have students make a practical contribution to a free software project.
+In the first semester of 2025, we identified an opportunity to leverage our tool within this pedagogical structure.
+As a contribution option, we proposed that students remove code duplications found in the Linux kernel as one possible way of contributing to a free software project.
 
-The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of 
-two or three members. To assist students in this task, we prepare multiple 
-alternatives for contributing to free software projects, including simpler contributions such as 
-updating documentation or fixing code style issues. In this context, two new options related to migitating
-duplications in the Linux kernel were proposed alongside the others.
+The 2025 course offering had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of two or three members.
+To assist students in this task, we prepared multiple alternatives for contributing to free software projects, including simpler contributions such as updating documentation or fixing code style issues.
+In this context, two options related to removing duplications in the Linux kernel were offered.
 
-For the first our tool-related alternative, we ran the tool on the IIO subsystem, creating a list of duplicated function pair.
-Then, the following process happened to the list to prepare more actionable duplications to the students:
+In the first option, we ran the tool on the IIO subsystem, creating a list of duplicated function pairs.
+Then, we curated the list even more to get more actionable duplications for the students using this approach:
 
 \begin{enumerate}
-    \item We filtered the list to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different devices may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
-    \item We removed duplications that were too short and would not represent a significant change for a patch application;
+    \item We filtered the list to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different drivers may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
+    \item We removed duplications that were too short and would not represent a significant contribution;
     \item We manually ranked code duplications and marked some duplications with insights based on personal opinion and perceptions over the remaining entries.
@@
 
-The students' task was to select a recommended entry from the list, refactor the code to eliminate these specific duplications, and then submit their patches to the IIO maintainers for review.
+The students' task was to select a recommended entry from the list, refactor the code to eliminate these duplications, and then submit their patches to the IIO mailing list for the maintainers' review.
 
-For the second alternative, students were offered an experience similar to our initial participant 
-observation. In this model, the students were responsible for the entire workflow: executing our tool on the AMD display driver, independently analyzing the results to identify a duplication 
-to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.
+For the second alternative, students were offered an experience similar to our initial complete participant observation.
+In this model, students were responsible for the entire workflow: executing our tool on the AMD Display drivers, independently analyzing the results to identify a potential deduplication, creating a patch with it, and sending the patch to the driver's maintainers.
 
-Of the students who chose to work on our tool-related tasks, 23 (16 undergraduate and 7 graduate) 
-formed 11 groups to pursue the first alternative.
-One graduate student also opted for the second alternative.
+Of the students who chose to work on our tool-related tasks, 23 (16 undergraduate and 7 graduate) formed 11 groups to pursue the first alternative.
+One graduate student opted for the second alternative.
 Notably, none of these students had prior experience contributing to the Linux kernel, making this their first attempt to submit a patch as newcomers to the project.
-%TODO: ver na versão final incluir o patch do Lucas (Octavio e Fernando)
 
-The groups were requested to document their experiences and approaches to refactoring the duplications, as well as their experiences submitting a patch to the Linux kernel, in blogs.
-We analyzed these blogs to understand the refactoring patterns used to mitigate duplications, the experience of sending patches, and the opinions of subsystem maintainers on the mitigations.
+Using blog posts, the groups documented their experiences and approaches to refactoring the duplications, as well as their experiences submitting a patch to the Linux kernel.
+We also analyzed these posts to understand the refactoring patterns used in deduplications, the experience of sending patches, and the opinions of subsystem maintainers on them.
 
@@
 In the \textbf{complete participant observation} study, we chose one function pair found on AMD Display Driver to approach and apply the systematic approach.
-
-%TODO (Luan): nao terah mitigation 1 e 2, apenas a contribuicao enviada mesmo.
 

### 2025-10-24 paper/techdebt2026/techdebt.tex
techdebt2026: techdebt: revise results section
Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 \section{Results}
-%TODO: ver como inicar melhor aqui ...
 
@@
 
-We reached a point where we had a scratch on how to mitigate the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver.
+We reached a point where we had an idea of how to remove the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver.
 We found that the configuration files on the AMD Display driver are designed to be imported at compilation time using \textit{\#define} macros.
-The configuration files' design choice makes refactoring duplications on generic approaches trickier, as refactoring code that depends on these files requires significant refactoring of the configuration files' design and deep knowledge of the codebase, which we, as first-time contributors, do not have.
-Thus, we opted not to continue investigating this mitigation.
-We found that the duplicated function from the code redution exists in five code files, 
-and two other functions are duplicated in those five code files. All functions across the files were exactly equal, 
-so there was no need to apply the refactoring methods presented in the literature. 
+The configuration files' design choice makes refactoring duplications on generic approaches trickier, as refactoring code that depends on these files requires significant modification of the configuration files' design and deep knowledge of the codebase, which we, as first-time contributors, did not have.
+Thus, we opted not to continue investigating this deduplication.
+We found that the duplicated function from the code reduction existed in five code files, and two other functions were duplicated in those five.
+All functions across the files were exactly equal, so there was no need to apply the refactoring methods presented in the literature.
 The refactoring was resumed to create a generic library and fix the compilation targets.
 
-Refactoring the code to reduce the duplications reached a point we judged was good enough to submit as a 
-patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
-the refactoring to the driver maintainers as a patch, documenting our findings and experiences 
-on the process.
+In this attempt, we reached a point where the deduplication was good enough to submit it as a contribution to the AMD Display driver code quality.
+Thus, we moved to sending the refactoring to the driver maintainers as a patch.
 
-We submitted the initial patch: the maintainers only asked for minor changes to align with code style, correct license use, and best practices that were initially unknown.
-We submitted a second version: the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
+In the feedback on the first submitted version, the maintainers only asked for minor changes to align with the coding style, correct license use, and align with initially unknown conventions.
+We submitted a second version, in which the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
 We sent a third version, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
@@
 
-%TODO: deve ficar claro que isto aqui, mesmo que não dito explicitamente pelo mantenedores, é o resultado da Complete Participant Observation
-\noindent\textbf{Takeaway T1:}
-\textbf{We came to understand that AMD Display driver developers do not view duplicated code negatively, 
-which seems to contradict standard software engineering practices.} 
-In our attempt to understand the reason for the delay in the review process, 
-we contacted a maintainer. It was mentioned to us that one reason for the delay 
-is that duplicated code enhances the independence of GPU driver code. This allows 
-developers to make changes to a specific GPU without needing to test compatibility 
-with others, which helps save a significant amount of time and effort. This example of 
-hardware variations, also mentioned by Kapser and Godfrey~\cite{cloneharm}, 
-corroborates our understanding that driver developers do not view duplicated code negatively by default.
+\begin{highlight-box}{Takeaway T1:}
+  \textbf{As the primary result of our complete participant observation we
+  understood that AMD Display driver developers do not necessarily view
+  duplicated code negatively, which contradicts standard software engineering
+  practices}. We suffered a delay in the review process, and to understand why,
+  we directly contacted a maintainer. It was mentioned to us that one reason for
+  the delay was that duplicated code enhances the independence of GPU driver
+  code. This allows developers to make changes to a specific GPU without needing
+  to test compatibility with others, which helps save a significant amount of
+  time and effort. This example of hardware variations, also mentioned by Kapser
+  and Godfrey~\cite{cloneharm}, corroborates our understanding that GPU driver
+  developers do not view duplicated code negatively by default.
+\end{highlight-box}
 
-%TODO: ver se transicao estah boa; ou se eh melhor ter subsecoes
-Regarding our analysis from the \textbf{participant-as-Observer} phase, Table~\ref{tab:stu} summarizes the student groups' mitigation approaches.
-It contains the information about the chosen subsystem (SS) and the similarity threshold for groups (SML) that opted for the first approach.
-Each entry also has the patch status, and code diff corresponding to the final patch versions before it was dropped or accepted into the code base.
-The code diff for a patchset corresponds to the sum of added and removed lines of each of its patches.
+Regarding our analysis from the \textbf{participant-as-observer} phase, Table~\ref{tab:stu} summarizes all the contributions sent by newcomer groups (GID)~\footnote{As in the first phase, we were also newcomers, we included our experience in the table as Group 0.
+Nevertheless, the subsequent analysis will focus more on the patches sent by students.}.
+It contains the information about the chosen subsystem (SS) and the similarity threshold (SML).
+Each entry also has the patch status, takeaways that may have surfaced from the interaction, and a code differential corresponding to the final patch versions before it was dropped or accepted into the code base.
+The code differential for a set of patches corresponds to the sum of added and removed lines of each patch.
 
@@
 \centering
-\caption{Refactor methods used and metrics of the students' mitigations.}
+\caption{Summary of newcomer deduplication contributions.}
 \begin{tabular}{ |c |c |c |c |c | c| }
@@
 \hline
-0 & AMD & ???\% & Accepted (v3) & T1 & +114/-520 \\ \hline
+0 & AMD & 100\% & Accepted (v3) & T1 & +114/-520 \\ \hline
 1 & IIO & 100\% & Dropped (v2) & T2, T3 & +7/-14 \\ \hline
@@
 
-Groups 1 and 4 have worked in similar duplications, where two functions are involved: one to determine if a register is volatile and the other to determine if it is writable.
-These are two functions that return opposite Boolean results.
+Groups 1 and 4 worked on similar duplications, where two functions are involved: one to determine if a register was volatile and the other to determine if it was writable.
+These were two functions that returned opposite Boolean results.
 Group 1 initially addressed the problem using the extract method, while Group 2 used macros.
@@
 However, by the end of the feedback cycle, maintainers considered the patch from Group 1 irrelevant, while the patch from Group 4 was accepted.
-The reason is that Group 1 has chosen to apply the refactoring to a device that contains a register that is both read-only and non-volatile.
+The reason is that Group 1 chose to apply the refactoring to a device that contained a register that was both read-only and non-volatile.
 Thus, the technical debt selected by Group 1 was not merely a duplication but an incorrect implementation of the method.
 
-Groups 3, 6, and 10 created a struct to encapsulate the parameters and mitigate duplication when applying the parameterized method.
+Groups 3, 6, and 10 created a struct to encapsulate the parameters and remove the duplication by applying the parameterized method.
 The patches from Groups 6 and 10 were rejected by the same maintainer who approved the patch from Group 3.
-The maintainer's justification for rejecting the patch from Group 6 is that it proposes to fix duplication while adding more lines than removing, and is not convincing enough in other aspects, such as readability.
-The reason for rejecting the patch from Group 10 is that they refactored code files with simple context; therefore, adding a layer of abstraction to remove duplication decreases code readability without clear gains in codebase quality.
-The maintainers' feedback for Group 10 was:
-\begin{quote}
-
-\textit{``A few comments inline, but overall I'm not sure the code reduction is sufficient to cover the resulting loss of readability.
-Sometimes a switch is simply clearer than a partial look-up table.''}
-
-\textit{``So nice idea, but I'm not seeing it as being a good move for this particular code.
-I'd be slightly interested to see the optimized output of the two approaches, but this is far from a high-performance path, so we care a lot more about readability here.''}
-\end{quote}
-
-Interestingly, both groups addressed duplications across initialization functions for different devices, where code readability is vulnerable to refactoring methods.
+The maintainer's justification for rejecting the patch from Group 6 was that it proposed to fix a duplication while adding more lines than removing, and that it was not convincing enough in other aspects, such as readability.
+The reason for rejecting the patch from Group 10 was that they refactored code files with simple context; therefore, adding a layer of abstraction to remove duplication decreased code readability without clear gains in codebase quality.
+Interestingly, both rejected groups addressed duplications across initialization functions for different devices, where code readability was vulnerable to refactoring methods.
 In contrast, Group 3 tackled the duplication of a simple iterative algorithm, which also contained a bug that could be inferred by inspecting both functions.
@@
 
-Groups 7 and 8 used the extract method to approach the mitigations.
-Group 8 sent the mitigations to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed.
-For Group 7 mitigation, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function.
+Groups 7 and 8 used the extract method to approach the deduplications.
+Group 8 sent their patch to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed.
+In the case of Group 7, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function.
 After the feedback cycle, maintainers accepted the patch in version 3.
@@
 Group 9 initially approached the selected duplication with a parameterized method, but after entering the feedback cycle, it was revealed that the problem represented a much more complex technical debt.
-The patch is still under development, currently at version 6, and has been separated into a 4-patch patch set, where half of them have been applied, and the remaining entries are receiving suggestions.
-Fortunately, maintainers are receptive and provide constant feedback; the contribution is expected to be fully merged if the students do not abandon the current work.
+From the first version to the last (version six), the patch was split into a series with four patches, where two of them have been applied, and the remaining entries are receiving suggestions.
+Even though the patch became a series of four, only one of them (one of the accepted) was strictly tackling duplication.
 
-Group 11 is the student who executed the tool and completed the process of identifying duplications.
+Group 11 was a single student who took the second option and executed the tool, completing the entire process of identifying duplications himself.
 The student identified a duplication in the driver that occurred across eighteen code files, creating a patch to remove approximately four hundred lines of duplicated code.
-The student sent the mitigations to the maintainers, and their feedback pointed to fixing the warning in the code without arguing about removing the duplications.
+The student sent the deduplications to the maintainers, and their feedback pointed to fixing warnings in the code without arguing about the merit of the duplication removals.
 After addressing the issue, the patch was applied.
@@
 
-Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach duplications found by our tool and become contributors to the kernel.
-We observed that not all duplications
-are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code, code readability, and the levels of abstraction added to address the duplications.
-
-%\subsection{Statistics of the contributions}
+Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach the duplications found by our tool as a comprehensive and more straightforward way to become contributors to the kernel.
+We observed that not all duplications are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code and many other factors.
 
@@
 
-Finally, the AMD contribution from Group 11, which involves the complete use of our developed tool toolchain, got a completely different outcome.
-With a total of 90 lines inserted and 489 lines removed, there are some points to be considered for the success and high modification extension of the contribution:
+Finally, the AMD contribution from Group 11, which involved the complete use of our developed tool components, got a completely different outcome.
+With a total of 90 lines inserted and 489 lines removed, there are some points to be considered for the success and high modification extension of the contribution: (1) the AMD Display subsystem has some self-reported problems regarding overall technical debt; (2) the AMD Display subsystem has a more centralized code structure and maintainer hierarchy, enabling contributors to tackle code duplications across a broader scope and multiple files, which happened to Group 11's contribution;
 
-\begin{itemize}
-    \item The AMD Display subsystem has some self-reported problems regarding overall technical debt;
-    \item The AMD Display subsystem has a more centralized code structure and maintainer hierarchy, enabling contributors to tackle code duplications across a broader scope and multiple files, which happened to Group 11's contribution;
-    \item The complete control over the process of finding code duplication gave Group 11 the opportunity to identify and reason better over the duplications found by our tool, enabling a more strategic approach to searching for promising patch possibilities.
-\end{itemize}
+We derived the following takeaways from \textbf{participant-as-observer observation}.
+These takeaways are accompanied by example quotes from the maintainers that originated during the review process~\footnote{It is paramount to state that we \textbf{are not affirming} that these quotes represent the complete positioning of the maintainers who wrote them on code duplication.}.
 
-%TODO REMOVE ME!!! Essa seção é só para fazer o dump e ver como fica o tamanho.
-\section{Takeaways (temporary)}
+\begin{highlight-box}{Takeaway T2 (Readability)}
+  Maintainers notably prioritized readability over removing duplications in
+  many different contexts. In their view, going against the \textbf{DRY
+  principle} was worth it if it resulted in code that is easier to read and
+  understand.
+\hfill\\
 
-%TODO achar uma forma mais visual de colocar as aspas
 \noindent
@@
 \end{footnotesize}
+\end{highlight-box}
 
-\begin{highlight-box}{Takeaway T2 (Readability)}
-  Maintainers notably prioritized readability over removing duplications in
-  many different contexts. In their view, going against the \textbf{DRY
-  principle} was worth it if it resulted in code that is easier to read and
-  understand.
-\end{highlight-box}
+\begin{highlight-box}{Takeaway T3 (Integration Overhead)}
+  A maintainer suggested that the effort put into the contribution after it was
+  accepted (i.e., the process of propagating the change to upper maintainers)
+  was considered. In other words, maintainers may \textbf{evaluate the
+  trade-off between the impact of contributions (not only deduplications) and
+  the struggle to integrate them into the complex Linux project}.
+\hfill\\
 
@@
 \end{footnotesize}
-
-\begin{highlight-box}{Takeaway T3 (Integration Overhead)}
-  A maintainer suggested that the effort put into the contribution after it was
-  accepted (i.e., the process of propagating the change to upper maintainers)
-  was considered. In other words, maintainers may \textbf{evaluate the
-  trade-off between the impact of contributions (not only deduplications) and
-  the struggle to integrate them into the complex Linux project}.
 \end{highlight-box}
-
-\noindent
-\begin{footnotesize}
-\textit{
-    Quote: ``Wrapping this up doesn't provide any real advantage, requiring as
-    it does the reviewer to look at this function AND where the value is set
-    rather than seeing them in one place.''
-  }
-\end{footnotesize}
 
@@
   parsing through the code~\cite{skylines-cognitive-load}.
+\hfill\\
+
+\noindent
+\begin{footnotesize}
+\textit{
+  Quote: ``Wrapping this up doesn't provide any real advantage, requiring as
+  it does the reviewer to look at this function AND where the value is set
+  rather than seeing them in one place.''
+}
+\end{footnotesize}
 \end{highlight-box}
+
+\begin{highlight-box}{Takeaway T5 (Performance)}
+  Given the low-level development context of Linux and the IIO subsystem, which
+  is usually related to devices with limited resources, we detected hints that
+  duplication may be forgivable if it improves performance.
+\hfill\\
 
@@
 \end{footnotesize}
+\end{highlight-box}
 
-\begin{highlight-box}{Takeaway T5 (Performance)}
-  Given the low-level development context of Linux and the IIO subsystem, which
-  is usually related to devices with limited resources, we detected hints that
-  duplication may be forgivable if it improves performance.
-\end{highlight-box}
+%TODO ADICIONADO APENAS PARA FECHAR 5 PGS CERTINHO. ME REMOVA!!!
+\hfill\\
 

### 2025-10-24 paper/techdebt2026/techdebt.tex
techdebt2026: techdebt: revise concluding remarks and add replication package
Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-The ethnographic studies, which included participant observation and experimentation with students, offered a realistic perspective on the opportunities and significant challenges in mitigating code duplications found by our tool.
-The tool was effectively used by first-time contributors to the Linux kernel to identify duplications and submit patches.
-%TODO: disponibilizar artefato da mesma forma que mandamos para a ICSE Education Track~\footnote{To avoid violating the double-anonymized review process, we will add the tool repository, patches, and artifact links in the final camera-ready version of the paper, if it is accepted for publication.}
+Ethnographic studies, including full participant and participant-as-observer observations, provided a realistic view of the opportunities and challenges in removing code duplications detected by our tool.
+First-time Linux kernel contributors could effectively use the tool to identify duplications and submit patches.
 
-While this research included 1 successfully merged patch by the main author, and 7 out of 11 student groups had their patches accepted and merged into Linux, a portion of the students' efforts encountered hurdles: 4 student groups had their patches rejected.
-%TODO: avaliar se não estamos sendo muito negativos ao destacar aqui as 4 rejeições.
-Despite reducing duplication, these difficulties often stemmed from maintainer feedback, where proposed changes were rejected or required substantial rework.
-This was often due to concerns about code readability, added abstraction, or the specific context of the code.
-Technical complexities, such as using C macros in configuration files, also contributed to the challenges.
-Additionally, lengthy patch review times further compounded these difficulties.
+While this research included one successfully merged patch by the authors, and 7 out of 12 patches sent by students were accepted and merged into Linux, a portion of the students' efforts encountered hurdles: 4 student groups had their patches rejected.
 
-Furthermore, the studies indicated that maintainers may not always view code duplication as inherently unfavorable, especially when it might improve clarity or facilitate independent hardware support (in the IIO subsystem matter).
-This suggests a nuanced understanding of code quality within the Linux kernel community, where practicality and maintainability in specific contexts can sometimes outweigh the general principle of eliminating all code duplication.
+Nevertheless, our work suggests that maintainers do not always view code duplication as inherently unfavorable, especially when it improves clarity, performance, and more.
+This suggests a nuanced understanding of code quality within the Linux kernel community that can go against the DRY principle, where practicality and maintainability in specific contexts sometimes outweigh the general principle of eliminating all code duplication.
 
-%\section*{Acknowledgments}
+%TODO ADICIONADO APENAS PARA FECHAR 5 PGS CERTINHO. ME REMOVA!!!
+\hfill\\
+
+\section*{Replication Package}
+
+All anonymized threads containing the entire review process and interactions with maintainers are available in this Zenodo repository: \href{https://doi.org/10.5281/zenodo.17432193}{https://doi.org/10.5281/zenodo.17432193}.
+
+%TODO ADD ANONYMIZED ARKANJO REPO
+
+To avoid violating the double-anonymized review process, we will add the tool repository, artifact links, and omitted references in the final camera-ready version of the paper if it is accepted for publication.
 

### 2025-10-25 paper/techdebt2026/techdebt.tex
techdebt2026: add reference and improve tool anonymization
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 The vast scale and continuous evolution of the Linux kernel make its maintenance a complex undertaking, where code duplication remains a persistent challenge that can hinder development and introduce bugs.
+%
 This paper presents a multimethod ethnographic study investigating the socio-technical dynamics of contributing to the Linux kernel by reducing identified code duplication.
+%
 To support this study, we developed a command-line utility tool for detecting function-level duplications, offering a concrete entry point for new contributors.
+%
 Our ethnographic approach, including complete participant observation and participant-as-observer analyses, demonstrated that addressing duplications is viable for lowering the contribution barrier.
 This point is evidenced by 8 of 13 (62\%) accepted contributions (patches) from 24 newcomers (16 undergraduate and 8 graduate students) in the Linux kernel project, collectively removing 1.397 lines of duplicated code.
+%
 Nevertheless, the study reveals a more intriguing and complex reality beyond mindlessly removing clones.
@@
 
-To address code duplication in the Linux kernel, we developed a command-line tool capable of detecting and analyzing function-level duplications.
-We conducted a multimethod ethnographic study to test the tool capabilities by identifying duplications and submitting contributions, directly or indirectly, to the Linux kernel, removing the redundant code.
+To address code duplication in the Linux kernel, we developed a command-line tool capable of detecting and analyzing function-level duplications\footnote{The tool and its documentation will be referenced after the review process.}.
+We conducted a multimethod ethnographic study leveraging the tool capabilities to identify duplications and submit contributions, directly or indirectly, to the Linux kernel, aiming to remove redundant code.
+%
 Through interactions with maintainers, this study gained insights into how the Linux community perceives the code quality related to code clones.
-Contrary to the DRY principle, our findings indicate that, in some contexts, maintainers can tolerate duplications for better readability, performance increases, or even to avoid integration overhead.\looseness=-1
+Contrary to the DRY principle, our findings indicate that, in some contexts, maintainers can tolerate duplications for better readability, performance increases, or even to avoid integration overhead.
+%\looseness=-1
 
 Ethnography is a well-established qualitative method for understanding people, their cultures, and work practices~\cite{bookethno}.
-It provides insights into community members’ values, beliefs, and practices~\cite{ethnosoft}.
-In this study, we first employed a complete participant observation approach, actively contributing by performing deduplications while documenting the process.
-We then conducted a participant-as-observer study with students resolving duplications identified by the tool.
-Beyond demonstrating that our tool lowered the entry barrier for new contributors, these studies yielded intriguing observations on kernel deduplications.
+It provides insights into community members' values, beliefs, and practices~\cite{ethnosoft}.
+In this study, we first employed a complete participant observation approach~\cite{gold1958roles}, actively contributing by performing deduplications while documenting the process.
+We then conducted a participant-as-observer study~\cite{gold1958roles} with students resolving duplications identified by the tool.
+Beyond demonstrating that the tool lowered the entry barrier for new contributors, these studies yielded intriguing observations on kernel Linux deduplications.
 This paper describes the aforementioned studies, along with our visions and reflections from diving into the development process.
@@
 We employed a multi-method research approach.
-We developed and evaluated a proposed tool using a method from the literature~[X].
-Given this validation, we proceeded to understand the viability of reducing duplications found by the tool with an ethnographic study using a participant observation methodology~\cite{}.
-In the first phase, we interacted with the Linux kernel community and got their feedback (\textit{complete participant observation}~\cite{}) through sending deduplication patches to the AMD Display driver.
+We developed and evaluated a proposed tool using a method from the literature~\cite{bigclonebench}.
+Given this validation, we proceeded to understand the viability of reducing duplications found by the tool with an ethnographic study using a participant observation methodology~\cite{gold1958roles}.
+In the first phase, we interacted with the Linux kernel community and got their feedback (\textit{complete participant observation}~\cite{gold1958roles}) through sending deduplication patches to the AMD Display driver.
 In the second phase, during a university course, we proposed a similar activity of sending patches to AMD Display or IIO to students.
-At the same time, we guided and mentored them throughout the process (\textit{participant-as-observer}~\cite{}).
-
-%Uniformizar este entendimento abaixo ao longo do texto:
-%Fase 1:Complete Participant
-%O pesquisador não estava observando outros desenvolvedores passarem pelo processo; ele se tornou o desenvolvedor.
-Ele se imergiu totalmente no papel de um contribuidor do kernel, enfrentando os mesmos desafios, normas sociais e barreiras técnicas que qualquer membro da comunidade enfrentaria.
-% O objetivo aqui é ganhar uma visão "êmica" (de dentro), entendendo a cultura e o processo por meio da experiência vivida.
-Isso é, por vezes, também chamado de autoetnografia.
-Ele não estava "fingindo", ele estava genuinamente tentando contribuir para entender as dores.
-
-%Fase 2: Participant-as-Observer
-%O papel do pesquisador (no caso pesquisadores) mudou fundamentalmente. Ele deixou de ser o ator principal (o contribuidor) e passou a ser o facilitador (mentor).
-%O foco dos dados coletados não é mais a autoexperiência do pesquisador, mas sim a experiência dos estudantes.
-%A participação ainda é alta (selecionar, direcionar, revisar), o que influencia ativamente o cenário.
-No entanto, a identidade principal perante os participantes (os estudantes) é a de pesquisador/mentor, e o objetivo é observar o processo deles.
+At the same time, we guided and mentored them throughout the process (\textit{participant-as-observer}~\cite{gold1958roles}).
 
@@
 Academic research on code duplication often focuses on pairwise comparison of code artifacts, which is not directly applicable to comprehensive codebase analyses.
-Other existing free software tools we explored in practice suffered from limited functionality~[X].
-Thus, we developed a command-line tool~[X] designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation.
-Released under the LGPL license, it employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
+Existing free software tools we explored in practice suffered from limited functionality~[omitted].
+Thus, we developed a command-line tool\footnote{Anonymous repository:  \url{https://anonymous.4open.science/r/arkanjo-C548}} designed to detect and analyze function-level code duplication within large codebases, such as the Linux kernel.
+The tool is released under the LGPLv3 ( GNU Lesser General Public License, version 3) and employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
 The first stage analyzes the entire codebase, parses C functions, and stores identified duplicates in a database.
-The second stage uses this database to quickly retrieve information about the identified duplications.
+The second stage uses this database to retrieve information about the identified duplications quickly.
 %The tool is available at \textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.
 
-Our tool is designed for the specific challenges of the kernel, with a focus on scalability to handle its size and prioritizing duplications that impact maintainability.
-The tool uses a text similarity method that applies TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine distance similarity to detect clones.
+The tool was designed to address the specific challenges posed by large-scale codebases, focusing on scalability and prioritization of duplications that may impact maintainability.
+It uses a text similarity method based on TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.
 
 We compared our tool against the BigCloneBench dataset~\cite{bigclonebench}.
-This validation showed that it can detect the majority of Type-1 and Type-2 duplications (100\% of Type-1 and 85\% of Type-2 with a cosine similarity threshold of 0.9)~[X], demonstrating sufficient accuracy for conducting our ethnographic studies.
+This validation showed that the tool is capable of detecting the majority of Type-1 and Type-2 duplications (100\% of Type-1 and 85\% of Type-2 with a cosine similarity threshold of 0.9), demonstrating sufficient accuracy to support our ethnographic studies.
 
@@
 
-To investigate whether we could use our tool to find ways to contribute to the Linux kernel, we conducted a complete participant observation experiment, where we acted as first-time contributors to the AMD Display driver, collecting artifacts of our experience in the process.
+To investigate whether we could use the tool to find ways to contribute to the Linux kernel, we conducted a complete participant observation experiment, where we acted as first-time contributors to the AMD Display driver, collecting artifacts of our experience in the process.
 
-We executed our tool in the AMD Display driver codebase.
+We executed the tool in the AMD Display driver codebase.
 We manually analyzed the biggest duplications regarding the number of lines, finding one pair that we judged promising to try to deduplicate.
@@
 Since the Linux kernel is primarily written in C, our systematic strategy consisted of eliminating code duplications through consolidating the duplicated code into a single library, which would replace instances of duplication across the codebase.
-For each duplicated function pair, for each function in the pair, we used our tool to locate all duplications of the function in the codebase, as it could have been duplicated more times than the two occurrences detected initially.
+For each duplicated function pair, for each function in the pair, we used the tool to locate all duplications of the function in the codebase, as it could have been duplicated more times than the two occurrences detected initially.
 This strategy resulted in a collection of duplicated functions and corresponding code files.
@@
 A core dynamic of the university course where we conducted the participant-as-observer study was to have students make a practical contribution to a free software project.
-In the first semester of 2025, we identified an opportunity to leverage our tool within this pedagogical structure.
+In the first semester of 2025, we identified an opportunity to leverage the tool within this pedagogical structure.
 As a contribution option, we proposed that students remove code duplications found in the Linux kernel as one possible way of contributing to a free software project.
@@
 For the second alternative, students were offered an experience similar to our initial complete participant observation.
-In this model, students were responsible for the entire workflow: executing our tool on the AMD Display drivers, independently analyzing the results to identify a potential deduplication, creating a patch with it, and sending the patch to the driver's maintainers.
+In this model, students were responsible for the entire workflow: executing the tool on the AMD Display drivers, independently analyzing the results to identify a potential deduplication, creating a patch with it, and sending the patch to the driver's maintainers.
 
-Of the students who chose to work on our tool-related tasks, 23 (16 undergraduate and 7 graduate) formed 11 groups to pursue the first alternative.
+Of the students who chose to work on the tool-related tasks, 23 (16 undergraduate and 7 graduate) formed 11 groups to pursue the first alternative.
 One graduate student opted for the second alternative.
@@
 
-Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach the duplications found by our tool as a comprehensive and more straightforward way to become contributors to the kernel.
+Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach the duplications found by the tool as a comprehensive and more straightforward way to become contributors to the kernel.
 We observed that not all duplications are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code and many other factors.
@@
 For a 90\% value, the result was a total of +333/-515 with an average of +55.5/-85.83.
-The 90\% value enabled our tool to identify less strict duplication cases, and, therefore, code segments with a greater extension, compared to the 100\% similarity entries, which were shorter segments, causing a difference in the average number of modified lines.
+The 90\% value enabled the tool to identify less strict duplication cases, and, therefore, code segments with a greater extension, compared to the 100\% similarity entries, which were shorter segments, causing a difference in the average number of modified lines.
 
@@
 
-Ethnographic studies, including full participant and participant-as-observer observations, provided a realistic view of the opportunities and challenges in removing code duplications detected by our tool.
+Ethnographic studies, including full participant and participant-as-observer observations, provided a realistic view of the opportunities and challenges in removing code duplications detected by the tool.
 First-time Linux kernel contributors could effectively use the tool to identify duplications and submit patches.
@@
 
-%TODO ADD ANONYMIZED ARKANJO REPO
-
-To avoid violating the double-anonymized review process, we will add the tool repository, artifact links, and omitted references in the final camera-ready version of the paper if it is accepted for publication.
+To avoid violating the double-anonymized review process, the anonymous repository of the developed tool is available at \url{https://anonymous.4open.science/r/arkanjo-C548}.
+The tool website link and omitted references will be added in the camera-ready version of the paper.
 

### 2025-10-27 paper/techdebt2026/techdebt.tex
techdebt2026: techdebt: change takeaway 4 to use correct reference
As pointed out in a TODO, the reference being used to base the claim
that jumping across code fragments increases cognitive load was
not the best one, and we needed to resolve the "base" reference the one
used cited. After some research, we settled with the Sillito et. al [1]
reference, as it better fits what we are trying to claim.

[1]: https://doi.org/10.1145/1181775.1181779

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-%TODO Na citação abaixo, trazida pelo Arthur, pensar se não vale colocar as
-%referências base do paper do Lanza: ``Developers need to mentally reconstruct
-%the role and behavior of a class from scattered fragments of code [8],
-%presented as one page of text at a time. This process is cognitively demanding
-%[9], slow, and error-prone, especially in large systems with many
-%interdependent components [10].''
-%
-% Referências do paper do Lanza para esse trecho:
-%[8] M.-A. Storey, “Theories, methods and tools in program comprehension:
-%Past, present and future,” in Proceedings of IWPC 2005. IEEE, 2005,
-%pp. 181–191.
-%[9] J. Sillito, G. C. Murphy, and K. De Volder, “Questions programmers
-%ask during software evolution tasks,” in Proceedings FSE 2006. ACM,
-%2006, pp. 23––34.
-%[10] G. C. Murphy, D. Notkin, and K. Sullivan, “Software reflexion models:
-%Bridging the gap between source and high-level models,” in Proceedings
-%of FSE 1995. ACM, 1995, pp. 18–28.
 \begin{highlight-box}{Takeaway T4 (Cognitive Load)}
-  Duplications that made it possible to read the code without needing to jump
-  to the function definition (to see its behavior) or the function call (to see
-  its concrete parameters) were justifiable to reduce cognitive load when
-  parsing through the code~\cite{skylines-cognitive-load}.
+  Needing to jump to a function definition (to see its behavior) or a function
+  call (to see its concrete parameters) in order to understand code is a
+  cognitively demanding task for developers~\cite{questions-programmers-ask}.
+  Duplications that avoided this were justifiable to reduce cognitive load when
+  parsing through the code.
 \hfill\\

### 2025-10-28 paper/techdebt2026/techdebt.tex
techdebt2026: (draft) organize takeaways according to result discussions
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 However, by the end of the feedback cycle, maintainers considered the patch from Group 1 irrelevant, while the patch from Group 4 was accepted.
+
+\noindent
+\begin{footnotesize}
+\textit{
+  Quote: ``I think the old code is more readable than hiding the values in a macro
+  even if it is duplicating a few lines of code.''
+}
+\end{footnotesize}
+
+\begin{highlight-box}{Takeaway T2 (Readability)}
+  Maintainers notably prioritized readability over removing duplications in
+  many different contexts. In their view, going against the \textbf{DRY
+  principle} was worth it if it resulted in code that is easier to read and
+  understand.
+\end{highlight-box}
+
 The reason is that Group 1 chose to apply the refactoring to a device that contained a register that was both read-only and non-volatile.
@@
 
+\noindent
+\begin{footnotesize}
+\textit{
+  Quote: ``[...] such patches might not worth it since the proposed
+  improvement is very small (and questionable) while the upstreaming process
+  still requires some effort.''
+}
+\end{footnotesize}
+
+\begin{highlight-box}{Takeaway T3 (Integration Overhead)}
+  A maintainer suggested that the effort put into the contribution after it was
+  accepted (i.e., the process of propagating the change to upper maintainers)
+  was considered. In other words, maintainers may \textbf{evaluate the
+  trade-off between the impact of contributions (not only deduplications) and
+  the struggle to integrate them into the complex Linux project}.
+\end{highlight-box}
+
 Groups 3, 6, and 10 created a struct to encapsulate the parameters and remove the duplication by applying the parameterized method.
-The patches from Groups 6 and 10 were rejected by the same maintainer who approved the patch from Group 3.
-The maintainer's justification for rejecting the patch from Group 6 was that it proposed to fix a duplication while adding more lines than removing, and that it was not convincing enough in other aspects, such as readability.
-The reason for rejecting the patch from Group 10 was that they refactored code files with simple context; therefore, adding a layer of abstraction to remove duplication decreased code readability without clear gains in codebase quality.
-Interestingly, both rejected groups addressed duplications across initialization functions for different devices, where code readability was vulnerable to refactoring methods.
-In contrast, Group 3 tackled the duplication of a simple iterative algorithm, which also contained a bug that could be inferred by inspecting both functions.
+The patches from Groups 6 and 10 were rejected by the same maintainer who approved the patch from Group 3, tackling the duplication of a simple iterative algorithm, which also contained a bug that could be inferred by inspecting both functions.
 In this case, the group refactored code from a more critical technical debt and took the opportunity to fix an unnoticed incorrect behavior.
 
+The maintainer's justification for rejecting the patch from Group 6 was that it proposed to fix a duplication while adding more lines than removing, and that it was not convincing enough in other aspects, such as readability.
+
+\begin{highlight-box}{Takeaway T4 (Cognitive Load)}
+  Needing to jump to a function definition (to see its behavior) or a function
+  call (to see its concrete parameters) in order to understand code is a
+  cognitively demanding task for developers~\cite{questions-programmers-ask}.
+  Duplications that avoided this were justifiable to reduce cognitive load when
+  parsing through the code.
+\hfill\\
+
+\noindent
+\begin{footnotesize}
+\textit{
+  Quote: ``Wrapping this up doesn't provide any real advantage, requiring as
+  it does the reviewer to look at this function AND where the value is set
+  rather than seeing them in one place.''
+}
+\end{footnotesize}
+\end{highlight-box}
+
+The reason for rejecting the patch from Group 10 was that they refactored code files with simple context; therefore, adding a layer of abstraction to remove duplication decreased code readability without clear gains in codebase quality.
+Interestingly, both rejected groups addressed duplications across initialization functions for different devices, where code readability was vulnerable to refactoring methods (T2  + T5).
+[blabla] Group 9 initially approached the selected duplication with a parameterized method, but after entering the feedback cycle, it was revealed that the problem represented a much more complex technical debt.
+From the first version to the last (version six), the patch was split into a series with four patches, where two of them have been applied, and the remaining entries are receiving suggestions.
+Even though the patch became a series of four, only one of them (one of the accepted) was strictly tackling duplication.
+
+\noindent
+\begin{footnotesize}
+\textit{
+  Quote: ``Note that is not an appropriate change for the large reads though
+  as \texttt{spi\_write\_then\_read()} bounces all buffers and so would add a
+  copy to those high(ish) performance paths.''
+}
+\end{footnotesize}
+
+\begin{highlight-box}{Takeaway T5 (Performance)}
+  Given the low-level development context of Linux and the IIO subsystem, which
+  is usually related to devices with limited resources, we detected hints that
+  duplication may be forgivable if it improves performance.
+\hfill\\
+\end{highlight-box}
+
 Groups 7 and 8 used the extract method to approach the deduplications.
-Group 8 sent their patch to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed.
+Group 8 sent their patch to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed (++T5).
 In the case of Group 7, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function.
 After the feedback cycle, maintainers accepted the patch in version 3.
-
-Group 9 initially approached the selected duplication with a parameterized method, but after entering the feedback cycle, it was revealed that the problem represented a much more complex technical debt.
-From the first version to the last (version six), the patch was split into a series with four patches, where two of them have been applied, and the remaining entries are receiving suggestions.
-Even though the patch became a series of four, only one of them (one of the accepted) was strictly tackling duplication.
 
@@
 
-We derived the following takeaways from \textbf{participant-as-observer observation}.
+We derived four takeaways from \textbf{participant-as-observer observation}.
 These takeaways are accompanied by example quotes from the maintainers that originated during the review process~\footnote{It is paramount to state that we \textbf{are not affirming} that these quotes represent the complete positioning of the maintainers who wrote them on code duplication.}.
-
-\begin{highlight-box}{Takeaway T2 (Readability)}
-  Maintainers notably prioritized readability over removing duplications in
-  many different contexts. In their view, going against the \textbf{DRY
-  principle} was worth it if it resulted in code that is easier to read and
-  understand.
-\hfill\\
-
-\noindent
-\begin{footnotesize}
-\textit{
-  Quote: ``I think the old code is more readable than hiding the values in a macro
-  even if it is duplicating a few lines of code.''
-}
-\end{footnotesize}
-\end{highlight-box}
-
-\begin{highlight-box}{Takeaway T3 (Integration Overhead)}
-  A maintainer suggested that the effort put into the contribution after it was
-  accepted (i.e., the process of propagating the change to upper maintainers)
-  was considered. In other words, maintainers may \textbf{evaluate the
-  trade-off between the impact of contributions (not only deduplications) and
-  the struggle to integrate them into the complex Linux project}.
-\hfill\\
-
-\noindent
-\begin{footnotesize}
-\textit{
-  Quote: ``[...] such patches might not worth it since the proposed
-  improvement is very small (and questionable) while the upstreaming process
-  still requires some effort.''
-}
-\end{footnotesize}
-\end{highlight-box}
-
-\begin{highlight-box}{Takeaway T4 (Cognitive Load)}
-  Needing to jump to a function definition (to see its behavior) or a function
-  call (to see its concrete parameters) in order to understand code is a
-  cognitively demanding task for developers~\cite{questions-programmers-ask}.
-  Duplications that avoided this were justifiable to reduce cognitive load when
-  parsing through the code.
-\hfill\\
-
-\noindent
-\begin{footnotesize}
-\textit{
-  Quote: ``Wrapping this up doesn't provide any real advantage, requiring as
-  it does the reviewer to look at this function AND where the value is set
-  rather than seeing them in one place.''
-}
-\end{footnotesize}
-\end{highlight-box}
-
-\begin{highlight-box}{Takeaway T5 (Performance)}
-  Given the low-level development context of Linux and the IIO subsystem, which
-  is usually related to devices with limited resources, we detected hints that
-  duplication may be forgivable if it improves performance.
-\hfill\\
-
-\noindent
-\begin{footnotesize}
-\textit{
-  Quote: ``Note that is not an appropriate change for the large reads though
-  as \texttt{spi\_write\_then\_read()} bounces all buffers and so would add a
-  copy to those high(ish) performance paths.''
-}
-\end{footnotesize}
-\end{highlight-box}
-
-%TODO ADICIONADO APENAS PARA FECHAR 5 PGS CERTINHO. ME REMOVA!!!
-\hfill\\
 

### 2025-10-28 paper/techdebt2026/techdebt.tex
techdebt2026: general review and add some TODO
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 Contrary to the DRY principle, our findings indicate that, in some contexts, maintainers can tolerate duplications for better readability, performance increases, or even to avoid integration overhead.
-%\looseness=-1
 
@@
 Beyond demonstrating that the tool lowered the entry barrier for new contributors, these studies yielded intriguing observations on kernel Linux deduplications.
-This paper describes the aforementioned studies, along with our visions and reflections from diving into the development process.
+This paper describes the aforementioned studies, along with our visions and reflections from diving into the Linux kernel development process.
 
@@
 Each subsystem typically has a dedicated maintainer or a team of maintainers overseeing its development and managing contributions.
-The development process is coordinated through Git, and contributions are formatted as patches, text documents that outline the differences between two source code versions.
+The development process is coordinated through Git (also designed by Linus Torvald, focused on the Linux kernel development), and contributions are formatted as patches, text documents that outline the differences between two source code versions.
 These patches are then submitted to the relevant Linux mailing lists for public review and discussion.
@@
 This subsystem is substantial, with over 391,000 lines of code across more than 1,089 files.
-We also investigate the \textit{Industrial I/O (IIO)} subsystem, often considered an accessible entry point for new contributors.
+We also investigate the \textit{Industrial Input/Output (IIO)} subsystem, often considered an accessible entry point for new contributors.
 The IIO subsystem provides support for devices that perform analog-to-digital or digital-to-analog conversions\footnote{\href{https://kernel.org/doc/html/v6.17/driver-api/iio/intro.html}{https://kernel.org/doc/html/v6.17/driver-api/iio/intro.html}} and contains over 281,772 lines of code in more than 755 files.
@@
 The tool was designed to address the specific challenges posed by large-scale codebases, focusing on scalability and prioritization of duplications that may impact maintainability.
-It uses a text similarity method based on TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.
+It uses a text similarity method based on TF-IDF (Term Frequency-Inverse Document Frequency) vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.
 
@@
 
-To investigate whether we could use the tool to find ways to contribute to the Linux kernel, we conducted a complete participant observation experiment, where we acted as first-time contributors to the AMD Display driver, collecting artifacts of our experience in the process.
+To investigate whether we could use the tool to find ways to contribute to the Linux kernel, we first conducted a complete participant observation experiment, where the main author (a graduate student) acted as a first-time contributor to the AMD Display driver, collecting artifacts of our experience in the process.
 
@@
 Thus, we proposed a simple systematic approach to mitigate all the functions duplicated in the context, not just the duplicated function pairs.
-After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions.
+After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our Replication Package).
 
@@
 
-A core dynamic of the university course where we conducted the participant-as-observer study was to have students make a practical contribution to a free software project.
-In the first semester of 2025, we identified an opportunity to leverage the tool within this pedagogical structure.
-As a contribution option, we proposed that students remove code duplications found in the Linux kernel as one possible way of contributing to a free software project.
+To broaden our study and observations, we adopted a participant-as-observer approach, involving students as free software project newcomers to make practical contributions to the Linux kernel.
+This approach formed the core activity of a university course, in which we guided students to remove code duplications in the Linux kernel as one possible pathway for contributing to the project in the first phase of the course~[omitted].
 
-The 2025 course offering had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of two or three members.
-To assist students in this task, we prepared multiple alternatives for contributing to free software projects, including simpler contributions such as updating documentation or fixing code style issues.
-In this context, two options related to removing duplications in the Linux kernel were offered.
+The 2025 course offering had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of two or three members, but graduate students could work individually.
+To assist students in this task, we (a professor and three more experienced graduate students) prepared several alternatives for contributing to the Linux kernel, including simpler contributions such as fixing code style issues.
+In this context, we specified two options for removing duplications in the Linux kernel.
 
-In the first option, we ran the tool on the IIO subsystem, creating a list of duplicated function pairs.
+As the first option, we ran the tool on the IIO subsystem, creating a list of duplicated function pairs.
 Then, we curated the list even more to get more actionable duplications for the students using this approach:
@@
 The students' task was to select a recommended entry from the list, refactor the code to eliminate these duplications, and then submit their patches to the IIO mailing list for the maintainers' review.
+%TODO: explicar o processo do envio para o mentores/monitores, o tipo de feedback que recebiam antes enviarem.
 
 For the second alternative, students were offered an experience similar to our initial complete participant observation.
-In this model, students were responsible for the entire workflow: executing the tool on the AMD Display drivers, independently analyzing the results to identify a potential deduplication, creating a patch with it, and sending the patch to the driver's maintainers.
+In this approach, students were responsible for the entire workflow: executing the tool, independently analyzing the results to identify a potential deduplication, creating a patch, and sending the patch to the driver's maintainers.
+%TODO: aqui, no caso do Sebastian, também houve a revisão intermediária dos mentores?
 
@@
 
-Using blog posts, the groups documented their experiences and approaches to refactoring the duplications, as well as their experiences submitting a patch to the Linux kernel.
+Using blog posts, the groups documented their experiences and approaches to refactoring the duplications, as well as their experiences submitting a patch to the Linux kernel~[omitted].
 We also analyzed these posts to understand the refactoring patterns used in deduplications, the experience of sending patches, and the opinions of subsystem maintainers on them.
+%TODO: é fácil e queremos disponibilizar esses blogs também no pacote de replicação?
 
@@
 
-In the \textbf{complete participant observation} study, we chose one function pair found on AMD Display Driver to approach and apply the systematic approach.
-
+In the \textbf{complete participant observation} study (first phase), we chose one function pair found on AMD Display Driver to approach and apply the systematic approach.
+%
 We reached a point where we had an idea of how to remove the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver.
@@
 
+%TODO: dado que falamos da demora do feedback, colocar a data de envia da V1.
 In the feedback on the first submitted version, the maintainers only asked for minor changes to align with the coding style, correct license use, and align with initially unknown conventions.
@@
 
-\begin{highlight-box}{Takeaway T1:}
-  \textbf{As the primary result of our complete participant observation we
-  understood that AMD Display driver developers do not necessarily view
+Since we suffered a delay in the review process, and to understand why, we directly contacted a maintainer.
+It was mentioned to us that one reason for the delay was that duplicated code enhances the independence of GPU driver code.
+This allows developers to make changes to a specific GPU without needing to test compatibility with others, which helps save a significant amount of time and effort.
+This example of hardware variations, also mentioned by Kapser and Godfrey~\cite{cloneharm}, corroborates our understanding that GPU driver developers do not view duplicated code negatively by default.
+Thus, as the primary result of our complete participant observation we understood that:
+
+\begin{highlight-box}{Takeaway T1}
+  Driver developers do not necessarily view
   duplicated code negatively, which contradicts standard software engineering
-  practices}. We suffered a delay in the review process, and to understand why,
-  we directly contacted a maintainer. It was mentioned to us that one reason for
-  the delay was that duplicated code enhances the independence of GPU driver
-  code. This allows developers to make changes to a specific GPU without needing
-  to test compatibility with others, which helps save a significant amount of
-  time and effort. This example of hardware variations, also mentioned by Kapser
-  and Godfrey~\cite{cloneharm}, corroborates our understanding that GPU driver
-  developers do not view duplicated code negatively by default.
+  practices.
 \end{highlight-box}
 
-Regarding our analysis from the \textbf{participant-as-observer} phase, Table~\ref{tab:stu} summarizes all the contributions sent by newcomer groups (GID)~\footnote{As in the first phase, we were also newcomers, we included our experience in the table as Group 0.
+Table~\ref{tab:stu} summarizes all the contributions sent by the main author and student groups (GID)~\footnote{As in the first phase, the first author is also a newcomer, we included this experience in the table as Group 0.
 Nevertheless, the subsequent analysis will focus more on the patches sent by students.}.
@@
 
-We observed that most groups applied the parameterized refactor method while a few used the extractor method.
+Regarding our analysis from the \textbf{participant-as-observer} phase, we observed that most groups applied the parameterized refactor method while a few used the extractor method.
 

### 2025-10-28 masterthesis/conteudo/apendices/contribution.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

\chapter{Students Contribution to ArKanjo Tool}

\label{app:contribution}

This appendix includes the references to the students contributions,
the issue list proposed to the students and figures of the documentation webpage.

This appendix includes the header message of the resended patch and the 
patch initial reply.

\section{Issue List}

\begin{figure}
\includegraphics[scale=0.63]{patch_header}
\caption{List of issues in ArKanjo proposed to students to contribute}
\label{fig:issue_list}
\end{figure}

\newpage

\section{Students Contribution Reference}

Group 1 contribution was made through two change request.
The first change request can be accessed in the url: 
\url{https://github.com/LipArcanjo/arkanjo/pull/8}. 
The second change request can be accessed in the url: 
\url{https://github.com/LipArcanjo/arkanjo/pull/8}.

Group 2 contribution was splited in three change requests. The first change request
can be accessed in the url: \url{https://github.com/LipArcanjo/arkanjo/pull/7}.
The second change request can be accessed in the url:
\url{https://github.com/LipArcanjo/arkanjo/pull/9}. 
The third change request can be accessed in the 
url: \url{https://github.com/LipArcanjo/arkanjo/pull/11}.

\section{Documentation Webpage}

\begin{figure}
\includegraphics[scale=0.45]{webpage_1}
\caption{Documentation webpage - Main Page}
\label{fig:webpage1}
\end{figure}

\begin{figure}
\includegraphics[scale=0.45]{webpage_2}
\caption{Documentation webpage - Class List Page}
\label{fig:webpage2}
\end{figure}

### 2025-10-28 masterthesis/conteudo/apendices/students-artifacts.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

\chapter{Students Artifacts}

\label{app:stuart}

This appendix includes the artifacts produced by the students at University of São Paulo.

\paragraph{Group 1}

\

\textbf{Blog 1:} \textit{linux.ime.usp.br/~juliacalixto/software-livre.html}

\textbf{Blog 2:} \textit{linux.ime.usp.br/~soueuabianca/post2.html}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250426194135.244774-1-biancalvao@gmail.com/}}

\paragraph{Group 2}

\

\textbf{Blog 1:} \textit{mikhael-abdallah.github.io/software-livre/posts/aula-06/}

\textbf{Blog 2:} \textit{augustomb02.github.io/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/all/20250429093923.347370-1-mikhael\_abdallah@usp.br/}}

\paragraph{Group 3}

\

\textbf{Blog 1:} \textit{beavct.github.io/}

\textbf{Blog 2:} \textit{gabriela-victor-github-io.vercel.app/blog/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250424002144.23260-1-beatrizvianacosta16@gmail.com/}}

\textbf{Accepted PATCH (v1):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250424002144.23260-1-beatrizvianacosta16@gmail.com/}}

\paragraph{Group 4}

\

\textbf{Blog 1:} \textit{linux.ime.usp.br/~aschwarz}

\textbf{Blog 2:} \textit{\seqsplit{whip-foxtrot-b57.notion.site/MAC0470-Desenvolvimento-de-Software-Livre-2025-1dcde8837d}}

\textbf{Blog 3:} \textit{www.linux.ime.usp.br/~artpilone/sl\_blog.html}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250421145534.91146-1-arthurpilone@usp.br/}}

\textbf{Accepted PATCH (v3):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250426163311.546500bc@jic23-huawei/}}

\paragraph{Group 5}

\

\textbf{Blog 1:} \textit{gust-vaz.github.io/}

\textbf{Blog 2:} \textit{xivor.github.io/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250426153307.59689da0@jic23-huawei/T/\#t}}

\textbf{Accepted PATCH (v1):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250426153307.59689da0@jic23-huawei/T/\#t}}

\paragraph{Group 6}

\

\textbf{Blog 1:} \textit{gabriellimmaa.github.io/}

\textbf{Blog 2:} \textit{vitormarquesr.github.io/blog/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/all/20250625101857.00004dbf@huawei.com/}}

\paragraph{Group 7}

\

\textbf{Blog 1:} \textit{isacaselli.github.io/}

\textbf{Blog 2:} \textit{rodrigocmichelassi.github.io/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250428132551.176788-1-bellacaselli20@gmail.com/}}

\textbf{Accepted PATCH (v3):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250531160807.39e54096@jic23-huawei/}}

\paragraph{Group 8}

\

\textbf{Blog 1:} \textit{linux.ime.usp.br/~nattan/mac470/mac470.html}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/all/20250422191114.32832-1-nattanferreira58@gmail.com/\#t}}

\textbf{Accepted PATCH (v4):} \textit{\seqsplit{lore.kernel.org/all/20250614124401.32b17eda@jic23-huawei/}}

\paragraph{Group 9}

\

\textbf{Blog 1:} \textit{andrewijano.github.io/mac5856/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250423170526.15143-1-andrew.lopes@alumni.usp.br/}}

\textbf{Accepted PATCH (partially accepted on v6):} \textit{\seqsplit{lore.kernel.org/linux-iio/20250621183843.2f8bcb48@jic23-huawei/}}

\paragraph{Group 10}

\

\textbf{Blog 1:} \textit{linux.ime.usp.br/~apolo/mac0470/}

\textbf{Blog 2:} \textit{ncbor.github.io/}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/linux-iio/20250428034222.318898-1-gabrielgeraldinosouza@gmail.com/T/\#u}}

\paragraph{Group 11}

\

\textbf{Blog 1:} \textit{\seqsplit{saguileran.github.io/blog/2025/free-software-development/}}

\textbf{PATCH in Linux lore:} \textit{\seqsplit{lore.kernel.org/amd-gfx/f78f68ba-80e0-47ca-97a2-37a15d076e68@amd.com/T/\#t}}

\textbf{Accepted PATCH (v2):} \textit{\seqsplit{https://lore.kernel.org/amd-gfx/CADAW+GDN-4Lwd87h793csWF6r5cgqpgrpEPzRyDV9+rfkNPmTA@mail.gmail.com/}}

### 2025-10-28 masterthesis/conteudo/apendices/tool/oitenta.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

\section{Function pairs with similarity between 79\% and 81\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/dpp/dcn401/dcn401\_dpp.c

\textbf{Function name 1:} dpp401\_dpp\_setup

\textbf{File 2:} dc/dpp/dcn10/dcn10\_dpp.c

\textbf{Function name 2:} dpp1\_cnv\_setup

\textbf{Similarity:} 79.60\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 2 }

\

\textbf{File 1:} dc/dml/dcn32/dcn32\_fpu.c

\textbf{Function name 1:} dcn32\_calculate\_wm\_and\_dlg\_fpu

\textbf{File 2:} dc/dml/dcn31/dcn31\_fpu.c

\textbf{Function name 2:} dcn31\_calculate\_wm\_and\_dlg\_fp

\textbf{Similarity:} 80.41\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/resource/dcn316/dcn316\_resource.c

\textbf{Function name 1:} dcn316\_resource\_construct

\textbf{File 2:} dc/resource/dce80/dce80\_resource.c

\textbf{Function name 2:} dce81\_construct

\textbf{Similarity:} 79.37\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/mpc/dcn401/dcn401\_mpc.c

\textbf{Function name 1:} get3dlut\_config

\textbf{File 2:} dc/dpp/dcn20/dcn20\_dpp\_cm.c

\textbf{Function name 2:} get3dlut\_config

\textbf{Similarity:} 80.62\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/resource/dce100/dce100\_resource.c

\textbf{Function name 1:} dce100\_hwseq\_create

\textbf{File 2:} dc/resource/dce120/dce120\_resource.c

\textbf{Function name 2:} dce121\_hwseq\_create

\textbf{Similarity:} 79.34\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 6 }

\

\textbf{File 1:} dc/resource/dce80/dce80\_resource.c

\textbf{Function name 1:} dce80\_construct

\textbf{File 2:} dc/resource/dcn314/dcn314\_resource.c

\textbf{Function name 2:} dcn314\_resource\_construct

\textbf{Similarity:} 79.19\%

\textbf{Duplication type:} Type-4

\paragraph{Function pair 7 }

\

\textbf{File 1:} dc/clk\_mgr/dcn315/dcn315\_clk\_mgr.c

\textbf{Function name 1:} dcn315\_clk\_mgr\_destroy

\textbf{File 2:} dc/clk\_mgr/dcn35/dcn35\_clk\_mgr.c

\textbf{Function name 2:} dcn35\_clk\_mgr\_destroy

\textbf{Similarity:} 79.16\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/clk\_mgr/dcn401/dcn401\_clk\_mgr\_smu\_msg.c

\textbf{Function name 1:} dcn401\_smu\_send\_msg\_with\_param

\textbf{File 2:} dc/clk\_mgr/dcn30/dcn30\_clk\_mgr\_smu\_msg.c

\textbf{Function name 2:} dcn30\_smu\_send\_msg\_with\_param

\textbf{Similarity:} 80.30\%

\textbf{Duplication type:} Type-3

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/link/link\_dpms.c

\textbf{Function name 1:} link\_set\_dpms\_off

\textbf{File 2:} dc/link/link\_dpms.c

\textbf{Function name 2:} link\_set\_dpms\_on

\textbf{Similarity:} 80.74\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/clk\_mgr/dcn30/dcn30\_clk\_mgr\_smu\_msg.c

\textbf{Function name 1:} dcn30\_smu\_set\_hard\_min\_by\_freq

\textbf{File 2:} dc/clk\_mgr/dcn401/dcn401\_clk\_mgr\_smu\_msg.c

\textbf{Function name 2:} dcn401\_smu\_set\_hard\_min\_by\_freq

\textbf{Similarity:} 80.97\%

\textbf{Duplication type:} Type-3

### 2025-10-28 masterthesis/conteudo/apendices/tool/setenta.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

\section{Function pairs with similarity between 69\% and 71\%}

\paragraph{Function pair 1 }

\

\textbf{File 1:} dc/resource/dcn21/dcn21\_resource.c

\textbf{Function name 1:} read\_dce\_straps

\textbf{File 2:} dc/resource/dcn321/dcn321\_resource.c

\textbf{Function name 2:} read\_dce\_straps

\textbf{Similarity:} 69.52\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 2 }

\

\textbf{File 1:} amdgpu\_dm/amdgpu\_dm\_pp\_smu.c

\textbf{Function name 1:} pp\_rv\_set\_min\_deep\_sleep\_dcfclk

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm\_pp\_smu.c

\textbf{Function name 2:} pp\_rv\_set\_hard\_min\_fclk\_by\_freq

\textbf{Similarity:} 69.11\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 3 }

\

\textbf{File 1:} dc/hubbub/dcn35/dcn35\_hubbub.c

\textbf{Function name 1:} convert\_and\_clamp

\textbf{File 2:} dc/dml/dml1\_display\_rq\_dlg\_calc.c

\textbf{Function name 2:} is\_dual\_plane

\textbf{Similarity:} 69.41\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 4 }

\

\textbf{File 1:} dc/irq/dcn35/irq\_service\_dcn35.c

\textbf{Function name 1:} to\_dal\_irq\_source\_dcn35

\textbf{File 2:} dc/dml/calcs/dcn\_calcs.c

\textbf{Function name 2:} tl\_sw\_mode\_to\_bw\_defs

\textbf{Similarity:} 70.65\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 5 }

\

\textbf{File 1:} dc/spl/dc\_spl\_scl\_filters.c

\textbf{Function name 1:} spl\_get\_filter\_4tap\_16p

\textbf{File 2:} dc/spl/dc\_spl\_scl\_filters.c

\textbf{Function name 2:} spl\_get\_filter\_5tap\_64p

\textbf{Similarity:} 70.47\%

\textbf{Duplication type:} Type-2

\paragraph{Function pair 6 }

\

\textbf{File 1:} include/fixed31\_32.h

\textbf{Function name 1:} dc\_fixpt\_max

\textbf{File 2:} dc/dml/calcs/dcn\_calc\_math.c

\textbf{Function name 2:} dcn\_bw\_max

\textbf{Similarity:} 70.50\%

\textbf{Duplication type:} Type-3

\paragraph{Function pair 7 }

\

\textbf{File 1:} amdgpu\_dm/amdgpu\_dm\_crc.c

\textbf{Function name 1:} dm\_need\_crc\_dither

\textbf{File 2:} amdgpu\_dm/amdgpu\_dm\_crc.c

\textbf{Function name 2:} dm\_is\_crc\_source\_dprx

\textbf{Similarity:} 69.73\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 8 }

\

\textbf{File 1:} dc/optc/dcn35/dcn35\_optc.c

\textbf{Function name 1:} optc35\_set\_drr

\textbf{File 2:} dc/optc/dcn35/dcn35\_optc.c

\textbf{Function name 2:} optc35\_set\_long\_vtotal

\textbf{Similarity:} 70.46\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 9 }

\

\textbf{File 1:} dc/resource/dcn201/dcn201\_resource.c

\textbf{Function name 1:} dcn201\_hwseq\_create

\textbf{File 2:} dc/resource/dcn10/dcn10\_resource.c

\textbf{Function name 2:} dcn10\_hwseq\_create

\textbf{Similarity:} 69.21\%

\textbf{Duplication type:} Negative

\paragraph{Function pair 10 }

\

\textbf{File 1:} dc/resource/dce80/dce80\_resource.c

\textbf{Function name 1:} dce80\_link\_encoder\_create

\textbf{File 2:} dc/resource/dcn10/dcn10\_resource.c

\textbf{Function name 2:} dcn10\_link\_encoder\_create

\textbf{Similarity:} 70.72\%

\textbf{Duplication type:} Type-4

### 2025-10-28 masterthesis/conteudo/background/codeduplication.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-08-06 conteudo/background/codeduplication.tex
@@
 
-We use the \textit{diff} command as the base of one of the code duplication detection 
+We use the \textit{diff} as the base of one of the code duplication detection 
 methods on the tool proposed in this work, although it is not explored as the text 

### 2025-10-28 masterthesis/conteudo/background/nlpsimilarity.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-08-05 conteudo/background/nlpsimilarity.tex
@@
 
-Gensim is a FLOSS Python library \citep{gensim}, which its creators
+Gensim is a FLOSS Python library~\citep{gensim}, which its creators
 describe as ``The fastest library for training of vector embeddings -- Python or

### 2025-10-28 masterthesis/conteudo/background/refactor.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-07-14 conteudo/background/refactor.tex
@@
 
-Fowler defines refactoring as "a controlled technique for improving the design
+Martin Fowler defines refactoring as ``a controlled technique for improving the design
 of an existing code base. Its essence is applying a series of small
@@
 doing. However, the cumulative effect of each of these transformations is quite
-significant" \citep{refactorbook}. The elimination of detected code clones and
+significant''~\citep{refactorbook}. Eliminating detected code clones and
 duplication is a refactoring task that requires the application of refactoring
@@
 \textit{parameterize method}, \textit{replace parameter with explicit methods},
-and \textit{replace parameter with method} \citep{refactorbook}. Below is a
+and \textit{replace parameter with method}~\citep{refactorbook}. Below is a
 brief explanation of the methods used in our research.
@@
 
+\paragraph{Parameterize Method.}
+The Parameterize Method involves replacing multiple similar code blocks with a
+single function. This function incorporates parameters to account for the
+different behaviors across similar code snippets, allowing it to adjust
+functionality based on the parameter values~\citep{refactorbook}.
+Figure~\ref{fig:parameter} illustrates an example of the Parameterize Method.
+
 \begin{figure}[ht]
@@
 
-\paragraph{Parameterize Method.}
-
-The Parameterize Method involves replacing multiple similar code blocks with a
-single function. This function incorporates parameters to account for the
-different behaviors across similar code snippets, allowing it to adjust
-functionality based on the parameter values \citep{refactorbook}. Figure
-\ref{fig:parameter} illustrates an example of the Parameterize Method.
-
 \paragraph{Extract Method.}
-
 The Extract Method entails breaking down a complex code artifact into multiple,
@@
 contrast, a complex code artifact might not be entirely reusable, potentially
-resulting in further code duplication \citep{refactorbook}. Figure
-\ref{fig:extract} demonstrates an example of the Extract Method.
+resulting in further code duplication~\citep{refactorbook}.
+Figure~\ref{fig:extract} demonstrates an example of the Extract Method.
 
@@
 advantages of this method is reducing the number of unnecessary methods and 
-abstraction in the code, making the code more straightforward
-\citep{refactorbook}.
+abstraction in the code, making the code more straightforward~\citep{refactorbook}.

### 2025-10-28 masterthesis/conteudo/introduction/intro.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-08-03 conteudo/introduction/intro.tex
@@
 Maintaining the kernel is an enormous task, involving more than 28 million lines of 
-code and contributions from over twenty thousand developers
+code
 \footnote{Data directly extracted from Linux github repository~\citep{linuxrepo}.
 Counted  28783390 lines of code and 20699 contributors on May 29th, 2025.
-Number of contributors counted as unique full names listed as a patch author.}.
+Number of contributors counted as unique full names listed as a patch author.}
+and over 2,000 developers are actively contributing to the kernel
+\footnote{Data from \url{https://lwn.net/Articles/1022414/}}.
 Due to the scale of the project, not all contributions adhere to best practices, resulting 
@@
 they find. Despite searching formal and grey literature, existing solutions failed to address
-addressing these specific challenges. The list of related tools examined in this exploratory 
+these specific challenges. The list of related tools examined in this exploratory 
 research is available in Appendix~\ref{app:gray}.

### 2025-10-28 masterthesis/conteudo/methods/methods.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-08-05 conteudo/methods/methods.tex
@@
 
-\section{Evaluation of the Tool by the BigCloneBench dataset}
+\section{Tool Evaluation}
+
+\subsection{Evaluation of the Tool by the BigCloneBench dataset}
 \label{sec:metbig}
@@
 
-\section{Evaluation of the Tool by Empirical Analysis}
+\subsection{Evaluation of the Tool by Empirical Analysis}
 
@@
 We empirically evaluated these pairs to determine if they were duplications, comparing these results with the literature-based method to assess our tool effectiveness and understand the impact of the minimum similarity parameter.
+
+\subsection{Performance Evaluation of the Tool}
+
+We have not conducted a deep performance benchmark to understand the tool's performance across various 
+computer configurations. To gain an understanding of the time consumption of the tool's Preprocessor and 
+Query Responder steps, we conducted a brief benchmark using a machine equipped with a Ryzen 5700X processor 
+and 32GB of RAM. We executed both steps five times for both the AMD Display driver and the IIO subsystem.
 

### 2025-10-28 masterthesis/conteudo/results/results.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-06-12 conteudo/results/results.tex
@@
 In this chapter, we present and analyze the results of the experiments and methods 
-used in our research. Following the chronology of the study, we start by presenting the 
-evaluation of the ArKanjo tool through both a method from the literature and an empirical 
+used in our research. Following the chronology of the study, we start by presenting 
+the evaluation of the ArKanjo tool through both a method from the literature and an empirical
 analysis. Then, we present and discuss the experience of the initial participant 
-observation methodology, where we mitigate a code duplication on the AMD Display Driver 
-and send it as a patch. Finally, we show and discuss our discoveries of the non-participant observation
-of the students approaches to mitigate duplications and patches's discussions.
+observation methodology, where we mitigate one code duplication on the AMD Display 
+driver and send it as a patch. Finally, we show and discuss our discoveries of 
+the non-participant observation of the students' approaches to mitigate duplications 
+and patches' discussions.
 

### 2025-10-28 masterthesis/conteudo/resumoabstract.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-07-24 conteudo/resumoabstract.tex
@@
 % definidas antes do resumo/abstract. Acrescente quantas forem necessárias.
-\palavraschave{Software livre, kernel Linux, duplicação de código, refatoração, manutenção de software}
+\palavraschave{Software Livre, Kernel Linux, Duplicação de Código, Refatoração, Manutenção de Software}
 
-\keywords{Free software,Linux kernel, Code duplication, Refactoring, code maintenance}
+\keywords{Free Software,Linux Kernel, Code Duplication, Refactoring, Code Maintenance}
 
@@
 
-O kernel Linux é um projeto de software livre amplamente utilizado, alimentando 96,3\% 
-dos um milhão maiores servidores web e 85\% dos smartphones em todo o mundo. Dada a sua extensa 
+O kernel Linux é um projeto de software livre amplamente utilizado, alimentando a
+maioria dos servidores web e smartphones em todo o mundo.
+Dada a sua extensa 
 base de usuários, qualquer adição de recursos, correção de bugs ou tratamento de 
 vulnerabilidades de segurança pode impactar milhões de pessoas. Manter o kernel é 
-uma tarefa enorme, envolvendo mais de 19 milhões de linhas de código e contribuições 
-de mais de mil desenvolvedores. Devido à escala do projeto, nem todas as contribuições 
-seguem as melhores práticas, resultando em artefatos de código de baixa qualidade que 
-podem complicar a manutenção e o desenvolvimento de novos recursos. Um desses artefatos 
-é a duplicação de código, que é o foco desta pesquisa. As ferramentas existentes 
+uma tarefa enorme, envolvendo mais de 28 milhões de linhas de código e contribuições 
+de mais de vinte mil desenvolvedores. Devido à escala do projeto, nem todas as contribuições 
+seguem as melhores práticas, resultando em um código de qualidade inferior que pode complicar 
+a manutenção e o desenvolvimento de funcionalidades.
+Um dos cenários que pode levar à baixa qualidade de código é a duplicação de código,
+que é o foco deste trabalho As ferramentas existentes 
 para detectar duplicação de código normalmente comparam dois artefatos de código para 
@@
 conseguem resolver os desafios específicos do kernel Linux. 
-Esta pesquisa apresenta o ArKanjo, uma ferramenta de linha de comando para a manutenção 
-do kernel Linux, projetada para detectar e analisar duplicações em nível de função. 
+Este trabalho apresenta uma abordagem para tratar o problema da duplicação de código no contexto do 
+kernel Linux; para apoiar essa abordagem, desenvolvemos o ArKanjo, uma ferramenta de 
+linha de comando projetada para detectar e analisar duplicações em nível de função.
 Lançado sob a licença MIT, o ArKanjo emprega uma arquitetura de dois estágios que consiste 
@@
 orientação a futuros contribuidores. Avaliamos o ArKanjo em casos de duplicação do mundo 
-real em versões recentes do kernel, demonstrando sua eficácia na identificação de clones 
+real em versões recentes do kernel, demonstrando sua capacidade de identificação de clones 
 problemáticos que ferramentas genéricas frequentemente ignoram. Ao identificar instâncias 
 de duplicação bem definidas e gerenciáveis, o ArKanjo reduz efetivamente a barreira para 
-novos contribuidores, uma capacidade evidenciada por seu papel em orientar estudantes a 
+novos contribuidores, um aspecto evidenciado por seu papel em orientar estudantes a 
 fazerem suas primeiras melhorias de código no kernel.
@@
 The Linux kernel is a widely used Free/Libre/Open Source Software (FLOSS) project, powering
-96.3\% of the top one million web servers 
-and 85\% of smartphones worldwide. Given its extensive user base, any addition of features, 
+the majority of webservers and smartphones worldwide.
+Given its extensive user base, any addition of features, 
 bug fixes, or handling of security vulnerabilities can impact millions. Maintaining the kernel 
@@
 twenty thousand developers. Due to the scale of the project, not all contributions adhere to best 
-practices, resulting in poor-quality code artifacts that may complicate maintenance and feature 
-development. One such artifact is code duplication, which is the focus of this research. 
+practices, resulting in code of lower quality that may complicate maintenance and feature development.
+One of the scenarios that can lead to low code quality is code duplication, 
+which is the focus of this work. 
 Existing tools for detecting code duplication typically compare two code artifacts to determine 
@@
 short of addressing the specific challenges of the Linux kernel.
-This research presents ArKanjo  , a command-line tool for Linux kernel maintenance designed to detect and
-analyze function-level duplications. Released under the MIT
-license, ArKanjo employs a two-stage architecture consisting
+This work presents an approach to address the code duplication issue in the Linux kernel context; 
+to support this approach, we developed ArKanjo, a command-line tool designed 
+to detect and analyze function-level duplications.
+Released under the MIT license, ArKanjo employs a two-stage architecture consisting
 of a Preprocessor and a Query Responder that separates
@@
 against real-world duplication cases in recent kernel versions, demonstrating
-its effectiveness in identifying problematic clones that generic
+its capability on identifying problematic clones that generic
 tools often overlook. By identifying well-defined, manageable
 duplication instances, ArKanjo effectively lowers the barrier
-for new contributors, a capability evidenced by its role in guiding students to make their first 
+for new contributors, a aspect evidenced by its role in guiding students to make their first 
 code improvements to the kernel.

### 2025-10-28 masterthesis/conteudo/tool/functionalities.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-05-25 conteudo/tool/functionalities.tex
@@
 Each functionality accepts specific parameters to perform its operations, with an optional parameter shared by all: \textbf{minimum similarity per query}.
-This minimum similarity per query is distinct from the minimum similarity setting introduced in Section \ref{subsec:setup}, as it enables a dynamic adjustment of the similarity threshold within a query-specific context.
+This minimum similarity per query is distinct from the minimum similarity setting introduced in Section~\ref{subsec:setup}, as it enables a dynamic adjustment of the similarity threshold within a query-specific context.
 Below is an explanation of each of these main functionalities.
@@
 We implement optional filters to facilitate more complex queries, as described below.
-Figure \ref{fig:explorer_ex} demonstrates an example of this functionality in use.
+Figure~\ref{fig:explorer_ex} demonstrates an example of this functionality in use.
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[scale=0.30]{explorer_example}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[scale=0.3]{function_example}
@@
 A function matches a pattern if the concatenated string of the relative path and function name contains the pattern as a substring.
-Figure \ref{fig:function_ex} demonstrates this functionality in action.
+Figure~\ref{fig:function_ex} demonstrates this functionality in action.
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[scale=0.3]{relatory_example}
@@
 The tool utilizes the Code Duplication Database to list duplicated functions and the temporary codebase to determine the line count for these functions to generate this report.
-Figure \ref{fig:relatory_ex} demonstrates an example of this functionality in action.
+Figure~\ref{fig:relatory_ex} demonstrates an example of this functionality in action.

### 2025-10-28 masterthesis/conteudo/tool/tool.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-06-01 conteudo/tool/tool.tex
@@
 help developers identify code clone duplication at the function level.
-Specifically, it enables the detection of pairs of functions within a codebase
-that are clones of each other.
+Specifically, it enables the detection of pairs of functions that are clones of 
+each other within a codebase.
 
 The tool operates in two main parts: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
-The \textbf{Preprocessor} is responsible for performing heavy computations to find the code 
+The \textbf{Preprocessor} performs heavy computations to find the code 
 duplications across a codebase and produce artifacts to consult information related to the 
@@
 The ArKanjo tool has been implemented and is available for reference at
-\url{https://github.com/LipArcanjo/arkanjo}.
+\url{github.com/arkanjo-tool/arkanjo}.

### 2025-10-28 papers/cbsoftVEM2025/main.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

\documentclass[12pt]{article}

\usepackage{sbc-template}

\usepackage{graphicx,url}
\usepackage{hyperref}
\usepackage{nicematrix}

\usepackage{array}

%\usepackage[brazil]{babel}   
\usepackage[utf8]{inputenc}

\usepackage{seqsplit}

\sloppy

\title{Mitigating Code Duplication in the Linux Kernel}

%% Double-Blind review
%%%%%%%%%%%%%%%%%%%%%%
%\author{Luan Arcanjo\inst{1}, Marcelo Spessoto\inst{1}, David Tadokoro\inst{1}, Paulo Meirelles\inst{1}}

%\address{FLOSS Competence Center\\
%        Institute of Mathematics and Statistics\\
%        University of São Paulo
%  \email{\{luanicaro,marcelomspessoto\}@usp.br,paulormm@ime.usp.br}
%}

\begin{document}

\maketitle

\begin{abstract}
The vast scale and continuous evolution of the Linux kernel make its maintenance a complex undertaking, where code duplication remains a persistent challenge that can hinder development and introduce bugs. This paper presents a multimethod ethnographic study investigating the socio-technical dynamics of contributing to the Linux kernel by mitigating identified redundancies. To support this investigation, we developed \texttt{anonymous-tool}, a command-line utility for detecting function-level duplications, offering a concrete entry point for new contributors. Our ethnographic approach, including participant and non-participant observations, demonstrated that addressing duplications is viable for lowering the contribution barrier. This point is evidenced by the main author and 7 of 11 (64\%) student groups having their patches accepted into the kernel, collectively removing 991 lines of duplicated code. Moreover, the study also reveals a more complex reality beyond simple clone removal: an analysis of maintainer feedback on both accepted and rejected patches highlights a nuanced understanding of code quality within the Linux kernel community, where the benefits of eliminating duplication are carefully weighed against factors such as readability, the introduction of new abstractions, and the specific context of the code.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Coordinating hundreds or thousands of developers working on a software product is complex. Creating similar or even completely duplicated code is a frequent and often unavoidable outcome. Excessive duplication can negatively impact the maintainability of a project. When duplications occur, any change to a repeated code segment typically requires manual replication across all other copies. Developers must, therefore, remain constantly vigilant to synchronize different parts of the codebase, which is a highly error-prone and seemingly simple responsibility. With many duplications, modifications become slow, complex, and fragile. Removing or refactoring duplicated code demands great care to preserve the original behavior. Even minor mistakes can introduce subtle state changes that may lead to serious problems in the future.

The Linux kernel is a foundational Free/Libre and Open Source Software (FLOSS) project, a vital component of the world's digital infrastructure. Its maintenance is a massive undertaking, encompassing over 28 million lines of code and involving contributions from over twenty thousand developers. Within this context, code duplication remains a significant challenge,  reducing the code readability and increasing the risk of introducing bugs~\cite{harmone,harmtwo}. This issue is particularly prominent in kernel device drivers, which account for over 66\% of the Linux kernel source code.

The detection of code duplication, or \textit{code clones}, has been a research topic for decades~\cite{firstman}. The literature offers a widely accepted taxonomy that classifies clones into four types based on their similarity, from exact copies (Type-1) to semantically equivalent but syntactically different fragments (Type-4)~\cite{litreview}. Several detection techniques have emerged, including textual, token-based, tree-based, and graph-based approaches~\cite{litreview}. These efforts have culminated in state-of-the-art techniques, such as the graph-based approach proposed by Liu et al.~\cite{tailor}.

To address the code duplication issue, we developed \texttt{anonymous-tool}, a command-line tool designed to support Linux kernel maintenance by detecting and analyzing function-level duplications. This paper presents a multimethod ethnographic study to validate the tool capabilities in identifying duplications and supporting contributions to the Linux kernel by mitigating redundant code. Therefore, the study investigates how the Linux kernel community perceives the code quality related to code clones. Our findings indicate that maintainers consider not only duplication but also factors such as readability and appropriate levels of abstraction regarding code maintainability.

Ethnography is a well-established qualitative research method used to understand people, their cultures, and their work practices~\cite{bookethno}. It enables researchers to gain insights into values, beliefs, and practices from the perspective of community members~\cite{ethnosoft}.  
In this study, we first adopted a participant observation approach, where the main author actively engaged in mitigating code duplications while documenting the process and outcomes. Later, we followed a non-participant observation involving students tasked with resolving duplications identified by the tool. These combined studies showed that \texttt{anonymous-tool} can collaborate to reduce the entry barrier for new contributors. The student groups and the main author used the tool to guide their initial contributions to the Linux kernel.

\section{The Linux Kernel}

The Linux kernel is a monolithic system organized into subsystems, such as the process scheduler, memory management, and device drivers. Each subsystem typically has a dedicated maintainer or a team of maintainers overseeing its development and managing contributions. The development process is coordinated through the Git version control system. Contributions are formatted as patches, text documents that outline the differences between two source code versions. These patches are then submitted to the relevant Linux mailing lists for public review and discussion.

This study focuses on two of these subsystems, selected to represent different contexts for contribution. We explore the AMD Display driver, chosen for its importance within the hardware ecosystem, which is responsible for enabling AMD GPU functionality in Linux. This subsystem is substantial, with over 391,000 lines of code across more than 1,089 files. We also investigate the Industrial I/O (IIO) subsystem, often considered an accessible entry point for new contributors. The IIO subsystem provides support for devices that perform analog-to-digital or digital-to-analog conversions\footnote{\href{https://kernel.org/doc/html/v4.14/driver-api/iio/intro.html}{kernel.org/doc/html/v4.14/driver-api/iio/intro.html}} and contains over 281,772 lines of code in more than 755 files.

\section{Anonymous-tool}

We developed \texttt{anonymous-tool}, a command-line tool designed to detect and analyze function-level code duplication within the Linux kernel to support our investigation. Released under the MIT license, \texttt{anonymous-tool} employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results. The first stage, the Preprocessor, analyzes the entire codebase, parses C functions, and stores identified duplicates in a database. The second stage, the Query Responder, uses this database to retrieve information about the identified duplications quickly.
%The tool is available at \textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.

\texttt{Anonymous-tool} is designed for the specific challenges of the kernel, with a focus on scalability to handle its size and prioritizing duplications that impact maintainability. The tool's primary role in this study was to provide a concrete entry point for new contributors by identifying function-level duplications to address. Figure~\ref{fig:diagrama} illustrates the tool's architecture. The tool uses a text similarity method that applies TF-IDF vector embedding, implemented with the Gensim library~\cite{gensim}, and computes cosine similarity to detect clones.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.33]{diagrama_mestrado}
\caption{Architecture diagram with the tool components
}
\label{fig:diagrama}
\end{figure}

\section{Participant Observation Study}
\label{sec:participant}

To investigate whether we can use the \texttt{anonymous-tool} tool to find ways to contribute to the Linux kernel, we conducted a participant observation experiment in the second semester of 2024, where we acted as first-time contributors to the AMD Display driver, collecting artifacts of our experience in the process.

In the AMD Display driver codebase, we executed the \texttt{anonymous-tool}. We manually analyzed the biggest duplications regarding the number of lines and found two duplicated function pairs that we judged interesting to try to mitigate. We have chosen to approach the biggest duplications as proof of concept so that we can make significant impacts.

The first duplicated function pair chosen is function \textit{offset\_to\_id} on files \textit{dc/gpio/dcn32/hw\_translate\_dcn32.c} and \textit{dc/gpio/dcn315/hw\_translate\_dcn315.c}. The second duplicated function pair is function \textit{phy\_id\_to\_atom} in file 
\textit{ \seqsplit{dc/bios/dce110/command\_table\_helper\_dce110.c}} and \textit{\seqsplit{dc/bios/dce60/command\_table\_helper\_dce60.c}}.
Hereafter, the mitigation on \textit{offset\_to\_id} will be referred to as mitigation 1, and the mitigation on 
\textit{phy\_id\_to\_atom} will be referred to as mitigation 2.

Given the two duplicated function pairs, we observed that the files in the functions 
contain multiple other duplications. Thus, we proposed a simple systematic approach 
to mitigate all the functions duplicated in the context, not just the duplicated 
function pairs. 
After we approached the duplications with the systematic strategy, for the duplicated 
function pairs that had successfully mitigated the duplication, we sent patches to the 
AMD Display driver email list to validate the mitigations and see the maintainer's opinion 
while documenting the process and our impressions.
The systematic approach is defined below.

In C programming, communication between source files is achieved by creating header files that specify 
libraries~\cite{Cbook}. Since the Linux kernel is primarily written in C, we eliminated code duplications 
by consolidating duplicated code into a single library, which would replace instances of 
duplication across the codebase.

For each duplicated function pair, for each function in the pair, we used the \texttt{anonymous-tool} to locate all duplications of the function in the codebase, as it may have been duplicated more times than the two occurrences contained in the duplicated function pair. 
This strategy resulted in a collection of duplicated functions and corresponding code files.

To identify shared code more effectively, we extended this approach to search for other common 
functions across all collection files. We then applied specific refactoring methods to each shared 
function. If the functions were identical across files, we removed the duplicates and created a single 
function in the library. If modifications existed, we applied case-specific refactoring.

\subsection{Results}
\label{subsec:participantresults}

Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
files and lines changed and the refactoring method used. The following sections present our experience and learning for each mitigation.

\begin{table}
\centering
\caption{Refactor methods used and mitigation metrics in the participant-observation experiment.}
\begin{tabular}{ | m{20mm} | m{15mm} | m{15mm} | m{15mm} | m{40mm} | }

\hline

\textbf{Mitigation} & \textbf{Files changed} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
\\ \hline

1 & 13 & +224 & -753 & Parameterize Method, Extract Method  \\ \hline
2 & 8 & +114 & -520 & None \\ \hline

\hline
\end{tabular}

\label{tab:patch}
\end{table}

\subsubsection{Mitigation 1}

In the systematic approach, we found that the \textit{offset\_to\_id} function 
exists in 10 code files, enclosing multiple GPU architectures. The duplicated 
functions are not exactly equal, as some minimal specific logic is applied to each GPU architecture. To address this issue, we used the extract method to 
break the function into smaller parts and the parameterized method to address 
the specific logic for the GPUs.

We reached a point where we had a scratch on how to mitigate the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver. We found that the configuration files on the AMD Display driver are designed to be imported at compilation time using \textit{\#define} macros. The configuration files' design choice makes refactoring duplications on generic approaches trickier, as refactoring code that depends on these files requires significant refactoring of the configuration files' design and deep knowledge of the codebase, which we, as first-time contributors, do not have. Thus, we opted not to continue investigating this mitigation.

\subsubsection{Mitigation 2}

Using the systematic approach, we found that the \textit{phy\_id\_to\_atom} function exists in five code files, and two other functions are duplicated 
in those five code files. All functions across the files were exactly equal, 
so there was no need to apply the refactoring methods presented in the literature. 
The refactoring was resumed to create a generic library and fix the compilation targets.

The code files in the second mitigation do not depend on configuration files. Thus, 
we did not face the same issues during the first mitigation. Refactoring the code 
to mitigate duplications reached a point we judged was good enough to submit as a 
patch to contribute to the AMD Display driver code quality. Thus, we moved to send 
the refactoring to the driver maintainers as a patch, documenting our findings and experiences 
on the process.

\begin{figure}
\centering
\includegraphics[scale=0.45]{timeline_patch}
\caption{Timeline of the PATCH submission process}
\label{fig:timeline}
\end{figure}

The experience of submitting a patch was not as we expected. The process took 7 months and 16 days, significantly slower than expected. We initially submitted the patch on August 9, 2024, but we did not receive replies. Thus, we had to resend it on October 9, 2024. We received an initial response on November 3, 2024, requesting minimal changes. We submitted a second version with the requested changes on November 16, 2024. We did not receive replies initially and will need to resend the patch on February 7, 2025. We received a reply on February 11, 2025, asking for additional changes. We sent a third version on February 24, 2025, and the patch was accepted and integrated into the kernel codebase on February 25, 2025. Figure \ref{fig:timeline} illustrates the timeline of the patch submission.

After the initial reply from the maintainers, we observed that the review process became more responsive. We hypothesize that once a patch receives initial attention, it is more closely tracked by maintainers. Another contributing factor to the initial delay could be the maintainers' workload at the time of the first submission.

We discussed the Display driver with our point of contact at AMD to understand the reason for the slow patch submission process. We learned that some of the code in the driver is shared across all operating systems that support the implemented GPUs. This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors. Additionally, we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering. One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others. This approach helps save a significant amount of time and effort.

Regarding the patch changes requested for the second version, the maintainers only asked for minor changes to align with code style, correct license use, and best practices that were initially unknown. For the change requested on the third version, the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.

\section{Non-participant Observation Study}

A core dynamic of the course (concurrently for undergraduate and postgraduate studies) is to have students make a practical contribution to a free software project. In the first semester of 2025, the professor teaching the course, who is also the advisor of this work, identified an opportunity to leverage the \texttt{anonymous-tool} tool within this existing pedagogical structure. As a contribution option, the professor proposed that students mitigate code duplications found in the Linux kernel. The professor drove this initiative as a practical course activity, and the author of this thesis was not involved in using the tool to identify duplications or in developing approaches to mitigate them. The tasks were carried out entirely by the course staff and the students.

The course had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of 
two or three members. To assist students in this task, the professor and teaching assistants prepare multiple 
alternatives for contributing to free software projects, including simpler contributions such as 
updating documentation or fixing code style issues. In this context, two new options related to the 
\texttt{anonymous-tool} were proposed alongside the others.

For the first anonymous-tool-related alternative, a teaching assistant ran the tool on the IIO subsystem twice, creating two lists of duplicated function pairs, one with a 100\% similarity threshold and the other with a 90\% similarity threshold. Then, the following process happened to both lists:

\begin{enumerate}
    \item The teaching assistant filtered the lists to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different devices may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
    \item The teacher assistant removed duplications that were too short and would not represent a significant change for a patch application;
    \item The teacher assistant manually ranked code duplications and marked some duplications with insights based on personal opinion and perceptions over the remaining entries. This was done to guide students so that they could approach the technical debt with a greater chance of having their contribution merged into the code.
\end{enumerate}

The students' task was to select a recommended entry from the list, refactor the code to eliminate these specific duplications, and then submit their patches to the IIO maintainers for review.

For the second alternative, students were offered an experience similar to our initial participant 
observation. In this model, the students were responsible for the entire workflow: executing the 
\texttt{anonymous-tool} on the AMD display driver, independently analyzing the results to identify a duplication 
to fix, creating a patch to mitigate it, and sending the patch to the driver's maintainers.

Of the students who chose to work on the \texttt{anonymous-tool}-related tasks, 23 (16 undergraduate and 7 graduate) 
formed 11 groups to pursue the first alternative. One graduate student also opted for the second alternative. Notably, none of these students had prior experience contributing to the Linux kernel, making this their first attempt to submit a patch as newcomers to the project.
%TODO: ver na versão final incluir o patch do Lucas (Octavio e Fernando)

The groups were requested to document their experiences and approaches to refactoring the duplications, as well as their experiences submitting a patch to the Linux kernel, in blogs. We analyzed these blogs to understand the refactoring patterns used to mitigate duplications, the experience of sending patches, and the opinions of subsystem maintainers on the mitigations.

\subsection{Overview of the submitted patches}

Table \ref{tab:stu} summarizes the student groups' mitigation approaches. It contains the information about the chosen subsystem and the similarity threshold for groups that opted for the first approach. Each entry also has the patch status, refactoring method, and code diff corresponding to the final patch versions before it was dropped or accepted into the code base. The code diff for a patchset corresponds to the sum of added and removed lines of each of its patches.

\begin{table}[ht]
\centering
\caption{Refactor methods used and metrics of the students' mitigations.}
\resizebox{1\linewidth}{!}{%
\begin{NiceTabular}{ | c | m{20mm} | m{25mm} | m{30mm} | m{46mm} | m{28mm} | }[hvlines]
\hline
\textbf{Group} & \textbf{Subsystem} & \textbf{Similarity} & \textbf{Patch Status} & \textbf{Refactoring Method} & \textbf{Code Diff} \\
\hline

1 & IIO & 100\% & Dropped (v2) & Undefined & +7/-14 \\ \hline
2 & IIO & 100\% & Dropped (v1) & Parameterize method & +23/-27 \\ \hline
3 & IIO & 100\% & Accepted (v1) & Parameterize method & +12/-30 \\ \hline
4 & IIO & 100\% & Accepted (v3) & Undefined & +2/-7 \\ \hline
5 & IIO & 90\% & Accepted (v1) & Parameterize method & +17/-70 \\ \hline
6 & IIO & 90\% & Dropped (v1) & Parameterize method & +95/-92 \\ \hline
7 & IIO & 90\% & Accepted (v3) & Extract method & +20/-36 \\ \hline
8 & IIO & 90\% & Accepted (v4) & Extract method & +20/-16 \\ \hline
9 & IIO & 90\% & Accepted (v6) -- Partially & Parameterize method & +149/-243 \\ \hline
10 & IIO & 90\% & Dropped (v1) & Undefined & +32/-58 \\ \hline
\Block{2-1}{11} & IIO & 100\% & Dropped (v1) & Inline method & +2/-12 \\ 
& AMD & 90\% & Accepted (v2) & Undefined & +90/-489 \\ \hline

\end{NiceTabular}%
}
\label{tab:stu}
\end{table}

Seven groups applied the parameterized refactor method, two groups used the extractor method, and one group applied the inline method to their approaches to mitigate their duplications. Four patches do not contain a refactoring technique defined in the literature (undefined). Six groups had their patches accepted and fully merged into Linux; four groups had their patches dropped; one group has its patchset partially applied and is still working on patch submission, currently in the sixth version of the patch when this document was written, but we can also consider this case as accepted.

Groups 1 and 4 have worked in similar duplications, where two functions are involved: one to determine if a register is volatile and the other to determine if it is writable. These are two functions that return opposite Boolean results. Group 1 initially addressed the problem using the extract method, while Group 2 used macros. After feedback from maintainers, both solutions converged to keep one of the functions while changing the content of the other to return a negated query of the former function. However, by the end of the feedback cycle, maintainers considered the patch from Group 1 irrelevant, while the patch from Group 4 was accepted. The reason is that Group 1 has chosen to apply the refactoring to a device that contains a register that is both read-only and non-volatile. Thus, the technical debt selected by Group 1 was not merely a duplication but an incorrect implementation of the method.

Groups 3, 6, and 10 created a struct to encapsulate the parameters and mitigate duplication when applying the parameterized method. The patches from Groups 6 and 10 were rejected by the same maintainer who approved the patch from Group 3. The maintainer's justification for rejecting the patch from Group 6 is that it proposes to fix duplication while adding more lines than removing, and is not convincing enough in other aspects, such as readability. The reason for rejecting the patch from Group 10 is that they refactored code files with simple context; therefore, adding a layer of abstraction to remove duplication decreases code readability without clear gains in codebase quality. The maintainers' feedback for Group 10 was:
\begin{quote}

\textit{``A few comments inline, but overall I'm not sure the code reduction is sufficient to cover the resulting loss of readability. Sometimes a switch is simply clearer than a partial look-up table.''}

\textit{``So nice idea, but I'm not seeing it as being a good move for this particular code. I'd be slightly interested to see the optimized output of the two approaches, but this is far from a high-performance path, so we care a lot more about readability here.''} 
\end{quote}

Interestingly, both groups addressed duplications across initialization functions for different devices, where code readability is vulnerable to refactoring methods. In contrast, Group 3 tackled the duplication of a simple iterative algorithm, which also contained a bug that could be inferred by inspecting both functions. In this case, the group refactored code from a more critical technical debt and took the opportunity to fix an unnoticed incorrect behavior.

%\begin{figure}
%\centering
%\includegraphics[scale=0.8]{group_10_reply}
%\caption[Maintainers' Feedback on Group 10 Patch]{
% Maintainers' Feedback on Group Ten's Patch
%}
%\label{fig:group_10_reply}
%\end{figure}

Groups 7 and 8 used the extract method to approach the mitigations. Group 8 sent the mitigations to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed. For Group 7 mitigation, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function. After the feedback cycle, maintainers accepted the patch in version 3.

Group 9 initially approached the selected duplication with a parameterized method, but after entering the feedback cycle, it was revealed that the problem represented a much more complex technical debt. The patch is still under development, currently at version 6, and has been separated into a 4-patch patch set, where half of them have been applied, and the remaining entries are receiving suggestions. Fortunately, maintainers are receptive and provide constant feedback; the contribution is expected to be fully merged if the students do not abandon the current work.

Group 11 is the student who executed the tool and completed the process of identifying duplications. The student identified a duplication in the driver that occurred across eighteen code files, creating a patch to remove approximately four hundred lines of duplicated code. The student sent the mitigations to the maintainers, and their feedback pointed to fixing the warning in the code without arguing about removing the duplications. After addressing the issue, the patch was applied. This student also contributed to IIO, tackling a code duplication with inline method refactoring. This one, however, was received with more skepticism by the maintainer, and the student dropped the contribution because he did not think the feedback instructions were clear enough.

Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach duplications found by the \texttt{anonymous-tool} and become contributors to the kernel. We observed that not all duplications 
are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code, code readability, and the levels of abstraction added to address the duplications.

\subsection{Statistics of the contributions}

The eleven groups, which have submitted 11 contributions to the IIO subsystem, have collectively requested the insertion of 379 lines and the deletion of 605 lines in the subsystem, representing an average of 34.45 lines added and 55 lines removed per contribution.

Considering the sum of fully merged contributions, we have a code diff value of +51/-143, with an average of +12.75/-35.75. This ratio is better than dropped contributions, which have a total of +179/-219 with an average of +29.83/-36.5.

It is also possible to notice a difference based on the similarity threshold used to detect duplications. The proposed patches to address 100\%-similarity duplications summed up to +46/-90 (average of +9.2/-18). For a 90\% value, the result was a total of +333/-515 with an average of +55.5/-85.83. The 90\% value enabled \texttt{anonymous-tool} to identify less strict duplication cases, and, therefore, code segments with a greater extension, compared to the 100\% similarity entries, which were shorter segments, causing a difference in the average number of modified lines.

Finally, the AMD contribution from Group 11, which involves the complete use of the \texttt{anonymous-tool} toolchain, got a completely different outcome. With a total of 90 lines inserted and 489 lines removed, there are some points to be considered for the success and high modification extension of the contribution:

\begin{itemize}
    \item The AMD Display subsystem has some self-reported problems regarding overall technical debt;
    \item The AMD Display subsystem has a more centralized code structure and maintainer hierarchy, enabling contributors to tackle code duplications across a broader scope and multiple files, which happened to Group 11's contribution;
    \item The complete control over the process of finding code duplication gave Group 11 the opportunity to identify and reason better over the duplications found by \texttt{anonymous-tool}, enabling a more strategic approach to searching for promising patch possibilities. 
\end{itemize}

\section{Concluding Remarks}

The ethnographic studies, which included participant observation and experimentation with students, offered a realistic perspective on the opportunities and significant challenges in mitigating code duplications found by the \texttt{anonymous-tool} tool. The tool was effectively used by first-time contributors to the Linux kernel to identify duplications and submit patches~\footnote{To avoid violating the double-anonymized review process, we will add the tool repository, patches, and artifact links in the final camera-ready version of the paper, if it is accepted for publication.}.

While this research included 1 successfully merged patch by the main author, and 7 out of 11 student groups had their patches accepted and merged into Linux, a portion of the students' efforts encountered hurdles: 4 student groups had their patches rejected.
%TODO: avaliar se não estamos sendo muito negativos ao destacar aqui as 4 rejeições.
Despite reducing duplication, these difficulties often stemmed from maintainer feedback, where proposed changes were rejected or required substantial rework. This was often due to concerns about code readability, added abstraction, or the specific context of the code. Technical complexities, such as using C macros in configuration files, also contributed to the challenges. Additionally, lengthy patch review times further compounded these difficulties.

Furthermore, the studies indicated that maintainers may not always view code duplication as inherently unfavorable, especially when it might improve clarity or facilitate independent hardware support (in the IIO subsystem matter). This suggests a nuanced understanding of code quality within the Linux kernel community, where practicality and maintainability in specific contexts can sometimes outweigh the general principle of eliminating all code duplication.

%\section*{Acknowledgments}

\newpage
\bibliographystyle{sbc}
\bibliography{sbc-template}

\end{document}

### 2025-10-28 papers/debconf2025/paper.tex
organize folders
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-07-03 paper/paper.tex
@@
 The students' efforts demonstrated that newcomers could use the tool to contribute effectively. 
-Of the 11 students' patches, 5 were accepted, 5 were rejected, and 1 required follow-up 
+Of the 11 students' patches, 6 were accepted, 4 were rejected, and 1 required follow-up 
 fixes based on maintainer feedback. To remove the duplicated code, 7 student groups used the 
@@
 
-This paper presented ArKanjo, a novel tool designed to detect effectively and facilitate the refactoring of function-level code duplication within the Linux kernel. 
+This paper presented ArKanjo, a novel tool designed to detect effectively and facilitate the refactoring of 
+function-level code duplication within the Linux kernel. 
 Evaluations demonstrated that ArKanjo is highly effective at identifying Type-1 and Type-2 
 clones, which are often indicative of propagated issues such as copied buggy logic. We validated the practical 
-value of the tool through an ethnographic study in which the main author and student groups used ArKanjo to identify actionable duplications.
-This effort led to 13 patch submissions, including 1 from the main author and 5 from student groups that were successfully merged into the kernel
+value of the tool through an ethnographic study in which the main author and student groups used ArKanjo 
+to identify actionable duplications. This effort led to 13 patch submissions, including 1 
+from the main author and 6 from student groups that were successfully merged into the kernel
 \footnote{Reference for the accepted patches:  
 \textit{\href{https://github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}{github.com/arkanjo-tool/arkanjo/blob/main/PATCHES.md}}.
-}, collectively removing 991 lines of duplicated code. While some refactoring efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 
+}, collectively removing 987 lines of duplicated code. While some refactoring efforts were declined due to maintainer priorities, ArKanjo successfully lowered the barrier 
 for new contributors to make meaningful improvements. The tool provides immediate value for 

### 2025-10-28 papers/debconf2025/paper.tex
debconf2025: update ArKanjo's License
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 This paper presents \textit{ArKanjo}, a novel command-line tool for Linux kernel maintenance 
-designed to detect and analyze function-level duplications. Released under the MIT 
-license, ArKanjo employs a two-stage architecture consisting of a Preprocessor and 
+designed to detect and analyze function-level duplications. Released under the GNU Lesser General Public License (LGPL), ArKanjo employs a two-stage architecture consisting of a Preprocessor and 
 a Query Responder that separates computationally intensive analysis from efficient 
@@
 ArKanjo, our proposed tool, is a Command Line Interface (CLI) application designed to
-help developers identify code clone duplication at the function level. Released under the
-MIT license, ArKanjo is available at 
+help developers identify code clone duplication at the function level. Released under
+LGPLv3 (GNU Lesser General Public License, version 3), ArKanjo is available at 
 \textit{\href{https://github.com/arkanjo-tool/arkanjo}{github.com/arkanjo-tool/arkanjo}}.

### 2025-10-28 papers/techdebt2026/techdebt.tex
techdebt2026: minor adjustments on the results (part 1)
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

--- 2025-10-28 paper/techdebt2026/techdebt.tex
@@
 Thus, we developed a command-line tool\footnote{Anonymous repository:  \url{https://anonymous.4open.science/r/arkanjo-C548}} designed to detect and analyze function-level code duplication within large codebases, such as the Linux kernel.
-The tool is released under the LGPLv3 ( GNU Lesser General Public License, version 3) and employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
+The tool is released under LGPLv3 (GNU Lesser General Public License, version 3) and employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
 The first stage analyzes the entire codebase, parses C functions, and stores identified duplicates in a database.
@@
 
-In the \textbf{complete participant observation} study (first phase), we chose one function pair found on AMD Display Driver to approach and apply the systematic approach.
+In the \textbf{complete participant observation} study (first phase), we chose one function pair found on the AMD Display Driver to approach and apply the systematic approach.
 %
-We reached a point where we had an idea of how to remove the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver.
+We reached a point where we had an idea of removing the duplications found, but we could not complete the refactoring to a state capable of submitting as a patch to the driver.
 We found that the configuration files on the AMD Display driver are designed to be imported at compilation time using \textit{\#define} macros.
@@
 Since we suffered a delay in the review process, and to understand why, we directly contacted a maintainer.
-It was mentioned to us that one reason for the delay was that duplicated code enhances the independence of GPU driver code.
-This allows developers to make changes to a specific GPU without needing to test compatibility with others, which helps save a significant amount of time and effort.
-This example of hardware variations, also mentioned by Kapser and Godfrey~\cite{cloneharm}, corroborates our understanding that GPU driver developers do not view duplicated code negatively by default.
-Thus, as the primary result of our complete participant observation we understood that:
+From this conversation, we understand that one reason for the delay was that duplicated code enhances the independence of GPU driver code.
+This approach allows developers to make changes to a specific GPU without needing to test compatibility with others, which helps save a significant amount of time and effort.
+This example of hardware variations, also mentioned by Kapser and Godfrey~\cite{cloneharm}, leads us to the primary conclusion of our complete participant observation:
 
@@
 
+%TODO: precisamos fazer um gancho, para sinalizar que a partir do que observamos no T1, vimos a necessidade de investigar a visão dos mantenedores a partir de outros patches relacionados a redução de duplicação...
 Table~\ref{tab:stu} summarizes all the contributions sent by the main author and student groups (GID)~\footnote{As in the first phase, the first author is also a newcomer, we included this experience in the table as Group 0.

### 2025-10-28 papers/techdebt2026/techdebt.tex
techdebt2025: update discussion regarding Takeaway 2
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 
-Since we suffered a delay in the review process, and to understand why, we directly contacted a maintainer.
+Since we suffered a delay in the review process, we directly contacted a maintainer to understand why.
 From this conversation, we understand that one reason for the delay was that duplicated code enhances the independence of GPU driver code.
@@
 \begin{highlight-box}{Takeaway T1}
-  Driver developers do not necessarily view
-  duplicated code negatively, which contradicts standard software engineering
-  practices.
+Driver developers do not necessarily view duplicated code negatively, which contradicts standard software engineering practices.
 \end{highlight-box}
 
-%TODO: precisamos fazer um gancho, para sinalizar que a partir do que observamos no T1, vimos a necessidade de investigar a visão dos mantenedores a partir de outros patches relacionados a redução de duplicação...
+Building on this insight, we realized that the perception identified in T1 might not be limited to driver developers alone.
+To deepen our understanding, we systematically examined other contributions related to code duplication reduction and analyzed how different maintainers responded to such patches.
+This broader analysis aimed to capture diverse maintainer perspectives and contextual factors influencing their decisions.
+
 Table~\ref{tab:stu} summarizes all the contributions sent by the main author and student groups (GID)~\footnote{As in the first phase, the first author is also a newcomer, we included this experience in the table as Group 0.
@@
 
-Regarding our analysis from the \textbf{participant-as-observer} phase, we observed that most groups applied the parameterized refactor method while a few used the extractor method.
+Regarding our analysis from the \textbf{participant-as-observer} phase, we identified that most groups applied the parameterized refactor method, while a few used the extractor method.
+For instance, Groups 1 and 4 worked on similar duplications, where two functions were involved: one to determine if a register was volatile and another to determine if it was writable.
+These two functions returned opposite Boolean results.
+Group 1 initially addressed the problem using the extract method, while Groups 2 and 4 used macros.
+After feedback from maintainers, both solutions converged toward keeping one of the functions while modifying the other to return a negated query of the former function.
 
-Groups 1 and 4 worked on similar duplications, where two functions are involved: one to determine if a register was volatile and the other to determine if it was writable.
-These were two functions that returned opposite Boolean results.
-Group 1 initially addressed the problem using the extract method, while Group 2 used macros.
-After feedback from maintainers, both solutions converged to keep one of the functions while changing the content of the other to return a negated query of the former function.
-However, by the end of the feedback cycle, maintainers considered the patch from Group 1 irrelevant, while the patch from Group 4 was accepted.
-
-\noindent
-\begin{footnotesize}
-\textit{
-  Quote: ``I think the old code is more readable than hiding the values in a macro
-  even if it is duplicating a few lines of code.''
-}
-\end{footnotesize}
+By the end of the feedback cycle, the patch from Group 4 was accepted, but the final implementation differed substantially from the original.
+The change was primarily driven by a maintainer's feedback: \textit{``I think the old code is more readable than hiding the values in a macro even if it is duplicating a few lines of code.''}
+%
+This response category revealed a recurring concern among maintainers that went beyond functional correctness or redundancy reduction.
+Instead, it reflected a preference for code clarity and comprehension.
+Building on these consistent insights from multiple review interactions, we formulated our second takeaway:
 
 \begin{highlight-box}{Takeaway T2 (Readability)}
-  Maintainers notably prioritized readability over removing duplications in
-  many different contexts. In their view, going against the \textbf{DRY
-  principle} was worth it if it resulted in code that is easier to read and
-  understand.
+Maintainers prioritized readability over removing duplications in many different contexts.
+In their view, going against the \textbf{DRY principle} was worth it if it resulted in easier-to-read and understand code.
 \end{highlight-box}
 
+Maintainers considered the patch from Group 1 irrelevant.
 The reason is that Group 1 chose to apply the refactoring to a device that contained a register that was both read-only and non-volatile.

### 2025-10-28 papers/techdebt2026/techdebt.tex
techdebt2025: update discussion regarding Takeaway 3
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 This example of hardware variations, also mentioned by Kapser and Godfrey~\cite{cloneharm}, leads us to the primary conclusion of our complete participant observation:
-
+%
 \begin{highlight-box}{Takeaway T1}
@@
 Group 1 initially addressed the problem using the extract method, while Groups 2 and 4 used macros.
-After feedback from maintainers, both solutions converged toward keeping one of the functions while modifying the other to return a negated query of the former function.
+After feedback from maintainers, all solutions converged toward keeping one of the functions while modifying the other to return a negated query of the former function.
 
-By the end of the feedback cycle, the patch from Group 4 was accepted, but the final implementation differed substantially from the original.
-The change was primarily driven by a maintainer's feedback: \textit{``I think the old code is more readable than hiding the values in a macro even if it is duplicating a few lines of code.''}
+By the end of the feedback cycle, the patch from Group 4 was accepted, but the final implementation differed substantially from the original without using macros.
+The change was primarily driven by a maintainer's feedback: \textit{``I think the old code is more readable than hiding the values in a macro even if it is duplicating a few lines of code.''}.
 %
-This response category revealed a recurring concern among maintainers that went beyond functional correctness or redundancy reduction.
+This response category (also from groups 1, 2, and 10) revealed a recurring concern among maintainers that went beyond functional correctness or redundancy reduction.
 Instead, it reflected a preference for code clarity and comprehension.
 Building on these consistent insights from multiple review interactions, we formulated our second takeaway:
-
+%\vspace{-.5cm}
 \begin{highlight-box}{Takeaway T2 (Readability)}
@@
 
-Maintainers considered the patch from Group 1 irrelevant.
-The reason is that Group 1 chose to apply the refactoring to a device that contained a register that was both read-only and non-volatile.
+Still focusing on Group 1, maintainers considered the patch irrelevant.
+The reason was that Group 1 chose to apply the refactoring to a device containing a register that was both read-only and non-volatile.
 Thus, the technical debt selected by Group 1 was not merely a duplication but an incorrect implementation of the method.
-
-\noindent
-\begin{footnotesize}
-\textit{
-  Quote: ``[...] such patches might not worth it since the proposed
-  improvement is very small (and questionable) while the upstreaming process
-  still requires some effort.''
-}
-\end{footnotesize}
-
+%
+While the previous interactions highlighted how maintainers often favor readability over strict code deduplication, this case added a new dimension to our understanding: the \textit{cost-benefit} evaluation of proposed changes.
+A maintainer pointed out that the effort required to propagate such modifications after acceptance (i.e., during the upstreaming process) could outweigh the perceived benefits: ``\textit{[...] such patches might not worth it since the proposed improvement is very small (and questionable) while the upstreaming process still requires some effort.}''
+From this feedback, we crafted our third takeaway:
+%\vspace{-.5cm}
 \begin{highlight-box}{Takeaway T3 (Integration Overhead)}
-  A maintainer suggested that the effort put into the contribution after it was
-  accepted (i.e., the process of propagating the change to upper maintainers)
-  was considered. In other words, maintainers may \textbf{evaluate the
-  trade-off between the impact of contributions (not only deduplications) and
-  the struggle to integrate them into the complex Linux project}.
+Maintainers may evaluate the trade-off between the impact of contributions
+(not only deduplications) and the struggle to integrate them into the complex
+Linux project.
 \end{highlight-box}
 
+%TODO: David, consolidar a partir daqui
 Groups 3, 6, and 10 created a struct to encapsulate the parameters and remove the duplication by applying the parameterized method.
@@
 The maintainer's justification for rejecting the patch from Group 6 was that it proposed to fix a duplication while adding more lines than removing, and that it was not convincing enough in other aspects, such as readability.
+...
+``Wrapping this up doesn't provide any real advantage, requiring as
+  it does the reviewer to look at this function AND where the value is set
+  rather than seeing them in one place.''
 
@@
   parsing through the code.
-\hfill\\
-
-\noindent
-\begin{footnotesize}
-\textit{
-  Quote: ``Wrapping this up doesn't provide any real advantage, requiring as
-  it does the reviewer to look at this function AND where the value is set
-  rather than seeing them in one place.''
-}
-\end{footnotesize}
 \end{highlight-box}
@@
 
-%TODO ADICIONADO APENAS PARA FECHAR 5 PGS CERTINHO. ME REMOVA!!!
-\hfill\\
-
 \section*{Replication Package}

### 2025-10-28 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: add mentoring of students description in participant-as-observer
We were not citing how we supported students in their patch submissions
to the Linux kernel, which omitted a great deal of how the second phase
can be considered a participant-as-observer observation. So, add it to
the corresponding subsection, making adjustments for fluidity.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-As the first option, we ran the tool on the IIO subsystem, creating a list of duplicated function pairs.
-Then, we curated the list even more to get more actionable duplications for the students using this approach:
+\paragraph*{\textbf{Deduplication Option 1}} We ran the tool on the IIO subsystem to generate a list of duplicated function pairs, then curated it to highlight actionable cases for students.
+Specifically, we (1) kept only pairs within the same file, since cross-driver duplications could involve multiple maintainers or distinct interfaces; (2) removed very short duplications; and (3) manually ranked the remaining pairs, adding annotations to guide students toward entries likely to be accepted.
+In this option, students were tasked with selecting a recommended entry, devising a way to remove the duplication, and submitting their patches to the IIO mailing list for review.
 
-\begin{enumerate}
-    \item We filtered the list to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different drivers may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
-    \item We removed duplications that were too short and would not represent a significant contribution;
-    \item We manually ranked code duplications and marked some duplications with insights based on personal opinion and perceptions over the remaining entries.
-This was done to guide students so that they could approach the technical debt with a greater chance of having their contribution merged into the code.
-\end{enumerate}
-
-The students' task was to select a recommended entry from the list, refactor the code to eliminate these duplications, and then submit their patches to the IIO mailing list for the maintainers' review.
-%TODO: explicar o processo do envio para o mentores/monitores, o tipo de feedback que recebiam antes enviarem.
-
-For the second alternative, students were offered an experience similar to our initial complete participant observation.
-In this approach, students were responsible for the entire workflow: executing the tool, independently analyzing the results to identify a potential deduplication, creating a patch, and sending the patch to the driver's maintainers.
-%TODO: aqui, no caso do Sebastian, também houve a revisão intermediária dos mentores?
+\paragraph*{\textbf{Deduplication Option 2}} We offered an experience similar to our initial complete participant observation, where students managed the entire workflow: running the tool, analyzing results to identify potential deduplications, creating a patch, and submitting it to the driver's maintainers.
+This option provided freedom of choice but made the students' task more complex.
 
@@
 
-Using blog posts, the groups documented their experiences and approaches to refactoring the duplications, as well as their experiences submitting a patch to the Linux kernel~[omitted].
-We also analyzed these posts to understand the refactoring patterns used in deduplications, the experience of sending patches, and the opinions of subsystem maintainers on them.
-%TODO: é fácil e queremos disponibilizar esses blogs também no pacote de replicação?
+It is important to stress that, \textbf{no matter the chosen approach}, we supported all groups by providing review cycles on their contributions before submission, to prepare students for the interactions with the maintainers and alleviate rough, but natural, newcomers' mistakes.
+After the submission, we helped students understand the feedback from maintainers and gave directions on how to move forward.
+This close, yet indirect, participation through intimate mentoring characterizes the second phase as our participant-as-observer observation.
+
+For this second phase, our primary data source was the mailing list threads, where all development occurred (available in our Replication Package).
+Using blog posts, the groups documented their experiences submitting a patch to the Linux kernel~[omitted], and we also leveraged those to complement our observations.
+% (available in our Replication Package).
+%TODO: Add blogs ao pacote de replicação
 
@@
 %
-\begin{highlight-box}{Takeaway T1}
+\begin{highlight-box}{Takeaway T0}
 Driver developers do not necessarily view duplicated code negatively, which contradicts standard software engineering practices.
@@
 Instead, it reflected a preference for code clarity and comprehension.
-Building on these consistent insights from multiple review interactions, we formulated our second takeaway:
+Building on these consistent insights from multiple review interactions, we formulated our first takeaway:
 %\vspace{-.5cm}
-\begin{highlight-box}{Takeaway T2 (Readability)}
+\begin{highlight-box}{Takeaway T1 (Readability)}
 Maintainers prioritized readability over removing duplications in many different contexts.
@@
 A maintainer pointed out that the effort required to propagate such modifications after acceptance (i.e., during the upstreaming process) could outweigh the perceived benefits: ``\textit{[...] such patches might not worth it since the proposed improvement is very small (and questionable) while the upstreaming process still requires some effort.}''
-From this feedback, we crafted our third takeaway:
+From this feedback, we crafted our second takeaway:
 %\vspace{-.5cm}
-\begin{highlight-box}{Takeaway T3 (Integration Overhead)}
+\begin{highlight-box}{Takeaway T2 (Integration Overhead)}
 Maintainers may evaluate the trade-off between the impact of contributions
@@
   duplication may be forgivable if it improves performance.
-\hfill\\
 \end{highlight-box}
@@
 
-We derived four takeaways from \textbf{participant-as-observer observation}.
-These takeaways are accompanied by example quotes from the maintainers that originated during the review process~\footnote{It is paramount to state that we \textbf{are not affirming} that these quotes represent the complete positioning of the maintainers who wrote them on code duplication.}.
-
 \section{Concluding Remarks}
 
-Ethnographic studies, including full participant and participant-as-observer observations, provided a realistic view of the opportunities and challenges in removing code duplications detected by the tool.
-First-time Linux kernel contributors could effectively use the tool to identify duplications and submit patches.
+This ethnographic study, combining complete participant and participant-as-observer observations, provided a realistic view of the opportunities and challenges in removing code duplications detected by the tool.
+First-time Linux kernel contributors could effectively use the tool to identify duplicated functions and submit meaningful patches.
 
-While this research included one successfully merged patch by the authors, and 7 out of 12 patches sent by students were accepted and merged into Linux, a portion of the students' efforts encountered hurdles: 4 student groups had their patches rejected.
+Throughout this process, we observed that maintainers do not necessarily regard code duplication as inherently harmful.
+Instead, their decisions reflect a nuanced and context-dependent understanding of software quality.
+As evidenced by our takeaways\footnote{It is paramount to state that we \textbf{are not affirming} that these quotes represent the complete positioning of the maintainers who wrote them on code duplication.}, readability often outweighed strict adherence to the DRY principle (T1), while the effort required to propagate changes also influenced maintainers’ acceptance decisions (T2).
+Moreover, avoiding cognitive overhead when navigating the code (T3) and preserving performance in resource-constrained environments (T4) emerged as additional justifications for tolerating certain forms of duplication.
 
-Nevertheless, our work suggests that maintainers do not always view code duplication as inherently unfavorable, especially when it improves clarity, performance, and more.
-This suggests a nuanced understanding of code quality within the Linux kernel community that can go against the DRY principle, where practicality and maintainability in specific contexts sometimes outweigh the general principle of eliminating all code duplication.
+Overall, these findings reveal that, within the Linux kernel community, the evaluation of code duplication extends beyond technical redundancy.
+Maintainability, clarity, and practical trade-offs play a decisive role in shaping maintainers’ judgments, suggesting that DRY is sometimes less valuable than keeping the code understandable, efficient, and integrated smoothly into a large and evolving system.
 

### 2025-10-28 papers/techdebt2026/techdebt.tex
techdebt2026: update final remarks
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 
-\paragraph*{\textbf{Deduplication Option 1}} We ran the tool on the IIO subsystem to generate a list of duplicated function pairs, then curated it to highlight actionable cases for students.
-Specifically, we (1) kept only pairs within the same file, since cross-driver duplications could involve multiple maintainers or distinct interfaces; (2) removed very short duplications; and (3) manually ranked the remaining pairs, adding annotations to guide students toward entries likely to be accepted.
-In this option, students were tasked with selecting a recommended entry, devising a way to remove the duplication, and submitting their patches to the IIO mailing list for review.
+As the first option, we ran the tool on the IIO subsystem, creating a list of duplicated function pairs.
+Then, we curated the list even more to get more actionable duplications for the students using this approach:
 
-\paragraph*{\textbf{Deduplication Option 2}} We offered an experience similar to our initial complete participant observation, where students managed the entire workflow: running the tool, analyzing results to identify potential deduplications, creating a patch, and submitting it to the driver's maintainers.
-This option provided freedom of choice but made the students' task more complex.
+\begin{enumerate}
+    \item We filtered the list to keep only duplicated pairs that happened in the same file, because the IIO subsystem is very decentralized, and a duplication between two different drivers may involve the cooperation of different maintainers and/or the unification of totally different interfaces;
+    \item We removed duplications that were too short and would not represent a significant contribution;
+    \item We manually ranked code duplications and marked some duplications with insights based on personal opinion and perceptions over the remaining entries.
+This was done to guide students so that they could approach the technical debt with a greater chance of having their contribution merged into the code.
+\end{enumerate}
+
+The students' task was to select a recommended entry from the list, refactor the code to eliminate these duplications, and then submit their patches to the IIO mailing list for the maintainers' review.
+%TODO: explicar o processo do envio para o mentores/monitores, o tipo de feedback que recebiam antes enviarem.
+
+For the second alternative, students were offered an experience similar to our initial complete participant observation.
+In this approach, students were responsible for the entire workflow: executing the tool, independently analyzing the results to identify a potential deduplication, creating a patch, and sending the patch to the driver's maintainers.
+%TODO: aqui, no caso do Sebastian, também houve a revisão intermediária dos mentores?
 
@@
 
-It is important to stress that, \textbf{no matter the chosen approach}, we supported all groups by providing review cycles on their contributions before submission, to prepare students for the interactions with the maintainers and alleviate rough, but natural, newcomers' mistakes.
-After the submission, we helped students understand the feedback from maintainers and gave directions on how to move forward.
-This close, yet indirect, participation through intimate mentoring characterizes the second phase as our participant-as-observer observation.
-
-For this second phase, our primary data source was the mailing list threads, where all development occurred (available in our Replication Package).
-Using blog posts, the groups documented their experiences submitting a patch to the Linux kernel~[omitted], and we also leveraged those to complement our observations.
-% (available in our Replication Package).
-%TODO: Add blogs ao pacote de replicação
+Using blog posts, the groups documented their experiences and approaches to refactoring the duplications, as well as their experiences submitting a patch to the Linux kernel~[omitted].
+We also analyzed these posts to understand the refactoring patterns used in deduplications, the experience of sending patches, and the opinions of subsystem maintainers on them.
+%TODO: é fácil e queremos disponibilizar esses blogs também no pacote de replicação?
 

### 2025-10-28 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: standardize the term "coding style"
Throughout the text, there are two instances of "code style", which
isn't incorrect, yet, it diverges from what we were using ("coding
style"). In this sense, standardize this term, by replacing the former
occurences with the latter.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

--- 2025-10-28 papers/techdebt2026/techdebt.tex
@@
 The 2025 course offering had 37 students (25 undergraduate and 12 graduate), who were asked to form groups of two or three members, but graduate students could work individually.
-To assist students in this task, we (a professor and three more experienced graduate students) prepared several alternatives for contributing to the Linux kernel, including simpler contributions such as fixing code style issues.
+To assist students in this task, we (a professor and three more experienced graduate students) prepared several alternatives for contributing to the Linux kernel, including simpler contributions such as fixing coding style issues.
 In this context, we specified two options for removing duplications in the Linux kernel.
@@
 
-In this attempt, we reached a point where the deduplication was good enough to submit it as a contribution to the AMD Display driver code quality.
-Thus, we moved to sending the refactoring to the driver maintainers as a patch.
-
-%TODO: dado que falamos da demora do feedback, colocar a data de envia da V1.
-In the feedback on the first submitted version, the maintainers only asked for minor changes to align with the coding style, correct license use, and align with initially unknown conventions.
-We submitted a second version, in which the maintainers asked us to move the functions created in the generic library to an existing file instead of creating a new one.
-We sent a third version, and the patch was accepted and integrated into the kernel codebase on February 25, 2025.
-%\footnote{Patch submitted: \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br}}.
+At this stage, we considered the deduplication proposal good enough to submit it to the AMD Display driver as a patch.
+We sent the first patch on August 9, 2025, and a resend followed on October 9, 2025, due to a lack of initial feedback.
+Nevertheless, for the first two versions, the maintainers' response requested minor changes for coding style, license use, and alignment with existing conventions.
+In the third version, the feedback was to move the new generic library functions to an existing file, instead of creating a new one.
+Then, the fourth version was accepted and integrated into the kernel on February 25, 2025.
+%\footnote{V1: \href{https://lore.kernel.org/amd-gfx/20240809210122.23814-1-luanicaro@usp.br/}{https://lore.kernel.org/amd-gfx/20240809210122.23814-1-luanicaro@usp.br/}, V1 (resend): \href{https://lore.kernel.org/amd-gfx/20241009153424.15256-1-luanicaro@usp.br/}{https://lore.kernel.org/amd-gfx/20241009153424.15256-1-luanicaro@usp.br/}, V4 (approved): \href{https://lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}{lore.kernel.org/all/20250225015532.303032-1-luanicaro@usp.br/}}.
 
@@
 Groups 7 and 8 used the extract method to approach the deduplications.
-Group 8 sent their patch to the maintainers, and the code changes were approved, with minor requested changes in code style and smaller errors fixed (++T5).
+Group 8 sent their patch to the maintainers, and the code changes were approved, with minor requested changes in coding style and smaller errors fixed (++T5).
 In the case of Group 7, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function.

### 2025-10-29 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: revise results section
Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-Building on this insight, we realized that the perception identified in T1 might not be limited to driver developers alone.
-To deepen our understanding, we systematically examined other contributions related to code duplication reduction and analyzed how different maintainers responded to such patches.
+Building on this insight, we realized that the perception identified in takeaway T0 might not be limited to AMD Display alone.
+To deepen our understanding, we examined other contributions regarding code duplication reduction and analyzed how different maintainers responded to such patches.
 This broader analysis aimed to capture diverse maintainer perspectives and contextual factors influencing their decisions.
 
-Table~\ref{tab:stu} summarizes all the contributions sent by the main author and student groups (GID)~\footnote{As in the first phase, the first author is also a newcomer, we included this experience in the table as Group 0.
-Nevertheless, the subsequent analysis will focus more on the patches sent by students.}.
+Table~\ref{tab:stu} summarizes all the contributions sent by the main author and student groups (GID)~\footnote{Since the first author was also a newcomer, the first phase experience was included as Group 0.}.
 It contains the information about the chosen subsystem (SS) and the similarity threshold (SML).
@@
 \hline
-0 & AMD & 100\% & Accepted (v3) & T1 & +114/-520 \\ \hline
-1 & IIO & 100\% & Dropped (v2) & T2, T3 & +7/-14 \\ \hline
-2 & IIO & 100\% & Dropped (v1) & T2 & +23/-27 \\ \hline
+0 & AMD & 100\% & Accepted (v3) & T0 & +114/-520 \\ \hline
+1 & IIO & 100\% & Dropped (v2) & T1, T2 & +7/-14 \\ \hline
+2 & IIO & 100\% & Dropped (v1) & T1 & +23/-27 \\ \hline
 3 & IIO & 100\% & Accepted (v1) & None & +12/-30 \\ \hline
-4 & IIO & 100\% & Accepted (v3) & T2 & +2/-7 \\ \hline
+4 & IIO & 100\% & Accepted (v3) & T1 & +2/-7 \\ \hline
 5 & IIO & 90\% & Accepted (v1) & None & +17/-70 \\ \hline
-6 & IIO & 90\% & Dropped (v1) & T4 & +95/-92 \\ \hline
+6 & IIO & 90\% & Dropped (v1) & T3 & +95/-92 \\ \hline
 7 & IIO & 90\% & Accepted (v3) & None & +20/-36 \\ \hline
-8 & IIO & 90\% & Accepted (v4) & T5 & +20/-16 \\ \hline
-9 & IIO & 90\% & Accepted (v6) & T5 & +149/-243 \\ \hline
-10 & IIO & 90\% & Dropped (v1) & T2, T5 & +32/-58 \\ \hline
+8 & IIO & 90\% & Accepted (v4) & T4 & +20/-16 \\ \hline
+9 & IIO & 90\% & Accepted (v6) & T4 & +149/-243 \\ \hline
+10 & IIO & 90\% & Dropped (v1) & T1, T4 & +32/-58 \\ \hline
 11 & IIO & 100\% & Dropped (v1) & None & +2/-12 \\
@@
 
-Regarding our analysis from the \textbf{participant-as-observer} phase, we identified that most groups applied the parameterized refactor method, while a few used the extractor method.
-For instance, Groups 1 and 4 worked on similar duplications, where two functions were involved: one to determine if a register was volatile and another to determine if it was writable.
-These two functions returned opposite Boolean results.
-Group 1 initially addressed the problem using the extract method, while Groups 2 and 4 used macros.
-After feedback from maintainers, all solutions converged toward keeping one of the functions while modifying the other to return a negated query of the former function.
+We derived four takeaways from our \textbf{participant-as-observer observation}.
+We substantiate these takeaways with quotes from the maintainers that originated during the review process~\footnote{We \textbf{do not claim} these quotes fully represent the maintainers’ stance on code duplication.}.
 
-By the end of the feedback cycle, the patch from Group 4 was accepted, but the final implementation differed substantially from the original without using macros.
-The change was primarily driven by a maintainer's feedback: \textit{``I think the old code is more readable than hiding the values in a macro even if it is duplicating a few lines of code.''}.
-%
-This response category (also from groups 1, 2, and 10) revealed a recurring concern among maintainers that went beyond functional correctness or redundancy reduction.
+Groups 1 and 4 worked on similar duplications using the extract method on predicate functions (functions that return a boolean) that, essentially, were the negation of one another.
+Group 4 initially used macros, but, after feedback from maintainers, both solutions converged toward keeping one of the functions while modifying the other to return the negation of the former.
+Both cases suffered substantial mutations, even though they were correct in their first iterations, and these changes were primarily driven by the feedback of maintainers regarding readability.
+For Group 1, a maintainer said that \textit{``[...] the naming as \texttt{\_reg\_check()} is not helpful as it doesn't indicate anything specific is being checked.''}, while for Group 4, the feedback explicitly highlighted the precedence of readability over deduplications: \textit{``I think the old code is more readable than hiding the values in a macro even if it is duplicating a few lines of code.''}.
+
+Even though the deduplication of Group 2 was not similar to that of Group 1 and Group 4 in terms of implementation (it used the parametized method), it also brought forth the readability factor: \textit{``In my view this isn't a significant enough reduction to justify the more complex code.''}.
+
+These experiences revealed a recurring concern among maintainers beyond functional correctness or redundancy reduction.
 Instead, it reflected a preference for code clarity and comprehension.
-Building on these consistent insights from multiple review interactions, we formulated our first takeaway:
-%\vspace{-.5cm}
+From these consistent insights that originated from multiple review interactions, we formulated our first takeaway:
+
 \begin{highlight-box}{Takeaway T1 (Readability)}
@@
 
-Still focusing on Group 1, maintainers considered the patch irrelevant.
-The reason was that Group 1 chose to apply the refactoring to a device containing a register that was both read-only and non-volatile.
-Thus, the technical debt selected by Group 1 was not merely a duplication but an incorrect implementation of the method.
-%
+Still focusing on Group 1, maintainers questioned the merit of the proposed patch.
 While the previous interactions highlighted how maintainers often favor readability over strict code deduplication, this case added a new dimension to our understanding: the \textit{cost-benefit} evaluation of proposed changes.
-A maintainer pointed out that the effort required to propagate such modifications after acceptance (i.e., during the upstreaming process) could outweigh the perceived benefits: ``\textit{[...] such patches might not worth it since the proposed improvement is very small (and questionable) while the upstreaming process still requires some effort.}''
+A maintainer pointed out that the effort required to propagate such modifications after acceptance (i.e., during the upstreaming process) could outweigh the perceived benefits: ``\textit{[...] such patches might not worth it since the proposed improvement is very small (and questionable) while the upstreaming process still requires some effort.}''.
 From this feedback, we crafted our second takeaway:
-%\vspace{-.5cm}
+
 \begin{highlight-box}{Takeaway T2 (Integration Overhead)}
@@
 
-%TODO: David, consolidar a partir daqui
-Groups 3, 6, and 10 created a struct to encapsulate the parameters and remove the duplication by applying the parameterized method.
-The patches from Groups 6 and 10 were rejected by the same maintainer who approved the patch from Group 3, tackling the duplication of a simple iterative algorithm, which also contained a bug that could be inferred by inspecting both functions.
-In this case, the group refactored code from a more critical technical debt and took the opportunity to fix an unnoticed incorrect behavior.
+Groups 3, 6, and 10 used the parameterized method by introducing a new struct used as a parameter to a generic function.
+Notwithstanding that the same maintainer reviewed all three contributions, only Group 3 had its patch accepted.
+Our understanding was that this change, in particular, acted on a simple iterative algorithm that (colaterally) also fixed a bug, which favored its straightforward acceptance.
 
-The maintainer's justification for rejecting the patch from Group 6 was that it proposed to fix a duplication while adding more lines than removing, and that it was not convincing enough in other aspects, such as readability.
-...
-``Wrapping this up doesn't provide any real advantage, requiring as
-  it does the reviewer to look at this function AND where the value is set
-  rather than seeing them in one place.''
+Although removing duplications does not necessarily mean reducing lines of code, the patch from Group 6 added more net lines, which the maintainer pointed out as a drawback.
+Still, the real reason for the rejection was that the deduplication scattered parts of a complex code that resided close together before: \textit{``Wrapping this up doesn't provide any real advantage, requiring as it does the reviewer to look at this function AND where the value is set rather than seeing them in one place.''}.
+Needing to jump to a function definition (to see its behavior) or a function call (to see its concrete parameters) in order to understand code is a cognitively demanding task for developers~\cite{questions-programmers-ask}.
+Merging these informations, we derived our third takeaway:
 
-\begin{highlight-box}{Takeaway T4 (Cognitive Load)}
-  Needing to jump to a function definition (to see its behavior) or a function
-  call (to see its concrete parameters) in order to understand code is a
-  cognitively demanding task for developers~\cite{questions-programmers-ask}.
-  Duplications that avoided this were justifiable to reduce cognitive load when
-  parsing through the code.
+\begin{highlight-box}{Takeaway T3 (Cognitive Load)}
+Maintainers can justify duplications if they avoid having to jump back and forth in the code when parsing through the file.
 \end{highlight-box}
 
-The reason for rejecting the patch from Group 10 was that they refactored code files with simple context; therefore, adding a layer of abstraction to remove duplication decreased code readability without clear gains in codebase quality.
-Interestingly, both rejected groups addressed duplications across initialization functions for different devices, where code readability was vulnerable to refactoring methods (T2  + T5).
-[blabla] Group 9 initially approached the selected duplication with a parameterized method, but after entering the feedback cycle, it was revealed that the problem represented a much more complex technical debt.
-From the first version to the last (version six), the patch was split into a series with four patches, where two of them have been applied, and the remaining entries are receiving suggestions.
-Even though the patch became a series of four, only one of them (one of the accepted) was strictly tackling duplication.
+The reason for rejecting the patch from Group 10 involved the deduplication lowering readability, as the maintainer was \textit{``[...] not sure the code reduction is sufficiently to cover the resulting loss of readability.''}, reinforcing takeaway T1.
 
-\noindent
-\begin{footnotesize}
-\textit{
-  Quote: ``Note that is not an appropriate change for the large reads though
-  as \texttt{spi\_write\_then\_read()} bounces all buffers and so would add a
-  copy to those high(ish) performance paths.''
-}
-\end{footnotesize}
+In the interaction of Group 10, there was also a hint that performance could be considered when evaluating the merit of a deduplication.
+Nonetheless, this notion surfaced aggressively with Group 9, where the contribution could completely remove both duplicated functions using the inline method by leveraging helpers; however, one of the functions had a call that, if eliminated, would result in a significant performance impact: \textit{``Note that is not an appropriate change for the large reads though [...] This is the one case were \texttt{spi\_write\_then\_read()} is probably not appropriate due to the large buffers that are potentially involved.''}.
+These interactions, coupled with the low-level development context of Linux and driver development, which is usually related to devices with limited resources, produced our fourth and final takeaway.
 
-\begin{highlight-box}{Takeaway T5 (Performance)}
-  Given the low-level development context of Linux and the IIO subsystem, which
-  is usually related to devices with limited resources, we detected hints that
-  duplication may be forgivable if it improves performance.
+\begin{highlight-box}{Takeaway T4 (Performance)}
+Interactions with maintainers hinted that duplications can be forgiven if they improve performance in contexts where resources are limited.
 \end{highlight-box}
 
-Groups 7 and 8 used the extract method to approach the deduplications.
-Group 8 sent their patch to the maintainers, and the code changes were approved, with minor requested changes in coding style and smaller errors fixed (++T5).
-In the case of Group 7, the maintainer initially rejected the proposed patch but suggested a new refactoring approach: applying the extraction method refactoring to the duplicated code to an existing helper function.
-After the feedback cycle, maintainers accepted the patch in version 3.
+Groups 7 and 8 used the extract method using helpers to remove duplicated code in the function's body.
+Group 7 refined its patch from the maintainers' feedback, resulting in a smooth acceptance process that did not produce any insights.
+Group 8 followed a similar course, but generated an interaction reinforcing takeaway T4.
+In the feedback of version 1, a maintainer claimed that \textit{``Even though there is less code repetition, we now have an extra comparison [...]''}, with another maintainer corroborating this in version three by saying \textit{``There is always a balance/trade-off between modularity and execution speed.
+I agree with \texttt{anonymous-maintainer}'s reply in the first patch [...]''}, regardless of the fact that the patch was eventually accepted in version four.
 
-Group 11 was a single student who took the second option and executed the tool, completing the entire process of identifying duplications himself.
-The student identified a duplication in the driver that occurred across eighteen code files, creating a patch to remove approximately four hundred lines of duplicated code.
-The student sent the deduplications to the maintainers, and their feedback pointed to fixing warnings in the code without arguing about the merit of the duplication removals.
-After addressing the issue, the patch was applied.
-This student also contributed to IIO, tackling a code duplication with inline method refactoring.
-This one, however, was received with more skepticism by the maintainer, and the student dropped the contribution because he did not think the feedback instructions were clear enough.
+We did not explore the contributions of Group 5 and Group 11 in detail, as they did not result in any insights regarding these conflicting perspectives on software engineering practices.
+However, they were still meaningful as they produced impactful contributions that reduced a lot of duplicated code.
+Especially in Group 11, where a single student executed the whole workflow we did in the first phase, and managed to remove four hundred net lines of code.
 
-Analyzing the students' approaches to the duplications and the maintainers' feedback, we saw that people without previous experience in Linux kernel development could approach the duplications found by the tool as a comprehensive and more straightforward way to become contributors to the kernel.
-We observed that not all duplications are viewed as code of bad quality, with maintainers analyzing the trade-off between the purpose of the duplicated code and many other factors.
+In this sense, we discuss another aspect of importance stemming from this work: using the developed tool to create deduplication opportunities in the Linux kernel can be an excellent introduction for newcomers.
 
-The eleven groups, which have submitted 11 contributions to the IIO subsystem, have collectively requested the insertion of 379 lines and the deletion of 605 lines in the subsystem, representing an average of 34.45 lines added and 55 lines removed per contribution.
+To illustrate this point, we display some statistics concerning the 13 contributions contemplated in this paper:
 
-Considering the sum of fully merged contributions, we have a code diff value of +51/-143, with an average of +12.75/-35.75.
-This ratio is better than dropped contributions, which have a total of +179/-219 with an average of +29.83/-36.5.
-
-It is also possible to notice a difference based on the similarity threshold used to detect duplications.
-The proposed patches to address 100\%-similarity duplications summed up to +46/-90 (average of +9.2/-18).
-For a 90\% value, the result was a total of +333/-515 with an average of +55.5/-85.83.
-The 90\% value enabled the tool to identify less strict duplication cases, and, therefore, code segments with a greater extension, compared to the 100\% similarity entries, which were shorter segments, causing a difference in the average number of modified lines.
-
-Finally, the AMD contribution from Group 11, which involved the complete use of our developed tool components, got a completely different outcome.
-With a total of 90 lines inserted and 489 lines removed, there are some points to be considered for the success and high modification extension of the contribution: (1) the AMD Display subsystem has some self-reported problems regarding overall technical debt; (2) the AMD Display subsystem has a more centralized code structure and maintainer hierarchy, enabling contributors to tackle code duplications across a broader scope and multiple files, which happened to Group 11's contribution;
+\begin{itemize}
+  \item Collectively, the patches proposed the insertion of 585 lines and the removal of 1626 lines.
+  \item 8 out of 13 contributions were accepted, resulting in a success rate of 62\%.
+  \item For accepted contributions, the code diff value was +255/-1152, averaging +31.9/-144.0 additions/removals.
+  \item For rejected contributions, the code diff value was +181/-231, averaging +36.2/-46.2 additions/removals.
+\end{itemize}
 

### 2025-10-29 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: remove redundant footnote
In the concluding remarks, a footnote that was moved to page 4, was
still present in page 5, so remove it.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 Instead, their decisions reflect a nuanced and context-dependent understanding of software quality.
-As evidenced by our takeaways\footnote{It is paramount to state that we \textbf{are not affirming} that these quotes represent the complete positioning of the maintainers who wrote them on code duplication.}, readability often outweighed strict adherence to the DRY principle (T1), while the effort required to propagate changes also influenced maintainers’ acceptance decisions (T2).
+As evidenced by our takeaways, readability often outweighed strict adherence to the DRY principle (T1), while the effort required to propagate changes also influenced maintainers’ acceptance decisions (T2).
 Moreover, avoiding cognitive overhead when navigating the code (T3) and preserving performance in resource-constrained environments (T4) emerged as additional justifications for tolerating certain forms of duplication.

### 2025-10-29 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: make DRY highlight box unique
The highlight box of the DRY principle on the introduction was of the
same color as the takeaway boxes. To differentiate, make it blue-ish.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 \usepackage{todonotes}
+\usepackage[most]{tcolorbox}
 
 %% Takeaway Boxes %%
-\usepackage[most]{tcolorbox}
-\tcbset{colback=gray!10, colframe=gray!50!black, boxrule=0.5pt,
-        left=1mm, right=1mm, top=1mm, bottom=1mm, sharp corners}
-
 \newenvironment{highlight-box}[1]{%
-  \begin{tcolorbox}
+  \begin{tcolorbox}[colback=gray!10, colframe=gray!50!black, boxrule=0.5pt,
+                    left=1mm, right=1mm, top=1mm, bottom=1mm, sharp corners]
   \textbf{#1:} \itshape}{\end{tcolorbox}}
 %% Takeaway Boxes %%
+
+%% DRY Highlight Box %%
+\newenvironment{dry-highlight-box}[1]{%
+  \begin{tcolorbox}[colback=blue!5!white, colframe=blue!60!black, boxrule=0.5pt,
+                    left=1mm, right=1mm, top=1mm, bottom=1mm, sharp corners]
+  \textbf{#1:} \itshape}{\end{tcolorbox}}
+%% DRY Highlight Box %%
 
@@
 
-\begin{highlight-box}{DRY principle}
+\begin{dry-highlight-box}{DRY principle}
   ``We feel that the only way to develop software reliably, and to make our
@@
   authoritative representation within a system''.
-\end{highlight-box}
+\end{dry-highlight-box}
 

### 2025-10-29 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: better style DRY highlight box
The DRY box contained a nested colon (foo: is bar and bar: is foobar)
and a full stop after the quote. Fix both by styling the DRY box to
remove this colon, and take the opportunity to insert the full stop
inside the quote.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
                     left=1mm, right=1mm, top=1mm, bottom=1mm, sharp corners]
-  \textbf{#1:} \itshape}{\end{tcolorbox}}
+    \textbf{[#1]} \itshape}{\end{tcolorbox}}
 %% DRY Highlight Box %%
@@
   DRY principle: Every piece of knowledge must have a single, unambiguous,
-  authoritative representation within a system''.
+  authoritative representation within a system.''
 \end{dry-highlight-box}

### 2025-10-29 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: include a quote for Group 10 and takeaway T4
The patch from Group 10 produced an interesting quote related to
takeaway T4, which is not present in the paper. Given that we have quite
a few space left, include it.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-In the interaction of Group 10, there was also a hint that performance could be considered when evaluating the merit of a deduplication.
-Nonetheless, this notion surfaced aggressively with Group 9, where the contribution could completely remove both duplicated functions using the inline method by leveraging helpers; however, one of the functions had a call that, if eliminated, would result in a significant performance impact: \textit{``Note that is not an appropriate change for the large reads though [...] This is the one case were \texttt{spi\_write\_then\_read()} is probably not appropriate due to the large buffers that are potentially involved.''}.
-These interactions, coupled with the low-level development context of Linux and driver development, which is usually related to devices with limited resources, produced our fourth and final takeaway.
+In the interaction of Group 10, there was also a hint that performance could be considered when evaluating the worth of a deduplication, whilst reaffirming that readability was a priority in that context: \textit{``I'd be slightly interested to see the optimized output of the two approaches, but this is far from a high performance path so we care a lot more about readability here.''}.
+Nonetheless, this notion surfaced aggressively with Group 10, where the contribution could completely remove both duplicated functions using the inline method by leveraging helpers; however, one of the functions had a call that, if eliminated, would result in a significant performance impact: \textit{``Note that is not an appropriate change for the large reads though [...] This is the one case were \texttt{spi\_write\_then\_read()} is probably not appropriate due to the large buffers that are potentially involved.''}.
+These interactions, coupled with the low-level development context of Linux and driver development, which is usually related to devices with limited resources, produced our fourth and final takeaway:
 

### 2025-10-29 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: rephrase paragraph about undetailed groups
Group 5 and Group 11 were not detailed in the same sense as the other
groups, as they did not produce or influence the takeaways and the core
idea of the paper. However, the way it was put, made it feel like they
had literally no meaning in this regard, which was not true, so rephrase
this paragraph for a better contextualization.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-We did not explore the contributions of Group 5 and Group 11 in detail, as they did not result in any insights regarding these conflicting perspectives on software engineering practices.
+We did not explore the contributions of Group 5 and Group 11 in detail, as they did not result in (or influence) any insights regarding these fascinating perspectives from the Linux community that are dissonant with the apparently unquestionable DRY principle.
 However, they were still meaningful as they produced impactful contributions that reduced a lot of duplicated code.

### 2025-10-29 papers/techdebt2026/techdebt.tex
xtechdebt2026: minor adjustments
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 
-We executed the tool in the AMD Display driver codebase.
-We manually analyzed the biggest duplications regarding the number of lines, finding one pair that we judged promising to try to deduplicate.
+We executed the tool in the AMD Display driver codebase and manually analyzed the biggest duplications regarding the number of lines, finding one pair that we judged promising to try to deduplicate.
 We approached big duplications as proof of concept to gauge the tool's potential for enabling significant impacts.
@@
 Thus, we proposed a simple systematic approach to mitigate all the functions duplicated in the context, not just the duplicated function pairs.
-After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our Replication Package).
+After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our replication package\footnote{\url{https://doi.org/10.5281/zenodo.17432193}}	).
 
@@
 
-It is important to stress that, \textbf{no matter the chosen approach}, we supported all groups by providing review cycles on their contributions before submission, to prepare students for the interactions with the maintainers and alleviate rough, but natural, newcomers' mistakes.
+It is important to stress that, \textbf{no matter the chosen approach, we supported all groups by providing review cycles on their contributions before submission}, to prepare students for the interactions with the maintainers and alleviate rough, but natural, newcomers' mistakes.
 After the submission, we helped students understand the feedback from maintainers and gave directions on how to move forward.
@@
 
-For this second phase, our primary data source was the mailing list threads, where all development occurred (available in our Replication Package).
+For this second phase, our primary data source was the mailing list threads, where all development occurred (also available in our replication package).
 Using blog posts, the groups documented their experiences submitting a patch to the Linux kernel~[omitted], and we also leveraged those to complement our observations.
@@
 
+\vspace{-.2cm}
 \section{Results}
@@
 
-Table~\ref{tab:stu} summarizes all the contributions sent by the main author and student groups (GID)~\footnote{Since the first author was also a newcomer, the first phase experience was included as Group 0.}.
-It contains the information about the chosen subsystem (SS) and the similarity threshold (SML).
-Each entry also has the patch status, takeaways that may have surfaced from the interaction, and a code differential corresponding to the final patch versions before it was dropped or accepted into the code base.
-The code differential for a set of patches corresponds to the sum of added and removed lines of each patch.
-
 \begin{table}[ht]
@@
 
+Table~\ref{tab:stu} summarizes all the contributions sent by the main author and student groups (GID)~\footnote{Since the first author was also a newcomer, the first phase experience was included as Group 0.}.
+It contains the information about the chosen subsystem (SS) and the similarity threshold (SML).
+Each entry also has the patch status, takeaways that may have surfaced from the interaction, and a code differential corresponding to the final patch versions before it was dropped or accepted into the code base.
+The code differential for a set of patches corresponds to the sum of added and removed lines of each patch.
+
 We derived four takeaways from our \textbf{participant-as-observer observation}.
-We substantiate these takeaways with quotes from the maintainers that originated during the review process~\footnote{We \textbf{do not claim} these quotes fully represent the maintainers’ stance on code duplication.}.
+We substantiate these takeaways with quotes from the maintainers that originated during the review process\footnote{We \textbf{do not claim} these quotes fully represent the maintainers’ stance on code duplication.}.
 
@@
 Maintainers prioritized readability over removing duplications in many different contexts.
-In their view, going against the \textbf{DRY principle} was worth it if it resulted in easier-to-read and understand code.
+In their view, going against the DRY principle was worth it if it resulted in easier-to-read and understand code.
 \end{highlight-box}
@@
 
-Although removing duplications does not necessarily mean reducing lines of code, the patch from Group 6 added more net lines, which the maintainer pointed out as a drawback.
+Although removing duplications does not necessarily mean reducing lines of code, the patch from Group 6 resulted in a positive net change in lines of code, which the maintainer pointed out as a drawback.
 Still, the real reason for the rejection was that the deduplication scattered parts of a complex code that resided close together before: \textit{``Wrapping this up doesn't provide any real advantage, requiring as it does the reviewer to look at this function AND where the value is set rather than seeing them in one place.''}.
-Needing to jump to a function definition (to see its behavior) or a function call (to see its concrete parameters) in order to understand code is a cognitively demanding task for developers~\cite{questions-programmers-ask}.
+Needing to jump to a function definition (to see its behavior) or a function call (to see its concrete parameters) to understand code is a cognitively demanding task for developers~\cite{questions-programmers-ask}.
 Merging these informations, we derived our third takeaway:
@@
 
-The reason for rejecting the patch from Group 10 involved the deduplication lowering readability, as the maintainer was \textit{``[...] not sure the code reduction is sufficiently to cover the resulting loss of readability.''}, reinforcing takeaway T1.
+The reason for rejecting the patch from Group 10 involved the deduplication lowering readability, as the maintainer was \textit{``[...] not sure the code reduction is sufficiently to cover the resulting loss of readability.''}, reinforcing the takeaway T1.
 
@@
 Group 7 refined its patch from the maintainers' feedback, resulting in a smooth acceptance process that did not produce any insights.
-Group 8 followed a similar course, but generated an interaction reinforcing takeaway T4.
-In the feedback of version 1, a maintainer claimed that \textit{``Even though there is less code repetition, we now have an extra comparison [...]''}, with another maintainer corroborating this in version three by saying \textit{``There is always a balance/trade-off between modularity and execution speed.
+Group 8 followed a similar course, but generated an interaction reinforcing the takeaway T4.
+In the feedback of version 1, a maintainer claimed that \textit{``Even though there is less code repetition, we now have an extra comparison [...]''}, with another maintainer corroborating this in version three by saying: \textit{``There is always a balance/trade-off between modularity and execution speed.
 I agree with \texttt{anonymous-maintainer}'s reply in the first patch [...]''}, regardless of the fact that the patch was eventually accepted in version four.
@@
 Overall, these findings reveal that, within the Linux kernel community, the evaluation of code duplication extends beyond technical redundancy.
-Maintainability, clarity, and practical trade-offs play a decisive role in shaping maintainers’ judgments, suggesting that DRY is sometimes less valuable than keeping the code understandable, efficient, and integrated smoothly into a large and evolving system.
+Maintainability, clarity, and practical trade-offs play a decisive role in shaping maintainers’ judgments, suggesting that the DRY principle is sometimes less valuable than keeping the code understandable, efficient, and integrated smoothly into a large and evolving system.
 

### 2025-10-29 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: update Zenodo link to replication package v2
Given that we updated the replication package with Luan's threads,
update the link in the paper.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 
-All anonymized threads containing the entire review process and interactions with maintainers are available in this Zenodo repository: \href{https://doi.org/10.5281/zenodo.17432193}{https://doi.org/10.5281/zenodo.17432193}.
+All anonymized threads containing the entire review process and interactions with maintainers are available in this Zenodo repository: \href{https://doi.org/10.5281/zenodo.17474124}{https://doi.org/10.5281/zenodo.17474124}.
 

### 2025-10-29 papers/techdebt2026/techdebt.tex
techdebt2026: update the zenodo footnote link
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 Thus, we proposed a simple systematic approach to mitigate all the functions duplicated in the context, not just the duplicated function pairs.
-After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our replication package\footnote{\url{https://doi.org/10.5281/zenodo.17432193}}	).
+After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our replication package\footnote{\url{https://doi.org/10.5281/zenodo.17474124}}).
 

### 2025-11-01 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: update URL of anonymized ArKanjo repository
The URL of the GitHub Anonymous of ArKanjo contains the substring
`arkanjo`, which violated the double-blind. We changed the URL to hide
any identifiable information and, so, update it in the paper.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 Existing free software tools we explored in practice suffered from limited functionality~[omitted].
-Thus, we developed a command-line tool\footnote{Anonymous repository:  \url{https://anonymous.4open.science/r/arkanjo-C548}} designed to detect and analyze function-level code duplication within large codebases, such as the Linux kernel.
+Thus, we developed a command-line tool\footnote{Anonymized repository:  \url{https://anonymous.4open.science/r/anonymous-repo-tech-debt/}} designed to detect and analyze function-level code duplication within large codebases, such as the Linux kernel.
 The tool is released under LGPLv3 (GNU Lesser General Public License, version 3) and employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
@@
 
-To avoid violating the double-anonymized review process, the anonymous repository of the developed tool is available at \url{https://anonymous.4open.science/r/arkanjo-C548}.
+To avoid violating the double-anonymized review process, the anonymous repository of the developed tool is available at \url{https://anonymous.4open.science/r/anonymous-repo-tech-debt/}.
 The tool website link and omitted references will be added in the camera-ready version of the paper.

### 2025-11-01 papers/techdebt2026/techdebt.tex
techdebt2026: techdebt: update Zenodo URL for replication package
Given that we had to update the package on Zenodo due to the author's
name being present, update the Zenodo URL.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 Thus, we proposed a simple systematic approach to mitigate all the functions duplicated in the context, not just the duplicated function pairs.
-After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our replication package\footnote{\url{https://doi.org/10.5281/zenodo.17474124}}).
+After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our replication package\footnote{\url{https://doi.org/10.5281/zenodo.17503684}}).
 
@@
 
-All anonymized threads containing the entire review process and interactions with maintainers are available in this Zenodo repository: \href{https://doi.org/10.5281/zenodo.17474124}{https://doi.org/10.5281/zenodo.17474124}.
+All anonymized threads containing the entire review process and interactions with maintainers are available in this Zenodo repository: \href{https://doi.org/10.5281/zenodo.17503684}{https://doi.org/10.5281/zenodo.17503684}.
 

### 2025-11-01 papers/techdebt2026/techdebt.tex
techdebt2026: patchsets: anom-threads: add spacing for styling
There is small spacing remaining at the end of page 5, so add some
`\hfill` just for styling.

Signed-off-by: David Tadokoro <davidbtadokoro@usp.br>

@@
 Thus, we proposed a simple systematic approach to mitigate all the functions duplicated in the context, not just the duplicated function pairs.
-After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our replication package\footnote{\url{https://doi.org/10.5281/zenodo.17503684}}).
+After approaching the duplications with a systematic strategy, we sent a patch to the AMD development mailing list for the feedback of maintainers, while documenting the process, interactions, and impressions (available in our replication package\footnote{\url{https://doi.org/10.5281/zenodo.17503685}}).
 
@@
 
+\hfill\\
+
 \section{Concluding Remarks}
@@
 Maintainability, clarity, and practical trade-offs play a decisive role in shaping maintainers’ judgments, suggesting that the DRY principle is sometimes less valuable than keeping the code understandable, efficient, and integrated smoothly into a large and evolving system.
+
+\hfill
 

### 2025-11-01 papers/techdebt2026/techdebt.tex
techdebt2026: minor adjustment
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

@@
 Existing free software tools we explored in practice suffered from limited functionality~[omitted].
-Thus, we developed a command-line tool\footnote{Anonymized repository:  \url{https://anonymous.4open.science/r/anonymous-repo-tech-debt/}} designed to detect and analyze function-level code duplication within large codebases, such as the Linux kernel.
+Thus, we developed a command-line tool\footnote{\url{https://anonymous.4open.science/r/anonymous-repo-tech-debt/}} designed to detect and analyze function-level code duplication within large codebases, such as the Linux kernel.
 The tool is released under LGPLv3 (GNU Lesser General Public License, version 3) and employs a two-stage architecture that separates the computationally intensive task of finding duplicates from the process of querying the results.
